### 2019-05-24 manuscript.tex

\documentclass{article}
\usepackage{amsmath, amssymb, graphicx, setspace, multirow,multicol,units,float,fullpage}

\doublespacing

\newcommand{\mathsym}[1]{{}}
\newcommand{\unicode}[1]{{}}

\newcounter{mathematicapage}

\graphicspath{{C:/Users/a.quinones/Dropbox/LosAndes/investigacion/Learning_bagdes}}

\begin{document}

\title{The social costs of badges of status and learning} \author{Andr\'es Qui\~nones} \date{May 2019}
\maketitle

\section{Simple learning model}
\label{simple_model}

\begin{table}
	\centering
\begin{tabular}{c|c|c|cl}
	Option 1 & Option 2 & Action & Estimate & ID \\ \hline
	\multirow{2}{*}{V} & \multirow{2}{*}{R}  & V & Q(VR,V) & 0 \\ \cline{3-4} & & R & Q(VR,R) & 1\\ \cline{3-4} \hline
	\multirow{2}{*}{V} & \multirow{2}{*}{0}  & V & Q(V0,V) & 2 \\ \cline{3-4} & & 0 & Q(V0,0) & 3\\ \cline{3-4} \hline
	\multirow{2}{*}{R} & \multirow{2}{*}{0}  & R & Q(R0,R) & 4\\ \cline{3-4} & & 0 & Q(R0,0) & 5 \\ \cline{3-4} \hline
	\multirow{2}{*}{V} & \multirow{2}{*}{V}  & \multirow{2}{*}{V} & \multirow{2}{*}{Q(VV,V)} & 6 \\ & & \\  \hline
	\multirow{2}{*}{R} & \multirow{2}{*}{R}  & \multirow{2}{*}{R} & \multirow{2}{*}{Q(RR,R)} & 7 \\ & & \\  \hline
	\multirow{2}{*}{0} & \multirow{2}{*}{0}  & \multirow{2}{*}{0} & \multirow{2}{*}{Q(00,0)} & 8 \\ & & 
\end{tabular}
	\label{table:States}
	\caption{ State-space and potential action that agent can take. V and R stand for visitor and resident client respectively. Option 1 and 2 are given by the environment, and the action is chosen by the agent among the two options(see main text). The agent learns through estimating rewards obtained from each one of this State-action pairs $Q(s,a)$ depicted in the last column.}
\end{table}

The agent makes this estimates by starting with an incorrect value for each state-action pair and progressively update it as it faces the different states. Specifically, the update is given by
\begin{equation}
	Q(s_t,a_t)\leftarrow Q(s_t,a_t) +\alpha [R_{t+1}+\gamma Q(s_{t+1},a_{t+1})-Q(s_t,a_t)],
	\label{equation:updateRule}
\end{equation}

where $R_{t+1}$ is the reward obtained by the agent from playing the action $a$ given it has faced state $s$ at time $t$, and $Q(s_{t+1},a_{t+1})$ is the current estimate of the reward that will be obtained by the agent from playing action $a$ given it will face state $s$ at time $t+1$. It's worth noting that the update rule in expression  includes the use of an existing estimate as well as the return from current reward; thus, the update uses a form of \textit{bootstrapping}. As the agent faces the different states in the environment, the accuracy of the estimates increases and the agent learns to take actions that maximize the rewards it obtains. The update procedure just described correspond to the \textit{Sarsa} algorithm, part of the \textit{time difference} methods in reinforcement learning \cite{sutton_reinforcement_1998}.

So far, we have described how the agent makes estimates of values for the states. But, in order to collect the rewards the agent must have a policy to choose actions from the available options. Obviously, such policy should use the values estimated, and tend to choose actions with higher value. However, the policy must also explore alternative actions, instead of choosing the one with the highest value. That is because exploration will lead to better estimates of values, and in a changing environment, to finding actions with higher value. Thus, the policy of a agent must balance the exploration and exploitation of available actions. For this simple model, we will use as action choosing policy the \textit{soft-max} function. In the case of the market experiment, the choosing policy is given by the probability of choosing action \textit{a} over the alternative \textit{b} given the state encountered is \textit{s}, formally

\begin{align}
	P(A_t=a|S_t=s)&=\frac{e^{\nicefrac{Q(s,a)}{\tau}}}{e^{\nicefrac{Q(s,a)}{\tau}}+e^{\nicefrac{Q(s,b)}{\tau}}} \\
	P(A_t=b|S_t=s)&=1-P(A_t=a|S_t=s),
\end{align}

where $\tau$ is a parameter that determines how important is the difference in value between the two options when the agent makes the decision of which action to take. The higher the value of $\tau$ the less important the difference is; and thus, the more exploratory the agent is. Figure  shows the probability that the agent chooses one of the two options as a function of the difference in estimated value of the two actions, for different values of $\tau$.

In this simple model three parameters define the learning strategy of agents: $alpha$ the speed of learning, $\gamma$ the importance of future rewards in the estimation of value, and $\tau$ the importance of value in action choosing policy. In a first exploration, we will explore how variation in those parameters influences the capacity of agents to find a solution to the market experiment. Furthermore, using the natural history of cleaner wrasse, we could asses how natural selection changes the values in those parameters - assuming they are inherited - and the learning strategy.

\subsection*{Ecological environment}

As a first approximation to the ecological environment faced by learning agents, I will assume that potential cleaning options are replaced by one of the three options (resident, visitor, absence) with a certain constant probability. In the preliminary results presented here I will assume that resident and visitor are equally likely to fill the vacant position, they are both picked with probability 0.2; thus the position remains empty with the complementary probability 0.6. As described before, the market experiment assumes that visitors and residents leave the cleaning station with different probabilities when they are not immediately served. To capture this, I will set two parameters that define the probabilities that they leave when the cleaner fails to serve them. If they are served, I assume both types leave.

\subsection*{Results}
\subsubsection{State-action pairs value estimation}

Figure  shows the change in the estimated value of choosing a visitor and a resident when both are a cleaning option. Different panels present learning series with combinations of values for parameters $\gamma$ and $\tau$. The bottom row of panels, that corresponds to learning series with $\gamma = 0$, show that the estimated values for the two options do not differ from each other. This result is expected given that I assumed both of them to provide the same reward. The top two panel rows where $\gamma > 0 $, in contrast, show that the value estimated for the two options differ. So, visitor and resident are distinguishable as having different values  only when the learning agent includes future rewards as part of the estimate of value of current actions ($\gamma>0$). The difference in estimate between the options is, however, more clear as the value of $\tau$ increases (see Fig.  centre and right columns). With low $\tau$ there is more variation in the estimated values between different learning rounds. Low values of $\tau$ result in learning agents whose decision making gives a height weight to estimated values, avoiding actions with low estimated value. Therefore, the initial rewards obtained during training are more prone to bias the estimated values. The variation in the estimated values in the first column of panels in figure  is evidence that stochasticity plays a big role in the learning process.

The values of state-action pairs translates into probabilities of taking actions when agents face each state. Figure  depicts the frequency of taking the visitor option, when agents face the resident-visitor state, in the second half of the training when most values have reached a stable state. Scenarios where there is wide variation between the estimates of different learning series (fig.  correspond to variation in the frequency of of choosing a visitor (fig. ). The wide boxes in the boxplot, evidences that the variation stems from the learning series reaching one or two outcomes; either the agent chooses always a visitor or always a resident. In contrast, scenarios where the estimation is consistent among series, the probability of choosing a visitor reaches intermediate values. When the agent does not consider the future in the estimates of value ($\gamma = 0$, red boxes in fig. ) this intermediate value is 0.5, which corresponds to an equal estimate for both actions (fig. ). While,  when the agent considers future moves in the estimates, the visitor option has a higher value than a resident, and the probability of choosing a visitor is above 0.5.

Figure  shows the rewards accumulated in the second half of the learning series, corresponding to the dynamics of figure , and the frequency of a visitor choice in fig. , discriminated by the values used in parameters $\gamma$ and $\tau$. Scenarios where the agent learns to choose visitors over residents do not necessarily yield higher accumulated rewards. Furthermore, the scenarios that reached more accurate and precise value estimations also yielded lower rewards, particularly when $\tau=5$ (see fig.  $\gamma=0$, $\tau = 5$; and $\gamma=0.5$, $\tau=5$). Note that under the current implementation, agents do not only learn to choose visitors over residents; but also they learn to choose visitor and residents over absence of client. Thus, the differences in reward can stem from agents that do not always choose any client over the absence of client.

In order to find out whether the low accumulated rewards of some of the scenarios was due to the 'silly' decision of choosing nothing over a client; I run learning series 'hard-wiring' the correct decision when agent face either of the two types of client over the absence of clients. Figures , and  show the dynamics of estimated values, the probability of choosing a visitor and the accumulated rewards respectively, of this new implementation. Under this new set-up, the probability of choosing a visitor explains the accumulated rewards of the different scenarios. As for the variation found in the different scenarios, we retain the result from the original set-up, that higher values of $\tau$, and lower values of $\gamma$ favour more accurate estimations of value.

\subsubsection{Action value estimation}
\label{sec:actions}

\subsubsection{Negative reinforcement}

\section{A more realistic model}

\begin{equation}
Q=\sum_{i}^{n} f_{i} w_{i}
\label{equation:action_value_gen}
\end{equation}

where $f_i$ are quantifications of client's morphological features; $w_i$ are the estimated weights of those features and $n$ is the number of features that the agent is using for the estimation.

Each one of the weights is updated after the agent obtains reward and sees the following state. This update is determined by the difference between the estimated value and the 'observed' value, weighted by the morphological feature of the current client it is associated with. This way, the change is larger when the estimated value is far from the observed value, and weights associated with higher quantities of  the morphological features are blame more for the mismatch. It's worth noting that the 'observed' value is a combination of the reward obtained from the current choice and the estimation of value for the future choice. In that sense, just like in the simple case, the 'observed' value is not strictly observed. It is partly bootstrapped by the estimate of the future value. Formally, the update is given by

\begin{equation}
w_i\leftarrow w_i +\alpha [R_{t+1}+\gamma Q(s_{t+1},a_{t+1})-Q(s_t,a_t)]f_i,
\end{equation}

\subsubsection{State-action pair value estimation}
Given that for estimating value of state-action action pairs agents must assign a value to the pair of clients, rather than just to an individual client, we allow the function approximator to use as inputs the morphological features of both clients that are presented at the same time. Thus, the agent updates weights for the morphological features of two clients. Furthermore, the agent must estimate value for two potential actions; that is choosing each one of the two clients. To allow these two estimates to be different, the agents estimates different weights for the two actions. All in all, if a client is characterized by $n$ morphological features, the estimation of value for the state-action pairs must include the update $4n$ weights. Formally,

\begin{align}
Q(S_t,a)&=\sum_{j}^{2}\sum_{i}^{n} f_{ij} w(a)_{ij} \\
Q(S_t,b)&=\sum_{j}^{2}\sum_{i}^{n} f_{ij} w(b)_{ij} 
\end{align}

\subsubsection{Action value estimation}

%\begin{thebibliography}{1}
%	
%\end{thebibliography}

\bibliography{Cleanerlearning}
\bibliographystyle{plain}

\end{document}

### 2019-06-20 QD_report.Rmd
Rmd file with a quick and dirty report that runs and produces pdf and html

---
title: "The evolution of badges of status with learners"
author: "Andrés Quiñones"
output: 
  pdf_document: 
    fig_caption: yes
    keep_tex: yes
header-includes: \usepackage{float}
                \floatplacement{figure}{H}
---

# The Hawk-Dove game

Classical Hawk-Dove game. Hawks compete with doves in a population of 
reproductive individuals.

![Hawk-dove game](Simulations//mutType_//basicHawkDove.png "H-Dgame")

# The effect of learning

A population of individuals that estimate a value and a preference for a 
particular behaviour. In our case they specify the probability of playing dove 
in the Hawk-Dove game.

## Result when individuals do NOT vary in their quality

![Learning equilibrium when individuals do not vary in quality](Simulations//QualStDv_//hawkDoveLearn.png "HD learn qual. invariance")

## Results when individuals vary in their quality

![Learning equilibrium when individuals vary in quality](Simulations//QualStDv_//hawkDoveLearn_0.1.png "HD learn qual. 0.1 ")

## Over all effect of variance

![ Effect of variation in quality for the learning equilibrium](Simulations//QualStDv_//effectQualVariance.png "Quality variance effect")

## How does learning work?

![Function approximation for2 the actor and the critic](cartoonRBF.png "Learning cartoonRBF")

## How does learning look like in the Hawk-Dove game

![Within population variation in the learning parameters at the end of learning for three scenarios of variation in the quality of individuals](Simulations//QualStDv_//WeightsVarQualSt.png "Quality invariance")

# How do learners fare against the "pure" types

![Evolutionary dynamics of the learners competing against pure strategies](Simulations//mutType_//evolDynTypes.png "Types competition")

## How do the learning dynamics look like

![Learning dynamics of the individuals in a population](Simulations//mutType_//learnDyn200.png "Learning dynamics")

# What's next?

Let reaction norm evolve, under different initial conditions.

### 2019-06-20 QD_report.tex
Rmd file with a quick and dirty report that runs and produces pdf and html

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={The evolution of badges of status with learners},
            pdfauthor={Andres Quiñones},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\providecommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{The evolution of badges of status with learners}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Andres Quiñones}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
    \date{}
    \predate{}\postdate{}

\usepackage{float}

\floatplacement{figure}{H}

\begin{document}
\maketitle

\hypertarget{the-hawk-dove-game}{%
\section{The Hawk-Dove game}\label{the-hawk-dove-game}}

Classical Hawk-Dove game. Hawks compete with doves in a population of
reproductive individuals.

\begin{figure}
\centering
\includegraphics{Simulations//mutType_//basicHawkDove.png}
\caption{Hawk-dove game}
\end{figure}

\hypertarget{the-effect-of-learning}{%
\section{The effect of learning}\label{the-effect-of-learning}}

A population of individuals that estimate a value and a preference for a
particular behaviour. In our case they specify the probability of
playing dove in the Hawk-Dove game.

\hypertarget{result-when-individuals-do-not-vary-in-their-quality}{%
\subsection{Result when individuals do NOT vary in their
quality}\label{result-when-individuals-do-not-vary-in-their-quality}}

\begin{figure}
\centering
\includegraphics{Simulations//QualStDv_//hawkDoveLearn.png}
\caption{Learning equilibrium when individuals do not vary in quality}
\end{figure}

\hypertarget{results-when-individuals-vary-in-their-quality}{%
\subsection{Results when individuals vary in their
quality}\label{results-when-individuals-vary-in-their-quality}}

\begin{figure}
\centering
\includegraphics{Simulations//QualStDv_//hawkDoveLearn_0.1.png}
\caption{Learning equilibrium when individuals vary in quality}
\end{figure}

\hypertarget{over-all-effect-of-variance}{%
\subsection{Over all effect of
variance}\label{over-all-effect-of-variance}}

\begin{figure}
\centering
\includegraphics{Simulations//QualStDv_//effectQualVariance.png}
\caption{Effect of variation in quality for the learning equilibrium}
\end{figure}

\hypertarget{how-does-learning-work}{%
\subsection{How does learning work?}\label{how-does-learning-work}}

\begin{figure}
\centering
\includegraphics{cartoonRBF.png}
\caption{Function approximation for0 the actor and the critic}
\end{figure}

\hypertarget{how-does-learning-look-like-in-the-hawk-dove-game}{%
\subsection{How does learning look like in the Hawk-Dove
game}\label{how-does-learning-look-like-in-the-hawk-dove-game}}

\begin{figure}
\centering
\includegraphics{Simulations//QualStDv_//WeightsVarQualSt.png}
\caption{Within population variation in the learning parameters at the
end of learning for three scenarios of variation in the quality of
individuals}
\end{figure}

\hypertarget{how-do-learners-fare-against-the-pure-types}{%
\section{How do learners fare against the ``pure''
types}\label{how-do-learners-fare-against-the-pure-types}}

\begin{figure}
\centering
\includegraphics{Simulations//mutType_//evolDynTypes.png}
\caption{Evolutionary dynamics of the learners competing against pure
strategies}
\end{figure}

\hypertarget{how-do-the-learning-dynamics-look-like}{%
\subsection{How do the learning dynamics look
like}\label{how-do-the-learning-dynamics-look-like}}

\begin{figure}
\centering
\includegraphics{Simulations//mutType_//learnDyn200.png}
\caption{Learning dynamics of the individuals in a population}
\end{figure}

\hypertarget{whats-next}{%
\section{What's next?}\label{whats-next}}

Let reaction norm evolve, under different initial conditions.

\end{document}

### 2019-06-20 CartoonRBF.R

@@ -5,14 +5,23 @@ here()
 source("C:/Users/a.quinones/Dropbox/R_files/posPlots.R")
 
 
-
+png(here("cartoonRBF.png"),width=800,height = 800)
 nCenters<-5
 interv<-1/nCenters
 centers<-interv*0.5+interv*seq(0,nCenters-1)
-weights<-runif(nCenters)#RBF(centers,0.5,0.05)
+weights<-runif(nCenters,min=-1,max=1)#RBF(centers,0.5,0.05)
 rangx<-seq(0,1,length=1000)
 
-par(plt=posPlot())
+par(plt=posPlot(numploty = 2,idploty = 2))
+plot(logist(totRBF(rangx,centers,0.01,weights),alpha = 0,beta=1)~rangx,type='l',col=1,
+     xlab="",ylab="p(Dove)",ylim=c(0,1),lwd=3,xaxt="n",cex.lab=1.5)
+points(y=logist(weights,alpha = 0,beta=1),x=centers,cex=3)
+
+
+weights<-runif(nCenters)#RBF(centers,0.5,0.05)
+
+par(plt=posPlot(numploty = 2,idploty = 1),new=T)
 plot(totRBF(rangx,centers,0.01,weights)~rangx,type='l',col=1,
-     xlab="x",ylab="response",ylim=c(0,3),lwd=3)
-points(y=weights,x=centers,cex=3)
+     xlab="Badge",ylab="Estimated value",ylim=c(0,1),lwd=3,cex.lab=1.5)
+points(y=logist(weights,alpha = 0,beta=1),x=centers,cex=3)
+dev.off()

### 2019-06-20 dynSevFiles.R

@@ -62,7 +62,7 @@ evolStats<-evol[,.(m.freqGenHawk=mean(freqGenHawks),
 
 # Plot mean and IQRs of the genotypes and phenotypes ----------------------------
 
-png(here("Simulations",scenario,"basidHawkDove.png"),height = 800,width = 800)
+png(here("Simulations",paste0(scenario,"_"),"effectQualVariance.png"),height = 800,width = 800)
 
 timeChoi<-max(evolStats[,time])
 par(plt=posPlot(numploty = 2,idploty = 2),xaxt="n",las=1)

### 2019-06-20 dynamics.R

@@ -11,7 +11,7 @@ source(here("AccFunc.R"))
 
 # Scenario to be plotted - corresponds to folders where simulations are stored
 
-scenario<-"mutType_"
+scenario<-"QualStDv_"
 
 
 # Load files -------------------------------------------------------------------
@@ -19,7 +19,7 @@ scenario<-"mutType_"
 (listTest<-list.files(here("Simulations",scenario)))
 (sdList<-grep("evol",listTest,value=TRUE))
 
-evol<-fread(here("Simulations",scenario,sdList[4]))
+evol<-fread(here("Simulations",scenario,sdList[2]))
 
 # Extract means and IQR for the dynamic variables ------------------------------
 
@@ -66,7 +66,7 @@ evolStats<-evol[,.(m.freqGenHawk=mean(freqGenHawks),
 
 # Plot mean and IQRs of the genotypes and phenotypes ----------------------------
 
-png(here("Simulations",scenario,"evolDynTypes.png"),height = 800,width = 800)
+png(here("Simulations",scenario,"hawkDoveLearn_0.1.png"),height = 800,width = 800)
 
 par(plt=posPlot(numploty = 2,idploty = 2),xaxt="s",las=2)
 plot(x=c(0,max(evolStats$time)),y=c(0.5,0.5),type="l",lwd=2,col="grey",
@@ -92,8 +92,8 @@ with(evolStats,{
 })
 
 
-legend("topleft",legend=c("Hawk","Dove","Learner"),
-       lty=c(1,1,1),lwd=2,col=colTypesLin,bty="o",cex=1)
+legend("topleft",legend=c("Hawk","Dove"),
+       lty=c(1,1),lwd=2,col=colTypesLin[1:2],bty="o",cex=1)
 
 
 par(plt=posPlot(numploty = 2,idploty = 1),xaxt="n",las=1,new=TRUE,xpd=T)
@@ -130,7 +130,8 @@ legend("topright",legend=c("HH","HD","DD"),
 
 par(plt=posPlot(numploty = 2,idploty = 1),xaxt="n",las=1,new=TRUE,xpd=T)
 plot(x=c(0,max(evolStats$time)),y=c(0.5,0.5),type="l",lwd=2,col="grey",
-     ylim=c(0,1),xlab="",ylab="",cex.lab=1.5,cex.axis=1.2,xaxt='s')
[…]

### 2019-06-26 QD_report.Rmd
Elaborate and reoganize the QD report

@@
 
-Classical Hawk-Dove game. Hawks compete with doves in a population of 
-reproductive individuals.
+Classical Hawk-Dove game.
+Individuals have one fo two genetically determined phenotypic strategies.
+_Hawks_ are willing to start a conflict over resources, while _doves_ prefer to stand down in the hope to share the resource without an aggressive contest.
+Individuals interact randomly with each other over their lifetimes and collect resources depending on the strategy of theirs and their partners.
+In contrast to the classical _Hawk-dove_, in here when to _hawks_ interact with each other one of them wins the contest and gets $V-\frac{C}{2}$ as payoff; while the loser gets $-\frac{C}{2}$.
+Where $V$ is the value of the resource individuals are competing for, and $C$ is the cost of an agressive contest.
+individuals share the cost, but only the winner gets the value of the resource.The rest of payoff matrix follows the standard game.
 
-![Hawk-dove game](Simulations//mutType_//basicHawkDove.png "H-Dgame")
+Individuals may or may not vary in their quality.
+Quality is a numerical value from 0 to 1, drawn from a truncated normal distribution when each individual is born.
+Quality represents the different conditions under which individuals might be raised, and may influence their competitive abilities.
+Differences in quality between two individuals partly determine who wins in a conflict when both players play hawk.
+The probability of individual A wining a conflict over individual B is given by
+
+\begin{equation}
+p(A)=\frac{1}{1+e^{-\beta(Q_A-Q_B)}},
+\end{equation}
+
+where $Q_A$ and $Q_b$ are the qualities of individuals A and B respectively, and $\beta$ is how important quality is in determining who wins the contest.
+Individuals' reproductive success is proportional to the payoff accumulated
+throughout their lifes. Figure \ref{fig:HD_game} shows the evolutionary dynamics of  the two genotypes, which fit the game theoretical prediction.
+
+```{r fig1, echo=FALSE, fig.cap="\\label{fig:HD_game}Hawk-dove game. Dashed line is the  game theoretical prediction for frequency of hawks.", out.width = '80%'}
+knitr::include_graphics("Simulations//mutType_//basicHawkDove.png")
+```
 
@@
 
-A population of individuals that estimate a value and a preference for a 
-particular behaviour. In our case they specify the probability of playing dove 
-in the Hawk-Dove game.
+We let individuals have a _badge of status_ that potentially signals their quality.
+To allow for this signalling, we let the size of the _badge of status_ of an individual be dependent on her quality through a reaction norm.
+we assume the size of the _badge_ is a value from 0 to 1.
+This value is given by
+\begin{equation}
+s_i = \frac{1}{1+e^{-(\alpha_i+\gamma_iQ_i)}}
+\label{eq:react_norm}
+\end{equation}
+The parameters $\alpha$ and $\gamma$ in equation \ref{eq:react_norm} determine the
+shape of the reaction norm, and hence, how honest the _badge_ is in signalling the
+quality of the individual.Eventually, the aim is to allow the signalling system to
+evolve, by letting the parameters of the reaction norm be inherited from parent to offspring and be change by a mutational process.
 
-## Result when individuals do NOT vary in their quality
+## How does learning work?
 
-![Learning equilibrium when individuals do not vary in quality](Simulations//QualStDv_//hawkDoveLearn.png "HD learn qual. invariance")
+To complete the receiver side of the signalling system, we introduce a new strategy in the game that we call _learner_.
+_Learners_ use a mixed strategy in the hawk-dove game.
+Which means, they play hawk (dove) with a certain probability.
+Furthermore, the probability with which the play each strategy changes during their lifetime according to the payoff they recieve in previous interactions.
+In other words, _learners_ learn to behave in the game.
+To implement the learning
+process, we use the reinforcement learning formalism.
+Specifically, we use the actor-critic algorithm.
+Individuals ( _learners_ ) perceive the _badge_ of their counterpart as the environmental state in which they find themselves.
+They estimate the value of that state and a preference for playing hawk (dove) in that state.
+They update their estimation and preference after each interaction they have.
+Given that the badge size is a real number between 0 and 1, there are infinitely many environmental states.
+Thus, for their value estimation and preference computation, individual must use a form of generalization.
+We let them
+use as generalization a function approximation method based on radial basis functions.
+Each individual uses five _centers_ spread uniformly throughout the range of badge size variation (from 0 to 1).
+Each time an individual interacts with a conspecific the size of the badge of the interacting partner triggers are response from each one of the _centers_.
+This response decreases in strength as the badge size is further from the _center_.
+The decrease follows a radial basis function with gaussian shape.
+The sum of the responses given by all the centers provide the value estimation and the preference for one of the two pure strategies (hawk or dove).
+During the learning process individuals update the response that each one of these centers triggers.
+In the case of the preference for the strategies the numerical value resulting from the sum of the responses is translated to a probability of one of the two pure strategies through a logistic function.
+Figure \ref{fig:learning_cartoonRBF} shows a made up example of how the function approximation works for the actor (top panel - probability of choosing dove) and the critic (bottom panel - expected payoff from facing an individual with a given badge size).
+
+```{r fig2, echo=FALSE, fig.cap="\\label{fig:learning_cartoonRBF}Function approximation for the actor and the critic.
+Using random values for the responses of each center.
+Circles show the location of the centers and the maximum response they trigger, while the lines show the total response along the badge size continium.", out.width = '80%'}
+knitr::include_graphics("cartoonRBF.png")
+```
+
+## Results when individuals do NOT vary in their quality
+
+In Figure \ref{fig:learning_invar} we show the phenotypic frequencies and interaction-type frequencies in 4 generations of a population composed of only _learners_ that do not vary in their quality.
+The population is also monomorphic for the reaction norm parameters ($\alpha=0$ and $\gamma=0$).
+With the values chosen the badge size distribution is the same as the quality distribution.
+Thus, the four generations respresent just replicates of the learning process.
+The results show that under quality invariance, the learning algorithm develops similar phenotypic frequencies as the evolutionary dynamics of the pure strategies.
+
+```{r fig3, echo=FALSE, fig.cap="\\label{fig:learning_invar}Learning equilibrium when individuals do not vary in quality.
+Dashed line is the  game theoretical prediction for frequency of hawks.", out.width = '80%'}
+knitr::include_graphics("Simulations//QualStDv_//hawkDoveLearn.png")
+```
 
@@
 
-![Learning equilibrium when individuals vary in quality](Simulations//QualStDv_//hawkDoveLearn_0.1.png "HD learn qual. 0.1 ")
+```{r fig4, echo=FALSE, fig.cap="\\label{fig:learning_var0.1}Learning equilibrium when individuals vary in quality.
+Dashed lines are the  game theoretical predictions in the standard game.", out.width = '80%'}
+knitr::include_graphics("Simulations//QualStDv_//hawkDoveLearn_0.1.png")
+```
 
-## Over all effect of variance
+## Overal effect of variance
 
-![ Effect of variation in quality for the learning equilibrium](Simulations//QualStDv_//effectQualVariance.png "Quality variance effect")
-
-## How does learning work?
-
-![Function approximation for2 the actor and the critic](cartoonRBF.png "Learning cartoonRBF")
+```{r fig5, echo=FALSE, fig.cap="\\label{fig:learning_all_var}Effect of variation in quality for the learning equilibrium.", out.width = '80%'}
+knitr::include_graphics("Simulations//QualStDv_//effectQualVariance.png")
+```
 
@@
 
-![Within population variation in the learning parameters at the end of learning for three scenarios of variation in the quality of individuals](Simulations//QualStDv_//WeightsVarQualSt.png "Quality invariance")
-
-# How do learners fare against the "pure" types
-
-![Evolutionary dynamics of the learners competing against pure strategies](Simulations//mutType_//evolDynTypes.png "Types competition")
+```{r fig6, echo=FALSE, fig.cap="\\label{fig:AC_HD}Within population variation in the learning parameters at the end of learning for three scenarios of variation in the quality of individuals.", out.width = '80%'}
+knitr::include_graphics("Simulations//QualStDv_//WeightsVarQualSt.png")
+```
 
@@
 
-![Learning dynamics of the individuals in a population](Simulations//mutType_//learnDyn200.png "Learning dynamics")
+```{r fig7, echo=FALSE, fig.cap="\\label{fig:learn_dyn}Learning dynamics of the individuals in a population.", out.width = '80%'}
+knitr::include_graphics("Simulations//mutType_//learnDyn200.png")
+```
+
+# How do learners fare against the "pure" strategies
+
+```{r fig8, echo=FALSE, fig.cap="\\label{fig:types_comp}Evolutionary dynamics of the learners competing against pure strategies.", out.width = '80%'}
+knitr::include_graphics("Simulations//mutType_//evolDynTypes.png")
+```
 

### 2019-06-26 QD_report.tex
Elaborate and reoganize the QD report

@@
             pdftitle={The evolution of badges of status with learners},
-            pdfauthor={Andres Quiñones},
+            pdfauthor={Andrés Quiñones},
             pdfborder={0 0 0},
@@
   \posttitle{\par}
-    \author{Andres Quiñones}
+    \author{Andrés Quiñones}
     \preauthor{\centering\large\emph}
@@
 
-Classical Hawk-Dove game. Hawks compete with doves in a population of
-reproductive individuals.
+Classical Hawk-Dove game. Individuals have one fo two genetically
+determined phenotypic strategies. \emph{Hawks} are willing to start a
+conflict over resources, while \emph{doves} prefer to stand down in the
+hope to share the resource without an aggressive contest. Individuals
+interact randomly with each other over their lifetimes and collect
+resources depending on the strategy of theirs and their partners. In
+contrast to the classical \emph{Hawk-dove}, in here when to \emph{hawks}
+interact with each other one of them wins the contest and gets
+\(V-\frac{C}{2}\) as payoff; while the loser gets \(-\frac{C}{2}\).
+Where \(V\) is the value of the resource individuals are competing for,
+and \(C\) is the cost of an agressive contest. individuals share the
+cost, but only the winner gets the value of the resource.The rest of
+payoff matrix follows the standard game.
+
+Individuals may or may not vary in their quality. Quality is a numerical
+value from 0 to 1, drawn from a truncated normal distribution when each
+individual is born. Quality represents the different conditions under
+which individuals might be raised, and may influence their competitive
+abilities. Differences in quality between two individuals partly
+determine who wins in a conflict when both players play hawk. The
+probability of individual A wining a conflict over individual B is given
+by
+
+\begin{equation}
+p(A)=\frac{1}{1+e^{-\beta(Q_A-Q_B)}},
+\end{equation}
+
+where \(Q_A\) and \(Q_b\) are the qualities of individuals A and B
+respectively, and \(\beta\) is how important quality is in determining
+who wins the contest. Individuals' reproductive success is proportional
+to the payoff accumulated throughout their lifes. Figure
+\ref{fig:HD_game} shows the evolutionary dynamics of the two genotypes,
+which fit the game theoretical prediction.
 
 \begin{figure}
-\centering
-\includegraphics{Simulations//mutType_//basicHawkDove.png}
-\caption{Hawk-dove game}
+\includegraphics[width=0.8\linewidth]{Simulations//mutType_//basicHawkDove} \caption{\label{fig:HD_game}Hawk-dove game.
+Dashed line is the  game theoretical prediction for frequency of hawks.}\label{fig:fig1}
 \end{figure}
@@
 
-A population of individuals that estimate a value and a preference for a
-particular behaviour. In our case they specify the probability of
-playing dove in the Hawk-Dove game.
+We let individuals have a \emph{badge of status} that potentially
+signals their quality. To allow for this signalling, we let the size of
+the \emph{badge of status} of an individual be dependent on her quality
+through a reaction norm. we assume the size of the \emph{badge} is a
+value from 0 to 1. This value is given by \begin{equation}
+s_i = \frac{1}{1+e^{-(\alpha_i+\gamma_iQ_i)}}
+\label{eq:react_norm}
+\end{equation} The parameters \(\alpha\) and \(\gamma\) in equation
+\ref{eq:react_norm} determine the shape of the reaction norm, and hence,
+how honest the \emph{badge} is in signalling the quality of the
+individual.Eventually, the aim is to allow the signalling system to
+evolve, by letting the parameters of the reaction norm be inherited from
+parent to offspring and be change by a mutational process.
 
-\hypertarget{result-when-individuals-do-not-vary-in-their-quality}{%
-\subsection{Result when individuals do NOT vary in their
-quality}\label{result-when-individuals-do-not-vary-in-their-quality}}
+\hypertarget{how-does-learning-work}{%
+\subsection{How does learning work?}\label{how-does-learning-work}}
+
+To complete the receiver side of the signalling system, we introduce a
+new strategy in the game that we call \emph{learner}. \emph{Learners}
+use a mixed strategy in the hawk-dove game. Which means, they play hawk
+(dove) with a certain probability. Furthermore, the probability with
+which the play each strategy changes during their lifetime according to
+the payoff they recieve in previous interactions. In other words,
+\emph{learners} learn to behave in the game. To implement the learning
+process, we use the reinforcement learning formalism. Specifically, we
+use the actor-critic algorithm. Individuals ( \emph{learners} ) perceive
+the \emph{badge} of their counterpart as the environmental state in
+which they find themselves. They estimate the value of that state and a
+preference for playing hawk (dove) in that state. They update their
+estimation and preference after each interaction they have. Given that
+the badge size is a real number between 0 and 1, there are infinitely
+many environmental states. Thus, for their value estimation and
+preference computation, individual must use a form of generalization. We
+let them use as generalization a function approximation method based on
+radial basis functions. Each individual uses five \emph{centers} spread
+uniformly throughout the range of badge size variation (from 0 to 1).
+Each time an individual interacts with a conspecific the size of the
+badge of the interacting partner triggers are response from each one of
+the \emph{centers}. This response decreases in strength as the badge
+size is further from the \emph{center}. The decrease follows a radial
+basis function with gaussian shape. The sum of the responses given by
+all the centers provide the value estimation and the preference for one
+of the two pure strategies (hawk or dove). During the learning process
+individuals update the response that each one of these centers triggers.
+In the case of the preference for the strategies the numerical value
+resulting from the sum of the responses is translated to a probability
+of one of the two pure strategies through a logistic function. Figure
+\ref{fig:learning_cartoonRBF} shows a made up example of how the
+function approximation works for the actor (top panel - probability of
+choosing dove) and the critic (bottom panel - expected payoff from
+facing an individual with a given badge size).
 
 \begin{figure}
-\centering
-\includegraphics{Simulations//QualStDv_//hawkDoveLearn.png}
-\caption{Learning equilibrium when individuals do not vary in quality}
+\includegraphics[width=0.8\linewidth]{cartoonRBF} \caption{\label{fig:learning_cartoonRBF}Function approximation for the actor and the critic.
+Using random values for the responses of each center.
+Circles show the location of the centers and the maximum response they trigger, while the lines show the total response along the badge size continium.}\label{fig:fig2}
+\end{figure}
+
+\hypertarget{results-when-individuals-do-not-vary-in-their-quality}{%
+\subsection{Results when individuals do NOT vary in their
+quality}\label{results-when-individuals-do-not-vary-in-their-quality}}
+
+In Figure \ref{fig:learning_invar} we show the phenotypic frequencies
+and interaction-type frequencies in 4 generations of a population
+composed of only \emph{learners} that do not vary in their quality. The
+population is also monomorphic for the reaction norm parameters
+(\(\alpha=0\) and \(\beta=0\)). With the values chosen the badge size
+distribution is the same as the quality distribution. Thus, the four
+generations respresent just replicates of the learning process. The
+results show that under quality invariance, the learning algorithm
+develops similar phenotypic frequencies as the evolutionary dynamics of
+the pure strategies.
+
+\begin{figure}
+\includegraphics[width=0.8\linewidth]{Simulations//QualStDv_//hawkDoveLearn} \caption{\label{fig:learning_invar}Learning equilibrium when individuals do not vary in quality.
+Dashed line is the  game theoretical prediction for frequency of hawks.}\label{fig:fig3}
 \end{figure}
@@
 \begin{figure}
-\centering
-\includegraphics{Simulations//QualStDv_//hawkDoveLearn_0.1.png}
-\caption{Learning equilibrium when individuals vary in quality}
+\includegraphics[width=0.8\linewidth]{Simulations//QualStDv_//hawkDoveLearn_0.1} \caption{\label{fig:learning_var0.1}Learning equilibrium when individuals vary in quality.
+Dashed lines are the  game theoretical predictions in the standard game.}\label{fig:fig4}
 \end{figure}
 
-\hypertarget{over-all-effect-of-variance}{%
-\subsection{Over all effect of
-variance}\label{over-all-effect-of-variance}}
+\hypertarget{overal-effect-of-variance}{%
+\subsection{Overal effect of variance}\label{overal-effect-of-variance}}
 
 \begin{figure}
-\centering
-\includegraphics{Simulations//QualStDv_//effectQualVariance.png}
-\caption{Effect of variation in quality for the learning equilibrium}
-\end{figure}
-
-\hypertarget{how-does-learning-work}{%
-\subsection{How does learning work?}\label{how-does-learning-work}}
-
-\begin{figure}
-\centering
-\includegraphics{cartoonRBF.png}
-\caption{Function approximation for0 the actor and the critic}
+\includegraphics[width=0.8\linewidth]{Simulations//QualStDv_//effectQualVariance} \caption{\label{fig:learning_all_var}Effect of variation in quality for the learning equilibrium.}\label{fig:fig5}
 \end{figure}
@@
 \begin{figure}
-\centering
-\includegraphics{Simulations//QualStDv_//WeightsVarQualSt.png}
-\caption{Within population variation in the learning parameters at the
-end of learning for three scenarios of variation in the quality of
-individuals}
-\end{figure}
-
-\hypertarget{how-do-learners-fare-against-the-pure-types}{%
-\section{How do learners fare against the ``pure''
-types}\label{how-do-learners-fare-against-the-pure-types}}
-
-\begin{figure}
-\centering
-\includegraphics{Simulations//mutType_//evolDynTypes.png}
-\caption{Evolutionary dynamics of the learners competing against pure
-strategies}
+\includegraphics[width=0.8\linewidth]{Simulations//QualStDv_//WeightsVarQualSt} \caption{\label{fig:AC_HD}Within population variation in the learning parameters at the end of learning for three scenarios of variation in the quality of individuals.}\label{fig:fig6}
 \end{figure}
@@
 \begin{figure}
-\centering
-\includegraphics{Simulations//mutType_//learnDyn200.png}
-\caption{Learning dynamics of the individuals in a population}
+\includegraphics[width=0.8\linewidth]{Simulations//mutType_//learnDyn200} \caption{\label{fig:learn_dyn}Learning dynamics of the individuals in a population.}\label{fig:fig7}
+\end{figure}
+
+\hypertarget{how-do-learners-fare-against-the-pure-strategies}{%
+\section{How do learners fare against the ``pure''
+strategies}\label{how-do-learners-fare-against-the-pure-strategies}}
+
+\begin{figure}
+\includegraphics[width=0.8\linewidth]{Simulations//mutType_//evolDynTypes} \caption{\label{fig:types_comp}Evolutionary dynamics of the learners competing against pure strategies.}\label{fig:fig8}
 \end{figure}

### 2019-06-26 parTOjson.R

@@ -5,24 +5,24 @@ library("jsonlite")
 library("here")
 
 
-fileName<-"parameters.json"
+fileName<-"parameters2.json"
 
 
-param<-list(totGen   = 100000,   nRep    = 15,
-            printGen = 200,   printLearn = 200,
-            printLearnInt = 200,
+param<-list(totGen   = 10000,   nRep    = 15,
+            printGen = 400,   printLearn = 400,
+            printLearnInt = 150,
             popSize  = 1000,  baselineFit = 2,   
             MutSd    = 0.1,
-            nInt        = 1000,  init     = c(0,0,1),
+            nInt        = 750,  init     = c(0,0,1),
             mutRate  = 0.02,  mutType  = 0,
-            sampleSize = 10,   strQual  = 10,
+            sampleSize = 5,   strQual  = 10,
             alphaBad	 = 0,    betaBad	 = 0,
             alphaCrit  = 0.2,  alphaAct = 0.2,
             sigSq   	 = 0.01, nCenters = 5,
             QualStDv   = 0.2,  
             payoff_matrix = c(1.5,1,0,0.5),
             namParam = "baselineFit",
-            rangParam = c(2),
+            rangParam = c(2.1),
             folder=paste(here("Simulations"),"/",sep=""))
 
 

### 2019-07-01 QD_report.Rmd
Changes in the text of the Rmd, and adding the bibtex file with references.

@@
 author: "Andrés Quiñones"
+bibliography: SocialMorphCue.bib
 output: 
@@
 Individuals' reproductive success is proportional to the payoff accumulated
-throughout their lifes. Figure \ref{fig:HD_game} shows the evolutionary dynamics of  the two genotypes, which fit the game theoretical prediction.
+throughout their lifes. Figure \ref{fig:HD_game} shows the evolutionary dynamics 
+of  the two genotypes, which fit the game theoretical prediction.
 
 ```{r fig1, echo=FALSE, fig.cap="\\label{fig:HD_game}Hawk-dove game. Dashed line is the  game theoretical prediction for frequency of hawks.", out.width = '80%'}
-knitr::include_graphics("Simulations//mutType_//basicHawkDove.png")
+knitr::include_graphics("Simulations/mutType_/basicHawkDove.png")
 ```
@@
 
-We let individuals have a _badge of status_ that potentially signals their quality.
-To allow for this signalling, we let the size of the _badge of status_ of an individual be dependent on her quality through a reaction norm.
-we assume the size of the _badge_ is a value from 0 to 1.
-This value is given by
+We let individuals have a _badge of status_ that potentially signals their 
+quality. To allow for this signalling, we let the size of the _badge of status_ 
+of an individual be dependent on her quality through a reaction norm. we assume the size of the _badge_ is a value from 0 to 1. This value is given by
 \begin{equation}
@@
 \end{equation}
-The parameters $\alpha$ and $\gamma$ in equation \ref{eq:react_norm} determine the
-shape of the reaction norm, and hence, how honest the _badge_ is in signalling the
-quality of the individual.Eventually, the aim is to allow the signalling system to
-evolve, by letting the parameters of the reaction norm be inherited from parent to offspring and be change by a mutational process.
+The parameters $\alpha$ and $\gamma$ in equation \ref{eq:react_norm} determine 
+the shape of the reaction norm, and hence, how honest the _badge_ is in 
+signalling the quality of the individual.Eventually, the aim is to allow the 
+signalling system to evolve, by letting the parameters of the reaction norm be inherited from parent to offspring and be change by a mutational process.
 
@@
 
-To complete the receiver side of the signalling system, we introduce a new strategy in the game that we call _learner_.
-_Learners_ use a mixed strategy in the hawk-dove game.
-Which means, they play hawk (dove) with a certain probability.
+To complete the receiver side of the signalling system, we introduce a new 
+strategy in the game that we call _learner_. _Learners_ use a mixed strategy in 
+the hawk-dove game. Which means, they play hawk (dove) with a certain probability.
 Furthermore, the probability with which the play each strategy changes during their lifetime according to the payoff they recieve in previous interactions.
@@
 process, we use the reinforcement learning formalism.
-Specifically, we use the actor-critic algorithm.
+Specifically, we use the actor-critic algorithm [@sutton_reinforcement_2018].
 Individuals ( _learners_ ) perceive the _badge_ of their counterpart as the environmental state in which they find themselves.
@@
 The sum of the responses given by all the centers provide the value estimation and the preference for one of the two pure strategies (hawk or dove).
-During the learning process individuals update the response that each one of these centers triggers.
+During the learning process individuals update the _weight_ of the response that each one of these centers triggers.
 In the case of the preference for the strategies the numerical value resulting from the sum of the responses is translated to a probability of one of the two pure strategies through a logistic function.
@@
 
+In figure \ref{fig:learning_var0.1} we show results of phenotypic and interaction types frequencies from simulations in which individuals vary in their quality. 
+The value of quality is drawn from a truncated normal distribution with mean 0.5 
+and standard deviation 0.1.
+Compared to the expectation from evolutionary dynamics of the standard game, the learning scenario with qualiy variation shows lower levels of hawks and therefore lower levels of escalated conflict.
+At this point individuals do not respond adaptively to the quality of their partner, instead their preference for a given action is dependent on their own quality.
+
 ```{r fig4, echo=FALSE, fig.cap="\\label{fig:learning_var0.1}Learning equilibrium when individuals vary in quality.
@@
 
+In figure \ref{fig:learning_all_var} we show how the effect of increasing 
+variance in quality further deacrases the amount of individuals in a population 
+expressing the hawk phenotype.
+
 ```{r fig5, echo=FALSE, fig.cap="\\label{fig:learning_all_var}Effect of variation in quality for the learning equilibrium.", out.width = '80%'}
@@
 
-## How does learning look like in the Hawk-Dove game
+## How does learning look like in the Hawk-Dove game?
+
+Figures \ref{fig:learning_var0.1} and \ref{fig:learning_all_var} only show how learners behave at a population level, but they do not show how inter-invidivual variation behind those frequencies.
+In figure \ref{fig:AC_HD} we show the result of the function aproximation mechanism arising at the end of a generation for a sample of the individuals of the populations.
+These results are show under three different levels of variation in quality.
+All three panes show that through their learning processes individuals split up in two groups, one which mostly behaves like a hawk and a second that mostly behaves like a dove.
+The group of "hawks" is mostly, but no exclusively, composed of individuals whose quality is on the upper half of the quality distribution.
+Concordantly, the group of "doves" is mostly composed of individuals whose qualiy is on the lower half of the quality distribution.
+The relation between quality and the prefered behaviour of individuals obviously only holds in scenarios with variance in quality.
+In the scenario with quality invariance, the split of individuals in two groups with a tenedency for each behaviour nevertheless remains.
 
@@
 
+In figure \ref{fig:learn_dyn} we show the learning trajectory, for the actor and the critic, of a sample of individuals.
+These dynamics show how individuals split in the two groups according to their experiences.
+Despite the robustness of the bimodal distribution along the learning dynamics, it is not uncommon that individual go from one group to the other.
+
 ```{r fig7, echo=FALSE, fig.cap="\\label{fig:learn_dyn}Learning dynamics of the individuals in a population.", out.width = '80%'}
@@
 # How do learners fare against the "pure" strategies
+
+Finally, in a new set of simulations we explore the question of how competitive learners are in competition with the pure strategies. Figure \ref{fig:types_comp}
+show the results of competition among the three strategies.
+The population starts with a 1:1 ratio of pure hawks anf doves, and learners are intruced by mutation.
+Learners manage to increase in frequency.
+They, however, do not manage to outcompete the pure streategies.
+In the end, there is a polymorphic equilibrium with the three strategies coexisting, with learners being the most frequen of the threee.
 
@@
 Let reaction norm evolve, under different initial conditions.
+
+# Cited literature

### 2019-07-01 QD_report.tex
Changes in the text of the Rmd, and adding the bibtex file with references.

@@
 
-\usepackage{float}
-
-\floatplacement{figure}{H}
+\usepackage{float} \floatplacement{figure}{H}
 
@@
 
-\hypertarget{the-hawk-dove-game}{%
-\section{The Hawk-Dove game}\label{the-hawk-dove-game}}
+\section{The Hawk-Dove game}\label{the-hawk-dove-game}
 
@@
 \begin{figure}
-\includegraphics[width=0.8\linewidth]{Simulations//mutType_//basicHawkDove} \caption{\label{fig:HD_game}Hawk-dove game.
+\includegraphics[width=0.8\linewidth]{Simulations/mutType_/basicHawkDove} \caption{\label{fig:HD_game}Hawk-dove game.
 Dashed line is the  game theoretical prediction for frequency of hawks.}\label{fig:fig1}
@@
 
-\hypertarget{the-effect-of-learning}{%
-\section{The effect of learning}\label{the-effect-of-learning}}
+\section{The effect of learning}\label{the-effect-of-learning}
 
@@
 through a reaction norm. we assume the size of the \emph{badge} is a
-value from 0 to 1. This value is given by \begin{equation}
+value from 0 to 1. This value is given by
+
+\begin{equation}
 s_i = \frac{1}{1+e^{-(\alpha_i+\gamma_iQ_i)}}
 \label{eq:react_norm}
-\end{equation} The parameters \(\alpha\) and \(\gamma\) in equation
-\ref{eq:react_norm} determine the shape of the reaction norm, and hence,
-how honest the \emph{badge} is in signalling the quality of the
-individual.Eventually, the aim is to allow the signalling system to
-evolve, by letting the parameters of the reaction norm be inherited from
-parent to offspring and be change by a mutational process.
+\end{equation}
 
-\hypertarget{how-does-learning-work}{%
-\subsection{How does learning work?}\label{how-does-learning-work}}
+The parameters \(\alpha\) and \(\gamma\) in equation \ref{eq:react_norm}
+determine the shape of the reaction norm, and hence, how honest the
+\emph{badge} is in signalling the quality of the individual.Eventually,
+the aim is to allow the signalling system to evolve, by letting the
+parameters of the reaction norm be inherited from parent to offspring
+and be change by a mutational process.
+
+\subsection{How does learning work?}\label{how-does-learning-work}
 
@@
 process, we use the reinforcement learning formalism. Specifically, we
-use the actor-critic algorithm. Individuals ( \emph{learners} ) perceive
-the \emph{badge} of their counterpart as the environmental state in
-which they find themselves. They estimate the value of that state and a
-preference for playing hawk (dove) in that state. They update their
-estimation and preference after each interaction they have. Given that
-the badge size is a real number between 0 and 1, there are infinitely
-many environmental states. Thus, for their value estimation and
-preference computation, individual must use a form of generalization. We
-let them use as generalization a function approximation method based on
-radial basis functions. Each individual uses five \emph{centers} spread
-uniformly throughout the range of badge size variation (from 0 to 1).
-Each time an individual interacts with a conspecific the size of the
-badge of the interacting partner triggers are response from each one of
-the \emph{centers}. This response decreases in strength as the badge
-size is further from the \emph{center}. The decrease follows a radial
-basis function with gaussian shape. The sum of the responses given by
-all the centers provide the value estimation and the preference for one
-of the two pure strategies (hawk or dove). During the learning process
-individuals update the response that each one of these centers triggers.
-In the case of the preference for the strategies the numerical value
-resulting from the sum of the responses is translated to a probability
-of one of the two pure strategies through a logistic function. Figure
+use the actor-critic algorithm Sutton and Barto (2018). Individuals (
+\emph{learners} ) perceive the \emph{badge} of their counterpart as the
+environmental state in which they find themselves. They estimate the
+value of that state and a preference for playing hawk (dove) in that
+state. They update their estimation and preference after each
+interaction they have. Given that the badge size is a real number
+between 0 and 1, there are infinitely many environmental states. Thus,
+for their value estimation and preference computation, individual must
+use a form of generalization. We let them use as generalization a
+function approximation method based on radial basis functions. Each
+individual uses five \emph{centers} spread uniformly throughout the
+range of badge size variation (from 0 to 1). Each time an individual
+interacts with a conspecific the size of the badge of the interacting
+partner triggers are response from each one of the \emph{centers}. This
+response decreases in strength as the badge size is further from the
+\emph{center}. The decrease follows a radial basis function with
+gaussian shape. The sum of the responses given by all the centers
+provide the value estimation and the preference for one of the two pure
+strategies (hawk or dove). During the learning process individuals
+update the response that each one of these centers triggers. In the case
+of the preference for the strategies the numerical value resulting from
+the sum of the responses is translated to a probability of one of the
+two pure strategies through a logistic function. Figure
 \ref{fig:learning_cartoonRBF} shows a made up example of how the
@@
 
-\hypertarget{results-when-individuals-do-not-vary-in-their-quality}{%
 \subsection{Results when individuals do NOT vary in their
-quality}\label{results-when-individuals-do-not-vary-in-their-quality}}
+quality}\label{results-when-individuals-do-not-vary-in-their-quality}
 
@@
 population is also monomorphic for the reaction norm parameters
-(\(\alpha=0\) and \(\beta=0\)). With the values chosen the badge size
+(\(\alpha=0\) and \(\gamma=0\)). With the values chosen the badge size
 distribution is the same as the quality distribution. Thus, the four
@@
 
-\hypertarget{results-when-individuals-vary-in-their-quality}{%
 \subsection{Results when individuals vary in their
-quality}\label{results-when-individuals-vary-in-their-quality}}
+quality}\label{results-when-individuals-vary-in-their-quality}
 
@@
 
-\hypertarget{overal-effect-of-variance}{%
-\subsection{Overal effect of variance}\label{overal-effect-of-variance}}
+\subsection{Overal effect of variance}\label{overal-effect-of-variance}
 
@@
 
-\hypertarget{how-does-learning-look-like-in-the-hawk-dove-game}{%
 \subsection{How does learning look like in the Hawk-Dove
-game}\label{how-does-learning-look-like-in-the-hawk-dove-game}}
+game}\label{how-does-learning-look-like-in-the-hawk-dove-game}
 
@@
 
-\hypertarget{how-do-the-learning-dynamics-look-like}{%
 \subsection{How do the learning dynamics look
-like}\label{how-do-the-learning-dynamics-look-like}}
+like}\label{how-do-the-learning-dynamics-look-like}
 
@@
 
-\hypertarget{how-do-learners-fare-against-the-pure-strategies}{%
-\section{How do learners fare against the ``pure''
-strategies}\label{how-do-learners-fare-against-the-pure-strategies}}
+\section{\texorpdfstring{How do learners fare against the ``pure''
+strategies}{How do learners fare against the pure strategies}}\label{how-do-learners-fare-against-the-pure-strategies}
 
@@
 
-\hypertarget{whats-next}{%
-\section{What's next?}\label{whats-next}}
+\section{What's next?}\label{whats-next}
 
@@
 
+\section*{Cited literature}\label{cited-literature}
+\addcontentsline{toc}{section}{Cited literature}
+
+\hypertarget{refs}{}
+\hypertarget{ref-sutton_reinforcement_2018}{}
+Sutton, Richard S., and Andrew G. Barto. 2018. \emph{Reinforcement
+Learning: An Introduction}. Edited by Francis Bach. Second edition
+edition. Cambridge, MA: A Bradford Book.
+
 \end{document}

### 2019-07-02 QD_report.Rmd
FInalize QD report. Plot of evol simulations

@@
 header-includes: \usepackage{float}
-                \floatplacement{figure}{H}
+                #\floatplacement{figure}{H}
 ---
@@
 
-Classical Hawk-Dove game.
-Individuals have one fo two genetically determined phenotypic strategies.
+Individuals have one of two genetically determined phenotypic strategies.
 _Hawks_ are willing to start a conflict over resources, while _doves_ prefer to stand down in the hope to share the resource without an aggressive contest.
 Individuals interact randomly with each other over their lifetimes and collect resources depending on the strategy of theirs and their partners.
-In contrast to the classical _Hawk-dove_, in here when to _hawks_ interact with each other one of them wins the contest and gets $V-\frac{C}{2}$ as payoff; while the loser gets $-\frac{C}{2}$.
-Where $V$ is the value of the resource individuals are competing for, and $C$ is the cost of an agressive contest.
+In contrast to the classical _Hawk-dove_, in here when two _hawks_ interact with each other one of them wins the contest and gets $V-\frac{C}{2}$ as payoff; while the loser gets $-\frac{C}{2}$.
+Where $V$ is the value of the resource individuals are competing for, and $C$ is the cost of an aggressive contest.
 individuals share the cost, but only the winner gets the value of the resource.The rest of payoff matrix follows the standard game.
@@
 Individuals' reproductive success is proportional to the payoff accumulated
-throughout their lifes. Figure \ref{fig:HD_game} shows the evolutionary dynamics 
+throughout their life. Figure \ref{fig:HD_game} shows the evolutionary dynamics 
 of  the two genotypes, which fit the game theoretical prediction.
@@
 the hawk-dove game. Which means, they play hawk (dove) with a certain probability.
-Furthermore, the probability with which the play each strategy changes during their lifetime according to the payoff they recieve in previous interactions.
+Furthermore, the probability with which the play each strategy changes during their lifetime according to the payoff they receive in previous interactions.
 In other words, _learners_ learn to behave in the game.
@@
 Each individual uses five _centers_ spread uniformly throughout the range of badge size variation (from 0 to 1).
-Each time an individual interacts with a conspecific the size of the badge of the interacting partner triggers are response from each one of the _centers_.
+Each time an individual interacts with a con-specific the size of the badge of the interacting partner triggers a response from each one of the _centers_.
 This response decreases in strength as the badge size is further from the _center_.
-The decrease follows a radial basis function with gaussian shape.
+The decrease follows a radial basis function with Gaussian shape.
 The sum of the responses given by all the centers provide the value estimation and the preference for one of the two pure strategies (hawk or dove).
@@
 With the values chosen the badge size distribution is the same as the quality distribution.
-Thus, the four generations respresent just replicates of the learning process.
+Thus, the four generations represent just replicates of the learning process.
 The results show that under quality invariance, the learning algorithm develops similar phenotypic frequencies as the evolutionary dynamics of the pure strategies.
@@
 and standard deviation 0.1.
-Compared to the expectation from evolutionary dynamics of the standard game, the learning scenario with qualiy variation shows lower levels of hawks and therefore lower levels of escalated conflict.
+Compared to the expectation from evolutionary dynamics of the standard game, the learning scenario with quality variation shows lower levels of hawks and therefore lower levels of escalated conflict.
 At this point individuals do not respond adaptively to the quality of their partner, instead their preference for a given action is dependent on their own quality.
@@
 
-## Overal effect of variance
+## Overall effect of variance
 
 In figure \ref{fig:learning_all_var} we show how the effect of increasing 
-variance in quality further deacrases the amount of individuals in a population 
+variance in quality further decreases the amount of individuals in a population 
 expressing the hawk phenotype.
@@
 
-Figures \ref{fig:learning_var0.1} and \ref{fig:learning_all_var} only show how learners behave at a population level, but they do not show how inter-invidivual variation behind those frequencies.
-In figure \ref{fig:AC_HD} we show the result of the function aproximation mechanism arising at the end of a generation for a sample of the individuals of the populations.
-These results are show under three different levels of variation in quality.
+Figures \ref{fig:learning_var0.1} and \ref{fig:learning_all_var} only show how learners behave at a population level, but they do not show the inter-individual variation behind those frequencies.
+In figure \ref{fig:AC_HD} we show the result of the function approximation mechanism arising at the end of a generation for a sample of the individuals of the populations.
+These results are shown under three different levels of variation in quality.
 All three panes show that through their learning processes individuals split up in two groups, one which mostly behaves like a hawk and a second that mostly behaves like a dove.
 The group of "hawks" is mostly, but no exclusively, composed of individuals whose quality is on the upper half of the quality distribution.
-Concordantly, the group of "doves" is mostly composed of individuals whose qualiy is on the lower half of the quality distribution.
-The relation between quality and the prefered behaviour of individuals obviously only holds in scenarios with variance in quality.
-In the scenario with quality invariance, the split of individuals in two groups with a tenedency for each behaviour nevertheless remains.
+In accordance, the group of "doves" is mostly composed of individuals whose quality is on the lower half of the quality distribution.
+The relation between quality and the preferred behavior of individuals obviously only holds in scenarios with variance in quality.
+In the scenario with quality invariance, the split of individuals in two groups with a tendency for each behavior nevertheless remains.
 
-```{r fig6, echo=FALSE, fig.cap="\\label{fig:AC_HD}Within population variation in the learning parameters at the end of learning for three scenarios of variation in the quality of individuals.", out.width = '80%'}
+```{r fig6, echo=FALSE, fig.cap="\\label{fig:AC_HD}Within population variation in the learning parameters at the end of learning for three scenarios of variation in the quality of individuals.", out.width = '70%'}
 knitr::include_graphics("Simulations//QualStDv_//WeightsVarQualSt.png")
@@
 These dynamics show how individuals split in the two groups according to their experiences.
-Despite the robustness of the bimodal distribution along the learning dynamics, it is not uncommon that individual go from one group to the other.
+Despite the robustness of the bi-modal distribution along the learning dynamics, it is not uncommon that individual go from one group to the other.
 
@@
 show the results of competition among the three strategies.
-The population starts with a 1:1 ratio of pure hawks anf doves, and learners are intruced by mutation.
+The population starts with a 1:1 ratio of pure hawks and doves, and learners are introduced by mutation.
 Learners manage to increase in frequency.
-They, however, do not manage to outcompete the pure streategies.
-In the end, there is a polymorphic equilibrium with the three strategies coexisting, with learners being the most frequen of the threee.
+They, however, do not manage to outcompete the pure strategies.
+In the end, there is a polymorphic equilibrium with the three strategies coexisting, with learners being the most frequent of the three.
 

### 2019-07-02 QD_report.tex
FInalize QD report. Plot of evol simulations

--- 2019-06-26 QD_report.tex
@@
 
-\floatplacement{figure}{H}
-
 \begin{document}
@@
 
-Classical Hawk-Dove game. Individuals have one fo two genetically
-determined phenotypic strategies. \emph{Hawks} are willing to start a
-conflict over resources, while \emph{doves} prefer to stand down in the
-hope to share the resource without an aggressive contest. Individuals
-interact randomly with each other over their lifetimes and collect
-resources depending on the strategy of theirs and their partners. In
-contrast to the classical \emph{Hawk-dove}, in here when to \emph{hawks}
-interact with each other one of them wins the contest and gets
-\(V-\frac{C}{2}\) as payoff; while the loser gets \(-\frac{C}{2}\).
-Where \(V\) is the value of the resource individuals are competing for,
-and \(C\) is the cost of an agressive contest. individuals share the
-cost, but only the winner gets the value of the resource.The rest of
-payoff matrix follows the standard game.
+Individuals have one of two genetically determined phenotypic
+strategies. \emph{Hawks} are willing to start a conflict over resources,
+while \emph{doves} prefer to stand down in the hope to share the
+resource without an aggressive contest. Individuals interact randomly
+with each other over their lifetimes and collect resources depending on
+the strategy of theirs and their partners. In contrast to the classical
+\emph{Hawk-dove}, in here when two \emph{hawks} interact with each other
+one of them wins the contest and gets \(V-\frac{C}{2}\) as payoff; while
+the loser gets \(-\frac{C}{2}\). Where \(V\) is the value of the
+resource individuals are competing for, and \(C\) is the cost of an
+aggressive contest. individuals share the cost, but only the winner gets
+the value of the resource.The rest of payoff matrix follows the standard
+game.
 
@@
 who wins the contest. Individuals' reproductive success is proportional
-to the payoff accumulated throughout their lifes. Figure
+to the payoff accumulated throughout their life. Figure
 \ref{fig:HD_game} shows the evolutionary dynamics of the two genotypes,
@@
 which the play each strategy changes during their lifetime according to
-the payoff they recieve in previous interactions. In other words,
+the payoff they receive in previous interactions. In other words,
 \emph{learners} learn to behave in the game. To implement the learning
 process, we use the reinforcement learning formalism. Specifically, we
-use the actor-critic algorithm. Individuals ( \emph{learners} ) perceive
-the \emph{badge} of their counterpart as the environmental state in
-which they find themselves. They estimate the value of that state and a
-preference for playing hawk (dove) in that state. They update their
-estimation and preference after each interaction they have. Given that
-the badge size is a real number between 0 and 1, there are infinitely
-many environmental states. Thus, for their value estimation and
-preference computation, individual must use a form of generalization. We
-let them use as generalization a function approximation method based on
-radial basis functions. Each individual uses five \emph{centers} spread
-uniformly throughout the range of badge size variation (from 0 to 1).
-Each time an individual interacts with a conspecific the size of the
-badge of the interacting partner triggers are response from each one of
-the \emph{centers}. This response decreases in strength as the badge
-size is further from the \emph{center}. The decrease follows a radial
-basis function with gaussian shape. The sum of the responses given by
-all the centers provide the value estimation and the preference for one
-of the two pure strategies (hawk or dove). During the learning process
-individuals update the response that each one of these centers triggers.
-In the case of the preference for the strategies the numerical value
-resulting from the sum of the responses is translated to a probability
-of one of the two pure strategies through a logistic function. Figure
-\ref{fig:learning_cartoonRBF} shows a made up example of how the
-function approximation works for the actor (top panel - probability of
-choosing dove) and the critic (bottom panel - expected payoff from
-facing an individual with a given badge size).
+use the actor-critic algorithm (Sutton and Barto 2018). Individuals (
+\emph{learners} ) perceive the \emph{badge} of their counterpart as the
+environmental state in which they find themselves. They estimate the
+value of that state and a preference for playing hawk (dove) in that
+state. They update their estimation and preference after each
+interaction they have. Given that the badge size is a real number
+between 0 and 1, there are infinitely many environmental states. Thus,
+for their value estimation and preference computation, individual must
+use a form of generalization. We let them use as generalization a
+function approximation method based on radial basis functions. Each
+individual uses five \emph{centers} spread uniformly throughout the
+range of badge size variation (from 0 to 1). Each time an individual
+interacts with a con-specific the size of the badge of the interacting
+partner triggers a response from each one of the \emph{centers}. This
+response decreases in strength as the badge size is further from the
+\emph{center}. The decrease follows a radial basis function with
+Gaussian shape. The sum of the responses given by all the centers
+provide the value estimation and the preference for one of the two pure
+strategies (hawk or dove). During the learning process individuals
+update the \emph{weight} of the response that each one of these centers
+triggers. In the case of the preference for the strategies the numerical
+value resulting from the sum of the responses is translated to a
+probability of one of the two pure strategies through a logistic
+function. Figure \ref{fig:learning_cartoonRBF} shows a made up example
+of how the function approximation works for the actor (top panel -
+probability of choosing dove) and the critic (bottom panel - expected
+payoff from facing an individual with a given badge size).
 
@@
 population is also monomorphic for the reaction norm parameters
-(\(\alpha=0\) and \(\beta=0\)). With the values chosen the badge size
+(\(\alpha=0\) and \(\gamma=0\)). With the values chosen the badge size
 distribution is the same as the quality distribution. Thus, the four
-generations respresent just replicates of the learning process. The
+generations represent just replicates of the learning process. The
 results show that under quality invariance, the learning algorithm
@@
 
+In figure \ref{fig:learning_var0.1} we show results of phenotypic and
+interaction types frequencies from simulations in which individuals vary
+in their quality. The value of quality is drawn from a truncated normal
+distribution with mean 0.5 and standard deviation 0.1. Compared to the
+expectation from evolutionary dynamics of the standard game, the
+learning scenario with quality variation shows lower levels of hawks and
+therefore lower levels of escalated conflict. At this point individuals
+do not respond adaptively to the quality of their partner, instead their
+preference for a given action is dependent on their own quality.
+
 \begin{figure}
@@
 
-\hypertarget{overal-effect-of-variance}{%
-\subsection{Overal effect of variance}\label{overal-effect-of-variance}}
+\hypertarget{overall-effect-of-variance}{%
+\subsection{Overall effect of
+variance}\label{overall-effect-of-variance}}
+
+In figure \ref{fig:learning_all_var} we show how the effect of
+increasing variance in quality further decreases the amount of
+individuals in a population expressing the hawk phenotype.
 
@@
 \subsection{How does learning look like in the Hawk-Dove
-game}\label{how-does-learning-look-like-in-the-hawk-dove-game}}
+game?}\label{how-does-learning-look-like-in-the-hawk-dove-game}}
+
+Figures \ref{fig:learning_var0.1} and \ref{fig:learning_all_var} only
+show how learners behave at a population level, but they do not show the
+inter-individual variation behind those frequencies. In figure
+\ref{fig:AC_HD} we show the result of the function approximation
+mechanism arising at the end of a generation for a sample of the
+individuals of the populations. These results are shown under three
+different levels of variation in quality. All three panes show that
+through their learning processes individuals split up in two groups, one
+which mostly behaves like a hawk and a second that mostly behaves like a
+dove. The group of ``hawks'' is mostly, but no exclusively, composed of
+individuals whose quality is on the upper half of the quality
+distribution. In accordance, the group of ``doves'' is mostly composed
+of individuals whose quality is on the lower half of the quality
+distribution. The relation between quality and the preferred behavior of
+individuals obviously only holds in scenarios with variance in quality.
+In the scenario with quality invariance, the split of individuals in two
+groups with a tendency for each behavior nevertheless remains.
 
 \begin{figure}
-\includegraphics[width=0.8\linewidth]{Simulations//QualStDv_//WeightsVarQualSt} \caption{\label{fig:AC_HD}Within population variation in the learning parameters at the end of learning for three scenarios of variation in the quality of individuals.}\label{fig:fig6}
+\includegraphics[width=0.7\linewidth]{Simulations//QualStDv_//WeightsVarQualSt} \caption{\label{fig:AC_HD}Within population variation in the learning parameters at the end of learning for three scenarios of variation in the quality of individuals.}\label{fig:fig6}
 \end{figure}
@@
 like}\label{how-do-the-learning-dynamics-look-like}}
+
+In figure \ref{fig:learn_dyn} we show the learning trajectory, for the
+actor and the critic, of a sample of individuals. These dynamics show
+how individuals split in the two groups according to their experiences.
+Despite the robustness of the bi-modal distribution along the learning
+dynamics, it is not uncommon that individual go from one group to the
+other.
 
@@
 strategies}\label{how-do-learners-fare-against-the-pure-strategies}}
+
+Finally, in a new set of simulations we explore the question of how
+competitive learners are in competition with the pure strategies. Figure
+\ref{fig:types_comp} show the results of competition among the three
+strategies. The population starts with a 1:1 ratio of pure hawks and
+doves, and learners are introduced by mutation. Learners manage to
+increase in frequency. They, however, do not manage to outcompete the
+pure strategies. In the end, there is a polymorphic equilibrium with the
+three strategies coexisting, with learners being the most frequent of
+the three.
 
@@
 
+\hypertarget{cited-literature}{%
+\section*{Cited literature}\label{cited-literature}}
+\addcontentsline{toc}{section}{Cited literature}
+
+\hypertarget{refs}{}
+\leavevmode\hypertarget{ref-sutton_reinforcement_2018}{}%
+Sutton, Richard S., and Andrew G. Barto. 2018. \emph{Reinforcement
+Learning: An Introduction}. Edited by Francis Bach. Second edition
+edition. Cambridge, MA: A Bradford Book.
+
 \end{document}

### 2019-07-02 dynamics.R

@@ -11,15 +11,15 @@ source(here("AccFunc.R"))
 
 # Scenario to be plotted - corresponds to folders where simulations are stored
 
-scenario<-"QualStDv_"
+scenario<-"baselineFit"
 
 
 # Load files -------------------------------------------------------------------
 
-(listTest<-list.files(here("Simulations",scenario)))
+(listTest<-list.files(here("Simulations",paste0(scenario,"_"))))
 (sdList<-grep("evol",listTest,value=TRUE))
 
-evol<-fread(here("Simulations",scenario,sdList[2]))
+evol<-fread(here("Simulations",paste0(scenario,"_"),sdList[1]))
 
 # Extract means and IQR for the dynamic variables ------------------------------
 
@@ -66,7 +66,8 @@ evolStats<-evol[,.(m.freqGenHawk=mean(freqGenHawks),
 
 # Plot mean and IQRs of the genotypes and phenotypes ----------------------------
 
-png(here("Simulations",scenario,"hawkDoveLearn_0.1.png"),height = 800,width = 800)
+# png(here("Simulations",scenario,"hawkDoveLearn_0.1.png"),
+# height = 800,width = 800)
 
 par(plt=posPlot(numploty = 2,idploty = 2),xaxt="s",las=2)
 plot(x=c(0,max(evolStats$time)),y=c(0.5,0.5),type="l",lwd=2,col="grey",
@@ -152,7 +153,6 @@ with(evolStats,{
 
 # Plot the weights of the actor ------------------------------------------------
 
-
 nCenters<-5
 interv<-1/nCenters
 centers<-interv*0.5+interv*seq(0,nCenters-1)
@@ -201,6 +201,32 @@ with(evolStats,{
 })
 legend("topleft",legend = c(expression(alpha),expression(beta)),
        col=colGenesLin,lwd=2,bty = "n")
+
+
+hist(evol[time==9600,meanBeta])
+hist(evol[time==9600,meanAlpha])
+plot(meanAlpha~meanBeta,data=evol[time==9600])
+linReg<-lm(meanAlpha~meanBeta,data=evol[time==9600])
+linReg$coefficients[2]
+
+
+# Plot trajectory of the trait evolution for all runs separetely ---------------
+
+traitsTrajs<-dcast(evol,time~seed,value.var = c("meanAlpha","meanBeta"))
+
[…]

### 2019-07-02 popSample_1.R

@@ -11,7 +11,7 @@ source(here("AccFunc.R"))
 
 # Scenario to be plotted - corresponds to folders where simulations are stored
 
-scenario<-"mutType"
+scenario<-"baselineFit"
 
 
 # Load files -------------------------------------------------------------------
@@ -19,7 +19,7 @@ scenario<-"mutType"
 (listTest<-list.files(here("Simulations",paste0(scenario,"_"))))
 (List<-grep("indLearn",listTest,value=TRUE))
 
-pop<-fread(here("Simulations",paste0(scenario,"_"),List[4]))
+pop<-fread(here("Simulations",paste0(scenario,"_"),List[1]))
 
 pop<-pop[,.SD[nInteract==max(nInteract)],by=.(seed,time,indId)]
 
@@ -39,10 +39,10 @@ popStats<-pop[, as.list(unlist(lapply(.SD,
 
 # Plot variation of the weights ------------------------------------------------
 
-png(here("Simulations",paste0(scenario,"_"),"weightsVarQualSt.png"),
-    width = 1000,height = 800)
+# png(here("Simulations",paste0(scenario,"_"),"weightsVarQualSt.png"),
+#     width = 1000,height = 800)
 
-gener<-pop[,unique(time)][10]
+gener<-pop[,unique(time)][25]
 nCenters<-5
 interv<-1/nCenters
 centers<-interv*0.5+interv*seq(0,nCenters-1)
@@ -75,12 +75,12 @@ for(ind in pop[time==gener&genotype==2,unique(indId)]){
                                   Quality],colorbreaksQual)],lwd=0.5)
 }
 # Critic 
-weightsCrit<-as.double(evolStats[time==min(time),.SD,
+weightsCrit<-as.double(evolStats[time==gener,.SD,
                                  .SDcols=grep("m.weightCrit",
                                               names(evolStats),value = TRUE)])
 par(plt=posPlot(numploty = 2,idploty = 1),new=TRUE)
 plot(totRBF(rangx,centers,0.01,weightsCrit)~rangx,type='l',col=1,xaxt="s",
-     yaxt="s",xlab="Badge",ylab="value",ylim=c(-0.20,1),lwd=3)
+     yaxt="s",xlab="Badge",ylab="value",ylim=c(-1,1.5),lwd=3)
 lines(totRBF(rangx,centers,0.01,
              as.double(popStats[time==gener,.SD,
[…]

### 2019-07-26 QD_report.Rmd
Various edits in report and graph, trying (unsuccesfully) to include results in the Arollo talk

@@
 ---
+
+```{r loadData, echo=FALSE,include=FALSE}
+# Required libraries -----------------------------------------------------------
+
+library(here)
+here()
+source(here("AccFunc.R"))
+# Scenario to be plotted - corresponds to folders where simulations are stored
+scenario<-"mutType"
+(listTest<-list.files(here("Simulations",paste0(scenario,"_"))))
+(evolList<-grep(".txt",grep("evol",listTest,value=TRUE),value=TRUE))
+(indList<-grep("ind",listTest,value=TRUE))
+
+fileId<-1
+evol<-fread(here("Simulations",paste0(scenario,"_"),evolList[fileId]))
+pop<-fread(here("Simulations",paste0(scenario,"_"),indList[fileId]))
+evolStats<-evol[,.(m.freqGenHawk=mean(freqGenHawks),
+                   upIQR.freqGenHawk=fivenum(freqGenHawks)[4],
+                   lowIQR.freqGenHawk=fivenum(freqGenHawks)[2],
+                   m.freqGenDove=mean(freqGenDove),
+                   upIQR.freqGenDove=fivenum(freqGenDove)[4],
+                   lowIQR.freqGenDove=fivenum(freqGenDove)[2],
+                   m.freqEval=mean(freqGenEval),
+                   upIQR.freqGenEval=fivenum(freqGenEval)[4],
+                   lowIQR.freqGenEval=fivenum(freqGenEval)[2], 
+                   m.freqFenHawk=mean(freqFenHawks),
+                   upIQR.freqFenHawk=fivenum(freqFenHawks)[4],
+                   lowIQR.freqFenHawk=fivenum(freqFenHawks)[2],
+                   m.freqFenDove=mean(freqFenDoves),
+                   upIQR.freqFenDove=fivenum(freqFenDoves)[4],
+                   lowIQR.freqFenDove=fivenum(freqFenDoves)[2],
+                   m.meanAlpha=mean(meanAlpha),
+                   upIQR.alpha=fivenum(meanAlpha)[4],
+                   lowIQR.alpha=fivenum(meanAlpha)[2],
+                   m.meanBeta=mean(meanBeta),
+                   upIQR.beta=fivenum(meanBeta)[4],
+                   lowIQR.beta=fivenum(meanBeta)[2],
+                   m.freqHH = mean(freqHH),
+                   m.freqHD = mean(freqHD),
+                   m.freqDD = mean(freqDD),
+                   upIQR.freqHH = fivenum(freqHH)[4],
+                   upIQR.freqHD = fivenum(freqHD)[4],
+                   upIQR.freqDD = fivenum(freqDD)[4],
+                   lowIQR.freqHH = fivenum(freqHH)[4],
+                   lowIQR.freqHD = fivenum(freqHD)[2],
+                   lowIQR.freqDD = fivenum(freqDD)[2],
+                   m.weightAct_0=mean(WeightAct_0),
+                   m.weightAct_1=mean(WeightAct_1),
+                   m.weightAct_2=mean(WeightAct_2),
+                   m.weightAct_3=mean(WeightAct_3),
+                   m.weightAct_4=mean(WeightAct_4),
+                   m.weightCrit_0=mean(WeightCrit_0),
+                   m.weightCrit_1=mean(WeightCrit_1),
+                   m.weightCrit_2=mean(WeightCrit_2),
+                   m.weightCrit_3=mean(WeightCrit_3),
+                   m.weightCrit_4=mean(WeightCrit_4)),by=time]
+# Extract means and IQR for the dynamic variables ------------------------------
+
+popStats<-pop[, as.list(unlist(lapply(.SD, 
+                                      function(x) list(m = mean(x),
+                                                       upIQR = fivenum(x)[4],
+                                                       downIQR = fivenum(x)[2]
+                                      )))),
+              by = time, 
+              .SDcols=c("Quality","alpha","beta","Badge","nInteract","WeightAct_0",
+                        "WeightAct_1","WeightAct_2","WeightAct_3","WeightAct_4",
+                        "WeightCrit_0","WeightCrit_1","WeightCrit_2",
+                        "WeightCrit_3","WeightCrit_4")]
+
+# knitr::include_graphics("Simulations/mutType_/basicHawkDove.png")
+```
 

### 2019-07-26 QD_report.tex
Various edits in report and graph, trying (unsuccesfully) to include results in the Arollo talk

@@
 
-\hypertarget{overall-effect-of-variance}{%
 \subsection{Overall effect of
-variance}\label{overall-effect-of-variance}}
+variance}\label{overall-effect-of-variance}
 
@@
 
-\hypertarget{how-does-learning-look-like-in-the-hawk-dove-game}{%
 \subsection{How does learning look like in the Hawk-Dove
-game?}\label{how-does-learning-look-like-in-the-hawk-dove-game}}
+game?}\label{how-does-learning-look-like-in-the-hawk-dove-game}
 
@@
 
-\hypertarget{cited-literature}{%
-\section*{Cited literature}\label{cited-literature}}
+\section*{Cited literature}\label{cited-literature}
 \addcontentsline{toc}{section}{Cited literature}
@@
 \hypertarget{refs}{}
-\leavevmode\hypertarget{ref-sutton_reinforcement_2018}{}%
+\hypertarget{ref-sutton_reinforcement_2018}{}
 Sutton, Richard S., and Andrew G. Barto. 2018. \emph{Reinforcement

### 2019-07-26 AccFunc.R

@@ -79,3 +79,5 @@ filesScenar<-function(filename,scenario){
   tmp[,eval(scenario):=parVal]
   return(tmp)
 }
+
+

### 2019-07-26 CartoonRBF.R

@@ -2,7 +2,7 @@ library(data.table)
 library(here)
 here()
 
-source("C:/Users/a.quinones/Dropbox/R_files/posPlots.R")
+source("../R_files/posPlots.R")
 
 
 png(here("cartoonRBF.png"),width=800,height = 800)

### 2019-07-26 learnDynamics.R

@@ -11,7 +11,7 @@ source(here("AccFunc.R"))
 
 # Scenario to be plotted - corresponds to folders where simulations are stored
 
-scenario<-"learHonest_/alphaAct"
+scenario<-"mutType"
 
 
 # Load files -------------------------------------------------------------------
@@ -19,7 +19,7 @@ scenario<-"learHonest_/alphaAct"
 (listTest<-list.files(here("Simulations",paste0(scenario,"_"))))
 (List<-grep("ind",listTest,value=TRUE))
 
-indLearn<-fread(here("Simulations",paste0(scenario,"_"),List[3]))
+indLearn<-fread(here("Simulations",paste0(scenario,"_"),List[1]))
 
 
 

### 2019-08-15 QD_report.Rmd
New graph in the QD report

@@
 
-# How do learners fare against the "pure" strategies
+# How do learners fare against the "pure" strategies?
 
@@
 
+# How do learners behave when signals are honest?
+
+```{r fig9, echo=FALSE,fig.cap="\\label{fig:honest_learn}Actor and critic developed when signals are honest.
+Left panel: honest signal.
+Right panel: learners behavioral response ",out.width='100%',fig.show='hold'}
+knitr::include_graphics("Simulations//learHonest_//QualStDv_//weightsVarQualSt.png")
+
+```
+
 # What's next?

### 2019-08-15 QD_report.tex
New graph in the QD report

--- 2019-07-02 QD_report.tex
@@
 \section{How do learners fare against the ``pure''
-strategies}\label{how-do-learners-fare-against-the-pure-strategies}}
+strategies?}\label{how-do-learners-fare-against-the-pure-strategies}}
 
@@
 \includegraphics[width=0.8\linewidth]{Simulations//mutType_//evolDynTypes} \caption{\label{fig:types_comp}Evolutionary dynamics of the learners competing against pure strategies.}\label{fig:fig8}
+\end{figure}
+
+\hypertarget{how-do-learners-behave-when-signals-are-honest}{%
+\section{How do learners behave when signals are
+honest?}\label{how-do-learners-behave-when-signals-are-honest}}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{Simulations//learHonest_//QualStDv_//weightsVarQualSt} \caption{\label{fig:honest_learn}Actor and critic developed when signals are honest.
+Left panel: honest signal.
+Right panel: learners behavioral response }\label{fig:fig9}
 \end{figure}

### 2019-08-15 learnDynamics.R

@@ -11,7 +11,7 @@ source(here("AccFunc.R"))
 
 # Scenario to be plotted - corresponds to folders where simulations are stored
 
-scenario<-"mutType"
+scenario<-"learHonest_/QualStDv"
 
 
 # Load files -------------------------------------------------------------------

### 2019-09-05 QD_presentation.Rmd
starting the presentation for the group meeting

--- 2019-07-26 QD_report.Rmd
@@
 output: 
-  pdf_document: 
+    beamer_presentation: 
       fig_caption: yes
       keep_tex: yes
-header-includes: \usepackage{float}
+header-includes: 
+    \usepackage{float}
                 #\floatplacement{figure}{H}
@@
 
-# The Hawk-Dove game
+## The Hawk-Dove game
 
@@
 _Hawks_ are willing to start a conflict over resources, while _doves_ prefer to stand down in the hope to share the resource without an aggressive contest.
-Individuals interact randomly with each other over their lifetimes and collect resources depending on the strategy of theirs and their partners.
+
+```{r,echo=FALSE,fig.align="center",out.width='50%'}
+knitr::include_graphics("Images//hawk_dove.jpeg")
+```
+
+## The hawk-dove game
+
+\begin{align*}
+\end{align*}
+
+```{r, echo=FALSE, out.width = '80%'}
+knitr::include_graphics("Images/Hawk_dove_plot.png")
+```
+
+## The hawk-dove game
+
+```{r, echo=FALSE, fig.cap="\\label{fig:HD_game}Hawk-dove game. Dashed line is the  game theoretical prediction for frequency of hawks.", out.width = '80%'}
+knitr::include_graphics("Simulations/mutType_/basicHawkDove.png")
+```
+
+<!-- Individuals interact randomly with each other over their lifetimes and collect resources depending on the strategy of theirs and their partners.
 In contrast to the classical _Hawk-dove_, in here when two _hawks_ interact with each other one of them wins the contest and gets $V-\frac{C}{2}$ as payoff; while the loser gets $-\frac{C}{2}$.
@@
 individuals share the cost, but only the winner gets the value of the resource.The rest of payoff matrix follows the standard game.
+-->
 
-Individuals may or may not vary in their quality.
+<!-- Individuals may or may not vary in their quality.
 Quality is a numerical value from 0 to 1, drawn from a truncated normal distribution when each individual is born.
@@
 Differences in quality between two individuals partly determine who wins in a conflict when both players play hawk.
-The probability of individual A wining a conflict over individual B is given by
+The probability of individual A wining a conflict over individual B is given by -->
 
-\begin{equation}
-p(A)=\frac{1}{1+e^{-\beta(Q_A-Q_B)}},
-\end{equation}
+<!-- \begin{equation} -->
+<!-- p(A)=\frac{1}{1+e^{-\beta(Q_A-Q_B)}}, -->
+<!-- \end{equation} -->
 
-where $Q_A$ and $Q_b$ are the qualities of individuals A and B respectively, and $\beta$ is how important quality is in determining who wins the contest. 
-Individuals' reproductive success is proportional to the payoff accumulated
-throughout their life. Figure \ref{fig:HD_game} shows the evolutionary dynamics 
-of  the two genotypes, which fit the game theoretical prediction.
+<!-- where $Q_A$ and $Q_b$ are the qualities of individuals A and B respectively, and $\beta$ is how important quality is in determining who wins the contest.  -->
+<!-- Individuals' reproductive success is proportional to the payoff accumulated -->
+<!-- throughout their life. Figure \ref{fig:HD_game} shows the evolutionary dynamics  -->
+<!-- of  the two genotypes, which fit the game theoretical prediction. -->
 
-```{r fig1, echo=FALSE, fig.cap="\\label{fig:HD_game}Hawk-dove game. Dashed line is the  game theoretical prediction for frequency of hawks.", out.width = '80%'}
-knitr::include_graphics("Simulations/mutType_/basicHawkDove.png")
-```
+<!-- # The effect of learning  -->
 
-# The effect of learning
+<!-- We let individuals have a _badge of status_ that potentially signals their  -->
+<!-- quality. To allow for this signalling, we let the size of the _badge of status_  -->
+<!-- of an individual be dependent on her quality through a reaction norm. we assume the size of the _badge_ is a value from 0 to 1. This value is given by -->
+<!-- \begin{equation} -->
+<!-- s_i = \frac{1}{1+e^{-(\alpha_i+\gamma_iQ_i)}} -->
+<!-- \label{eq:react_norm} -->
+<!-- \end{equation} -->
+<!-- The parameters $\alpha$ and $\gamma$ in equation \ref{eq:react_norm} determine  -->
+<!-- the shape of the reaction norm, and hence, how honest the _badge_ is in  -->
+<!-- signalling the quality of the individual.Eventually, the aim is to allow the  -->
+<!-- signalling system to evolve, by letting the parameters of the reaction norm be inherited from parent to offspring and be change by a mutational process. -->
 
-We let individuals have a _badge of status_ that potentially signals their 
-quality. To allow for this signalling, we let the size of the _badge of status_ 
-of an individual be dependent on her quality through a reaction norm. we assume the size of the _badge_ is a value from 0 to 1. This value is given by
-\begin{equation}
-s_i = \frac{1}{1+e^{-(\alpha_i+\gamma_iQ_i)}}
-\label{eq:react_norm}
-\end{equation}
-The parameters $\alpha$ and $\gamma$ in equation \ref{eq:react_norm} determine 
-the shape of the reaction norm, and hence, how honest the _badge_ is in 
-signalling the quality of the individual.Eventually, the aim is to allow the 
-signalling system to evolve, by letting the parameters of the reaction norm be inherited from parent to offspring and be change by a mutational process.
+<!-- ## How does learning work? -->
 
-## How does learning work?
-
-To complete the receiver side of the signalling system, we introduce a new 
-strategy in the game that we call _learner_. _Learners_ use a mixed strategy in 
-the hawk-dove game. Which means, they play hawk (dove) with a certain probability.
-Furthermore, the probability with which the play each strategy changes during their lifetime according to the payoff they receive in previous interactions.
+<!-- To complete the receiver side of the signalling system, we introduce a new  -->
+<!-- strategy in the game that we call _learner_. _Learners_ use a mixed strategy in  -->
+<!-- the hawk-dove game. Which means, they play hawk (dove) with a certain probability. -->
+<!-- Furthermore, the probability with which the play each strategy changes during their lifetime according to the payoff they receive in previous interactions.
 In other words, _learners_ learn to behave in the game.
-To implement the learning
-process, we use the reinforcement learning formalism.
+To implement the learning -->
+<!-- process, we use the reinforcement learning formalism.
 Specifically, we use the actor-critic algorithm [@sutton_reinforcement_2018].
@@
 Thus, for their value estimation and preference computation, individual must use a form of generalization.
-We let them
-use as generalization a function approximation method based on radial basis functions.
+We let them -->
+<!-- use as generalization a function approximation method based on radial basis functions.
 Each individual uses five _centers_ spread uniformly throughout the range of badge size variation (from 0 to 1).
@@
 Figure \ref{fig:learning_cartoonRBF} shows a made up example of how the function approximation works for the actor (top panel - probability of choosing dove) and the critic (bottom panel - expected payoff from facing an individual with a given badge size).
+-->
 
-```{r fig2, echo=FALSE, fig.cap="\\label{fig:learning_cartoonRBF}Function approximation for the actor and the critic.
+<!-- ```{r fig2, echo=FALSE, fig.cap="\\label{fig:learning_cartoonRBF}Function approximation for the actor and the critic.
 Using random values for the responses of each center.
-Circles show the location of the centers and the maximum response they trigger, while the lines show the total response along the badge size continium.", out.width = '80%'}
-knitr::include_graphics("cartoonRBF.png")
-```
+Circles show the location of the centers and the maximum response they trigger, while the lines show the total response along the badge size continium.", out.width = '80%'} -->
+<!-- knitr::include_graphics("cartoonRBF.png") -->
+<!-- ``` -->
 
-## Results when individuals do NOT vary in their quality
+<!-- ## Results when individuals do NOT vary in their quality -->
 
-In Figure \ref{fig:learning_invar} we show the phenotypic frequencies and interaction-type frequencies in 4 generations of a population composed of only _learners_ that do not vary in their quality.
+<!-- In Figure \ref{fig:learning_invar} we show the phenotypic frequencies and interaction-type frequencies in 4 generations of a population composed of only _learners_ that do not vary in their quality.
 The population is also monomorphic for the reaction norm parameters ($\alpha=0$ and $\gamma=0$).
@@
 The results show that under quality invariance, the learning algorithm develops similar phenotypic frequencies as the evolutionary dynamics of the pure strategies.
+-->
 
-```{r fig3, echo=FALSE, fig.cap="\\label{fig:learning_invar}Learning equilibrium when individuals do not vary in quality.
-Dashed line is the  game theoretical prediction for frequency of hawks.", out.width = '80%'}
-knitr::include_graphics("Simulations//QualStDv_//hawkDoveLearn.png")
-```
+<!-- ```{r fig3, echo=FALSE, fig.cap="\\label{fig:learning_invar}Learning equilibrium when individuals do not vary in quality.
+Dashed line is the  game theoretical prediction for frequency of hawks.", out.width = '80%'} -->
+<!-- knitr::include_graphics("Simulations//QualStDv_//hawkDoveLearn.png") -->
+<!-- ``` -->
 
-## Results when individuals vary in their quality
+<!-- ## Results when individuals vary in their quality -->
 
-In figure \ref{fig:learning_var0.1} we show results of phenotypic and interaction types frequencies from simulations in which individuals vary in their quality. 
-The value of quality is drawn from a truncated normal distribution with mean 0.5 
-and standard deviation 0.1.
+<!-- In figure \ref{fig:learning_var0.1} we show results of phenotypic and interaction types frequencies from simulations in which individuals vary in their quality.  -->
+<!-- The value of quality is drawn from a truncated normal distribution with mean 0.5  -->
+<!-- and standard deviation 0.1.
 Compared to the expectation from evolutionary dynamics of the standard game, the learning scenario with quality variation shows lower levels of hawks and therefore lower levels of escalated conflict.
 At this point individuals do not respond adaptively to the quality of their partner, instead their preference for a given action is dependent on their own quality.
+-->
 
-```{r fig4, echo=FALSE, fig.cap="\\label{fig:learning_var0.1}Learning equilibrium when individuals vary in quality.
-Dashed lines are the  game theoretical predictions in the standard game.", out.width = '80%'}
-knitr::include_graphics("Simulations//QualStDv_//hawkDoveLearn_0.1.png")
-```
+<!-- ```{r fig4, echo=FALSE, fig.cap="\\label{fig:learning_var0.1}Learning equilibrium when individuals vary in quality.
+Dashed lines are the  game theoretical predictions in the standard game.", out.width = '80%'} -->
+<!-- knitr::include_graphics("Simulations//QualStDv_//hawkDoveLearn_0.1.png") -->
+<!-- ``` -->
 
-## Overall effect of variance
+<!-- ## Overall effect of variance -->
 
-In figure \ref{fig:learning_all_var} we show how the effect of increasing 
-variance in quality further decreases the amount of individuals in a population 
-expressing the hawk phenotype.
+<!-- In figure \ref{fig:learning_all_var} we show how the effect of increasing  -->
+<!-- variance in quality further decreases the amount of individuals in a population  -->
+<!-- expressing the hawk phenotype. -->
 
-```{r fig5, echo=FALSE, fig.cap="\\label{fig:learning_all_var}Effect of variation in quality for the learning equilibrium.", out.width = '80%'}
-knitr::include_graphics("Simulations//QualStDv_//effectQualVariance.png")
-```
+<!-- ```{r fig5, echo=FALSE, fig.cap="\\label{fig:learning_all_var}Effect of variation in quality for the learning equilibrium.", out.width = '80%'} -->
+<!-- knitr::include_graphics("Simulations//QualStDv_//effectQualVariance.png") -->
+<!-- ``` -->
 
-## How does learning look like in the Hawk-Dove game?
+<!-- ## How does learning look like in the Hawk-Dove game? -->
 
-Figures \ref{fig:learning_var0.1} and \ref{fig:learning_all_var} only show how learners behave at a population level, but they do not show the inter-individual variation behind those frequencies.
+<!-- Figures \ref{fig:learning_var0.1} and \ref{fig:learning_all_var} only show how learners behave at a population level, but they do not show the inter-individual variation behind those frequencies.
 In figure \ref{fig:AC_HD} we show the result of the function approximation mechanism arising at the end of a generation for a sample of the individuals of the populations.
@@
 In the scenario with quality invariance, the split of individuals in two groups with a tendency for each behavior nevertheless remains.
+-->
 
-```{r fig6, echo=FALSE, fig.cap="\\label{fig:AC_HD}Within population variation in the learning parameters at the end of learning for three scenarios of variation in the quality of individuals.", out.width = '70%'}
-knitr::include_graphics("Simulations//QualStDv_//WeightsVarQualSt.png")
-```
+<!-- ```{r fig6, echo=FALSE, fig.cap="\\label{fig:AC_HD}Within population variation in the learning parameters at the end of learning for three scenarios of variation in the quality of individuals.", out.width = '70%'} -->
+<!-- knitr::include_graphics("Simulations//QualStDv_//WeightsVarQualSt.png") -->
+<!-- ``` -->
 
-## How do the learning dynamics look like
+<!-- ## How do the learning dynamics look like -->
 
-In figure \ref{fig:learn_dyn} we show the learning trajectory, for the actor and the critic, of a sample of individuals.
+<!-- In figure \ref{fig:learn_dyn} we show the learning trajectory, for the actor and the critic, of a sample of individuals.
 These dynamics show how individuals split in the two groups according to their experiences.
 Despite the robustness of the bi-modal distribution along the learning dynamics, it is not uncommon that individual go from one group to the other.
+-->
 
-```{r fig7, echo=FALSE, fig.cap="\\label{fig:learn_dyn}Learning dynamics of the individuals in a population.", out.width = '80%'}
-knitr::include_graphics("Simulations//mutType_//learnDyn200.png")
-```
+<!-- ```{r fig7, echo=FALSE, fig.cap="\\label{fig:learn_dyn}Learning dynamics of the individuals in a population.", out.width = '80%'} -->
+<!-- knitr::include_graphics("Simulations//mutType_//learnDyn200.png") -->
+<!-- ``` -->
 
-# How do learners fare against the "pure" strategies
+<!-- # How do learners fare against the "pure" strategies? -->
 
-Finally, in a new set of simulations we explore the question of how competitive learners are in competition with the pure strategies. Figure \ref{fig:types_comp}
-show the results of competition among the three strategies.
+<!-- Finally, in a new set of simulations we explore the question of how competitive learners are in competition with the pure strategies. Figure \ref{fig:types_comp} -->
+<!-- show the results of competition among the three strategies.
 The population starts with a 1:1 ratio of pure hawks and doves, and learners are introduced by mutation.
@@
 In the end, there is a polymorphic equilibrium with the three strategies coexisting, with learners being the most frequent of the three.
+-->
 
-```{r fig8, echo=FALSE, fig.cap="\\label{fig:types_comp}Evolutionary dynamics of the learners competing against pure strategies.", out.width = '80%'}
-knitr::include_graphics("Simulations//mutType_//evolDynTypes.png")
-```
+<!-- ```{r fig8, echo=FALSE, fig.cap="\\label{fig:types_comp}Evolutionary dynamics of the learners competing against pure strategies.", out.width = '80%'} -->
+<!-- knitr::include_graphics("Simulations//mutType_//evolDynTypes.png") -->
+<!-- ``` -->
 
-# What's next?
+<!-- # How do learners behave when signals are honest? -->
 
-Let reaction norm evolve, under different initial conditions.
+<!-- ```{r fig9, echo=FALSE,fig.cap="\\label{fig:honest_learn}Actor and critic developed when signals are honest.
+Left panel: honest signal.
+Right panel: learners behavioral response ",out.width='100%',fig.show='hold'} -->
+<!-- knitr::include_graphics("Simulations//learHonest_//QualStDv_//weightsVarQualSt.png") -->
 
-# Cited literature
+<!-- ``` -->
+
+<!-- # What's next? -->
+
+<!-- Let reaction norm evolve, under different initial conditions.  -->
+
+<!-- # Cited literature -->

### 2019-09-05 QD_presentation.Rmd
more on the QD presentation

@@
     \usepackage{float}
+    \usepackage{multicol}
+    \usepackage{amsmath} 
+    \usepackage{graphicx}
                 #\floatplacement{figure}{H}
@@
 ```{r,echo=FALSE,fig.align="center",out.width='50%'}
-knitr::include_graphics("Images//hawk_dove.jpeg")
+knitr::include_graphics("Images//HDgame.jpeg")
 ```
@@
 \begin{align*}
+w_H &= p_H \frac{V-C}{2}+(1-p_h) V\\
+w_D &= p_H 0 + (1-p_H)\frac{V}{2}
 \end{align*}
-
-```{r, echo=FALSE, out.width = '80%'}
-knitr::include_graphics("Images/Hawk_dove_plot.png")
+\vspace{-0.8cm}
+```{r, echo=FALSE,fig.align="center", out.width = '55%'}
+knitr::include_graphics("Images//Hawk_dove_plot.png")
 ```
@@
 knitr::include_graphics("Simulations/mutType_/basicHawkDove.png")
+```
+
+## What about signals?
+
+```{r,echo=FALSE,fig.align='center',out.height='80%'}
+knitr::include_graphics("Images//hosp_alexandra_mackenzie.jpg")
+
+```
+
+## What about signals?
+
+### When are signals honest?
+
+* Impossible to fake
+* Individuals have common interests
+* Handicap principle (signal´s cost is proportional to quality)
+  + Social costs?
+
+```{r,echo=FALSE, fig.align='center',out.width='100%'}
+knitr::include_graphics("Images//social_costs.png")
+```
+
+## What about signals?
+
+```{r,echo=FALSE,fig.align='center',out.width='90%'}
+knitr::include_graphics("Images//comSystemBot.png")
+```
+
+## What about learning?
+
+```{r,echo=FALSE,fig.align='center',out.width='90%'}
+knitr::include_graphics("Images//pavlov.jpg")
+```
+
+# Associative learning
+
+## Reinforcement learning theory
+
+\begin{columns}[T]
+ \column{0.5\linewidth}
+ \begin{equation*}
+  \Delta V_{t(s)}=\alpha \underbrace{(R_t-V_t)}_\text{prediction error}
+  \end{equation*}
+  \pause
+  \includegraphics[width=1\textwidth]{Images//learn_curve.png}
+ \column{0.5\linewidth}
+  \pause
+  \includegraphics[height=.9\textheight]{Images//predErrorNeur.png}
+\end{columns}
+
+## Reinforcement learning theory
+
+```{r,echo=FALSE,fig.align='center',out.width='90%'}
+knitr::include_graphics("Images//learnUpdate.png")
+```
+
+## Environmental states
+
+### Discrete states
+\includegraphics[width=0.8\textwidth]{Images//discreteStates.png}
+
+\pause
+
+### Continuos states
+\includegraphics[width=0.8\textwidth]{Images//contStates.png}
+
+## Continuos environmental states
+
+```{r fig2, echo=FALSE, fig.cap="\\label{fig:learning_cartoonRBF}Function approximation for the actor and the critic.
+Using random values for the responses of each center.
+Circles show the location of the centers and the maximum response they trigger, while the lines show the total response along the badge size continium.", out.width = '80%'}
+knitr::include_graphics("cartoonRBF.png")
 ```
@@
 -->
-
-<!-- ```{r fig2, echo=FALSE, fig.cap="\\label{fig:learning_cartoonRBF}Function approximation for the actor and the critic.
-Using random values for the responses of each center.
-Circles show the location of the centers and the maximum response they trigger, while the lines show the total response along the badge size continium.", out.width = '80%'} -->
-<!-- knitr::include_graphics("cartoonRBF.png") -->
-<!-- ``` -->
 

### 2019-09-05 CartoonRBF.R

@@ -53,9 +53,11 @@ lines(x=rangx,y=totRBF(rangx,centers,0.01,means),col=2,lwd=2)
 
 # Cartoon Logist
 
-logist(0.5,alpha = 3,beta = 6)
+logist(0.5,alpha = 5,beta = 10)
 
 plot(logist(seq(0,1,length.out = 1000),alpha = 3,beta = 6)~
        seq(0,1,length.out = 1000),type="l")
 
+plot(logist(seq(-1,1,length.out = 1000),alpha = 0,beta = 2)~
+       seq(-1,1,length.out = 1000),type="l")
 

### 2019-09-05 parTOjson.R

@@ -17,13 +17,13 @@ param<-list(totGen   = 5,   nRep    = 30,
             nInt        = 5000,  init     = c(0,0,1),
             mutRate  = 0.02,  mutType  = 0,
             sampleSize = 20,   strQual  = 10,
-            alphaBad	 = 5,    betaBad	 = 10,
+            alphaBad	 = 3,    betaBad	 = 6,
             alphaCrit  = 0.01,  alphaAct = 0.01,
             sigSq   	 = 0.01, nCenters = 5,
             QualStDv   = 1.1,  
             payoff_matrix = c(1.5,1,0,0.5),
-            namParam = "alphaBad",
-            rangParam = c(1,2,3),
+            namParam = "strQual",
+            rangParam = c(1,2,5),
             folder=paste(here("Simulations"),"/",sep=""))
 
 
@@ -43,7 +43,7 @@ listfolders<-check_create.dir(here("Simulations","alphaAC_"),
 
 for (i in 1:1) {
   param$folder<-paste0(here("Simulations","learHonest_",
-                            paste0(param$namParam,1)),"_/")
+                            paste0(param$namParam,"1")),"_/")
   outParam<-toJSON(param,auto_unbox = TRUE,pretty = TRUE)
   if(file.exists(paste(param$folder,fileName,sep = ''))){
     currFile<-fromJSON(paste(param$folder,fileName,sep = ''))

### 2019-09-05 popSample_1.R

@@ -11,7 +11,7 @@ source(here("AccFunc.R"))
 
 # Scenario to be plotted - corresponds to folders where simulations are stored
 
-scenario<-"learHonest_/alphaBad1"
+scenario<-"learHonest_/strQual1"
 
 
 # Load files -------------------------------------------------------------------
@@ -20,7 +20,7 @@ scenario<-"learHonest_/alphaBad1"
 (evolList<-grep("evol",listTest,value=TRUE))
 (indList<-grep("ind",listTest,value=TRUE))
 
-fileId<-1
+fileId<-2
 evol<-fread(here("Simulations",paste0(scenario,"_"),evolList[fileId]))
 pop<-fread(here("Simulations",paste0(scenario,"_"),indList[fileId]))
 

### 2019-09-06 QD_presentation.tex

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  ignorenonframetext,
]{beamer}
\usepackage{pgfpages}
\setbeamertemplate{caption}[numbered]
\setbeamertemplate{caption label separator}{: }
\setbeamercolor{caption name}{fg=normal text.fg}
\beamertemplatenavigationsymbolsempty
% Prevent slide breaks in the middle of a paragraph:
\widowpenalties 1 10000
\raggedbottom
\setbeamertemplate{part page}{
  \centering
  \begin{beamercolorbox}[sep=16pt,center]{part title}
    \usebeamerfont{part title}\insertpart\par
  \end{beamercolorbox}
}
\setbeamertemplate{section page}{
  \centering
  \begin{beamercolorbox}[sep=12pt,center]{part title}
    \usebeamerfont{section title}\insertsection\par
  \end{beamercolorbox}
}
\setbeamertemplate{subsection page}{
  \centering
  \begin{beamercolorbox}[sep=8pt,center]{part title}
    \usebeamerfont{subsection title}\insertsubsection\par
  \end{beamercolorbox}
}
\AtBeginPart{
  \frame{\partpage}
}
\AtBeginSection{
  \ifbibliography
  \else
    \frame{\sectionpage}
  \fi
}
\AtBeginSubsection{
  \frame{\subsectionpage}
}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={The evolution of badges of status with learners},
  pdfauthor={Andrés Quiñones},
  pdfborder={0 0 0},
  breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\newif\ifbibliography
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-2}

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{float} \usepackage{multicol} \usepackage{amsmath} \usepackage{graphicx}

\title{The evolution of badges of status with learners}
\author{Andrés Quiñones}
\date{}

\begin{document}
\frame{\titlepage}

\begin{frame}{The Hawk-Dove game}
\protect\hypertarget{the-hawk-dove-game}{}

Individuals have one of two genetically determined phenotypic
strategies. \emph{Hawks} are willing to start a conflict over resources,
while \emph{doves} prefer to stand down in the hope to share the
resource without an aggressive contest.

\begin{center}\includegraphics[width=0.5\linewidth]{Images//HDgame} \end{center}

\end{frame}

\begin{frame}{The hawk-dove game}
\protect\hypertarget{the-hawk-dove-game-1}{}

\begin{align*}
w_H &= p_H \frac{V-C}{2}+(1-p_h) V\\
w_D &= p_H 0 + (1-p_H)\frac{V}{2}
\end{align*} \vspace{-0.8cm}

\begin{center}\includegraphics[width=0.55\linewidth]{Images//Hawk_dove_plot} \end{center}

\end{frame}

\begin{frame}{The hawk-dove game}
\protect\hypertarget{the-hawk-dove-game-2}{}

\begin{figure}
\includegraphics[width=0.8\linewidth]{Simulations/mutType_/basicHawkDove} \caption{\label{fig:HD_game}Hawk-dove game. Dashed line is the  game theoretical prediction for frequency of hawks.}\label{fig:unnamed-chunk-3}
\end{figure}

\end{frame}

\begin{frame}{What about signals?}
\protect\hypertarget{what-about-signals}{}

\begin{center}\includegraphics[height=0.8\textheight]{Images//hosp_alexandra_mackenzie} \end{center}

\end{frame}

\begin{frame}{What about signals?}
\protect\hypertarget{what-about-signals-1}{}

\begin{block}{When are signals honest?}

\begin{itemize}
\tightlist
\item
  Impossible to fake
\item
  Individuals have common interests
\item
  Handicap principle (signal´s cost is proportional to quality)

  \begin{itemize}
  \tightlist
  \item
    Social costs?
  \end{itemize}
\end{itemize}

\begin{center}\includegraphics[width=1\linewidth]{Images//social_costs} \end{center}

\end{block}

\end{frame}

\begin{frame}{What about signals?}
\protect\hypertarget{what-about-signals-2}{}

\begin{center}\includegraphics[width=0.9\linewidth]{Images//comSystemBot} \end{center}

\end{frame}

\begin{frame}{What about learning?}
\protect\hypertarget{what-about-learning}{}

\begin{center}\includegraphics[width=0.9\linewidth]{Images//pavlov} \end{center}

\end{frame}

\hypertarget{associative-learning}{%
\section{Associative learning}\label{associative-learning}}

\begin{frame}{Reinforcement learning theory}
\protect\hypertarget{reinforcement-learning-theory}{}

\begin{columns}[T]
 \column{0.5\linewidth}
 \begin{equation*}
  \Delta V_{t(s)}=\alpha \underbrace{(R_t-V_t)}_\text{prediction error}
  \end{equation*}
  \pause
  \includegraphics[width=1\textwidth]{Images//learn_curve.png}
 \column{0.5\linewidth}
  \pause
  \includegraphics[height=.9\textheight]{Images//predErrorNeur.png}
\end{columns}

\end{frame}

\begin{frame}{Reinforcement learning theory}
\protect\hypertarget{reinforcement-learning-theory-1}{}

\begin{center}\includegraphics[width=0.9\linewidth]{Images//learnUpdate} \end{center}

\end{frame}

\begin{frame}{Environmental states}
\protect\hypertarget{environmental-states}{}

\begin{block}{Discrete states}

\includegraphics[width=0.8\textwidth]{Images//discreteStates.png}

\pause

\end{block}

\begin{block}{Continuos states}

\includegraphics[width=0.8\textwidth]{Images//contStates.png}

\end{block}

\end{frame}

\begin{frame}{Continuos environmental states}
\protect\hypertarget{continuos-environmental-states}{}

\begin{figure}
\includegraphics[width=0.8\linewidth]{cartoonRBF} \caption{\label{fig:learning_cartoonRBF}Function approximation for the actor and the critic. Using random values for the responses of each center. Circles show the location of the centers and the maximum response they trigger, while the lines show the total response along the badge size continium.}\label{fig:fig2}
\end{figure}

\end{frame}

\end{document}

### 2019-09-06 QD_presentation.Rmd
almost donde with QD presentation

@@
     \usepackage{graphicx}
+    \usepackage{array}
                 #\floatplacement{figure}{H}
@@
 
-```{r loadData, echo=FALSE,include=FALSE}
+```{r loadData, echo=FALSE,include=FALSE,warning=FALSE,message=FALSE}
 # Required libraries -----------------------------------------------------------
@@
 ## Continuos environmental states
+\begin{equation*}
+  \Delta V_{t(s)}=\alpha \underbrace{(R_t-V_t)}_\text{prediction error}
+\end{equation*}
 
@@
 Using random values for the responses of each center.
-Circles show the location of the centers and the maximum response they trigger, while the lines show the total response along the badge size continium.", out.width = '80%'}
+Circles show the location of the centers and the maximum response they trigger, while the lines show the total response along the badge size continium.", out.width = '60%'}
 knitr::include_graphics("cartoonRBF.png")
@@
 
-<!-- Individuals interact randomly with each other over their lifetimes and collect resources depending on the strategy of theirs and their partners.
-In contrast to the classical _Hawk-dove_, in here when two _hawks_ interact with each other one of them wins the contest and gets $V-\frac{C}{2}$ as payoff; while the loser gets $-\frac{C}{2}$.
-Where $V$ is the value of the resource individuals are competing for, and $C$ is the cost of an aggressive contest.
-individuals share the cost, but only the winner gets the value of the resource.The rest of payoff matrix follows the standard game.
--->
+# Our model
 
-<!-- Individuals may or may not vary in their quality.
-Quality is a numerical value from 0 to 1, drawn from a truncated normal distribution when each individual is born.
-Quality represents the different conditions under which individuals might be raised, and may influence their competitive abilities.
-Differences in quality between two individuals partly determine who wins in a conflict when both players play hawk.
-The probability of individual A wining a conflict over individual B is given by -->
+## The hawk-dove game 2.0
 
-<!-- \begin{equation} -->
-<!-- p(A)=\frac{1}{1+e^{-\beta(Q_A-Q_B)}}, -->
-<!-- \end{equation} -->
+\begin{center}
+\begin{tabular}{ >{\centering\arraybackslash}p{1cm} | >{\centering\arraybackslash}p{4.5cm} | >{\centering\arraybackslash}p{4.5cm} }
+& H & D \\ \hline
+H & $p_w V\frac{-C}{2} + (1-p_w) \frac{-C}{2}$ & $V$ \\ \hline
+V & $0$ & $\frac{V}{2}$\\
+\end{tabular}
+\end{center}
 
-<!-- where $Q_A$ and $Q_b$ are the qualities of individuals A and B respectively, and $\beta$ is how important quality is in determining who wins the contest.  -->
-<!-- Individuals' reproductive success is proportional to the payoff accumulated -->
-<!-- throughout their life. Figure \ref{fig:HD_game} shows the evolutionary dynamics  -->
-<!-- of  the two genotypes, which fit the game theoretical prediction. -->
+\begin{equation*}
+p_w=\frac{1}{1+e^{-\beta(Q_i-Q_j)}}
+\end{equation*}
 
-<!-- # The effect of learning  -->
+```{r,echo=FALSE,fig.align='center',out.width='50%'}
+par(plt=posPlot(),las=1)
+plot(logist(seq(-1,1,length.out = 1000),alpha = 0,beta = 2)~
+       seq(-1,1,length.out = 1000),type="l",xlab="Difference in quality",
+     ylab = expression(p[w]),cex.lab=2,lwd=3,col="red",cex.axis=2,ylim=c(0,1))
+lines(logist(seq(-1,1,length.out = 1000),alpha = 0,beta = 10)~
+       seq(-1,1,length.out = 1000),lwd=3,col="green")
+legend("topleft",legend=c(2,10),col=c("red","green"),lwd=3,title = expression(beta),cex=2,bty="n")
 
-<!-- We let individuals have a _badge of status_ that potentially signals their  -->
-<!-- quality. To allow for this signalling, we let the size of the _badge of status_  -->
-<!-- of an individual be dependent on her quality through a reaction norm. we assume the size of the _badge_ is a value from 0 to 1. This value is given by -->
-<!-- \begin{equation} -->
-<!-- s_i = \frac{1}{1+e^{-(\alpha_i+\gamma_iQ_i)}} -->
-<!-- \label{eq:react_norm} -->
-<!-- \end{equation} -->
-<!-- The parameters $\alpha$ and $\gamma$ in equation \ref{eq:react_norm} determine  -->
-<!-- the shape of the reaction norm, and hence, how honest the _badge_ is in  -->
-<!-- signalling the quality of the individual.Eventually, the aim is to allow the  -->
-<!-- signalling system to evolve, by letting the parameters of the reaction norm be inherited from parent to offspring and be change by a mutational process. -->
+```
 
-<!-- ## How does learning work? -->
+## Sender Code
 
-<!-- To complete the receiver side of the signalling system, we introduce a new  -->
-<!-- strategy in the game that we call _learner_. _Learners_ use a mixed strategy in  -->
-<!-- the hawk-dove game. Which means, they play hawk (dove) with a certain probability. -->
-<!-- Furthermore, the probability with which the play each strategy changes during their lifetime according to the payoff they receive in previous interactions.
-In other words, _learners_ learn to behave in the game.
-To implement the learning -->
-<!-- process, we use the reinforcement learning formalism.
-Specifically, we use the actor-critic algorithm [@sutton_reinforcement_2018].
-Individuals ( _learners_ ) perceive the _badge_ of their counterpart as the environmental state in which they find themselves.
-They estimate the value of that state and a preference for playing hawk (dove) in that state.
-They update their estimation and preference after each interaction they have.
-Given that the badge size is a real number between 0 and 1, there are infinitely many environmental states.
-Thus, for their value estimation and preference computation, individual must use a form of generalization.
-We let them -->
-<!-- use as generalization a function approximation method based on radial basis functions.
-Each individual uses five _centers_ spread uniformly throughout the range of badge size variation (from 0 to 1).
-Each time an individual interacts with a con-specific the size of the badge of the interacting partner triggers a response from each one of the _centers_.
-This response decreases in strength as the badge size is further from the _center_.
-The decrease follows a radial basis function with Gaussian shape.
-The sum of the responses given by all the centers provide the value estimation and the preference for one of the two pure strategies (hawk or dove).
-During the learning process individuals update the _weight_ of the response that each one of these centers triggers.
-In the case of the preference for the strategies the numerical value resulting from the sum of the responses is translated to a probability of one of the two pure strategies through a logistic function.
-Figure \ref{fig:learning_cartoonRBF} shows a made up example of how the function approximation works for the actor (top panel - probability of choosing dove) and the critic (bottom panel - expected payoff from facing an individual with a given badge size).
--->
+\begin{equation*}
+s_i = \frac{1}{1+e^{-(\epsilon_i+\gamma_iQ_i)}}
+\label{eq:react_norm}
+\end{equation*}
 
-<!-- ## Results when individuals do NOT vary in their quality -->
+```{r,echo=FALSE,fig.align='center',out.width='50%'}
+par(plt=posPlot(),las=1)
+plot(logist(seq(0,1,length.out = 1000),alpha = 2,beta = 4)~
+       seq(0,1,length.out = 1000),type="l",xlab="Quality",
+     ylab = expression(p[w]),cex.lab=2,lwd=3,col="red",cex.axis=2,ylim=c(0,1))
+lines(logist(seq(0,1,length.out = 1000),alpha = 5,beta = 10)~
+       seq(0,1,length.out = 1000),lwd=3,col="green")
+legend("topleft",legend=c(4,10),col=c("red","green"),lwd=3,title = expression(gamma),cex=2,bty="n")
 
-<!-- In Figure \ref{fig:learning_invar} we show the phenotypic frequencies and interaction-type frequencies in 4 generations of a population composed of only _learners_ that do not vary in their quality.
-The population is also monomorphic for the reaction norm parameters ($\alpha=0$ and $\gamma=0$).
-With the values chosen the badge size distribution is the same as the quality distribution.
-Thus, the four generations represent just replicates of the learning process.
-The results show that under quality invariance, the learning algorithm develops similar phenotypic frequencies as the evolutionary dynamics of the pure strategies.
--->
+```
 
-<!-- ```{r fig3, echo=FALSE, fig.cap="\\label{fig:learning_invar}Learning equilibrium when individuals do not vary in quality.
-Dashed line is the  game theoretical prediction for frequency of hawks.", out.width = '80%'} -->
-<!-- knitr::include_graphics("Simulations//QualStDv_//hawkDoveLearn.png") -->
-<!-- ``` -->
+## Receiver code
 
-<!-- ## Results when individuals vary in their quality -->
+\begin{equation*}
+  \Delta V_{t(s)}=\alpha (R_t-V_t)
+\end{equation*}
 
-<!-- In figure \ref{fig:learning_var0.1} we show results of phenotypic and interaction types frequencies from simulations in which individuals vary in their quality.  -->
-<!-- The value of quality is drawn from a truncated normal distribution with mean 0.5  -->
-<!-- and standard deviation 0.1.
-Compared to the expectation from evolutionary dynamics of the standard game, the learning scenario with quality variation shows lower levels of hawks and therefore lower levels of escalated conflict.
-At this point individuals do not respond adaptively to the quality of their partner, instead their preference for a given action is dependent on their own quality.
--->
+```{r, echo=FALSE,fig.align='center', out.width = '60%'}
+knitr::include_graphics("cartoonRBF.png")
+```
 
-<!-- ```{r fig4, echo=FALSE, fig.cap="\\label{fig:learning_var0.1}Learning equilibrium when individuals vary in quality.
-Dashed lines are the  game theoretical predictions in the standard game.", out.width = '80%'} -->
-<!-- knitr::include_graphics("Simulations//QualStDv_//hawkDoveLearn_0.1.png") -->
-<!-- ``` -->
+# Results
 
-<!-- ## Overall effect of variance -->
+## When individuals do _NOT_ vary in quality
 
-<!-- In figure \ref{fig:learning_all_var} we show how the effect of increasing  -->
-<!-- variance in quality further decreases the amount of individuals in a population  -->
-<!-- expressing the hawk phenotype. -->
+```{r , echo=FALSE, fig.align='center',out.width = '80%'}
+knitr::include_graphics("Simulations//QualStDv_//hawkDoveLearn.png")
+```
 
-<!-- ```{r fig5, echo=FALSE, fig.cap="\\label{fig:learning_all_var}Effect of variation in quality for the learning equilibrium.", out.width = '80%'} -->
-<!-- knitr::include_graphics("Simulations//QualStDv_//effectQualVariance.png") -->
-<!-- ``` -->
+## When individuals _DO_ vary in quality
 
-<!-- ## How does learning look like in the Hawk-Dove game? -->
+```{r, echo=FALSE, fig.align='center', out.width = '80%'}
+knitr::include_graphics("Simulations//QualStDv_//hawkDoveLearn_1.png")
+```
 
-<!-- Figures \ref{fig:learning_var0.1} and \ref{fig:learning_all_var} only show how learners behave at a population level, but they do not show the inter-individual variation behind those frequencies.
-In figure \ref{fig:AC_HD} we show the result of the function approximation mechanism arising at the end of a generation for a sample of the individuals of the populations.
-These results are shown under three different levels of variation in quality.
-All three panes show that through their learning processes individuals split up in two groups, one which mostly behaves like a hawk and a second that mostly behaves like a dove.
-The group of "hawks" is mostly, but no exclusively, composed of individuals whose quality is on the upper half of the quality distribution.
-In accordance, the group of "doves" is mostly composed of individuals whose quality is on the lower half of the quality distribution.
-The relation between quality and the preferred behavior of individuals obviously only holds in scenarios with variance in quality.
-In the scenario with quality invariance, the split of individuals in two groups with a tendency for each behavior nevertheless remains.
--->
+## Overal effect of quality variation
 
-<!-- ```{r fig6, echo=FALSE, fig.cap="\\label{fig:AC_HD}Within population variation in the learning parameters at the end of learning for three scenarios of variation in the quality of individuals.", out.width = '70%'} -->
-<!-- knitr::include_graphics("Simulations//QualStDv_//WeightsVarQualSt.png") -->
-<!-- ``` -->
+```{r fig5, echo=FALSE, fig.align='center', out.width = '80%'}
+knitr::include_graphics("Simulations//QualStDv_//effectQualVariance.png")
+```
 
-<!-- ## How do the learning dynamics look like -->
+## How does variation among _learners_ look like?
 
-<!-- In figure \ref{fig:learn_dyn} we show the learning trajectory, for the actor and the critic, of a sample of individuals.
-These dynamics show how individuals split in the two groups according to their experiences.
-Despite the robustness of the bi-modal distribution along the learning dynamics, it is not uncommon that individual go from one group to the other.
--->
+```{r fig6, echo=FALSE, fig.align='center',out.width = '90%'}
+knitr::include_graphics("Simulations//QualStDv_//WeightsVarQualSt.png")
+```
 
-<!-- ```{r fig7, echo=FALSE, fig.cap="\\label{fig:learn_dyn}Learning dynamics of the individuals in a population.", out.width = '80%'} -->
-<!-- knitr::include_graphics("Simulations//mutType_//learnDyn200.png") -->
-<!-- ``` -->
+## How do the learning dynamics look like?
 
-<!-- # How do learners fare against the "pure" strategies? -->
+```{r fig7, echo=FALSE, out.width = '80%'}
+knitr::include_graphics("Simulations//mutType_//learnDyn200.png")
+```
 
-<!-- Finally, in a new set of simulations we explore the question of how competitive learners are in competition with the pure strategies. Figure \ref{fig:types_comp} -->
-<!-- show the results of competition among the three strategies.
-The population starts with a 1:1 ratio of pure hawks and doves, and learners are introduced by mutation.
-Learners manage to increase in frequency.
-They, however, do not manage to outcompete the pure strategies.
-In the end, there is a polymorphic equilibrium with the three strategies coexisting, with learners being the most frequent of the three.
--->
+## How do learners behave when signals are honest?
 
-<!-- ```{r fig8, echo=FALSE, fig.cap="\\label{fig:types_comp}Evolutionary dynamics of the learners competing against pure strategies.", out.width = '80%'} -->
-<!-- knitr::include_graphics("Simulations//mutType_//evolDynTypes.png") -->
-<!-- ``` -->
+```{r fig9, echo=FALSE,out.width='90%',fig.show='hold'}
+knitr::include_graphics("Simulations//learHonest_//QualStDv_//weightsVarQualSt.png")
+```
 
-<!-- # How do learners behave when signals are honest? -->
+## How do learners fare against pure strategies?
 
-<!-- ```{r fig9, echo=FALSE,fig.cap="\\label{fig:honest_learn}Actor and critic developed when signals are honest.
-Left panel: honest signal.
-Right panel: learners behavioral response ",out.width='100%',fig.show='hold'} -->
-<!-- knitr::include_graphics("Simulations//learHonest_//QualStDv_//weightsVarQualSt.png") -->
+```{r fig8, echo=FALSE, fig.align='center', out.width = '80%'}
+knitr::include_graphics("Simulations//mutType_//evolDynTypes.png")
+```
 
-<!-- ``` -->
+## What's next?
 
-<!-- # What's next? -->
-
-<!-- Let reaction norm evolve, under different initial conditions.  -->
-
-<!-- # Cited literature -->
+* Let reaction norm evolve, under different initial conditions.
+* Let learning parmeters evolve
+* Let the communication system co-evolve

### 2019-09-06 QD_presentation.tex
almost donde with QD presentation

@@
-\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
-\PassOptionsToPackage{hyphens}{url}
-%
-\documentclass[
-  ignorenonframetext,
-]{beamer}
-\usepackage{pgfpages}
+\documentclass[ignorenonframetext,]{beamer}
 \setbeamertemplate{caption}[numbered]
@@
 \beamertemplatenavigationsymbolsempty
-% Prevent slide breaks in the middle of a paragraph:
-\widowpenalties 1 10000
-\raggedbottom
-\setbeamertemplate{part page}{
-  \centering
-  \begin{beamercolorbox}[sep=16pt,center]{part title}
-    \usebeamerfont{part title}\insertpart\par
-  \end{beamercolorbox}
-}
-\setbeamertemplate{section page}{
-  \centering
-  \begin{beamercolorbox}[sep=12pt,center]{part title}
-    \usebeamerfont{section title}\insertsection\par
-  \end{beamercolorbox}
-}
-\setbeamertemplate{subsection page}{
-  \centering
-  \begin{beamercolorbox}[sep=8pt,center]{part title}
-    \usebeamerfont{subsection title}\insertsubsection\par
-  \end{beamercolorbox}
-}
-\AtBeginPart{
-  \frame{\partpage}
-}
-\AtBeginSection{
-  \ifbibliography
-  \else
-    \frame{\sectionpage}
-  \fi
-}
-\AtBeginSubsection{
-  \frame{\subsectionpage}
-}
 \usepackage{lmodern}
@@
 \usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
 \ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
@@
   \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provides euro and other symbols
 \else % if luatex or xelatex
-  \usepackage{unicode-math}
-  \defaultfontfeatures{Scale=MatchLowercase}
-  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+  \ifxetex
+    \usepackage{mathspec}
+  \else
+    \usepackage{fontspec}
+  \fi
+  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
 \fi
@@
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\IfFileExists{microtype.sty}{% use microtype if available
-  \usepackage[]{microtype}
+% use microtype if available
+\IfFileExists{microtype.sty}{%
+\usepackage{microtype}
 \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
 }{}
-\makeatletter
-\@ifundefined{KOMAClassName}{% if non-KOMA class
-  \IfFileExists{parskip.sty}{%
-    \usepackage{parskip}
-  }{% else
-    \setlength{\parindent}{0pt}
-    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
-}{% if KOMA class
-  \KOMAoptions{parskip=half}}
-\makeatother
-\usepackage{xcolor}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\newif\ifbibliography
 \hypersetup{
@@
 \urlstyle{same}  % don't use monospace font for urls
-\newif\ifbibliography
+
+% Prevent slide breaks in the middle of a paragraph:
+\widowpenalties 1 10000
+\raggedbottom
+
+\AtBeginPart{
+  \let\insertpartnumber\relax
+  \let\partname\relax
+  \frame{\partpage}
+}
+\AtBeginSection{
+  \ifbibliography
+  \else
+    \let\insertsectionnumber\relax
+    \let\sectionname\relax
+    \frame{\sectionpage}
+  \fi
+}
+\AtBeginSubsection{
+  \let\insertsubsectionnumber\relax
+  \let\subsectionname\relax
+  \frame{\subsectionpage}
+}
+
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
 \setlength{\emergencystretch}{3em}  % prevent overfull lines
@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{-2}
-
-% set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
-
-\usepackage{float} \usepackage{multicol} \usepackage{amsmath} \usepackage{graphicx}
+\setcounter{secnumdepth}{0}
+\usepackage{float} \usepackage{multicol} \usepackage{amsmath}
+\usepackage{graphicx} \usepackage{array}
 
@@
 \begin{frame}{The Hawk-Dove game}
-\protect\hypertarget{the-hawk-dove-game}{}
 
@@
 \begin{frame}{The hawk-dove game}
-\protect\hypertarget{the-hawk-dove-game-1}{}
 
@@
 w_D &= p_H 0 + (1-p_H)\frac{V}{2}
-\end{align*} \vspace{-0.8cm}
+\end{align*}
+
+\vspace{-0.8cm}
 
@@
 \begin{frame}{The hawk-dove game}
-\protect\hypertarget{the-hawk-dove-game-2}{}
 
@@
 \begin{frame}{What about signals?}
-\protect\hypertarget{what-about-signals}{}
 
@@
 \begin{frame}{What about signals?}
-\protect\hypertarget{what-about-signals-1}{}
 
@@
   Handicap principle (signal´s cost is proportional to quality)
-
-  \begin{itemize}
-  \tightlist
 \item
   Social costs?
-  \end{itemize}
 \end{itemize}
@@
 \begin{frame}{What about signals?}
-\protect\hypertarget{what-about-signals-2}{}
 
@@
 \begin{frame}{What about learning?}
-\protect\hypertarget{what-about-learning}{}
 
@@
 
-\hypertarget{associative-learning}{%
-\section{Associative learning}\label{associative-learning}}
+\section{Associative learning}\label{associative-learning}
 
 \begin{frame}{Reinforcement learning theory}
-\protect\hypertarget{reinforcement-learning-theory}{}
 
@@
 \begin{frame}{Reinforcement learning theory}
-\protect\hypertarget{reinforcement-learning-theory-1}{}
 
@@
 \begin{frame}{Environmental states}
-\protect\hypertarget{environmental-states}{}
 
@@
 \begin{frame}{Continuos environmental states}
-\protect\hypertarget{continuos-environmental-states}{}
+
+\begin{equation*}
+  \Delta V_{t(s)}=\alpha \underbrace{(R_t-V_t)}_\text{prediction error}
+\end{equation*}
 
 \begin{figure}
-\includegraphics[width=0.8\linewidth]{cartoonRBF} \caption{\label{fig:learning_cartoonRBF}Function approximation for the actor and the critic.
+\includegraphics[width=0.6\linewidth]{cartoonRBF} \caption{\label{fig:learning_cartoonRBF}Function approximation for the actor and the critic.
 Using random values for the responses of each center.
@@
 
+\section{Our model}\label{our-model}
+
+\begin{frame}{The hawk-dove game 2.0}
+
+\begin{center}
+\begin{tabular}{ >{\centering\arraybackslash}p{1cm} | >{\centering\arraybackslash}p{4.5cm} | >{\centering\arraybackslash}p{4.5cm} }
+& H & D \\ \hline
+H & $p_w V\frac{-C}{2} + (1-p_w) \frac{-C}{2}$ & $V$ \\ \hline
+V & $0$ & $\frac{V}{2}$\\
+\end{tabular}
+\end{center}
+
+\begin{equation*}
+p_w=\frac{1}{1+e^{-\beta(Q_i-Q_j)}}
+\end{equation*}
+
+\begin{center}\includegraphics[width=0.5\linewidth]{QD_presentation_files/figure-beamer/unnamed-chunk-9-1} \end{center}
+
+\end{frame}
+
+\begin{frame}{Sender Code}
+
+\begin{equation*}
+s_i = \frac{1}{1+e^{-(\epsilon_i+\gamma_iQ_i)}}
+\label{eq:react_norm}
+\end{equation*}
+
+\begin{center}\includegraphics[width=0.5\linewidth]{QD_presentation_files/figure-beamer/unnamed-chunk-10-1} \end{center}
+
+\end{frame}
+
+\begin{frame}{Receiver code}
+
+\begin{equation*}
+  \Delta V_{t(s)}=\alpha (R_t-V_t)
+\end{equation*}
+
+\begin{center}\includegraphics[width=0.6\linewidth]{cartoonRBF} \end{center}
+
+\end{frame}
+
+\section{Results}\label{results}
+
+\begin{frame}{When individuals do \emph{NOT} vary in quality}
+
+\begin{center}\includegraphics[width=0.8\linewidth]{Simulations//QualStDv_//hawkDoveLearn} \end{center}
+
+\end{frame}
+
+\begin{frame}{When individuals \emph{DO} vary in quality}
+
+\begin{center}\includegraphics[width=0.8\linewidth]{Simulations//QualStDv_//hawkDoveLearn_1} \end{center}
+
+\end{frame}
+
+\begin{frame}{Overal effect of quality variation}
+
+\begin{center}\includegraphics[width=0.8\linewidth]{Simulations//QualStDv_//effectQualVariance} \end{center}
+
+\end{frame}
+
+\begin{frame}{How does variation among \emph{learners} look like?}
+
+\begin{center}\includegraphics[width=0.7\linewidth]{Simulations//QualStDv_//WeightsVarQualSt} \end{center}
+
+\end{frame}
+
+\begin{frame}{How do the learning dynamics look like?}
+
+\includegraphics[width=0.8\linewidth]{Simulations//mutType_//learnDyn200}
+
+\end{frame}
+
+\begin{frame}{How do learners behave when signals are honest?}
+
+\includegraphics[width=1\linewidth]{Simulations//learHonest_//QualStDv_//weightsVarQualSt}
+
+\end{frame}
+
+\begin{frame}{How do learners fare against pure strategies?}
+
+\begin{center}\includegraphics[width=0.8\linewidth]{Simulations//mutType_//evolDynTypes} \end{center}
+
+\end{frame}
+
+\begin{frame}{What's next?}
+
+\begin{itemize}
+\tightlist
+\item
+  Let reaction norm evolve, under different initial conditions.
+\item
+  Let learning parmeters evolve
+\item
+  Let the communication system co-evolve
+\end{itemize}
+
+\end{frame}
+
 \end{document}

### 2019-09-06 AccFunc.R

@@ -73,11 +73,16 @@ color.bar.aeqp <- function(lut, min, max=-min, nticks=11, ticks=seq(min, max, le
 # not robust yet to filename variation CHECK!!! --------------------------------
 
 filesScenar<-function(filename,scenario){
-  extPar<-strsplit(filename,scenario)[[1]][2]
+  splitScen<-tail(strsplit(scenario,"_")[[1]],1)
+  splitScen<-gsub("/",splitScen,replacement = '')
+  extPar<-strsplit(filename,splitScen)[[1]][2]
   parVal<-as.numeric(substr(gsub("[[:alpha:]]",extPar,replacement = ''),1,3))
   tmp<-fread(here("Simulations",paste0(scenario,"_"),filename))
-  tmp[,eval(scenario):=parVal]
+  tmp[,eval(splitScen):=parVal]
   return(tmp)
 }
-
+splitScen<-tail(strsplit(scenario,"_")[[1]],1)
+splitScen<-gsub("/",splitScen,replacement = '')
+strsplit(indList[1],'QualStDv')[[1]][2]
+as.numeric(substr(gsub("[[:alpha:]]","0.1.txt",replacement = ''),1,3))
 

### 2019-09-06 compPar.R

@@ -1,0 +1,81 @@
+# Varying parameter of a single scenario ---------------------------------------
+
+# Required libraries -----------------------------------------------------------
+
+library(here)
+here()
+source(here("AccFunc.R"))
+
+# Scenario to be plotted - corresponds to folders where simulations are stored
+
+scenario<-"learHonest_/QualStDv"
+
+# Load files -------------------------------------------------------------------
+
+(listTest<-list.files(here("Simulations",paste0(scenario,"_"))))
+(sdList<-grep("evol",listTest,value=TRUE))
+(indList<-grep("ind",listTest,value=TRUE))
+
+evol<-do.call(rbind,lapply(sdList,filesScenar,scenario))
+inds<-do.call(rbind,lapply(indList,filesScenar,scenario))
+
+
+# Graphs -----------------------------------------------------------------------
+
+gener<-tail(pop[,unique(time)],1)
+lastInt<-tail(pop[,unique(nInteract)],3)
+runChoi<-0
+nCenters<-5
+interv<-1/nCenters
+centers<-interv*0.5+interv*seq(0,nCenters-1)
+rangx<-seq(0,1,length=1000)
+tempPop<-inds[time==gener&nInteract==lastInt[1],.SD[.N],
+             .SDcol=c(grep("Weight",
+                           names(evol),value = TRUE),"Quality","QualStDv"),
+             by=indId]
+dataIndAct_0.1<-sapply(as.list(tempPop[QualStDv==0.1,indId]),
+                   function(x){x=
+                     logist(totRBF(rangx,
+                                   centers,0.01,
+                                   as.double(
+                                     tempPop[indId==x,.SD,
+                                             .SDcol=grep("WeightAct",
+                                                         names(tempPop),
+                                                         value = TRUE)
+                                             ])),alpha=0,beta = 1)})
+
+dataIndAct_1.1<-sapply(as.list(tempPop[QualStDv==1.1,indId]),
+                       function(x){x=
+                         logist(totRBF(rangx,
+                                       centers,0.01,
+                                       as.double(
+                                         tempPop[indId==x,.SD,
+                                                 .SDcol=grep("WeightAct",
[…]

### 2019-09-06 QD_presentation.Rmd
done with QD presentation

@@
 
+## What parameters influence responsiveness?
+
+### Distribution of quality
+
+```{r,echo=FALSE,fig.align='center',out.width='90%'}
+knitr::include_graphics("Simulations//learHonest_//QualStDv_//compDistrNorm1Unif.png")
+```
+
+## What parameters influence responsiveness?
+
+### Distribution of quality
+
+```{r,echo=FALSE,fig.align='center',out.width='90%'}
+knitr::include_graphics("Simulations//learHonest_//QualStDv_//compDistrDiff.png")
+```
+
+## What parameters influence responsiveness?
+
+### Strength of quality influence on competition outcome
+
+```{r,echo=FALSE,fig.align='center',out.width='90%'}
+knitr::include_graphics("Simulations//learHonest_//strQual_//compStr1_last.png")
+```
+
+## What parameters influence responsiveness?
+
+### Strength of quality influence on competition outcome
+
+```{r,echo=FALSE,out.width='45%',fig.show='hold'}
+knitr::include_graphics("Simulations//learHonest_//strQual_//compStrDiff.png")
+knitr::include_graphics("Simulations//learHonest_//strQual_//compStrpDove.png")
+```
+
+# Why the hell is responsiveness so low?
+
 ## How do learners fare against pure strategies?
@@
 
+## Concluding
+
+* Learning is good, even if you do not learn about your opponent. You still learn about you
+* Learning reduces aggressive interactions
+* Learning splits up individuals into _types_
+
 ## What's next?

### 2019-09-06 QD_presentation.tex
done with QD presentation

--- 2019-09-06 QD_presentation.tex
@@
 
-\usepackage{float} \usepackage{multicol} \usepackage{amsmath} \usepackage{graphicx}
+\usepackage{float} \usepackage{multicol} \usepackage{amsmath} \usepackage{graphicx} \usepackage{array}
 
@@
 
+\begin{equation*}
+  \Delta V_{t(s)}=\alpha \underbrace{(R_t-V_t)}_\text{prediction error}
+\end{equation*}
+
 \begin{figure}
-\includegraphics[width=0.8\linewidth]{cartoonRBF} \caption{\label{fig:learning_cartoonRBF}Function approximation for the actor and the critic.
+\includegraphics[width=0.6\linewidth]{cartoonRBF} \caption{\label{fig:learning_cartoonRBF}Function approximation for the actor and the critic.
 Using random values for the responses of each center.
@@
 
+\hypertarget{our-model}{%
+\section{Our model}\label{our-model}}
+
+\begin{frame}{The hawk-dove game 2.0}
+\protect\hypertarget{the-hawk-dove-game-2.0}{}
+
+\begin{center}
+\begin{tabular}{ >{\centering\arraybackslash}p{1cm} | >{\centering\arraybackslash}p{4.5cm} | >{\centering\arraybackslash}p{4.5cm} }
+& H & D \\ \hline
+H & $p_w V\frac{-C}{2} + (1-p_w) \frac{-C}{2}$ & $V$ \\ \hline
+V & $0$ & $\frac{V}{2}$\\
+\end{tabular}
+\end{center}
+
+\begin{equation*}
+p_w=\frac{1}{1+e^{-\beta(Q_i-Q_j)}}
+\end{equation*}
+
+\begin{center}\includegraphics[width=0.5\linewidth]{QD_presentation_files/figure-beamer/unnamed-chunk-9-1} \end{center}
+
+\end{frame}
+
+\begin{frame}{Sender Code}
+\protect\hypertarget{sender-code}{}
+
+\begin{equation*}
+s_i = \frac{1}{1+e^{-(\epsilon_i+\gamma_iQ_i)}}
+\label{eq:react_norm}
+\end{equation*}
+
+\begin{center}\includegraphics[width=0.5\linewidth]{QD_presentation_files/figure-beamer/unnamed-chunk-10-1} \end{center}
+
+\end{frame}
+
+\begin{frame}{Receiver code}
+\protect\hypertarget{receiver-code}{}
+
+\begin{equation*}
+  \Delta V_{t(s)}=\alpha (R_t-V_t)
+\end{equation*}
+
+\begin{center}\includegraphics[width=0.6\linewidth]{cartoonRBF} \end{center}
+
+\end{frame}
+
+\hypertarget{results}{%
+\section{Results}\label{results}}
+
+\begin{frame}{When individuals do \emph{NOT} vary in quality}
+\protect\hypertarget{when-individuals-do-not-vary-in-quality}{}
+
+\begin{center}\includegraphics[width=0.8\linewidth]{Simulations//QualStDv_//hawkDoveLearn} \end{center}
+
+\end{frame}
+
+\begin{frame}{When individuals \emph{DO} vary in quality}
+\protect\hypertarget{when-individuals-do-vary-in-quality}{}
+
+\begin{center}\includegraphics[width=0.8\linewidth]{Simulations//QualStDv_//hawkDoveLearn_1} \end{center}
+
+\end{frame}
+
+\begin{frame}{Overal effect of quality variation}
+\protect\hypertarget{overal-effect-of-quality-variation}{}
+
+\begin{center}\includegraphics[width=0.8\linewidth]{Simulations//QualStDv_//effectQualVariance} \end{center}
+
+\end{frame}
+
+\begin{frame}{How does variation among \emph{learners} look like?}
+\protect\hypertarget{how-does-variation-among-learners-look-like}{}
+
+\begin{center}\includegraphics[width=0.9\linewidth]{Simulations//QualStDv_//WeightsVarQualSt} \end{center}
+
+\end{frame}
+
+\begin{frame}{How do the learning dynamics look like?}
+\protect\hypertarget{how-do-the-learning-dynamics-look-like}{}
+
+\includegraphics[width=0.8\linewidth]{Simulations//mutType_//learnDyn200}
+
+\end{frame}
+
+\begin{frame}{How do learners behave when signals are honest?}
+\protect\hypertarget{how-do-learners-behave-when-signals-are-honest}{}
+
+\includegraphics[width=0.9\linewidth]{Simulations//learHonest_//QualStDv_//weightsVarQualSt}
+
+\end{frame}
+
+\begin{frame}{What parameters influence responsiveness?}
+\protect\hypertarget{what-parameters-influence-responsiveness}{}
+
+\begin{block}{Distribution of quality}
+
+\begin{center}\includegraphics[width=0.9\linewidth]{Simulations//learHonest_//QualStDv_//compDistrNorm1Unif} \end{center}
+
+\end{block}
+
+\end{frame}
+
+\begin{frame}{What parameters influence responsiveness?}
+\protect\hypertarget{what-parameters-influence-responsiveness-1}{}
+
+\begin{block}{Distribution of quality}
+
+\begin{center}\includegraphics[width=0.9\linewidth]{Simulations//learHonest_//QualStDv_//compDistrDiff} \end{center}
+
+\end{block}
+
+\end{frame}
+
+\begin{frame}{What parameters influence responsiveness?}
+\protect\hypertarget{what-parameters-influence-responsiveness-2}{}
+
+\begin{block}{Strength of quality influence on competition outcome}
+
+\begin{center}\includegraphics[width=0.9\linewidth]{Simulations//learHonest_//strQual_//compStr1_last} \end{center}
+
+\end{block}
+
+\end{frame}
+
+\begin{frame}{What parameters influence responsiveness?}
+\protect\hypertarget{what-parameters-influence-responsiveness-3}{}
+
+\begin{block}{Strength of quality influence on competition outcome}
+
+\includegraphics[width=0.45\linewidth]{Simulations//learHonest_//strQual_//compStrDiff}
+\includegraphics[width=0.45\linewidth]{Simulations//learHonest_//strQual_//compStrpDove}
+
+\end{block}
+
+\end{frame}
+
+\hypertarget{why-the-hell-is-responsiveness-so-low}{%
+\section{Why the hell is responsiveness so
+low?}\label{why-the-hell-is-responsiveness-so-low}}
+
+\begin{frame}{How do learners fare against pure strategies?}
+\protect\hypertarget{how-do-learners-fare-against-pure-strategies}{}
+
+\begin{center}\includegraphics[width=0.8\linewidth]{Simulations//mutType_//evolDynTypes} \end{center}
+
+\end{frame}
+
+\begin{frame}{Concluding}
+\protect\hypertarget{concluding}{}
+
+\begin{itemize}
+\tightlist
+\item
+  Learning is good, even if you do not learn about your opponent. You
+  still learn about you
+\item
+  Learning reduces aggressive interactions
+\item
+  Learning splits up individuals into \emph{types}
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}{What's next?}
+\protect\hypertarget{whats-next}{}
+
+\begin{itemize}
+\tightlist
+\item
+  Let reaction norm evolve, under different initial conditions.
+\item
+  Let learning parmeters evolve
+\item
+  Let the communication system co-evolve
+\end{itemize}
+
+\end{frame}
+
 \end{document}

### 2019-09-06 compPar.R

@@ -8,7 +8,9 @@ source(here("AccFunc.R"))
 
 # Scenario to be plotted - corresponds to folders where simulations are stored
 
-scenario<-"learHonest_/QualStDv"
+scenario<-"learHonest_/alphaAct"
+
+shortSce<-tail(strsplit(scenario,split = "_/")[[1]],1)
 
 # Load files -------------------------------------------------------------------
 
@@ -16,14 +18,14 @@ scenario<-"learHonest_/QualStDv"
 (sdList<-grep("evol",listTest,value=TRUE))
 (indList<-grep("ind",listTest,value=TRUE))
 
-evol<-do.call(rbind,lapply(sdList,filesScenar,scenario))
 inds<-do.call(rbind,lapply(indList,filesScenar,scenario))
 
+inds[,diffActWeight:=abs(WeightAct_0-WeightAct_4)]
 
 # Graphs -----------------------------------------------------------------------
 
-gener<-tail(pop[,unique(time)],1)
-lastInt<-tail(pop[,unique(nInteract)],3)
+gener<-tail(inds[,unique(time)],1)
+lastInt<-tail(inds[,unique(nInteract)],3)
 runChoi<-0
 nCenters<-5
 interv<-1/nCenters
@@ -31,9 +33,10 @@ centers<-interv*0.5+interv*seq(0,nCenters-1)
 rangx<-seq(0,1,length=1000)
 tempPop<-inds[time==gener&nInteract==lastInt[1],.SD[.N],
              .SDcol=c(grep("Weight",
-                           names(evol),value = TRUE),"Quality","QualStDv"),
+                           names(evol),value = TRUE),"Quality",
+                      as.character(shortSce),"diffActWeight"),
              by=indId]
-dataIndAct_0.1<-sapply(as.list(tempPop[QualStDv==0.1,indId]),
+dataIndAct_1<-sapply(as.list(tempPop[get(shortSce)==unique(get(shortSce)[1]),indId]),
                    function(x){x=
                      logist(totRBF(rangx,
                                    centers,0.01,
@@ -44,7 +47,7 @@ dataIndAct_0.1<-sapply(as.list(tempPop[QualStDv==0.1,indId]),
                                                          value = TRUE)
                                              ])),alpha=0,beta = 1)})
 
[…]

### 2019-09-06 popSample_1.R

@@ -11,7 +11,7 @@ source(here("AccFunc.R"))
 
 # Scenario to be plotted - corresponds to folders where simulations are stored
 
-scenario<-"learHonest_/QualStDv"
+scenario<-"learHonest_/strQual1"
 
 
 # Load files -------------------------------------------------------------------

### 2020-06-24 QD_report.Rmd
Writing the first version of the manuscript

--- 2019-07-02 QD_report.Rmd
@@
 ---
+
+<!-- # ```{r loadData, echo=FALSE,include=FALSE} -->
+<!-- # Required libraries ----------------------------------------------------------- -->
+
+<!-- library(here) -->
+<!-- here() -->
+<!-- source(here("AccFunc.R")) -->
+<!-- # Scenario to be plotted - corresponds to folders where simulations are stored -->
+<!-- scenario<-"mutType" -->
+<!-- (listTest<-list.files(here("Simulations",paste0(scenario,"_")))) -->
+<!-- (evolList<-grep(".txt",grep("evol",listTest,value=TRUE),value=TRUE)) -->
+<!-- (indList<-grep("ind",listTest,value=TRUE)) -->
+
+<!-- fileId<-1 -->
+<!-- evol<-fread(here("Simulations",paste0(scenario,"_"),evolList[fileId])) -->
+<!-- pop<-fread(here("Simulations",paste0(scenario,"_"),indList[fileId])) -->
+<!-- evolStats<-evol[,.(m.freqGenHawk=mean(freqGenHawks), -->
+<!--                    upIQR.freqGenHawk=fivenum(freqGenHawks)[4], -->
+<!--                    lowIQR.freqGenHawk=fivenum(freqGenHawks)[2], -->
+<!--                    m.freqGenDove=mean(freqGenDove), -->
+<!--                    upIQR.freqGenDove=fivenum(freqGenDove)[4], -->
+<!--                    lowIQR.freqGenDove=fivenum(freqGenDove)[2], -->
+<!--                    m.freqEval=mean(freqGenEval), -->
+<!--                    upIQR.freqGenEval=fivenum(freqGenEval)[4], -->
+<!--                    lowIQR.freqGenEval=fivenum(freqGenEval)[2],  -->
+<!--                    m.freqFenHawk=mean(freqFenHawks), -->
+<!--                    upIQR.freqFenHawk=fivenum(freqFenHawks)[4], -->
+<!--                    lowIQR.freqFenHawk=fivenum(freqFenHawks)[2], -->
+<!--                    m.freqFenDove=mean(freqFenDoves), -->
+<!--                    upIQR.freqFenDove=fivenum(freqFenDoves)[4], -->
+<!--                    lowIQR.freqFenDove=fivenum(freqFenDoves)[2], -->
+<!--                    m.meanAlpha=mean(meanAlpha), -->
+<!--                    upIQR.alpha=fivenum(meanAlpha)[4], -->
+<!--                    lowIQR.alpha=fivenum(meanAlpha)[2], -->
+<!--                    m.meanBeta=mean(meanBeta), -->
+<!--                    upIQR.beta=fivenum(meanBeta)[4], -->
+<!--                    lowIQR.beta=fivenum(meanBeta)[2], -->
+<!--                    m.freqHH = mean(freqHH), -->
+<!--                    m.freqHD = mean(freqHD), -->
+<!--                    m.freqDD = mean(freqDD), -->
+<!--                    upIQR.freqHH = fivenum(freqHH)[4], -->
+<!--                    upIQR.freqHD = fivenum(freqHD)[4], -->
+<!--                    upIQR.freqDD = fivenum(freqDD)[4], -->
+<!--                    lowIQR.freqHH = fivenum(freqHH)[4], -->
+<!--                    lowIQR.freqHD = fivenum(freqHD)[2], -->
+<!--                    lowIQR.freqDD = fivenum(freqDD)[2], -->
+<!--                    m.weightAct_0=mean(WeightAct_0), -->
+<!--                    m.weightAct_1=mean(WeightAct_1), -->
+<!--                    m.weightAct_2=mean(WeightAct_2), -->
+<!--                    m.weightAct_3=mean(WeightAct_3), -->
+<!--                    m.weightAct_4=mean(WeightAct_4), -->
+<!--                    m.weightCrit_0=mean(WeightCrit_0), -->
+<!--                    m.weightCrit_1=mean(WeightCrit_1), -->
+<!--                    m.weightCrit_2=mean(WeightCrit_2), -->
+<!--                    m.weightCrit_3=mean(WeightCrit_3), -->
+<!--                    m.weightCrit_4=mean(WeightCrit_4)),by=time] -->
+<!-- # Extract means and IQR for the dynamic variables ------------------------------ -->
+
+<!-- # popStats<-pop[, as.list(unlist(lapply(.SD, -->
+<!-- #                                       function(x) list(m = mean(x), -->
+<!-- #                                                        upIQR = fivenum(x)[4], -->
+<!-- #                                                        downIQR = fivenum(x)[2] -->
+<!-- #                                       )))), -->
+<!-- #               by = time, -->
+<!-- #               .SDcols=c("Quality","alpha","beta","Badge","nInteract","WeightAct_0", -->
+<!-- #                         "WeightAct_1","WeightAct_2","WeightAct_3","WeightAct_4", -->
+<!-- #                         "WeightCrit_0","WeightCrit_1","WeightCrit_2", -->
+<!-- #                         "WeightCrit_3","WeightCrit_4")] -->
+
+<!-- # knitr::include_graphics("Simulations/mutType_/basicHawkDove.png") -->
+<!-- ``` -->
 
@@
 
-```{r fig1, echo=FALSE, fig.cap="\\label{fig:HD_game}Hawk-dove game. Dashed line is the  game theoretical prediction for frequency of hawks.", out.width = '80%'}
-knitr::include_graphics("Simulations/mutType_/basicHawkDove.png")
-```
+<!-- ```{r fig1, echo=FALSE, fig.cap="\\label{fig:HD_game}Hawk-dove game. Dashed line is the  game theoretical prediction for frequency of hawks.", out.width = '80%'} -->
+<!-- knitr::include_graphics("Simulations/mutType_/basicHawkDove.png") -->
+<!-- ``` -->
 
@@
 
-```{r fig2, echo=FALSE, fig.cap="\\label{fig:learning_cartoonRBF}Function approximation for the actor and the critic.
+<!-- # ```{r fig2, echo=FALSE, fig.cap="\\label{fig:learning_cartoonRBF}Function approximation for the actor and the critic.
 Using random values for the responses of each center.
-Circles show the location of the centers and the maximum response they trigger, while the lines show the total response along the badge size continium.", out.width = '80%'}
-knitr::include_graphics("cartoonRBF.png")
-```
+Circles show the location of the centers and the maximum response they trigger, while the lines show the total response along the badge size continium.", out.width = '80%'} -->
+<!-- # knitr::include_graphics("cartoonRBF.png") -->
+<!-- ``` -->
 
@@
 
-```{r fig3, echo=FALSE, fig.cap="\\label{fig:learning_invar}Learning equilibrium when individuals do not vary in quality.
-Dashed line is the  game theoretical prediction for frequency of hawks.", out.width = '80%'}
-knitr::include_graphics("Simulations//QualStDv_//hawkDoveLearn.png")
-```
+<!-- ```{r fig3, echo=FALSE, fig.cap="\\label{fig:learning_invar}Learning equilibrium when individuals do not vary in quality.
+Dashed line is the  game theoretical prediction for frequency of hawks.", out.width = '80%'} -->
+<!-- knitr::include_graphics("Simulations//QualStDv_//hawkDoveLearn.png") -->
+<!-- ``` -->
 
@@
 
-```{r fig4, echo=FALSE, fig.cap="\\label{fig:learning_var0.1}Learning equilibrium when individuals vary in quality.
-Dashed lines are the  game theoretical predictions in the standard game.", out.width = '80%'}
-knitr::include_graphics("Simulations//QualStDv_//hawkDoveLearn_0.1.png")
-```
+<!-- ```{r fig4, echo=FALSE, fig.cap="\\label{fig:learning_var0.1}Learning equilibrium when individuals vary in quality.
+Dashed lines are the  game theoretical predictions in the standard game.", out.width = '80%'} -->
+<!-- knitr::include_graphics("Simulations//QualStDv_//hawkDoveLearn_0.1.png") -->
+<!-- ``` -->
 
@@
 
-```{r fig5, echo=FALSE, fig.cap="\\label{fig:learning_all_var}Effect of variation in quality for the learning equilibrium.", out.width = '80%'}
-knitr::include_graphics("Simulations//QualStDv_//effectQualVariance.png")
-```
+<!-- ```{r fig5, echo=FALSE, fig.cap="\\label{fig:learning_all_var}Effect of variation in quality for the learning equilibrium.", out.width = '80%'} -->
+<!-- knitr::include_graphics("Simulations//QualStDv_//effectQualVariance.png") -->
+<!-- ``` -->
 
@@
 
-```{r fig6, echo=FALSE, fig.cap="\\label{fig:AC_HD}Within population variation in the learning parameters at the end of learning for three scenarios of variation in the quality of individuals.", out.width = '70%'}
-knitr::include_graphics("Simulations//QualStDv_//WeightsVarQualSt.png")
-```
+<!-- ```{r fig6, echo=FALSE, fig.cap="\\label{fig:AC_HD}Within population variation in the learning parameters at the end of learning for three scenarios of variation in the quality of individuals.", out.width = '70%'} -->
+<!-- knitr::include_graphics("Simulations//QualStDv_//WeightsVarQualSt.png") -->
+<!-- ``` -->
 
@@
 
-```{r fig7, echo=FALSE, fig.cap="\\label{fig:learn_dyn}Learning dynamics of the individuals in a population.", out.width = '80%'}
-knitr::include_graphics("Simulations//mutType_//learnDyn200.png")
-```
+<!-- ```{r fig7, echo=FALSE, fig.cap="\\label{fig:learn_dyn}Learning dynamics of the individuals in a population.", out.width = '80%'} -->
+<!-- knitr::include_graphics("Simulations//mutType_//learnDyn200.png") -->
+<!-- ``` -->
 
-# How do learners fare against the "pure" strategies
+# How do learners fare against the "pure" strategies?
 
@@
 
-```{r fig8, echo=FALSE, fig.cap="\\label{fig:types_comp}Evolutionary dynamics of the learners competing against pure strategies.", out.width = '80%'}
-knitr::include_graphics("Simulations//mutType_//evolDynTypes.png")
-```
+<!-- ```{r fig8, echo=FALSE, fig.cap="\\label{fig:types_comp}Evolutionary dynamics of the learners competing against pure strategies.", out.width = '80%'} -->
+<!-- knitr::include_graphics("Simulations//mutType_//evolDynTypes.png") -->
+<!-- ``` -->
+
+# How do learners behave when signals are honest?
+
+<!-- ```{r fig9, echo=FALSE,fig.cap="\\label{fig:honest_learn}Actor and critic developed when signals are honest.
+Left panel: honest signal.
+Right panel: learners behavioral response ",out.width='100%',fig.show='hold'} -->
+<!-- knitr::include_graphics("Simulations//learHonest_//QualStDv_//weightsVarQualSt.png") -->
+
+<!-- ``` -->
 

### 2020-06-24 QD_report.tex
Writing the first version of the manuscript

@@
-\documentclass[]{article}
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+%
+\documentclass[
+]{article}
 \usepackage{lmodern}
@@
 \usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
 \ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
@@
   \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-  \else
-    \usepackage{fontspec}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
 \fi
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
-\fi
-% use upquote if available, for straight quotes in verbatim environments
+% Use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage{microtype}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
   \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
 }{}
-\usepackage[margin=1in]{geometry}
-\usepackage{hyperref}
-\hypersetup{unicode=true,
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\hypersetup{
   pdftitle={The evolution of badges of status with learners},
   pdfauthor={Andrés Quiñones},
-            pdfborder={0 0 0},
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
+\usepackage[margin=1in]{geometry}
 \usepackage{graphicx,grffile}
@@
 \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-\IfFileExists{parskip.sty}{%
-\usepackage{parskip}
-}{% else
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
 \setlength{\emergencystretch}{3em} % prevent overfull lines
@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{0}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-
-%%% Use protect on footnotes to avoid problems with footnotes in titles
-\let\rmarkdownfootnote\footnote%
-\def\footnote{\protect\rmarkdownfootnote}
-
-%%% Change title format to be more compact
-\usepackage{titling}
-
-% Create subtitle command for use in maketitle
-\providecommand{\subtitle}[1]{
-  \posttitle{
-    \begin{center}\large#1\end{center}
-    }
-}
-
-\setlength{\droptitle}{-2em}
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+\usepackage{float}
 
 \title{The evolution of badges of status with learners}
-    \pretitle{\vspace{\droptitle}\centering\huge}
-  \posttitle{\par}
 \author{Andrés Quiñones}
-    \preauthor{\centering\large\emph}
-  \postauthor{\par}
 \date{}
-    \predate{}\postdate{}
-
-\usepackage{float}
 
@@
 which fit the game theoretical prediction.
-
-\begin{figure}
-\includegraphics[width=0.8\linewidth]{Simulations/mutType_/basicHawkDove} \caption{\label{fig:HD_game}Hawk-dove game.
-Dashed line is the  game theoretical prediction for frequency of hawks.}\label{fig:fig1}
-\end{figure}
 
@@
 
-\begin{figure}
-\includegraphics[width=0.8\linewidth]{cartoonRBF} \caption{\label{fig:learning_cartoonRBF}Function approximation for the actor and the critic.
-Using random values for the responses of each center.
-Circles show the location of the centers and the maximum response they trigger, while the lines show the total response along the badge size continium.}\label{fig:fig2}
-\end{figure}
-
 \hypertarget{results-when-individuals-do-not-vary-in-their-quality}{%
@@
 
-\begin{figure}
-\includegraphics[width=0.8\linewidth]{Simulations//QualStDv_//hawkDoveLearn} \caption{\label{fig:learning_invar}Learning equilibrium when individuals do not vary in quality.
-Dashed line is the  game theoretical prediction for frequency of hawks.}\label{fig:fig3}
-\end{figure}
-
 \hypertarget{results-when-individuals-vary-in-their-quality}{%
@@
 
-\begin{figure}
-\includegraphics[width=0.8\linewidth]{Simulations//QualStDv_//hawkDoveLearn_0.1} \caption{\label{fig:learning_var0.1}Learning equilibrium when individuals vary in quality.
-Dashed lines are the  game theoretical predictions in the standard game.}\label{fig:fig4}
-\end{figure}
-
 \hypertarget{overall-effect-of-variance}{%
@@
 individuals in a population expressing the hawk phenotype.
-
-\begin{figure}
-\includegraphics[width=0.8\linewidth]{Simulations//QualStDv_//effectQualVariance} \caption{\label{fig:learning_all_var}Effect of variation in quality for the learning equilibrium.}\label{fig:fig5}
-\end{figure}
 
@@
 
-\begin{figure}
-\includegraphics[width=0.7\linewidth]{Simulations//QualStDv_//WeightsVarQualSt} \caption{\label{fig:AC_HD}Within population variation in the learning parameters at the end of learning for three scenarios of variation in the quality of individuals.}\label{fig:fig6}
-\end{figure}
-
 \hypertarget{how-do-the-learning-dynamics-look-like}{%
@@
 other.
-
-\begin{figure}
-\includegraphics[width=0.8\linewidth]{Simulations//mutType_//learnDyn200} \caption{\label{fig:learn_dyn}Learning dynamics of the individuals in a population.}\label{fig:fig7}
-\end{figure}
 
@@
 
-\begin{figure}
-\includegraphics[width=0.8\linewidth]{Simulations//mutType_//evolDynTypes} \caption{\label{fig:types_comp}Evolutionary dynamics of the learners competing against pure strategies.}\label{fig:fig8}
-\end{figure}
-
 \hypertarget{how-do-learners-behave-when-signals-are-honest}{%
@@
 honest?}\label{how-do-learners-behave-when-signals-are-honest}}
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{Simulations//learHonest_//QualStDv_//weightsVarQualSt} \caption{\label{fig:honest_learn}Actor and critic developed when signals are honest.
-Left panel: honest signal.
-Right panel: learners behavioral response }\label{fig:fig9}
-\end{figure}
 

### 2020-06-24 manuscript_1.0.Rmd
Writing the first version of the manuscript

---
title: |
  | Learning can mediate both badges of status and individual recognition
author:
  - name: Andrés E. Quiñones
    affiliation: "1"
  - name: C. Daniel Cadena
    affiliation: "1"
  - name: Olof Leimar
    affiliation: "2"
  - name: Redouan Bshary
    affiliation: "3"
address:
  - code: "1"
    address: Laboratorio de Biología Evolutiva de Vertebrados, Departamento de Ciencias Biológicas, Universidad de los Andes, Bogotá, Colombia
  - code: "2"
    address: Department of Zoology, Stockholm University, Stockholm, Sweden
  - code: "3"
    address: Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland
date: "`r format(Sys.time(), '%B %d, %Y')`"
bibliography: SocialMorphCue.bib
header-includes: 
  - \usepackage{float}
  - \usepackage{setspace}\doublespacing
  - \linespread{2}
  - \usepackage{lineno}
  - \linenumbers
spacing: double
abstract: My abstract which has to be long enough to take multiple
  lines otherwise one does not see the effect of single-spacing.
keywords: "My keywords"
fig_caption: yes
keep_tex: yes
fontsize: 12pt
linespread: 2
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

For social animals the outcome of interactions with conspecifics is an important 
determinant of fitness. Irrespective of whether such interactions are of cooperative or competitive nature, individual's actions will partly determine their success in the reproductive arena. However, the best action for an invidual might vary depending on it's 
own conditions, and on who they are interacting with. Thus, an individual making their action dependent on information about interacting partners is typically adaptive [@quinones_Negotiation_2016].The source of this information however is far from trivial. In some cases interacting
partners (signallers) might be 'willing' to provide accurate information, but in 
others it might be in their own interest to conceal information [@johnstone_Recognition_1997],
or to provide erroneous information. Typically, in a given context, some proportion of 
individuals will be interested in broadcasting accurate information, and some others 
in hiding it. Take for instance an interaction between two individuals where one can help the other. Given some conditions, the donor will be interested in helping a related individual. So for a relative of the donor, broadcasting kinship is advantageous; while for an unrelated
individual, conceling the lack of kinship will be better. This same pattern applies to 
many aspects social life such as finding mates, feeding offspring, dominance relationships and aggressive interactions. In dominance and aggressive interactions an important 
information to make actions dependent on is interacting partners fighting ability, 
which is sometimes refered to as Resource Holding Potential (RHP) [@parker_Assesment_1974] , or simply quality. Stable communication systems can evolve where a signal conveys quality can be 
evolutionarily stable whenever the signal imposes a fitness cost which increases with the quality of an individual

## References

### 2020-06-24 learnDynamics.R

@@ -11,22 +11,22 @@ source(here("AccFunc.R"))
 
 # Scenario to be plotted - corresponds to folders where simulations are stored
 
-scenario<-"nIntGroupEvolLear"
+scenario<-"test"
 extSimsDir<-paste0("e:/BadgeSims/",scenario,"_")
 
 
 # Load files -------------------------------------------------------------------
 
 (listTest<-list.files(here("Simulations",paste0(scenario,"_"))))
-(listTest<-list.files(extSimsDir,full.names = TRUE))
+# (listTest<-list.files(extSimsDir,full.names = TRUE))
 
 
-(List<-grep("ind",listTest,value=TRUE))
+(indList<-grep("ind",listTest,value=TRUE))
 
-fileId<-2
+fileId<-1
 
-# indLearn<-fread(here("Simulations",paste0(scenario,"_"),indList[fileId]))
-indLearn<-fread(List[fileId])
+indLearn<-fread(here("Simulations",paste0(scenario,"_"),indList[fileId]))
+# indLearn<-fread(indList[fileId])
 
 # new columns ------------------------------------------------------------------
 
@@ -36,10 +36,31 @@ indLearn[,`:=`(freqHH=nint_HH/ntotInteract,
                freqHD=(ntotInteract-nint_HH-nint_DD)/ntotInteract,
                freqDD=nint_DD/ntotInteract)]
 
-length(indLearn[seed==8&time==0,unique(indId)])
 
-indLearn[(seed==1&time==0)&indId==833]
+tempPop[nInteract>1500,indId]
+par(plt=posPlot())
 
+indLearn[seed==8,unique(indId)]
+
+maxInt<-indLearn[,max(nInteract),by=.(seed,indId,time)]
+
+hist(maxInt[,V1])
+
+maxInt[V1>2000,length(indId),by=.(time,seed)]
+
+indLearn[,unique(indId-rivalId)]
+indLearn[10<indId & indId<90,unique(indId-rivalId)]
+
+temphist<-hist(indLearn[10<indId & indId<90,abs(indId-rivalId)],breaks = 80)
+temphist<-hist(indLearn[,abs(indId-rivalId)],breaks = 80)
+
+indLearn[,dist:=abs(indId-rivalId)]
+
+weirddata<-indLearn[dist>10&dist<90,]
+
+hist(weirddata[,dist])
+plot(indLearn[,.(indId,rivalId)])
[…]

### 2020-07-15 manuscript_1.0.Rmd
Found the bug on the implementation. The random number generator (sander's) is not thread safe. The problem is solved only for the interactions function. Must either adapt the RNG or change the engine.

@@
 author:
-  - name: Andrés E. Quiñones
-    affiliation: "1"
-  - name: C. Daniel Cadena
-    affiliation: "1"
-  - name: Olof Leimar
-    affiliation: "2"
-  - name: Redouan Bshary
-    affiliation: "3"
-address:
-  - code: "1"
-    address: Laboratorio de Biología Evolutiva de Vertebrados, Departamento de Ciencias Biológicas, Universidad de los Andes, Bogotá, Colombia
-  - code: "2"
-    address: Department of Zoology, Stockholm University, Stockholm, Sweden
-  - code: "3"
-    address: Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland
+  - Andrés E. Quiñones^[Laboratorio de Biología Evolutiva de Vertebrados, Departamento de Ciencias Biológicas, Universidad de los Andes, Bogotá, Colombia]
+  - C. Daniel Cadena^[Laboratorio de Biología Evolutiva de Vertebrados, Departamento de Ciencias Biológicas, Universidad de los Andes, Bogotá, Colombia]
+  - Olof Leimar^[Department of Zoology, Stockholm University, Stockholm, Sweden]
+  - Redouan Bshary^[Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland]
 date: "`r format(Sys.time(), '%B %d, %Y')`"
 bibliography: SocialMorphCue.bib
+style: C:\Users\andre\Zotero\styles\elife.csl
 header-includes: 
@@
 
-For social animals the outcome of interactions with conspecifics is an important 
-determinant of fitness.
-Irrespective of whether such interactions are of cooperative or competitive nature, individual's actions will partly determine their success in the reproductive arena.
-However, the best action for an invidual might vary depending on it's
-own conditions, and on who they are interacting with.
-Thus, an individual making their action dependent on information about interacting partners is typically adaptive [@quinones_Negotiation_2016].The source of this information however is far from trivial.
-In some cases interacting
-partners (signallers) might be 'willing' to provide accurate information, but in 
-others it might be in their own interest to conceal information [@johnstone_Recognition_1997],
-or to provide erroneous information. Typically, in a given context, some proportion of 
-individuals will be interested in broadcasting accurate information, and some others 
-in hiding it.
+For social animals the outcome of interactions with con-specifics is an important 
+determinant of fitness. Irrespective of whether such interactions are of cooperative 
+or competitive nature, individual's actions and their partners will partly determine 
+their success in the reproductive arena. However, the best action for an individual 
+might vary depending on it's own conditions, and with whom they interact. 
+Thus, an individual making their action dependent on information about interacting 
+partners is typically adaptive [@quinones_Negotiation_2016].
+The source of this information however is far from trivial. 
+In some cases interacting partners (signalers) might be 'willing' to provide accurate
+information, but in others it might be in their own interest to conceal information
+[@johnstone_Recognition_1997], or to provide erroneous information [@johnstone_Badges_1993]. 
+Typically, in a given context, some proportion of individuals will be interested 
+in broadcasting accurate information, and some others in hiding it. 
 Take for instance an interaction between two individuals where one can help the other. 
-Given some conditions, the donor will be interested in helping a related individual.
-So for a relative of the donor, broadcasting kinship is advantageous; while for an unrelated
-individual, conceling the lack of kinship will be better. This same pattern applies to 
-many aspects social life such as finding mates, feeding offspring, dominance relationships and aggressive interactions. In dominance and aggressive interactions an important 
+Given some conditions of cost and benefits, the donor will be interested in helping 
+a related individual. So for a relative of the donor, 
+broadcasting kinship is advantageous; while for an unrelated individual, 
+concealing the lack of kinship will be better. This same pattern applies to 
+many aspects of social life such as finding mates, feeding offspring, dominance relationships 
+and aggressive interactions. In dominance and aggressive interactions an important 
 information to make actions dependent on is interacting partners fighting ability, 
-which is sometimes refered to as Resource Holding Potential (RHP) [@parker_Assesment_1974] , or simply quality. Stable communication systems can evolve where a signal conveys quality can be 
-evolutionarily stable whenever the signal imposes a fitness cost which increases with the quality of an individual
+which is sometimes referred to as Resource Holding Potential (RHP) [@parker_Assessment_1974], 
+or as we will here simply quality. This responsiveness to the quality of the partner is achieved in
+communication systems such as badges of status (BS), where an arbitrary signal conveys quality. 
+BS can be evolutionarily stable whenever the signal imposes a fitness cost which 
+decreases with the quality of an individual  [@botero_Evolution_2010; @johnstone_Badges_1993]. 
+So, low quality individuals can't afford the cost of the signal, 
+and it's no longer in their interest to fake quality. 
+The cost of the honest signal is a more general principle of communications systems, 
+usually referred to as the handicap principle [@grafen_Biological_1990; @zahavi_Mate_1975]. 
+An alternative system of communication is one where individuals gather and store information
+about each individual they interact with. This information, directly collected by 
+each individual, should reliably allow them to react adaptively to the 
+quality difference with their opponent. This type of assessment strategy is often referred to as 
+individual recognition [IR, @whitfield_Plumage_1986]. Evidently, IR is restricted to 
+certain interaction structures, individuals are limited by cognitive adaptations in how many
+individuals they can recognize and track. Thus, IR and BS have been assumed to be alternative
+communication systems fulfilling the same function, but expected to evolve in populations 
+under different interaction structures [@dale_Signaling_2001; @sheehan_There_2016].
 

### 2020-07-15 cue_compt.cpp

@@ -519,12 +519,14 @@ void printLearnDynamics(ofstream &genoutput, vector<individual> &pop,
 
 void interactions(vector<individual>& population, int nint,
 	vector<double> payoff_matrix, double strQual, bool trackPopLearn,
-	int printLearnInt, int sampleSize, int generat, int seed, int nIntGroup,   
-	printingObj &localPrint) {
+	int printLearnInt, int sampleSize, int generat, int nIntGroup,   
+	printingObj &localPrint, std::mt19937 rngT) {
 	int ind1 = population.size();
 	int ind2 = population.size();
 	int intType = 0;
 	bool ind1win = 0;
+	std::uniform_int_distribution<int> randInd(0,population.size()-1);
+	std::uniform_int_distribution<int> randPeer(0,nIntGroup-1);
 	vector<int> sample;
 	vector<int>::iterator itSamp1 = sample.begin();
 	//cout << "track " << generat << endl;
@@ -535,14 +537,17 @@ void interactions(vector<individual>& population, int nint,
 		/*localPrint.countIntTypesGen[0];
 		localPrint.countIntTypesGen[1], localPrint.countIntTypesGen[2] = 0;*/
 		for (int countSam = 0; countSam < sampleSize; ++countSam) {
-			sample.emplace_back(rnd::integer(population.size()));
+			sample.emplace_back(randInd(rngT));
 			localPrint.sampledInd[countSam] = sample[countSam];
 			localPrint.counterRecords[countSam] = 0;
 		}
 	}
 	for (int i = 0; i < nint*population.size(); ++i) {
-		ind1 = rnd::integer(population.size());
-		ind2 = ind1 + 1 + rnd::integer(nIntGroup);
+		ind1 = randInd(rngT);
+		ind2 = ind1 + 1 + randPeer(rngT);
+		if (abs(ind2 - ind1) > nIntGroup) {
+			cout << "WTF!!!!!" << endl;
+		}
 		if (ind2 >= population.size()) ind2 = ind2 - population.size();
 		intType = 0;
 		intType += population[ind1].set_phenotype(population[ind2],localPrint);
@@ -634,27 +639,27 @@ void printStats(int popsize,ofstream &evolOutput,
 	//cout << endl;
 }
 
-void printPopSample(vector<individual> &population, ofstream &popOutput,
-	int time, int seed, int sampleSize, int nFeat = 5) {
-	int sample;
-	for (int countSample = 0; countSample < sampleSize; ++countSample) {
-		sample = rnd::integer(population.size());
[…]

### 2020-07-15 learnDynamics.R

@@ -23,7 +23,7 @@ extSimsDir<-paste0("e:/BadgeSims/",scenario,"_")
 
 (indList<-grep("ind",listTest,value=TRUE))
 
-fileId<-1
+fileId<-5
 
 indLearn<-fread(here("Simulations",paste0(scenario,"_"),indList[fileId]))
 # indLearn<-fread(indList[fileId])
@@ -37,21 +37,24 @@ indLearn[,`:=`(freqHH=nint_HH/ntotInteract,
                freqDD=nint_DD/ntotInteract)]
 
 
-tempPop[nInteract>1500,indId]
-par(plt=posPlot())
 
-indLearn[seed==8,unique(indId)]
+par(plt=posPlot())
 
 maxInt<-indLearn[,max(nInteract),by=.(seed,indId,time)]
 
 hist(maxInt[,V1])
 
-maxInt[V1>2000,length(indId),by=.(time,seed)]
 
 indLearn[,unique(indId-rivalId)]
+
 indLearn[10<indId & indId<90,unique(indId-rivalId)]
+indLearn[20<indId & indId<80,unique(indId-rivalId)]
+indLearn[30<indId & indId<70,unique(indId-rivalId)]
 
 temphist<-hist(indLearn[10<indId & indId<90,abs(indId-rivalId)],breaks = 80)
+temphist<-hist(indLearn[20<indId & indId<80,abs(indId-rivalId)],breaks = 80)
+temphist<-hist(indLearn[30<indId & indId<70,abs(indId-rivalId)],breaks = 80)
+
 temphist<-hist(indLearn[,abs(indId-rivalId)],breaks = 80)
 
 indLearn[,dist:=abs(indId-rivalId)]
@@ -59,9 +62,9 @@ indLearn[,dist:=abs(indId-rivalId)]
 weirddata<-indLearn[dist>10&dist<90,]
 
 hist(weirddata[,dist])
-plot(indLearn[,.(indId,rivalId)])
-points(weirddata[,.(indId,rivalId)],col=2,pch=20)
-
+plot(indLearn[,.(indId,rivalId)],col=indLearn[,seed]+1)
+points(weirddata[,.(indId,rivalId)],col=weirddata[,seed]+1,pch=20)
+legend("top",legend = weirddata[,unique(seed)],col = weirddata[,unique(seed)]+1,pch=20)
 # Changes in learning parameters for several individuals -----------------------
 
[…]

### 2020-08-20 EvolvertPresent.Rmd
Plots for the group meeting

--- 2019-09-06 QD_presentation.Rmd
@@
 ---
-title: "The evolution of badges of status with learners"
+title: "Badges of status and IR mediated by learning"
 author: "Andrés Quiñones"
 bibliography: SocialMorphCue.bib
+date: "`r format(Sys.time(), '%B %d, %Y')`"
 output: 
@@
     \usepackage{array}
-                #\floatplacement{figure}{H}
 ---
 
-```{r loadData, echo=FALSE,include=FALSE,warning=FALSE,message=FALSE}
-# Required libraries -----------------------------------------------------------
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = FALSE)
+source("../R_files/posPlots.R")
+require("here")
+source(here("AccFunc.R"))
 
-library(here)
-here()
-source(here("AccFunc.R"))
-# Scenario to be plotted - corresponds to folders where simulations are stored
-scenario<-"mutType"
-(listTest<-list.files(here("Simulations",paste0(scenario,"_"))))
-(evolList<-grep(".txt",grep("evol",listTest,value=TRUE),value=TRUE))
-(indList<-grep("ind",listTest,value=TRUE))
-
-fileId<-1
-evol<-fread(here("Simulations",paste0(scenario,"_"),evolList[fileId]))
-pop<-fread(here("Simulations",paste0(scenario,"_"),indList[fileId]))
-evolStats<-evol[,.(m.freqGenHawk=mean(freqGenHawks),
-                   upIQR.freqGenHawk=fivenum(freqGenHawks)[4],
-                   lowIQR.freqGenHawk=fivenum(freqGenHawks)[2],
-                   m.freqGenDove=mean(freqGenDove),
-                   upIQR.freqGenDove=fivenum(freqGenDove)[4],
-                   lowIQR.freqGenDove=fivenum(freqGenDove)[2],
-                   m.freqEval=mean(freqGenEval),
-                   upIQR.freqGenEval=fivenum(freqGenEval)[4],
-                   lowIQR.freqGenEval=fivenum(freqGenEval)[2], 
-                   m.freqFenHawk=mean(freqFenHawks),
-                   upIQR.freqFenHawk=fivenum(freqFenHawks)[4],
-                   lowIQR.freqFenHawk=fivenum(freqFenHawks)[2],
-                   m.freqFenDove=mean(freqFenDoves),
-                   upIQR.freqFenDove=fivenum(freqFenDoves)[4],
-                   lowIQR.freqFenDove=fivenum(freqFenDoves)[2],
-                   m.meanAlpha=mean(meanAlpha),
-                   upIQR.alpha=fivenum(meanAlpha)[4],
-                   lowIQR.alpha=fivenum(meanAlpha)[2],
-                   m.meanBeta=mean(meanBeta),
-                   upIQR.beta=fivenum(meanBeta)[4],
-                   lowIQR.beta=fivenum(meanBeta)[2],
-                   m.freqHH = mean(freqHH),
-                   m.freqHD = mean(freqHD),
-                   m.freqDD = mean(freqDD),
-                   upIQR.freqHH = fivenum(freqHH)[4],
-                   upIQR.freqHD = fivenum(freqHD)[4],
-                   upIQR.freqDD = fivenum(freqDD)[4],
-                   lowIQR.freqHH = fivenum(freqHH)[4],
-                   lowIQR.freqHD = fivenum(freqHD)[2],
-                   lowIQR.freqDD = fivenum(freqDD)[2],
-                   m.weightAct_0=mean(WeightAct_0),
-                   m.weightAct_1=mean(WeightAct_1),
-                   m.weightAct_2=mean(WeightAct_2),
-                   m.weightAct_3=mean(WeightAct_3),
-                   m.weightAct_4=mean(WeightAct_4),
-                   m.weightCrit_0=mean(WeightCrit_0),
-                   m.weightCrit_1=mean(WeightCrit_1),
-                   m.weightCrit_2=mean(WeightCrit_2),
-                   m.weightCrit_3=mean(WeightCrit_3),
-                   m.weightCrit_4=mean(WeightCrit_4)),by=time]
-# Extract means and IQR for the dynamic variables ------------------------------
-
-popStats<-pop[, as.list(unlist(lapply(.SD, 
-                                      function(x) list(m = mean(x),
-                                                       upIQR = fivenum(x)[4],
-                                                       downIQR = fivenum(x)[2]
-                                      )))),
-              by = time, 
-              .SDcols=c("Quality","alpha","beta","Badge","nInteract","WeightAct_0",
-                        "WeightAct_1","WeightAct_2","WeightAct_3","WeightAct_4",
-                        "WeightCrit_0","WeightCrit_1","WeightCrit_2",
-                        "WeightCrit_3","WeightCrit_4")]
-
-# knitr::include_graphics("Simulations/mutType_/basicHawkDove.png")
 ```
@@
 ```{r,echo=FALSE,fig.align="center",out.width='50%'}
-knitr::include_graphics("Images//HDgame.jpeg")
+knitr::include_graphics("Images//HDgame.png")
 ```
@@
 
-## The hawk-dove game
+<!-- ## The hawk-dove game -->
 
-```{r, echo=FALSE, fig.cap="\\label{fig:HD_game}Hawk-dove game. Dashed line is the  game theoretical prediction for frequency of hawks.", out.width = '80%'}
-knitr::include_graphics("Simulations/mutType_/basicHawkDove.png")
-```
+<!-- ```{r, echo=FALSE, fig.cap="\\label{fig:HD_game}Hawk-dove game. Dashed line is the  game theoretical prediction for frequency of hawks.", out.width = '80%'} -->
+<!-- knitr::include_graphics("Simulations/mutType_/basicHawkDove.png") -->
+<!-- ``` -->
 
@@
 Circles show the location of the centers and the maximum response they trigger, while the lines show the total response along the badge size continium.", out.width = '60%'}
-knitr::include_graphics("cartoonRBF.png")
+knitr::include_graphics("Images/cartoonRBF.png")
 ```
@@
 ```{r, echo=FALSE,fig.align='center', out.width = '60%'}
-knitr::include_graphics("cartoonRBF.png")
+knitr::include_graphics("Images/cartoonRBF.png")
 ```
@@
 
-## When individuals do _NOT_ vary in quality
+## Learning under honest signalling
 
-```{r , echo=FALSE, fig.align='center',out.width = '80%'}
-knitr::include_graphics("Simulations//QualStDv_//hawkDoveLearn.png")
+```{r, echo=FALSE,fig.align='center', out.width = '100%'}
+knitr::include_graphics("Simulations/alphaAct_/alphaAct0.6learnDyn.png")
 ```
 
-## When individuals _DO_ vary in quality
-
-```{r, echo=FALSE, fig.align='center', out.width = '80%'}
-knitr::include_graphics("Simulations//QualStDv_//hawkDoveLearn_1.png")
+## Learning can mediate the evolution of Badges of Status
+### Variation among replicates
+```{r, echo=FALSE,fig.align='center', out.width='100%'}
+knitr::include_graphics("Simulations/betCostEvol1_/evolDynAll_betCost5.png")
 ```
 
-## Overal effect of quality variation
-
-```{r fig5, echo=FALSE, fig.align='center', out.width = '80%'}
-knitr::include_graphics("Simulations//QualStDv_//effectQualVariance.png")
+## Learning can mediate the evolution of Badges of Status
+### Individual replicates: Badges
+```{r, echo=FALSE,fig.align='center', out.width='100%'}
+knitr::include_graphics("Simulations/betCostEvol1_/evolDyn0_betCost5.png")
 ```
 
-## How does variation among _learners_ look like?
-
-```{r fig6, echo=FALSE, fig.align='center',out.width = '90%'}
-knitr::include_graphics("Simulations//QualStDv_//WeightsVarQualSt.png")
+## Learning can mediate the evolution of Badges of Status
+### Individual replicates: no badges
+```{r, echo=FALSE,fig.align='center', out.width='100%'}
+knitr::include_graphics("Simulations/betCostEvol1_/evolDyn5_betCost5.png")
 ```
 
-## How do the learning dynamics look like?
+## The effect of costs on Badges of Status
+### Behavioural interactions
 
-```{r fig7, echo=FALSE, out.width = '80%'}
-knitr::include_graphics("Simulations//mutType_//learnDyn200.png")
+```{r, echo=FALSE,fig.align='center', out.width='100%'}
+knitr::include_graphics("Simulations/betCostEvol1_/BehavIntALL.png")
 ```
-
-## How do learners behave when signals are honest?
-
-```{r fig9, echo=FALSE,out.width='90%',fig.show='hold'}
-knitr::include_graphics("Simulations//learHonest_//QualStDv_//weightsVarQualSt.png")
-```
-
-## How do learners fare against pure strategies?
-
-```{r fig8, echo=FALSE, fig.align='center', out.width = '80%'}
-knitr::include_graphics("Simulations//mutType_//evolDynTypes.png")
-```
-
-## What's next?
-
-* Let reaction norm evolve, under different initial conditions.
-* Let learning parmeters evolve
-* Let the communication system co-evolve

### 2020-08-20 EvolvertPresent.tex
Plots for the group meeting

--- 2019-09-06 QD_presentation.tex
@@
-\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
 \PassOptionsToPackage{hyphens}{url}
@@
 \beamertemplatenavigationsymbolsempty
-% Prevent slide breaks in the middle of a paragraph:
+% Prevent slide breaks in the middle of a paragraph
 \widowpenalties 1 10000
@@
   \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provides euro and other symbols
-\else % if luatex or xelatex
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
   \usepackage{unicode-math}
@@
 \fi
-% use upquote if available, for straight quotes in verbatim environments
+% Use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
@@
 \hypersetup{
-  pdftitle={The evolution of badges of status with learners},
+  pdftitle={Badges of status and IR mediated by learning},
   pdfauthor={Andrés Quiñones},
-  pdfborder={0 0 0},
-  breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
 \newif\ifbibliography
@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{-2}
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+\usepackage{float} \usepackage{multicol} \usepackage{amsmath} \usepackage{graphicx} \usepackage{array}
 
-% set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
-
-\usepackage{float} \usepackage{multicol} \usepackage{amsmath} \usepackage{graphicx}
-
-\title{The evolution of badges of status with learners}
+\title{Badges of status and IR mediated by learning}
 \author{Andrés Quiñones}
-\date{}
+\date{August 20, 2020}
 
@@
 \begin{center}\includegraphics[width=0.55\linewidth]{Images//Hawk_dove_plot} \end{center}
-
-\end{frame}
-
-\begin{frame}{The hawk-dove game}
-\protect\hypertarget{the-hawk-dove-game-2}{}
-
-\begin{figure}
-\includegraphics[width=0.8\linewidth]{Simulations/mutType_/basicHawkDove} \caption{\label{fig:HD_game}Hawk-dove game.
-Dashed line is the  game theoretical prediction for frequency of hawks.}\label{fig:unnamed-chunk-3}
-\end{figure}
 
@@
 
+\begin{equation*}
+  \Delta V_{t(s)}=\alpha \underbrace{(R_t-V_t)}_\text{prediction error}
+\end{equation*}
+
 \begin{figure}
-\includegraphics[width=0.8\linewidth]{cartoonRBF} \caption{\label{fig:learning_cartoonRBF}Function approximation for the actor and the critic.
+\includegraphics[width=0.6\linewidth]{Images/cartoonRBF} \caption{\label{fig:learning_cartoonRBF}Function approximation for the actor and the critic.
 Using random values for the responses of each center.
@@
 
+\hypertarget{our-model}{%
+\section{Our model}\label{our-model}}
+
+\begin{frame}{The hawk-dove game 2.0}
+\protect\hypertarget{the-hawk-dove-game-2.0}{}
+
+\begin{center}
+\begin{tabular}{ >{\centering\arraybackslash}p{1cm} | >{\centering\arraybackslash}p{4.5cm} | >{\centering\arraybackslash}p{4.5cm} }
+& H & D \\ \hline
+H & $p_w V\frac{-C}{2} + (1-p_w) \frac{-C}{2}$ & $V$ \\ \hline
+V & $0$ & $\frac{V}{2}$\\
+\end{tabular}
+\end{center}
+
+\begin{equation*}
+p_w=\frac{1}{1+e^{-\beta(Q_i-Q_j)}}
+\end{equation*}
+
+\begin{center}\includegraphics[width=0.5\linewidth]{EvolvertPresent_files/figure-beamer/unnamed-chunk-8-1} \end{center}
+
+\end{frame}
+
+\begin{frame}{Sender Code}
+\protect\hypertarget{sender-code}{}
+
+\begin{equation*}
+s_i = \frac{1}{1+e^{-(\epsilon_i+\gamma_iQ_i)}}
+\label{eq:react_norm}
+\end{equation*}
+
+\begin{center}\includegraphics[width=0.5\linewidth]{EvolvertPresent_files/figure-beamer/unnamed-chunk-9-1} \end{center}
+
+\end{frame}
+
+\begin{frame}{Receiver code}
+\protect\hypertarget{receiver-code}{}
+
+\begin{equation*}
+  \Delta V_{t(s)}=\alpha (R_t-V_t)
+\end{equation*}
+
+\begin{center}\includegraphics[width=0.6\linewidth]{Images/cartoonRBF} \end{center}
+
+\end{frame}
+
+\hypertarget{results}{%
+\section{Results}\label{results}}
+
+\begin{frame}{Learning under honest signalling}
+\protect\hypertarget{learning-under-honest-signalling}{}
+
+\begin{center}\includegraphics[width=1\linewidth]{Simulations/alphaAct_/alphaAct0.6learnDyn} \end{center}
+
+\end{frame}
+
+\begin{frame}{Learning can mediate the evolution of Badges of Status}
+\protect\hypertarget{learning-can-mediate-the-evolution-of-badges-of-status}{}
+
+\begin{block}{Variation among replicates}
+
+\begin{center}\includegraphics[width=1\linewidth]{Simulations/betCostEvol1_/evolDynAll_betCost5} \end{center}
+
+\end{block}
+
+\end{frame}
+
+\begin{frame}{Learning can mediate the evolution of Badges of Status}
+\protect\hypertarget{learning-can-mediate-the-evolution-of-badges-of-status-1}{}
+
+\begin{block}{Individual replicates: Badges}
+
+\begin{center}\includegraphics[width=1\linewidth]{Simulations/betCostEvol1_/evolDyn0_betCost5} \end{center}
+
+\end{block}
+
+\end{frame}
+
+\begin{frame}{Learning can mediate the evolution of Badges of Status}
+\protect\hypertarget{learning-can-mediate-the-evolution-of-badges-of-status-2}{}
+
+\begin{block}{Individual replicates: no badges}
+
+\begin{center}\includegraphics[width=1\linewidth]{Simulations/betCostEvol1_/evolDyn5_betCost5} \end{center}
+
+\end{block}
+
+\end{frame}
+
+\begin{frame}{The effect of costs on Badges of Status}
+\protect\hypertarget{the-effect-of-costs-on-badges-of-status}{}
+
+\begin{block}{Behavioural interactions}
+
+\begin{center}\includegraphics[width=1\linewidth]{Simulations/betCostEvol1_/BehavIntALL} \end{center}
+
+\end{block}
+
+\end{frame}
+
 \end{document}

### 2020-08-20 manuscript_1.0.Rmd
Plots for the group meeting

@@
 author:
-  - Andrés E. Quiñones^[Laboratorio de Biología Evolutiva de Vertebrados, Departamento de Ciencias Biológicas, Universidad de los Andes, Bogotá, Colombia]
-  - C. Daniel Cadena^[Laboratorio de Biología Evolutiva de Vertebrados, Departamento de Ciencias Biológicas, Universidad de los Andes, Bogotá, Colombia]
+- Andrés E. Quiñones^[Laboratorio de Biología Evolutiva de Vertebrados, Departamento
+  de Ciencias Biológicas, Universidad de los Andes, Bogotá, Colombia]
+- C. Daniel Cadena*
 - Olof Leimar^[Department of Zoology, Stockholm University, Stockholm, Sweden]
@@
 date: "`r format(Sys.time(), '%B %d, %Y')`"
-bibliography: SocialMorphCue.bib
-style: C:\Users\andre\Zotero\styles\elife.csl
+output:
+  word_document: default
+  pdf_document: default
+fig_caption: yes
+fontsize: 12pt
 header-includes:
@@
 - \linenumbers
+keep_tex: yes
+keywords: My keywords
+linespread: 2
+bibliography: SocialMorphCue.bib
 spacing: double
-abstract: My abstract which has to be long enough to take multiple
-  lines otherwise one does not see the effect of single-spacing.
-keywords: "My keywords"
-fig_caption: yes
-keep_tex: yes
-fontsize: 12pt
-linespread: 2
-output: pdf_document
+style: C:\Users\andre\Zotero\styles\elife.csl
+abstract: Learning rules.
 ---
@@
 which is sometimes referred to as Resource Holding Potential (RHP) [@parker_Assessment_1974], 
-or as we will here simply quality. This responsiveness to the quality of the partner is achieved in
-communication systems such as badges of status (BS), where an arbitrary signal conveys quality. 
+or as we will here simply quality. This responsiveness to the quality of the 
+partner is achieved in communication systems such as badges of status (BS), 
+where an arbitrary signal conveys quality. 
 BS can be evolutionarily stable whenever the signal imposes a fitness cost which 
@@
 each individual, should reliably allow them to react adaptively to the 
-quality difference with their opponent. This type of assessment strategy is often referred to as 
-individual recognition [IR, @whitfield_Plumage_1986]. Evidently, IR is restricted to 
-certain interaction structures, individuals are limited by cognitive adaptations in how many
-individuals they can recognize and track. Thus, IR and BS have been assumed to be alternative
-communication systems fulfilling the same function, but expected to evolve in populations 
-under different interaction structures [@dale_Signaling_2001; @sheehan_There_2016].
+quality difference with their opponent. This type of assessment strategy is often 
+referred to as individual recognition [IR, @whitfield_Plumage_1986]. Evidently, 
+IR is restricted to certain interaction structures, individuals are limited by cognitive 
+abilities in how many individuals they can recognize and track. Thus, IR and BS 
+have been assumed to be alternative communication systems fulfilling the same function, 
+but expected to evolve in populations under different conditions
+[@dale_Signaling_2001; @sheehan_There_2016]. There is no clear cut reason, however, 
+why both systems could not work together, in fact there is some evidence
+that they can [@chaine_Manipulating_2018].
+
+A somewhat ignored aspect of communication systems, in the context of aggressive 
+interactions, is the cognitive aspects of the receiver module. 
+BS are often assume to be cognitively less demanding than IR 
+[@sheehan_There_2016]. The idea is that using IR, individuals need to store identity 
+and associated information about each individual they interact with. All this 
+information requires more memory and therefore is more demanding. In contrast,
+the response using BS have been thought as an innate response 
+[@botero_Evolution_2010; @johnstone_Badges_1993], where individuals use their 
+own quality and the opponent's badge to determine whether to aggressively contest 
+a resource. An alternative view of BS is that individuals learn to react to the signal
+based on their experiences [@guilford_Receiver_1991]. In line with that view, both
+communication systems would require the receivers to learn an association between
+the signal and the fighting ability of the bearer. In BS, this association would be a 
+a simple monotonic increase of fighting ability with the size of the badge, 
+and would be reinforced by every interaction, so in  principle it would be 
+quicker to learn. While in IR, the association would 
+be different for every individual, depending on the interacting partners. 
+Learning could be a unifying cognitive mechanism behind both types of communication 
+systems. This idea of learning being a general response mechanism has not been 
+thoroughly explored, nor in the empirical or theoretical literature.
+
+Associative learning is a key cognitive mechanism that allows individuals to associate
+rewards with environmental stimuli and appropriate behavior [@staddon_Adaptive_2016]. 
+Since it was originally described, associative learning has been found in all major
+vertebrate taxon, and many invertebrates [@heyes_Simple_2012;
+@macphail_Brain_1982; @staddon_Adaptive_2016; @behrens_Associative_2008].
+Using these associations to respond to environmental changes is particularly adaptive in
+complex environments [@dridi_Environmental_2016]. Besides its wide taxonomic and
+ecological relevance, associative learning has shown to be a flexible cognitive mechanism
+that can be tuned to different roles depending on the ecological situation 
+[@enquist_Power_2016; @quinones_Reinforcement_2019]. Despite all this, associative 
+learning is not often included in evolutionary explanations 
+[@fawcett_Exposing_2013; @kamil_Optimal_1983; @mcauliffe_Psychology_2015]. 
+Computational models of evolution can play an important role to overcome the 
+lack of integration between learning and evolution. Reinforcement learning theory 
+encompases a series of computational methods inspired on the psychological and 
+neurological mechanisms of associative learning [@sutton_Reinforcement_2018].
+These set of algorithms allow the implementation of biologically realistic problems, 
+capturing the essence of learning processes [@frankenhuis_Enriching_2018; @quinones_Reinforcement_2019]. 
+Furthermore, these algorithms can be embebed in evolutionary simulations in order 
+to have theoretical predictions of the effect that learning can have in behavioral [@leimar_Learning_2019]
+and morphological evolution.
+
+In here we present an evolutionary model where individuals use associative learning 
+to develop a tendency to behave aggressively (or  peacefully), in the context of
+competition over resoureces, depending on the quantitative morphological trait 
+they perceive in their opponent (badge). Over evolutionary time individuals evolve
+the size of their badge, and whether it depends on their quality. Under this simple
+set up, individuals can use the badge both as a signal of quality (BS)
+or as a trait for IR. We use the model to asses under what conditions of interaction
+structure we expect different communication signals evolve.
 

### 2020-08-20 aesth.R

@@ -28,4 +28,6 @@ colboxes<- c('#31a354','#d7191c','#fdae61','#2b83ba',"black")
 
 colRuns<-c('#e41a1c','#377eb8','#4daf4a', '#984ea3','#ff7f00')
 
+colSeeds<-brewer.pal(8,name = "Set2")
+
 colorbreaksQual<-seq(0,1,length=100)

### 2020-08-20 dynSevFiles.R

@@ -9,7 +9,7 @@ library("plotrix")
 
 # Scenario to be plotted - corresponds to folders where simulations are stored
 
-scenario<-"initAct"
+scenario<-"betCostEvol1"
 variable<-gsub("[^[:alpha:]]",scenario,replacement = '')
 variable<-gsub("Evol",variable,replacement = '')
 extSimsDir<-paste0("e:/BadgeSims/",scenario,"_")
@@ -72,9 +72,9 @@ evolStats<-evol[,.(m.freqGenHawk=mean(freqGenHawks),
                    m.weightCrit_2=mean(WeightCrit_2),
                    m.weightCrit_3=mean(WeightCrit_3),
                    m.weightCrit_4=mean(WeightCrit_4)),
-                by=.(time,get(variable))]
+                by=.(time,get(nampar))]
 
-setnames(evolStats,"get",variable)
+setnames(evolStats,"get",nampar)
 
 timeChoi<-max(evolStats[,time])
 
@@ -82,17 +82,25 @@ timeChoi<-max(evolStats[,time])
 
 # png(here("Simulations",paste0(scenario,"_"),"effectQualVariance.png"),height = 800,width = 800)
 
+evol[time==timeChoi,]
+with(evol[time==timeChoi],{
+boxplot(freqFenHawks~get(nampar),
+       xlab='',ylab='',cex.axis=1.3,yaxt='s')
+points(x=posX,y=freqFenHawks,pch=21)
+  })
+
+
 
 par(plt=posPlot(numploty = 2,idploty = 2),xaxt="n",las=1)
 with(evolStats[time==timeChoi],{
-  plotCI(x=as.factor(get(variable)),y=m.freqFenHawk,
+  plotCI(x=as.factor(get(nampar)),y=m.freqFenHawk,
          ui = upIQR.freqFenHawk,li=lowIQR.freqFenHawk,
          pch=16,xlab='',ylab='',
          col=colTypesLin[1],
          sfrac=0.005,cex.axis=1.3,yaxt='s',ylim=c(0,0.8))
-  lines(x=c(0,max(get(variable))),y=c(0.66,0.66),col='grey')
+  lines(x=c(0,max(get(nampar))),y=c(0.66,0.66),col='grey')
   par(new=T)
-  plotCI(x=as.factor(get(variable)),y=m.freqFenDove,new=T,
[…]

### 2020-08-20 learnDynamics.R

@@ -11,64 +11,46 @@ source(here("AccFunc.R"))
 
 # Scenario to be plotted - corresponds to folders where simulations are stored
 
-scenario<-"test"
+scenario<-"alphaAct"
 extSimsDir<-paste0("e:/BadgeSims/",scenario,"_")
+
 
 
 # Load files -------------------------------------------------------------------
 
-(listTest<-list.files(here("Simulations",paste0(scenario,"_"))))
-# (listTest<-list.files(extSimsDir,full.names = TRUE))
-
-
+# (listTest<-list.files(here("Simulations",paste0(scenario,"_"))))
+(listTest<-list.files(extSimsDir,full.names = TRUE))
 (indList<-grep("ind",listTest,value=TRUE))
+# parameter values from project folder
+paramName<-list.files(here("Simulations",paste0(scenario,"_")))
+paramName<-grep(".json",paramName,value=TRUE)
+param<-fromJSON(here("Simulations",paste0(scenario,"_"),paramName[2]))
 
-fileId<-5
 
-indLearn<-fread(here("Simulations",paste0(scenario,"_"),indList[fileId]))
-# indLearn<-fread(indList[fileId])
+fileId<-6
+
+
+# indLearn<-fread(here("Simulations",paste0(scenario,"_"),indList[fileId]))
+indLearn<-fread(indList[fileId])
+
+# names(indLearn)[29:35]<-c("Quality", "genotype","alpha","beta","Badge","initCrit","initAct")
+# names(indLearn)[9:28]<-do.call(rbind,as.list(sapply(0:9,FUN = function(x){
+#   return(rbind(paste0("WeightAct_",x),paste0("WeightCrit_",x)))
+# })))
+
+Valpar<-gsub("[[:alpha:]]",gsub(".txt","",tail(strsplit(indList[fileId],"_")[[1]],1)),
+             replacement = "")
+nampar<-gsub("[^[:alpha:]]",gsub(".txt","",tail(strsplit(indList[fileId],"_")[[1]],1)),
+             replacement = "")
 
 # new columns ------------------------------------------------------------------
 
-indLearn[,diffActWeights:=abs(WeightAct_0-WeightAct_4)]
 
-indLearn[,`:=`(freqHH=nint_HH/ntotInteract,
-               freqHD=(ntotInteract-nint_HH-nint_DD)/ntotInteract,
-               freqDD=nint_DD/ntotInteract)]
-
-
-
-par(plt=posPlot())
-
[…]

### 2020-08-20 parTOjson.R

@@ -9,30 +9,30 @@ source(here("AccFunc.R"))
 fileName<-"parameters"
 here()
 
+param1<-fromJSON(here("Simulations","alphaAct_","parameters21.json"))
 
-param<-list(totGen   =10000,   nRep    = 16,
-            printGen = 1000,   printLearn = 1000,
-            printLearnInt = 500,
+param<-list(totGen   =5,   nRep    = 30,
+            printGen = 1,   printLearn = 1,
+            printLearnInt = 50,
             popSize  = 2000,  baselineFit = 2,   
             MutSd    = 0.3,
             nInt        = 1000,  init     = c(0,0,1),
             mutRate  = 0.001,  mutType  = 0,
             sampleSize = 50,   strQual  = 10,
-            alphaBad	 = 0,    betaBad	 = 0,
-            alphaCrit  = 0.1,  alphaAct = 0.1,
-            sigSq   	 = 0.01, nCenters = 6,
-            initCrit = 0,      initAct=0,
+            errorQual = 0,    alphaBad	 = 3,    betaBad	 = 6,
+            alphaCrit  = 0.5,  alphaAct = 0.5,
+            sigSq   	 = 0.001, nCenters = 10,
+            initCrit = 0,      initAct=0,   gamma = 0,
             QualStDv   = 1.1, betCost = 0,
-            alphCost	 = 3, mutLearn = TRUE,
+            alphCost	 = 3, mutLearn = FALSE,
             nIntGroup  = 2000,
             payoff_matrix = c(1.5,1,0,0.5),
             namParam = "nIntGroup",
-            rangParam = c(10,50,2000),
+            rangParam = c(4,6,8,10,20,50,1000),
             folderL=paste(here("Simulations"),"/",sep=""))
 
 
-
-apendScenar<-"EvolLear"
+apendScenar<-"LearnIR"
 param$folderL<-paste0(param$folderL,
        param$namParam,apendScenar,"_/")
 # runTime<-"360:00:00"# "10:00:00"# 
@@ -44,13 +44,13 @@ param$folderL<-paste0(param$folderL,
 # nodes<-10
 
 ## For the cluster Uniandes
-param$folder<-paste0("/hpcfs/home/a.quinones/BadgeStatus/",
-                    param$namParam,apendScenar,"_/")
+# param$folder<-paste0("/hpcfs/home/a.quinones/BadgeStatus/",
+#                     param$namParam,apendScenar,"_/")
 ## for the cluster unine
[…]

### 2020-10-01 EvolvertPresent.Rmd
Random changes to improve flow, and comments on the presentation for Olle to follow

@@
 ---
-title: "Badges of status and IR mediated by learning"
+title: "Badges of status and Individual Recognition mediated by learning"
 author: "Andrés Quiñones"
@@
 
+<!-- ## The need to choose a partner and an action -->
+
+<!-- ```{r,echo=FALSE,fig.align="center",out.width='40%',fig.show='hold'} -->
+<!-- knitr::include_graphics("Images/sexSelec.jpg") -->
+<!-- knitr::include_graphics("Images/begging.jpg") -->
+<!-- knitr::include_graphics("Images/broodparasite.jpg") -->
+<!-- knitr::include_graphics("Images/dominance.jpg") -->
+<!-- ``` -->
+
+<!-- ## Animal communication -->
+
+<!-- ```{r,echo=FALSE,fig.align="center",out.width='80%'} -->
+<!-- knitr::include_graphics("Images/monkey_mobile.jpeg") -->
+<!-- ``` -->
+
+<!-- ## The cost of aggression -->
+
+<!-- ```{r,echo=FALSE,fig.align="center",out.width='80%'} -->
+<!-- knitr::include_graphics("Images/harm.jpg") -->
+<!-- ``` -->
+
+<!-- ## What about signals? -->
+
+<!-- ### Badge of status -->
+
+<!-- ```{r,echo=FALSE,fig.align='center',out.height='80%'} -->
+<!-- knitr::include_graphics("Images//hosp_alexandra_mackenzie.jpg") -->
+
+<!-- ``` -->
+
+<!-- ## What about signals? -->
+
+<!-- ### When are signals honest? -->
+
+<!-- * Impossible to fake -->
+<!-- * Individuals have common interests -->
+<!-- * Handicap principle (signal´s cost is proportional to quality) -->
+
+<!-- ## What about signals? -->
+
+<!-- ### Individual recognition -->
+
+<!-- ```{r,echo=FALSE,fig.align='center',out.height='70%'} -->
+<!-- knitr::include_graphics("Images//indRecog.png") -->
+
+<!-- ``` -->
+<!-- [@tibbetts_Individual_2007] -->
+
+<!-- ## What about signals? -->
+
+<!-- ### BS or IR? -->
+
+<!-- ```{r,echo=FALSE,fig.align='center',out.height='75%'} -->
+<!-- knitr::include_graphics("Images//IRvsBS.png") -->
+
+<!-- ``` -->
+<!-- [@sheehan_There_2016] -->
+
+<!-- ## What about signals? -->
+
+<!-- ```{r,echo=FALSE,fig.align='center',out.width='90%'} -->
+<!-- knitr::include_graphics("Images//comSystemBot.png") -->
+<!-- ``` -->
+
+<!-- ## What about learning? -->
+
+<!-- ```{r,echo=FALSE,fig.align='center',out.width='90%'} -->
+<!-- knitr::include_graphics("Images//pavlov.jpg") -->
+<!-- ``` -->
+
+<!-- # Associative learning -->
+
+<!-- ## Reinforcement learning theory -->
+
+<!-- \begin{columns}[T] -->
+<!--  \column{0.5\linewidth} -->
+<!--  \begin{equation*} -->
+<!--   \Delta V_{t(s)}=\alpha \underbrace{(R_t-V_t)}_\text{prediction error} -->
+<!--   \end{equation*} -->
+<!--   \pause -->
+<!--   \includegraphics[width=1\textwidth]{Images//learn_curve.png} -->
+<!--  \column{0.5\linewidth} -->
+<!--   \pause -->
+<!--   \includegraphics[height=.9\textheight]{Images//predErrorNeur.png} -->
+<!-- \end{columns} -->
+
+<!-- ## Reinforcement learning theory -->
+
+<!-- ```{r,echo=FALSE,fig.align='center',out.width='90%'} -->
+<!-- knitr::include_graphics("Images//learnUpdate.png") -->
+<!-- ``` -->
+
+<!-- ## The Hawk-Dove game -->
+
+<!-- Individuals have one of two genetically determined phenotypic strategies.
+_Hawks_ are willing to start a conflict over resources, while _doves_ prefer to stand down in the hope to share the resource without an aggressive contest.
+-->
+
+<!-- ```{r,echo=FALSE,fig.align="center",out.width='50%'} -->
+<!-- knitr::include_graphics("Images//HDgame.png") -->
+<!-- ``` -->
+
+<!-- ## The hawk-dove game -->
+
+<!-- \begin{align*} -->
+<!-- w_H &= p_H \frac{V-C}{2}+(1-p_h) V\\ -->
+<!-- w_D &= p_H 0 + (1-p_H)\frac{V}{2} -->
+<!-- \end{align*} -->
+<!-- \vspace{-0.8cm} -->
+<!-- ```{r, echo=FALSE,fig.align="center", out.width = '55%'} -->
+<!-- knitr::include_graphics("Images//Hawk_dove_plot.png") -->
+<!-- ``` -->
+
+<!-- ## The hawk-dove game -->
+
+<!-- ```{r, echo=FALSE, fig.cap="\\label{fig:HD_game}Hawk-dove game. Dashed line is the  game theoretical prediction for frequency of hawks.", out.width = '80%'} -->
+<!-- knitr::include_graphics("Simulations/mutType_/basicHawkDove.png") -->
+<!-- ``` -->
+
+<!-- ## Environmental states -->
+
+<!-- ### Discrete states -->
+<!-- \includegraphics[width=0.8\textwidth]{Images//discreteStates.png} -->
+
+<!-- \pause -->
+
+<!-- ### Continuos states -->
+<!-- \includegraphics[width=0.8\textwidth]{Images//contStates.png} -->
+
+<!-- ## Continuos environmental states -->
+<!-- \begin{equation*} -->
+<!--   \Delta V_{t(s)}=\alpha \underbrace{(R_t-V_t)}_\text{prediction error} -->
+<!-- \end{equation*} -->
+
+<!-- ```{r fig2, echo=FALSE, fig.cap="\\label{fig:learning_cartoonRBF}Function approximation for the actor and the critic.
+Using random values for the responses of each center.
+Circles show the location of the centers and the maximum response they trigger, while the lines show the total response along the badge size continium.", out.width = '60%'} -->
+<!-- knitr::include_graphics("Images/cartoonRBF.png") -->
+<!-- ``` -->
+
+<!-- # Our model -->
+
+<!-- ## The hawk-dove game 2.0 -->
+
+<!-- \begin{center} -->
+<!-- \begin{tabular}{ >{\centering\arraybackslash}p{1cm} | >{\centering\arraybackslash}p{4.5cm} | >{\centering\arraybackslash}p{4.5cm} } -->
+<!-- & H & D \\ \hline -->
+<!-- H & $p_w V\frac{-C}{2} + (1-p_w) \frac{-C}{2}$ & $V$ \\ \hline -->
+<!-- V & $0$ & $\frac{V}{2}$\\ -->
+<!-- \end{tabular} -->
+<!-- \end{center} -->
+
+<!-- \begin{equation*} -->
+<!-- p_w=\frac{1}{1+e^{-\beta(Q_i-Q_j)}} -->
+<!-- \end{equation*} -->
+
+<!-- ```{r,echo=FALSE,fig.align='center',out.width='50%'} -->
+<!-- par(plt=posPlot(),las=1) -->
+<!-- plot(logist(seq(-1,1,length.out = 1000),alpha = 0,beta = 2)~ -->
+<!--        seq(-1,1,length.out = 1000),type="l",xlab="Difference in quality", -->
+<!--      ylab = expression(p[w]),cex.lab=2,lwd=3,col="red",cex.axis=2,ylim=c(0,1)) -->
+<!-- lines(logist(seq(-1,1,length.out = 1000),alpha = 0,beta = 10)~ -->
+<!--        seq(-1,1,length.out = 1000),lwd=3,col="green") -->
+<!-- legend("topleft",legend=c(2,10),col=c("red","green"),lwd=3,title = expression(beta),cex=2,bty="n") -->
+
+<!-- ``` -->
+
+<!-- ## Sender Code -->
+
+<!-- \begin{equation*} -->
+<!-- s_i = \frac{1}{1+e^{-(\epsilon_i+\gamma_iQ_i)}} -->
+<!-- \label{eq:react_norm} -->
+<!-- \end{equation*} -->
+
+<!-- ```{r,echo=FALSE,fig.align='center',out.width='50%'} -->
+<!-- par(plt=posPlot(),las=1) -->
+<!-- plot(logist(seq(0,1,length.out = 1000),alpha = 2,beta = 4)~ -->
+<!--        seq(0,1,length.out = 1000),type="l",xlab="Quality", -->
+<!--      ylab = expression(p[w]),cex.lab=2,lwd=3,col="red",cex.axis=2,ylim=c(0,1)) -->
+<!-- lines(logist(seq(0,1,length.out = 1000),alpha = 5,beta = 10)~ -->
+<!--        seq(0,1,length.out = 1000),lwd=3,col="green") -->
+<!-- legend("topleft",legend=c(4,10),col=c("red","green"),lwd=3,title = expression(gamma),cex=2,bty="n") -->
+
+<!-- ``` -->
+
+<!-- ## Receiver code -->
+
+<!-- \begin{equation*} -->
+<!--   \Delta V_{t(s)}=\alpha (R_t-V_t) -->
+<!-- \end{equation*} -->
+
+<!-- ```{r, echo=FALSE,fig.align='center', out.width = '60%'} -->
+<!-- knitr::include_graphics("Images/cartoonRBF.png") -->
+<!-- ``` -->
+
+# Results
+
 ```{r setup, include=FALSE}
@@
 
-## The Hawk-Dove game
+## Learning under honest signalling
 
-Individuals have one of two genetically determined phenotypic strategies.
-_Hawks_ are willing to start a conflict over resources, while _doves_ prefer to stand down in the hope to share the resource without an aggressive contest.
-
-```{r,echo=FALSE,fig.align="center",out.width='50%'}
-knitr::include_graphics("Images//HDgame.png")
-```
-
-## The hawk-dove game
-
-\begin{align*}
-w_H &= p_H \frac{V-C}{2}+(1-p_h) V\\
-w_D &= p_H 0 + (1-p_H)\frac{V}{2}
-\end{align*}
-\vspace{-0.8cm}
-```{r, echo=FALSE,fig.align="center", out.width = '55%'}
-knitr::include_graphics("Images//Hawk_dove_plot.png")
-```
-
-<!-- ## The hawk-dove game -->
-
-<!-- ```{r, echo=FALSE, fig.cap="\\label{fig:HD_game}Hawk-dove game. Dashed line is the  game theoretical prediction for frequency of hawks.", out.width = '80%'} -->
-<!-- knitr::include_graphics("Simulations/mutType_/basicHawkDove.png") -->
-<!-- ``` -->
-
-## What about signals?
-
-```{r,echo=FALSE,fig.align='center',out.height='80%'}
-knitr::include_graphics("Images//hosp_alexandra_mackenzie.jpg")
-
-```
-
-## What about signals?
-
-### When are signals honest?
-
-* Impossible to fake
-* Individuals have common interests
-* Handicap principle (signal´s cost is proportional to quality)
-  + Social costs?
-
-```{r,echo=FALSE, fig.align='center',out.width='100%'}
-knitr::include_graphics("Images//social_costs.png")
-```
-
-## What about signals?
-
-```{r,echo=FALSE,fig.align='center',out.width='90%'}
-knitr::include_graphics("Images//comSystemBot.png")
-```
-
-## What about learning?
-
-```{r,echo=FALSE,fig.align='center',out.width='90%'}
-knitr::include_graphics("Images//pavlov.jpg")
-```
-
-# Associative learning
-
-## Reinforcement learning theory
-
-\begin{columns}[T]
- \column{0.5\linewidth}
- \begin{equation*}
-  \Delta V_{t(s)}=\alpha \underbrace{(R_t-V_t)}_\text{prediction error}
-  \end{equation*}
-  \pause
-  \includegraphics[width=1\textwidth]{Images//learn_curve.png}
- \column{0.5\linewidth}
-  \pause
-  \includegraphics[height=.9\textheight]{Images//predErrorNeur.png}
-\end{columns}
-
-## Reinforcement learning theory
-
-```{r,echo=FALSE,fig.align='center',out.width='90%'}
-knitr::include_graphics("Images//learnUpdate.png")
-```
-
-## Environmental states
-
-### Discrete states
-\includegraphics[width=0.8\textwidth]{Images//discreteStates.png}
-
-\pause
-
-### Continuos states
-\includegraphics[width=0.8\textwidth]{Images//contStates.png}
-
-## Continuos environmental states
-\begin{equation*}
-  \Delta V_{t(s)}=\alpha \underbrace{(R_t-V_t)}_\text{prediction error}
-\end{equation*}
-
-```{r fig2, echo=FALSE, fig.cap="\\label{fig:learning_cartoonRBF}Function approximation for the actor and the critic.
-Using random values for the responses of each center.
-Circles show the location of the centers and the maximum response they trigger, while the lines show the total response along the badge size continium.", out.width = '60%'}
-knitr::include_graphics("Images/cartoonRBF.png")
-```
-
-# Our model
-
-## The hawk-dove game 2.0
-
-\begin{center}
-\begin{tabular}{ >{\centering\arraybackslash}p{1cm} | >{\centering\arraybackslash}p{4.5cm} | >{\centering\arraybackslash}p{4.5cm} }
-& H & D \\ \hline
-H & $p_w V\frac{-C}{2} + (1-p_w) \frac{-C}{2}$ & $V$ \\ \hline
-V & $0$ & $\frac{V}{2}$\\
-\end{tabular}
-\end{center}
-
-\begin{equation*}
-p_w=\frac{1}{1+e^{-\beta(Q_i-Q_j)}}
-\end{equation*}
-
-```{r,echo=FALSE,fig.align='center',out.width='50%'}
-par(plt=posPlot(),las=1)
-plot(logist(seq(-1,1,length.out = 1000),alpha = 0,beta = 2)~
-       seq(-1,1,length.out = 1000),type="l",xlab="Difference in quality",
-     ylab = expression(p[w]),cex.lab=2,lwd=3,col="red",cex.axis=2,ylim=c(0,1))
-lines(logist(seq(-1,1,length.out = 1000),alpha = 0,beta = 10)~
-       seq(-1,1,length.out = 1000),lwd=3,col="green")
-legend("topleft",legend=c(2,10),col=c("red","green"),lwd=3,title = expression(beta),cex=2,bty="n")
-
-```
-
-## Sender Code
-
-\begin{equation*}
-s_i = \frac{1}{1+e^{-(\epsilon_i+\gamma_iQ_i)}}
-\label{eq:react_norm}
-\end{equation*}
-
-```{r,echo=FALSE,fig.align='center',out.width='50%'}
-par(plt=posPlot(),las=1)
-plot(logist(seq(0,1,length.out = 1000),alpha = 2,beta = 4)~
-       seq(0,1,length.out = 1000),type="l",xlab="Quality",
-     ylab = expression(p[w]),cex.lab=2,lwd=3,col="red",cex.axis=2,ylim=c(0,1))
-lines(logist(seq(0,1,length.out = 1000),alpha = 5,beta = 10)~
-       seq(0,1,length.out = 1000),lwd=3,col="green")
-legend("topleft",legend=c(4,10),col=c("red","green"),lwd=3,title = expression(gamma),cex=2,bty="n")
-
-```
-
-## Receiver code
-
-\begin{equation*}
-  \Delta V_{t(s)}=\alpha (R_t-V_t)
-\end{equation*}
-
-```{r, echo=FALSE,fig.align='center', out.width = '60%'}
-knitr::include_graphics("Images/cartoonRBF.png")
-```
-
-# Results
-
-## Learning under honest signalling
+Learning simulations (without evolution yet), the sender code is set as an honest signal.
+Receivers develop a threshold like response according to their quality
 
@@
 ### Variation among replicates
+\small
+set of simulations with the badge working as a handicap (cost inversely proportional to quality)
+
 ```{r, echo=FALSE,fig.align='center', out.width='100%'}
@@
 ### Individual replicates: Badges
+\small
+One of the replicates of the previous slide, the badge as a handicap
 ```{r, echo=FALSE,fig.align='center', out.width='100%'}
@@
 ## Learning can mediate the evolution of Badges of Status
+### Individual replicates: Badges
+\small
+Panels show the changes in frequency distribution along evolution ($\alpha$, $\beta$, and resulting badge respectively) for the replicate shown in the previous slide
+
+```{r, echo=FALSE,fig.align='left', out.height='28%',fig.show='hold', fig.cap="alpha - beta -Badge"}
+knitr::include_graphics("Simulations/betCostEvol1_/evolDistAlpha0_betCost5.png")
+knitr::include_graphics("Simulations/betCostEvol1_/evolDistBeta0_betCost5.png")
+knitr::include_graphics("Simulations/betCostEvol1_/evolDistBadge0_betCost5.png")
+```
+
+## Learning can mediate the evolution of Badges of Status
 ### Individual replicates: no badges
+\small
+One of the replicates where the badge did not evolve
+
 ```{r, echo=FALSE,fig.align='center', out.width='100%'}
@@
 ### Behavioural interactions
+\small
+Colours correspond to replicates, each point is one snapshot on the second half of the simulations.
+Higher costs seem to favour the signalling equilibrium
 
-```{r, echo=FALSE,fig.align='center', out.width='100%'}
+```{r, echo=FALSE,fig.align='center', out.width='80%'}
 knitr::include_graphics("Simulations/betCostEvol1_/BehavIntALL.png")
 ```
+## The effect of costs on Badges of Status
+### Badge size
+\small
+Same as previous slide but for the resulting signal. Higher costs yield smaller averga sizes and smaller ranges.
+
+```{r, echo=FALSE,fig.align='center', out.width='80%'}
+knitr::include_graphics("Simulations/betCostEvol1_/CueALL.png")
+```
+
+## The evolution of cheap Badges of Status
+### Variation among replicates
+\small
+Here simulations where the signal has no cost
+
+```{r, echo=FALSE,fig.align='center', out.width='80%'}
+knitr::include_graphics("Simulations/nIntGroupEvol1_/evolDynALL_nIntGroup2000.png")
+```
+
+## The evolution of cheap Badges of Status
+### Variation within a replicate
+\small
+Here simulations where the signal has no cost.
+
+There seems to be some consistency in the three type of sender codes
+
+```{r, echo=FALSE,fig.align='center', out.width='80%'}
+knitr::include_graphics("Simulations/nIntGroupEvol1_/evolDyn8_nIntGroup2000.png")
+```
+
+## The evolution of cheap Badges of Status
+### Variation within a replicate
+
+\tiny
+Panels show the changes in frequency distribution along evolution ($\alpha$, $\beta$, and resulting badge respectively) for the replicate shown in the previous slide
+
+Branching events in the reaction norm parameters yield three types of sender codes. One (more or less)
+honest, two dishonest (one with a big badge and one with a small badge)
+
+```{r, echo=FALSE,fig.align='left', out.height='28%',fig.show='hold', fig.cap="alpha - beta -Badge"}
+knitr::include_graphics("Simulations/nIntGroupEvol1_/evolDistAlpha8_nIntGroup2000.png")
+knitr::include_graphics("Simulations/nIntGroupEvol1_/evolDistBeta8_nIntGroup2000.png")
+knitr::include_graphics("Simulations/nIntGroupEvol1_/evolDistBadge8_nIntGroup2000.png")
+```
+
+## The evolution of cheap Badges of Status
+### How often does it happen?
+\tiny
+Values of $\alpha$ and $\beta$ for each individual in the different replicates. Each dot an individual,
+each panel a replicate. Individuals are sampled from populations from the second half of the simulations.
+
+Individuals in some simulations split up in three groups, others in two groups; 
+in one simulation the population does not split up
+
+```{r, echo=FALSE,fig.align='left', out.width='90%',fig.show='hold'}
+knitr::include_graphics("Simulations/nIntGroupEvol1_/corrAlphBet_nIntGroup2000.png")
+```
+
+## The evolution of cheap Badges of Status
+### Does it matter?
+\small
+Behavioural interactions splitting up the replicates according to the number of "types" that 
+evolve
+
+```{r, echo=FALSE,fig.align='left', out.width='90%',fig.show='hold'}
+knitr::include_graphics("Simulations/nIntGroupEvol1_/BehavIntAllClusters_nIntGroup2000.png")
+```
+
+## Interaction structures - do they matter?
+
+```{r, echo=FALSE,fig.align='center', out.width='100%',}
+knitr::include_graphics("Images/networks.png")
+```
+[@chaine_Manipulating_2018]
+
+## Localized interaction structures
+
+### Costly signals 
+\tiny
+Behavioural interactions and badge size for simulations with costs where individuals interact in small groups. On the $x$ axes number of different individuals each focal interacts with.
+
+```{r, echo=FALSE,fig.align='center', out.width='90%',fig.show='hold'}
+knitr::include_graphics("Simulations/nIntGroupCost_/BehavIntALL.png")
+knitr::include_graphics("Simulations/nIntGroupCost_/CueALL.png")
+```
+
+## Localized interaction structures
+
+### Cheap signals (free)
+\small
+Variation among individuals in different replicates (just like before) in replicates where individuals interact with 10 other individuals.
+
+A part from replicates with two clusters, the clustering seems less straightforward as before
+
+```{r, echo=FALSE,fig.align='center', out.width='85%',fig.show='hold'}
+knitr::include_graphics("Simulations/nIntGroupEvol1_/corrAlphBet_nIntGroup10.png")
+```
+
+## Localized interaction structures
+
+### Cheap signals (free)
+
+\small
+Behavioural interactions for the replicates form previous slide. I split up the replicates according to clusters (despite clusters not being clear cut).
+
+Only replicates with only one cluster seems to have less aggressive interactions
+
+```{r, echo=FALSE,fig.align='center', out.width='90%',fig.show='hold'}
+knitr::include_graphics("Simulations/nIntGroupEvol1_/BehavIntAllClusters_nIntGroup10.png")
+```
+
+## Localized interaction structures
+
+### Cheap signals or Individual recognition?
+\tiny
+In these set, I changed the learning set up, I increased the number of centers (10 now, 6 before)
+along the signal axis; and I reduced the level or generalization.
+
+There seems to be an effect on the slope of the reaction norm. It stays very close to zero.
+
+```{r, echo=FALSE,fig.align='center', out.width='90%',fig.show='hold'}
+knitr::include_graphics("Simulations/nIntGroupCenters1_/evolDynALL_nIntGroup10.png")
+```
+
+## Localized interaction structures
+
+### Cheap signals or Individual recognition
+\small
+Same replicates as the previous slide. Variation in $\beta$ and $\alpha$ among individuals for different replicates.
+
+No clear clustering. Variation mostly along the $\alpha$ axis
+
+```{r, echo=FALSE,fig.align='center', out.width='90%',fig.show='hold'}
+knitr::include_graphics("Simulations/nIntGroupCenters1_/corrAlphBet_nIntGroup10.png")
+```
+## Localized interaction structures
+
+### Cheap signals or Individual recognition?
+\small
+Panels show the changes in frequency distribution along evolution ($\alpha$, $\beta$, and resulting badge respectively) for the replicate shown in the previous slide.
+
+$\beta$ around zero, $\alpha$ all over the place.
+
+```{r, echo=FALSE,fig.align='left', out.height='28%',fig.show='hold'}
+knitr::include_graphics("Simulations/nIntGroupCenters1_/evolDistAlpha5_nIntGroup10.png")
+knitr::include_graphics("Simulations/nIntGroupCenters1_/evolDistBeta5_nIntGroup10.png")
+knitr::include_graphics("Simulations/nIntGroupCenters1_/evolDistBadge5_nIntGroup10.png")
+```
+
+<!-- ## Aqui vamos! mirar correlaciones de 20 y 2000 luego interacciones -->
+
+<!-- ## Comparar scenarios -->
+
+## Why would you want to be recognized?
+
+\pause
+
+### The peaceful, the aggressive and the clever
+
+- Learning initial conditions
+
+\tiny
+Here, I vary the aggressive tendency for unlearned individuals in these three set ups.
+
+Agressive, peaceful, clever (starts at the ESS of the classic hawk-dove game)
+
+```{r, echo=FALSE,fig.align='left', out.width='90%',fig.show='hold'}
+knitr::include_graphics("Images/cartoonRBF_InitAct.png")
+```
+
+## Why would you want to be recognized?
+
+### The peaceful
+
+Variation among replicates
+
+```{r, echo=FALSE,fig.align='left', out.width='100%',fig.show='hold'}
+knitr::include_graphics("Simulations/initAct_/evolDynALL_initAct2.png")
+```
+
+## Why would you want to be recognized?
+
+### The peaceful
+
+Variation within one replicate
+
+```{r, echo=FALSE,fig.align='left', out.width='100%',fig.show='hold'}
+knitr::include_graphics("Simulations/initAct_/evolDyn2_initAct2.png")
+```
+
+## Why would you want to be recognized?
+
+### The peaceful
+
+Changes in the frequency distributions
+
+```{r, echo=FALSE,fig.align='left', out.height='28%',fig.show='hold'}
+knitr::include_graphics("Simulations/initAct_/evolDistAlpha2_initAct2.png")
+knitr::include_graphics("Simulations/initAct_/evolDistBeta2_initAct2.png")
+knitr::include_graphics("Simulations/initAct_/evolDistBadge2_initAct2.png")
+```
+
+## Why would you want to be recognized?
+
+### The peaceful
+
+Favours variation and cheap signals
+
+```{r, echo=FALSE,fig.align='left', out.width='100%',fig.show='hold'}
+knitr::include_graphics("Simulations/initAct_/corrAlphBet_initAct2.png")
+```
+
+## Why would you want to be recognized?
+
+### The aggressive
+
+Variation among replicates
+
+```{r, echo=FALSE,fig.align='left', out.width='100%',fig.show='hold'}
+knitr::include_graphics("Simulations/initAct_/evolDynALL_initAct-2.png")
+```
+
+## Why would you want to be recognized?
+
+### The aggressive
+
+Variation within one replicate
+
+```{r, echo=FALSE,fig.align='left', out.width='100%',fig.show='hold'}
+knitr::include_graphics("Simulations/initAct_/evolDyn0_initAct-2.png")
+```
+
+## Why would you want to be recognized?
+
+### The aggressive
+
+Disfavours variation and signal evolution - no signalling
+
+```{r, echo=FALSE,fig.align='left', out.width='100%',fig.show='hold'}
+knitr::include_graphics("Simulations/initAct_/corrAlphBet_initAct-2.png")
+```
+
+## Why would you want to be recognized?
+
+### The clever (ESS)
+
+Variation among replicates
+
+```{r, echo=FALSE,fig.align='left', out.width='100%',fig.show='hold'}
+knitr::include_graphics("Simulations/initAct_/evolDynALL_initAct-0.69.png")
+```
+
+## Why would you want to be recognized?
+
+### The clever (ESS)
+
+Variation within one replicate
+
+```{r, echo=FALSE,fig.align='left', out.width='100%',fig.show='hold'}
+  knitr::include_graphics("Simulations/initAct_/evolDyn0_initAct-0.69.png")
+```
+
+## Why would you want to be recognized?
+
+### The clever (ESS)
+
+Signal evolution seems to be driven by drift. Most replicates end up in one of the two extremes
+
+```{r, echo=FALSE,fig.align='left', out.width='100%',fig.show='hold'}
+  knitr::include_graphics("Simulations/initAct_/corrAlphBet_initAct-0.69.png")
+```
+
+## Take-home messages
+
+### Costly signals
+- Learning can mediate costly signals
+- More expensive signals seem be more likely to evolve
+- Less expensive signals provide more information
+- Local interactions do not impede the evolution of costly signals (as has been suggested)
+
+### Cheap signals
+- Learning mediates the evolution of polymorphisms with "honest" and "dishonest"
+individuals
+- Under local interactions learning seems to allow the evolution of IR
+- Cheap signals depend on the innate level of aggression (before learning) of individuals
+- Peace begets heterogeneity, aggression begets homogeneity
+
+## Questions?
+
+```{r, echo=FALSE,fig.align='center', out.height='80%',fig.show='hold'}
+  knitr::include_graphics("Images/BizarroParrot.jpg")
+```
+
+## References

### 2020-10-01 manuscript_1.0.Rmd
Random changes to improve flow, and comments on the presentation for Olle to follow

@@
 output:
+  pdf_document: default
   word_document: default
-  pdf_document: default
 fig_caption: yes
@@
 style: C:\Users\andre\Zotero\styles\elife.csl
-abstract: Learning rules.
+abstract: Adaptive behavioural responses often depend on qualities of the interacting partner of an individual.
+For example, when competing for resources, an individual might be better off escalating fights with individuals of lower quality, while restraining from fighting individuals of higher quality.
+Communication systems involving signals of quality allow individuals to reduce uncertainty regarding the fighting ability of their partners and make more adaptive behavioral decisions.
+However, dishonest individuals can destabilize such communications systems.
+One open question is whether cognitive mechanisms, such as learning, can maintain the honesty of signals, thus, favoring their evolutionary stability.
+Here we present evolutionary simulations, where individuals can produce a signal proportional to their quality; and learn along their lifetime the best response to the signal emitted by their peers.
+In these simulations, learning on the receiver side can mediate the evolution of signals of quality on the sender side.
+When the cost of the signal is proportional to the quality of the sender populations are only composed of honest signalers.
+When the cost is not proportional to the quality of the signaler, the population is composed of both honest and dishonest signalers.
+We argue that learning can be a general cognitive mechanism playing a role in a wide range of communication systems.
 ---
@@
 lack of integration between learning and evolution. Reinforcement learning theory 
-encompases a series of computational methods inspired on the psychological and 
+encompass a series of computational methods inspired on the psychological and 
 neurological mechanisms of associative learning [@sutton_Reinforcement_2018].
@@
 capturing the essence of learning processes [@frankenhuis_Enriching_2018; @quinones_Reinforcement_2019]. 
-Furthermore, these algorithms can be embebed in evolutionary simulations in order 
+Furthermore, these algorithms can be embedded in evolutionary simulations in order 
 to have theoretical predictions of the effect that learning can have in behavioral [@leimar_Learning_2019]
@@
 to develop a tendency to behave aggressively (or  peacefully), in the context of
-competition over resoureces, depending on the quantitative morphological trait 
+competition over resources, depending on the quantitative morphological trait 
 they perceive in their opponent (badge). Over evolutionary time individuals evolve

### 2020-10-01 AccFunc.R

@@ -2,10 +2,13 @@
 
 # Libraries
 source(here("aesth.R"))
+library("jsonlite")
 library('rmarkdown')
 source(here("..","R_files","posPlots.R"))
+source(here("..","R_files","Filled.contour3.R"))
 # library('plotrix')
 library(data.table)
+require(lattice)
 
 # Calculate response triggered by one RBF --------------------------------------
 
@@ -92,6 +95,24 @@ filesScenar<-function(filename,scenario,full.name=FALSE){
   return(tmp)
 }
 
+# Load all the files for different scenarios  varying one parameter ----------------------
+
+
+filesCompScenar<-function(filename,scenario,full.name=FALSE){
+  par<-gsub(".txt","",tail(strsplit(filename,"_")[[1]],1))
+  parVal<-as.numeric(gsub("[[:alpha:]]",par,replacement = ''))
+  parNam<-gsub("[^[:alpha:]]",par,replacement = '')
+  if(full.name){
+    tmp<-fread(filename)
+  }
+  else{
+    tmp<-fread(here("Simulations",paste0(scenario,"_"),filename)) 
+  }
+  tmp[,eval(parNam):=parVal]
+  return(tmp)
+}
+
+
 # Visualize difference in parameters between 2 JSON files ----------------------
 
 diffJsons<-function(json1,json2){
@@ -148,6 +169,60 @@ Critic<-function(weights,centers,sigSq=0.01,nx=1000){
                 centers,sigSq,as.double(weights))
 }
 
+# Functio to generate evolutionary dynamics in the form of frecuency distributions
+
+evolDist<-function(indData,variable,nbins,range=NULL,pal, nlevels =10, 
+                   numx=1,numy=1,idx=1,idy=1,NewP=FALSE,
+                   cexAxis=2,xlab="x",ylab=variable, keyTitle="", ...){
+  if(is.null(range)){
+    maxV<-indData[,max(get(variable))];minV<-indData[,min(get(variable))]
+  }
+  else {maxV<-range[1];minV<-range[2]}
+  Interv<-seq(minV,maxV,length.out = nbins+1)
+  ints<-(Interv[2]-Interv[1])/2
+  bins<-seq(minV+ints,maxV-ints,length.out = nbins)
+  timeseq<-indData[,unique(time)]
+  indData[,Int:=findInterval(get(variable),Interv)]
+  indData[,bin:=bins[Int]]
+  totalCounts<-indData[,length(bin),by=.(time)]
[…]

### 2020-10-01 Dyn_reac_learn.R

@@ -12,9 +12,9 @@ require("jsonlite")
 
 # Scenario to be plotted - corresponds to folders where simulations are stored
 
-scenario<-"betCostEvol1"
+scenario<-"nIntGroupEvolLear"
 
-extSimsDir<-#here(paste0(scenario,"_"))
+extSimsDir<-#here("Simulations",paste0(scenario,"_"))
   paste0("e:/BadgeSims/",scenario,"_")
 
 
@@ -28,18 +28,20 @@ extSimsDir<-#here(paste0(scenario,"_"))
 (indList<-grep("indLearn",listTest,value=TRUE))
 paramName<-list.files(here("Simulations",paste0(scenario,"_")))
 paramName<-grep(".json",paramName,value=TRUE)
-param<-fromJSON(here("Simulations",paste0(scenario,"_"),paramName))
+param<-fromJSON(here("Simulations",paste0(scenario,"_"),paramName[1]))
 
 
-numCores <- 3
+numCores <- length(indList)
 registerDoParallel(numCores)
 
-val<-3
+val<-2
 
 # loop to produce pdfs for parameter values
-foreach(val = 1:3,.packages = c("data.table","here")) %dopar% {
+foreach(val = 1:length(indList),.packages = c("data.table","here")) %dopar% {
 source(here("AccFunc.R"))
-fileId<-val
+
+  fileId<-val
+
 # Project folder
 # evol<-fread(here("Simulations",paste0(scenario,"_"),evolList[fileId]))
 # pop<-fread(here("Simulations",paste0(scenario,"_"),indList[fileId]))
@@ -47,20 +49,19 @@ fileId<-val
 evol<-fread(evolList[fileId])
 pop<-fread(indList[fileId])
 
-# names(pop)[29:35]<-c("Quality", "genotype","alpha","beta","Badge","initCrit","initAct")
-# names(pop)[9:28]<-do.call(rbind,as.list(sapply(0:9,FUN = function(x){
-#     return(rbind(paste0("WeightAct_",x),paste0("WeightCrit_",x)))
-#     })))
-
-
 Valpar<-gsub("[[:alpha:]]",gsub(".txt","",tail(strsplit(indList[val],"_")[[1]],1)),
-     replacement = "")
+             replacement = "")
 nampar<-gsub("[^[:alpha:]]",gsub(".txt","",tail(strsplit(indList[val],"_")[[1]],1)),
[…]

### 2020-10-01 aesth.R

@@ -28,6 +28,8 @@ colboxes<- c('#31a354','#d7191c','#fdae61','#2b83ba',"black")
 
 colRuns<-c('#e41a1c','#377eb8','#4daf4a', '#984ea3','#ff7f00')
 
-colSeeds<-brewer.pal(8,name = "Set2")
+colSeeds<-brewer.pal(10,name = "Paired")
+
+pal_dist<-colorRampPalette(brewer.pal(9,"Blues"))
 
 colorbreaksQual<-seq(0,1,length=100)

### 2020-10-01 compScen.R

@@ -1,0 +1,38 @@
+# Comparing results from different scenarios -----------------------------------
+
+# Required libraries -----------------------------------------------------------
+
+library(here)
+here()
+source(here("AccFunc.R"))
+
+
+# Choove scenarios to be plotted - corresponds to folders where simulations are stored
+
+scenarios<-c("nIntGroupEvol1","betCostEvol1")
+
+variable<-
+  lapply(scenarios,FUN=gsub,pattern="[^[:alpha:]]",replacement = '')
+variable1<-gsub("Evol",variable1,replacement = '')
+extSimsDir<-paste0("e:/BadgeSims/",scenario1,"_")
+
+variable2<-gsub("[^[:alpha:]]",scenario2,replacement = '')
+variable2<-gsub("Evol",variable2,replacement = '')
+extSimsDir<-paste0("e:/BadgeSims/",scenario2,"_")
+
+
+# Load files -------------------------------------------------------------------
+
+(listTest<-list.files(extSimsDir,full.names = TRUE))
+(evolList<-grep("evolLearn",listTest,value=TRUE))
+(indList<-grep("indLearn",listTest,value=TRUE))
+
+nampar<-gsub("[^[:alpha:]]",gsub(".txt","",tail(strsplit(indList[1],"_")[[1]],1)),
+             replacement = "")
+
+# (listTest<-list.files(here("Simulations",paste0(scenario,"_"))))
+# (sdList<-grep("evol",listTest,value=TRUE))
+
+evol<-do.call(rbind,lapply(evolList,filesScenar,scenario,full.name=TRUE))
+
+pop<-do.call(rbind,lapply(indList, filesScenar,scenario,full.name=TRUE))

### 2020-10-01 cue_compt.cpp

@@ -69,6 +69,8 @@ public:
 	double featActMean[20];
 	// records the distribution of Actor features at the end of learning for evolutionary dyn
 	double featCritMean[20];
+	// records the mean fitness of the populations
+	double meanFit;
 	// records the distribution of Critic features at the end of learning for evolutionary dyn
 	//vector<vector<int> > countIntTypesGen;
 	// records the frequency of interaction types for learning dyn
@@ -85,6 +87,8 @@ public:
 	// Historical record of total # ints at the point of sample records
 	vector<vector<int>> nTotDDIntsHist;
 	// Historical record of total # ints at the point of sample records
+	vector<vector<double>> cumPayoffs;
+	// Historical record of the cumpayoff along life of inds
 	vector<vector<int>> rivals;
 	// Historical record of the rivals faced at the point of sample records
 	vector<vector<vector<double> > > actFeatHistory;
@@ -108,6 +112,7 @@ printingObj::printingObj(int nSamples, int nInt, int printLearnInt,
 		nTotHHIntsHist.emplace_back(0);
 		nTotDDIntsHist.emplace_back(0);
 		rivals.emplace_back(0);
+		cumPayoffs.emplace_back(0);
 		for (int countSamples = 0; countSamples < nSamples; ++countSamples) {
 			if (countRecords == 0) {
 				sampledInd.emplace_back(0), counterRecords.emplace_back(0);
@@ -118,6 +123,7 @@ printingObj::printingObj(int nSamples, int nInt, int printLearnInt,
 			nTotHHIntsHist[countRecords].emplace_back(0);
 			nTotDDIntsHist[countRecords].emplace_back(0);
 			rivals[countRecords].emplace_back(0);
+			cumPayoffs[countRecords].emplace_back(0);
 			for (int countCenters = 0; countCenters < nCenters; ++countCenters) {
 				actFeatHistory[countRecords][countSamples].emplace_back(0);
 				critFeatHistory[countRecords][countSamples].emplace_back(0);
[…]

### 2020-10-01 dynSevFiles.R

@@ -9,7 +9,7 @@ library("plotrix")
 
 # Scenario to be plotted - corresponds to folders where simulations are stored
 
-scenario<-"betCostEvol1"
+scenario<-"nIntGroupEvolLear"
 variable<-gsub("[^[:alpha:]]",scenario,replacement = '')
 variable<-gsub("Evol",variable,replacement = '')
 extSimsDir<-paste0("e:/BadgeSims/",scenario,"_")
@@ -22,6 +22,7 @@ extSimsDir<-paste0("e:/BadgeSims/",scenario,"_")
 
 nampar<-gsub("[^[:alpha:]]",gsub(".txt","",tail(strsplit(indList[1],"_")[[1]],1)),
              replacement = "")
+
 
 # (listTest<-list.files(here("Simulations",paste0(scenario,"_"))))
 # (sdList<-grep("evol",listTest,value=TRUE))
@@ -142,46 +143,48 @@ legend("bottomright",legend=c("HH","HD","DD"),
 
 pdf(paste0(extSimsDir,"/summ_",nampar,".pdf"))
 
+# png(here("Simulations",paste0(scenario,"_"),
+#          paste0("BehavIntALL",".png")),
+#     width = 1400,height = 500)
+
+
 # Jitter for the replicates
-
-png(here("Simulations",paste0(scenario,"_"),
-         paste0("BehavIntALL",".png")),
-    width = 1400,height = 1000)
-
 evol[,posX:=match(get(nampar),sort(unique(get(nampar))))+
        runif(length(get(nampar)),
-             min = -0.2,max = 0.2)]
-cexAxis<-2;cexPoints<-2
+             min = -0.3,max = 0.3)]
+cexAxis<-1.5;cexPoints<-1.5
 yLim<-as.numeric(evol[,.(min(c(freqHH,freqHD,freqDD)),
               max(c(freqHH,freqHD,freqDD)))])+c(0,0.02)
-par(mfrow=c(1,3),xaxt="s",las=1)
+par(mfrow=c(1,3),xaxt="s",las=1,plt=posPlot())
 
 # frequency of interactions types
 
-bbHH<-boxplot(freqHH~get(nampar),data=evol[time>max(time)/2],
-         pch=16,xlab="Badge cost",cex.lab=2,
-         #xlab=variable,
+bbHH<-boxplot(freqHH~get(nampar),data=evol,
+         pch=16,cex.lab=2,
+         xlab=variable,
[…]

### 2020-10-01 learnDynamics.R

@@ -11,7 +11,7 @@ source(here("AccFunc.R"))
 
 # Scenario to be plotted - corresponds to folders where simulations are stored
 
-scenario<-"alphaAct"
+scenario<-"nIntGroupEvol2"
 extSimsDir<-paste0("e:/BadgeSims/",scenario,"_")
 
 
@@ -21,13 +21,14 @@ extSimsDir<-paste0("e:/BadgeSims/",scenario,"_")
 # (listTest<-list.files(here("Simulations",paste0(scenario,"_"))))
 (listTest<-list.files(extSimsDir,full.names = TRUE))
 (indList<-grep("ind",listTest,value=TRUE))
+
 # parameter values from project folder
 paramName<-list.files(here("Simulations",paste0(scenario,"_")))
 paramName<-grep(".json",paramName,value=TRUE)
-param<-fromJSON(here("Simulations",paste0(scenario,"_"),paramName[2]))
+param<-fromJSON(here("Simulations",paste0(scenario,"_"),paramName))
 
 
-fileId<-6
+fileId<-1
 
 
 # indLearn<-fread(here("Simulations",paste0(scenario,"_"),indList[fileId]))
@@ -48,7 +49,7 @@ nampar<-gsub("[^[:alpha:]]",gsub(".txt","",tail(strsplit(indList[fileId],"_")[[1
 
 # Changes in learning parameters for several individuals -----------------------
 
-gener<-tail(indLearn[,unique(time)],2)[1]
+gener<-tail(indLearn[,unique(time)],2)[2]
 nCenters<-param$nCenters
 sigSquar<-param$sigSq
 interv<-1/(nCenters-1)
@@ -56,13 +57,14 @@ centers<-interv*seq(0,nCenters-1)
 # nCenters<-5
 # interv<-1/nCenters
 # centers<-interv*0.5+interv*seq(0,nCenters-1)
-rangx<-seq(0,1,length=1000)
+rangx<-seq(0,1,length=100)
 colorbreaksQual<-seq(0,1,length=100)
 
 # Actor 
 
 finReps<-indLearn[time==max(time),unique(seed)]
-seedCh<-finReps[round(runif(1,0,length(finReps)))+1]
+seedCh<-3
+  finReps[round(runif(1,0,length(finReps)))+1]
 
 
 # Select run and generation to plot
@@ -71,8 +73,8 @@ tempPop<-indLearn[time==gener&seed==seedCh]
[…]

### 2020-10-01 parTOjson.R

@@ -9,30 +9,31 @@ source(here("AccFunc.R"))
 fileName<-"parameters"
 here()
 
-param1<-fromJSON(here("Simulations","alphaAct_","parameters21.json"))
+param1<-fromJSON(paste0("E:/BadgeSims","/nIntGroupEvol1_","/parameters.json"))
+  #fromJSON(here("Simulations","nIntGroupEvol1","parameters.json"))
 
-param<-list(totGen   =5,   nRep    = 30,
-            printGen = 1,   printLearn = 1,
+param<-list(totGen   =15000,   nRep    = 8,
+            printGen = 1000,   printLearn = 1000,
             printLearnInt = 50,
             popSize  = 2000,  baselineFit = 2,   
-            MutSd    = 0.3,
-            nInt        = 1000,  init     = c(0,0,1),
+            MutSd    = 0.2,
+            nInt        = 300,  init     = c(0,0,1),
             mutRate  = 0.001,  mutType  = 0,
             sampleSize = 50,   strQual  = 10,
-            errorQual = 0,    alphaBad	 = 3,    betaBad	 = 6,
+            errorQual = 0,    alphaBad	 = 0,    betaBad	 = 0,
             alphaCrit  = 0.5,  alphaAct = 0.5,
-            sigSq   	 = 0.001, nCenters = 10,
+            sigSq   	 = 0.01, nCenters = 6,
             initCrit = 0,      initAct=0,   gamma = 0,
             QualStDv   = 1.1, betCost = 0,
             alphCost	 = 3, mutLearn = FALSE,
             nIntGroup  = 2000,
             payoff_matrix = c(1.5,1,0,0.5),
             namParam = "nIntGroup",
-            rangParam = c(4,6,8,10,20,50,1000),
+            rangParam = c(2000,8),
             folderL=paste(here("Simulations"),"/",sep=""))
 
 
-apendScenar<-"LearnIR"
+apendScenar<-"Evol2"
 param$folderL<-paste0(param$folderL,
        param$namParam,apendScenar,"_/")
 # runTime<-"360:00:00"# "10:00:00"# 

### 2020-10-01 manuscript_1.0.Rmd
Changes in the manuscript based on CDC feedback

@@
 
-For social animals the outcome of interactions with con-specifics is an important 
-determinant of fitness. Irrespective of whether such interactions are of cooperative 
-or competitive nature, individual's actions and their partners will partly determine 
-their success in the reproductive arena. However, the best action for an individual 
-might vary depending on it's own conditions, and with whom they interact. 
-Thus, an individual making their action dependent on information about interacting 
+The outcome of interactions with con-specifics is an important 
+determinant of fitness in social animals. Irrespective of whether such interactions are cooperative 
+or competitive, the actions of interacting individual partly determine 
+their reproductive success. However, the best action for an individual 
+might vary depending on its condition, and that of individuals with whom it interacts. 
+Thus, acting based on information about interacting 
 partners is typically adaptive [@quinones_Negotiation_2016].
-The source of this information however is far from trivial. 
-In some cases interacting partners (signalers) might be 'willing' to provide accurate
+Acquiring this information, however, is far from trivial. 
+In some cases, interacting partners (signalers) might be 'willing' to provide accurate
 information, but in others it might be in their own interest to conceal information
 [@johnstone_Recognition_1997], or to provide erroneous information [@johnstone_Badges_1993]. 
-Typically, in a given context, some proportion of individuals will be interested 
-in broadcasting accurate information, and some others in hiding it. 
-Take for instance an interaction between two individuals where one can help the other. 
-Given some conditions of cost and benefits, the donor will be interested in helping 
-a related individual. So for a relative of the donor, 
-broadcasting kinship is advantageous; while for an unrelated individual, 
-concealing the lack of kinship will be better. This same pattern applies to 
+Typically, in a given context, some proportion of individuals will beenefit 
+from broadcasting accurate information, whereas others will benefit from hiding it. 
+Take, for instance, an interaction between two individuals where one can help the other. 
+Given some costs and benefits, a potential donor will be interested in helping 
+related individuals. Therefore, for a relative of the donor, 
+broadcasting kinship is advantageous, whereas for an unrelated individual, 
+concealing the lack of kinship is better. Similar scenarios may apply to 
 many aspects of social life such as finding mates, feeding offspring, dominance relationships 
-and aggressive interactions. In dominance and aggressive interactions an important 
-information to make actions dependent on is interacting partners fighting ability, 
+and aggressive interactions [@bradbury_Principles_2011; @moller_Female_1988; @tibbetts_Socially_2004].
+In dominance and aggressive interactions, an important piece of
+information to guide individual actions is interacting partners fighting ability, 
 which is sometimes referred to as Resource Holding Potential (RHP) [@parker_Assessment_1974], 
-or as we will here simply quality. This responsiveness to the quality of the 
-partner is achieved in communication systems such as badges of status (BS), 
-where an arbitrary signal conveys quality. 
-BS can be evolutionarily stable whenever the signal imposes a fitness cost which 
-decreases with the quality of an individual  [@botero_Evolution_2010; @johnstone_Badges_1993]. 
-So, low quality individuals can't afford the cost of the signal, 
-and it's no longer in their interest to fake quality. 
-The cost of the honest signal is a more general principle of communications systems, 
+or simply as quality. Responsiveness to the quality of the 
+partner is central to communication systems such as badges of status, 
+where an arbitrary signal conveys quality [@johnstone_Badges_1993; @rohwer_Social_1975a]. 
+Here, arbitrary refers to a signal which is not ontogeneteically correlated with quality. 
+Badges of status can be evolutionarily stable whenever the signal imposes a fitness cost which 
+decreases with the quality of an individual  [@botero_Evolution_2010; @johnstone_Badges_1993], 
+such that it's no longer in the interest of low quality individuals to fake quality by producing a
+large badge. The cost of honest signals is a more general principle of communications systems, 
 usually referred to as the handicap principle [@grafen_Biological_1990; @zahavi_Mate_1975]. 
 An alternative system of communication is one where individuals gather and store information
-about each individual they interact with. This information, directly collected by 
-each individual, should reliably allow them to react adaptively to the 
-quality difference with their opponent. This type of assessment strategy is often 
-referred to as individual recognition [IR, @whitfield_Plumage_1986]. Evidently, 
-IR is restricted to certain interaction structures, individuals are limited by cognitive 
-abilities in how many individuals they can recognize and track. Thus, IR and BS 
-have been assumed to be alternative communication systems fulfilling the same function, 
-but expected to evolve in populations under different conditions
-[@dale_Signaling_2001; @sheehan_There_2016]. There is no clear cut reason, however, 
-why both systems could not work together, in fact there is some evidence
+about each individual they interact with. This information should reliably allow 
+each individual to react adaptively to the quality difference with their opponent. 
+This type of assessment strategy is often referred to as individual recognition [@whitfield_Plumage_1986]. Evidently, individual recognition is restricted to 
+certain interaction structures, because individuals are limited by cognitive 
+abilities in how many individuals they can recognize and track. Thus, individual 
+recognition and badges of status have been assumed to be alternative communication 
+systems fulfilling the same function, reducing the costs of aggression in contests over resources. 
+Furthermore, badges of status and individual recognition are expected to evolve under different interaction structures, badges of status under global interactions and individual recognition
+under local interactions [@dale_Signaling_2001; @sheehan_There_2016].  There is no clear cut reason, however, why both systems could not work together, in fact there is some evidence
 that they can [@chaine_Manipulating_2018].
 
-A somewhat ignored aspect of communication systems, in the context of aggressive 
-interactions, is the cognitive aspects of the receiver module. 
-BS are often assume to be cognitively less demanding than IR 
-[@sheehan_There_2016]. The idea is that using IR, individuals need to store identity 
-and associated information about each individual they interact with. All this 
-information requires more memory and therefore is more demanding. In contrast,
-the response using BS have been thought as an innate response 
-[@botero_Evolution_2010; @johnstone_Badges_1993], where individuals use their 
-own quality and the opponent's badge to determine whether to aggressively contest 
-a resource. An alternative view of BS is that individuals learn to react to the signal
-based on their experiences [@guilford_Receiver_1991]. In line with that view, both
-communication systems would require the receivers to learn an association between
-the signal and the fighting ability of the bearer. In BS, this association would be a 
-a simple monotonic increase of fighting ability with the size of the badge, 
-and would be reinforced by every interaction, so in  principle it would be 
-quicker to learn. While in IR, the association would 
-be different for every individual, depending on the interacting partners. 
-Learning could be a unifying cognitive mechanism behind both types of communication 
-systems. This idea of learning being a general response mechanism has not been 
-thoroughly explored, nor in the empirical or theoretical literature.
+A somewhat ignored component of communication systems, in the context of aggressive 
+interactions, is the cognitive aspects of the receiver module. Communication systems
+relying on badges of status are often assumed to be cognitively less demanding 
+than those involving individual recognition [@sheehan_There_2016] because in the latter, 
+individuals need to store identity and associated information about each individual 
+they interact with. In contrast, the response towards badges of status is often 
+thought to be innate [@botero_Evolution_2010; @johnstone_Badges_1993], with 
+individuals using their own quality and the opponent's badge to determine whether to aggressively
+engage in contest. An alternative view of badges of status is that individuals learn 
+to react to them based on their experiences [@guilford_Receiver_1991], which would imply that receivers must learn to associate signals with the fighting ability of bearers just as in systems based on individual recognition.
+In cases where badges of status vary quantitatively (e.g.
+in size or intensity), fighting ability may increase monotonically with attributes of the badge and would be reinforced by every interaction, so in principle learning would be faster than in systems involving individual recognition in which the association between signals and their meaning would vary depending on the interacting partners.
+In any case, learning could be a central cognitive mechanism in both types of communication systems, but the role of learning in these contexts has not been thoroughly explored in the empirical nor theoretical literature.
 
 Associative learning is a key cognitive mechanism that allows individuals to associate
-rewards with environmental stimuli and appropriate behavior [@staddon_Adaptive_2016]. 
-Since it was originally described, associative learning has been found in all major
-vertebrate taxon, and many invertebrates [@heyes_Simple_2012;
-@macphail_Brain_1982; @staddon_Adaptive_2016; @behrens_Associative_2008].
-Using these associations to respond to environmental changes is particularly adaptive in
-complex environments [@dridi_Environmental_2016]. Besides its wide taxonomic and
-ecological relevance, associative learning has shown to be a flexible cognitive mechanism
-that can be tuned to different roles depending on the ecological situation 
-[@enquist_Power_2016; @quinones_Reinforcement_2019]. Despite all this, associative 
-learning is not often included in evolutionary explanations 
-[@fawcett_Exposing_2013; @kamil_Optimal_1983; @mcauliffe_Psychology_2015]. 
+rewards with environmental stimuli and thus behave adaptively [@staddon_Adaptive_2016]. 
+Associative learning exists in all major vertebrate taxa, and in many invertebrates [@heyes_Simple_2012;@macphail_Brain_1982; @staddon_Adaptive_2016; @behrens_Associative_2008].
+Theory has shown that natural selection favors these associations in complex environments
+where conditions are hard to predict [@dridi_Environmental_2016]. Besides its wide taxonomic and
+ecological relevance, associative learning is a flexible cognitive mechanism whose underpinings
+show interspecific variation [@enquist_Power_2016; @quinones_Reinforcement_2019], so 
+presumably, it has been modified by natural selection. Despite all this, associative 
+learning is not often included in the narrative of evolutionary explanations of 
+behavioural patterns [@fawcett_Exposing_2013; @kamil_Optimal_1983; @mcauliffe_Psychology_2015]. 
 Computational models of evolution can play an important role to overcome the 
 lack of integration between learning and evolution. Reinforcement learning theory 
-encompass a series of computational methods inspired on the psychological and 
+encompases a series of computational methods inspired on the psychological and 
 neurological mechanisms of associative learning [@sutton_Reinforcement_2018].
@@
 capturing the essence of learning processes [@frankenhuis_Enriching_2018; @quinones_Reinforcement_2019]. 
-Furthermore, these algorithms can be embedded in evolutionary simulations in order 
-to have theoretical predictions of the effect that learning can have in behavioral [@leimar_Learning_2019]
-and morphological evolution.
+Furthermore, these algorithms can be embedded in evolutionary simulations to 
+generate theoretical predictions of the effect that learning can have in behavioral
+evolution [@leimar_Learning_2019].
 
@@
 set up, individuals can use the badge both as a signal of quality (BS)
-or as a trait for IR. We use the model to asses under what conditions of interaction
-structure we expect different communication signals evolve.
+or as a trait for IR. We use the model to assess under what conditions of interaction
+structure we expect different communication signals to evolve.
 

### 2020-10-22 Presentations/EvolvertPresent.tex
New folder to contain presentation and manuscripts. Relevent files moved accordingly. Code needs to be updated to this change. New presentaion started for AGA symposium. New simulations varyinh alphaAct. Need to check why individuals of different quality are not asssorting in nIntGroupEvol4.

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  ignorenonframetext,
]{beamer}
\usepackage{pgfpages}
\setbeamertemplate{caption}[numbered]
\setbeamertemplate{caption label separator}{: }
\setbeamercolor{caption name}{fg=normal text.fg}
\beamertemplatenavigationsymbolsempty
% Prevent slide breaks in the middle of a paragraph
\widowpenalties 1 10000
\raggedbottom
\setbeamertemplate{part page}{
  \centering
  \begin{beamercolorbox}[sep=16pt,center]{part title}
    \usebeamerfont{part title}\insertpart\par
  \end{beamercolorbox}
}
\setbeamertemplate{section page}{
  \centering
  \begin{beamercolorbox}[sep=12pt,center]{part title}
    \usebeamerfont{section title}\insertsection\par
  \end{beamercolorbox}
}
\setbeamertemplate{subsection page}{
  \centering
  \begin{beamercolorbox}[sep=8pt,center]{part title}
    \usebeamerfont{subsection title}\insertsubsection\par
  \end{beamercolorbox}
}
\AtBeginPart{
  \frame{\partpage}
}
\AtBeginSection{
  \ifbibliography
  \else
    \frame{\sectionpage}
  \fi
}
\AtBeginSubsection{
  \frame{\subsectionpage}
}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Badges of status and Individual Recognition mediated by learning},
  pdfauthor={Andrés Quiñones},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\newif\ifbibliography
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{float} \usepackage{multicol} \usepackage{amsmath} \usepackage{graphicx} \usepackage{array}

\title{Badges of status and Individual Recognition mediated by learning}
\author{Andrés Quiñones}
\date{September 15, 2020}

\begin{document}
\frame{\titlepage}

\begin{frame}

\end{frame}

\hypertarget{results}{%
\section{Results}\label{results}}

\begin{frame}{Learning under honest signalling}
\protect\hypertarget{learning-under-honest-signalling}{}

Learning simulations (without evolution yet), the sender code is set as
an honest signal. Receivers develop a threshold like response according
to their quality

\begin{center}\includegraphics[width=1\linewidth]{Simulations/alphaAct_/alphaAct0.6learnDyn} \end{center}

\end{frame}

\begin{frame}{Learning can mediate the evolution of Badges of Status}
\protect\hypertarget{learning-can-mediate-the-evolution-of-badges-of-status}{}

\begin{block}{Variation among replicates}

\small

set of simulations with the badge working as a handicap (cost inversely
proportional to quality)

\begin{center}\includegraphics[width=1\linewidth]{Simulations/betCostEvol1_/evolDynAll_betCost5} \end{center}

\end{block}

\end{frame}

\begin{frame}{Learning can mediate the evolution of Badges of Status}
\protect\hypertarget{learning-can-mediate-the-evolution-of-badges-of-status-1}{}

\begin{block}{Individual replicates: Badges}

\small

One of the replicates of the previous slide, the badge as a handicap

\begin{center}\includegraphics[width=1\linewidth]{Simulations/betCostEvol1_/evolDyn0_betCost5} \end{center}

\end{block}

\end{frame}

\begin{frame}{Learning can mediate the evolution of Badges of Status}
\protect\hypertarget{learning-can-mediate-the-evolution-of-badges-of-status-2}{}

\begin{block}{Individual replicates: Badges}

\small

Panels show the changes in frequency distribution along evolution
(\(\alpha\), \(\beta\), and resulting badge respectively) for the
replicate shown in the previous slide

\begin{figure}

\includegraphics[height=0.28\textheight]{Simulations/betCostEvol1_/evolDistAlpha0_betCost5} \includegraphics[height=0.28\textheight]{Simulations/betCostEvol1_/evolDistBeta0_betCost5} \includegraphics[height=0.28\textheight]{Simulations/betCostEvol1_/evolDistBadge0_betCost5} \hfill{}

\caption{alpha - beta -Badge}\label{fig:unnamed-chunk-4}
\end{figure}

\end{block}

\end{frame}

\begin{frame}{Learning can mediate the evolution of Badges of Status}
\protect\hypertarget{learning-can-mediate-the-evolution-of-badges-of-status-3}{}

\begin{block}{Individual replicates: no badges}

\small

One of the replicates where the badge did not evolve

\begin{center}\includegraphics[width=1\linewidth]{Simulations/betCostEvol1_/evolDyn5_betCost5} \end{center}

\end{block}

\end{frame}

\begin{frame}{The effect of costs on Badges of Status}
\protect\hypertarget{the-effect-of-costs-on-badges-of-status}{}

\begin{block}{Behavioural interactions}

\small

Colours correspond to replicates, each point is one snapshot on the
second half of the simulations. Higher costs seem to favour the
signalling equilibrium

\begin{center}\includegraphics[width=0.8\linewidth]{Simulations/betCostEvol1_/BehavIntALL} \end{center}

\end{block}

\end{frame}

\begin{frame}{The effect of costs on Badges of Status}
\protect\hypertarget{the-effect-of-costs-on-badges-of-status-1}{}

\begin{block}{Badge size}

\small

Same as previous slide but for the resulting signal. Higher costs yield
smaller averga sizes and smaller ranges.

\begin{center}\includegraphics[width=0.8\linewidth]{Simulations/betCostEvol1_/CueALL} \end{center}

\end{block}

\end{frame}

\begin{frame}{The evolution of cheap Badges of Status}
\protect\hypertarget{the-evolution-of-cheap-badges-of-status}{}

\begin{block}{Variation among replicates}

\small

Here simulations where the signal has no cost

\begin{center}\includegraphics[width=0.8\linewidth]{Simulations/nIntGroupEvol1_/evolDynALL_nIntGroup2000} \end{center}

\end{block}

\end{frame}

\begin{frame}{The evolution of cheap Badges of Status}
\protect\hypertarget{the-evolution-of-cheap-badges-of-status-1}{}

\begin{block}{Variation within a replicate}

\small

Here simulations where the signal has no cost.

There seems to be some consistency in the three type of sender codes

\begin{center}\includegraphics[width=0.8\linewidth]{Simulations/nIntGroupEvol1_/evolDyn8_nIntGroup2000} \end{center}

\end{block}

\end{frame}

\begin{frame}{The evolution of cheap Badges of Status}
\protect\hypertarget{the-evolution-of-cheap-badges-of-status-2}{}

\begin{block}{Variation within a replicate}

\tiny

Panels show the changes in frequency distribution along evolution
(\(\alpha\), \(\beta\), and resulting badge respectively) for the
replicate shown in the previous slide

Branching events in the reaction norm parameters yield three types of
sender codes. One (more or less) honest, two dishonest (one with a big
badge and one with a small badge)

\begin{figure}

\includegraphics[height=0.28\textheight]{Simulations/nIntGroupEvol1_/evolDistAlpha8_nIntGroup2000} \includegraphics[height=0.28\textheight]{Simulations/nIntGroupEvol1_/evolDistBeta8_nIntGroup2000} \includegraphics[height=0.28\textheight]{Simulations/nIntGroupEvol1_/evolDistBadge8_nIntGroup2000} \hfill{}

\caption{alpha - beta -Badge}\label{fig:unnamed-chunk-10}
\end{figure}

\end{block}

\end{frame}

\begin{frame}{The evolution of cheap Badges of Status}
\protect\hypertarget{the-evolution-of-cheap-badges-of-status-3}{}

\begin{block}{How often does it happen?}

\tiny

Values of \(\alpha\) and \(\beta\) for each individual in the different
replicates. Each dot an individual, each panel a replicate. Individuals
are sampled from populations from the second half of the simulations.

Individuals in some simulations split up in three groups, others in two
groups; in one simulation the population does not split up

\begin{flushleft}\includegraphics[width=0.9\linewidth]{Simulations/nIntGroupEvol1_/corrAlphBet_nIntGroup2000} \end{flushleft}

\end{block}

\end{frame}

\begin{frame}{The evolution of cheap Badges of Status}
\protect\hypertarget{the-evolution-of-cheap-badges-of-status-4}{}

\begin{block}{Does it matter?}

\small

Behavioural interactions splitting up the replicates according to the
number of ``types'' that evolve

\begin{flushleft}\includegraphics[width=0.9\linewidth]{Simulations/nIntGroupEvol1_/BehavIntAllClusters_nIntGroup2000} \end{flushleft}

\end{block}

\end{frame}

\begin{frame}{Interaction structures - do they matter?}
\protect\hypertarget{interaction-structures---do-they-matter}{}

\begin{center}\includegraphics[width=1\linewidth]{Images/networks} \end{center}

(Chaine et al. 2018)

\end{frame}

\begin{frame}{Localized interaction structures}
\protect\hypertarget{localized-interaction-structures}{}

\begin{block}{Costly signals}

\tiny

Behavioural interactions and badge size for simulations with costs where
individuals interact in small groups. On the \(x\) axes number of
different individuals each focal interacts with.

\begin{center}\includegraphics[width=0.9\linewidth]{Simulations/nIntGroupCost_/BehavIntALL} \includegraphics[width=0.9\linewidth]{Simulations/nIntGroupCost_/CueALL} \end{center}

\end{block}

\end{frame}

\begin{frame}{Localized interaction structures}
\protect\hypertarget{localized-interaction-structures-1}{}

\begin{block}{Cheap signals (free)}

\small

Variation among individuals in different replicates (just like before)
in replicates where individuals interact with 10 other individuals.

A part from replicates with two clusters, the clustering seems less
straightforward as before

\begin{center}\includegraphics[width=0.85\linewidth]{Simulations/nIntGroupEvol1_/corrAlphBet_nIntGroup10} \end{center}

\end{block}

\end{frame}

\begin{frame}{Localized interaction structures}
\protect\hypertarget{localized-interaction-structures-2}{}

\begin{block}{Cheap signals (free)}

\small

Behavioural interactions for the replicates form previous slide. I split
up the replicates according to clusters (despite clusters not being
clear cut).

Only replicates with only one cluster seems to have less aggressive
interactions

\begin{center}\includegraphics[width=0.9\linewidth]{Simulations/nIntGroupEvol1_/BehavIntAllClusters_nIntGroup10} \end{center}

\end{block}

\end{frame}

\begin{frame}{Localized interaction structures}
\protect\hypertarget{localized-interaction-structures-3}{}

\begin{block}{Cheap signals or Individual recognition?}

\tiny

In these set, I changed the learning set up, I increased the number of
centers (10 now, 6 before) along the signal axis; and I reduced the
level or generalization.

There seems to be an effect on the slope of the reaction norm. It stays
very close to zero.

\begin{center}\includegraphics[width=0.9\linewidth]{Simulations/nIntGroupCenters1_/evolDynALL_nIntGroup10} \end{center}

\end{block}

\end{frame}

\begin{frame}{Localized interaction structures}
\protect\hypertarget{localized-interaction-structures-4}{}

\begin{block}{Cheap signals or Individual recognition}

\small

Same replicates as the previous slide. Variation in \(\beta\) and
\(\alpha\) among individuals for different replicates.

No clear clustering. Variation mostly along the \(\alpha\) axis

\begin{center}\includegraphics[width=0.9\linewidth]{Simulations/nIntGroupCenters1_/corrAlphBet_nIntGroup10} \end{center}

\end{block}

\end{frame}

\begin{frame}{Localized interaction structures}
\protect\hypertarget{localized-interaction-structures-5}{}

\begin{block}{Cheap signals or Individual recognition?}

\small

Panels show the changes in frequency distribution along evolution
(\(\alpha\), \(\beta\), and resulting badge respectively) for the
replicate shown in the previous slide.

\(\beta\) around zero, \(\alpha\) all over the place.

\begin{flushleft}\includegraphics[height=0.28\textheight]{Simulations/nIntGroupCenters1_/evolDistAlpha5_nIntGroup10} \includegraphics[height=0.28\textheight]{Simulations/nIntGroupCenters1_/evolDistBeta5_nIntGroup10} \includegraphics[height=0.28\textheight]{Simulations/nIntGroupCenters1_/evolDistBadge5_nIntGroup10} \end{flushleft}

\end{block}

\end{frame}

\begin{frame}{Why would you want to be recognized?}
\protect\hypertarget{why-would-you-want-to-be-recognized}{}

\pause

\begin{block}{The peaceful, the aggressive and the clever}

\begin{itemize}
\tightlist
\item
  Learning initial conditions
\end{itemize}

\tiny

Here, I vary the aggressive tendency for unlearned individuals in these
three set ups.

Agressive, peaceful, clever (starts at the ESS of the classic hawk-dove
game)

\begin{flushleft}\includegraphics[width=0.9\linewidth]{Images/cartoonRBF_InitAct} \end{flushleft}

\end{block}

\end{frame}

\begin{frame}{Why would you want to be recognized?}
\protect\hypertarget{why-would-you-want-to-be-recognized-1}{}

\begin{block}{The peaceful}

Variation among replicates

\begin{flushleft}\includegraphics[width=1\linewidth]{Simulations/initAct_/evolDynALL_initAct2} \end{flushleft}

\end{block}

\end{frame}

\begin{frame}{Why would you want to be recognized?}
\protect\hypertarget{why-would-you-want-to-be-recognized-2}{}

\begin{block}{The peaceful}

Variation within one replicate

\begin{flushleft}\includegraphics[width=1\linewidth]{Simulations/initAct_/evolDyn2_initAct2} \end{flushleft}

\end{block}

\end{frame}

\begin{frame}{Why would you want to be recognized?}
\protect\hypertarget{why-would-you-want-to-be-recognized-3}{}

\begin{block}{The peaceful}

Changes in the frequency distributions

\begin{flushleft}\includegraphics[height=0.28\textheight]{Simulations/initAct_/evolDistAlpha2_initAct2} \includegraphics[height=0.28\textheight]{Simulations/initAct_/evolDistBeta2_initAct2} \includegraphics[height=0.28\textheight]{Simulations/initAct_/evolDistBadge2_initAct2} \end{flushleft}

\end{block}

\end{frame}

\begin{frame}{Why would you want to be recognized?}
\protect\hypertarget{why-would-you-want-to-be-recognized-4}{}

\begin{block}{The peaceful}

Favours variation and cheap signals

\begin{flushleft}\includegraphics[width=1\linewidth]{Simulations/initAct_/corrAlphBet_initAct2} \end{flushleft}

\end{block}

\end{frame}

\begin{frame}{Why would you want to be recognized?}
\protect\hypertarget{why-would-you-want-to-be-recognized-5}{}

\begin{block}{The aggressive}

Variation among replicates

\begin{flushleft}\includegraphics[width=1\linewidth]{Simulations/initAct_/evolDynALL_initAct-2} \end{flushleft}

\end{block}

\end{frame}

\begin{frame}{Why would you want to be recognized?}
\protect\hypertarget{why-would-you-want-to-be-recognized-6}{}

\begin{block}{The aggressive}

Variation within one replicate

\begin{flushleft}\includegraphics[width=1\linewidth]{Simulations/initAct_/evolDyn0_initAct-2} \end{flushleft}

\end{block}

\end{frame}

\begin{frame}{Why would you want to be recognized?}
\protect\hypertarget{why-would-you-want-to-be-recognized-7}{}

\begin{block}{The aggressive}

Disfavours variation and signal evolution - no signalling

\begin{flushleft}\includegraphics[width=1\linewidth]{Simulations/initAct_/corrAlphBet_initAct-2} \end{flushleft}

\end{block}

\end{frame}

\begin{frame}{Why would you want to be recognized?}
\protect\hypertarget{why-would-you-want-to-be-recognized-8}{}

\begin{block}{The clever (ESS)}

Variation among replicates

\begin{flushleft}\includegraphics[width=1\linewidth]{Simulations/initAct_/evolDynALL_initAct-0.69} \end{flushleft}

\end{block}

\end{frame}

\begin{frame}{Why would you want to be recognized?}
\protect\hypertarget{why-would-you-want-to-be-recognized-9}{}

\begin{block}{The clever (ESS)}

Variation within one replicate

\begin{flushleft}\includegraphics[width=1\linewidth]{Simulations/initAct_/evolDyn0_initAct-0.69} \end{flushleft}

\end{block}

\end{frame}

\begin{frame}{Why would you want to be recognized?}
\protect\hypertarget{why-would-you-want-to-be-recognized-10}{}

\begin{block}{The clever (ESS)}

Signal evolution seems to be driven by drift. Most replicates end up in
one of the two extremes

\begin{flushleft}\includegraphics[width=1\linewidth]{Simulations/initAct_/corrAlphBet_initAct-0.69} \end{flushleft}

\end{block}

\end{frame}

\begin{frame}{Take-home messages}
\protect\hypertarget{take-home-messages}{}

\begin{block}{Costly signals}

\begin{itemize}
\tightlist
\item
  Learning can mediate costly signals
\item
  More expensive signals seem be more likely to evolve
\item
  Less expensive signals provide more information
\item
  Local interactions do not impede the evolution of costly signals (as
  has been suggested)
\end{itemize}

\end{block}

\begin{block}{Cheap signals}

\begin{itemize}
\tightlist
\item
  Learning mediates the evolution of polymorphisms with ``honest'' and
  ``dishonest'' individuals
\item
  Under local interactions learning seems to allow the evolution of IR
\item
  Cheap signals depend on the innate level of aggression (before
  learning) of individuals
\item
  Peace begets heterogeneity, aggression begets homogeneity
\end{itemize}

\end{block}

\end{frame}

\begin{frame}{Questions?}
\protect\hypertarget{questions}{}

\begin{center}\includegraphics[height=0.8\textheight]{Images/BizarroParrot} \end{center}

\end{frame}

\begin{frame}{References}
\protect\hypertarget{references}{}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-chaine_Manipulating_2018}{}%
Chaine, Alexis S., Daizaburo Shizuka, Theadora A. Block, Lynn Zhang, and
Bruce E. Lyon. 2018. ``Manipulating Badges of Status Only Fools
Strangers.'' \emph{Ecology Letters} 21 (10): 1477--85.
\url{https://doi.org/10.1111/ele.13128}.

\end{frame}

\end{document}

### 2020-10-22 Presentations/GenesAsEnviron.Rmd
New folder to contain presentation and manuscripts. Relevent files moved accordingly. Code needs to be updated to this change. New presentaion started for AGA symposium. New simulations varyinh alphaAct. Need to check why individuals of different quality are not asssorting in nIntGroupEvol4.

---
title: | 
    | Learning can mediate the evolution of 
    | (cheap and costly) signals of quality
author: "Andrés Quiñones PhD."
bibliography: ../SocialMorphCue.bib
date: | 
    | 
    | 
    | Genes as environment - AGA2020 virtual symposium
    | `r format(Sys.time(), '%B %d, %Y')`
output: 
    beamer_presentation: 
      fig_caption: yes
      keep_tex: yes
header-includes: 
    \usepackage{float}
    \usepackage{multicol}
    \usepackage{amsmath} 
    \usepackage{graphicx}
    \usepackage{array}
---

## Social interactions

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source("../../R_files/posPlots.R")
require("here")
source("../AccFunc.R")

```

### The need partners with actions

```{r,echo=FALSE,fig.align="center",out.width='40%',fig.show='hold'}

knitr::include_graphics("../Images/sexSelec.jpg")
knitr::include_graphics("../Images/begging.jpg")
knitr::include_graphics("../Images/broodparasite.jpg")
knitr::include_graphics("../Images/dominance.jpg")
```

## The effect of communication

```{r,echo=FALSE,fig.align="center",out.width='80%'}
knitr::include_graphics("../Images/harm.jpg")
```

## What about signals?

### Badge of status

```{r,echo=FALSE,fig.align='center',out.height='80%'}
knitr::include_graphics("../Images//hosp_alexandra_mackenzie.jpg")

```

## What about signals?

### When are signals honest?

* Impossible to fake
* Individuals have common interests
* Handicap principle (signal´s cost is proportional to quality)

## What about signals?

### Individual recognition

```{r,echo=FALSE,fig.align='center',out.height='70%'}
knitr::include_graphics("../Images//indRecog.png")

```
[@tibbetts_Individual_2007]

## What about signals?

### BS or IR?

```{r,echo=FALSE,fig.align='center',out.height='75%'}
knitr::include_graphics("../Images//IRvsBS.png")

```
[@sheehan_There_2016]

## What about signals?

```{r,echo=FALSE,fig.align='center',out.width='90%'}
knitr::include_graphics("../Images//comSystemBot.png")
```

## What about learning?

```{r,echo=FALSE,fig.align='center',out.width='90%'}
knitr::include_graphics("../Images//pavlov.jpg")
```

# Associative learning

## Reinforcement learning theory

\begin{columns}[T]
 \column{0.5\linewidth}
 \begin{equation*}
  \Delta V_{t(s)}=\alpha \underbrace{(R_t-V_t)}_\text{prediction error}
  \end{equation*}
  \pause
  \includegraphics[width=1\textwidth]{../Images//learn_curve.png}
 \column{0.5\linewidth}
  \pause
  \includegraphics[height=.9\textheight]{../Images//predErrorNeur.png}
\end{columns}

## Reinforcement learning theory

```{r,echo=FALSE,fig.align='center',out.width='90%'}
knitr::include_graphics("../Images//learnUpdate.png")
```

## The Hawk-Dove game

Individuals have one of two genetically determined phenotypic strategies. _Hawks_ are willing to start a conflict over resources, while _doves_ prefer to stand down in the hope to share the resource without an aggressive contest.

```{r,echo=FALSE,fig.align="center",out.width='50%'}
knitr::include_graphics("../Images//HDgame.png")
```

## The hawk-dove game

\begin{align*}
w_H &= p_H \frac{V-C}{2}+(1-p_h) V\\
w_D &= p_H 0 + (1-p_H)\frac{V}{2}
\end{align*}
\vspace{-0.8cm}
```{r, echo=FALSE,fig.align="center", out.width = '55%'}
knitr::include_graphics("../Images//Hawk_dove_plot.png")
```

<!-- ## The hawk-dove game -->

<!-- ```{r, echo=FALSE, fig.cap="\\label{fig:HD_game}Hawk-dove game. Dashed line is the  game theoretical prediction for frequency of hawks.", out.width = '80%'} -->
<!-- knitr::include_graphics("../Simulations/mutType_/basicHawkDove.png") -->
<!-- ``` -->

## Environmental states

### Discrete states
\includegraphics[width=0.8\textwidth]{../Images//discreteStates.png}

\pause

### Continuos states
\includegraphics[width=0.8\textwidth]{../Images//contStates.png}

## Continuos environmental states
\begin{equation*}
  \Delta V_{t(s)}=\alpha \underbrace{(R_t-V_t)}_\text{prediction error}
\end{equation*}

```{r fig2, echo=FALSE, fig.cap="\\label{fig:learning_cartoonRBF}Function approximation for the actor and the critic. Using random values for the responses of each center. Circles show the location of the centers and the maximum response they trigger, while the lines show the total response along the badge size continium.", out.width = '60%'}
knitr::include_graphics("../Images/cartoonRBF.png")
```

# Our model

## The hawk-dove game 2.0

\begin{center}
\begin{tabular}{ >{\centering\arraybackslash}p{1cm} | >{\centering\arraybackslash}p{4.5cm} | >{\centering\arraybackslash}p{4.5cm} }
& H & D \\ \hline
H & $p_w V\frac{-C}{2} + (1-p_w) \frac{-C}{2}$ & $V$ \\ \hline
V & $0$ & $\frac{V}{2}$\\
\end{tabular}
\end{center}

\begin{equation*}
p_w=\frac{1}{1+e^{-\beta(Q_i-Q_j)}}
\end{equation*}

```{r,echo=FALSE,fig.align='center',out.width='50%'}
par(plt=posPlot(),las=1)
plot(logist(seq(-1,1,length.out = 1000),alpha = 0,beta = 2)~
       seq(-1,1,length.out = 1000),type="l",xlab="Difference in quality",
     ylab = expression(p[w]),cex.lab=2,lwd=3,col="red",cex.axis=2,ylim=c(0,1))
lines(logist(seq(-1,1,length.out = 1000),alpha = 0,beta = 10)~
       seq(-1,1,length.out = 1000),lwd=3,col="green")
legend("topleft",legend=c(2,10),col=c("red","green"),lwd=3,title = expression(beta),cex=2,bty="n")

```

## Sender Code

\begin{equation*}
s_i = \frac{1}{1+e^{-(\epsilon_i+\gamma_iQ_i)}}
\label{eq:react_norm}
\end{equation*}

```{r,echo=FALSE,fig.align='center',out.width='50%'}
par(plt=posPlot(),las=1)
plot(logist(seq(0,1,length.out = 1000),alpha = 2,beta = 4)~
       seq(0,1,length.out = 1000),type="l",xlab="Quality",
     ylab = expression(p[w]),cex.lab=2,lwd=3,col="red",cex.axis=2,ylim=c(0,1))
lines(logist(seq(0,1,length.out = 1000),alpha = 5,beta = 10)~
       seq(0,1,length.out = 1000),lwd=3,col="green")
legend("topleft",legend=c(4,10),col=c("red","green"),lwd=3,title = expression(gamma),cex=2,bty="n")

```

## Receiver code

\begin{equation*}
  \Delta V_{t(s)}=\alpha (R_t-V_t)
\end{equation*}

```{r, echo=FALSE,fig.align='center', out.width = '60%'}
knitr::include_graphics("../Images/cartoonRBF.png")
```

# Results

## Learning under honest signalling

Learning simulations (without evolution yet), the sender code is set as an honest signal.
Receivers develop a threshold like response according to their quality

```{r, echo=FALSE,fig.align='center', out.width = '100%'}
knitr::include_graphics("../Simulations/alphaAct_/alphaAct0.6learnDyn.png")
```

## Learning can mediate the evolution of Badges of Status
### Variation among replicates
\small
set of simulations with the badge working as a handicap (cost inversely proportional to quality)

```{r, echo=FALSE,fig.align='center', out.width='100%'}
knitr::include_graphics("../Simulations/betCostEvol1_/evolDynAll_betCost5.png")
```

## Learning can mediate the evolution of Badges of Status
### Individual replicates: Badges
\small
One of the replicates of the previous slide, the badge as a handicap
```{r, echo=FALSE,fig.align='center', out.width='100%'}
knitr::include_graphics("../Simulations/betCostEvol1_/evolDyn0_betCost5.png")
```

## Learning can mediate the evolution of Badges of Status
### Individual replicates: Badges
\small
Panels show the changes in frequency distribution along evolution ($\alpha$, $\beta$, and resulting badge respectively) for the replicate shown in the previous slide

```{r, echo=FALSE,fig.align='left', out.height='28%',fig.show='hold', fig.cap="alpha - beta -Badge"}
knitr::include_graphics("../Simulations/betCostEvol1_/evolDistAlpha0_betCost5.png")
knitr::include_graphics("../Simulations/betCostEvol1_/evolDistBeta0_betCost5.png")
knitr::include_graphics("../Simulations/betCostEvol1_/evolDistBadge0_betCost5.png")
```

## Learning can mediate the evolution of Badges of Status
### Individual replicates: no badges
\small
One of the replicates where the badge did not evolve

```{r, echo=FALSE,fig.align='center', out.width='100%'}
knitr::include_graphics("../Simulations/betCostEvol1_/evolDyn5_betCost5.png")
```

## The effect of costs on Badges of Status
### Behavioural interactions
\small
Colours correspond to replicates, each point is one snapshot on the second half of the simulations.
Higher costs seem to favour the signalling equilibrium

```{r, echo=FALSE,fig.align='center', out.width='80%'}
knitr::include_graphics("../Simulations/betCostEvol1_/BehavIntALL.png")
```
## The effect of costs on Badges of Status
### Badge size
\small
Same as previous slide but for the resulting signal. Higher costs yield smaller averga sizes and smaller ranges.

```{r, echo=FALSE,fig.align='center', out.width='80%'}
knitr::include_graphics("../Simulations/betCostEvol1_/CueALL.png")
```

## The evolution of cheap Badges of Status
### Variation among replicates
\small
Here simulations where the signal has no cost

```{r, echo=FALSE,fig.align='center', out.width='80%'}
knitr::include_graphics("../Simulations/nIntGroupEvol1_/evolDynALL_nIntGroup2000.png")
```

## The evolution of cheap Badges of Status
### Variation within a replicate
\small
Here simulations where the signal has no cost.

There seems to be some consistency in the three type of sender codes

```{r, echo=FALSE,fig.align='center', out.width='80%'}
knitr::include_graphics("../Simulations/nIntGroupEvol1_/evolDyn8_nIntGroup2000.png")
```

## The evolution of cheap Badges of Status
### Variation within a replicate

\tiny
Panels show the changes in frequency distribution along evolution ($\alpha$, $\beta$, and resulting badge respectively) for the replicate shown in the previous slide

Branching events in the reaction norm parameters yield three types of sender codes. One (more or less)
honest, two dishonest (one with a big badge and one with a small badge)

```{r, echo=FALSE,fig.align='left', out.height='28%',fig.show='hold', fig.cap="alpha - beta -Badge"}
knitr::include_graphics("../Simulations/nIntGroupEvol1_/evolDistAlpha8_nIntGroup2000.png")
knitr::include_graphics("../Simulations/nIntGroupEvol1_/evolDistBeta8_nIntGroup2000.png")
knitr::include_graphics("../Simulations/nIntGroupEvol1_/evolDistBadge8_nIntGroup2000.png")
```

## The evolution of cheap Badges of Status
### How often does it happen?
\tiny
Values of $\alpha$ and $\beta$ for each individual in the different replicates. Each dot an individual,
each panel a replicate. Individuals are sampled from populations from the second half of the simulations.

Individuals in some simulations split up in three groups, others in two groups; 
in one simulation the population does not split up

```{r, echo=FALSE,fig.align='left', out.width='90%',fig.show='hold'}
knitr::include_graphics("../Simulations/nIntGroupEvol1_/corrAlphBet_nIntGroup2000.png")
```

## The evolution of cheap Badges of Status
### Does it matter?
\small
Behavioural interactions splitting up the replicates according to the number of "types" that 
evolve

```{r, echo=FALSE,fig.align='left', out.width='90%',fig.show='hold'}
knitr::include_graphics("../Simulations/nIntGroupEvol1_/BehavIntAllClusters_nIntGroup2000.png")
```

## Interaction structures - do they matter?

```{r, echo=FALSE,fig.align='center', out.width='100%',}
knitr::include_graphics("../Images/networks.png")
```
[@chaine_Manipulating_2018]

## Localized interaction structures

### Costly signals 
\tiny
Behavioural interactions and badge size for simulations with costs where individuals interact in small groups. On the $x$ axes number of different individuals each focal interacts with.

```{r, echo=FALSE,fig.align='center', out.width='90%',fig.show='hold'}
knitr::include_graphics("../Simulations/nIntGroupCost_/BehavIntALL.png")
knitr::include_graphics("../Simulations/nIntGroupCost_/CueALL.png")
```

## Localized interaction structures

### Cheap signals (free)
\small
Variation among individuals in different replicates (just like before) in replicates where individuals interact with 10 other individuals.

A part from replicates with two clusters, the clustering seems less straightforward as before

```{r, echo=FALSE,fig.align='center', out.width='85%',fig.show='hold'}
knitr::include_graphics("../Simulations/nIntGroupEvol1_/corrAlphBet_nIntGroup10.png")
```

## Localized interaction structures

### Cheap signals (free)

\small
Behavioural interactions for the replicates form previous slide. I split up the replicates according to clusters (despite clusters not being clear cut).

Only replicates with only one cluster seems to have less aggressive interactions

```{r, echo=FALSE,fig.align='center', out.width='90%',fig.show='hold'}
knitr::include_graphics("../Simulations/nIntGroupEvol1_/BehavIntAllClusters_nIntGroup10.png")
```

## Localized interaction structures

### Cheap signals or Individual recognition?
\tiny
In these set, I changed the learning set up, I increased the number of centers (10 now, 6 before)
along the signal axis; and I reduced the level or generalization.

There seems to be an effect on the slope of the reaction norm. It stays very close to zero.

```{r, echo=FALSE,fig.align='center', out.width='90%',fig.show='hold'}
knitr::include_graphics("../Simulations/nIntGroupCenters1_/evolDynALL_nIntGroup10.png")
```

## Localized interaction structures

### Cheap signals or Individual recognition
\small
Same replicates as the previous slide. Variation in $\beta$ and $\alpha$ among individuals for different replicates.

No clear clustering. Variation mostly along the $\alpha$ axis

```{r, echo=FALSE,fig.align='center', out.width='90%',fig.show='hold'}
knitr::include_graphics("../Simulations/nIntGroupCenters1_/corrAlphBet_nIntGroup10.png")
```
## Localized interaction structures

### Cheap signals or Individual recognition?
\small
Panels show the changes in frequency distribution along evolution ($\alpha$, $\beta$, and resulting badge respectively) for the replicate shown in the previous slide.

$\beta$ around zero, $\alpha$ all over the place.

```{r, echo=FALSE,fig.align='left', out.height='28%',fig.show='hold'}
knitr::include_graphics("../Simulations/nIntGroupCenters1_/evolDistAlpha5_nIntGroup10.png")
knitr::include_graphics("../Simulations/nIntGroupCenters1_/evolDistBeta5_nIntGroup10.png")
knitr::include_graphics("../Simulations/nIntGroupCenters1_/evolDistBadge5_nIntGroup10.png")
```

<!-- ## Aqui vamos! mirar correlaciones de 20 y 2000 luego interacciones -->

<!-- ## Comparar scenarios -->

## Why would you want to be recognized?

\pause

### The peaceful, the aggressive and the clever

- Learning initial conditions

\tiny
Here, I vary the aggressive tendency for unlearned individuals in these three set ups.

Agressive, peaceful, clever (starts at the ESS of the classic hawk-dove game)

```{r, echo=FALSE,fig.align='left', out.width='90%',fig.show='hold'}
knitr::include_graphics("../Images/cartoonRBF_InitAct.png")
```

## Why would you want to be recognized?

### The peaceful

Variation among replicates

```{r, echo=FALSE,fig.align='left', out.width='100%',fig.show='hold'}
knitr::include_graphics("../Simulations/initAct_/evolDynALL_initAct2.png")
```

## Why would you want to be recognized?

### The peaceful

Variation within one replicate

```{r, echo=FALSE,fig.align='left', out.width='100%',fig.show='hold'}
knitr::include_graphics("../Simulations/initAct_/evolDyn2_initAct2.png")
```

## Why would you want to be recognized?

### The peaceful

Changes in the frequency distributions

```{r, echo=FALSE,fig.align='left', out.height='28%',fig.show='hold'}
knitr::include_graphics("../Simulations/initAct_/evolDistAlpha2_initAct2.png")
knitr::include_graphics("../Simulations/initAct_/evolDistBeta2_initAct2.png")
knitr::include_graphics("../Simulations/initAct_/evolDistBadge2_initAct2.png")
```

## Why would you want to be recognized?

### The peaceful

Favours variation and cheap signals

```{r, echo=FALSE,fig.align='left', out.width='100%',fig.show='hold'}
knitr::include_graphics("../Simulations/initAct_/corrAlphBet_initAct2.png")
```

## Why would you want to be recognized?

### The aggressive

Variation among replicates

```{r, echo=FALSE,fig.align='left', out.width='100%',fig.show='hold'}
knitr::include_graphics("../Simulations/initAct_/evolDynALL_initAct-2.png")
```

## Why would you want to be recognized?

### The aggressive

Variation within one replicate

```{r, echo=FALSE,fig.align='left', out.width='100%',fig.show='hold'}
knitr::include_graphics("../Simulations/initAct_/evolDyn0_initAct-2.png")
```

## Why would you want to be recognized?

### The aggressive

Disfavours variation and signal evolution - no signalling

```{r, echo=FALSE,fig.align='left', out.width='100%',fig.show='hold'}
knitr::include_graphics("../Simulations/initAct_/corrAlphBet_initAct-2.png")
```

## Why would you want to be recognized?

### The clever (ESS)

Variation among replicates

```{r, echo=FALSE,fig.align='left', out.width='100%',fig.show='hold'}
knitr::include_graphics("../Simulations/initAct_/evolDynALL_initAct-0.69.png")
```

## Why would you want to be recognized?

### The clever (ESS)

Variation within one replicate

```{r, echo=FALSE,fig.align='left', out.width='100%',fig.show='hold'}
  knitr::include_graphics("../Simulations/initAct_/evolDyn0_initAct-0.69.png")
```

## Why would you want to be recognized?

### The clever (ESS)

Signal evolution seems to be driven by drift. Most replicates end up in one of the two extremes

```{r, echo=FALSE,fig.align='left', out.width='100%',fig.show='hold'}
  knitr::include_graphics("../Simulations/initAct_/corrAlphBet_initAct-0.69.png")
```

## Take-home messages

### Costly signals
- Learning can mediate costly signals
- More expensive signals seem be more likely to evolve
- Less expensive signals provide more information
- Local interactions do not impede the evolution of costly signals (as has been suggested)

### Cheap signals
- Learning mediates the evolution of polymorphisms with "honest" and "dishonest"
individuals
- Under local interactions learning seems to allow the evolution of IR
- Cheap signals depend on the innate level of aggression (before learning) of individuals
- Peace begets heterogeneity, aggression begets homogeneity

## Questions?

```{r, echo=FALSE,fig.align='center', out.height='80%',fig.show='hold'}
  knitr::include_graphics("../Images/BizarroParrot.jpg")
```

## References

### 2020-10-22 Presentations/GenesAsEnviron.tex
New folder to contain presentation and manuscripts. Relevent files moved accordingly. Code needs to be updated to this change. New presentaion started for AGA symposium. New simulations varyinh alphaAct. Need to check why individuals of different quality are not asssorting in nIntGroupEvol4.

--- 2020-10-22 Presentations/EvolvertPresent.tex
@@
 \hypersetup{
-  pdftitle={Badges of status and Individual Recognition mediated by learning},
-  pdfauthor={Andrés Quiñones},
+  pdfauthor={Andrés Quiñones PhD.},
   hidelinks,
@@
 
-\title{Badges of status and Individual Recognition mediated by learning}
-\author{Andrés Quiñones}
-\date{September 15, 2020}
+\title{Learning can mediate the evolution of\\
+(cheap and costly) signals of quality}
+\author{Andrés Quiñones PhD.}
+\date{Genes as environment - AGA2020 virtual symposium\\
+October 21, 2020}
 
@@
 
-\begin{frame}
+\begin{frame}{Social interactions}
+\protect\hypertarget{social-interactions}{}
+
+\begin{block}{The need partners with actions}
+
+\begin{center}\includegraphics[width=0.4\linewidth]{../Images/sexSelec} \includegraphics[width=0.4\linewidth]{../Images/begging} \includegraphics[width=0.4\linewidth]{../Images/broodparasite} \includegraphics[width=0.4\linewidth]{../Images/dominance} \end{center}
+
+\end{block}
+
+\end{frame}
+
+\begin{frame}{The effect of communication}
+\protect\hypertarget{the-effect-of-communication}{}
+
+\begin{center}\includegraphics[width=0.8\linewidth]{../Images/harm} \end{center}
+
+\end{frame}
+
+\begin{frame}{What about signals?}
+\protect\hypertarget{what-about-signals}{}
+
+\begin{block}{Badge of status}
+
+\begin{center}\includegraphics[height=0.8\textheight]{../Images//hosp_alexandra_mackenzie} \end{center}
+
+\end{block}
+
+\end{frame}
+
+\begin{frame}{What about signals?}
+\protect\hypertarget{what-about-signals-1}{}
+
+\begin{block}{When are signals honest?}
+
+\begin{itemize}
+\tightlist
+\item
+  Impossible to fake
+\item
+  Individuals have common interests
+\item
+  Handicap principle (signal´s cost is proportional to quality)
+\end{itemize}
+
+\end{block}
+
+\end{frame}
+
+\begin{frame}{What about signals?}
+\protect\hypertarget{what-about-signals-2}{}
+
+\begin{block}{Individual recognition}
+
+\begin{center}\includegraphics[height=0.7\textheight]{../Images//indRecog} \end{center}
+
+(Tibbetts and Dale 2007)
+
+\end{block}
+
+\end{frame}
+
+\begin{frame}{What about signals?}
+\protect\hypertarget{what-about-signals-3}{}
+
+\begin{block}{BS or IR?}
+
+\begin{center}\includegraphics[height=0.75\textheight]{../Images//IRvsBS} \end{center}
+
+(Sheehan and Bergman 2016)
+
+\end{block}
+
+\end{frame}
+
+\begin{frame}{What about signals?}
+\protect\hypertarget{what-about-signals-4}{}
+
+\begin{center}\includegraphics[width=0.9\linewidth]{../Images//comSystemBot} \end{center}
+
+\end{frame}
+
+\begin{frame}{What about learning?}
+\protect\hypertarget{what-about-learning}{}
+
+\begin{center}\includegraphics[width=0.9\linewidth]{../Images//pavlov} \end{center}
+
+\end{frame}
+
+\hypertarget{associative-learning}{%
+\section{Associative learning}\label{associative-learning}}
+
+\begin{frame}{Reinforcement learning theory}
+\protect\hypertarget{reinforcement-learning-theory}{}
+
+\begin{columns}[T]
+ \column{0.5\linewidth}
+ \begin{equation*}
+  \Delta V_{t(s)}=\alpha \underbrace{(R_t-V_t)}_\text{prediction error}
+  \end{equation*}
+  \pause
+  \includegraphics[width=1\textwidth]{../Images//learn_curve.png}
+ \column{0.5\linewidth}
+  \pause
+  \includegraphics[height=.9\textheight]{../Images//predErrorNeur.png}
+\end{columns}
+
+\end{frame}
+
+\begin{frame}{Reinforcement learning theory}
+\protect\hypertarget{reinforcement-learning-theory-1}{}
+
+\begin{center}\includegraphics[width=0.9\linewidth]{../Images//learnUpdate} \end{center}
+
+\end{frame}
+
+\begin{frame}{The Hawk-Dove game}
+\protect\hypertarget{the-hawk-dove-game}{}
+
+Individuals have one of two genetically determined phenotypic
+strategies. \emph{Hawks} are willing to start a conflict over resources,
+while \emph{doves} prefer to stand down in the hope to share the
+resource without an aggressive contest.
+
+\begin{center}\includegraphics[width=0.5\linewidth]{../Images//HDgame} \end{center}
+
+\end{frame}
+
+\begin{frame}{The hawk-dove game}
+\protect\hypertarget{the-hawk-dove-game-1}{}
+
+\begin{align*}
+w_H &= p_H \frac{V-C}{2}+(1-p_h) V\\
+w_D &= p_H 0 + (1-p_H)\frac{V}{2}
+\end{align*} \vspace{-0.8cm}
+
+\begin{center}\includegraphics[width=0.55\linewidth]{../Images//Hawk_dove_plot} \end{center}
+
+\end{frame}
+
+\begin{frame}{Environmental states}
+\protect\hypertarget{environmental-states}{}
+
+\begin{block}{Discrete states}
+
+\includegraphics[width=0.8\textwidth]{../Images//discreteStates.png}
+
+\pause
+
+\end{block}
+
+\begin{block}{Continuos states}
+
+\includegraphics[width=0.8\textwidth]{../Images//contStates.png}
+
+\end{block}
+
+\end{frame}
+
+\begin{frame}{Continuos environmental states}
+\protect\hypertarget{continuos-environmental-states}{}
+
+\begin{equation*}
+  \Delta V_{t(s)}=\alpha \underbrace{(R_t-V_t)}_\text{prediction error}
+\end{equation*}
+
+\begin{figure}
+\includegraphics[width=0.6\linewidth]{../Images/cartoonRBF} \caption{\label{fig:learning_cartoonRBF}Function approximation for the actor and the critic.
+Using random values for the responses of each center.
+Circles show the location of the centers and the maximum response they trigger, while the lines show the total response along the badge size continium.}\label{fig:fig2}
+\end{figure}
+
+\end{frame}
+
+\hypertarget{our-model}{%
+\section{Our model}\label{our-model}}
+
+\begin{frame}{The hawk-dove game 2.0}
+\protect\hypertarget{the-hawk-dove-game-2.0}{}
+
+\begin{center}
+\begin{tabular}{ >{\centering\arraybackslash}p{1cm} | >{\centering\arraybackslash}p{4.5cm} | >{\centering\arraybackslash}p{4.5cm} }
+& H & D \\ \hline
+H & $p_w V\frac{-C}{2} + (1-p_w) \frac{-C}{2}$ & $V$ \\ \hline
+V & $0$ & $\frac{V}{2}$\\
+\end{tabular}
+\end{center}
+
+\begin{equation*}
+p_w=\frac{1}{1+e^{-\beta(Q_i-Q_j)}}
+\end{equation*}
+
+\begin{center}\includegraphics[width=0.5\linewidth]{GenesAsEnviron_files/figure-beamer/unnamed-chunk-11-1} \end{center}
+
+\end{frame}
+
+\begin{frame}{Sender Code}
+\protect\hypertarget{sender-code}{}
+
+\begin{equation*}
+s_i = \frac{1}{1+e^{-(\epsilon_i+\gamma_iQ_i)}}
+\label{eq:react_norm}
+\end{equation*}
+
+\begin{center}\includegraphics[width=0.5\linewidth]{GenesAsEnviron_files/figure-beamer/unnamed-chunk-12-1} \end{center}
+
+\end{frame}
+
+\begin{frame}{Receiver code}
+\protect\hypertarget{receiver-code}{}
+
+\begin{equation*}
+  \Delta V_{t(s)}=\alpha (R_t-V_t)
+\end{equation*}
+
+\begin{center}\includegraphics[width=0.6\linewidth]{../Images/cartoonRBF} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{Simulations/alphaAct_/alphaAct0.6learnDyn} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{../Simulations/alphaAct_/alphaAct0.6learnDyn} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{Simulations/betCostEvol1_/evolDynAll_betCost5} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{../Simulations/betCostEvol1_/evolDynAll_betCost5} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{Simulations/betCostEvol1_/evolDyn0_betCost5} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{../Simulations/betCostEvol1_/evolDyn0_betCost5} \end{center}
 
@@
 
-\includegraphics[height=0.28\textheight]{Simulations/betCostEvol1_/evolDistAlpha0_betCost5} \includegraphics[height=0.28\textheight]{Simulations/betCostEvol1_/evolDistBeta0_betCost5} \includegraphics[height=0.28\textheight]{Simulations/betCostEvol1_/evolDistBadge0_betCost5} \hfill{}
+\includegraphics[height=0.28\textheight]{../Simulations/betCostEvol1_/evolDistAlpha0_betCost5} \includegraphics[height=0.28\textheight]{../Simulations/betCostEvol1_/evolDistBeta0_betCost5} \includegraphics[height=0.28\textheight]{../Simulations/betCostEvol1_/evolDistBadge0_betCost5} \hfill{}
 
-\caption{alpha - beta -Badge}\label{fig:unnamed-chunk-4}
+\caption{alpha - beta -Badge}\label{fig:unnamed-chunk-17}
 \end{figure}
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{Simulations/betCostEvol1_/evolDyn5_betCost5} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{../Simulations/betCostEvol1_/evolDyn5_betCost5} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=0.8\linewidth]{Simulations/betCostEvol1_/BehavIntALL} \end{center}
+\begin{center}\includegraphics[width=0.8\linewidth]{../Simulations/betCostEvol1_/BehavIntALL} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=0.8\linewidth]{Simulations/betCostEvol1_/CueALL} \end{center}
+\begin{center}\includegraphics[width=0.8\linewidth]{../Simulations/betCostEvol1_/CueALL} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=0.8\linewidth]{Simulations/nIntGroupEvol1_/evolDynALL_nIntGroup2000} \end{center}
+\begin{center}\includegraphics[width=0.8\linewidth]{../Simulations/nIntGroupEvol1_/evolDynALL_nIntGroup2000} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=0.8\linewidth]{Simulations/nIntGroupEvol1_/evolDyn8_nIntGroup2000} \end{center}
+\begin{center}\includegraphics[width=0.8\linewidth]{../Simulations/nIntGroupEvol1_/evolDyn8_nIntGroup2000} \end{center}
 
@@
 
-\includegraphics[height=0.28\textheight]{Simulations/nIntGroupEvol1_/evolDistAlpha8_nIntGroup2000} \includegraphics[height=0.28\textheight]{Simulations/nIntGroupEvol1_/evolDistBeta8_nIntGroup2000} \includegraphics[height=0.28\textheight]{Simulations/nIntGroupEvol1_/evolDistBadge8_nIntGroup2000} \hfill{}
+\includegraphics[height=0.28\textheight]{../Simulations/nIntGroupEvol1_/evolDistAlpha8_nIntGroup2000} \includegraphics[height=0.28\textheight]{../Simulations/nIntGroupEvol1_/evolDistBeta8_nIntGroup2000} \includegraphics[height=0.28\textheight]{../Simulations/nIntGroupEvol1_/evolDistBadge8_nIntGroup2000} \hfill{}
 
-\caption{alpha - beta -Badge}\label{fig:unnamed-chunk-10}
+\caption{alpha - beta -Badge}\label{fig:unnamed-chunk-23}
 \end{figure}
@@
 
-\begin{flushleft}\includegraphics[width=0.9\linewidth]{Simulations/nIntGroupEvol1_/corrAlphBet_nIntGroup2000} \end{flushleft}
+\begin{flushleft}\includegraphics[width=0.9\linewidth]{../Simulations/nIntGroupEvol1_/corrAlphBet_nIntGroup2000} \end{flushleft}
 
@@
 
-\begin{flushleft}\includegraphics[width=0.9\linewidth]{Simulations/nIntGroupEvol1_/BehavIntAllClusters_nIntGroup2000} \end{flushleft}
+\begin{flushleft}\includegraphics[width=0.9\linewidth]{../Simulations/nIntGroupEvol1_/BehavIntAllClusters_nIntGroup2000} \end{flushleft}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{Images/networks} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{../Images/networks} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=0.9\linewidth]{Simulations/nIntGroupCost_/BehavIntALL} \includegraphics[width=0.9\linewidth]{Simulations/nIntGroupCost_/CueALL} \end{center}
+\begin{center}\includegraphics[width=0.9\linewidth]{../Simulations/nIntGroupCost_/BehavIntALL} \includegraphics[width=0.9\linewidth]{../Simulations/nIntGroupCost_/CueALL} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=0.85\linewidth]{Simulations/nIntGroupEvol1_/corrAlphBet_nIntGroup10} \end{center}
+\begin{center}\includegraphics[width=0.85\linewidth]{../Simulations/nIntGroupEvol1_/corrAlphBet_nIntGroup10} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=0.9\linewidth]{Simulations/nIntGroupEvol1_/BehavIntAllClusters_nIntGroup10} \end{center}
+\begin{center}\includegraphics[width=0.9\linewidth]{../Simulations/nIntGroupEvol1_/BehavIntAllClusters_nIntGroup10} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=0.9\linewidth]{Simulations/nIntGroupCenters1_/evolDynALL_nIntGroup10} \end{center}
+\begin{center}\includegraphics[width=0.9\linewidth]{../Simulations/nIntGroupCenters1_/evolDynALL_nIntGroup10} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=0.9\linewidth]{Simulations/nIntGroupCenters1_/corrAlphBet_nIntGroup10} \end{center}
+\begin{center}\includegraphics[width=0.9\linewidth]{../Simulations/nIntGroupCenters1_/corrAlphBet_nIntGroup10} \end{center}
 
@@
 
-\begin{flushleft}\includegraphics[height=0.28\textheight]{Simulations/nIntGroupCenters1_/evolDistAlpha5_nIntGroup10} \includegraphics[height=0.28\textheight]{Simulations/nIntGroupCenters1_/evolDistBeta5_nIntGroup10} \includegraphics[height=0.28\textheight]{Simulations/nIntGroupCenters1_/evolDistBadge5_nIntGroup10} \end{flushleft}
+\begin{flushleft}\includegraphics[height=0.28\textheight]{../Simulations/nIntGroupCenters1_/evolDistAlpha5_nIntGroup10} \includegraphics[height=0.28\textheight]{../Simulations/nIntGroupCenters1_/evolDistBeta5_nIntGroup10} \includegraphics[height=0.28\textheight]{../Simulations/nIntGroupCenters1_/evolDistBadge5_nIntGroup10} \end{flushleft}
 
@@
 
-\begin{flushleft}\includegraphics[width=0.9\linewidth]{Images/cartoonRBF_InitAct} \end{flushleft}
+\begin{flushleft}\includegraphics[width=0.9\linewidth]{../Images/cartoonRBF_InitAct} \end{flushleft}
 
@@
 
-\begin{flushleft}\includegraphics[width=1\linewidth]{Simulations/initAct_/evolDynALL_initAct2} \end{flushleft}
+\begin{flushleft}\includegraphics[width=1\linewidth]{../Simulations/initAct_/evolDynALL_initAct2} \end{flushleft}
 
@@
 
-\begin{flushleft}\includegraphics[width=1\linewidth]{Simulations/initAct_/evolDyn2_initAct2} \end{flushleft}
+\begin{flushleft}\includegraphics[width=1\linewidth]{../Simulations/initAct_/evolDyn2_initAct2} \end{flushleft}
 
@@
 
-\begin{flushleft}\includegraphics[height=0.28\textheight]{Simulations/initAct_/evolDistAlpha2_initAct2} \includegraphics[height=0.28\textheight]{Simulations/initAct_/evolDistBeta2_initAct2} \includegraphics[height=0.28\textheight]{Simulations/initAct_/evolDistBadge2_initAct2} \end{flushleft}
+\begin{flushleft}\includegraphics[height=0.28\textheight]{../Simulations/initAct_/evolDistAlpha2_initAct2} \includegraphics[height=0.28\textheight]{../Simulations/initAct_/evolDistBeta2_initAct2} \includegraphics[height=0.28\textheight]{../Simulations/initAct_/evolDistBadge2_initAct2} \end{flushleft}
 
@@
 
-\begin{flushleft}\includegraphics[width=1\linewidth]{Simulations/initAct_/corrAlphBet_initAct2} \end{flushleft}
+\begin{flushleft}\includegraphics[width=1\linewidth]{../Simulations/initAct_/corrAlphBet_initAct2} \end{flushleft}
 
@@
 
-\begin{flushleft}\includegraphics[width=1\linewidth]{Simulations/initAct_/evolDynALL_initAct-2} \end{flushleft}
+\begin{flushleft}\includegraphics[width=1\linewidth]{../Simulations/initAct_/evolDynALL_initAct-2} \end{flushleft}
 
@@
 
-\begin{flushleft}\includegraphics[width=1\linewidth]{Simulations/initAct_/evolDyn0_initAct-2} \end{flushleft}
+\begin{flushleft}\includegraphics[width=1\linewidth]{../Simulations/initAct_/evolDyn0_initAct-2} \end{flushleft}
 
@@
 
-\begin{flushleft}\includegraphics[width=1\linewidth]{Simulations/initAct_/corrAlphBet_initAct-2} \end{flushleft}
+\begin{flushleft}\includegraphics[width=1\linewidth]{../Simulations/initAct_/corrAlphBet_initAct-2} \end{flushleft}
 
@@
 
-\begin{flushleft}\includegraphics[width=1\linewidth]{Simulations/initAct_/evolDynALL_initAct-0.69} \end{flushleft}
+\begin{flushleft}\includegraphics[width=1\linewidth]{../Simulations/initAct_/evolDynALL_initAct-0.69} \end{flushleft}
 
@@
 
-\begin{flushleft}\includegraphics[width=1\linewidth]{Simulations/initAct_/evolDyn0_initAct-0.69} \end{flushleft}
+\begin{flushleft}\includegraphics[width=1\linewidth]{../Simulations/initAct_/evolDyn0_initAct-0.69} \end{flushleft}
 
@@
 
-\begin{flushleft}\includegraphics[width=1\linewidth]{Simulations/initAct_/corrAlphBet_initAct-0.69} \end{flushleft}
+\begin{flushleft}\includegraphics[width=1\linewidth]{../Simulations/initAct_/corrAlphBet_initAct-0.69} \end{flushleft}
 
@@
 
-\begin{center}\includegraphics[height=0.8\textheight]{Images/BizarroParrot} \end{center}
+\begin{center}\includegraphics[height=0.8\textheight]{../Images/BizarroParrot} \end{center}
 
@@
 
+\leavevmode\hypertarget{ref-sheehan_There_2016}{}%
+Sheehan, Michael J., and Thore J. Bergman. 2016. ``Is There an
+Evolutionary Trade-Off Between Quality Signaling and Social
+Recognition?'' \emph{Behav Ecol} 27 (1): 2--13.
+\url{https://doi.org/10.1093/beheco/arv109}.
+
+\leavevmode\hypertarget{ref-tibbetts_Individual_2007}{}%
+Tibbetts, Elizabeth A., and James Dale. 2007. ``Individual Recognition:
+It Is Good to Be Different.'' \emph{Trends in Ecology \& Evolution} 22
+(10): 529--37. \url{https://doi.org/10.1016/j.tree.2007.09.001}.
+
 \end{frame}

### 2020-10-22 Dyn_reac_learn.R

@@ -12,7 +12,7 @@ require("jsonlite")
 
 # Scenario to be plotted - corresponds to folders where simulations are stored
 
-scenario<-"test"
+scenario<-"nIntGroupEvol4"
 
 extSimsDir<-#here("Simulations",paste0(scenario,"_"))
   paste0("e:/BadgeSims/",scenario,"_")
@@ -359,7 +359,7 @@ finReps<-evol[time==max(time),seed]
 
 
 for(runChoi in finReps){
- runChoi<-0
+ # runChoi<-0
 
 # Average trajectory
 par(plt=posPlot(numploty = 3,idploty = 2,numplotx = 3,idplotx = 1,

### 2020-10-22 learnDynamics.R

@@ -11,7 +11,7 @@ source(here("AccFunc.R"))
 
 # Scenario to be plotted - corresponds to folders where simulations are stored
 
-scenario<-"nIntGroupEvol2"
+scenario<-"alphaAct2"
 extSimsDir<-paste0("e:/BadgeSims/",scenario,"_")
 
 
@@ -28,7 +28,7 @@ paramName<-grep(".json",paramName,value=TRUE)
 param<-fromJSON(here("Simulations",paste0(scenario,"_"),paramName))
 
 
-fileId<-1
+fileId<-3
 
 
 # indLearn<-fread(here("Simulations",paste0(scenario,"_"),indList[fileId]))
@@ -63,14 +63,14 @@ colorbreaksQual<-seq(0,1,length=100)
 # Actor 
 
 finReps<-indLearn[time==max(time),unique(seed)]
-seedCh<-3
+seedCh<-0
   finReps[round(runif(1,0,length(finReps)))+1]
 
 
 # Select run and generation to plot
 tempPop<-indLearn[time==gener&seed==seedCh]
 
-timePoints<-round(seq(1,length(unique(tempPop[,nInteract]))-1,
+timePoints<-round(seq(1,length(unique(tempPop[,nInteract]))-40,
                       length.out = 10))
 
 # png(here("Simulations",paste0(scenario,"_"),paste0(nampar,Valpar,"learnDyn.png")),
@@ -96,8 +96,10 @@ plot.new()
                          logist(totRBF(rangx,
                                 centers,sigSquar,
                                 as.double(
-                                  tempPop[(nInteract==behavTime&
-                                             Quality<0.5)&indId==x,.SD,
+                                  tempPop[(nInteract==behavTime
+                                           &indId==x)
+                                          # &(Quality>0.4&Quality<0.6)
+                                          ,.SD,
                                            .SDcol=grep("WeightAct",
                                                        names(tempPop),
                                                        value = TRUE)

### 2020-10-27 Presentations/GenesAsEnviron.Rmd
cpp code: a bug fixed in the records of interactions in each generations. New simulations started (par2Json):starting  values for 3d reaction norms had the opposite sign. Moving forward in the presentation for AGA

@@
 
-```{r,echo=FALSE,fig.align="center",out.width='80%'}
-knitr::include_graphics("../Images/harm.jpg")
-```
-
-## What about signals?
-
 ### Badge of status
@@
 
-### When are signals honest?
+### The handicap principle
 
-* Impossible to fake
-* Individuals have common interests
-* Handicap principle (signal´s cost is proportional to quality)
+```{r,echo=FALSE,fig.align='center',out.height='35%',fig.show='hold'}
+knitr::include_graphics("../Images/Amotz_Zahavi_profile.jpg")
+knitr::include_graphics("../Images/sexualSelection.jpg")
+knitr::include_graphics("../Images/Grafen.jpg")
+```
 
-## What about signals?
+```{r,echo=FALSE,fig.align='center',out.height='35%'}
 
-### Individual recognition
-
-```{r,echo=FALSE,fig.align='center',out.height='70%'}
-knitr::include_graphics("../Images//indRecog.png")
+badQualRang<-seq(0,1,length=1000)
+alpha<-0
+beta<-6
+costFunc<-function(beta,alpha,x) {d<-1/(1+exp(-alpha-beta*x))}
+par(plt=posPlot())
+betRang<-c(5)
+matplot(y=sapply(betRang,FUN = costFunc,alpha=alpha,x=badQualRang),x=badQualRang,
+        type="l",cex.lab=1.5, cex.axis=2,lwd=3,lty=1,las=1,
+     col=c(2,3),ylim=c(0.5,1.1),xaxt="s",yaxt="s",ylab="",xlab="") #,col=i
+# lines(badgeFunc(qualRange,3,-60,1.05)~qualRange,cex.lab=1.5,cex.axis=2,lwd=3,
+#       col="blue")
+# legend("bottomright",legend = betRang,title = expression(beta),col=1:3,
+       # lty=1)
+mtext(side=1,text="Quality",line=3.5,cex=3)
+mtext(side=2,text="Survival",line=3.5,cex=3)
 
 ```
-[@tibbetts_Individual_2007]
+[@zahavi_Mate_1975;@grafen_Biological_1990]
 
-## What about signals?
+## The handicap principle
 
-### BS or IR?
-
-```{r,echo=FALSE,fig.align='center',out.height='75%'}
-knitr::include_graphics("../Images//IRvsBS.png")
-
-```
-[@sheehan_There_2016]
-
-## What about signals?
+### Badges of status
 
@@
 
-```{r,echo=FALSE,fig.align='center',out.width='90%'}
+```{r,echo=FALSE,fig.align='center',out.width='70%'}
 knitr::include_graphics("../Images//pavlov.jpg")

### 2020-10-27 Presentations/GenesAsEnviron.tex
cpp code: a bug fixed in the records of interactions in each generations. New simulations started (par2Json):starting  values for 3d reaction norms had the opposite sign. Moving forward in the presentation for AGA

@@
 \date{Genes as environment - AGA2020 virtual symposium\\
-October 21, 2020}
+October 26, 2020}
 
@@
 
-\begin{center}\includegraphics[width=0.8\linewidth]{../Images/harm} \end{center}
-
-\end{frame}
-
-\begin{frame}{What about signals?}
-\protect\hypertarget{what-about-signals}{}
-
 \begin{block}{Badge of status}
@@
 \begin{frame}{What about signals?}
-\protect\hypertarget{what-about-signals-1}{}
+\protect\hypertarget{what-about-signals}{}
 
-\begin{block}{When are signals honest?}
+\begin{block}{The handicap principle}
 
-\begin{itemize}
-\tightlist
-\item
-  Impossible to fake
-\item
-  Individuals have common interests
-\item
-  Handicap principle (signal´s cost is proportional to quality)
-\end{itemize}
+\begin{center}\includegraphics[height=0.35\textheight]{../Images/Amotz_Zahavi_profile} \includegraphics[height=0.35\textheight]{../Images/sexualSelection} \includegraphics[height=0.35\textheight]{../Images/Grafen} \end{center}
+
+\begin{center}\includegraphics[height=0.35\textheight]{GenesAsEnviron_files/figure-beamer/unnamed-chunk-4-1} \end{center}
+
+(Zahavi 1975; Grafen 1990)
 
@@
 
-\begin{frame}{What about signals?}
-\protect\hypertarget{what-about-signals-2}{}
+\begin{frame}{The handicap principle}
+\protect\hypertarget{the-handicap-principle-1}{}
 
-\begin{block}{Individual recognition}
+\begin{block}{Badges of status}
 
-\begin{center}\includegraphics[height=0.7\textheight]{../Images//indRecog} \end{center}
-
-(Tibbetts and Dale 2007)
+\begin{center}\includegraphics[width=0.9\linewidth]{../Images//comSystemBot} \end{center}
 
 \end{block}
-
-\end{frame}
-
-\begin{frame}{What about signals?}
-\protect\hypertarget{what-about-signals-3}{}
-
-\begin{block}{BS or IR?}
-
-\begin{center}\includegraphics[height=0.75\textheight]{../Images//IRvsBS} \end{center}
-
-(Sheehan and Bergman 2016)
-
-\end{block}
-
-\end{frame}
-
-\begin{frame}{What about signals?}
-\protect\hypertarget{what-about-signals-4}{}
-
-\begin{center}\includegraphics[width=0.9\linewidth]{../Images//comSystemBot} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=0.9\linewidth]{../Images//pavlov} \end{center}
+\begin{center}\includegraphics[width=0.7\linewidth]{../Images//pavlov} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=0.5\linewidth]{GenesAsEnviron_files/figure-beamer/unnamed-chunk-11-1} \end{center}
+\begin{center}\includegraphics[width=0.5\linewidth]{GenesAsEnviron_files/figure-beamer/unnamed-chunk-10-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=0.5\linewidth]{GenesAsEnviron_files/figure-beamer/unnamed-chunk-12-1} \end{center}
+\begin{center}\includegraphics[width=0.5\linewidth]{GenesAsEnviron_files/figure-beamer/unnamed-chunk-11-1} \end{center}
 
@@
 
-\caption{alpha - beta -Badge}\label{fig:unnamed-chunk-17}
+\caption{alpha - beta -Badge}\label{fig:unnamed-chunk-16}
 \end{figure}
@@
 
-\caption{alpha - beta -Badge}\label{fig:unnamed-chunk-23}
+\caption{alpha - beta -Badge}\label{fig:unnamed-chunk-22}
 \end{figure}
@@
 
-\leavevmode\hypertarget{ref-sheehan_There_2016}{}%
-Sheehan, Michael J., and Thore J. Bergman. 2016. ``Is There an
-Evolutionary Trade-Off Between Quality Signaling and Social
-Recognition?'' \emph{Behav Ecol} 27 (1): 2--13.
-\url{https://doi.org/10.1093/beheco/arv109}.
+\leavevmode\hypertarget{ref-grafen_Biological_1990}{}%
+Grafen, Alan. 1990. ``Biological Signals as Handicaps.'' \emph{Journal
+of Theoretical Biology} 144 (4): 517--46.
+\url{https://doi.org/10.1016/S0022-5193(05)80088-8}.
 
-\leavevmode\hypertarget{ref-tibbetts_Individual_2007}{}%
-Tibbetts, Elizabeth A., and James Dale. 2007. ``Individual Recognition:
-It Is Good to Be Different.'' \emph{Trends in Ecology \& Evolution} 22
-(10): 529--37. \url{https://doi.org/10.1016/j.tree.2007.09.001}.
+\leavevmode\hypertarget{ref-zahavi_Mate_1975}{}%
+Zahavi, Amotz. 1975. ``Mate Selection-A Selection for a Handicap.''
+\emph{Journal of Theoretical Biology} 53 (1): 205--14.
+\url{https://doi.org/10.1016/0022-5193(75)90111-3}.
 

### 2020-10-27 Dyn_reac_learn.R

@@ -20,35 +20,36 @@ extSimsDir<-#here("Simulations",paste0(scenario,"_"))
 
 # Load files -------------------------------------------------------------------
 # Project folder
-(listTest<-list.files(here("Simulations",paste0(scenario,"_"))))
+# (listTest<-list.files(here("Simulations",paste0(scenario,"_"))))
 # External sims folder
-# (listTest<-list.files(extSimsDir,full.names = TRUE))
+(listTest<-list.files(extSimsDir,full.names = TRUE))
 
 (evolList<-grep("evolLearn",listTest,value=TRUE))
 (indList<-grep("indLearn",listTest,value=TRUE))
 paramName<-list.files(here("Simulations",paste0(scenario,"_")))
+paramName<-list.files(extSimsDir,full.names = TRUE)
 paramName<-grep(".json",paramName,value=TRUE)
-param<-fromJSON(here("Simulations",paste0(scenario,"_"),paramName[1]))
-
+param<-#fromJSON(here("Simulations",paste0(scenario,"_"),paramName))
+fromJSON(paramName)
 
 
 val<-1
 
 # loop to produce pdfs for parameter values
-# numCores <- length(indList)
-# registerDoParallel(numCores)
-# foreach(val = 1:length(indList),.packages = c("data.table","here")) %dopar% {
-# source(here("AccFunc.R"))
+numCores <- length(indList)
+registerDoParallel(numCores)
+foreach(val = 1:length(indList),.packages = c("data.table","here")) %dopar% {
+source(here("AccFunc.R"))
 
   fileId<-val
 
 # Project folder
-evol<-fread(here("Simulations",paste0(scenario,"_"),evolList[fileId]))
-pop<-fread(here("Simulations",paste0(scenario,"_"),indList[fileId]))
+# evol<-fread(here("Simulations",paste0(scenario,"_"),evolList[fileId]))
+# pop<-fread(here("Simulations",paste0(scenario,"_"),indList[fileId]))
 
 # External sims folder
-# evol<-fread(evolList[fileId])
-# pop<-fread(indList[fileId])
+evol<-fread(evolList[fileId])
+pop<-fread(indList[fileId])
 
 Valpar<-gsub("[[:alpha:]]",gsub(".txt","",tail(strsplit(indList[val],"_")[[1]],1)),
              replacement = "")
@@ -69,8 +70,9 @@ names(evol)
[…]

### 2020-10-27 cartoons_reactNorms.R

@@ -46,21 +46,21 @@ persp(x = qualRange,y=rivalBadge,z=perspTmp,phi = 35, theta=40 ,
 
 # png(here("reacNorm.png"),width = 800,height = 800)
 
-badQualRang<-seq(-1,1,length=1000)
-alpha<-4
+badQualRang<-seq(0,1,length=1000)
+alpha<-1
 beta<-6
-costFunc<-function(beta,alpha,x) {d<-1/(1+exp(alpha-beta*x))}
+costFunc<-function(beta,alpha,x) {d<-1/(1+exp(-alpha-beta*x))}
 par(plt=posPlot())
-betRang<-c(2,4,6)
+betRang<-c(20)
 matplot(y=sapply(betRang,FUN = costFunc,alpha=alpha,x=badQualRang),x=badQualRang,
-        type="l",cex.lab=1.5, cex.axis=2,lwd=3,lty=1,
-     col=c(1,2,3),ylim=c(0,1.1),xaxt="s",yaxt="s",ylab="",xlab="") #,col=i
+        type="l",cex.lab=1.5, cex.axis=2,lwd=3,lty=1,las=1,
+     col=c(2,3),ylim=c(0.5,1.1),xaxt="s",yaxt="s",ylab="",xlab="") #,col=i
 # lines(badgeFunc(qualRange,3,-60,1.05)~qualRange,cex.lab=1.5,cex.axis=2,lwd=3,
 #       col="blue")
-legend("bottomright",legend = betRang,title = expression(beta),col=1:3,
-       lty=1)
-mtext(side=1,text="Quality-badge",line=2,cex=5)
-mtext(side=2,text="Survival probability",line=2,cex=5)
+# legend("bottomright",legend = betRang,title = expression(beta),col=1:3,
+       # lty=1)
+mtext(side=1,text="Quality",line=2.8,cex=3)
+mtext(side=2,text="Cost",line=3,cex=3)
 # text(0.01,1.17,"A", cex=2)
 dev.off()
 

### 2020-10-27 cue_compt.cpp

@@ -578,10 +578,10 @@ void interactions(vector<individual>& population, int nint,
 	vector<int> sample;
 	vector<int>::iterator itSamp1 = sample.begin();
 	//cout << "track " << generat << endl;
+	localPrint.countIntTypes[0] = 0, localPrint.countIntTypes[1] = 0;
+	localPrint.countIntTypes[2] = 0;
 	if ((population[0].get_strat() == learner) && (trackPopLearn)) {
 		localPrint.countPhenotypes[0] = 0, localPrint.countPhenotypes[1] = 0;
-		localPrint.countIntTypes[0] = 0, localPrint.countIntTypes[1] = 0;
-		localPrint.countIntTypes[2] = 0;
 		/*localPrint.countIntTypesGen[0];
 		localPrint.countIntTypesGen[1], localPrint.countIntTypesGen[2] = 0;*/
 		for (int countSam = 0; countSam < sampleSize; ++countSam) {

### 2020-10-27 learnDynamics.R

@@ -11,7 +11,8 @@ source(here("AccFunc.R"))
 
 # Scenario to be plotted - corresponds to folders where simulations are stored
 
-scenario<-"alphaAct2"
+scenario<-"nIntGroupEvol4"
+
 extSimsDir<-paste0("e:/BadgeSims/",scenario,"_")
 
 
@@ -20,15 +21,18 @@ extSimsDir<-paste0("e:/BadgeSims/",scenario,"_")
 
 # (listTest<-list.files(here("Simulations",paste0(scenario,"_"))))
 (listTest<-list.files(extSimsDir,full.names = TRUE))
+
 (indList<-grep("ind",listTest,value=TRUE))
 
 # parameter values from project folder
-paramName<-list.files(here("Simulations",paste0(scenario,"_")))
+# paramName<-list.files(here("Simulations",paste0(scenario,"_")))
+paramName<-list.files(extSimsDir,full.names = TRUE)
+
 paramName<-grep(".json",paramName,value=TRUE)
-param<-fromJSON(here("Simulations",paste0(scenario,"_"),paramName))
+param<-#fromJSON(here("Simulations",paste0(scenario,"_"),paramName[2]))
+fromJSON(paramName)
 
-
-fileId<-3
+fileId<-1
 
 
 # indLearn<-fread(here("Simulations",paste0(scenario,"_"),indList[fileId]))
@@ -49,7 +53,7 @@ nampar<-gsub("[^[:alpha:]]",gsub(".txt","",tail(strsplit(indList[fileId],"_")[[1
 
 # Changes in learning parameters for several individuals -----------------------
 
-gener<-tail(indLearn[,unique(time)],2)[2]
+gener<-0#tail(indLearn[,unique(time)],2)[2]
 nCenters<-param$nCenters
 sigSquar<-param$sigSq
 interv<-1/(nCenters-1)
@@ -63,14 +67,14 @@ colorbreaksQual<-seq(0,1,length=100)
 # Actor 
 
 finReps<-indLearn[time==max(time),unique(seed)]
-seedCh<-0
+seedCh<-
   finReps[round(runif(1,0,length(finReps)))+1]
 
 
 # Select run and generation to plot
 tempPop<-indLearn[time==gener&seed==seedCh]
 
-timePoints<-round(seq(1,length(unique(tempPop[,nInteract]))-40,
+timePoints<-round(seq(1,length(unique(tempPop[,nInteract]))-1,
                       length.out = 10))
 
[…]

### 2020-10-27 parTOjson.R

@@ -10,17 +10,17 @@ fileName<-"parameters"
 here()
 
 param1<-#fromJSON(paste0("E:/BadgeSims","/betCostNoLearn_","/parameters.json"))
-     fromJSON(here("Simulations","alphaAct_","parameters1.json"))
+     fromJSON(here("Simulations","betCostNoLearn_","parameters.json"))
 
-param<-list(totGen   =10,   nRep    = 8,
-            printGen = 1,   printLearn = 1,
+param<-list(totGen   = 100,   nRep    = 5,
+            printGen = 10,   printLearn = 10,
             printLearnInt = 50,
             popSize  = 2000,  baselineFit = 2,   
-            MutSd    = 0.4,  nInt        = 1500,  init     = c(0,0,1,0),
-            mutRate  = 0.1,  mutType  = 0,
-            sampleSize = 100,   strQual  = 10,
-            errorQual = 0,    alphaBad	 = 3,    betaBad	 = 6,
-            alphaRes =0,betaRes=0, gammaRes=0,
+            MutSd    = 0.3,  nInt        = 500,  init     = c(0,0,0,1),
+            mutRate  = 0.001,  mutType  = 0,
+            sampleSize = 50,   strQual  = 10,
+            errorQual = 0,    alphaBad	 = 0,    betaBad	 = 0,
+            alphaRes =10,betaRes=20, gammaRes=0,
             alphaCrit  = 0.5,  alphaAct = 0.5,
             sigSq   	 = 0.01, nCenters = 6,
             initCrit = 0,      initAct=0,   gamma = 0,
@@ -28,14 +28,14 @@ param<-list(totGen   =10,   nRep    = 8,
             alphCost	 = 3, mutLearn = FALSE,
             nIntGroup  = 2000, 
             payoff_matrix = c(1.5,1,0,0.5),
-            namParam = "alphaAct",
-            rangParam = c(0.1,0.3,0.5,0.6,0.7),
-            typeAgent = 2, #//0. hawk 1.dove 2. learner 3. evaluator
-            folderL=paste(here("Simulations"),"/",sep="")) # comment for debug
-# folderL=paste(here("Simulations","test_"),"/",sep="")) # comment for realease
+            namParam = "betCost",
+            rangParam = c(2,3,5),
+            typeAgent = 3, #//0. hawk 1.dove 2. learner 3. evaluator
+            # folderL=paste(here("Simulations"),"/",sep="")) # comment for debug
[…]

### 2020-11-03 Presentations/GenesAsEnviron.Rmd
New file with the code for the AGA present. figures. New files using cluster analysis to classify inds withon populations. cpp modified to run also replicates serially, or parallely externally. Improved some of the plots.

@@
     \usepackage{array}
+    \usepackage{xcolor}
 ---
@@
 source("../AccFunc.R")
+source("../aesth.R")
 
@@
 
-### The need partners with actions
+### The need to match partners with actions
 
@@
 
+[@botero_Evolution_2010]
+
+## Are bibs really badges?
+
+```{r,echo=FALSE,fig.align='left',out.width='30%'}
+knitr::include_graphics("../Images/Sanchez-Tojar_2018.png")
+```
+
 ## What about learning?
 
-```{r,echo=FALSE,fig.align='center',out.width='70%'}
+```{r,echo=FALSE,fig.align='center',out.height='40%'}
 knitr::include_graphics("../Images//pavlov.jpg")
@@
 
-# Associative learning
-
-## Reinforcement learning theory
-
-\begin{columns}[T]
- \column{0.5\linewidth}
- \begin{equation*}
-  \Delta V_{t(s)}=\alpha \underbrace{(R_t-V_t)}_\text{prediction error}
-  \end{equation*}
-  \pause
-  \includegraphics[width=1\textwidth]{../Images//learn_curve.png}
- \column{0.5\linewidth}
-  \pause
-  \includegraphics[height=.9\textheight]{../Images//predErrorNeur.png}
-\end{columns}
-
-## Reinforcement learning theory
-
-```{r,echo=FALSE,fig.align='center',out.width='90%'}
-knitr::include_graphics("../Images//learnUpdate.png")
+```{r,echo=FALSE,fig.align='center',out.height='35%'}
+knitr::include_graphics("../Images/learn_cartoon_0.png")
 ```
 
-## The Hawk-Dove game
+## What about learning?
 
-Individuals have one of two genetically determined phenotypic strategies.
-_Hawks_ are willing to start a conflict over resources, while _doves_ prefer to stand down in the hope to share the resource without an aggressive contest.
-
-```{r,echo=FALSE,fig.align="center",out.width='50%'}
-knitr::include_graphics("../Images//HDgame.png")
+```{r,echo=FALSE,fig.align='center',out.height='40%'}
+knitr::include_graphics("../Images//pavlov.jpg")
 ```
 
-## The hawk-dove game
+```{r,echo=FALSE,fig.align='center',out.height='35%'}
+knitr::include_graphics("../Images/learn_cartoon_1.png")
+```
 
-\begin{align*}
-w_H &= p_H \frac{V-C}{2}+(1-p_h) V\\
-w_D &= p_H 0 + (1-p_H)\frac{V}{2}
-\end{align*}
-\vspace{-0.8cm}
-```{r, echo=FALSE,fig.align="center", out.width = '55%'}
-knitr::include_graphics("../Images//Hawk_dove_plot.png")
+## What about learning?
+
+```{r,echo=FALSE,fig.align='center',out.height='40%'}
+knitr::include_graphics("../Images//pavlov.jpg")
 ```
+
+```{r,echo=FALSE,fig.align='center',out.height='35%'}
+knitr::include_graphics("../Images/learn_cartoon_2.png")
+```
+
+## What about learning?
+
+```{r,echo=FALSE,fig.align='center',out.height='40%'}
+knitr::include_graphics("../Images//pavlov.jpg")
+```
+
+```{r,echo=FALSE,fig.align='center',out.height='35%'}
+knitr::include_graphics("../Images/learn_cartoon_3.png")
+```
+
+## What about learning?
+
+```{r,echo=FALSE,fig.align='center',out.height='40%'}
+knitr::include_graphics("../Images//pavlov.jpg")
+```
+
+```{r,echo=FALSE,fig.align='center',out.height='35%'}
+knitr::include_graphics("../Images/learn_cartoon_5.png")
+```
+
+## What about learning?
+
+```{r,echo=FALSE,fig.align='center',out.height='40%'}
+knitr::include_graphics("../Images//pavlov.jpg")
+```
+
+```{r,echo=FALSE,fig.align='center',out.height='35%'}
+knitr::include_graphics("../Images/learn_cartoon_6.png")
+```
+
+<!-- ## How to implement learning -->
+
+<!-- ### Reinforcement learning theory -->
+
+<!-- \begin{columns}[T] -->
+<!--  \column{0.5\linewidth} -->
+<!--  \begin{equation*} -->
+<!--   \Delta V_{t(s)}=\alpha \underbrace{(R_t-V_t)}_\text{prediction error} -->
+<!--   \end{equation*} -->
+<!--   \pause -->
+<!--   \includegraphics[width=1\textwidth]{../Images//learn_curve.png} -->
+<!--  \column{0.5\linewidth} -->
+<!--   \pause -->
+<!--   \includegraphics[height=.9\textheight]{../Images//predErrorNeur.png} -->
+<!-- \end{columns} -->
+
+<!-- ## Reinforcement learning theory -->
+
+<!-- ```{r,echo=FALSE,fig.align='center',out.width='90%'} -->
+<!-- knitr::include_graphics("../Images//learnUpdate.png") -->
+<!-- ``` -->
+
+<!-- ## The Hawk-Dove game -->
+
+<!-- Individuals have one of two genetically determined phenotypic strategies.
+_Hawks_ are willing to start a conflict over resources, while _doves_ prefer to stand down in the hope to share the resource without an aggressive contest.
+-->
+
+<!-- ```{r,echo=FALSE,fig.align="center",out.width='50%'} -->
+<!-- knitr::include_graphics("../Images//HDgame.png") -->
+<!-- ``` -->
 
@@
 
-## Environmental states
+## A model of communication in animal conflict
 
-### Discrete states
-\includegraphics[width=0.8\textwidth]{../Images//discreteStates.png}
+\vspace{-0.5cm}
 
-\pause
+### Comunication
 
-### Continuos states
-\includegraphics[width=0.8\textwidth]{../Images//contStates.png}
+\begin{columns}[T]
 
-## Continuos environmental states
+\column{0.45\linewidth}
+
+The sender code
+
+\tiny
 \begin{equation*}
-  \Delta V_{t(s)}=\alpha \underbrace{(R_t-V_t)}_\text{prediction error}
+B_i = \frac{1}{1+e^{-(\textcolor{red}{\alpha_s} +\textcolor{red}{\beta_s} Q_i)}}
 \end{equation*}
 
-```{r fig2, echo=FALSE, fig.cap="\\label{fig:learning_cartoonRBF}Function approximation for the actor and the critic.
-Using random values for the responses of each center.
-Circles show the location of the centers and the maximum response they trigger, while the lines show the total response along the badge size continium.", out.width = '60%'}
-knitr::include_graphics("../Images/cartoonRBF.png")
+```{r,echo=FALSE,fig.align='center',out.width='50%'}
+ par(plt=posPlot(),las=1)
+ rangeBeta<-c(4,10,-4,-10)
+ rangQual<-seq(0,1,by = 0.01)
+ linesSender<-sapply(rangeBeta,function(x)logist(rangQual,alpha = x/2,beta = x))
+ plot(logist(rangQual,alpha = 0,beta = 0)~rangQual,type="l",xlab="Quality",
+      ylab = expression(p[w]),cex.lab=2,lwd=2,col="grey",cex.axis=2,ylim=c(0,1))
+ matlines(x=rangQual,y=linesSender,lwd=4,col=colRuns,lty=1)
+ legend("left",legend=rangeBeta,col=colRuns,lwd=3,title = expression(beta),cex=2,bty="n")
+
 ```
 
-# Our model
+\column{0.45\linewidth}
 
-## The hawk-dove game 2.0
+The receiver code (learning)
 
+\tiny
+\begin{equation*}
+  \Delta V_{t(s)}=\alpha (R_t-V_t)
+\end{equation*}
+
+```{r, echo=FALSE,fig.align='center', out.width = '60%'}
+
+nCenters<-6
+interv<-1/(nCenters-1)
+centers<-interv*seq(0,nCenters-1)
+sigSq<-0.01
+weights<-rep(0,nCenters)
+#  Random weights
+weights<-runif(nCenters,min=-1,max=1)#RBF(centers,0.5,0.05)
+rangx<-seq(0,1,length=1000)
+
+par(plt=posPlot(),las=1)
+plot(logist(totRBF(rangx,centers,sigSq,weights),alpha = 0,beta=1)~rangx,type='l',col=1,
+     xlab="",ylab="p(retreat)",ylim=c(0,1),lwd=3,xaxt="n",cex.lab=2,cex.axis=2)
+points(y=logist(weights,alpha = 0,beta=1),x=centers,cex=3,pch=20)
+
+```
+
+\end{columns}
+
+### Conflict
+
+\vspace{0.3cm}
+\begin{columns}
+\column{0.5\linewidth}
+hawk-dove game
+\tiny
 \begin{center}
-\begin{tabular}{ >{\centering\arraybackslash}p{1cm} | >{\centering\arraybackslash}p{4.5cm} | >{\centering\arraybackslash}p{4.5cm} }
+\begin{tabular}{ >{\centering\arraybackslash}p{0.2cm} | >{\centering\arraybackslash}p{2.5cm} | >{\centering\arraybackslash}p{0.8cm} }
 & H & D \\ \hline
@@
 
-```{r,echo=FALSE,fig.align='center',out.width='50%'}
+\column{0.5\linewidth}
+
+```{r,echo=FALSE,fig.align='center',out.width='60%'}
 par(plt=posPlot(),las=1)
@@
 legend("topleft",legend=c(2,10),col=c("red","green"),lwd=3,title = expression(beta),cex=2,bty="n")
-
 ```
 
-## Sender Code
-
-\begin{equation*}
-s_i = \frac{1}{1+e^{-(\epsilon_i+\gamma_iQ_i)}}
-\label{eq:react_norm}
-\end{equation*}
-
-```{r,echo=FALSE,fig.align='center',out.width='50%'}
-par(plt=posPlot(),las=1)
-plot(logist(seq(0,1,length.out = 1000),alpha = 2,beta = 4)~
-       seq(0,1,length.out = 1000),type="l",xlab="Quality",
-     ylab = expression(p[w]),cex.lab=2,lwd=3,col="red",cex.axis=2,ylim=c(0,1))
-lines(logist(seq(0,1,length.out = 1000),alpha = 5,beta = 10)~
-       seq(0,1,length.out = 1000),lwd=3,col="green")
-legend("topleft",legend=c(4,10),col=c("red","green"),lwd=3,title = expression(gamma),cex=2,bty="n")
-
-```
-
-## Receiver code
-
-\begin{equation*}
-  \Delta V_{t(s)}=\alpha (R_t-V_t)
-\end{equation*}
-
-```{r, echo=FALSE,fig.align='center', out.width = '60%'}
-knitr::include_graphics("../Images/cartoonRBF.png")
-```
+\end{columns}
 
@@
 
-Learning simulations (without evolution yet), the sender code is set as an honest signal.
-Receivers develop a threshold like response according to their quality
+<!-- Learning simulations (without evolution yet), the sender code is set as an honest signal. -->
+<!-- Receivers develop a threshold like response according to their quality -->
 
+\pause
 ```{r, echo=FALSE,fig.align='center', out.width = '100%'}
@@
 
-## Learning can mediate the evolution of Badges of Status
-### Variation among replicates
-\small
-set of simulations with the badge working as a handicap (cost inversely proportional to quality)
+<!-- ## Learning can mediate the evolution of Badges of Status -->
+<!-- ### Variation among replicates -->
+<!-- \small -->
+<!-- set of simulations with the badge working as a handicap (cost inversely proportional to quality) -->
 
-```{r, echo=FALSE,fig.align='center', out.width='100%'}
-knitr::include_graphics("../Simulations/betCostEvol1_/evolDynAll_betCost5.png")
-```
+<!-- ```{r, echo=FALSE,fig.align='center', out.width='100%'} -->
+<!-- knitr::include_graphics("../Simulations/betCostEvol1_/evolDynAll_betCost5.png") -->
+<!-- ``` -->
 
-## Learning can mediate the evolution of Badges of Status
-### Individual replicates: Badges
-\small
-One of the replicates of the previous slide, the badge as a handicap
+## Learning mediates the evolution of badges of status (handicaps)
+
+<!-- \small -->
+<!-- One of the replicates of the previous slide, the badge as a handicap -->
 ```{r, echo=FALSE,fig.align='center', out.width='100%'}
@@
 
-## Learning can mediate the evolution of Badges of Status
-### Individual replicates: Badges
-\small
-Panels show the changes in frequency distribution along evolution ($\alpha$, $\beta$, and resulting badge respectively) for the replicate shown in the previous slide
+<!-- ## Learning can mediate the evolution of Badges of Status -->
+<!-- ### Individual replicates: Badges -->
+<!-- \small -->
+<!-- Panels show the changes in frequency distribution along evolution ($\alpha$, $\beta$, and resulting badge respectively) for the replicate shown in the previous slide -->
 
-```{r, echo=FALSE,fig.align='left', out.height='28%',fig.show='hold', fig.cap="alpha - beta -Badge"}
-knitr::include_graphics("../Simulations/betCostEvol1_/evolDistAlpha0_betCost5.png")
-knitr::include_graphics("../Simulations/betCostEvol1_/evolDistBeta0_betCost5.png")
-knitr::include_graphics("../Simulations/betCostEvol1_/evolDistBadge0_betCost5.png")
-```
+<!-- ```{r, echo=FALSE,fig.align='left', out.height='28%',fig.show='hold', fig.cap="alpha - beta -Badge"} -->
+<!-- knitr::include_graphics("../Simulations/betCostEvol1_/evolDistAlpha0_betCost5.png") -->
+<!-- knitr::include_graphics("../Simulations/betCostEvol1_/evolDistBeta0_betCost5.png") -->
+<!-- knitr::include_graphics("../Simulations/betCostEvol1_/evolDistBadge0_betCost5.png") -->
+<!-- ``` -->
 
@@
 ### Individual replicates: no badges
-\small
-One of the replicates where the badge did not evolve
+<!-- \small -->
+<!-- One of the replicates where the badge did not evolve -->
 
@@
 
-## The effect of costs on Badges of Status
+## Badges of Status: do they matter?
 ### Behavioural interactions
-\small
-Colours correspond to replicates, each point is one snapshot on the second half of the simulations.
-Higher costs seem to favour the signalling equilibrium
 
-```{r, echo=FALSE,fig.align='center', out.width='80%'}
-knitr::include_graphics("../Simulations/betCostEvol1_/BehavIntALL.png")
+<!-- \small -->
+<!-- Colours correspond to replicates, each point is one snapshot on the second half of the simulations. -->
+<!-- Higher costs seem to favour the signalling equilibrium -->
+
+```{r, echo=FALSE,fig.align='center', out.width='90%'}
+knitr::include_graphics("../Simulations/betCostEvol1_/compClustRep_betCost5.png")
 ```
-## The effect of costs on Badges of Status
-### Badge size
-\small
-Same as previous slide but for the resulting signal. Higher costs yield smaller averga sizes and smaller ranges.
 
-```{r, echo=FALSE,fig.align='center', out.width='80%'}
-knitr::include_graphics("../Simulations/betCostEvol1_/CueALL.png")
+<!-- ## The effect of costs on Badges of Status -->
+<!-- ### Badge size -->
+<!-- \small -->
+<!-- Same as previous slide but for the resulting signal. Higher costs yield smaller averga sizes and smaller ranges.  -->
+
+<!-- ```{r, echo=FALSE,fig.align='center', out.width='80%'} -->
+<!-- knitr::include_graphics("../Simulations/betCostEvol1_/CueALL.png") -->
+<!-- ``` -->
+
+<!-- ## The evolution of cheap Badges of Status -->
+
+<!-- <!-- ### Variation among replicates --> -->
+<!-- <!-- \small --> -->
+<!-- <!-- Here simulations where the signal has no cost --> -->
+
+<!-- ```{r, echo=FALSE,fig.align='center', out.width='80%'} -->
+<!-- knitr::include_graphics("../Simulations/nIntGroupEvol1_/evolDynALL_nIntGroup2000.png") -->
+<!-- ``` -->
+
+## The evolution of cheap Badges of Status
+
+<!-- \small -->
+<!-- Here simulations where the signal has no cost. -->
+
+<!-- There seems to be some consistency in the three type of sender codes -->
+
+```{r, echo=FALSE,fig.align='center', out.width='90%'}
+knitr::include_graphics("../Simulations/nIntGroupEvol4_/evolDyn13_nIntGroup2000.png")
 ```
@@
 ## The evolution of cheap Badges of Status
-### Variation among replicates
-\small
-Here simulations where the signal has no cost
 
-```{r, echo=FALSE,fig.align='center', out.width='80%'}
-knitr::include_graphics("../Simulations/nIntGroupEvol1_/evolDynALL_nIntGroup2000.png")
-```
+<!-- \tiny -->
+<!-- Panels show the changes in frequency distribution along evolution ($\alpha$, $\beta$, and resulting badge respectively) for the replicate shown in the previous slide -->
 
-## The evolution of cheap Badges of Status
-### Variation within a replicate
-\small
-Here simulations where the signal has no cost.
-
-There seems to be some consistency in the three type of sender codes
-
-```{r, echo=FALSE,fig.align='center', out.width='80%'}
-knitr::include_graphics("../Simulations/nIntGroupEvol1_/evolDyn8_nIntGroup2000.png")
-```
-
-## The evolution of cheap Badges of Status
-### Variation within a replicate
-
-\tiny
-Panels show the changes in frequency distribution along evolution ($\alpha$, $\beta$, and resulting badge respectively) for the replicate shown in the previous slide
-
-Branching events in the reaction norm parameters yield three types of sender codes. One (more or less)
-honest, two dishonest (one with a big badge and one with a small badge)
+<!-- Branching events in the reaction norm parameters yield three types of sender codes. One (more or less) -->
+<!-- honest, two dishonest (one with a big badge and one with a small badge) -->
 
 ```{r, echo=FALSE,fig.align='left', out.height='28%',fig.show='hold', fig.cap="alpha - beta -Badge"}
-knitr::include_graphics("../Simulations/nIntGroupEvol1_/evolDistAlpha8_nIntGroup2000.png")
-knitr::include_graphics("../Simulations/nIntGroupEvol1_/evolDistBeta8_nIntGroup2000.png")
-knitr::include_graphics("../Simulations/nIntGroupEvol1_/evolDistBadge8_nIntGroup2000.png")
+knitr::include_graphics("../Simulations/nIntGroupEvol4_/evolDistAlpha13_nIntGroup2000.png")
+knitr::include_graphics("../Simulations/nIntGroupEvol4_/evolDistBeta13_nIntGroup2000.png")
+knitr::include_graphics("../Simulations/nIntGroupEvol4_/evolDistBadge13_nIntGroup2000.png")
 ```
@@
 ### How often does it happen?
-\tiny
-Values of $\alpha$ and $\beta$ for each individual in the different replicates. Each dot an individual,
-each panel a replicate. Individuals are sampled from populations from the second half of the simulations.
+<!-- \tiny -->
+<!-- Values of $\alpha$ and $\beta$ for each individual in the different replicates. Each dot an individual, -->
+<!-- each panel a replicate. Individuals are sampled from populations from the second half of the simulations. -->
 
-Individuals in some simulations split up in three groups, others in two groups; 
-in one simulation the population does not split up
+<!-- Individuals in some simulations split up in three groups, others in two groups;  -->
+<!-- in one simulation the population does not split up -->
 
@@
 
-## Interaction structures - do they matter?
+<!-- ## Interaction structures - do they matter? -->
 
-```{r, echo=FALSE,fig.align='center', out.width='100%',}
-knitr::include_graphics("../Images/networks.png")
-```
-[@chaine_Manipulating_2018]
+<!-- ```{r, echo=FALSE,fig.align='center', out.width='100%',} -->
+<!-- knitr::include_graphics("../Images/networks.png") -->
+<!-- ``` -->
+<!-- [@chaine_Manipulating_2018] -->
 
-## Localized interaction structures
+<!-- ## Localized interaction structures -->
 
-### Costly signals 
-\tiny
-Behavioural interactions and badge size for simulations with costs where individuals interact in small groups. On the $x$ axes number of different individuals each focal interacts with.
+<!-- ### Costly signals  -->
+<!-- \tiny -->
+<!-- Behavioural interactions and badge size for simulations with costs where individuals interact in small groups. On the $x$ axes number of different individuals each focal interacts with. -->
 
-```{r, echo=FALSE,fig.align='center', out.width='90%',fig.show='hold'}
-knitr::include_graphics("../Simulations/nIntGroupCost_/BehavIntALL.png")
-knitr::include_graphics("../Simulations/nIntGroupCost_/CueALL.png")
-```
+<!-- ```{r, echo=FALSE,fig.align='center', out.width='90%',fig.show='hold'} -->
+<!-- knitr::include_graphics("../Simulations/nIntGroupCost_/BehavIntALL.png") -->
+<!-- knitr::include_graphics("../Simulations/nIntGroupCost_/CueALL.png") -->
+<!-- ``` -->
 
-## Localized interaction structures
+<!-- ## Localized interaction structures -->
 
-### Cheap signals (free)
-\small
-Variation among individuals in different replicates (just like before) in replicates where individuals interact with 10 other individuals.
+<!-- ### Cheap signals (free) -->
+<!-- \small -->
+<!-- Variation among individuals in different replicates (just like before) in replicates where individuals interact with 10 other individuals. -->
 
-A part from replicates with two clusters, the clustering seems less straightforward as before
+<!-- A part from replicates with two clusters, the clustering seems less straightforward as before -->
 
-```{r, echo=FALSE,fig.align='center', out.width='85%',fig.show='hold'}
-knitr::include_graphics("../Simulations/nIntGroupEvol1_/corrAlphBet_nIntGroup10.png")
-```
+<!-- ```{r, echo=FALSE,fig.align='center', out.width='85%',fig.show='hold'} -->
+<!-- knitr::include_graphics("../Simulations/nIntGroupEvol1_/corrAlphBet_nIntGroup10.png") -->
+<!-- ``` -->
 
-## Localized interaction structures
+<!-- ## Localized interaction structures -->
 
-### Cheap signals (free)
+<!-- ### Cheap signals (free)  -->
 
-\small
-Behavioural interactions for the replicates form previous slide. I split up the replicates according to clusters (despite clusters not being clear cut).
+<!-- \small -->
+<!-- Behavioural interactions for the replicates form previous slide. I split up the replicates according to clusters (despite clusters not being clear cut). -->
 
-Only replicates with only one cluster seems to have less aggressive interactions
+<!-- Only replicates with only one cluster seems to have less aggressive interactions -->
 
-```{r, echo=FALSE,fig.align='center', out.width='90%',fig.show='hold'}
-knitr::include_graphics("../Simulations/nIntGroupEvol1_/BehavIntAllClusters_nIntGroup10.png")
-```
+<!-- ```{r, echo=FALSE,fig.align='center', out.width='90%',fig.show='hold'} -->
+<!-- knitr::include_graphics("../Simulations/nIntGroupEvol1_/BehavIntAllClusters_nIntGroup10.png") -->
+<!-- ``` -->
 
-## Localized interaction structures
+<!-- ## Localized interaction structures -->
 
-### Cheap signals or Individual recognition?
-\tiny
-In these set, I changed the learning set up, I increased the number of centers (10 now, 6 before)
-along the signal axis; and I reduced the level or generalization.
+<!-- ### Cheap signals or Individual recognition? -->
+<!-- \tiny -->
+<!-- In these set, I changed the learning set up, I increased the number of centers (10 now, 6 before) -->
+<!-- along the signal axis; and I reduced the level or generalization.  -->
 
-There seems to be an effect on the slope of the reaction norm. It stays very close to zero.
+<!-- There seems to be an effect on the slope of the reaction norm. It stays very close to zero.  -->
 
-```{r, echo=FALSE,fig.align='center', out.width='90%',fig.show='hold'}
-knitr::include_graphics("../Simulations/nIntGroupCenters1_/evolDynALL_nIntGroup10.png")
-```
+<!-- ```{r, echo=FALSE,fig.align='center', out.width='90%',fig.show='hold'} -->
+<!-- knitr::include_graphics("../Simulations/nIntGroupCenters1_/evolDynALL_nIntGroup10.png") -->
+<!-- ``` -->
 
-## Localized interaction structures
+<!-- ## Localized interaction structures -->
 
-### Cheap signals or Individual recognition
-\small
-Same replicates as the previous slide. Variation in $\beta$ and $\alpha$ among individuals for different replicates.
+<!-- ### Cheap signals or Individual recognition -->
+<!-- \small -->
+<!-- Same replicates as the previous slide. Variation in $\beta$ and $\alpha$ among individuals for different replicates.  -->
 
-No clear clustering. Variation mostly along the $\alpha$ axis
+<!-- No clear clustering. Variation mostly along the $\alpha$ axis -->
 
-```{r, echo=FALSE,fig.align='center', out.width='90%',fig.show='hold'}
-knitr::include_graphics("../Simulations/nIntGroupCenters1_/corrAlphBet_nIntGroup10.png")
-```
-## Localized interaction structures
+<!-- ```{r, echo=FALSE,fig.align='center', out.width='90%',fig.show='hold'} -->
+<!-- knitr::include_graphics("../Simulations/nIntGroupCenters1_/corrAlphBet_nIntGroup10.png") -->
+<!-- ``` -->
+<!-- ## Localized interaction structures -->
 
-### Cheap signals or Individual recognition?
-\small
-Panels show the changes in frequency distribution along evolution ($\alpha$, $\beta$, and resulting badge respectively) for the replicate shown in the previous slide.
+<!-- ### Cheap signals or Individual recognition? -->
+<!-- \small -->
+<!-- Panels show the changes in frequency distribution along evolution ($\alpha$, $\beta$, and resulting badge respectively) for the replicate shown in the previous slide.  -->
 
-$\beta$ around zero, $\alpha$ all over the place.
+<!-- $\beta$ around zero, $\alpha$ all over the place.  -->
 
-```{r, echo=FALSE,fig.align='left', out.height='28%',fig.show='hold'}
-knitr::include_graphics("../Simulations/nIntGroupCenters1_/evolDistAlpha5_nIntGroup10.png")
-knitr::include_graphics("../Simulations/nIntGroupCenters1_/evolDistBeta5_nIntGroup10.png")
-knitr::include_graphics("../Simulations/nIntGroupCenters1_/evolDistBadge5_nIntGroup10.png")
-```
+<!-- ```{r, echo=FALSE,fig.align='left', out.height='28%',fig.show='hold'} -->
+<!-- knitr::include_graphics("../Simulations/nIntGroupCenters1_/evolDistAlpha5_nIntGroup10.png") -->
+<!-- knitr::include_graphics("../Simulations/nIntGroupCenters1_/evolDistBeta5_nIntGroup10.png") -->
+<!-- knitr::include_graphics("../Simulations/nIntGroupCenters1_/evolDistBadge5_nIntGroup10.png") -->
+<!-- ``` -->
 
-<!-- ## Aqui vamos! mirar correlaciones de 20 y 2000 luego interacciones -->
+<!-- <!-- ## Aqui vamos! mirar correlaciones de 20 y 2000 luego interacciones --> -->
 
-<!-- ## Comparar scenarios -->
+<!-- <!-- ## Comparar scenarios --> -->
 

### 2020-11-03 Presentations/GenesAsEnviron.tex
New file with the code for the AGA present. figures. New files using cluster analysis to classify inds withon populations. cpp modified to run also replicates serially, or parallely externally. Improved some of the plots.

@@
 \setcounter{secnumdepth}{-\maxdimen} % remove section numbering
-\usepackage{float} \usepackage{multicol} \usepackage{amsmath} \usepackage{graphicx} \usepackage{array}
+\usepackage{float} \usepackage{multicol} \usepackage{amsmath} \usepackage{graphicx} \usepackage{array} \usepackage{xcolor}
 
@@
 \date{Genes as environment - AGA2020 virtual symposium\\
-October 26, 2020}
+November 02, 2020}
 
@@
 
-\begin{block}{The need partners with actions}
+\begin{block}{The need to match partners with actions}
 
@@
 
+(Botero et al. 2010)
+
 \end{block}
+
+\end{frame}
+
+\begin{frame}{Are bibs really badges?}
+\protect\hypertarget{are-bibs-really-badges}{}
+
+\begin{flushleft}\includegraphics[width=0.3\linewidth]{../Images/Sanchez-Tojar_2018} \end{flushleft}
 
@@
 
-\begin{center}\includegraphics[width=0.7\linewidth]{../Images//pavlov} \end{center}
+\begin{center}\includegraphics[height=0.4\textheight]{../Images//pavlov} \end{center}
+
+\begin{center}\includegraphics[height=0.35\textheight]{../Images/learn_cartoon_0} \end{center}
 
@@
 
-\hypertarget{associative-learning}{%
-\section{Associative learning}\label{associative-learning}}
+\begin{frame}{What about learning?}
+\protect\hypertarget{what-about-learning-1}{}
 
-\begin{frame}{Reinforcement learning theory}
-\protect\hypertarget{reinforcement-learning-theory}{}
+\begin{center}\includegraphics[height=0.4\textheight]{../Images//pavlov} \end{center}
 
-\begin{columns}[T]
- \column{0.5\linewidth}
- \begin{equation*}
-  \Delta V_{t(s)}=\alpha \underbrace{(R_t-V_t)}_\text{prediction error}
-  \end{equation*}
-  \pause
-  \includegraphics[width=1\textwidth]{../Images//learn_curve.png}
- \column{0.5\linewidth}
-  \pause
-  \includegraphics[height=.9\textheight]{../Images//predErrorNeur.png}
-\end{columns}
+\begin{center}\includegraphics[height=0.35\textheight]{../Images/learn_cartoon_1} \end{center}
 
@@
 
-\begin{frame}{Reinforcement learning theory}
-\protect\hypertarget{reinforcement-learning-theory-1}{}
+\begin{frame}{What about learning?}
+\protect\hypertarget{what-about-learning-2}{}
 
-\begin{center}\includegraphics[width=0.9\linewidth]{../Images//learnUpdate} \end{center}
+\begin{center}\includegraphics[height=0.4\textheight]{../Images//pavlov} \end{center}
+
+\begin{center}\includegraphics[height=0.35\textheight]{../Images/learn_cartoon_2} \end{center}
 
@@
 
-\begin{frame}{The Hawk-Dove game}
-\protect\hypertarget{the-hawk-dove-game}{}
+\begin{frame}{What about learning?}
+\protect\hypertarget{what-about-learning-3}{}
 
-Individuals have one of two genetically determined phenotypic
-strategies. \emph{Hawks} are willing to start a conflict over resources,
-while \emph{doves} prefer to stand down in the hope to share the
-resource without an aggressive contest.
+\begin{center}\includegraphics[height=0.4\textheight]{../Images//pavlov} \end{center}
 
-\begin{center}\includegraphics[width=0.5\linewidth]{../Images//HDgame} \end{center}
+\begin{center}\includegraphics[height=0.35\textheight]{../Images/learn_cartoon_3} \end{center}
 
@@
 
-\begin{frame}{The hawk-dove game}
-\protect\hypertarget{the-hawk-dove-game-1}{}
+\begin{frame}{What about learning?}
+\protect\hypertarget{what-about-learning-4}{}
 
-\begin{align*}
-w_H &= p_H \frac{V-C}{2}+(1-p_h) V\\
-w_D &= p_H 0 + (1-p_H)\frac{V}{2}
-\end{align*} \vspace{-0.8cm}
+\begin{center}\includegraphics[height=0.4\textheight]{../Images//pavlov} \end{center}
 
-\begin{center}\includegraphics[width=0.55\linewidth]{../Images//Hawk_dove_plot} \end{center}
+\begin{center}\includegraphics[height=0.35\textheight]{../Images/learn_cartoon_5} \end{center}
 
@@
 
-\begin{frame}{Environmental states}
-\protect\hypertarget{environmental-states}{}
+\begin{frame}{What about learning?}
+\protect\hypertarget{what-about-learning-5}{}
 
-\begin{block}{Discrete states}
+\begin{center}\includegraphics[height=0.4\textheight]{../Images//pavlov} \end{center}
 
-\includegraphics[width=0.8\textwidth]{../Images//discreteStates.png}
+\begin{center}\includegraphics[height=0.35\textheight]{../Images/learn_cartoon_6} \end{center}
 
-\pause
+\end{frame}
+
+\begin{frame}{A model of communication in animal conflict}
+\protect\hypertarget{a-model-of-communication-in-animal-conflict}{}
+
+\vspace{-0.5cm}
+
+\begin{block}{Comunication}
+
+\begin{columns}[T]
+
+\column{0.45\linewidth}
+
+The sender code
+
+\tiny
+\begin{equation*}
+B_i = \frac{1}{1+e^{-(\textcolor{red}{\alpha_s} +\textcolor{red}{\beta_s} Q_i)}}
+\end{equation*}
+
+\begin{center}\includegraphics[width=0.5\linewidth]{GenesAsEnviron_files/figure-beamer/unnamed-chunk-19-1} \end{center}
+
+\column{0.45\linewidth}
+
+The receiver code (learning)
+
+\tiny
+\begin{equation*}
+  \Delta V_{t(s)}=\alpha (R_t-V_t)
+\end{equation*}
+
+\begin{center}\includegraphics[width=0.6\linewidth]{GenesAsEnviron_files/figure-beamer/unnamed-chunk-20-1} \end{center}
+
+\end{columns}
 
@@
 
-\begin{block}{Continuos states}
+\begin{block}{Conflict}
 
-\includegraphics[width=0.8\textwidth]{../Images//contStates.png}
-
-\end{block}
-
-\end{frame}
-
-\begin{frame}{Continuos environmental states}
-\protect\hypertarget{continuos-environmental-states}{}
-
-\begin{equation*}
-  \Delta V_{t(s)}=\alpha \underbrace{(R_t-V_t)}_\text{prediction error}
-\end{equation*}
-
-\begin{figure}
-\includegraphics[width=0.6\linewidth]{../Images/cartoonRBF} \caption{\label{fig:learning_cartoonRBF}Function approximation for the actor and the critic.
-Using random values for the responses of each center.
-Circles show the location of the centers and the maximum response they trigger, while the lines show the total response along the badge size continium.}\label{fig:fig2}
-\end{figure}
-
-\end{frame}
-
-\hypertarget{our-model}{%
-\section{Our model}\label{our-model}}
-
-\begin{frame}{The hawk-dove game 2.0}
-\protect\hypertarget{the-hawk-dove-game-2.0}{}
-
+\vspace{0.3cm}
+\begin{columns}
+\column{0.5\linewidth}
+hawk-dove game
+\tiny
 \begin{center}
-\begin{tabular}{ >{\centering\arraybackslash}p{1cm} | >{\centering\arraybackslash}p{4.5cm} | >{\centering\arraybackslash}p{4.5cm} }
+\begin{tabular}{ >{\centering\arraybackslash}p{0.2cm} | >{\centering\arraybackslash}p{2.5cm} | >{\centering\arraybackslash}p{0.8cm} }
 & H & D \\ \hline
@@
 
-\begin{center}\includegraphics[width=0.5\linewidth]{GenesAsEnviron_files/figure-beamer/unnamed-chunk-10-1} \end{center}
+\column{0.5\linewidth}
 
-\end{frame}
+\begin{center}\includegraphics[width=0.6\linewidth]{GenesAsEnviron_files/figure-beamer/unnamed-chunk-21-1} \end{center}
 
-\begin{frame}{Sender Code}
-\protect\hypertarget{sender-code}{}
+\end{columns}
 
-\begin{equation*}
-s_i = \frac{1}{1+e^{-(\epsilon_i+\gamma_iQ_i)}}
-\label{eq:react_norm}
-\end{equation*}
-
-\begin{center}\includegraphics[width=0.5\linewidth]{GenesAsEnviron_files/figure-beamer/unnamed-chunk-11-1} \end{center}
-
-\end{frame}
-
-\begin{frame}{Receiver code}
-\protect\hypertarget{receiver-code}{}
-
-\begin{equation*}
-  \Delta V_{t(s)}=\alpha (R_t-V_t)
-\end{equation*}
-
-\begin{center}\includegraphics[width=0.6\linewidth]{../Images/cartoonRBF} \end{center}
+\end{block}
 
@@
 
-Learning simulations (without evolution yet), the sender code is set as
-an honest signal. Receivers develop a threshold like response according
-to their quality
+\pause
 
 \begin{center}\includegraphics[width=1\linewidth]{../Simulations/alphaAct_/alphaAct0.6learnDyn} \end{center}
+
+\end{frame}
+
+\begin{frame}{Learning mediates the evolution of badges of status
+(handicaps)}
+\protect\hypertarget{learning-mediates-the-evolution-of-badges-of-status-handicaps}{}
+
+\begin{center}\includegraphics[width=1\linewidth]{../Simulations/betCostEvol1_/evolDyn0_betCost5} \end{center}
 
@@
 
-\begin{block}{Variation among replicates}
-
-\small
-
-set of simulations with the badge working as a handicap (cost inversely
-proportional to quality)
-
-\begin{center}\includegraphics[width=1\linewidth]{../Simulations/betCostEvol1_/evolDynAll_betCost5} \end{center}
-
-\end{block}
-
-\end{frame}
-
-\begin{frame}{Learning can mediate the evolution of Badges of Status}
-\protect\hypertarget{learning-can-mediate-the-evolution-of-badges-of-status-1}{}
-
-\begin{block}{Individual replicates: Badges}
-
-\small
-
-One of the replicates of the previous slide, the badge as a handicap
-
-\begin{center}\includegraphics[width=1\linewidth]{../Simulations/betCostEvol1_/evolDyn0_betCost5} \end{center}
-
-\end{block}
-
-\end{frame}
-
-\begin{frame}{Learning can mediate the evolution of Badges of Status}
-\protect\hypertarget{learning-can-mediate-the-evolution-of-badges-of-status-2}{}
-
-\begin{block}{Individual replicates: Badges}
-
-\small
-
-Panels show the changes in frequency distribution along evolution
-(\(\alpha\), \(\beta\), and resulting badge respectively) for the
-replicate shown in the previous slide
-
-\begin{figure}
-
-\includegraphics[height=0.28\textheight]{../Simulations/betCostEvol1_/evolDistAlpha0_betCost5} \includegraphics[height=0.28\textheight]{../Simulations/betCostEvol1_/evolDistBeta0_betCost5} \includegraphics[height=0.28\textheight]{../Simulations/betCostEvol1_/evolDistBadge0_betCost5} \hfill{}
-
-\caption{alpha - beta -Badge}\label{fig:unnamed-chunk-16}
-\end{figure}
-
-\end{block}
-
-\end{frame}
-
-\begin{frame}{Learning can mediate the evolution of Badges of Status}
-\protect\hypertarget{learning-can-mediate-the-evolution-of-badges-of-status-3}{}
-
 \begin{block}{Individual replicates: no badges}
-
-\small
-
-One of the replicates where the badge did not evolve
 
@@
 
-\begin{frame}{The effect of costs on Badges of Status}
-\protect\hypertarget{the-effect-of-costs-on-badges-of-status}{}
+\begin{frame}{Badges of Status: do they matter?}
+\protect\hypertarget{badges-of-status-do-they-matter}{}
 
@@
 
-\small
+\begin{center}\includegraphics[width=0.9\linewidth]{../Simulations/betCostEvol1_/compClustRep_betCost5} \end{center}
 
-Colours correspond to replicates, each point is one snapshot on the
-second half of the simulations. Higher costs seem to favour the
-signalling equilibrium
-
-\begin{center}\includegraphics[width=0.8\linewidth]{../Simulations/betCostEvol1_/BehavIntALL} \end{center}
-
-\end{block}
-
-\end{frame}
-
-\begin{frame}{The effect of costs on Badges of Status}
-\protect\hypertarget{the-effect-of-costs-on-badges-of-status-1}{}
-
-\begin{block}{Badge size}
-
-\small
-
-Same as previous slide but for the resulting signal. Higher costs yield
-smaller averga sizes and smaller ranges.
-
-\begin{center}\includegraphics[width=0.8\linewidth]{../Simulations/betCostEvol1_/CueALL} \end{center}
+--\textgreater{} --\textgreater{} --\textgreater{}
 
@@
 
-\begin{block}{Variation among replicates}
-
-\small
-
-Here simulations where the signal has no cost
-
-\begin{center}\includegraphics[width=0.8\linewidth]{../Simulations/nIntGroupEvol1_/evolDynALL_nIntGroup2000} \end{center}
-
-\end{block}
+\begin{center}\includegraphics[width=0.9\linewidth]{../Simulations/nIntGroupEvol4_/evolDyn13_nIntGroup2000} \end{center}
 
@@
 
-\begin{block}{Variation within a replicate}
+\begin{figure}
 
-\small
+\includegraphics[height=0.28\textheight]{../Simulations/nIntGroupEvol4_/evolDistAlpha13_nIntGroup2000} \includegraphics[height=0.28\textheight]{../Simulations/nIntGroupEvol4_/evolDistBeta13_nIntGroup2000} \includegraphics[height=0.28\textheight]{../Simulations/nIntGroupEvol4_/evolDistBadge13_nIntGroup2000} \hfill{}
 
-Here simulations where the signal has no cost.
-
-There seems to be some consistency in the three type of sender codes
-
-\begin{center}\includegraphics[width=0.8\linewidth]{../Simulations/nIntGroupEvol1_/evolDyn8_nIntGroup2000} \end{center}
-
-\end{block}
+\caption{alpha - beta -Badge}\label{fig:unnamed-chunk-27}
+\end{figure}
 
@@
 
-\begin{block}{Variation within a replicate}
-
-\tiny
-
-Panels show the changes in frequency distribution along evolution
-(\(\alpha\), \(\beta\), and resulting badge respectively) for the
-replicate shown in the previous slide
-
-Branching events in the reaction norm parameters yield three types of
-sender codes. One (more or less) honest, two dishonest (one with a big
-badge and one with a small badge)
-
-\begin{figure}
-
-\includegraphics[height=0.28\textheight]{../Simulations/nIntGroupEvol1_/evolDistAlpha8_nIntGroup2000} \includegraphics[height=0.28\textheight]{../Simulations/nIntGroupEvol1_/evolDistBeta8_nIntGroup2000} \includegraphics[height=0.28\textheight]{../Simulations/nIntGroupEvol1_/evolDistBadge8_nIntGroup2000} \hfill{}
-
-\caption{alpha - beta -Badge}\label{fig:unnamed-chunk-22}
-\end{figure}
-
-\end{block}
-
-\end{frame}
-
-\begin{frame}{The evolution of cheap Badges of Status}
-\protect\hypertarget{the-evolution-of-cheap-badges-of-status-3}{}
-
 \begin{block}{How often does it happen?}
-
-\tiny
-
-Values of \(\alpha\) and \(\beta\) for each individual in the different
-replicates. Each dot an individual, each panel a replicate. Individuals
-are sampled from populations from the second half of the simulations.
-
-Individuals in some simulations split up in three groups, others in two
-groups; in one simulation the population does not split up
 
@@
 \begin{frame}{The evolution of cheap Badges of Status}
-\protect\hypertarget{the-evolution-of-cheap-badges-of-status-4}{}
+\protect\hypertarget{the-evolution-of-cheap-badges-of-status-3}{}
 
@@
 
-\end{block}
+--\textgreater{}
 
-\end{frame}
-
-\begin{frame}{Interaction structures - do they matter?}
-\protect\hypertarget{interaction-structures---do-they-matter}{}
-
-\begin{center}\includegraphics[width=1\linewidth]{../Images/networks} \end{center}
-
-(Chaine et al. 2018)
-
-\end{frame}
-
-\begin{frame}{Localized interaction structures}
-\protect\hypertarget{localized-interaction-structures}{}
-
-\begin{block}{Costly signals}
-
-\tiny
-
-Behavioural interactions and badge size for simulations with costs where
-individuals interact in small groups. On the \(x\) axes number of
-different individuals each focal interacts with.
-
-\begin{center}\includegraphics[width=0.9\linewidth]{../Simulations/nIntGroupCost_/BehavIntALL} \includegraphics[width=0.9\linewidth]{../Simulations/nIntGroupCost_/CueALL} \end{center}
-
-\end{block}
-
-\end{frame}
-
-\begin{frame}{Localized interaction structures}
-\protect\hypertarget{localized-interaction-structures-1}{}
-
-\begin{block}{Cheap signals (free)}
-
-\small
-
-Variation among individuals in different replicates (just like before)
-in replicates where individuals interact with 10 other individuals.
-
-A part from replicates with two clusters, the clustering seems less
-straightforward as before
-
-\begin{center}\includegraphics[width=0.85\linewidth]{../Simulations/nIntGroupEvol1_/corrAlphBet_nIntGroup10} \end{center}
-
-\end{block}
-
-\end{frame}
-
-\begin{frame}{Localized interaction structures}
-\protect\hypertarget{localized-interaction-structures-2}{}
-
-\begin{block}{Cheap signals (free)}
-
-\small
-
-Behavioural interactions for the replicates form previous slide. I split
-up the replicates according to clusters (despite clusters not being
-clear cut).
-
-Only replicates with only one cluster seems to have less aggressive
-interactions
-
-\begin{center}\includegraphics[width=0.9\linewidth]{../Simulations/nIntGroupEvol1_/BehavIntAllClusters_nIntGroup10} \end{center}
-
-\end{block}
-
-\end{frame}
-
-\begin{frame}{Localized interaction structures}
-\protect\hypertarget{localized-interaction-structures-3}{}
-
-\begin{block}{Cheap signals or Individual recognition?}
-
-\tiny
-
-In these set, I changed the learning set up, I increased the number of
-centers (10 now, 6 before) along the signal axis; and I reduced the
-level or generalization.
-
-There seems to be an effect on the slope of the reaction norm. It stays
-very close to zero.
-
-\begin{center}\includegraphics[width=0.9\linewidth]{../Simulations/nIntGroupCenters1_/evolDynALL_nIntGroup10} \end{center}
-
-\end{block}
-
-\end{frame}
-
-\begin{frame}{Localized interaction structures}
-\protect\hypertarget{localized-interaction-structures-4}{}
-
-\begin{block}{Cheap signals or Individual recognition}
-
-\small
-
-Same replicates as the previous slide. Variation in \(\beta\) and
-\(\alpha\) among individuals for different replicates.
-
-No clear clustering. Variation mostly along the \(\alpha\) axis
-
-\begin{center}\includegraphics[width=0.9\linewidth]{../Simulations/nIntGroupCenters1_/corrAlphBet_nIntGroup10} \end{center}
-
-\end{block}
-
-\end{frame}
-
-\begin{frame}{Localized interaction structures}
-\protect\hypertarget{localized-interaction-structures-5}{}
-
-\begin{block}{Cheap signals or Individual recognition?}
-
-\small
-
-Panels show the changes in frequency distribution along evolution
-(\(\alpha\), \(\beta\), and resulting badge respectively) for the
-replicate shown in the previous slide.
-
-\(\beta\) around zero, \(\alpha\) all over the place.
-
-\begin{flushleft}\includegraphics[height=0.28\textheight]{../Simulations/nIntGroupCenters1_/evolDistAlpha5_nIntGroup10} \includegraphics[height=0.28\textheight]{../Simulations/nIntGroupCenters1_/evolDistBeta5_nIntGroup10} \includegraphics[height=0.28\textheight]{../Simulations/nIntGroupCenters1_/evolDistBadge5_nIntGroup10} \end{flushleft}
+--\textgreater{}
 
@@
 \hypertarget{refs}{}
-\leavevmode\hypertarget{ref-chaine_Manipulating_2018}{}%
-Chaine, Alexis S., Daizaburo Shizuka, Theadora A. Block, Lynn Zhang, and
-Bruce E. Lyon. 2018. ``Manipulating Badges of Status Only Fools
-Strangers.'' \emph{Ecology Letters} 21 (10): 1477--85.
-\url{https://doi.org/10.1111/ele.13128}.
+\leavevmode\hypertarget{ref-botero_Evolution_2010}{}%
+Botero, C. A, I. Pen, J. Komdeur, and F. J Weissing. 2010. ``The
+Evolution of Individual Variation in Communication Strategies.''
+\emph{Evolution} 64 (11): 3123--33.
 

### 2020-11-03 Dyn_reac_learn_noLearn.R

@@ -12,7 +12,7 @@ require("jsonlite")
 
 # Scenario to be plotted - corresponds to folders where simulations are stored
 
-scenario<-"betCostnoLearn"
+scenario<-"betCostNoLearn"
 SimDir<-"Simulations"
 
 extSimsDir<-#here("Simulations",paste0(scenario,"_"))
@@ -35,7 +35,7 @@ param<-fromJSON(here(SimDir,paste0(scenario,"_"),paramName[1]))
 numCores <- length(evolList)
 registerDoParallel(numCores)
 
-val<-4
+val<-2
 
 
 # loop to produce pdfs for parameter values
@@ -249,7 +249,7 @@ traitsTrajs<-dcast(evol,time~seed,
 
 
 for(runChoi in finReps){
-   runChoi<-0
+   # runChoi<-6
   
 # Dynamics of genotypic traits (reaction norm) -  signaller
 par(plt=posPlot(numploty = 3,idploty = 2,numplotx = 3,idplotx = 1,
@@ -323,7 +323,6 @@ polygon(x=c(traitsTrajs[,time],rev(traitsTrajs[,time])),
                                       paste0("sdGammaAtt_",runChoi))])),
         col=colGenesPol[3],border = NA)
 
-traitsTrajs[,.SD,.SDcol=grep("_1",names(traitsTrajs))]
 
 matlines(x=traitsTrajs[,time],
          y=traitsTrajs[,.SD,
@@ -339,35 +338,30 @@ axis(4,cex=0.8)
 # dynamics of behavioural interactions 
 
 par(plt=posPlot(numploty = 3,idploty = 2,numplotx = 3, idplotx = 3,
-                lowboundx = 8,upboundx = 93),
+                lowboundx = 8, upboundx = 93),
     xaxt="s",las=1,new=TRUE)
-plot(x=c(0,max(evolStats$time)),y=c(0,0),type="l",lwd=2,col=0,
-     ylim=c(0,1),xlab="",ylab="",yaxt="n",
+plot(x=c(0,max(evol$time)),y=c(0,0),type="l",lwd=2,col=0,
+     ylim=c(0,0.6),xlab="",ylab="",yaxt="n",
      cex.lab=1.5,cex.axis=1,xaxt='n',las=1)
[…]

### 2020-11-03 Figure_dyn_AGA.R

@@ -1,0 +1,342 @@
+# Evolutionary dynamics for averages and individual runs, reaction norms of sampled 
+# individuals and their learning profiles
+
+# Required libraries -----------------------------------------------------------
+
+require(here)
+here()
+source(here("AccFunc.R"))
+require("foreach")
+require("doParallel")
+require("jsonlite")
+
+# Scenario to be plotted - corresponds to folders where simulations are stored
+
+scenario<-"nIntGroupEvol4TestSeed"
+
+extSimsDir<-#here("Simulations",paste0(scenario,"_"))
+  paste0("e:/BadgeSims/",scenario,"_")
+
+
+# Load files -------------------------------------------------------------------
+# Project folder
+# (listTest<-list.files(here("Simulations",paste0(scenario,"_"))))
+# External sims folder
+(listTest<-list.files(extSimsDir,full.names = TRUE))
+
+(evolList<-grep("evolLearn",listTest,value=TRUE))
+(indList<-grep("indLearn",listTest,value=TRUE))
+paramName<-list.files(here("Simulations",paste0(scenario,"_")))
+paramName<-list.files(extSimsDir,full.names = TRUE)
+paramName<-grep(".json",paramName,value=TRUE)
+param<-#fromJSON(here("Simulations",paste0(scenario,"_"),paramName))
+  fromJSON(paramName[1])
+
+
+val<-1
+
+# loop to produce pdfs for parameter values
+# numCores <- length(indList)
+# registerDoParallel(numCores)
+# foreach(val = 1:length(indList),.packages = c("data.table","here")) %dopar% {
+#   source(here("AccFunc.R"))
+  
+  fileId<-val
+  
+  # Project folder
+  # evol<-fread(here("Simulations",paste0(scenario,"_"),evolList[fileId]))
+  # pop<-fread(here("Simulations",paste0(scenario,"_"),indList[fileId]))
+  
+  # External sims folder
+  evol<-fread(evolList[fileId])
+  pop<-fread(indList[fileId])
+  
+  Valpar<-gsub("[[:alpha:]]",gsub(".txt","",tail(strsplit(indList[val],"_")[[1]],1)),
+               replacement = "")
+  nampar<-gsub("[^[:alpha:]]",gsub(".txt","",tail(strsplit(indList[val],"_")[[1]],1)),
+               replacement = "")
+  
+  nCenters<-param$nCenters
[…]

### 2020-11-03 clusterAnalysis.R

@@ -1,0 +1,199 @@
+require(here)
+here()
+source(here("AccFunc.R"))
+require("foreach")
+require("doParallel")
+require("jsonlite")
+require("factoextra")
+require("cluster")
+
+# Scenario to be plotted - corresponds to folders where simulations are stored
+
+scenario<-"nIntGroupEvol4"
+
+extSimsDir<-#here("Simulations",paste0(scenario,"_"))
+  paste0("e:/BadgeSims/",scenario,"_")
+
+
+# Load files -------------------------------------------------------------------
+# Project folder
+# (listTest<-list.files(here("Simulations",paste0(scenario,"_"))))
+# External sims folder
+(listTest<-list.files(extSimsDir,full.names = TRUE))
+
+(evolList<-grep("evolLearn",listTest,value=TRUE))
+(indList<-grep("indLearn",listTest,value=TRUE))
+paramName<-list.files(here("Simulations",paste0(scenario,"_")))
+paramName<-list.files(extSimsDir,full.names = TRUE)
+paramName<-grep(".json",paramName,value=TRUE)
+param<-#fromJSON(here("Simulations",paste0(scenario,"_"),paramName))
+  fromJSON(paramName)
+
+
+val<-1
+
+# loop to produce pdfs for parameter values
+# numCores <- length(indList)
+# registerDoParallel(numCores)
+# foreach(val = 1:length(indList),.packages = c("data.table","here")) %dopar% {
+#   source(here("AccFunc.R"))
+  
+  fileId<-val
+  
+  # Project folder
+  # evol<-fread(here("Simulations",paste0(scenario,"_"),evolList[fileId]))
+  # pop<-fread(here("Simulations",paste0(scenario,"_"),indList[fileId]))
+  
+  # External sims folder
+  evol<-fread(evolList[fileId])
+  pop<-fread(indList[fileId])
+  
+  Valpar<-gsub("[[:alpha:]]",gsub(".txt","",tail(strsplit(indList[val],"_")[[1]],1)),
+               replacement = "")
+  nampar<-gsub("[^[:alpha:]]",gsub(".txt","",tail(strsplit(indList[val],"_")[[1]],1)),
+               replacement = "")
+  
[…]

### 2020-11-03 compPar.R

@@ -8,15 +8,21 @@ source(here("AccFunc.R"))
 
 # Scenario to be plotted - corresponds to folders where simulations are stored
 
-scenario<-"betCost"
+scenario<-"betCostEvol1"
+
+extSimsDir<-#here("Simulations",paste0(scenario,"_"))
+  paste0("e:/BadgeSims/",scenario,"_")
 
 # Load files -------------------------------------------------------------------
 
 (listTest<-list.files(here("Simulations",paste0(scenario,"_"))))
+
+(listTest<-list.files(extSimsDir,full.names = TRUE))
+
 (sdList<-grep("evol",listTest,value=TRUE))
 (indList<-grep("ind",listTest,value=TRUE))
 
-inds<-do.call(rbind,lapply(indList,filesScenar,scenario))
+inds<-do.call(rbind,lapply(indList,filesScenar,scenario,full.name=TRUE))
 
 inds[,diffActWeight:=abs(WeightAct_0-WeightAct_4)]
 

### 2020-11-03 cue_compt.cpp

@@ -27,7 +27,7 @@ Start date :
 #include <math.h>	
 #include <vector>
 #include <string.h>
-#include <omp.h>
+//#include <omp.h>
 //#include "tchar.h"   //Eliminate for g++
 #include "../Cpp/Routines/C++/RandomNumbers/random.h"
 #include "../Cpp/json.hpp"       
@@ -580,8 +580,8 @@ void interactions(vector<individual>& population, int nint,
 	//cout << "track " << generat << endl;
 	localPrint.countIntTypes[0] = 0, localPrint.countIntTypes[1] = 0;
 	localPrint.countIntTypes[2] = 0;
+	localPrint.countPhenotypes[0] = 0, localPrint.countPhenotypes[1] = 0;
 	if ((population[0].get_strat() == learner) && (trackPopLearn)) {
-		localPrint.countPhenotypes[0] = 0, localPrint.countPhenotypes[1] = 0;
 		/*localPrint.countIntTypesGen[0];
 		localPrint.countIntTypesGen[1], localPrint.countIntTypesGen[2] = 0;*/
 		for (int countSam = 0; countSam < sampleSize; ++countSam) {
@@ -740,6 +740,11 @@ string create_filename(std::string filename, json param, int idRangPar) {
 	std::string namParam = param["namParam"];
 	filename.append(namParam);
 	filename.append(douts(param["rangParam"][idRangPar]));
+	if ((int)(param["nRep"]) == 1) {
+		filename.append("_");
+		filename.append("seed");
+		filename.append(douts(param["seed"]));
+	}
 	filename.append(".txt");
 	return(filename);
 }
@@ -816,6 +821,7 @@ int main(int argc, char* argv[]) {
 	//json param;
 	//param["totGen"]            = 10;   // Total number of generations
 	//param["nRep"]			   = 5;     // Number of replicates
+	//param["seed"]				=0;
 	//param["printGen"]          = 1;     // How often data is printed	
 	//param["printLearn"]        = 1;	  // how often learning dyn are printed
 	//param["printLearnInt"]	   = 1;   // How often are learning parameters printed
@@ -871,40 +877,46 @@ int main(int argc, char* argv[]) {
 	else{ omp_get_max_threads(); }*/
 	
 	//pointParam = &param;
-
 	string namParam = param["namParam"];
 	vector<ofstream>  evolOutput(param["rangParam"].size());
[…]

### 2020-11-03 parTOjson.R

@@ -10,13 +10,13 @@ fileName<-"parameters"
 here()
 
 param1<-#fromJSON(paste0("E:/BadgeSims","/betCostNoLearn_","/parameters.json"))
-     fromJSON(here("Simulations","betCostNoLearn_","parameters.json"))
+     fromJSON(here("Simulations","nIntGroupEvol4_","parameters.json"))
 
-param<-list(totGen   = 15000,   nRep    = 8,
+param<-list(totGen   = 15000,   nRep    = 1, seed = 1,
             printGen = 1000,   printLearn = 1000,
             printLearnInt = 50,
             popSize  = 2000,  baselineFit = 2,   
-            MutSd    = 0.3,  nInt        = 100,  init     = c(0,0,0,1),
+            MutSd    = 0.3,  nInt        = 500,  init     = c(0,0,1,0),
             mutRate  = 0.001,  mutType  = 0,
             sampleSize = 50,   strQual  = 10,
             errorQual = 0,    alphaBad	 = 0,    betaBad	 = 0,
@@ -28,14 +28,14 @@ param<-list(totGen   = 15000,   nRep    = 8,
             alphCost	 = 3, mutLearn = FALSE,
             nIntGroup  = 2000, 
             payoff_matrix = c(1.5,1,0,0.5),
-            namParam = "betCost",
-            rangParam = c(0,5),
-            typeAgent = 3, #//0. hawk 1.dove 2. learner 3. evaluator
+            namParam = "nIntGroup",
+            rangParam = c(2000,8),
+            typeAgent = 2, #//0. hawk 1.dove 2. learner 3. evaluator
             folderL=paste(here("Simulations"),"/",sep="")) # comment for debug
 # folderL=paste(here("Simulations","test_"),"/",sep="")) # comment for realease
 
 
-apendScenar<-"NoLearn"
+apendScenar<-"Evol5"
 param$folderL<-paste0(param$folderL,
        param$namParam,apendScenar,"_/")
 # runTime<-"360:00:00"# "10:00:00"# 
@@ -53,7 +53,7 @@ param$folderL<-paste0(param$folderL,
 param$folder<-paste0("/home/ubuntu/BadgeStatus/",
                      param$namParam,apendScenar,"_/")
 ## For the local pc
-param$folder<-paste0("e:/BadgeSims/",param$namParam,apendScenar,"_/")
[…]

### 2020-11-25 Presentations/GenesAsEnviron.Rmd
AccFunc: use countour3 in evolDist to include several panels in a figure. Figure files: use several files to upload one scenario. Aga_figures: finished Improve cluster analysis cue_compt.cpp: Allow to insert types at the beggining of the simulations New simulation with different sd for the mutation kernel

@@
     | Genes as environment - AGA2020 virtual symposium
-    | `r format(Sys.time(), '%B %d, %Y')`
+    | November 17 - 2020
 output: 
@@
 ---
+
+## Acknowledgements
+
+| ![Red1](../Images/redouan.jpg){width=33%} | ![Olle](../Images/olof-leimar.png){width=23%} | ![Daniel](../Images/danielcadena.jpg){width=32%}  |
+| :---:        |    :----:   |    :---: |
+| *R. Bshary* |*O. Leimar*|*D. Cadena*|
+
+![unine](../Images/UNINE_cmjn.jpg){width=35%}
+![uniandes](../Images/logo_largo.png){width=45%}
 
@@
 
-```{r,echo=FALSE,fig.align="center",out.width='40%',fig.show='hold'}
+```{r,echo=FALSE,fig.align="center",out.width='40%'}
 
 knitr::include_graphics("../Images/sexSelec.jpg")
-knitr::include_graphics("../Images/begging.jpg")
-knitr::include_graphics("../Images/broodparasite.jpg")
 knitr::include_graphics("../Images/dominance.jpg")
@@
 
-```{r,echo=FALSE,fig.align='center',out.height='80%'}
-knitr::include_graphics("../Images//hosp_alexandra_mackenzie.jpg")
-
+```{r,echo=FALSE,fig.align='center',out.width='45%',fig.show='hold'}
+knitr::include_graphics("../Images/HouseSparrow_strong.jpg")
+knitr::include_graphics("../Images/HouseSparrow_weak.jpg")
 ```
@@
 
-## The handicap principle
+<!-- ## The handicap principle -->
 
-### Badges of status
+<!-- ### Badges of status -->
 
-```{r,echo=FALSE,fig.align='center',out.width='90%'}
-knitr::include_graphics("../Images//comSystemBot.png")
-```
+<!-- ```{r,echo=FALSE,fig.align='center',out.width='90%'} -->
+<!-- knitr::include_graphics("../Images//comSystemBot.png") -->
+<!-- ``` -->
 
-[@botero_Evolution_2010]
+<!-- [@botero_Evolution_2010] -->
 
-## Are bibs really badges?
+<!-- ## Are bibs really badges? -->
 
-```{r,echo=FALSE,fig.align='left',out.width='30%'}
-knitr::include_graphics("../Images/Sanchez-Tojar_2018.png")
-```
+<!-- ```{r,echo=FALSE,fig.align='left',out.width='30%'} -->
+<!-- knitr::include_graphics("../Images/Sanchez-Tojar_2018.png") -->
+<!-- ``` -->
+
+<!-- [@sanchez-tojar_Metaanalysis_2018a] -->
 
@@
 
-\vspace{-0.5cm}
+\vspace{-0.3cm}
 
-### Comunication
+### Conflict
+
+\vspace{0.3cm}
+\begin{columns}
+\column{0.5\linewidth}
+hawk-dove game
+\tiny
+\begin{center}
+\begin{tabular}{ >{\centering\arraybackslash}p{0.2cm} | >{\centering\arraybackslash}p{2.5cm} | >{\centering\arraybackslash}p{0.8cm} }
+& H & D \\ \hline
+H & $p_w V\frac{-C}{2} + (1-p_w) \frac{-C}{2}$ & $V$ \\ \hline
+V & $0$ & $\frac{V}{2}$\\
+\end{tabular}
+\end{center}
+
+\begin{equation*}
+p_w=\frac{1}{1+e^{-\beta(Q_i-Q_j)}}
+\end{equation*}
+
+\column{0.5\linewidth}
+
+```{r,echo=FALSE,fig.align='center',out.width='60%'}
+par(plt=posPlot(),las=1)
+plot(logist(seq(-1,1,length.out = 1000),alpha = 0,beta = 2)~
+       seq(-1,1,length.out = 1000),type="l",xlab="",
+     ylab = expression(p[w]),cex.lab=3,lwd=3,col="red",cex.axis=2,ylim=c(0,1))
+ mtext("Difference in quality",1,line = 4,cex = 3,las = 0)
+lines(logist(seq(-1,1,length.out = 1000),alpha = 0,beta = 10)~
+       seq(-1,1,length.out = 1000),lwd=3,col="springgreen4")
+legend("topleft",legend=c(2,10),col=c("red","springgreen4"),lwd=3,title = expression(beta),cex=2,bty="n")
+```
+
+\end{columns}
+
+\pause
+
+### Communication
 
@@
  linesSender<-sapply(rangeBeta,function(x)logist(rangQual,alpha = x/2,beta = x))
- plot(logist(rangQual,alpha = 0,beta = 0)~rangQual,type="l",xlab="Quality",
-      ylab = expression(p[w]),cex.lab=2,lwd=2,col="grey",cex.axis=2,ylim=c(0,1))
+ plot(logist(rangQual,alpha = 0,beta = 0)~rangQual,type="l",xlab="",
+      ylab ="",cex.lab=2,lwd=2,col="grey",cex.axis=2,ylim=c(0,1))
+ mtext("Quality",1,line = 4,cex = 3,las = 0)
  matlines(x=rangQual,y=linesSender,lwd=4,col=colRuns,lty=1)
+ mtext("Badge size",2,line = 4,cex = 3,las = 0)
  legend("left",legend=rangeBeta,col=colRuns,lwd=3,title = expression(beta),cex=2,bty="n")
@@
 plot(logist(totRBF(rangx,centers,sigSq,weights),alpha = 0,beta=1)~rangx,type='l',col=1,
-     xlab="",ylab="p(retreat)",ylim=c(0,1),lwd=3,xaxt="n",cex.lab=2,cex.axis=2)
+     xlab="",ylab="",ylim=c(0,1),lwd=3,xaxt="s",cex.lab=2,cex.axis=2)
+# test<-sapply(1:6, function(x){RBF()})
 points(y=logist(weights,alpha = 0,beta=1),x=centers,cex=3,pch=20)
+# xRBFs<-sapply(centers, function(x){seq(x-0.2,x+0.2,by=0.001)})
+# RBFS<-sapply(1:6, function(x){
+#   logist(weights[x]*RBF(seq(centers[x]-0.2,centers[x]+0.2,by=0.001),xCenter = centers[x],
+#                         sigSq = 0.01))
+# })
+# matlines(x=xRBFs,y=RBFS,col="grey",lty=2,lwd=2)
+
+mtext("p(dove)",2,line = 4,cex = 3,las = 0)
+mtext("Badge size",1,line = 4,cex = 3,las = 0)
 
@@
 \end{columns}
-
-### Conflict
-
-\vspace{0.3cm}
-\begin{columns}
-\column{0.5\linewidth}
-hawk-dove game
-\tiny
-\begin{center}
-\begin{tabular}{ >{\centering\arraybackslash}p{0.2cm} | >{\centering\arraybackslash}p{2.5cm} | >{\centering\arraybackslash}p{0.8cm} }
-& H & D \\ \hline
-H & $p_w V\frac{-C}{2} + (1-p_w) \frac{-C}{2}$ & $V$ \\ \hline
-V & $0$ & $\frac{V}{2}$\\
-\end{tabular}
-\end{center}
-
-\begin{equation*}
-p_w=\frac{1}{1+e^{-\beta(Q_i-Q_j)}}
-\end{equation*}
-
-\column{0.5\linewidth}
-
-```{r,echo=FALSE,fig.align='center',out.width='60%'}
-par(plt=posPlot(),las=1)
-plot(logist(seq(-1,1,length.out = 1000),alpha = 0,beta = 2)~
-       seq(-1,1,length.out = 1000),type="l",xlab="Difference in quality",
-     ylab = expression(p[w]),cex.lab=2,lwd=3,col="red",cex.axis=2,ylim=c(0,1))
-lines(logist(seq(-1,1,length.out = 1000),alpha = 0,beta = 10)~
-       seq(-1,1,length.out = 1000),lwd=3,col="green")
-legend("topleft",legend=c(2,10),col=c("red","green"),lwd=3,title = expression(beta),cex=2,bty="n")
-```
-
-\end{columns}
-
-# Results
 
@@
 
-\pause
 ```{r, echo=FALSE,fig.align='center', out.width = '100%'}
-knitr::include_graphics("../Simulations/alphaAct_/alphaAct0.6learnDyn.png")
+knitr::include_graphics("../Simulations/alphaAct_/alphaAct0.4learnDyn.png")
 ```
 
-<!-- ## Learning can mediate the evolution of Badges of Status -->
-<!-- ### Variation among replicates -->
-<!-- \small -->
-<!-- set of simulations with the badge working as a handicap (cost inversely proportional to quality) -->
-
-<!-- ```{r, echo=FALSE,fig.align='center', out.width='100%'} -->
-<!-- knitr::include_graphics("../Simulations/betCostEvol1_/evolDynAll_betCost5.png") -->
-<!-- ``` -->
+# Results
 
@@
 <!-- One of the replicates of the previous slide, the badge as a handicap -->
-```{r, echo=FALSE,fig.align='center', out.width='100%'}
+```{r, echo=FALSE,fig.align='center', out.width='90%'}
 knitr::include_graphics("../Simulations/betCostEvol1_/evolDyn0_betCost5.png")
@@
 
-```{r, echo=FALSE,fig.align='center', out.width='100%'}
+```{r, echo=FALSE,fig.align='center', out.height='90%'}
 knitr::include_graphics("../Simulations/betCostEvol1_/evolDyn5_betCost5.png")
@@
 
-## Badges of Status: do they matter?
-### Behavioural interactions
+<!-- ## Badges of Status: do they matter? -->
+<!-- ### Behavioural interactions -->
 
@@
 
-```{r, echo=FALSE,fig.align='center', out.width='90%'}
-knitr::include_graphics("../Simulations/betCostEvol1_/compClustRep_betCost5.png")
-```
+<!-- ```{r, echo=FALSE,fig.align='center', out.width='90%'} -->
+<!-- knitr::include_graphics("../Simulations/betCostEvol1_/compClustRep_betCost5.png") -->
+<!-- ``` -->
 
@@
 
-<!-- <!-- ### Variation among replicates --> -->
-<!-- <!-- \small --> -->
-<!-- <!-- Here simulations where the signal has no cost --> -->
+<!-- <!-- ### Variation among replicates --> 
+<!-- <!-- \small --> 
+<!-- <!-- Here simulations where the signal has no cost -->
 
@@
 
-## The evolution of cheap Badges of Status
+## The evolution of cheap badges of Status
 
@@
 ```{r, echo=FALSE,fig.align='center', out.width='90%'}
-knitr::include_graphics("../Simulations/nIntGroupEvol4_/evolDyn13_nIntGroup2000.png")
+knitr::include_graphics("../Simulations/nIntGroupNormQual_/evolDyn7_nIntGroup2000.png")
 ```
 
-## The evolution of cheap Badges of Status
-
-<!-- \tiny -->
-<!-- Panels show the changes in frequency distribution along evolution ($\alpha$, $\beta$, and resulting badge respectively) for the replicate shown in the previous slide -->
-
-<!-- Branching events in the reaction norm parameters yield three types of sender codes. One (more or less) -->
-<!-- honest, two dishonest (one with a big badge and one with a small badge) -->
-
-```{r, echo=FALSE,fig.align='left', out.height='28%',fig.show='hold', fig.cap="alpha - beta -Badge"}
-knitr::include_graphics("../Simulations/nIntGroupEvol4_/evolDistAlpha13_nIntGroup2000.png")
-knitr::include_graphics("../Simulations/nIntGroupEvol4_/evolDistBeta13_nIntGroup2000.png")
-knitr::include_graphics("../Simulations/nIntGroupEvol4_/evolDistBadge13_nIntGroup2000.png")
-```
-
-## The evolution of cheap Badges of Status
-### How often does it happen?
+<!-- ## The evolution of cheap Badges of Status -->
+<!-- ### How often does it happen? -->
 <!-- \tiny -->
@@
 
-```{r, echo=FALSE,fig.align='left', out.width='90%',fig.show='hold'}
-knitr::include_graphics("../Simulations/nIntGroupEvol1_/corrAlphBet_nIntGroup2000.png")
-```
-
-## The evolution of cheap Badges of Status
-### Does it matter?
-\small
-Behavioural interactions splitting up the replicates according to the number of "types" that 
-evolve
-
-```{r, echo=FALSE,fig.align='left', out.width='90%',fig.show='hold'}
-knitr::include_graphics("../Simulations/nIntGroupEvol1_/BehavIntAllClusters_nIntGroup2000.png")
-```
-
-<!-- ## Interaction structures - do they matter? -->
-
-<!-- ```{r, echo=FALSE,fig.align='center', out.width='100%',} -->
-<!-- knitr::include_graphics("../Images/networks.png") -->
-<!-- ``` -->
-<!-- [@chaine_Manipulating_2018] -->
-
-<!-- ## Localized interaction structures -->
-
-<!-- ### Costly signals  -->
-<!-- \tiny -->
-<!-- Behavioural interactions and badge size for simulations with costs where individuals interact in small groups. On the $x$ axes number of different individuals each focal interacts with. -->
-
-<!-- ```{r, echo=FALSE,fig.align='center', out.width='90%',fig.show='hold'} -->
-<!-- knitr::include_graphics("../Simulations/nIntGroupCost_/BehavIntALL.png") -->
-<!-- knitr::include_graphics("../Simulations/nIntGroupCost_/CueALL.png") -->
+<!-- ```{r, echo=FALSE,fig.align='left', out.width='90%',fig.show='hold'} -->
+<!-- knitr::include_graphics("../Simulations/nIntGroupNormQual_/corrAlphBet_nIntGroup2000.png") -->
 <!-- ``` -->
 
-<!-- ## Localized interaction structures -->
+<!-- ## The evolution of cheap Badges of Status -->
+<!-- ### Does it matter? -->
+<!-- \small -->
+<!-- Behavioural interactions splitting up the replicates according to the number of "types" that  -->
+<!-- evolve -->
 
-<!-- ### Cheap signals (free) -->
-<!-- \small -->
-<!-- Variation among individuals in different replicates (just like before) in replicates where individuals interact with 10 other individuals. -->
-
-<!-- A part from replicates with two clusters, the clustering seems less straightforward as before -->
-
-<!-- ```{r, echo=FALSE,fig.align='center', out.width='85%',fig.show='hold'} -->
-<!-- knitr::include_graphics("../Simulations/nIntGroupEvol1_/corrAlphBet_nIntGroup10.png") -->
+<!-- ```{r, echo=FALSE,fig.align='left', out.width='90%',fig.show='hold'} -->
+<!-- knitr::include_graphics("../Simulations/nIntGroupEvol1_/BehavIntAllClusters_nIntGroup2000.png") -->
 <!-- ``` -->
-
-<!-- ## Localized interaction structures -->
-
-<!-- ### Cheap signals (free)  -->
-
-<!-- \small -->
-<!-- Behavioural interactions for the replicates form previous slide. I split up the replicates according to clusters (despite clusters not being clear cut). -->
-
-<!-- Only replicates with only one cluster seems to have less aggressive interactions -->
-
-<!-- ```{r, echo=FALSE,fig.align='center', out.width='90%',fig.show='hold'} -->
-<!-- knitr::include_graphics("../Simulations/nIntGroupEvol1_/BehavIntAllClusters_nIntGroup10.png") -->
-<!-- ``` -->
-
-<!-- ## Localized interaction structures -->
-
-<!-- ### Cheap signals or Individual recognition? -->
-<!-- \tiny -->
-<!-- In these set, I changed the learning set up, I increased the number of centers (10 now, 6 before) -->
-<!-- along the signal axis; and I reduced the level or generalization.  -->
-
-<!-- There seems to be an effect on the slope of the reaction norm. It stays very close to zero.  -->
-
-<!-- ```{r, echo=FALSE,fig.align='center', out.width='90%',fig.show='hold'} -->
-<!-- knitr::include_graphics("../Simulations/nIntGroupCenters1_/evolDynALL_nIntGroup10.png") -->
-<!-- ``` -->
-
-<!-- ## Localized interaction structures -->
-
-<!-- ### Cheap signals or Individual recognition -->
-<!-- \small -->
-<!-- Same replicates as the previous slide. Variation in $\beta$ and $\alpha$ among individuals for different replicates.  -->
-
-<!-- No clear clustering. Variation mostly along the $\alpha$ axis -->
-
-<!-- ```{r, echo=FALSE,fig.align='center', out.width='90%',fig.show='hold'} -->
-<!-- knitr::include_graphics("../Simulations/nIntGroupCenters1_/corrAlphBet_nIntGroup10.png") -->
-<!-- ``` -->
-<!-- ## Localized interaction structures -->
-
-<!-- ### Cheap signals or Individual recognition? -->
-<!-- \small -->
-<!-- Panels show the changes in frequency distribution along evolution ($\alpha$, $\beta$, and resulting badge respectively) for the replicate shown in the previous slide.  -->
-
-<!-- $\beta$ around zero, $\alpha$ all over the place.  -->
-
-<!-- ```{r, echo=FALSE,fig.align='left', out.height='28%',fig.show='hold'} -->
-<!-- knitr::include_graphics("../Simulations/nIntGroupCenters1_/evolDistAlpha5_nIntGroup10.png") -->
-<!-- knitr::include_graphics("../Simulations/nIntGroupCenters1_/evolDistBeta5_nIntGroup10.png") -->
-<!-- knitr::include_graphics("../Simulations/nIntGroupCenters1_/evolDistBadge5_nIntGroup10.png") -->
-<!-- ``` -->
-
-<!-- <!-- ## Aqui vamos! mirar correlaciones de 20 y 2000 luego interacciones --> -->
-
-<!-- <!-- ## Comparar scenarios --> -->
 
@@
 
-\pause
+\vspace{-1cm}
 
@@
 
-\tiny
-Here, I vary the aggressive tendency for unlearned individuals in these three set ups.
+<!-- \tiny -->
+<!-- Here, I vary the aggressive tendency for unlearned individuals in these three set ups. -->
 
-Agressive, peaceful, clever (starts at the ESS of the classic hawk-dove game)
+<!-- Agressive, peaceful, clever (starts at the ESS of the classic hawk-dove game) -->
 
-```{r, echo=FALSE,fig.align='left', out.width='90%',fig.show='hold'}
-knitr::include_graphics("../Images/cartoonRBF_InitAct.png")
+```{r,out.width='45%',fig.show='hold'}
+knitr::include_graphics("../Simulations/alphaAct_/alphaAct0.4learnDyn.png")
+# knitr::include_graphics("../Images/cartoonRBF_InitAct.png")
+nCenters<-6
+interv<-1/(nCenters-1)
+centers<-interv*seq(0,nCenters-1)
+sigSq<-0.01
+weights<-data.table(neutral=rep(0,nCenters),agressive=rep(-2,nCenters),
+        peaceful=rep(2,nCenters),clever=rep(-0.69,nCenters))
+rangx<-seq(0,1,length=1000)
+
+par(plt=posPlot(numploty = 2,idploty = 2),las=1)
+actor<-sapply(weights, Actor,centers=centers)
+
+par(plt=posPlot()*0.95)
+matplot(x=rangx,y=as.matrix(actor),type='l',col=c("black","red","blue","springgreen4"),lty=1,
+     xlab="",ylab="",ylim=c(0,1.1),lwd=5,xaxt="s",cex.lab=1.5,cex.axis=2)
+mtext(text = "p(dove)",2,line = 4,cex = 2.5,las=0)
+mtext(text = "Badge size",1,line = 3.5,cex = 2.5)
+points(y=logist(weights$neutral,alpha = 0,beta=1),x=centers,cex=3)
+text(y = actor[1,]+0.1,x=rep(0.5,4),col=c("black","red","blue","springgreen4"),
+     labels = names(weights),cex=2)
 ```
@@
 
-### The peaceful
+### The peaceful, the aggressive and the clever
 
-Variation among replicates
+```{r,echo=FALSE,out.width='90%',fig.show='hold'}
+knitr::include_graphics("../Simulations/initAct_/indVarScatter_1.png")
 
-```{r, echo=FALSE,fig.align='left', out.width='100%',fig.show='hold'}
-knitr::include_graphics("../Simulations/initAct_/evolDynALL_initAct2.png")
 ```
@@
 
-### The peaceful
+### The peaceful, the aggressive and the clever
 
-Variation within one replicate
+```{r,out.width='90%',fig.show='hold'}
+knitr::include_graphics("../Simulations/initAct_/indVarScatter_.png")
 
-```{r, echo=FALSE,fig.align='left', out.width='100%',fig.show='hold'}
-knitr::include_graphics("../Simulations/initAct_/evolDyn2_initAct2.png")
-```
-
-## Why would you want to be recognized?
-
-### The peaceful
-
-Changes in the frequency distributions
-
-```{r, echo=FALSE,fig.align='left', out.height='28%',fig.show='hold'}
-knitr::include_graphics("../Simulations/initAct_/evolDistAlpha2_initAct2.png")
-knitr::include_graphics("../Simulations/initAct_/evolDistBeta2_initAct2.png")
-knitr::include_graphics("../Simulations/initAct_/evolDistBadge2_initAct2.png")
-```
-
-## Why would you want to be recognized?
-
-### The peaceful
-
-Favours variation and cheap signals
-
-```{r, echo=FALSE,fig.align='left', out.width='100%',fig.show='hold'}
-knitr::include_graphics("../Simulations/initAct_/corrAlphBet_initAct2.png")
-```
-
-## Why would you want to be recognized?
-
-### The aggressive
-
-Variation among replicates
-
-```{r, echo=FALSE,fig.align='left', out.width='100%',fig.show='hold'}
-knitr::include_graphics("../Simulations/initAct_/evolDynALL_initAct-2.png")
-```
-
-## Why would you want to be recognized?
-
-### The aggressive
-
-Variation within one replicate
-
-```{r, echo=FALSE,fig.align='left', out.width='100%',fig.show='hold'}
-knitr::include_graphics("../Simulations/initAct_/evolDyn0_initAct-2.png")
-```
-
-## Why would you want to be recognized?
-
-### The aggressive
-
-Disfavours variation and signal evolution - no signalling
-
-```{r, echo=FALSE,fig.align='left', out.width='100%',fig.show='hold'}
-knitr::include_graphics("../Simulations/initAct_/corrAlphBet_initAct-2.png")
-```
-
-## Why would you want to be recognized?
-
-### The clever (ESS)
-
-Variation among replicates
-
-```{r, echo=FALSE,fig.align='left', out.width='100%',fig.show='hold'}
-knitr::include_graphics("../Simulations/initAct_/evolDynALL_initAct-0.69.png")
-```
-
-## Why would you want to be recognized?
-
-### The clever (ESS)
-
-Variation within one replicate
-
-```{r, echo=FALSE,fig.align='left', out.width='100%',fig.show='hold'}
-  knitr::include_graphics("../Simulations/initAct_/evolDyn0_initAct-0.69.png")
-```
-
-## Why would you want to be recognized?
-
-### The clever (ESS)
-
-Signal evolution seems to be driven by drift. Most replicates end up in one of the two extremes
-
-```{r, echo=FALSE,fig.align='left', out.width='100%',fig.show='hold'}
-  knitr::include_graphics("../Simulations/initAct_/corrAlphBet_initAct-0.69.png")
 ```
@@
 - Learning can mediate costly signals
-- More expensive signals seem be more likely to evolve
-- Less expensive signals provide more information
-- Local interactions do not impede the evolution of costly signals (as has been suggested)
 
@@
 individuals
-- Under local interactions learning seems to allow the evolution of IR
 - Cheap signals depend on the innate level of aggression (before learning) of individuals
-- Peace begets heterogeneity, aggression begets homogeneity
 

### 2020-11-25 Presentations/GenesAsEnviron.tex
AccFunc: use countour3 in evolDist to include several panels in a figure. Figure files: use several files to upload one scenario. Aga_figures: finished Improve cluster analysis cue_compt.cpp: Allow to insert types at the beggining of the simulations New simulation with different sd for the mutation kernel

@@
 \newif\ifbibliography
+\usepackage{longtable,booktabs}
+\usepackage{caption}
+% Make caption package work with longtable
+\makeatletter
+\def\fnum@table{\tablename~\thetable}
+\makeatother
+\usepackage{graphicx,grffile}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
 \setlength{\emergencystretch}{3em} % prevent overfull lines
@@
 \date{Genes as environment - AGA2020 virtual symposium\\
-November 02, 2020}
+November 17 - 2020}
 
@@
 \frame{\titlepage}
+
+\begin{frame}{Acknowledgements}
+\protect\hypertarget{acknowledgements}{}
+
+\begin{longtable}[]{@{}ccc@{}}
+\toprule
+\includegraphics[width=0.33\textwidth,height=\textheight]{../Images/redouan.jpg}
+&
+\includegraphics[width=0.23\textwidth,height=\textheight]{../Images/olof-leimar.png}
+&
+\includegraphics[width=0.32\textwidth,height=\textheight]{../Images/danielcadena.jpg}\tabularnewline
+\midrule
+\endhead
+\emph{R. Bshary} & \emph{O. Leimar} & \emph{D. Cadena}\tabularnewline
+\bottomrule
+\end{longtable}
+
+\includegraphics[width=0.35\textwidth,height=\textheight]{../Images/UNINE_cmjn.jpg}
+\includegraphics[width=0.45\textwidth,height=\textheight]{../Images/logo_largo.png}
+
+\end{frame}
 
@@
 
-\begin{center}\includegraphics[width=0.4\linewidth]{../Images/sexSelec} \includegraphics[width=0.4\linewidth]{../Images/begging} \includegraphics[width=0.4\linewidth]{../Images/broodparasite} \includegraphics[width=0.4\linewidth]{../Images/dominance} \end{center}
+\begin{center}\includegraphics[width=0.4\linewidth]{../Images/sexSelec} \end{center}
+
+\begin{center}\includegraphics[width=0.4\linewidth]{../Images/dominance} \end{center}
 
@@
 
-\begin{center}\includegraphics[height=0.8\textheight]{../Images//hosp_alexandra_mackenzie} \end{center}
+\begin{center}\includegraphics[width=0.45\linewidth]{../Images/HouseSparrow_strong} \includegraphics[width=0.45\linewidth]{../Images/HouseSparrow_weak} \end{center}
 
@@
 \end{block}
-
-\end{frame}
-
-\begin{frame}{The handicap principle}
-\protect\hypertarget{the-handicap-principle-1}{}
-
-\begin{block}{Badges of status}
-
-\begin{center}\includegraphics[width=0.9\linewidth]{../Images//comSystemBot} \end{center}
-
-(Botero et al. 2010)
-
-\end{block}
-
-\end{frame}
-
-\begin{frame}{Are bibs really badges?}
-\protect\hypertarget{are-bibs-really-badges}{}
-
-\begin{flushleft}\includegraphics[width=0.3\linewidth]{../Images/Sanchez-Tojar_2018} \end{flushleft}
 
@@
 
-\vspace{-0.5cm}
-
-\begin{block}{Comunication}
-
-\begin{columns}[T]
-
-\column{0.45\linewidth}
-
-The sender code
-
-\tiny
-\begin{equation*}
-B_i = \frac{1}{1+e^{-(\textcolor{red}{\alpha_s} +\textcolor{red}{\beta_s} Q_i)}}
-\end{equation*}
-
-\begin{center}\includegraphics[width=0.5\linewidth]{GenesAsEnviron_files/figure-beamer/unnamed-chunk-19-1} \end{center}
-
-\column{0.45\linewidth}
-
-The receiver code (learning)
-
-\tiny
-\begin{equation*}
-  \Delta V_{t(s)}=\alpha (R_t-V_t)
-\end{equation*}
-
-\begin{center}\includegraphics[width=0.6\linewidth]{GenesAsEnviron_files/figure-beamer/unnamed-chunk-20-1} \end{center}
-
-\end{columns}
-
-\end{block}
+\vspace{-0.3cm}
 
@@
 
-\begin{center}\includegraphics[width=0.6\linewidth]{GenesAsEnviron_files/figure-beamer/unnamed-chunk-21-1} \end{center}
+\begin{center}\includegraphics[width=0.6\linewidth]{GenesAsEnviron_files/figure-beamer/unnamed-chunk-17-1} \end{center}
+
+\end{columns}
+
+\pause
+
+\end{block}
+
+\begin{block}{Communication}
+
+\begin{columns}[T]
+
+\column{0.45\linewidth}
+
+The sender code
+
+\tiny
+\begin{equation*}
+B_i = \frac{1}{1+e^{-(\textcolor{red}{\alpha_s} +\textcolor{red}{\beta_s} Q_i)}}
+\end{equation*}
+
+\begin{center}\includegraphics[width=0.5\linewidth]{GenesAsEnviron_files/figure-beamer/unnamed-chunk-18-1} \end{center}
+
+\column{0.45\linewidth}
+
+The receiver code (learning)
+
+\tiny
+\begin{equation*}
+  \Delta V_{t(s)}=\alpha (R_t-V_t)
+\end{equation*}
+
+\begin{center}\includegraphics[width=0.6\linewidth]{GenesAsEnviron_files/figure-beamer/unnamed-chunk-19-1} \end{center}
 
@@
 
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
 \begin{frame}{Learning under honest signalling}
@@
 
-\pause
-
-\begin{center}\includegraphics[width=1\linewidth]{../Simulations/alphaAct_/alphaAct0.6learnDyn} \end{center}
+\begin{center}\includegraphics[width=1\linewidth]{../Simulations/alphaAct_/alphaAct0.4learnDyn} \end{center}
 
 \end{frame}
+
+\hypertarget{results}{%
+\section{Results}\label{results}}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{../Simulations/betCostEvol1_/evolDyn0_betCost5} \end{center}
+\begin{center}\includegraphics[width=0.9\linewidth]{../Simulations/betCostEvol1_/evolDyn0_betCost5} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{../Simulations/betCostEvol1_/evolDyn5_betCost5} \end{center}
+\begin{center}\includegraphics[height=0.9\textheight]{../Simulations/betCostEvol1_/evolDyn5_betCost5} \end{center}
 
@@
 
-\begin{frame}{Badges of Status: do they matter?}
-\protect\hypertarget{badges-of-status-do-they-matter}{}
-
-\begin{block}{Behavioural interactions}
-
-\begin{center}\includegraphics[width=0.9\linewidth]{../Simulations/betCostEvol1_/compClustRep_betCost5} \end{center}
-
---\textgreater{} --\textgreater{} --\textgreater{}
-
-\end{block}
-
-\end{frame}
-
-\begin{frame}{The evolution of cheap Badges of Status}
+\begin{frame}{The evolution of cheap badges of Status}
 \protect\hypertarget{the-evolution-of-cheap-badges-of-status}{}
 
-\begin{center}\includegraphics[width=0.9\linewidth]{../Simulations/nIntGroupEvol4_/evolDyn13_nIntGroup2000} \end{center}
-
-\end{frame}
-
-\begin{frame}{The evolution of cheap Badges of Status}
-\protect\hypertarget{the-evolution-of-cheap-badges-of-status-1}{}
-
-\begin{figure}
-
-\includegraphics[height=0.28\textheight]{../Simulations/nIntGroupEvol4_/evolDistAlpha13_nIntGroup2000} \includegraphics[height=0.28\textheight]{../Simulations/nIntGroupEvol4_/evolDistBeta13_nIntGroup2000} \includegraphics[height=0.28\textheight]{../Simulations/nIntGroupEvol4_/evolDistBadge13_nIntGroup2000} \hfill{}
-
-\caption{alpha - beta -Badge}\label{fig:unnamed-chunk-27}
-\end{figure}
-
-\end{frame}
-
-\begin{frame}{The evolution of cheap Badges of Status}
-\protect\hypertarget{the-evolution-of-cheap-badges-of-status-2}{}
-
-\begin{block}{How often does it happen?}
-
-\begin{flushleft}\includegraphics[width=0.9\linewidth]{../Simulations/nIntGroupEvol1_/corrAlphBet_nIntGroup2000} \end{flushleft}
-
-\end{block}
-
-\end{frame}
-
-\begin{frame}{The evolution of cheap Badges of Status}
-\protect\hypertarget{the-evolution-of-cheap-badges-of-status-3}{}
-
-\begin{block}{Does it matter?}
-
-\small
-
-Behavioural interactions splitting up the replicates according to the
-number of ``types'' that evolve
-
-\begin{flushleft}\includegraphics[width=0.9\linewidth]{../Simulations/nIntGroupEvol1_/BehavIntAllClusters_nIntGroup2000} \end{flushleft}
-
---\textgreater{}
-
---\textgreater{}
-
-\end{block}
+\begin{center}\includegraphics[width=0.9\linewidth]{../Simulations/nIntGroupNormQual_/evolDyn7_nIntGroup2000} \end{center}
 
@@
 
-\tiny
-
-Here, I vary the aggressive tendency for unlearned individuals in these
-three set ups.
-
-Agressive, peaceful, clever (starts at the ESS of the classic hawk-dove
-game)
-
-\begin{flushleft}\includegraphics[width=0.9\linewidth]{../Images/cartoonRBF_InitAct} \end{flushleft}
+\includegraphics[width=0.45\linewidth]{../Simulations/alphaAct_/alphaAct0.4learnDyn}
+\includegraphics[width=0.45\linewidth]{GenesAsEnviron_files/figure-beamer/unnamed-chunk-24-2}
 
@@
 
-\begin{block}{The peaceful}
+\begin{block}{The peaceful, the aggressive and the clever}
 
-Variation among replicates
-
-\begin{flushleft}\includegraphics[width=1\linewidth]{../Simulations/initAct_/evolDynALL_initAct2} \end{flushleft}
+\includegraphics[width=0.9\linewidth]{../Simulations/initAct_/indVarScatter_1}
 
@@
 
-\begin{block}{The peaceful}
+\begin{block}{The peaceful, the aggressive and the clever}
 
-Variation within one replicate
-
-\begin{flushleft}\includegraphics[width=1\linewidth]{../Simulations/initAct_/evolDyn2_initAct2} \end{flushleft}
-
-\end{block}
-
-\end{frame}
-
-\begin{frame}{Why would you want to be recognized?}
-\protect\hypertarget{why-would-you-want-to-be-recognized-3}{}
-
-\begin{block}{The peaceful}
-
-Changes in the frequency distributions
-
-\begin{flushleft}\includegraphics[height=0.28\textheight]{../Simulations/initAct_/evolDistAlpha2_initAct2} \includegraphics[height=0.28\textheight]{../Simulations/initAct_/evolDistBeta2_initAct2} \includegraphics[height=0.28\textheight]{../Simulations/initAct_/evolDistBadge2_initAct2} \end{flushleft}
-
-\end{block}
-
-\end{frame}
-
-\begin{frame}{Why would you want to be recognized?}
-\protect\hypertarget{why-would-you-want-to-be-recognized-4}{}
-
-\begin{block}{The peaceful}
-
-Favours variation and cheap signals
-
-\begin{flushleft}\includegraphics[width=1\linewidth]{../Simulations/initAct_/corrAlphBet_initAct2} \end{flushleft}
-
-\end{block}
-
-\end{frame}
-
-\begin{frame}{Why would you want to be recognized?}
-\protect\hypertarget{why-would-you-want-to-be-recognized-5}{}
-
-\begin{block}{The aggressive}
-
-Variation among replicates
-
-\begin{flushleft}\includegraphics[width=1\linewidth]{../Simulations/initAct_/evolDynALL_initAct-2} \end{flushleft}
-
-\end{block}
-
-\end{frame}
-
-\begin{frame}{Why would you want to be recognized?}
-\protect\hypertarget{why-would-you-want-to-be-recognized-6}{}
-
-\begin{block}{The aggressive}
-
-Variation within one replicate
-
-\begin{flushleft}\includegraphics[width=1\linewidth]{../Simulations/initAct_/evolDyn0_initAct-2} \end{flushleft}
-
-\end{block}
-
-\end{frame}
-
-\begin{frame}{Why would you want to be recognized?}
-\protect\hypertarget{why-would-you-want-to-be-recognized-7}{}
-
-\begin{block}{The aggressive}
-
-Disfavours variation and signal evolution - no signalling
-
-\begin{flushleft}\includegraphics[width=1\linewidth]{../Simulations/initAct_/corrAlphBet_initAct-2} \end{flushleft}
-
-\end{block}
-
-\end{frame}
-
-\begin{frame}{Why would you want to be recognized?}
-\protect\hypertarget{why-would-you-want-to-be-recognized-8}{}
-
-\begin{block}{The clever (ESS)}
-
-Variation among replicates
-
-\begin{flushleft}\includegraphics[width=1\linewidth]{../Simulations/initAct_/evolDynALL_initAct-0.69} \end{flushleft}
-
-\end{block}
-
-\end{frame}
-
-\begin{frame}{Why would you want to be recognized?}
-\protect\hypertarget{why-would-you-want-to-be-recognized-9}{}
-
-\begin{block}{The clever (ESS)}
-
-Variation within one replicate
-
-\begin{flushleft}\includegraphics[width=1\linewidth]{../Simulations/initAct_/evolDyn0_initAct-0.69} \end{flushleft}
-
-\end{block}
-
-\end{frame}
-
-\begin{frame}{Why would you want to be recognized?}
-\protect\hypertarget{why-would-you-want-to-be-recognized-10}{}
-
-\begin{block}{The clever (ESS)}
-
-Signal evolution seems to be driven by drift. Most replicates end up in
-one of the two extremes
-
-\begin{flushleft}\includegraphics[width=1\linewidth]{../Simulations/initAct_/corrAlphBet_initAct-0.69} \end{flushleft}
+\includegraphics[width=0.9\linewidth]{../Simulations/initAct_/indVarScatter_}
 
@@
   Learning can mediate costly signals
-\item
-  More expensive signals seem be more likely to evolve
-\item
-  Less expensive signals provide more information
-\item
-  Local interactions do not impede the evolution of costly signals (as
-  has been suggested)
 \end{itemize}
@@
 \item
-  Under local interactions learning seems to allow the evolution of IR
-\item
   Cheap signals depend on the innate level of aggression (before
   learning) of individuals
-\item
-  Peace begets heterogeneity, aggression begets homogeneity
 \end{itemize}
@@
 \hypertarget{refs}{}
-\leavevmode\hypertarget{ref-botero_Evolution_2010}{}%
-Botero, C. A, I. Pen, J. Komdeur, and F. J Weissing. 2010. ``The
-Evolution of Individual Variation in Communication Strategies.''
-\emph{Evolution} 64 (11): 3123--33.
-
 \leavevmode\hypertarget{ref-grafen_Biological_1990}{}%

### 2020-11-25 AccFunc.R

@@ -5,6 +5,7 @@ source(here("aesth.R"))
 library("jsonlite")
 library('rmarkdown')
 require("NbClust")
+source(here("..","R_files","NbClust.R"))
 source(here("..","R_files","posPlots.R"))
 source(here("..","R_files","Filled.contour3.R"))
 # library('plotrix')
@@ -179,7 +180,9 @@ Critic<-function(weights,centers,sigSq=0.01,nx=1000){
 
 evolDist<-function(indData,variable,nbins,range=NULL,pal, nlevels =10, 
                    numx=1,numy=1,idx=1,idy=1,NewP=FALSE,
-                   cexAxis=2,xlab="x",ylab=variable, keyTitle="", ...){
+                   ylim=range(indData[,get(variable)]),
+                   cexAxis=2,xlab="x",ylab=variable, keyTitle="", xaxt="s",
+                   yaxt="s",...){
   if(is.null(range)){
     maxV<-indData[,max(get(variable))];minV<-indData[,min(get(variable))]
   }
@@ -213,16 +216,20 @@ evolDist<-function(indData,variable,nbins,range=NULL,pal, nlevels =10,
   #           log(0.01+t(counts))~time*get(variable),col.regions =  pal)
   # # par(plt=posPlot(numplotx = numx,numploty = numy,idplotx = idx,idploty = idy),
   #                 new=NewP)
-  filled.contour(x=as.numeric(timeseq),y = as.numeric(bins),z = log(0.01+t(contour1)),
-                 color.palette = pal, nlevels = nlevels,plot.axes={
-                   axis(1,cex.axis=cexAxis)
-                   axis(2,cex.axis=cexAxis)
+  
+ 
+  filled.contour3(x=as.numeric(timeseq),y = as.numeric(bins),z = log(0.01+t(contour1)),
+                 color.palette = pal, nlevels = nlevels,ylim=ylim,
+                 plot.axes={
+                   axis(1,cex.axis=cexAxis,xaxt=xaxt)
+                   axis(2,cex.axis=cexAxis,yaxt=yaxt)
                  },
                  plot.title={
                    title(xlab=xlab,cex.lab=2)
                    mtext(ylab,2,cex=2,line=3,las=1)},
                  key.title = {par(cex.main=cexAxis);title(main=keyTitle)}
                  )
+  # return(list(x=timeseq,y=bins,z=contour1))
 }
 
 
[…]

### 2020-11-25 CartoonRBF.R

@@ -9,7 +9,7 @@ png(here("Images","cartoonRBF.png"),width=800,height = 800)
 nCenters<-6
 interv<-1/(nCenters-1)
 centers<-interv*seq(0,nCenters-1)
-sigSq<-0.01
+sigSq<-0.001
 weights<-rep(0,nCenters)
 #  Random weights
 weights<-runif(nCenters,min=-1,max=1)#RBF(centers,0.5,0.05)
@@ -25,7 +25,8 @@ weights<-runif(nCenters)#RBF(centers,0.5,0.05)
 
 par(plt=posPlot(numploty = 2,idploty = 1),new=T)
 plot(totRBF(rangx,centers,0.01,weights)~rangx,type='l',col=1,
-     xlab="Badge",ylab="Estimated value",lwd=3,cex.lab=1.5,cex.axis=1.3)
+     xlab="Badge",ylab="Estimated value",lwd=3,cex.lab=1.5,cex.axis=1.3,
+     ylim=range(c(weights,totRBF(rangx,centers,0.01,weights))))
 points(y=weights,x=centers,cex=3)
 # dev.off()
 

### 2020-11-25 Dyn_reac_learn.R

@@ -12,7 +12,7 @@ require("jsonlite")
 
 # Scenario to be plotted - corresponds to folders where simulations are stored
 
-scenario<-"nIntGroupEvol4"
+scenario<-"nCentersnInt8"
 
 extSimsDir<-#here("Simulations",paste0(scenario,"_"))
   paste0("e:/BadgeSims/",scenario,"_")
@@ -23,7 +23,7 @@ extSimsDir<-#here("Simulations",paste0(scenario,"_"))
 # when program was run with internal paralellization
 
 # Project folder
-# (listTest<-list.files(here("Simulations",paste0(scenario,"_"))))
+(listTest<-list.files(here("Simulations",paste0(scenario,"_"))))
 # External sims folder
 (listTest<-list.files(extSimsDir,full.names = TRUE))
 
@@ -37,22 +37,20 @@ param<-fromJSON(here("Simulations",paste0(scenario,"_"),paramName[1]))
 
 # Choose which parameter to plot
 
-val<-1
+val<-2
 
 
 # Load files -------------------------------------------------------------------
 # when program was run with external paralellization
 
-# evolList_runs<-grep(paste0(param$namParam,param$rangParam[val]),
-#                     evolList,value =TRUE)
-# indList_runs<-grep(paste0(param$namParam,param$rangParam[val]),
-#                     indList,value =TRUE)
 
 
 # loop to produce pdfs for parameter values
-numCores <- length(indList)
+numCores <- length(param$rangParam)
 registerDoParallel(numCores)
-foreach(val = 1:length(indList),.packages = c("data.table","here")) %dopar% {
+
+foreach(val = 1:2,#length(param$rangParam),
+        .packages = c("data.table","here")) %dopar% {
 source(here("AccFunc.R"))
 
   fileId<-val
@@ -61,28 +59,47 @@ source(here("AccFunc.R"))
 # evol<-fread(here("Simulations",paste0(scenario,"_"),evolList[fileId]))
 # pop<-fread(here("Simulations",paste0(scenario,"_"),indList[fileId]))
 
-# evol<-do.call(rbind,lapply(evolList_runs,function(x){
-#   fread(here("Simulations",paste0(scenario,"_"),x))
-# }))
[…]

### 2020-11-25 Dyn_reac_learn_noLearn.R

@@ -12,7 +12,7 @@ require("jsonlite")
 
 # Scenario to be plotted - corresponds to folders where simulations are stored
 
-scenario<-"betCostNoLearn"
+scenario<-"nIntGroupEvol6"
 SimDir<-"Simulations"
 
 extSimsDir<-#here("Simulations",paste0(scenario,"_"))
@@ -35,7 +35,7 @@ param<-fromJSON(here(SimDir,paste0(scenario,"_"),paramName[1]))
 numCores <- length(evolList)
 registerDoParallel(numCores)
 
-val<-2
+val<-1
 
 
 # loop to produce pdfs for parameter values
@@ -52,7 +52,7 @@ fileId<-val
 # pop<-fread(here(SimDir,paste0(scenario,"_"),indList[fileId]))
 # External sims folder
 
-evol<-fread(evolList[fileId])
+# evol<-fread(evolList[fileId])
 # pop<-fread(indList[fileId])
 
 Valpar<-gsub("[[:alpha:]]",gsub(".txt","",tail(strsplit(evolList[val],"_")[[1]],1)),

### 2020-11-25 Figure_dyn_AGA.R

@@ -12,7 +12,7 @@ require("jsonlite")
 
 # Scenario to be plotted - corresponds to folders where simulations are stored
 
-scenario<-"nIntGroupEvol4TestSeed"
+scenario<-"betCostEvol1"
 
 extSimsDir<-#here("Simulations",paste0(scenario,"_"))
   paste0("e:/BadgeSims/",scenario,"_")
@@ -27,46 +27,61 @@ extSimsDir<-#here("Simulations",paste0(scenario,"_"))
 (evolList<-grep("evolLearn",listTest,value=TRUE))
 (indList<-grep("indLearn",listTest,value=TRUE))
 paramName<-list.files(here("Simulations",paste0(scenario,"_")))
-paramName<-list.files(extSimsDir,full.names = TRUE)
+# paramName<-list.files(extSimsDir,full.names = TRUE)
 paramName<-grep(".json",paramName,value=TRUE)
-param<-#fromJSON(here("Simulations",paste0(scenario,"_"),paramName))
-  fromJSON(paramName[1])
+param<-fromJSON(here("Simulations",paste0(scenario,"_"),paramName[1]))
+  # fromJSON(paramName[1])
 
 
-val<-1
+val<-3
+fileId<-val
 
-# loop to produce pdfs for parameter values
-# numCores <- length(indList)
-# registerDoParallel(numCores)
-# foreach(val = 1:length(indList),.packages = c("data.table","here")) %dopar% {
-#   source(here("AccFunc.R"))
+
+evolList_runs<-grep(paste0(param$namParam,param$rangParam[val]),
+                    evolList,value =TRUE)
+indList_runs<-grep(paste0(param$namParam,param$rangParam[val]),
+                   indList,value =TRUE)
+
+
+evol<-do.call(rbind,lapply(evolList_runs,fread))
+pop<-do.call(rbind,lapply(indList_runs, fread))
   
-  fileId<-val
+
+  # fileId<-val
   
   # Project folder
   # evol<-fread(here("Simulations",paste0(scenario,"_"),evolList[fileId]))
   # pop<-fread(here("Simulations",paste0(scenario,"_"),indList[fileId]))
-  
+  # 
   # External sims folder
   evol<-fread(evolList[fileId])
   pop<-fread(indList[fileId])
+
+  # Valpar<-gsub("[[:alpha:]]",gsub(".txt","",tail(strsplit(indList[val],"_")[[1]],1)),
[…]

### 2020-11-25 aesth.R

@@ -34,4 +34,6 @@ colSeeds<-brewer.pal(10,name = "Paired")
 
 pal_dist<-colorRampPalette(brewer.pal(9,"Blues"))
 
+colReps<-c(brewer.pal(12,"Paired"),1:3)
+
 colorbreaksQual<-seq(0,1,length=100)

### 2020-11-25 compPar.R

@@ -8,7 +8,7 @@ source(here("AccFunc.R"))
 
 # Scenario to be plotted - corresponds to folders where simulations are stored
 
-scenario<-"betCostEvol1"
+scenario<-"initAct"
 
 extSimsDir<-#here("Simulations",paste0(scenario,"_"))
   paste0("e:/BadgeSims/",scenario,"_")
@@ -22,7 +22,9 @@ extSimsDir<-#here("Simulations",paste0(scenario,"_"))
 (sdList<-grep("evol",listTest,value=TRUE))
 (indList<-grep("ind",listTest,value=TRUE))
 
-inds<-do.call(rbind,lapply(indList,filesScenar,scenario,full.name=TRUE))
+fread(indList[2])
+
+inds<-rbindlist(lapply(indList,filesScenar,scenario,full.name=TRUE),fill = TRUE)
 
 inds[,diffActWeight:=abs(WeightAct_0-WeightAct_4)]
 
@@ -202,4 +204,59 @@ dev.off()
 par(plt=posPlot())
 hist(inds[time==gener&get(shortSce)==unique(get(shortSce))[1],Badge],xaxt="s")
 hist(inds[time==gener&get(shortSce)==unique(get(shortSce))[3],Badge],xaxt="s")
+
+## Individual variation for different parameters ------------------------------
+
+lastInt<-tail(inds[,unique(nInteract)],2)[1]
+
+popOneInd<-inds[nInteract==lastInt]
+
+png(here("Simulations",paste0(scenario,"_"),
+         paste0("indVarScatter","_1",".png")),
+    width = 1400,height = 1200)
+
+yaxtAll<-c("s","n","n")
+xlabAll<-c("",expression(beta[s]),"")
+ylabAll<-c(expression(alpha[s]),"","")
+titleAll<-c("Peaceful","Clever","Aggressive")
+plot.new()
+for(PAr in popOneInd[,sort(unique(get(shortSce)),decreasing = TRUE)]){
+  count<-match(PAr,popOneInd[,sort(unique(get(shortSce)),decreasing = TRUE)])
+  par(plt=posPlot(numploty = 2,idploty = 2,
+                  numplotx = length(popOneInd[,unique(get(shortSce))]),
+                  idplotx = count)+c(0,0,-0.05,-0.05),
+      las=1,new=TRUE,
[…]

### 2020-11-25 learnDynamics.R

@@ -11,7 +11,7 @@ source(here("AccFunc.R"))
 
 # Scenario to be plotted - corresponds to folders where simulations are stored
 
-scenario<-"nIntGroupEvol4"
+scenario<-"alphaAct"
 
 extSimsDir<-paste0("e:/BadgeSims/",scenario,"_")
 
@@ -25,12 +25,12 @@ extSimsDir<-paste0("e:/BadgeSims/",scenario,"_")
 (indList<-grep("ind",listTest,value=TRUE))
 
 # parameter values from project folder
-# paramName<-list.files(here("Simulations",paste0(scenario,"_")))
-paramName<-list.files(extSimsDir,full.names = TRUE)
+paramName<-list.files(here("Simulations",paste0(scenario,"_")))
+# paramName<-list.files(extSimsDir,full.names = TRUE)
 
 paramName<-grep(".json",paramName,value=TRUE)
-param<-#fromJSON(here("Simulations",paste0(scenario,"_"),paramName[2]))
-fromJSON(paramName)
+param<-fromJSON(here("Simulations",paste0(scenario,"_"),paramName[2]))
+# fromJSON(paramName[1])
 
 fileId<-1
 
@@ -67,14 +67,14 @@ colorbreaksQual<-seq(0,1,length=100)
 # Actor 
 
 finReps<-indLearn[time==max(time),unique(seed)]
-seedCh<-
+seedCh<-0
   finReps[round(runif(1,0,length(finReps)))+1]
 
 
 # Select run and generation to plot
 tempPop<-indLearn[time==gener&seed==seedCh]
 
-timePoints<-round(seq(1,length(unique(tempPop[,nInteract]))-1,
+timePoints<-round(seq(1,length(unique(tempPop[,nInteract]))-90,
                       length.out = 10))
 
 # png(here("Simulations",paste0(scenario,"_"),paste0(nampar,Valpar,"learnDyn.png")),

### 2020-11-25 parTOjson.R

@@ -9,8 +9,8 @@ source(here("AccFunc.R"))
 fileName<-"parameters"
 here()
 
-param1<-#fromJSON(paste0("E:/BadgeSims","/betCostNoLearn_","/parameters.json"))
-     fromJSON(here("Simulations","nIntGroupEvol4_","parameters.json"))
+# param1<-#fromJSON(paste0("E:/BadgeSims","/betCostNoLearn_","/parameters.json"))
+#      fromJSON(here("Simulations","alphaAct_","parameters1.json"))
 
 param<-list(totGen   = 15000,   nRep    = 1, seed = 1,
             printGen = 1000,   printLearn = 1000,
@@ -19,23 +19,26 @@ param<-list(totGen   = 15000,   nRep    = 1, seed = 1,
             MutSd    = 0.3,  nInt        = 500,  init     = c(0,0,1,0),
             mutRate  = 0.001,  mutType  = 0,
             sampleSize = 50,   strQual  = 10,
-            errorQual = 0,    alphaBad	 = 0,    betaBad	 = 0,
+            errorQual = 0,    alphaBad	 = I(c(0)),    
+            betaBad	 = I(c(0)),
             alphaRes =-10,betaRes=-20, gammaRes=0,
-            alphaCrit  = 0.5,  alphaAct = 0.5,
+            alphaCrit  = 0.4,  alphaAct = 0.4,
             sigSq   	 = 0.01, nCenters = 6,
             initCrit = 0,      initAct=0,   gamma = 0,
-            QualStDv   = 1.1, betCost = 0,
+            QualStDv   = 0.15, betCost = 0,
             alphCost	 = 3, mutLearn = FALSE,
-            nIntGroup  = 2000, 
+            nIntGroup  = 8, 
             payoff_matrix = c(1.5,1,0,0.5),
-            namParam = "nIntGroup",
-            rangParam = c(2000,8),
+            namParam = "nCenters",
+            rangParam = I(c(4,8,10)),
             typeAgent = 2, #//0. hawk 1.dove 2. learner 3. evaluator
             folderL=paste(here("Simulations"),"/",sep="")) # comment for debug
-# folderL=paste(here("Simulations","test_"),"/",sep="")) # comment for realease
+# folderL=paste(here("Simulations","test_"),"/",sep="")) # comment for release
 
 
-apendScenar<-"Evol5"
+
+
+apendScenar<-"nInt8"
[…]

### 2020-12-09 manuscript/manuscript_1.0.Rmd
New folder to contain presentation and manuscripts. Relevent files moved accordingly. Code needs to be updated to this change. New presentaion started for AGA symposium. New simulations varyinh alphaAct. Need to check why individuals of different quality are not asssorting in nIntGroupEvol4.
Starting the description of the model

--- 2020-10-01 manuscript_1.0.Rmd
@@
 linespread: 2
-bibliography: SocialMorphCue.bib
+bibliography: ../SocialMorphCue.bib
 spacing: double
@@
 where an arbitrary signal conveys quality [@johnstone_Badges_1993; @rohwer_Social_1975a]. 
-Here, arbitrary refers to a signal which is not ontogeneteically correlated with quality. 
+Here, arbitrary refers to a signal which is not ontogenetically correlated with quality. 
 Badges of status can be evolutionarily stable whenever the signal imposes a fitness cost which 
@@
 
+## The model
+
+We model the evolution of signals of quality in the context of agonistic interactions by 
+modelling a population of haploid individuals with non-overlaping generations. 
+Individuals are born every generation with a level of quality ($Q_i$, where $i$ is subscript 
+of the population vector of size $N$) given by a number between zero and one, which is drawn 
+from a truncated normal distribution $N(\mu = 0.5,\sigma)$. 
+An Individual with quality $0$ has the lowest RHP, 
+while an individual with quality $1$ has the highest. As part of development processes
+individuals produce a morphological signal (badge), the size of such depends on their level 
+of quality according to a reaction norm given by $B_i=1/(1+e^{\alpha_i-\beta_i Q_i})$;
+where are $\alpha_i$ and $\beta_i$ are individual specific traits that determine 
+the shape of the reaction norm. Variation in $\alpha_i$ and $beta_i$ means individuals
+can have either uninformative (flat) or informative (logistic) norms; but the size of 
+the signal is constrained between zero and one. We assume different values of these traits 
+are given by different alleles, and are inherited from mother to offspring. 
+Unless, with a small probability, mutation changes the allelic value of the offspring.
+After birth individuals go through a round of viavility selection. 
+Individual specific survival probability is given by $s_i=1/(1+e^{-k_1-k_2(Q_i-B_i)}$. 
+Where $k_1$ and $k_2$ are parameter values determining the shape of the survival function. 
+Importantly, if $k_2=0$ survival probability is independent of quality, while if $k_2>0$
+lower quality individuals pay a higher price for s similar-sized badge, fulfilling the core assumption 
+of the handicap principle [@botero_Evolution_2010; @grafen_Biological_1990; @johnstone_Badges_1993].
+
+Individuals who survive engage in a series of pairwise interactions 
+where they compete for resources. Individuals in real populations typically do not 
+interact at random nor with all individuals in the population [@kurvers_Evolutionary_2014]. 
+In order to allow for non-random interactions, we use the population vector to define 
+an interaction neighborhood for the focal individual, the size of the neighborhood (g) 
+is the number of positions around the focal´s from which the interactant is drawn. 
+So, if $g=N$ interactions are global; as $g$ gets smaller interactions are more local. 
+In an interaction each individual must decide whether to escalate a fight or not. 
+Following the classic hawk-dove game [@maynard-smith_Evolution_1982], if the focal individual fights and its partner does not, the focal gets as payoff the resource of value $V$ while its partner gets nothing; if both individuals restrain  from fighting they split up the resource in half; if both decide to fight they splitup the cost of fighting and the winner takes over the resource.
+We further assume that the probability of wining a fight for the focal individual depends on the difference in quality between her and her interacting partner; specifically it is given by $p_{ij}=1/(1+e^{-k_3(Q_i-Q_j)})$, where $k_3$ is a parameter defining
+how strong the quality difference determines the winning probability; 
+and $i$ and $j$ denote the the position in the population 
+vector of the focal and its interacting partner respectively.
+
+The decision of whether to escalate a fight against an interacting partner can be dependent 
+on the size of the partner´s badge, the actual dependency is determined by 
+the focal´s experiences through a learning process. We implement learning using 
+the actor-critic aproach from reinforcement learning theory [RL; @sutton_Reinforcement_2018;@quinones_Reinforcement_2019; @leimar_Learning_2019].
+Individuals estimate the reward (payoff) expected from  interacting with partners 
+of different badge sizes (the critic in RL terminology). 
+After every interaction they update the estimate of reward proportionally to the difference between their current estimate and the observed reward (prediction error $\delta$).
+Furthermore, they express different probabilities of escalatining (retreating) depending on
+the badge size of their opponent (the actor in RL). They update the probability of 
+retreating (escalating)  up or down depending on whether retreating (escalating) leads to an increase 
+in the reward estimation. So, if a focal individual decides to escalate a fight against 
+an individual with small badge, and that leads to an increase in the reward estimation,
+the focal individual will increase the probability of escalating fights with 
+individuals of small badges in the future. Given that badge size is a real number between 0 and 1,
+the reward estimation, as well as the probability of retreating (escalating), must be
+generalized across different values. To implement generalization we use the function
+approximation method based on radial basis functions [@sutton_Reinforcement_2018]. Specifically, we pick $c$ evenly spaced 
+values along the badge size interval,
+
 ## References

### 2020-12-09 CartoonRBF.R

@@ -6,10 +6,10 @@ source("../R_files/posPlots.R")
 source(here("AccFunc.R"))
 
 png(here("Images","cartoonRBF.png"),width=800,height = 800)
-nCenters<-6
+nCenters<-10
 interv<-1/(nCenters-1)
 centers<-interv*seq(0,nCenters-1)
-sigSq<-0.001
+sigSq<-0.005
 weights<-rep(0,nCenters)
 #  Random weights
 weights<-runif(nCenters,min=-1,max=1)#RBF(centers,0.5,0.05)

### 2020-12-09 Dyn_reac_learn.R

@@ -12,7 +12,7 @@ require("jsonlite")
 
 # Scenario to be plotted - corresponds to folders where simulations are stored
 
-scenario<-"nCentersnInt8"
+scenario<-"initAct"
 
 extSimsDir<-#here("Simulations",paste0(scenario,"_"))
   paste0("e:/BadgeSims/",scenario,"_")
@@ -23,7 +23,7 @@ extSimsDir<-#here("Simulations",paste0(scenario,"_"))
 # when program was run with internal paralellization
 
 # Project folder
-(listTest<-list.files(here("Simulations",paste0(scenario,"_"))))
+# (listTest<-list.files(here("Simulations",paste0(scenario,"_"))))
 # External sims folder
 (listTest<-list.files(extSimsDir,full.names = TRUE))
 
@@ -32,7 +32,7 @@ extSimsDir<-#here("Simulations",paste0(scenario,"_"))
 paramName<-list.files(here("Simulations",paste0(scenario,"_")))
 # paramName<-list.files(extSimsDir,full.names = TRUE)
 paramName<-grep(".json",paramName,value=TRUE)
-param<-fromJSON(here("Simulations",paste0(scenario,"_"),paramName[1]))
+param<-fromJSON(here("Simulations",paste0(scenario,"_"),paramName))
 # fromJSON(paramName)
 
 # Choose which parameter to plot
@@ -49,7 +49,7 @@ val<-2
 numCores <- length(param$rangParam)
 registerDoParallel(numCores)
 
-foreach(val = 1:2,#length(param$rangParam),
+foreach(val = 1:3,#length(param$rangParam),
         .packages = c("data.table","here")) %dopar% {
 source(here("AccFunc.R"))
 
@@ -64,13 +64,13 @@ source(here("AccFunc.R"))
 
   
 # External sims folder
-# evol<-fread(evolList[fileId])
-# pop<-fread(indList[fileId])
+evol<-fread(evolList[fileId])
+pop<-fread(indList[fileId])
 
-  evolList_runs<-grep(paste0(param$namParam,param$rangParam[val]),
-                      evolList,value =TRUE)
-  indList_runs<-grep(paste0(param$namParam,param$rangParam[val]),
-                     indList,value =TRUE)
+  # evolList_runs<-grep(paste0(param$namParam,param$rangParam[val]),
+  #                     evolList,value =TRUE)
[…]

### 2020-12-09 cartoons_reactNorms.R

@@ -47,11 +47,11 @@ persp(x = qualRange,y=rivalBadge,z=perspTmp,phi = 35, theta=40 ,
 # png(here("reacNorm.png"),width = 800,height = 800)
 
 badQualRang<-seq(0,1,length=1000)
-alpha<-1
-beta<-6
+alpha<-2
+beta<-0
 costFunc<-function(beta,alpha,x) {d<-1/(1+exp(-alpha-beta*x))}
 par(plt=posPlot())
-betRang<-c(20)
+betRang<-c(0)
 matplot(y=sapply(betRang,FUN = costFunc,alpha=alpha,x=badQualRang),x=badQualRang,
         type="l",cex.lab=1.5, cex.axis=2,lwd=3,lty=1,las=1,
      col=c(2,3),ylim=c(0.5,1.1),xaxt="s",yaxt="s",ylab="",xlab="") #,col=i

### 2020-12-09 parTOjson.R

@@ -9,10 +9,10 @@ source(here("AccFunc.R"))
 fileName<-"parameters"
 here()
 
-# param1<-#fromJSON(paste0("E:/BadgeSims","/betCostNoLearn_","/parameters.json"))
-#      fromJSON(here("Simulations","alphaAct_","parameters1.json"))
+param1<-#fromJSON(paste0("E:/BadgeSims","/betCostNoLearn_","/parameters.json"))
+     fromJSON(here("Simulations","initAct_","parameters2.json"))
 
-param<-list(totGen   = 15000,   nRep    = 1, seed = 1,
+param<-list(totGen   = 20000,   nRep    = 1, seed = 1,
             printGen = 1000,   printLearn = 1000,
             printLearnInt = 50,
             popSize  = 2000,  baselineFit = 2,   
@@ -23,14 +23,14 @@ param<-list(totGen   = 15000,   nRep    = 1, seed = 1,
             betaBad	 = I(c(0)),
             alphaRes =-10,betaRes=-20, gammaRes=0,
             alphaCrit  = 0.4,  alphaAct = 0.4,
-            sigSq   	 = 0.01, nCenters = 6,
+            sigSq   	 = 0.005, nCenters = 10,
             initCrit = 0,      initAct=0,   gamma = 0,
             QualStDv   = 0.15, betCost = 0,
             alphCost	 = 3, mutLearn = FALSE,
             nIntGroup  = 8, 
             payoff_matrix = c(1.5,1,0,0.5),
-            namParam = "nCenters",
-            rangParam = I(c(4,8,10)),
+            namParam = "initAct",
+            rangParam = I(c(-0.69,2,-2)),
             typeAgent = 2, #//0. hawk 1.dove 2. learner 3. evaluator
             folderL=paste(here("Simulations"),"/",sep="")) # comment for debug
 # folderL=paste(here("Simulations","test_"),"/",sep="")) # comment for release
@@ -38,7 +38,7 @@ param<-list(totGen   = 15000,   nRep    = 1, seed = 1,
 
 
 
-apendScenar<-"nInt8"
+apendScenar<-"nGroup8"
 param$folderL<-paste0(param$folderL,
        param$namParam,apendScenar,"_/")
 # runTime<-"360:00:00"# "10:00:00"# 

### 2020-12-19 manuscript/manuscript_1.0.Rmd
First draft of the methods done. Add a parameter, sharecost, determining which proportion of the cost is beared by the winner.

@@
 knitr::opts_chunk$set(echo = TRUE)
+require(png)
+require(here)
+source(here("../R_files/posPlots.r"))
+source(here("AccFunc.r"))
 ```
@@
 from a truncated normal distribution $N(\mu = 0.5,\sigma)$. 
-An Individual with quality $0$ has the lowest RHP, 
+An individual with quality $0$ has the lowest RHP, 
 while an individual with quality $1$ has the highest. As part of development processes
@@
 where are $\alpha_i$ and $\beta_i$ are individual specific traits that determine 
-the shape of the reaction norm. Variation in $\alpha_i$ and $beta_i$ means individuals
+the shape of the reaction norm. Variation in $\alpha_i$ and $\beta_i$ means individuals
 can have either uninformative (flat) or informative (logistic) norms; but the size of 
@@
 the reward estimation, as well as the probability of retreating (escalating), must be
-generalized across different values. To implement generalization we use the function
-approximation method based on radial basis functions [@sutton_Reinforcement_2018]. Specifically, we pick $c$ evenly spaced 
-values along the badge size interval,
+generalized across different values. To implement generalization we use the linear function
+approximation method based on radial basis functions [@sutton_Reinforcement_2018].
+Specifically, we pick $f$ feature centers, which are evenly spaced values along the badge size interval ([0,1]).
+Each one of these features centers is associated with a weight for reward estimation and
+tendency to play retreat in a given interaction. The reward estimation and the tendency 
+to retreat are calculated (in every interaction) as the sum of the weights associated 
+with each feature, weighted by the response triggered by the feature. 
+The response of each feature center dimishes as a gaussian function  with the distance between the feature center and the badge size of the partner. Formally the reward estimation 
+$\hat{R}$ when the focal individual ($i$) faces individual $j$ is given by,
+
+\begin{equation}
+\hat{R}_{ij} =  \sum_{z=1}^{c} x_z e^{(-\frac{|B_i-c_z|}{2\theta^2})} 
+\end{equation}
+
+where $x_z$ is the weight of feature center $z$ on the reward estimation; and $\theta$ is
+the width of the generalization function. Similarly, the tendency to retreat is given by sum 
+of feature weigths, and the probability of associated with the tendency is obtained
+by applying a logistic transformation. So formally, the log-odds to retreat
+when facing individual $j$ is
+
+\begin{equation}
+logit(q_{ij}) =  \sum_{z=1}^{c} y_z e^{(-\frac{|B_i-c_z|}{2\theta^2})} 
+\end{equation}
+
+where $y_z$ is the weight of feature center $z$ on the tendency to retreat.
+
+After all their interactions, individuals in the population reproduce proportionally to 
+their fitness $w_i$, which is a sum of the baseline fitness ($w_0$) and all the payoffs
+obtained throughout their life. Thus, the combination of natural selection and genetic drift
+changes the distribution of values in $\alpha$ and $\beta$ that segregates in the population, effectively the badges expressed and changing the communication system.
+
+```{r Figure 1,echo=FALSE,fig.width=8,fig.align='center',fig.height=4,fig.show='hold',fig.cap= "Model of communication in the context of aggressive interactions.
+In A the sender code: a reaction norm determines how the badge size is determined by the quality of the individual.
+Red shows an honest reaction norm, while the blue shows a dishonest reaction norm.
+B the receiver code: individuals have a behavioral reaction norm that determines their probability of retreating (black line).
+The reaction norm arises from generalizing the information from the feature weights (black dots).
+Generalization is represented by the grey line and its axis, which shows the response triggered by the fourth feature diminishes as the value evaluated is further from the feature center.
+The learning process moves the feature weights (black dots) up or down depending on whether that leads to an increase in the estimated reward."}
+honestBig<-readPNG(here("Images/cart_honestBig.png"))
+honestSmall<-readPNG(here("Images/cart_honestSmall.png"))
+dishonest<-readPNG(here("Images/cart_dishonest.png"))
+par(plt=posPlot(numplotx = 2,idplotx = 1,upboundx = 91),las=1)
+  rangPars<-data.frame(honest=c(4,2),dishonest=c(0,-1.8))
+ rangQual<-seq(0,1,by = 0.01)
+ linesSender<-sapply(rangPars,function(x)logist(rangQual,alpha = x[2],beta = x[1]))
+ plot(0,0,type="l",xlab="",xlim=c(0,1),
+      ylab ="",cex.lab=1,lwd=2,col="grey",cex.axis=1,ylim=c(0,1.1))
+ text(x = 0.08,y=1.05,"A",cex = 1.5)
+ mtext("Quality",1,line = 2,cex = 1,las = 0)
+ matlines(x=rangQual,y=linesSender,lwd=4,col=colRuns,lty=1)
+ mtext("Badge size",2,line = 2.3,cex = 1,las = 0)
+ rasterImage(honestBig,0.75,0.6,0.95,0.8)
+ rasterImage(honestSmall,0.05,0.01,0.25,0.21)
+ rasterImage(dishonest,0.75,0.87,0.95,1.07)
+ rasterImage(dishonest,0.05,0.87,0.25,1.07)
+ # legend("left",legend=rangeBeta,col=colRuns,lwd=3,title = expression(beta),cex=2,bty="n")
+
+ par(plt=posPlot(numplotx = 2,idplotx = 2,upboundx = 91),las=1,new=TRUE)
+
+ nCenters<-6
+interv<-1/(nCenters-1)
+centers<-interv*seq(0,nCenters-1)
+sigSq<-0.01
+weights<-rep(0,nCenters)
+weights<-c(-5,-2,-1,1,2,5)#runif(nCenters,min=-1,max=1)
+rangx<-seq(0,1,length=1000)
+
+plot(logist(totRBF(rangx,centers,sigSq,weights),alpha = 0,beta=1)~rangx,type='l',col=1,
+     xlab="",ylab="",ylim=c(0,1.1),lwd=3,xaxt="s",cex.lab=2,cex.axis=1,yaxt='n')
+axis(4,cex=1)
+text(x = 0.08,y=1.05,"B",cex = 1.5)
+# test<-sapply(1:6, function(x){RBF()})
+points(y=logist(weights,alpha = 0,beta=1),x=centers,cex=1.5,pch=20)
+respon<-logist(weights[4]*RBF(rangx,centers[4],0.01))
+lines(x=rangx,y=respon,col="grey",lwd=2)
+axis(side = 4,at = seq(range(respon)[1],range(respon)[2],length.out = 4),
+     labels = round(seq(0,1,length.out = 4),2),line = -4,col="grey")
+
+# xRBFs<-sapply(centers, function(x){seq(x-0.2,x+0.2,by=0.001)})
+# RBFS<-sapply(1:6, function(x){
+#   logist(weights[x]*RBF(seq(centers[x]-0.2,centers[x]+0.2,by=0.001),xCenter = centers[x],
+#                         sigSq = 0.01))
+# })
+# matlines(x=xRBFs,y=RBFS,col="grey",lty=2,lwd=2)
+
+mtext("prob. retreat",4,line = 2.3,cex = 1,las = 0)
+mtext("Badge size",1,line = 2,cex = 1,las = 0)
+
+```
+
+## Results
 

### 2020-12-19 Dyn_reac_learn.R

@@ -12,7 +12,7 @@ require("jsonlite")
 
 # Scenario to be plotted - corresponds to folders where simulations are stored
 
-scenario<-"initAct"
+scenario<-"test"
 
 extSimsDir<-#here("Simulations",paste0(scenario,"_"))
   paste0("e:/BadgeSims/",scenario,"_")
@@ -23,7 +23,7 @@ extSimsDir<-#here("Simulations",paste0(scenario,"_"))
 # when program was run with internal paralellization
 
 # Project folder
-# (listTest<-list.files(here("Simulations",paste0(scenario,"_"))))
+(listTest<-list.files(here("Simulations",paste0(scenario,"_"))))
 # External sims folder
 (listTest<-list.files(extSimsDir,full.names = TRUE))
 
@@ -32,12 +32,12 @@ extSimsDir<-#here("Simulations",paste0(scenario,"_"))
 paramName<-list.files(here("Simulations",paste0(scenario,"_")))
 # paramName<-list.files(extSimsDir,full.names = TRUE)
 paramName<-grep(".json",paramName,value=TRUE)
-param<-fromJSON(here("Simulations",paste0(scenario,"_"),paramName))
+param<-fromJSON(here("Simulations",paste0(scenario,"_"),paramName[1]))
 # fromJSON(paramName)
 
 # Choose which parameter to plot
 
-val<-2
+val<-1
 
 
 # Load files -------------------------------------------------------------------
@@ -64,20 +64,20 @@ source(here("AccFunc.R"))
 
   
 # External sims folder
-evol<-fread(evolList[fileId])
-pop<-fread(indList[fileId])
+# evol<-fread(evolList[fileId])
+# pop<-fread(indList[fileId])
 
-  # evolList_runs<-grep(paste0(param$namParam,param$rangParam[val]),
-  #                     evolList,value =TRUE)
-  # indList_runs<-grep(paste0(param$namParam,param$rangParam[val]),
-  #                    indList,value =TRUE)
+  evolList_runs<-grep(paste0(param$namParam,param$rangParam[val]),
+                      evolList,value =TRUE)
+  indList_runs<-grep(paste0(param$namParam,param$rangParam[val]),
+                     indList,value =TRUE)
   
-  # evol<-do.call(rbind,lapply(evolList_runs,function(x){
-  #   fread(here("Simulations",paste0(scenario,"_"),x))
-  # }))
[…]

### 2020-12-19 Dyn_reac_learn_noLearn.R

@@ -12,7 +12,7 @@ require("jsonlite")
 
 # Scenario to be plotted - corresponds to folders where simulations are stored
 
-scenario<-"nIntGroupEvol6"
+scenario<-"betCostNoLearnShare0"
 SimDir<-"Simulations"
 
 extSimsDir<-#here("Simulations",paste0(scenario,"_"))
@@ -29,7 +29,7 @@ extSimsDir<-#here("Simulations",paste0(scenario,"_"))
 
 paramName<-list.files(here(SimDir,paste0(scenario,"_")))
 paramName<-grep(".json",paramName,value=TRUE)
-param<-fromJSON(here(SimDir,paste0(scenario,"_"),paramName[1]))
+param<-fromJSON(here(SimDir,paste0(scenario,"_"),paramName[9]))
 
 
 numCores <- length(evolList)
@@ -39,7 +39,7 @@ val<-1
 
 
 # loop to produce pdfs for parameter values
-foreach(val = 1:length(evolList),.packages = c("data.table","here")) %dopar% {
+foreach(val = 1:2,.packages = c("data.table","here")) %dopar% {
   source(here("AccFunc.R"))
   
   
@@ -55,10 +55,20 @@ fileId<-val
 # evol<-fread(evolList[fileId])
 # pop<-fread(indList[fileId])
 
-Valpar<-gsub("[[:alpha:]]",gsub(".txt","",tail(strsplit(evolList[val],"_")[[1]],1)),
-             replacement = "")
-nampar<-gsub("[^[:alpha:]]",gsub(".txt","",tail(strsplit(evolList[val],"_")[[1]],1)),
-             replacement = "")
+
+(Valpar<-param$rangParam[val])
+
+(nampar<-param$namParam)
+
+evolList_runs<-grep(paste0(param$namParam,param$rangParam[val]),
+                    evolList,value =TRUE)
+
+evol<-do.call(rbind,lapply(evolList_runs,fread))
+
+# evol<-do.call(rbind,lapply(evolList_runs,function(x){
+#   fread(here("Simulations",paste0(scenario,"_"),x))
+# }))
+
 
 # pdf(paste0(extSimsDir,"/evolDyn_",nampar,Valpar,".pdf"))
 pdf(here(SimDir,paste0(scenario,"_"),paste0("evolDyn_",nampar,Valpar,".pdf")))
@@ -249,7 +259,7 @@ traitsTrajs<-dcast(evol,time~seed,
 
[…]

### 2020-12-19 cue_compt.cpp

@@ -185,7 +185,8 @@ class individual {
 			return(nCenters);
 		}
 		int set_phenotype(individual partner, printingObj &localPrint);
-		void get_payoff(individual partner, vector<double> param, bool win);
+		void get_payoff(individual partner, vector<double> param, bool win,
+			float shareCost);
 		void set_Badge(double stdDev, double errorQual);
 		bool Winfight(double otherQuality, double strQual);
 		strategy mutateStr(strategy genotype,double mutRate,int mutType);
@@ -411,7 +412,7 @@ double  individual::mutateDoub(double value, double mutRate, double mutSD) {
 
 
 void individual::get_payoff(individual partner,vector<double> payoff_matrix, 
-	bool win) {      
+	bool win, float shareCost) {      
 	/*  (This are IDs to the entries of the matrix, NOT payoffs)    
 	        Hawk      Dove     
 	   Hawk   0         1
@@ -419,12 +420,12 @@ void individual::get_payoff(individual partner,vector<double> payoff_matrix,
 	*/
 	int scenario = 2 * phenotype + partner.phenotype;
 	if (scenario == 0 && win) {
-		curr_payoff = payoff_matrix[1] - payoff_matrix[0] * 0.5;
-		cum_payoff += payoff_matrix[1] - payoff_matrix[0] * 0.5;
+		curr_payoff = payoff_matrix[1] - payoff_matrix[0] * shareCost;
+		cum_payoff += payoff_matrix[1] - payoff_matrix[0] * shareCost;
 	}
 	else if (scenario == 0) {
-		curr_payoff = - payoff_matrix[0] * 0.5;
-		cum_payoff += - payoff_matrix[0] * 0.5;
+		curr_payoff = - payoff_matrix[0] * (1-shareCost);
+		cum_payoff += - payoff_matrix[0] * (1 - shareCost);
 	}
 	else {
 		curr_payoff = payoff_matrix[scenario];
@@ -448,7 +449,12 @@ bool individual::viability(double alphaCost,double betaCost) {
 int individual::set_phenotype(individual partner, printingObj& localPrint) {
 	if (get_strat() == evaluator || get_strat() == learner) {
 		calcRespValPref(partner);
-		phenotype = static_cast<strategy>(rnd::bernoulli(logist(preferenceT, 0)));
+		if (get_strat() == evaluator)
+			phenotype = static_cast<strategy>(rnd::bernoulli(
+			logist(preferenceT, 0,-1)));
+		else
[…]

### 2020-12-19 parTOjson.R

@@ -9,14 +9,14 @@ source(here("AccFunc.R"))
 fileName<-"parameters"
 here()
 
-param1<-#fromJSON(paste0("E:/BadgeSims","/betCostNoLearn_","/parameters.json"))
-     fromJSON(here("Simulations","initAct_","parameters2.json"))
+(param1<-#fromJSON(paste0("E:/BadgeSims","/betCostNoLearn_","/parameters.json"))
+     fromJSON(here("Simulations","betCostNoLearnShare0_","parameters0.json")))
 
-param<-list(totGen   = 20000,   nRep    = 1, seed = 1,
-            printGen = 1000,   printLearn = 1000,
+param<-list(totGen   = 2000,   nRep    = 1, seed = 1,
+            printGen = 100,   printLearn = 100,
             printLearnInt = 50,
             popSize  = 2000,  baselineFit = 2,   
-            MutSd    = 0.3,  nInt        = 500,  init     = c(0,0,1,0),
+            MutSd    = 2,  nInt        = 50,  init     = c(0,0,0,1),
             mutRate  = 0.001,  mutType  = 0,
             sampleSize = 50,   strQual  = 10,
             errorQual = 0,    alphaBad	 = I(c(0)),    
@@ -25,20 +25,21 @@ param<-list(totGen   = 20000,   nRep    = 1, seed = 1,
             alphaCrit  = 0.4,  alphaAct = 0.4,
             sigSq   	 = 0.005, nCenters = 10,
             initCrit = 0,      initAct=0,   gamma = 0,
-            QualStDv   = 0.15, betCost = 0,
+            QualStDv   = 0.15, betCost = 6,
             alphCost	 = 3, mutLearn = FALSE,
-            nIntGroup  = 8, 
-            payoff_matrix = c(1.5,1,0,0.5),
-            namParam = "initAct",
-            rangParam = I(c(-0.69,2,-2)),
-            typeAgent = 2, #//0. hawk 1.dove 2. learner 3. evaluator
+            nIntGroup  = 2000, 
+            payoff_matrix = c(3,1,0,0.5),
+            shareCost = 0,
+            namParam = "betCost",
+            rangParam = I(c(0,6)),
+            typeAgent = 3, #//0. hawk 1.dove 2. learner 3. evaluator
[…]

### 2020-12-22 manuscript/manuscript_1.0.Rmd
Small changes on the manuscript.

@@
 determinant of fitness in social animals. Irrespective of whether such interactions are cooperative 
-or competitive, the actions of interacting individual partly determine 
+or competitive, the actions of interacting individuals partly determine 
 their reproductive success. However, the best action for an individual 
@@
 [@johnstone_Recognition_1997], or to provide erroneous information [@johnstone_Badges_1993]. 
-Typically, in a given context, some proportion of individuals will beenefit 
+Typically, in a given context, some proportion of individuals will benefit 
 from broadcasting accurate information, whereas others will benefit from hiding it. 
@@
 where are $\alpha_i$ and $\beta_i$ are individual specific traits that determine 
-the shape of the reaction norm. Variation in $\alpha_i$ and $\beta_i$ means individuals
-can have either uninformative (flat) or informative (logistic) norms; but the size of 
-the signal is constrained between zero and one. We assume different values of these traits 
+the shape of the reaction norm (Fig. \ref{fig:model-struc} A). Variation in $\alpha_i$ and $\beta_i$ means individuals can have either uninformative (flat) or informative (logistic) norms. 
+The size of the signal is constrained between zero and one. We assume different values of these traits 
 are given by different alleles, and are inherited from mother to offspring. 
@@
 After birth individuals go through a round of viavility selection. 
-Individual specific survival probability is given by $s_i=1/(1+e^{-k_1-k_2(Q_i-B_i)}$. 
+Individual specific survival probability is given by $s_i=1/(1+e^{-k_1-k_2(Q_i-B_i)})$. 
 Where $k_1$ and $k_2$ are parameter values determining the shape of the survival function. 
@@
 an interaction neighborhood for the focal individual, the size of the neighborhood (g) 
-is the number of positions around the focal´s from which the interactant is drawn. 
+is the number of positions around the focal's from which the interactant is drawn. 
 So, if $g=N$ interactions are global; as $g$ gets smaller interactions are more local. 
@@
 The decision of whether to escalate a fight against an interacting partner can be dependent 
-on the size of the partner´s badge, the actual dependency is determined by 
+on the size of the partner's badge, the actual dependency is determined by 
 the focal´s experiences through a learning process. We implement learning using 
@@
 After every interaction they update the estimate of reward proportionally to the difference between their current estimate and the observed reward (prediction error $\delta$).
-Furthermore, they express different probabilities of escalatining (retreating) depending on
+Furthermore, they express different probabilities of retreating (attacking) depending on
 the badge size of their opponent (the actor in RL). They update the probability of 
-retreating (escalating)  up or down depending on whether retreating (escalating) leads to an increase 
+retreating (attacking)  up or down depending on whether retreating (attacking) leads to an increase 
 in the reward estimation. So, if a focal individual decides to escalate a fight against 
@@
 the focal individual will increase the probability of escalating fights with 
-individuals of small badges in the future. Given that badge size is a real number between 0 and 1,
-the reward estimation, as well as the probability of retreating (escalating), must be
-generalized across different values. To implement generalization we use the linear function
+individuals of small badges in the future.
+Given that badge size is a real number between 0 and 1, there are infinitely many badge sizes.
+Thus, the reward estimation, as well as the probability of retreating (escalating), must be generalized across different values.
+To implement generalization we use the linear function
 approximation method based on radial basis functions [@sutton_Reinforcement_2018].
-Specifically, we pick $f$ feature centers, which are evenly spaced values along the badge size interval ([0,1]).
+Specifically, we pick $c$ feature centers, which are evenly spaced values along the badge size interval ([0,1]).
 Each one of these features centers is associated with a weight for reward estimation and
@@
 to retreat are calculated (in every interaction) as the sum of the weights associated 
-with each feature, weighted by the response triggered by the feature. 
-The response of each feature center dimishes as a gaussian function  with the distance between the feature center and the badge size of the partner. Formally the reward estimation 
+with each feature, weighted by the response triggered by the feature 
+(Fig. \ref{fig:model-struc} B, dots represent the feature weights) 
+The response of each feature center dimishes as a gaussian function  with the distance between the feature center and the badge size of the partner (Fig.
+\ref{fig:model-struc} B, grey line).
+Formally the reward estimation
 $\hat{R}$ when the focal individual ($i$) faces individual $j$ is given by,
@@
 where $x_z$ is the weight of feature center $z$ on the reward estimation; and $\theta$ is
-the width of the generalization function. Similarly, the tendency to retreat is given by sum 
-of feature weigths, and the probability of associated with the tendency is obtained
-by applying a logistic transformation. So formally, the log-odds to retreat
-when facing individual $j$ is
+the width of the generalization function. Similarly, the tendency to retreat is given by the sum 
+of feature weigths associated with the actor, and the probability is obtained
+by applying a logistic transformation (Fig. \ref{fig:model-struc} B, black line). 
+So formally, the log-odds to retreat when facing individual $j$ is
 
@@
 obtained throughout their life. Thus, the combination of natural selection and genetic drift
-changes the distribution of values in $\alpha$ and $\beta$ that segregates in the population, effectively the badges expressed and changing the communication system.
+changes the distribution of values in $\alpha$ and $\beta$ that segregates in the population, effectively changing  the badges expressed and the communication system.
 
-```{r Figure 1,echo=FALSE,fig.width=8,fig.align='center',fig.height=4,fig.show='hold',fig.cap= "Model of communication in the context of aggressive interactions.
+```{r model-struc,echo=FALSE,fig.width=8,fig.align='center',fig.height=4,fig.show='hold',fig.cap= "Model of communication in the context of aggressive interactions.
 In A the sender code: a reaction norm determines how the badge size is determined by the quality of the individual.

### 2020-12-22 Dyn_reac_learn_noLearn.R

@@ -35,7 +35,7 @@ param<-fromJSON(here(SimDir,paste0(scenario,"_"),paramName[9]))
 numCores <- length(evolList)
 registerDoParallel(numCores)
 
-val<-1
+val<-2
 
 
 # loop to produce pdfs for parameter values
@@ -259,7 +259,7 @@ traitsTrajs<-dcast(evol,time~seed,
 
 
 for(runChoi in finReps){
-   # runChoi<-0
+   runChoi<-0
   
 # Dynamics of genotypic traits (reaction norm) -  signaller
 par(plt=posPlot(numploty = 3,idploty = 2,numplotx = 3,idplotx = 1,
@@ -300,9 +300,10 @@ axis(side=1,padj = -3.5,cex=0.8,at=axTicks(1),labels = axTicks(1)/100)
 par(plt=posPlot(numploty = 3,idploty = 2,numplotx = 3,idplotx = 2,
                 lowboundx = 8,upboundx = 93),xaxt="s",las=1,new=TRUE)
 plot(x=c(0,max(evolStats$time)),y=c(0,0),type="l",lwd=2,col="grey",
-     ylim=fivenum(as.matrix(evol[,.(meanAlphaAtt+sdAlphaAtt,meanAlphaAtt-sdAlphaAtt,
-                                    meanBetaAtt+sdBetaAtt,meanBetaAtt-sdBetaAtt,
-                                    meanGammaAtt+sdGammaAtt,meanGammaAtt-sdGammaAtt)
+     # ylim=fivenum(as.matrix(evol[,.(meanAlphaAtt+sdAlphaAtt,meanAlphaAtt-sdAlphaAtt,
+     #                                meanBetaAtt+sdBetaAtt,meanBetaAtt-sdBetaAtt,
+     #                                meanGammaAtt+sdGammaAtt,meanGammaAtt-sdGammaAtt)
+     ylim=fivenum(as.matrix(evol[,.(meanAlphaAtt+sdAlphaAtt,meanAlphaAtt-sdAlphaAtt)
                                  ]))[c(1,5)],yaxt="n",
      xlab="",ylab="",cex.lab=1.2,cex.axis=1,xaxt='n',las=1)
 # polygon(x=c(evolStats$time,rev(evolStats$time)),
@@ -376,6 +377,7 @@ legend("topleft",legend = c("HH","DD","HD"),ncol = 3,
 
 
[…]

### 2020-12-22 cue_compt.cpp

@@ -424,7 +424,7 @@ void individual::get_payoff(individual partner,vector<double> payoff_matrix,
 		cum_payoff += payoff_matrix[1] - payoff_matrix[0] * shareCost;
 	}
 	else if (scenario == 0) {
-		curr_payoff = - payoff_matrix[0] * (1-shareCost);
+		curr_payoff = - payoff_matrix[0] * (1 - shareCost);
 		cum_payoff += - payoff_matrix[0] * (1 - shareCost);
 	}
 	else {

### 2021-01-19 manuscript/manuscript_1.0.Rmd
Fig 1, depicting the model and the effect of learning is finished.

@@
 where conditions are hard to predict [@dridi_Environmental_2016]. Besides its wide taxonomic and
-ecological relevance, associative learning is a flexible cognitive mechanism whose underpinings
+ecological relevance, associative learning is a flexible cognitive mechanism whose underpinnings
 show interspecific variation [@enquist_Power_2016; @quinones_Reinforcement_2019], so 
@@
 Unless, with a small probability, mutation changes the allelic value of the offspring.
-After birth individuals go through a round of viavility selection. 
+After birth individuals go through a round of viability selection. 
 Individual specific survival probability is given by $s_i=1/(1+e^{-k_1-k_2(Q_i-B_i)})$. 
@@
 In order to allow for non-random interactions, we use the population vector to define 
-an interaction neighborhood for the focal individual, the size of the neighborhood (g) 
+an interaction neighbourhood for the focal individual, the size of the neighbourhood (g) 
 is the number of positions around the focal's from which the interactant is drawn. 
@@
 In an interaction each individual must decide whether to escalate a fight or not. 
-Following the classic hawk-dove game [@maynard-smith_Evolution_1982], if the focal individual fights and its partner does not, the focal gets as payoff the resource of value $V$ while its partner gets nothing; if both individuals restrain  from fighting they split up the resource in half; if both decide to fight they splitup the cost of fighting and the winner takes over the resource.
+Following the classic hawk-dove game [@maynard-smith_Evolution_1982], if the focal individual fights and its partner does not, the focal gets as pay-off the resource of value $V$ while its partner gets nothing; if both individuals restrain  from fighting they split up the resource in half; if both decide to fight they split up the cost of fighting and the winner takes over the resource.
 We further assume that the probability of wining a fight for the focal individual depends on the difference in quality between her and her interacting partner; specifically it is given by $p_{ij}=1/(1+e^{-k_3(Q_i-Q_j)})$, where $k_3$ is a parameter defining
@@
 the focal´s experiences through a learning process. We implement learning using 
-the actor-critic aproach from reinforcement learning theory [RL; @sutton_Reinforcement_2018;@quinones_Reinforcement_2019; @leimar_Learning_2019].
-Individuals estimate the reward (payoff) expected from  interacting with partners 
+the actor-critic approach from reinforcement learning theory [RL; @sutton_Reinforcement_2018;@quinones_Reinforcement_2019; @leimar_Learning_2019].
+Individuals estimate the reward (pay-off) expected from  interacting with partners 
 of different badge sizes (the critic in RL terminology). 
-After every interaction they update the estimate of reward proportionally to the difference between their current estimate and the observed reward (prediction error $\delta$).
+After every interaction they update the estimate of reward proportionally to the difference between their current estimate and the observed reward (prediction error $\delta$) and to 
+the speed of learning ($\mathrm{A}$).
 Furthermore, they express different probabilities of retreating (attacking) depending on
@@
 approximation method based on radial basis functions [@sutton_Reinforcement_2018].
-Specifically, we pick $c$ feature centers, which are evenly spaced values along the badge size interval ([0,1]).
-Each one of these features centers is associated with a weight for reward estimation and
+Specifically, we pick $c$ feature centres, which are evenly spaced values along the badge size interval ([0,1]).
+Each one of these features centres is associated with a weight for reward estimation and
 tendency to play retreat in a given interaction. The reward estimation and the tendency 
@@
 (Fig. \ref{fig:model-struc} B, dots represent the feature weights) 
-The response of each feature center dimishes as a gaussian function  with the distance between the feature center and the badge size of the partner (Fig.
+The response of each feature centre diminishes as a Gaussian function  with the distance between the feature centre and the badge size of the partner (Fig.
 \ref{fig:model-struc} B, grey line).
@@
 
-where $x_z$ is the weight of feature center $z$ on the reward estimation; and $\theta$ is
+where $x_z$ is the weight of feature centre $z$ on the reward estimation; and $\theta$ is
 the width of the generalization function. Similarly, the tendency to retreat is given by the sum 
-of feature weigths associated with the actor, and the probability is obtained
+of feature weights associated with the actor, and the probability is obtained
 by applying a logistic transformation (Fig. \ref{fig:model-struc} B, black line). 
@@
 
-where $y_z$ is the weight of feature center $z$ on the tendency to retreat.
+where $y_z$ is the weight of feature centre $z$ on the tendency to retreat.
 
 After all their interactions, individuals in the population reproduce proportionally to 
-their fitness $w_i$, which is a sum of the baseline fitness ($w_0$) and all the payoffs
+their fitness $w_i$, which is a sum of the baseline fitness ($w_0$) and all the pay-offs
 obtained throughout their life. Thus, the combination of natural selection and genetic drift
@@
 
-```{r model-struc,echo=FALSE,fig.width=8,fig.align='center',fig.height=4,fig.show='hold',fig.cap= "Model of communication in the context of aggressive interactions.
+```{r model-struc,echo=FALSE,fig.width=8,fig.align='center',fig.height=8,fig.show='hold',fig.cap= "Model of communication in the context of aggressive interactions.
 In A the sender code: a reaction norm determines how the badge size is determined by the quality of the individual.
-Red shows an honest reaction norm, while the blue shows a dishonest reaction norm.
+Red shows an informative reaction norm, while the blue shows a uninformative reaction norm.
 B the receiver code: individuals have a behavioral reaction norm that determines their probability of retreating (black line).
@@
 Generalization is represented by the grey line and its axis, which shows the response triggered by the fourth feature diminishes as the value evaluated is further from the feature center.
-The learning process moves the feature weights (black dots) up or down depending on whether that leads to an increase in the estimated reward."}
+The learning process moves the feature weights (black dots) up or down depending on whether that leads to an increase in the estimated reward.
+In C and D, effect of learning on the receiver strategy.
+Receivers in C face signallers with uninformative reaction norms (blue line in A), while in D they face signallers with informative reaction norms (red line in A).
+Colour scale in C (applies also for D) indicates the quality of the individual"}
+
+# Top panels -  A the sender code and  B learning for the receiver code
+
 honestBig<-readPNG(here("Images/cart_honestBig.png"))
@@
 dishonest<-readPNG(here("Images/cart_dishonest.png"))
-par(plt=posPlot(numplotx = 2,idplotx = 1,upboundx = 91),las=1)
-  rangPars<-data.frame(honest=c(4,2),dishonest=c(0,-1.8))
+par(plt=posPlot(numplotx = 2,idplotx = 1,numploty = 1,idploty = 1,upboundx = 91),
+    las=1,mfrow=c(2,1))
+  rangPars<-data.frame(honest=c(4,2),dishonest=c(0,0))
  rangQual<-seq(0,1,by = 0.01)
@@
  plot(0,0,type="l",xlab="",xlim=c(0,1),
-      ylab ="",cex.lab=1,lwd=2,col="grey",cex.axis=1,ylim=c(0,1.1))
+      ylab ="",cex.lab=1,lwd=2,col="grey",cex.axis=1,ylim=c(0,1.1),xaxt="n")
+ axis(1)
  text(x = 0.08,y=1.05,"A",cex = 1.5)
@@
  mtext("Badge size",2,line = 2.3,cex = 1,las = 0)
- rasterImage(honestBig,0.75,0.6,0.95,0.8)
+ # include bird cartoons
+ rasterImage(honestBig,0.75,0.87,0.95,1.07)
  rasterImage(honestSmall,0.05,0.01,0.25,0.21)
- rasterImage(dishonest,0.75,0.87,0.95,1.07)
- rasterImage(dishonest,0.05,0.87,0.25,1.07)
+ rasterImage(dishonest,0.75,0.5,0.95,0.7)
+ rasterImage(dishonest,0.05,0.5,0.25,0.7)
  # legend("left",legend=rangeBeta,col=colRuns,lwd=3,title = expression(beta),cex=2,bty="n")
 
- par(plt=posPlot(numplotx = 2,idplotx = 2,upboundx = 91),las=1,new=TRUE)
-
+ par(plt=posPlot(numplotx = 2,idplotx = 2,numploty = 1,idploty = 1,
+                 upboundx = 91),las=1,new=TRUE)
   nCenters<-6
@@
 
-# xRBFs<-sapply(centers, function(x){seq(x-0.2,x+0.2,by=0.001)})
-# RBFS<-sapply(1:6, function(x){
-#   logist(weights[x]*RBF(seq(centers[x]-0.2,centers[x]+0.2,by=0.001),xCenter = centers[x],
-#                         sigSq = 0.01))
-# })
-# matlines(x=xRBFs,y=RBFS,col="grey",lty=2,lwd=2)
-
 mtext("prob. retreat",4,line = 2.3,cex = 1,las = 0)
 mtext("Badge size",1,line = 2,cex = 1,las = 0)
+
+## Lower panels - the effect of learning
+
+scenario1<-"alphaActDishonest"
+scenario2<-"alphaActHonest"
+
+extSimsDir<-paste0("e:/BadgeSims/",scenario1,"_")
+listTest<-list.files(extSimsDir,full.names = TRUE)
+indList<-grep("ind",listTest,value=TRUE)
+# parameter values from project folder
+paramName<-list.files(here("Simulations",paste0(scenario1,"_")))
+# paramName<-list.files(extSimsDir,full.names = TRUE)
+paramName<-grep(".json",paramName,value=TRUE)
+param<-fromJSON(here("Simulations",paste0(scenario1,"_"),paramName[2]))
+# fromJSON(paramName[1])
+fileId<-1 # choose the replicate
+
+indLearn<-fread(indList[fileId]) # load file
+Valpar<-gsub("[[:alpha:]]",gsub(".txt","",tail(strsplit(indList[fileId],"_")[[1]],1)),
+             replacement = "")
+nampar<-gsub("[^[:alpha:]]",gsub(".txt","",tail(strsplit(indList[fileId],"_")[[1]],1)),
+             replacement = "")
+
+gener<-0#tail(indLearn[,unique(time)],2)[2]
+nCenters<-param$nCenters
+sigSquar<-param$sigSq
+interv<-1/(nCenters-1)
+centers<-interv*seq(0,nCenters-1)
+# nCenters<-5
+# interv<-1/nCenters
+# centers<-interv*0.5+interv*seq(0,nCenters-1)
+rangx<-seq(0,1,length=100)
+cexAxis<-2
+colorbreaksQual<-seq(0,1,length=100)
+tempPop<-indLearn[time==gener]
+behavTime<- tempPop[,tail(unique(nInteract),2)[1],by=indId][,min(V1)]
+
+rangx<-seq(0,1,by=0.01)
+  dataIndsAct<-sapply(as.list(tempPop[nInteract==behavTime,indId]),
+                      function(x){x=
+                        logist(totRBF(rangx,
+                                      centers,sigSquar,
+                                      as.double(
+                                        tempPop[(nInteract==behavTime
+                                                 &indId==x)
+                                                # &(Quality>0.8&Quality<1)
+                                                ,.SD,
+                                                .SDcol=grep("WeightAct",
+                                                            names(tempPop),
+                                                            value = TRUE)
+                                                ])),alpha = 0,beta = 1)})
+  # par(plt=posPlot(numploty = 4,numplotx = 4,idploty = 2,idplotx = 1))
+  par(plt=posPlot(numploty = 1,numplotx = 2,idploty = 1,idplotx = 1,upboundx = 91))
+matplot(x=rangx,y=dataIndsAct,type='l',xlab="",ylab="prob. retreat",
+          xaxt="s",yaxt="s",lty = 1,
+          col=paletteMeans(100)[
+            findInterval(tempPop[nInteract==1700,Quality],colorbreaksQual)],
+          lwd=0.5,ylim=c(0,1.1))
+text(x = 0.08,y=1.05,"C",cex = 1.5)
+par(new=FALSE)
+color.bar.aeqp(paletteMeans1(100),min =min(colorbreaksQual),
+               max = max(colorbreaksQual),nticks = 3,
+               cex.tit = 0.9,title = "quality",
+               numplotx = 20,numploty = 13,idplotx =9,idploty = 2)
+
+extSimsDir<-paste0("e:/BadgeSims/",scenario2,"_")
+listTest<-list.files(extSimsDir,full.names = TRUE)
+indList<-grep("ind",listTest,value=TRUE)
+# parameter values from project folder
+paramName<-list.files(here("Simulations",paste0(scenario1,"_")))
+# paramName<-list.files(extSimsDir,full.names = TRUE)
+paramName<-grep(".json",paramName,value=TRUE)
+param<-fromJSON(here("Simulations",paste0(scenario1,"_"),paramName[2]))
+# fromJSON(paramName[1])
+fileId<-1 # choose the replicate
+
+indLearn<-fread(indList[fileId]) # load file
+Valpar<-gsub("[[:alpha:]]",gsub(".txt","",tail(strsplit(indList[fileId],"_")[[1]],1)),
+             replacement = "")
+nampar<-gsub("[^[:alpha:]]",gsub(".txt","",tail(strsplit(indList[fileId],"_")[[1]],1)),
+             replacement = "")
+
+gener<-0#tail(indLearn[,unique(time)],2)[2]
+nCenters<-param$nCenters
+sigSquar<-param$sigSq
+interv<-1/(nCenters-1)
+centers<-interv*seq(0,nCenters-1)
+# nCenters<-5
+# interv<-1/nCenters
+# centers<-interv*0.5+interv*seq(0,nCenters-1)
+rangx<-seq(0,1,length=100)
+cexAxis<-2
+colorbreaksQual<-seq(0,1,length=100)
+tempPop<-indLearn[time==gener]
+behavTime<- tempPop[,tail(unique(nInteract),2)[1],by=indId][,min(V1)]
+
+rangx<-seq(0,1,by=0.01)
+  dataIndsAct<-sapply(as.list(tempPop[nInteract==behavTime,indId]),
+                      function(x){x=
+                        logist(totRBF(rangx,
+                                      centers,sigSquar,
+                                      as.double(
+                                        tempPop[(nInteract==behavTime
+                                                 &indId==x)
+                                                # &(Quality>0.8&Quality<1)
+                                                ,.SD,
+                                                .SDcol=grep("WeightAct",
+                                                            names(tempPop),
+                                                            value = TRUE)
+                                                ])),alpha = 0,beta = 1)})
+  # par(plt=posPlot(numploty = 4,numplotx = 4,idploty = 2,idplotx = 1))
+  par(plt=posPlot(numploty = 1,numplotx = 2,idploty = 1,idplotx = 2,upboundx = 91)
+      ,new=TRUE,xpd=TRUE)
+matplot(x=rangx,y=dataIndsAct,type='l',xlab="",ylab="",
+          xaxt="s",yaxt="n",lty = 1,
+          col=paletteMeans(100)[
+            findInterval(tempPop[nInteract==1700,Quality],colorbreaksQual)],
+          lwd=0.5,ylim=c(0,1.1))
+text(x = 0.08,y=1.05,"D",cex = 1.5)
+text(x = -0.1,y = -0.23,labels = "Badge size")
 
@@
 
+We first present the outcome of simulations where we prevent the evolution of the 
+sender code (by setting the mutation rate to $0$), and assume that all individuals 
+in the population display either an uninformative or informative badge size with respect to their quality. These simulations show what type of receiver strategy develops through a learning process. 
+Figure \ref{fig:model-struc} (C and D) shows the receiver strategy developed 
+through learning; panel C is for receivers that faced uninformative signals, 
+and D informative ones. When learners face uninformative signals, they change their 
+probability of retreating depending on their own quality. Individuals with higher 
+quality (red tones), after the learning process, have a high probability of attacking; while 
+individuals of lower quality (blue tones) mostly retreat from confrontations. Thus, learning splits
+the population of receivers into the two classic pure strategies of hawks and doves. 
+Given that we have a assumed a monomorphic population with unresponsive reaction norms 
+on the signalling side, the changes triggered by learning only affect a small range 
+of badge sizes. In contrast, when receivers face informative reaction norms 
+on the side of the signaller (panel B in fig. \ref{fig:model-struc}), receivers use the 
+badge size of their interacting partners to determine whether to retreat or attack. 
+The relationship is given by a threashold-like reaction norm, where the change from 
+retreating to attacking depends on the quality of the receiver. As expected, 
+the higher the quality of the receiver the larger the badge size that triggers a retreat.
+
+When we allow the evolution of the sender code ($\alpha$ and $\beta$ change subject to natural selection and genetic drift), and the badge size works as a handicap 
+(i.e. the cost of the badge is inversely proportional to the quality of the individual), the sender code evolves to produce an honest signalFig \ref{fig:handicap}).
+
 ## References

### 2021-01-19 learnDynamics.R

@@ -11,7 +11,7 @@ source(here("AccFunc.R"))
 
 # Scenario to be plotted - corresponds to folders where simulations are stored
 
-scenario<-"alphaAct"
+scenario<-"alphaActHonest"
 
 extSimsDir<-paste0("e:/BadgeSims/",scenario,"_")
 
@@ -74,7 +74,7 @@ seedCh<-0
 # Select run and generation to plot
 tempPop<-indLearn[time==gener&seed==seedCh]
 
-timePoints<-round(seq(1,length(unique(tempPop[,nInteract]))-90,
+timePoints<-round(seq(1,length(unique(tempPop[,nInteract])),
                       length.out = 10))
 
 # png(here("Simulations",paste0(scenario,"_"),paste0(nampar,Valpar,"learnDyn.png")),
@@ -222,7 +222,7 @@ colorbreaksQual<-seq(0,1,length=100)
 # Actor 
 
 finReps<-indLearn[time==max(time),unique(seed)]
-seedCh<-5
+seedCh<-0
 finReps[round(runif(1,0,length(finReps)))+1]
 
 
@@ -245,7 +245,7 @@ xAxLabs[3]<-"Badge size"
 yAxLabs<-c("p(dove)","")
 
 
-par(mfrow=c(1,nIntPlot))
+par(mfrow=c(1,nIntPlot),plt=posPlot())
 
 
 for(countInts in 1:nIntPlot){

### 2021-01-19 parTOjson.R

@@ -10,36 +10,36 @@ fileName<-"parameters"
 here()
 
 (param1<-#fromJSON(paste0("E:/BadgeSims","/betCostNoLearn_","/parameters.json"))
-     fromJSON(here("Simulations","betCostNoLearnShare0_","parameters0.json")))
+     fromJSON(here("Simulations","alphaAct_","parameters1.json")))
 
-param<-list(totGen   = 20000,   nRep    = 1, seed = 1,
-            printGen = 100,   printLearn = 100,
+param<-list(totGen   = 1,   nRep    = 1, seed = 1,
+            printGen = 1,   printLearn = 1,
             printLearnInt = 50,
             popSize  = 2000,  baselineFit = 2,   
-            MutSd    = 2,  nInt        = 5,  init     = c(0,0,0,1),
-            mutRate  = 0.001,  mutType  = 0,
+            MutSd    = 2,  nInt        = 10000,  init     = c(0,0,1,0),
+            mutRate  = 0.00,  mutType  = 0,
             sampleSize = 50,   strQual  = 10,
-            errorQual = 0,    alphaBad	 = I(c(0)),    
-            betaBad	 = I(c(0)),
+            errorQual = 0,    alphaBad	 = I(c(3)),    
+            betaBad	 = I(c(6)),
             alphaRes =-10,betaRes=-20, gammaRes=0,
-            alphaCrit  = 0.4,  alphaAct = 0.4,
+            alphaCrit  = 0.05,  alphaAct = 0.05,
             sigSq   	 = 0.005, nCenters = 10,
             initCrit = 0,      initAct=0,   gamma = 0,
-            QualStDv   = 0.15, betCost = 6,
+            QualStDv   = 0.15, betCost = 0,
             alphCost	 = 3, mutLearn = FALSE,
             nIntGroup  = 2000, 
             payoff_matrix = c(3,1,0,0.5),
             shareCost = 0,
-            namParam = "betCost",
-            rangParam = I(c(0,6)),
-            typeAgent = 3, #//0. hawk 1.dove 2. learner 3. evaluator
+            namParam = "alphaAct",
+            rangParam = I(c(0.01)),
+            typeAgent = 2, #//0. hawk 1.dove 2. learner 3. evaluator
             folderL=paste(here("Simulations"),"/",sep="")) # comment for debug
 # folderL=paste(here("Simulations","test_"),"/",sep="")) # comment for release
 
 
 
[…]

### 2021-08-11 manuscript/manuscript_1.0.Rmd
Moving forward in the manuscript

@@
 - \linenumbers
+- \floatplacement{figure}{H}
 keep_tex: yes
@@
 Acquiring this information, however, is far from trivial. 
-In some cases, interacting partners (signalers) might be 'willing' to provide accurate
+In some cases, interacting partners (signallers) might be 'willing' to provide accurate
 information, but in others it might be in their own interest to conceal information
@@
 Associative learning exists in all major vertebrate taxa, and in many invertebrates [@heyes_Simple_2012;@macphail_Brain_1982; @staddon_Adaptive_2016; @behrens_Associative_2008].
-Theory has shown that natural selection favors these associations in complex environments
+Theory has shown that natural selection favours these associations in complex environments
 where conditions are hard to predict [@dridi_Environmental_2016]. Besides its wide taxonomic and
@@
 lack of integration between learning and evolution. Reinforcement learning theory 
-encompases a series of computational methods inspired on the psychological and 
+encompasses a series of computational methods inspired on the psychological and 
 neurological mechanisms of associative learning [@sutton_Reinforcement_2018].
@@
 Furthermore, these algorithms can be embedded in evolutionary simulations to 
-generate theoretical predictions of the effect that learning can have in behavioral
+generate theoretical predictions of the effect that learning can have in behavioural
 evolution [@leimar_Learning_2019].
@@
 We model the evolution of signals of quality in the context of agonistic interactions by 
-modelling a population of haploid individuals with non-overlaping generations. 
+modelling a population of haploid individuals with non-overlapping generations. 
 Individuals are born every generation with a level of quality ($Q_i$, where $i$ is subscript 
@@
 par(new=FALSE)
-color.bar.aeqp(paletteMeans1(100),min =min(colorbreaksQual),
+color.bar.aeqp(paletteMeans(100),min =min(colorbreaksQual),
                max = max(colorbreaksQual),nticks = 3,
                cex.tit = 0.9,title = "quality",
-               numplotx = 20,numploty = 13,idplotx =9,idploty = 2)
+               numplotx = 20,numploty = 10,idplotx =9,idploty = 2)
 
@@
 badge size of their interacting partners to determine whether to retreat or attack. 
-The relationship is given by a threashold-like reaction norm, where the change from 
+The relationship is given by a threshold-like reaction norm, where the change from 
 retreating to attacking depends on the quality of the receiver. As expected, 
@@
 
-When we allow the evolution of the sender code ($\alpha$ and $\beta$ change subject to natural selection and genetic drift), and the badge size works as a handicap 
-(i.e. the cost of the badge is inversely proportional to the quality of the individual), the sender code evolves to produce an honest signalFig \ref{fig:handicap}).
+```{r handicap,echo=FALSE,out.width='90%',fig.align='center',fig.show='hold',fig.cap="Evolution of the badge size as a handicap mediated by learning.
+Middle panels show the evolutionary dynamics of the sender code.
+On the left, changes in the distribution of values for the intercept of the reaction norm ($\\alpha$); on the right changes in the distribution of the slope ($\\beta$).
+Dashed lines in both panels show the mean of the distribution.
+Grey lines show the generation time corresponding to the panels above and below showing the sender and receiver code.
+In the panels above, the reaction norms correspond to individuals in the population after the interaction round.
+Color scale indicates quality just as in Fig.
+\\ref{fig:model-struc}.
+In the panels below, colors represent a cluster clasification performed by the k-means algorithm" }
+
+knitr::include_graphics(here("Simulations","betCostEvol1_","evolDyn0_betCost5.png"))
+
+```
+
+When we allow the evolution of the sender code ($\alpha$ and $\beta$ change subject to natural selection and genetic drift), and the badge size work as a handicap 
+(i.e. the cost of the badge is inversely proportional to the quality of the individual), the sender code evolves to produce an honest signal Fig \ref{fig:handicap}). The evolution of 
+the sender code does not happen immediately after the start of the evolutionary process. Instead, the evolutionary dynamics of the  reaction norm parameters (Fig. \ref{fig:handicap}, X) show a set of 
+steps. First, the intercept of the reaction norm ($\alpha$) evolves to higher values, reducing the 
+average badge size in the populations.
+That makes sense because reaction norms segregating in the population are flat at this point, and consequently the receiver code does not respond to the badge size, bigger badges are costlier and do not trigger lower attack probabilities.
+During those first generations, the slope of the sender reaction norm ($\beta$)  remains around the initial value of zero.
+At around generation 4000, the slope evolves toward positive values.
+Which implies that larger badges correlate with higher quality.
+Receivers learn to react to such correlation, reducing
+the probability of attack towards individual of larger badges. Hence, natural selection 
+favours larger values of $\beta$, eventually leading to an evolutionary equilibrium where 
+badge size is an honest signal of quality mediated by the learned responses of receivers. The
+evolutionary trajectory portrayed in figure \ref{fig:handicap} is not the only possible trajectory. 
+If the slope of the sender reaction norms ($\beta$) evolves toward negative values before badges become handicaps, receivers never learn to react to the size of badges.
+Thus, badges are only costly and do not provide information about the quality of an individual.
+Eventually, the badge disappears from the population.
+
+When production of the badge is not costly (i.e.
+does not work as a handicap), the parameters of the sender code go through evolutionary branching event, triggereing the evolution of different types of
+signals in the population. Fig. \ref{fig:branching}A and B shows changes in the distribution of the 
+sender reaction norm ($\alpha$ and $\beta$) along evolutionary time. Darker areas show trait values 
+with high frequency in the population.
+Populations start monomorphic with a value of zero on both traits, and mutations quickly buid up a normal distribution around the starting value.
+Within the first 2000 generations the unimodal distribution splits into a bimodal one.
+In the case of $\alpha$, one of the peaks splits further, so at the end of the evolutionary simulation the distribution of $\alpha$ value in the population shows three peaks.
+In the case of $\beta$, the two initial peaks coexist for around 4000 generations.
+Later on one of the types goes extinct and the distribution goes back to being unimodal.
+These changes in the distrubtuion of the sender reaction norm parameters imply that during the second half of the simualation individuals can by classify into three distinct types.
+Two types express a flat reaction norm, meaning that their badge size is not informative of their quality, while the third type shows intermediate badge sizes which are determined by the quality of the individual (Fig.
+\ref{fig:branching}).
+Furthermore, the receiver reaction norms develop through learning, particularly those of individuals of intermediate quality, respond to the signal of their sender type by increasing the probability of retreating from a fight with individuals with larger badges (Fig.
+\ref{fig:branching}).
+
+```{r branching,echo=FALSE,out.width='90%',fig.align='center',fig.show='hold',fig.cap= "The evolution of cheap signals.
+Portrait of the evolutionary dynamics of the sender code with snapsots of both sender and receiver codes just as in fig.
+\\ref{fig:handicap}.
+The middle panels show changes in the distribution of values fir $\\alpha$ and $\\beta$ along evolutionary time.
+Panels above and bellow correspond to snapshots of the sender and receiver codes, respectively, generation time of the snapshots are indicated by the grey lines in the middle panels."}
+
+knitr::include_graphics(here("Simulations","nIntGroupNormQual_","evolDyn7_nIntGroup2000.png"))
+
+```
+
+```{r startCond, echo=FALSE, out.width='90%',fig.align='center',fig.show='hold',fig.cap="WHy would you want to be recognized"}
+
+knitr::include_graphics("../Simulations/initAct_/indVarScatter_.png")
+
+```
 

### 2021-08-11 AccFunc.R

@@ -293,6 +293,22 @@ get_clusters<-function(DT,vars,k.max=5,Bsamples=500,iterMax=100){
   return(totClusters[,2])
 }
 
+## Function to calculate ESS in the classic gam from the pay-off matrix --------
+
+ESS<-function(pay_offsM){
+  # /*  (This are IDs to the entries of the matrix, NOT payoffs)    
+  # Hawk      Dove     
+  # Hawk   0         1      
+  # Dove   2         3
+  # */
+  V<-pay_offsM[2]; C<-pay_offsM[2]
+  hawk<-V/C; dove<-1-hawk
+  HH<-hawk**2 ; HD<-hawk*dove; DD<-dove**2
+  if(HH+HD+DD!=1)  warning("proportions do not add up to 1")
+  return(list(hawk=hawk,dove=dove,HH=HH,HD=HD,DD=DD))
+}
+
+
 
 # clus_silou<-c(0,clus_silou)
 # nclustersSil<-which.max(clus_silou)

### 2021-08-11 Dyn_reac_learn.R

@@ -12,7 +12,7 @@ require("jsonlite")
 
 # Scenario to be plotted - corresponds to folders where simulations are stored
 
-scenario<-"test"
+scenario<-"betCostEvolNorm"
 
 extSimsDir<-#here("Simulations",paste0(scenario,"_"))
   paste0("e:/BadgeSims/",scenario,"_")
@@ -37,7 +37,7 @@ param<-fromJSON(here("Simulations",paste0(scenario,"_"),paramName[1]))
 
 # Choose which parameter to plot
 
-val<-1
+val<-2
 
 
 # Load files -------------------------------------------------------------------
@@ -49,7 +49,7 @@ val<-1
 numCores <- length(param$rangParam)
 registerDoParallel(numCores)
 
-foreach(val = 1:3,#length(param$rangParam),
+foreach(val = 2:3,#length(param$rangParam),
         .packages = c("data.table","here")) %dopar% {
 source(here("AccFunc.R"))
 
@@ -72,17 +72,16 @@ source(here("AccFunc.R"))
   indList_runs<-grep(paste0(param$namParam,param$rangParam[val]),
                      indList,value =TRUE)
   
-  evol<-do.call(rbind,lapply(evolList_runs,function(x){
-    fread(here("Simulations",paste0(scenario,"_"),x))
-  }))
-  pop<-do.call(rbind,lapply(indList_runs,function(x){
-    fread(here("Simulations",paste0(scenario,"_"),x))
-  }))
+  # evol<-do.call(rbind,lapply(evolList_runs,function(x){
+  #   fread(here("Simulations",paste0(scenario,"_"),x))
+  # }))
+  # pop<-do.call(rbind,lapply(indList_runs,function(x){
+  #   fread(here("Simulations",paste0(scenario,"_"),x))
+  # }))
   
   
-  
-# evol<-do.call(rbind,lapply(evolList_runs,fread))
-# pop<-do.call(rbind,lapply(indList_runs, fread))
+evol<-do.call(rbind,lapply(evolList_runs,fread))
+pop<-do.call(rbind,lapply(indList_runs, fread))
 
 # temp fix to lack of name to the weights
 
@@ -96,7 +95,7 @@ source(here("AccFunc.R"))
 
 param<-fromJSON(here("Simulations",paste0(scenario,"_"),paramName[val]))
 
-Valpar<-param$rangParam
+Valpar<-param$rangParam[val]
 
 nampar<-param$namParam
 

### 2022-03-15 manuscript/manuscript_1.0.Rmd
new file with notes of to-dos. Small changes in the manuscript

@@
 style: C:\Users\andre\Zotero\styles\elife.csl
-abstract: Adaptive behavioural responses often depend on qualities of the interacting partner of an individual.
+abstract: | 
+  Adaptive behavioural responses often depend on qualities of the interacting partner of an individual.
 For example, when competing for resources, an individual might be better off escalating fights with individuals of lower quality, while restraining from fighting individuals of higher quality.
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE,cache=TRUE)
 require(png)
@@
 The outcome of interactions with con-specifics is an important 
-determinant of fitness in social animals. Irrespective of whether such interactions are cooperative 
-or competitive, the actions of interacting individuals partly determine 
+determinant of fitness in social animals. Irrespective of whether 
+such interactions are cooperative or competitive, the actions of 
+interacting individuals partly determine 
 their reproductive success. However, the best action for an individual 
-might vary depending on its condition, and that of individuals with whom it interacts. 
-Thus, acting based on information about interacting 
+might vary depending on its condition, and that of individuals with 
+whom it interacts. Thus, acting based on information about interacting 
 partners is typically adaptive [@quinones_Negotiation_2016].
 Acquiring this information, however, is far from trivial. 
-In some cases, interacting partners (signallers) might be 'willing' to provide accurate
-information, but in others it might be in their own interest to conceal information
-[@johnstone_Recognition_1997], or to provide erroneous information [@johnstone_Badges_1993]. 
+In some cases, interacting partners (signallers) might be 'willing' to 
+provide accurate information, but in others it might be in their 
+own interest to conceal information [@johnstone_Recognition_1997], 
+or to provide erroneous information [@johnstone_Badges_1993]. 
 Typically, in a given context, some proportion of individuals will benefit 
-from broadcasting accurate information, whereas others will benefit from hiding it. 
-Take, for instance, an interaction between two individuals where one can help the other. 
-Given some costs and benefits, a potential donor will be interested in helping 
+from broadcasting accurate information, whereas others will benefit 
+from hiding it. Take, for instance, an interaction between two individuals 
+where one can help the other. Given some costs and benefits, 
+a potential donor will be interested in helping 
 related individuals. Therefore, for a relative of the donor, 
@@
 concealing the lack of kinship is better. Similar scenarios may apply to 
-many aspects of social life such as finding mates, feeding offspring, dominance relationships 
-and aggressive interactions [@bradbury_Principles_2011; @moller_Female_1988; @tibbetts_Socially_2004].
+many aspects of social life such as finding mates, 
+feeding offspring, dominance relationships and aggressive interactions 
+[@bradbury_Principles_2011; @moller_Female_1988; @tibbetts_Socially_2004].
 In dominance and aggressive interactions, an important piece of
-information to guide individual actions is interacting partners fighting ability, 
-which is sometimes referred to as Resource Holding Potential (RHP) [@parker_Assessment_1974], 
-or simply as quality. Responsiveness to the quality of the 
-partner is central to communication systems such as badges of status, 
-where an arbitrary signal conveys quality [@johnstone_Badges_1993; @rohwer_Social_1975a]. 
-Here, arbitrary refers to a signal which is not ontogenetically correlated with quality. 
-Badges of status can be evolutionarily stable whenever the signal imposes a fitness cost which 
-decreases with the quality of an individual  [@botero_Evolution_2010; @johnstone_Badges_1993], 
-such that it's no longer in the interest of low quality individuals to fake quality by producing a
-large badge. The cost of honest signals is a more general principle of communications systems, 
-usually referred to as the handicap principle [@grafen_Biological_1990; @zahavi_Mate_1975]. 
-An alternative system of communication is one where individuals gather and store information
-about each individual they interact with. This information should reliably allow 
-each individual to react adaptively to the quality difference with their opponent. 
-This type of assessment strategy is often referred to as individual recognition [@whitfield_Plumage_1986]. Evidently, individual recognition is restricted to 
-certain interaction structures, because individuals are limited by cognitive 
-abilities in how many individuals they can recognize and track. Thus, individual 
-recognition and badges of status have been assumed to be alternative communication 
-systems fulfilling the same function, reducing the costs of aggression in contests over resources. 
-Furthermore, badges of status and individual recognition are expected to evolve under different interaction structures, badges of status under global interactions and individual recognition
-under local interactions [@dale_Signaling_2001; @sheehan_There_2016].  There is no clear cut reason, however, why both systems could not work together, in fact there is some evidence
+information to guide individual actions is interacting 
+partners fighting ability, which is sometimes referred to as 
+Resource Holding Potential (RHP) 
+[@parker_Assessment_1974], or simply as quality. 
+Responsiveness to the quality of the partner is central 
+to communication systems such as badges of status, 
+where an arbitrary signal conveys quality 
+[@johnstone_Badges_1993; @rohwer_Social_1975a]. 
+Here, arbitrary refers to a signal which is not 
+ontogenetically correlated with quality. 
+Badges of status can be evolutionarily stable whenever the 
+signal imposes a fitness cost which decreases with the 
+quality of an individual  [@botero_Evolution_2010; @johnstone_Badges_1993], 
+such that it's no longer in the interest of low quality individuals 
+to fake quality by producing a large badge. The cost of honest signals 
+is a more general principle of communications systems, 
+usually referred to as the handicap principle 
+[@grafen_Biological_1990; @zahavi_Mate_1975]. 
+<!-- should this last part of the paragraph stay ?-->
+An alternative system of communication is one where individuals 
+gather and store information about each individual they interact with. 
+This information should reliably allow each individual to react adaptively 
+to the quality difference with their opponent. This type of 
+assessment strategy is often referred to as individual 
+recognition [@whitfield_Plumage_1986]. Evidently, individual recognition 
+is restricted to  certain interaction structures, because individuals 
+are limited by cognitive abilities in how many individuals 
+they can recognize and track. Thus, individual recognition and 
+badges of status have been assumed to be alternative communication 
+systems fulfilling the same function, reducing the costs of 
+aggression in contests over resources.
+
+Furthermore, badges of status and individual recognition are expected 
+to evolve under different interaction structures, badges of status under 
+global interactions and individual recognition
+under local interactions [@dale_Signaling_2001; @sheehan_There_2016].  
+There is no clear cut reason, however, why both systems could not 
+work together, in fact there is some evidence
 that they can [@chaine_Manipulating_2018].
 
-A somewhat ignored component of communication systems, in the context of aggressive 
-interactions, is the cognitive aspects of the receiver module. Communication systems
-relying on badges of status are often assumed to be cognitively less demanding 
-than those involving individual recognition [@sheehan_There_2016] because in the latter, 
-individuals need to store identity and associated information about each individual 
-they interact with. In contrast, the response towards badges of status is often 
+A somewhat ignored component of communication systems, in the context 
+of aggressive interactions, is the cognitive aspects of the receiver module. 
+Communication systems relying on badges of status are often assumed 
+to be cognitively less demanding than those involving individual 
+recognition [@sheehan_There_2016] because in the latter, 
+individuals need to store identity and associated information 
+about each individual they interact with. In contrast, the response 
+towards badges of status is often 
 thought to be innate [@botero_Evolution_2010; @johnstone_Badges_1993], with 
-individuals using their own quality and the opponent's badge to determine whether to aggressively
-engage in contest. An alternative view of badges of status is that individuals learn 
-to react to them based on their experiences [@guilford_Receiver_1991], which would imply that receivers must learn to associate signals with the fighting ability of bearers just as in systems based on individual recognition.
-In cases where badges of status vary quantitatively (e.g.
-in size or intensity), fighting ability may increase monotonically with attributes of the badge and would be reinforced by every interaction, so in principle learning would be faster than in systems involving individual recognition in which the association between signals and their meaning would vary depending on the interacting partners.
-In any case, learning could be a central cognitive mechanism in both types of communication systems, but the role of learning in these contexts has not been thoroughly explored in the empirical nor theoretical literature.
+individuals using their own quality and the opponent's badge to determine 
+whether to aggressively engage in contest. An alternative view of 
+badges of status is that individuals learn 
+to react to them based on their experiences [@guilford_Receiver_1991], 
+which would imply that receivers must learn to associate signals with 
+the fighting ability of bearers just as in systems based on individual 
+recognition. In cases where badges of status vary quantitatively 
+(e.g. in size or intensity), fighting ability may increase monotonically 
+with attributes of the badge and would be reinforced by every interaction, 
+so in principle learning would be faster than in systems involving individual 
+recognition in which the association between signals and their meaning would 
+vary depending on the interacting partners. In any case, learning could be 
+a central cognitive mechanism in both types of communication systems, 
+but the role of learning in these contexts has not been thoroughly 
+explored in the empirical nor theoretical literature.
 
-Associative learning is a key cognitive mechanism that allows individuals to associate
-rewards with environmental stimuli and thus behave adaptively [@staddon_Adaptive_2016]. 
-Associative learning exists in all major vertebrate taxa, and in many invertebrates [@heyes_Simple_2012;@macphail_Brain_1982; @staddon_Adaptive_2016; @behrens_Associative_2008].
-Theory has shown that natural selection favours these associations in complex environments
-where conditions are hard to predict [@dridi_Environmental_2016]. Besides its wide taxonomic and
-ecological relevance, associative learning is a flexible cognitive mechanism whose underpinnings
-show interspecific variation [@enquist_Power_2016; @quinones_Reinforcement_2019], so 
-presumably, it has been modified by natural selection. Despite all this, associative 
-learning is not often included in the narrative of evolutionary explanations of 
-behavioural patterns [@fawcett_Exposing_2013; @kamil_Optimal_1983; @mcauliffe_Psychology_2015]. 
-Computational models of evolution can play an important role to overcome the 
-lack of integration between learning and evolution. Reinforcement learning theory 
-encompasses a series of computational methods inspired on the psychological and 
-neurological mechanisms of associative learning [@sutton_Reinforcement_2018].
-These set of algorithms allow the implementation of biologically realistic problems, 
+Associative learning is a key cognitive mechanism that allows 
+individuals to associate rewards with environmental stimuli 
+and thus behave adaptively [@staddon_Adaptive_2016]. Associative 
+learning exists in all major vertebrate taxa, and in many invertebrates [@heyes_Simple_2012;@macphail_Brain_1982; @staddon_Adaptive_2016; 
+@behrens_Associative_2008].
+Theory has shown that natural selection favours these associations 
+in complex environments where conditions are hard to predict 
+[@dridi_Environmental_2016]. Besides its wide taxonomic and
+ecological relevance, associative learning is a flexible cognitive 
+mechanism whose underpinnings show interspecific variation [@enquist_Power_2016; @quinones_Reinforcement_2019], so presumably, it has been modified 
+by natural selection. Despite all this, associative learning is 
+not often included in the narrative of evolutionary explanations of 
+behavioural patterns [@fawcett_Exposing_2013; @kamil_Optimal_1983; 
+@mcauliffe_Psychology_2015]. Computational models of evolution 
+can play an important role to overcome the lack of integration 
+between learning and evolution. Reinforcement learning theory 
+encompasses a series of computational methods inspired 
+on the psychological and neurological mechanisms of associative learning [@sutton_Reinforcement_2018]. These set of algorithms allow the 
+implementation of biologically realistic problems, 
 capturing the essence of learning processes [@frankenhuis_Enriching_2018; @quinones_Reinforcement_2019]. 
 Furthermore, these algorithms can be embedded in evolutionary simulations to 
-generate theoretical predictions of the effect that learning can have in behavioural
-evolution [@leimar_Learning_2019].
+generate theoretical predictions of the effect that learning can 
+have in behavioural evolution [@leimar_Learning_2019].
 
-In here we present an evolutionary model where individuals use associative learning 
-to develop a tendency to behave aggressively (or  peacefully), in the context of
-competition over resources, depending on the quantitative morphological trait 
-they perceive in their opponent (badge). Over evolutionary time individuals evolve
-the size of their badge, and whether it depends on their quality. Under this simple
+In here we present an evolutionary model where individuals use 
+associative learning to develop a tendency to behave aggressively 
+(or  peacefully), in the context of competition over resources, 
+depending on the quantitative morphological trait they 
+perceive in their opponent (badge). Over evolutionary time 
+individuals evolve the size of their badge, and whether it depends 
+on their quality. Under this simple
 set up, individuals can use the badge both as a signal of quality (BS)
-or as a trait for IR. We use the model to assess under what conditions of interaction
-structure we expect different communication signals to evolve.
+or as a trait for IR. We use the model to assess under what conditions 
+of interaction structure we expect different communication signals to evolve.
 
@@
 
-We model the evolution of signals of quality in the context of agonistic interactions by 
-modelling a population of haploid individuals with non-overlapping generations. 
-Individuals are born every generation with a level of quality ($Q_i$, where $i$ is subscript 
-of the population vector of size $N$) given by a number between zero and one, which is drawn 
-from a truncated normal distribution $N(\mu = 0.5,\sigma)$. 
-An individual with quality $0$ has the lowest RHP, 
-while an individual with quality $1$ has the highest. As part of development processes
-individuals produce a morphological signal (badge), the size of such depends on their level 
-of quality according to a reaction norm given by $B_i=1/(1+e^{\alpha_i-\beta_i Q_i})$;
-where are $\alpha_i$ and $\beta_i$ are individual specific traits that determine 
-the shape of the reaction norm (Fig. \ref{fig:model-struc} A). Variation in $\alpha_i$ and $\beta_i$ means individuals can have either uninformative (flat) or informative (logistic) norms. 
-The size of the signal is constrained between zero and one. We assume different values of these traits 
+We model the evolution of signals of quality in the context of agonistic 
+interactions. For simplicity we consider a population of haploid 
+individuals with non-overlapping generations. Individuals are born 
+every generation with a level of quality ($Q_i$, where $i$ is subscript 
+of the population vector of size $N$) given by a number between zero and one, 
+which is drawnfrom a truncated normal distribution 
+$N(\mu = 0.5,\sigma)$. An individual with quality $0$ has the lowest RHP, 
+while an individual with quality $1$ has the highest. As part of 
+development processes individuals produce a morphological signal (badge), 
+the size of such depends on their level 
+of quality according to a reaction norm given by 
+$B_i=1/(1+e^{\alpha_i-\beta_i Q_i})$; where are $\alpha_i$ and 
+$\beta_i$ are individual specific traits that determine the shape of 
+the reaction norm (Fig. \ref{fig:model-struc} A). Variation in $\alpha_i$ 
+and $\beta_i$ means individuals can have either uninformative (flat) or 
+informative (logistic) norms. The size of the signal is constrained 
+between zero and one. We assume different values of these traits 
 are given by different alleles, and are inherited from mother to offspring. 
-Unless, with a small probability, mutation changes the allelic value of the offspring.
-After birth individuals go through a round of viability selection. 
-Individual specific survival probability is given by $s_i=1/(1+e^{-k_1-k_2(Q_i-B_i)})$. 
-Where $k_1$ and $k_2$ are parameter values determining the shape of the survival function. 
-Importantly, if $k_2=0$ survival probability is independent of quality, while if $k_2>0$
-lower quality individuals pay a higher price for s similar-sized badge, fulfilling the core assumption 
-of the handicap principle [@botero_Evolution_2010; @grafen_Biological_1990; @johnstone_Badges_1993].
+Unless, with a small probability, mutation changes the allelic value 
+of the offspring. After birth individuals go through a round of 
+viability selection. Individual specific survival probability is given by $s_i=1/(1+e^{-k_1-k_2(Q_i-B_i)})$. Where $k_1$ and $k_2$ are parameter 
+values determining the shape of the survival function. 
+Importantly, if $k_2=0$ survival probability is independent of quality, 
+while if $k_2>0$ lower quality individuals pay a higher price for similar-sized 
+badge, fulfilling the core assumption of the handicap principle 
+[@botero_Evolution_2010; @grafen_Biological_1990; @johnstone_Badges_1993].
 
 Individuals who survive engage in a series of pairwise interactions 
-where they compete for resources. Individuals in real populations typically do not 
-interact at random nor with all individuals in the population [@kurvers_Evolutionary_2014]. 
-In order to allow for non-random interactions, we use the population vector to define 
-an interaction neighbourhood for the focal individual, the size of the neighbourhood (g) 
-is the number of positions around the focal's from which the interactant is drawn. 
-So, if $g=N$ interactions are global; as $g$ gets smaller interactions are more local. 
-In an interaction each individual must decide whether to escalate a fight or not. 
-Following the classic hawk-dove game [@maynard-smith_Evolution_1982], if the focal individual fights and its partner does not, the focal gets as pay-off the resource of value $V$ while its partner gets nothing; if both individuals restrain  from fighting they split up the resource in half; if both decide to fight they split up the cost of fighting and the winner takes over the resource.
-We further assume that the probability of wining a fight for the focal individual depends on the difference in quality between her and her interacting partner; specifically it is given by $p_{ij}=1/(1+e^{-k_3(Q_i-Q_j)})$, where $k_3$ is a parameter defining
-how strong the quality difference determines the winning probability; 
-and $i$ and $j$ denote the the position in the population 
-vector of the focal and its interacting partner respectively.
+where they compete for resources. Individuals in real populations typically 
+do not interact at random nor with all individuals in the population 
+[@kurvers_Evolutionary_2014]. In order to allow for non-random interactions, 
+we use the population vector to define an interaction neighbourhood 
+for the focal individual, the size of the neighbourhood ($g$)
+is the number of positions around the focal's from which the partner is drawn. 
+So, if $g=N$ interactions are global; as $g$ gets smaller 
+interactions are more local. In an interaction each individual must 
+decide whether to escalate a fight or not. 
+Following the classic *hawk-dove* game [@maynard-smith_Evolution_1982], 
+if the focal individual fights and its partner does not, the focal gets 
+as pay-off the resource of value $V$ while its partner gets nothing; 
+if both individuals restrain  from fighting they split up the resource 
+in half; if both decide to fight they split up the cost of fighting and 
+the winner takes over the resource. We further assume that the 
+probability of wining a fight for the focal individual depends on the 
+difference in quality between her and her interacting partner; specifically 
+it is given by $p_{ij}=1/(1+e^{-k_3(Q_i-Q_j)})$, where $k_3$ is a 
+parameter defining how strong the quality difference determines the 
+winning probability; and $i$ and $j$ denote the the position 
+in the population vector of the focal and its interacting 
+partner respectively.
 
-The decision of whether to escalate a fight against an interacting partner can be dependent 
-on the size of the partner's badge, the actual dependency is determined by 
-the focal´s experiences through a learning process. We implement learning using 
+The decision of whether to escalate a fight against an interacting 
+partner can be dependent on the size of the partner's badge, 
+the actual dependency is determined by the focal´s experiences 
+through a learning process. We implement learning using 
 the actor-critic approach from reinforcement learning theory [RL; @sutton_Reinforcement_2018;@quinones_Reinforcement_2019; @leimar_Learning_2019].
-Individuals estimate the reward (pay-off) expected from  interacting with partners 
-of different badge sizes (the critic in RL terminology). 
-After every interaction they update the estimate of reward proportionally to the difference between their current estimate and the observed reward (prediction error $\delta$) and to 
-the speed of learning ($\mathrm{A}$).
-Furthermore, they express different probabilities of retreating (attacking) depending on
-the badge size of their opponent (the actor in RL). They update the probability of 
-retreating (attacking)  up or down depending on whether retreating (attacking) leads to an increase 
-in the reward estimation. So, if a focal individual decides to escalate a fight against 
-an individual with small badge, and that leads to an increase in the reward estimation,
-the focal individual will increase the probability of escalating fights with 
-individuals of small badges in the future.
-Given that badge size is a real number between 0 and 1, there are infinitely many badge sizes.
-Thus, the reward estimation, as well as the probability of retreating (escalating), must be generalized across different values.
-To implement generalization we use the linear function
-approximation method based on radial basis functions [@sutton_Reinforcement_2018].
-Specifically, we pick $c$ feature centres, which are evenly spaced values along the badge size interval ([0,1]).
-Each one of these features centres is associated with a weight for reward estimation and
-tendency to play retreat in a given interaction. The reward estimation and the tendency 
-to retreat are calculated (in every interaction) as the sum of the weights associated 
-with each feature, weighted by the response triggered by the feature 
-(Fig. \ref{fig:model-struc} B, dots represent the feature weights) 
-The response of each feature centre diminishes as a Gaussian function  with the distance between the feature centre and the badge size of the partner (Fig.
-\ref{fig:model-struc} B, grey line).
-Formally the reward estimation
-$\hat{R}$ when the focal individual ($i$) faces individual $j$ is given by,
-
+Individuals estimate the reward (pay-off) expected from  interacting 
+with partners of different badge sizes (the critic in RL terminology). 
+After every interaction they update the estimate of reward proportionally 
+to the difference between their current estimate and the observed 
+reward (prediction error $\delta$) and to the speed of learning 
+($\mathrm{A}$). Furthermore, they express different probabilities of 
+retreating (attacking) depending on the badge size of their opponent 
+(the actor in RL). They update the probability of 
+retreating (attacking)  up or down depending on whether retreating 
+(attacking) leads to an increase in the reward estimation. 
+So, if a focal individual decides to escalate a fight against 
+an individual with small badge, and that leads to an increase in 
+the reward estimation, the focal individual will increase the probability 
+of escalating fights with individuals of small badges in the future. 
+Given that badge size is a real number between 0 and 1, there are infinitely 
+many badge sizes. Thus, the reward estimation, as well as the probability 
+of retreating (escalating), must be generalized across different values. 
+To implement generalization we use the linear function approximation 
+method based on radial basis functions [@sutton_Reinforcement_2018]. 
+Specifically, we pick $c$ feature centres, which are evenly spaced values 
+along the badge size interval ([0,1]). Each one of these features centres 
+is associated with a weight for reward estimation and tendency to play 
+retreat in a given interaction. The reward estimation and the tendency 
+to retreat are calculated (in every interaction) as the sum of the weights 
+associated with each feature, weighted by the response 
+triggered by the feature (Fig. \ref{fig:model-struc} B, dots 
+represent the feature weights) The response of each feature centre 
+diminishes as a Gaussian function  with the distance between the 
+feature centre and the badge size of 
+the partner (Fig. \ref{fig:model-struc} B, grey line). Formally the 
+reward estimation $\hat{R}$ when the focal individual ($i$) faces 
+individual $j$ is given by,
+\
 \begin{equation}
@@
 
-where $x_z$ is the weight of feature centre $z$ on the reward estimation; and $\theta$ is
-the width of the generalization function. Similarly, the tendency to retreat is given by the sum 
-of feature weights associated with the actor, and the probability is obtained
-by applying a logistic transformation (Fig. \ref{fig:model-struc} B, black line). 
+where $x_z$ is the weight of feature centre $z$ on the reward estimation; 
+and $\theta$ is the width of the generalization function. Similarly, 
+the tendency to retreat is given by the sum of feature weights associated 
+with the actor, and the probability is obtained by applying a logistic 
+transformation (Fig. \ref{fig:model-struc} B, black line). 
 So formally, the log-odds to retreat when facing individual $j$ is
@@
 
-After all their interactions, individuals in the population reproduce proportionally to 
-their fitness $w_i$, which is a sum of the baseline fitness ($w_0$) and all the pay-offs
-obtained throughout their life. Thus, the combination of natural selection and genetic drift
-changes the distribution of values in $\alpha$ and $\beta$ that segregates in the population, effectively changing  the badges expressed and the communication system.
+After all their interactions, individuals in the population 
+reproduce proportionally to their fitness $w_i$, which is a sum of the 
+baseline fitness ($w_0$) and all the pay-offs obtained throughout their 
+life. Thus, the combination of natural selection and genetic drift 
+changes the distribution of values in $\alpha$ and $\beta$ that segregates 
+in the population, effectively changing  the badges expressed and the 
+communication system.
 
@@
 
-extSimsDir<-paste0("e:/BadgeSims/",scenario1,"_")
+extSimsDir<-paste0("M:/BadgeSims/",scenario1,"_")
 listTest<-list.files(extSimsDir,full.names = TRUE)
@@
 
-extSimsDir<-paste0("e:/BadgeSims/",scenario2,"_")
+extSimsDir<-paste0("M:/BadgeSims/",scenario2,"_")
 listTest<-list.files(extSimsDir,full.names = TRUE)
@@
 
-We first present the outcome of simulations where we prevent the evolution of the 
-sender code (by setting the mutation rate to $0$), and assume that all individuals 
-in the population display either an uninformative or informative badge size with respect to their quality. These simulations show what type of receiver strategy develops through a learning process. 
-Figure \ref{fig:model-struc} (C and D) shows the receiver strategy developed 
-through learning; panel C is for receivers that faced uninformative signals, 
-and D informative ones. When learners face uninformative signals, they change their 
-probability of retreating depending on their own quality. Individuals with higher 
-quality (red tones), after the learning process, have a high probability of attacking; while 
-individuals of lower quality (blue tones) mostly retreat from confrontations. Thus, learning splits
-the population of receivers into the two classic pure strategies of hawks and doves. 
-Given that we have a assumed a monomorphic population with unresponsive reaction norms 
-on the signalling side, the changes triggered by learning only affect a small range 
-of badge sizes. In contrast, when receivers face informative reaction norms 
-on the side of the signaller (panel B in fig. \ref{fig:model-struc}), receivers use the 
-badge size of their interacting partners to determine whether to retreat or attack. 
-The relationship is given by a threshold-like reaction norm, where the change from 
-retreating to attacking depends on the quality of the receiver. As expected, 
-the higher the quality of the receiver the larger the badge size that triggers a retreat.
+We first present the outcome of simulations where we prevent the 
+evolution of the sender code (by setting the mutation rate to $0$), 
+and assume that all individuals in the population display either 
+an uninformative or informative badge size with respect to their quality. 
+These simulations show what type of receiver strategy develops through 
+a learning process. Figure \ref{fig:model-struc} (C and D) shows 
+the receiver strategy developed through learning; panel C is for 
+receivers that faced uninformative signals, and D informative ones. 
+When learners face uninformative signals, they change their 
+probability of retreating depending on their own quality. Individuals 
+with higher quality (red tones), after the learning process, have a 
+high probability of attacking; while individuals of lower quality 
+(blue tones) mostly retreat from confrontations. Thus, learning splits
+the population of receivers into the two classic pure strategies of 
+hawks and doves. Given that we have a assumed a monomorphic population 
+with unresponsive reaction norms on the signalling side, the changes 
+triggered by learning only affect a small range of badge sizes. 
+In contrast, when receivers face informative reaction norms 
+on the side of the signaller (panel B in fig. \ref{fig:model-struc}), 
+receivers use the badge size of their interacting partners to determine 
+whether to retreat or attack. The relationship is given by a 
+threshold-like reaction norm, where the change from retreating to 
+attacking depends on the quality of the receiver. As expected, 
+the higher the quality of the receiver the larger the badge size 
+that triggers a retreat.
 
@@
 
-When we allow the evolution of the sender code ($\alpha$ and $\beta$ change subject to natural selection and genetic drift), and the badge size work as a handicap 
-(i.e. the cost of the badge is inversely proportional to the quality of the individual), the sender code evolves to produce an honest signal Fig \ref{fig:handicap}). The evolution of 
-the sender code does not happen immediately after the start of the evolutionary process. Instead, the evolutionary dynamics of the  reaction norm parameters (Fig. \ref{fig:handicap}, X) show a set of 
-steps. First, the intercept of the reaction norm ($\alpha$) evolves to higher values, reducing the 
-average badge size in the populations.
-That makes sense because reaction norms segregating in the population are flat at this point, and consequently the receiver code does not respond to the badge size, bigger badges are costlier and do not trigger lower attack probabilities.
-During those first generations, the slope of the sender reaction norm ($\beta$)  remains around the initial value of zero.
+When we allow the evolution of the sender code ($\alpha$ and $\beta$ 
+change subject to natural selection and genetic drift), and the badge 
+size work as a handicap (i.e. the cost of the badge is inversely 
+proportional to the quality of the individual), the sender code evolves 
+to produce an honest signal Fig \ref{fig:handicap}). The evolution of 
+the sender code does not happen immediately after the start of the 
+evolutionary process. Instead, the evolutionary dynamics of the reaction 
+norm parameters (Fig. \ref{fig:handicap}, X) show a set of 
+steps. First, the intercept of the reaction norm ($\alpha$) evolves to 
+higher values, reducing the average badge size in the populations. 
+That makes sense because reaction norms segregating in the population are 
+flat at this point, and consequently the receiver code does not respond to 
+the badge size, bigger badges are costlier and do not trigger lower attack 
+probabilities. During those first generations, the slope of the sender 
+reaction norm ($\beta$)  remains around the initial value of zero. 
 At around generation 4000, the slope evolves toward positive values. 
-Which implies that larger badges correlate with higher quality.
-Receivers learn to react to such correlation, reducing
-the probability of attack towards individual of larger badges. Hence, natural selection 
-favours larger values of $\beta$, eventually leading to an evolutionary equilibrium where 
-badge size is an honest signal of quality mediated by the learned responses of receivers. The
-evolutionary trajectory portrayed in figure \ref{fig:handicap} is not the only possible trajectory. 
-If the slope of the sender reaction norms ($\beta$) evolves toward negative values before badges become handicaps, receivers never learn to react to the size of badges.
-Thus, badges are only costly and do not provide information about the quality of an individual.
-Eventually, the badge disappears from the population.
+Which implies that larger badges correlate with higher quality. Receivers 
+learn to react to such correlation, reducing the probability of attack towards 
+individual of larger badges. Hence, natural selection favours larger 
+values of $\beta$, eventually leading to an evolutionary equilibrium where 
+badge size is an honest signal of quality mediated by the learned responses 
+of receivers. The evolutionary trajectory portrayed in figure 
+\ref{fig:handicap} is not the only possible trajectory. If the slope of the 
+sender reaction norms ($\beta$) evolves toward negative values before badges 
+become handicaps, receivers never learn to react to the size of badges. 
+Thus, badges are only costly and do not provide information about the 
+quality of an individual. Eventually, the badge disappears from the population.
 
-When production of the badge is not costly (i.e.
-does not work as a handicap), the parameters of the sender code go through evolutionary branching event, triggereing the evolution of different types of
-signals in the population. Fig. \ref{fig:branching}A and B shows changes in the distribution of the 
-sender reaction norm ($\alpha$ and $\beta$) along evolutionary time. Darker areas show trait values 
-with high frequency in the population.
-Populations start monomorphic with a value of zero on both traits, and mutations quickly buid up a normal distribution around the starting value.
-Within the first 2000 generations the unimodal distribution splits into a bimodal one.
-In the case of $\alpha$, one of the peaks splits further, so at the end of the evolutionary simulation the distribution of $\alpha$ value in the population shows three peaks.
-In the case of $\beta$, the two initial peaks coexist for around 4000 generations.
-Later on one of the types goes extinct and the distribution goes back to being unimodal.
-These changes in the distrubtuion of the sender reaction norm parameters imply that during the second half of the simualation individuals can by classify into three distinct types.
-Two types express a flat reaction norm, meaning that their badge size is not informative of their quality, while the third type shows intermediate badge sizes which are determined by the quality of the individual (Fig.
-\ref{fig:branching}).
-Furthermore, the receiver reaction norms develop through learning, particularly those of individuals of intermediate quality, respond to the signal of their sender type by increasing the probability of retreating from a fight with individuals with larger badges (Fig.
-\ref{fig:branching}).
+When production of the badge is not costly (i.e. does not work as a handicap), 
+the parameters of the sender code go through evolutionary branching event, 
+triggering the evolution of different types of signals in the population. 
+Fig. \ref{fig:branching}A and B shows changes in the distribution of the 
+sender reaction norm ($\alpha$ and $\beta$) along evolutionary time. 
+Darker areas show trait values with high frequency in the population. 
+Populations start monomorphic with a value of zero on both traits, 
+and mutations quickly build up a normal distribution around the starting 
+value. Within the first 2000 generations the unimodal distribution splits 
+into a bimodal one. In the case of $\alpha$, one of the peaks splits further, 
+so at the end of the evolutionary simulation the distribution of $\alpha$ 
+value in the population shows three peaks. In the case of $\beta$, the two 
+initial peaks coexist for around 4000 generations. Later on one of the 
+types goes extinct and the distribution goes back to being unimodal. 
+These changes in the distribution of the sender reaction norm parameters 
+imply that during the second half of the simulation individuals can be 
+classified into three distinct types. Two types express a flat reaction 
+norm, meaning that their badge size is not informative of their quality, 
+while the third type shows intermediate badge sizes which are determined 
+by the quality of the individual (Fig. \ref{fig:branching}). Furthermore, 
+the receiver reaction norms develop through learning, particularly those 
+of individuals of intermediate quality, respond to the signal of their 
+sender type by increasing the probability of retreating from a fight with 
+individuals with larger badges (Fig. \ref{fig:branching}).
 

### 2022-03-15 Figure_dyn_AGA.R

@@ -15,7 +15,7 @@ require("jsonlite")
 scenario<-"betCostEvol1"
 
 extSimsDir<-#here("Simulations",paste0(scenario,"_"))
-  paste0("e:/BadgeSims/",scenario,"_")
+  paste0("M:/BadgeSims/",scenario,"_")
 
 
 # Load files -------------------------------------------------------------------
@@ -70,7 +70,7 @@ pop<-do.call(rbind,lapply(indList_runs, fread))
   
   ## Calculate clustering for all generations  -------------------
   
-  runChoi<-5
+  runChoi<-0
   
   # Choose which interaction to visualize
   lastInt<-tail(pop[,unique(nInteract)],2)[1]
@@ -124,7 +124,7 @@ pop<-do.call(rbind,lapply(indList_runs, fread))
   ## Plot mean and IQRs among individuals for each replicate --------------------
   
   
-  cexAxis<-2.5
+  cexAxis<-1.5
   
   # get the trajectories for individual runs
   traitsTrajs<-dcast(evol,time~seed,
@@ -139,15 +139,15 @@ pop<-do.call(rbind,lapply(indList_runs, fread))
   genstoPrint<-round(seq(1,length(unique(evolStats$time)),length.out = 5))[2:5]
   
   
-  png(here("Simulations",paste0(scenario,"_"),
-           paste0("evolDyn",runChoi,"_",nampar,Valpar,".png")),
-      width = 1400,height = 1200)
+  # png(here("Simulations",paste0(scenario,"_"),
+  #          paste0("evolDyn",runChoi,"_",nampar,Valpar,".png")),
+  #     width = 1400,height = 1200)
   
   
   # Evolutionary dynamics for alpha - intercept of the sender --------------------------------------------
   
   par(plt=posPlot(numploty = 3,idploty = 2,numplotx = 2,idplotx = 1,
-                  lowboundx = 8, upboundx = 93),mfrow=c(1,1),
+                  lowboundx = 8, upboundx=93,title = TRUE),mfrow=c(1,1),
       xaxt="s",las=1)
   
   evolDist(indData = pop[seed==runChoi],variable = "alpha",nbins = 20,pal = pal_dist,
@@ -158,9 +158,10 @@ pop<-do.call(rbind,lapply(indList_runs, fread))
[…]

### 2022-07-01 manuscript/manuscript_1.0.Rmd
New plots and edits in the manuscript

@@
 - C. Daniel Cadena*
-- Olof Leimar^[Department of Zoology, Stockholm University, Stockholm, Sweden]
 - Redouan Bshary^[Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland]
@@
 fontsize: 12pt
-header-includes:
-- \usepackage{float}
-- \usepackage{setspace}\doublespacing
-- \linespread{2}
-- \usepackage{lineno}
-- \linenumbers
-- \floatplacement{figure}{H}
+header-includes: |
+  ```{=latex}
+    \usepackage{float}
+    \usepackage{setspace}\doublespacing
+    \usepackage[backref=page]{hyperref}
+    \linespread{2}
+    \usepackage{lineno}
+    \linenumbers
+    \floatplacement{figure}{H}
+   \newcommand{\beginsupplement}{ \setcounter{table}{0}     \renewcommand{\thetable}{S\arabic{table}}\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
+  ```
 keep_tex: yes
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE,cache=TRUE)
+knitr::opts_chunk$set(echo = TRUE,cache=TRUE,warnings=FALSE)
 require(png)
@@
 <!-- should this last part of the paragraph stay ?-->
-An alternative system of communication is one where individuals 
-gather and store information about each individual they interact with. 
-This information should reliably allow each individual to react adaptively 
-to the quality difference with their opponent. This type of 
-assessment strategy is often referred to as individual 
-recognition [@whitfield_Plumage_1986]. Evidently, individual recognition 
-is restricted to  certain interaction structures, because individuals 
-are limited by cognitive abilities in how many individuals 
-they can recognize and track. Thus, individual recognition and 
-badges of status have been assumed to be alternative communication 
-systems fulfilling the same function, reducing the costs of 
-aggression in contests over resources.
+<!-- An alternative system of communication is one where individuals  -->
+<!-- gather and store information about each individual they interact with.  -->
+<!-- This information should reliably allow each individual to react adaptively  -->
+<!-- to the quality difference with their opponent. This type of  -->
+<!-- assessment strategy is often referred to as individual  -->
+<!-- recognition [@whitfield_Plumage_1986]. Evidently, individual recognition  -->
+<!-- is restricted to  certain interaction structures, because individuals  -->
+<!-- are limited by cognitive abilities in how many individuals  -->
+<!-- they can recognize and track. Thus, individual recognition and  -->
+<!-- badges of status have been assumed to be alternative communication  -->
+<!-- systems fulfilling the same function, reducing the costs of  -->
+<!-- aggression in contests over resources.  -->
 
-Furthermore, badges of status and individual recognition are expected 
-to evolve under different interaction structures, badges of status under 
-global interactions and individual recognition
-under local interactions [@dale_Signaling_2001; @sheehan_There_2016].  
-There is no clear cut reason, however, why both systems could not 
-work together, in fact there is some evidence
-that they can [@chaine_Manipulating_2018].
+<!-- Furthermore, badges of status and individual recognition are expected  -->
+<!-- to evolve under different interaction structures, badges of status under  -->
+<!-- global interactions and individual recognition -->
+<!-- under local interactions [@dale_Signaling_2001; @sheehan_There_2016].   -->
+<!-- There is no clear cut reason, however, why both systems could not  -->
+<!-- work together, in fact there is some evidence -->
+<!-- that they can [@chaine_Manipulating_2018]. -->
 
@@
 of aggressive interactions, is the cognitive aspects of the receiver module. 
-Communication systems relying on badges of status are often assumed 
-to be cognitively less demanding than those involving individual 
-recognition [@sheehan_There_2016] because in the latter, 
-individuals need to store identity and associated information 
-about each individual they interact with. In contrast, the response 
-towards badges of status is often 
-thought to be innate [@botero_Evolution_2010; @johnstone_Badges_1993], with 
+Theoretical models often assume that Communication systems relying on badges of 
+status to have reaction norms as their 
+mechanistic underpinning [@Botero_Evolution_2010], with 
 individuals using their own quality and the opponent's badge to determine 
-whether to aggressively engage in contest. An alternative view of 
-badges of status is that individuals learn 
+whether to aggressively engage in contest. Reaction norms 
+allow individuals to respond to the available information without a big
+cognitive burden. That is at least in contrast to other information processing 
+mechanisms like individual recognition. With individual recognition 
+individuals associate cue of their peers to their quality. These association 
+must be learn thoughout individuals lifes, thus it has the usual 
+cognitive requirements of an associative learning process. An alternative view 
+of badges of status is that individuals learn 
 to react to them based on their experiences [@guilford_Receiver_1991], 
@@
 and thus behave adaptively [@staddon_Adaptive_2016]. Associative 
-learning exists in all major vertebrate taxa, and in many invertebrates [@heyes_Simple_2012;@macphail_Brain_1982; @staddon_Adaptive_2016; 
+learning exists in all major vertebrate taxa, and in many invertebrates
+as well [@heyes_Simple_2012;@macphail_Brain_1982; @staddon_Adaptive_2016; 
 @behrens_Associative_2008].
@@
 ecological relevance, associative learning is a flexible cognitive 
-mechanism whose underpinnings show interspecific variation [@enquist_Power_2016; @quinones_Reinforcement_2019], so presumably, it has been modified 
+mechanism whose underpinnings show interspecific variation 
+[@enquist_Power_2016; @quinones_Reinforcement_2019], so presumably, it has been modified 
 by natural selection. Despite all this, associative learning is 
@@
 encompasses a series of computational methods inspired 
-on the psychological and neurological mechanisms of associative learning [@sutton_Reinforcement_2018]. These set of algorithms allow the 
+on the psychological and neurological mechanisms of associative 
+learning [@sutton_Reinforcement_2018]. These set of algorithms allow the 
 implementation of biologically realistic problems, 
-capturing the essence of learning processes [@frankenhuis_Enriching_2018; @quinones_Reinforcement_2019]. 
+capturing the essence of learning processes 
+[@frankenhuis_Enriching_2018; @quinones_Reinforcement_2019]. 
 Furthermore, these algorithms can be embedded in evolutionary simulations to 
@@
 on their quality. Under this simple
-set up, individuals can use the badge both as a signal of quality (BS)
-or as a trait for IR. We use the model to assess under what conditions 
+set up, individuals can use the badge as a signal of quality. 
+We use the model to assess under what conditions 
 of interaction structure we expect different communication signals to evolve.
@@
 of the population vector of size $N$) given by a number between zero and one, 
-which is drawnfrom a truncated normal distribution 
+which is drawn from a truncated normal distribution 
 $N(\mu = 0.5,\sigma)$. An individual with quality $0$ has the lowest RHP, 
@@
 of the offspring. After birth individuals go through a round of 
-viability selection. Individual specific survival probability is given by $s_i=1/(1+e^{-k_1-k_2(Q_i-B_i)})$. Where $k_1$ and $k_2$ are parameter 
+viability selection. Individual specific survival probability is 
+given by $s_i=1/(1+e^{-k_1-k_2(Q_i-B_i)})$. Where $k_1$ and $k_2$ are parameter 
 values determining the shape of the survival function. 
@@
 
-extSimsDir<-paste0("M:/BadgeSims/",scenario1,"_")
-listTest<-list.files(extSimsDir,full.names = TRUE)
+SimsDir<-here("Simulations",paste0(scenario1,"_"))
+listTest<-list.files(SimsDir,full.names = TRUE)
 indList<-grep("ind",listTest,value=TRUE)
@@
 
-extSimsDir<-paste0("M:/BadgeSims/",scenario2,"_")
-listTest<-list.files(extSimsDir,full.names = TRUE)
+SimsDir<-SimsDir<-here("Simulations",paste0(scenario2,"_"))
+listTest<-list.files(SimsDir,full.names = TRUE)
 indList<-grep("ind",listTest,value=TRUE)
@@
 Thus, badges are only costly and do not provide information about the 
-quality of an individual. Eventually, the badge disappears from the population.
+quality of an individual(Fig. \ref{fig:noBadge} . Eventually, the badge disappears from the population.
 
@@
 
+# Supplementary material
+
+```{=tex}
+\beginsupplement
+```
+
+```{r noBadge,echo=FALSE,warning=FALSE,out.width='90%',fig.align='center',fig.show='hold',fig.cap= "The evolution of cheap signals.
+Portrait of the evolutionary dynamics of the sender code with snapsots of both sender and receiver codes just as in fig.
+\\ref{fig:handicap}.
+The middle panels show changes in the distribution of values fir $\\alpha$ and $\\beta$ along evolutionary time.
+Panels above and bellow correspond to snapshots of the sender and receiver codes, respectively, generation time of the snapshots are indicated by the grey lines in the middle panels."}
+
+knitr::include_graphics(here("Simulations","betCostEvol1_","evolDyn5_betCost5.png"))
+
+```
+
 ## References

### 2022-07-01 AccFunc.R

@@ -242,6 +242,9 @@ get_clusters<-function(DT,vars,k.max=5,Bsamples=500,iterMax=100){
         rep(1:ncol(t(as.matrix(listTimeSeed))), 
             each = nrow(t(as.matrix(listTimeSeed)))))
   totClusters<-do.call(rbind,sapply(listTimeSeed, function(x){
+    if(x[1]==1&x[2]==20000){
+      print("STOPPP!!!")
+    }
       print(x)
       error<-FALSE
       error<-tryCatch(expr = 
@@ -266,7 +269,7 @@ get_clusters<-function(DT,vars,k.max=5,Bsamples=500,iterMax=100){
       }
       else{
         print(x)
-        if(x[1]==0&x[2]==6000){
+        if(x[1]==1&x[2]==20000){
           print("STOPPP!!!")
         }
         tryCatch(expr = invisible(capture.output(clus_Nb<-NbClust.AEQP(as.matrix(

### 2022-07-01 Figure_dyn_AGA.R

@@ -12,7 +12,7 @@ require("jsonlite")
 
 # Scenario to be plotted - corresponds to folders where simulations are stored
 
-scenario<-"betCostEvol1"
+scenario<-"betCostEvol2"
 
 extSimsDir<-#here("Simulations",paste0(scenario,"_"))
   paste0("M:/BadgeSims/",scenario,"_")
@@ -20,9 +20,9 @@ extSimsDir<-#here("Simulations",paste0(scenario,"_"))
 
 # Load files -------------------------------------------------------------------
 # Project folder
-# (listTest<-list.files(here("Simulations",paste0(scenario,"_"))))
+(listTest<-list.files(here("Simulations",paste0(scenario,"_")),full.names = TRUE))
 # External sims folder
-(listTest<-list.files(extSimsDir,full.names = TRUE))
+# (listTest<-list.files(extSimsDir,full.names = TRUE))
 
 (evolList<-grep("evolLearn",listTest,value=TRUE))
 (indList<-grep("indLearn",listTest,value=TRUE))
@@ -33,7 +33,7 @@ param<-fromJSON(here("Simulations",paste0(scenario,"_"),paramName[1]))
   # fromJSON(paramName[1])
 
 
-val<-3
+val<-2
 fileId<-val
 
 
@@ -70,7 +70,7 @@ pop<-do.call(rbind,lapply(indList_runs, fread))
   
   ## Calculate clustering for all generations  -------------------
   
-  runChoi<-0
+  runChoi<-1
   
   # Choose which interaction to visualize
   lastInt<-tail(pop[,unique(nInteract)],2)[1]

### 2022-07-01 parTOjson.R

@@ -12,26 +12,26 @@ here()
 (param1<-#fromJSON(paste0("E:/BadgeSims","/betCostNoLearn_","/parameters.json"))
      fromJSON(here("Simulations","betCostEvol1_","parameters.json")))
 
-param<-list(totGen   = 20000,   nRep    = 1, seed = 1,
+param<-list(totGen   = 20000,   nRep    = 1, seed = 10,
             printGen = 1000,   printLearn = 1000,
-            printLearnInt = 200,
+            printLearnInt = 500,
             popSize  = 2000,  baselineFit = 2,   
-            MutSd    = 1,  nInt        = 2000,  init     = c(0,0,1,0),
+            MutSd    = 0.3,  nInt        = 2000,  init     = c(0,0,1,0),
             mutRate  = 0.001,  mutType  = 0,
             sampleSize = 50,   strQual  = 10,
             errorQual = 0,    alphaBad	 = I(c(0)),    
             betaBad	 = I(c(0)),
             alphaRes =-10,betaRes=-20, gammaRes=0,
             alphaCrit  = 0.4,  alphaAct = 0.4,
-            sigSq   	 = 0.005, nCenters = 10,
+            sigSq   	 = 0.01, nCenters = 6,
             initCrit = 0,      initAct=0,   gamma = 0,
             QualStDv   = 0.15, betCost = 0,
             alphCost	 = 3, mutLearn = FALSE,
             nIntGroup  = 2000, 
-            payoff_matrix = c(3,1,0,0.5),
+            payoff_matrix = c(1.5,1,0,0.5),
             shareCost = 0,
             namParam = "betCost",
-            rangParam = I(c(0,2,5)),
+            rangParam = I(c(5)),
             typeAgent = 2, #//0. hawk 1.dove 2. learner 3. evaluator
             folderL=paste(here("Simulations"),"/",sep="")) # comment for debug
 # folderL=paste(here("Simulations","test_"),"/",sep="")) # comment for release
@@ -39,7 +39,7 @@ param<-list(totGen   = 20000,   nRep    = 1, seed = 1,
 
 
 
-apendScenar<-"EvolNorm"
+apendScenar<-"Evol2"
 param$folderL<-paste0(param$folderL,
        param$namParam,apendScenar,"_/")
 # runTime<-"360:00:00"# "10:00:00"# 
@@ -57,7 +57,7 @@ param$folderL<-paste0(param$folderL,
[…]

### 2022-07-20 Presentations/ISBE2022.Rmd
New presentation for ISBE2022

--- 2020-11-25 Presentations/GenesAsEnviron.Rmd
@@
     | (cheap and costly) signals of quality
-author: "Andrés Quiñones PhD."
+author: "Andrés Quiñones"
 bibliography: ../SocialMorphCue.bib
@@
     | 
-    | Genes as environment - AGA2020 virtual symposium
-    | November 17 - 2020
+    | ISBE - Stockholm 2022
+    | August 2 - 2020
 output: 
@@
 
-| ![Red1](../Images/redouan.jpg){width=33%} | ![Olle](../Images/olof-leimar.png){width=23%} | ![Daniel](../Images/danielcadena.jpg){width=32%}  |
-| :---:        |    :----:   |    :---: |
-| *R. Bshary* |*O. Leimar*|*D. Cadena*|
-
-![unine](../Images/UNINE_cmjn.jpg){width=35%}
-![uniandes](../Images/logo_largo.png){width=45%}
+ ![Red1](../Images/redouan.jpg){width=40%} |  ![Daniel](../Images/danielcadena.jpg){width=55%} |
+ :---: | :---:|
+ *R. Bshary* |*D. Cadena*|
+ ![unine](../Images/UNINE_cmjn.jpg){width=30%} | ![uniandes](../Images/logo_largo.png){width=30%} |
 

### 2022-07-20 Presentations/ISBE2022.tex
New presentation for ISBE2022

--- 2020-11-25 Presentations/GenesAsEnviron.tex
@@
 }
+\usepackage{amsmath,amssymb}
 \usepackage{lmodern}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+\usepackage{iftex}
+\ifPDFTeX
   \usepackage[T1]{fontenc}
@@
 \hypersetup{
-  pdfauthor={Andrés Quiñones PhD.},
+  pdfauthor={Andrés Quiñones},
   hidelinks,
@@
 \newif\ifbibliography
-\usepackage{longtable,booktabs}
+\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
 \usepackage{caption}
@@
 \makeatother
-\usepackage{graphicx,grffile}
+\usepackage{graphicx}
 \makeatletter
@@
 \setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newlength{\cslentryspacingunit} % times entry-spacing
+\setlength{\cslentryspacingunit}{\parskip}
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1
+  \let\oldpar\par
+  \def\par{\hangindent=\cslhangindent\oldpar}
+  \fi
+  % set entry spacing
+  \setlength{\parskip}{#2\cslentryspacingunit}
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 \usepackage{float} \usepackage{multicol} \usepackage{amsmath} \usepackage{graphicx} \usepackage{array} \usepackage{xcolor}
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
 
@@
 (cheap and costly) signals of quality}
-\author{Andrés Quiñones PhD.}
-\date{Genes as environment - AGA2020 virtual symposium\\
-November 17 - 2020}
+\author{Andrés Quiñones}
+\date{\hfill\break
+\hfill\break
+ISBE - Stockholm 2022\\
+August 2 - 2020}
 
@@
 \protect\hypertarget{acknowledgements}{}
-
-\begin{longtable}[]{@{}ccc@{}}
+\begin{longtable}[]{@{}
+  >{\centering\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.5000}}
+  >{\centering\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.5000}}@{}}
 \toprule
-\includegraphics[width=0.33\textwidth,height=\textheight]{../Images/redouan.jpg}
-&
-\includegraphics[width=0.23\textwidth,height=\textheight]{../Images/olof-leimar.png}
-&
-\includegraphics[width=0.32\textwidth,height=\textheight]{../Images/danielcadena.jpg}\tabularnewline
+\begin{minipage}[b]{\linewidth}\centering
+\includegraphics[width=0.4\textwidth,height=\textheight]{../Images/redouan.jpg}
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+\includegraphics[width=0.55\textwidth,height=\textheight]{../Images/danielcadena.jpg}
+\end{minipage} \\
 \midrule
 \endhead
-\emph{R. Bshary} & \emph{O. Leimar} & \emph{D. Cadena}\tabularnewline
+\emph{R. Bshary} & \emph{D. Cadena} \\
+\includegraphics[width=0.3\textwidth,height=\textheight]{../Images/UNINE_cmjn.jpg}
+&
+\includegraphics[width=0.3\textwidth,height=\textheight]{../Images/logo_largo.png} \\
 \bottomrule
 \end{longtable}
-
-\includegraphics[width=0.35\textwidth,height=\textheight]{../Images/UNINE_cmjn.jpg}
-\includegraphics[width=0.45\textwidth,height=\textheight]{../Images/logo_largo.png}
-
 \end{frame}
@@
 \protect\hypertarget{social-interactions}{}
-
 \begin{block}{The need to match partners with actions}
-
+\protect\hypertarget{the-need-to-match-partners-with-actions}{}
 \begin{center}\includegraphics[width=0.4\linewidth]{../Images/sexSelec} \end{center}
@@
 \begin{center}\includegraphics[width=0.4\linewidth]{../Images/dominance} \end{center}
-
 \end{block}
-
 \end{frame}
@@
 \protect\hypertarget{the-effect-of-communication}{}
-
 \begin{block}{Badge of status}
-
+\protect\hypertarget{badge-of-status}{}
 \begin{center}\includegraphics[width=0.45\linewidth]{../Images/HouseSparrow_strong} \includegraphics[width=0.45\linewidth]{../Images/HouseSparrow_weak} \end{center}
-
 \end{block}
-
 \end{frame}
@@
 \protect\hypertarget{what-about-signals}{}
-
 \begin{block}{The handicap principle}
-
+\protect\hypertarget{the-handicap-principle}{}
 \begin{center}\includegraphics[height=0.35\textheight]{../Images/Amotz_Zahavi_profile} \includegraphics[height=0.35\textheight]{../Images/sexualSelection} \includegraphics[height=0.35\textheight]{../Images/Grafen} \end{center}
 
-\begin{center}\includegraphics[height=0.35\textheight]{GenesAsEnviron_files/figure-beamer/unnamed-chunk-4-1} \end{center}
+\begin{center}\includegraphics[height=0.35\textheight]{ISBE2022_files/figure-beamer/unnamed-chunk-4-1} \end{center}
 
 (Zahavi 1975; Grafen 1990)
-
 \end{block}
-
 \end{frame}
@@
 \protect\hypertarget{what-about-learning}{}
-
 \begin{center}\includegraphics[height=0.4\textheight]{../Images//pavlov} \end{center}
 
-\begin{center}\includegraphics[height=0.35\textheight]{../Images/learn_cartoon_0} \end{center}
-
+\begin{center}\includegraphics[width=16.76in,height=0.35\textheight]{../Images/learn_cartoon_0} \end{center}
 \end{frame}
@@
 \protect\hypertarget{what-about-learning-1}{}
-
 \begin{center}\includegraphics[height=0.4\textheight]{../Images//pavlov} \end{center}
 
-\begin{center}\includegraphics[height=0.35\textheight]{../Images/learn_cartoon_1} \end{center}
-
+\begin{center}\includegraphics[width=16.76in,height=0.35\textheight]{../Images/learn_cartoon_1} \end{center}
 \end{frame}
@@
 \protect\hypertarget{what-about-learning-2}{}
-
 \begin{center}\includegraphics[height=0.4\textheight]{../Images//pavlov} \end{center}
 
-\begin{center}\includegraphics[height=0.35\textheight]{../Images/learn_cartoon_2} \end{center}
-
+\begin{center}\includegraphics[width=16.76in,height=0.35\textheight]{../Images/learn_cartoon_2} \end{center}
 \end{frame}
@@
 \protect\hypertarget{what-about-learning-3}{}
-
 \begin{center}\includegraphics[height=0.4\textheight]{../Images//pavlov} \end{center}
 
-\begin{center}\includegraphics[height=0.35\textheight]{../Images/learn_cartoon_3} \end{center}
-
+\begin{center}\includegraphics[width=16.76in,height=0.35\textheight]{../Images/learn_cartoon_3} \end{center}
 \end{frame}
@@
 \protect\hypertarget{what-about-learning-4}{}
-
 \begin{center}\includegraphics[height=0.4\textheight]{../Images//pavlov} \end{center}
 
-\begin{center}\includegraphics[height=0.35\textheight]{../Images/learn_cartoon_5} \end{center}
-
+\begin{center}\includegraphics[width=16.76in,height=0.35\textheight]{../Images/learn_cartoon_5} \end{center}
 \end{frame}
@@
 \protect\hypertarget{what-about-learning-5}{}
-
 \begin{center}\includegraphics[height=0.4\textheight]{../Images//pavlov} \end{center}
 
-\begin{center}\includegraphics[height=0.35\textheight]{../Images/learn_cartoon_6} \end{center}
-
+\begin{center}\includegraphics[width=16.76in,height=0.35\textheight]{../Images/learn_cartoon_6} \end{center}
 \end{frame}
@@
 \protect\hypertarget{a-model-of-communication-in-animal-conflict}{}
-
 \vspace{-0.3cm}
@@
 \begin{block}{Conflict}
-
+\protect\hypertarget{conflict}{}
 \vspace{0.3cm}
@@
 
-\begin{center}\includegraphics[width=0.6\linewidth]{GenesAsEnviron_files/figure-beamer/unnamed-chunk-17-1} \end{center}
+\begin{center}\includegraphics[width=0.6\linewidth]{ISBE2022_files/figure-beamer/unnamed-chunk-17-1} \end{center}
 
@@
 \pause
-
 \end{block}
@@
 \begin{block}{Communication}
-
+\protect\hypertarget{communication}{}
 \begin{columns}[T]
@@
 
-\begin{center}\includegraphics[width=0.5\linewidth]{GenesAsEnviron_files/figure-beamer/unnamed-chunk-18-1} \end{center}
+\begin{center}\includegraphics[width=0.5\linewidth]{ISBE2022_files/figure-beamer/unnamed-chunk-18-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=0.6\linewidth]{GenesAsEnviron_files/figure-beamer/unnamed-chunk-19-1} \end{center}
+\begin{center}\includegraphics[width=0.6\linewidth]{ISBE2022_files/figure-beamer/unnamed-chunk-19-1} \end{center}
 
 \end{columns}
-
 \end{block}
-
 \end{frame}
@@
 \protect\hypertarget{learning-under-honest-signalling}{}
-
 \begin{center}\includegraphics[width=1\linewidth]{../Simulations/alphaAct_/alphaAct0.4learnDyn} \end{center}
-
 \end{frame}
@@
 \protect\hypertarget{learning-mediates-the-evolution-of-badges-of-status-handicaps}{}
-
 \begin{center}\includegraphics[width=0.9\linewidth]{../Simulations/betCostEvol1_/evolDyn0_betCost5} \end{center}
-
 \end{frame}
@@
 \protect\hypertarget{learning-can-mediate-the-evolution-of-badges-of-status}{}
-
 \begin{block}{Individual replicates: no badges}
-
-\begin{center}\includegraphics[height=0.9\textheight]{../Simulations/betCostEvol1_/evolDyn5_betCost5} \end{center}
-
+\protect\hypertarget{individual-replicates-no-badges}{}
+\begin{center}\includegraphics[width=19.44in,height=0.9\textheight]{../Simulations/betCostEvol1_/evolDyn5_betCost5} \end{center}
 \end{block}
-
 \end{frame}
@@
 \protect\hypertarget{the-evolution-of-cheap-badges-of-status}{}
-
 \begin{center}\includegraphics[width=0.9\linewidth]{../Simulations/nIntGroupNormQual_/evolDyn7_nIntGroup2000} \end{center}
-
 \end{frame}
@@
 \protect\hypertarget{why-would-you-want-to-be-recognized}{}
-
 \vspace{-1cm}
@@
 \begin{block}{The peaceful, the aggressive and the clever}
-
+\protect\hypertarget{the-peaceful-the-aggressive-and-the-clever}{}
 \begin{itemize}
@@
 \includegraphics[width=0.45\linewidth]{../Simulations/alphaAct_/alphaAct0.4learnDyn}
-\includegraphics[width=0.45\linewidth]{GenesAsEnviron_files/figure-beamer/unnamed-chunk-24-2}
-
+\includegraphics[width=0.45\linewidth]{ISBE2022_files/figure-beamer/unnamed-chunk-24-2}
 \end{block}
-
 \end{frame}
@@
 \protect\hypertarget{why-would-you-want-to-be-recognized-1}{}
-
 \begin{block}{The peaceful, the aggressive and the clever}
-
+\protect\hypertarget{the-peaceful-the-aggressive-and-the-clever-1}{}
 \includegraphics[width=0.9\linewidth]{../Simulations/initAct_/indVarScatter_1}
-
 \end{block}
-
 \end{frame}
@@
 \protect\hypertarget{why-would-you-want-to-be-recognized-2}{}
-
 \begin{block}{The peaceful, the aggressive and the clever}
-
+\protect\hypertarget{the-peaceful-the-aggressive-and-the-clever-2}{}
 \includegraphics[width=0.9\linewidth]{../Simulations/initAct_/indVarScatter_}
-
 \end{block}
-
 \end{frame}
@@
 \protect\hypertarget{take-home-messages}{}
-
 \begin{block}{Costly signals}
-
+\protect\hypertarget{costly-signals}{}
 \begin{itemize}
@@
 \end{itemize}
-
 \end{block}
@@
 \begin{block}{Cheap signals}
-
+\protect\hypertarget{cheap-signals}{}
 \begin{itemize}
@@
 \end{itemize}
-
 \end{block}
-
 \end{frame}
@@
 \protect\hypertarget{questions}{}
-
 \begin{center}\includegraphics[height=0.8\textheight]{../Images/BizarroParrot} \end{center}
-
 \end{frame}
@@
 \protect\hypertarget{references}{}
-
 \hypertarget{refs}{}
-\leavevmode\hypertarget{ref-grafen_Biological_1990}{}%
-Grafen, Alan. 1990. ``Biological Signals as Handicaps.'' \emph{Journal
+\begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-grafen_Biological_1990}{}}%
+Grafen, Alan. 1990. {``Biological Signals as Handicaps.''} \emph{Journal
 of Theoretical Biology} 144 (4): 517--46.
@@
 
-\leavevmode\hypertarget{ref-zahavi_Mate_1975}{}%
-Zahavi, Amotz. 1975. ``Mate Selection-A Selection for a Handicap.''
-\emph{Journal of Theoretical Biology} 53 (1): 205--14.
+\leavevmode\vadjust pre{\hypertarget{ref-zahavi_Mate_1975}{}}%
+Zahavi, Amotz. 1975. {``Mate Selection\textemdash -{A} Selection for a
+Handicap.''} \emph{Journal of Theoretical Biology} 53 (1): 205--14.
 \url{https://doi.org/10.1016/0022-5193(75)90111-3}.
 
+\end{CSLReferences}
 \end{frame}

### 2022-07-20 manuscript/manuscript_1.0.Rmd
New presentation for ISBE2022

@@
 status to have reaction norms as their 
-mechanistic underpinning [@Botero_Evolution_2010], with 
+mechanistic underpinning [@botero_Evolution_2010], with 
 individuals using their own quality and the opponent's badge to determine 

### 2022-07-20 manuscript/manuscript_1.0.tex
New presentation for ISBE2022

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  12pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdfauthor={Andrés E. Quiñones; C. Daniel Cadena*; Redouan Bshary},
  pdfkeywords={My keywords},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
  \usepackage{float}
  \usepackage{setspace}\doublespacing
  \usepackage[backref=page]{hyperref}
  \linespread{2}
  \usepackage{lineno}
  \linenumbers
  \floatplacement{figure}{H}
 \newcommand{\beginsupplement}{ \setcounter{table}{0}     \renewcommand{\thetable}{S\arabic{table}}\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Learning can mediate both badges of status and individual
recognition}
\author{Andrés E. Quiñones\footnote{Laboratorio de Biología Evolutiva de
  Vertebrados, Departamento de Ciencias Biológicas, Universidad de los
  Andes, Bogotá, Colombia} \and C. Daniel Cadena* \and Redouan
Bshary\footnote{Institute of Biology, University of Neuchâtel,
  Neuchâtel, Switzerland}}
\date{July 18, 2022}

\begin{document}
\maketitle
\begin{abstract}
Adaptive behavioural responses often depend on qualities of the
interacting partner of an individual. For example, when competing for
resources, an individual might be better off escalating fights with
individuals of lower quality, while restraining from fighting
individuals of higher quality. Communication systems involving signals
of quality allow individuals to reduce uncertainty regarding the
fighting ability of their partners and make more adaptive behavioral
decisions. However, dishonest individuals can destabilize such
communications systems. One open question is whether cognitive
mechanisms, such as learning, can maintain the honesty of signals, thus,
favoring their evolutionary stability. Here we present evolutionary
simulations, where individuals can produce a signal proportional to
their quality; and learn along their lifetime the best response to the
signal emitted by their peers. In these simulations, learning on the
receiver side can mediate the evolution of signals of quality on the
sender side. When the cost of the signal is proportional to the quality
of the sender populations are only composed of honest signalers. When
the cost is not proportional to the quality of the signaler, the
population is composed of both honest and dishonest signalers. We argue
that learning can be a general cognitive mechanism playing a role in a
wide range of communication systems.
\end{abstract}

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

The outcome of interactions with con-specifics is an important
determinant of fitness in social animals. Irrespective of whether such
interactions are cooperative or competitive, the actions of interacting
individuals partly determine their reproductive success. However, the
best action for an individual might vary depending on its condition, and
that of individuals with whom it interacts. Thus, acting based on
information about interacting partners is typically adaptive (Quiñones
et al. 2016). Acquiring this information, however, is far from trivial.
In some cases, interacting partners (signallers) might be `willing' to
provide accurate information, but in others it might be in their own
interest to conceal information (Johnstone 1997), or to provide
erroneous information (Johnstone and Norris 1993). Typically, in a given
context, some proportion of individuals will benefit from broadcasting
accurate information, whereas others will benefit from hiding it. Take,
for instance, an interaction between two individuals where one can help
the other. Given some costs and benefits, a potential donor will be
interested in helping related individuals. Therefore, for a relative of
the donor, broadcasting kinship is advantageous, whereas for an
unrelated individual, concealing the lack of kinship is better. Similar
scenarios may apply to many aspects of social life such as finding
mates, feeding offspring, dominance relationships and aggressive
interactions (Bradbury and Vehrencamp 2011; Møller 1988; Tibbetts and
Dale 2004). In dominance and aggressive interactions, an important piece
of information to guide individual actions is interacting partners
fighting ability, which is sometimes referred to as Resource Holding
Potential (RHP) (Parker 1974), or simply as quality. Responsiveness to
the quality of the partner is central to communication systems such as
badges of status, where an arbitrary signal conveys quality (Johnstone
and Norris 1993; Rohwer 1975). Here, arbitrary refers to a signal which
is not ontogenetically correlated with quality. Badges of status can be
evolutionarily stable whenever the signal imposes a fitness cost which
decreases with the quality of an individual (Botero et al. 2010;
Johnstone and Norris 1993), such that it's no longer in the interest of
low quality individuals to fake quality by producing a large badge. The
cost of honest signals is a more general principle of communications
systems, usually referred to as the handicap principle (Grafen 1990;
Zahavi 1975).

A somewhat ignored component of communication systems, in the context of
aggressive interactions, is the cognitive aspects of the receiver
module. Theoretical models often assume that Communication systems
relying on badges of status to have reaction norms as their mechanistic
underpinning (Botero et al. 2010), with individuals using their own
quality and the opponent's badge to determine whether to aggressively
engage in contest. Reaction norms allow individuals to respond to the
available information without a big cognitive burden. That is at least
in contrast to other information processing mechanisms like individual
recognition. With individual recognition individuals associate cue of
their peers to their quality. These association must be learn thoughout
individuals lifes, thus it has the usual cognitive requirements of an
associative learning process. An alternative view of badges of status is
that individuals learn to react to them based on their experiences
(Guilford and Dawkins 1991), which would imply that receivers must learn
to associate signals with the fighting ability of bearers just as in
systems based on individual recognition. In cases where badges of status
vary quantitatively (e.g.~in size or intensity), fighting ability may
increase monotonically with attributes of the badge and would be
reinforced by every interaction, so in principle learning would be
faster than in systems involving individual recognition in which the
association between signals and their meaning would vary depending on
the interacting partners. In any case, learning could be a central
cognitive mechanism in both types of communication systems, but the role
of learning in these contexts has not been thoroughly explored in the
empirical nor theoretical literature.

Associative learning is a key cognitive mechanism that allows
individuals to associate rewards with environmental stimuli and thus
behave adaptively (Staddon 2016). Associative learning exists in all
major vertebrate taxa, and in many invertebrates as well (Heyes 2012;
Macphail 1982; Staddon 2016; Behrens et al. 2008). Theory has shown that
natural selection favours these associations in complex environments
where conditions are hard to predict (Dridi and Lehmann 2016). Besides
its wide taxonomic and ecological relevance, associative learning is a
flexible cognitive mechanism whose underpinnings show interspecific
variation (Enquist, Lind, and Ghirlanda 2016; Quiñones et al. 2019), so
presumably, it has been modified by natural selection. Despite all this,
associative learning is not often included in the narrative of
evolutionary explanations of behavioural patterns (Fawcett, Hamblin, and
Giraldeau 2013; Kamil 1983; McAuliffe and Thornton 2015). Computational
models of evolution can play an important role to overcome the lack of
integration between learning and evolution. Reinforcement learning
theory encompasses a series of computational methods inspired on the
psychological and neurological mechanisms of associative learning
(Sutton and Barto 2018). These set of algorithms allow the
implementation of biologically realistic problems, capturing the essence
of learning processes (Frankenhuis, Panchanathan, and Barto 2018;
Quiñones et al. 2019). Furthermore, these algorithms can be embedded in
evolutionary simulations to generate theoretical predictions of the
effect that learning can have in behavioural evolution (Leimar and
McNamara 2019).

In here we present an evolutionary model where individuals use
associative learning to develop a tendency to behave aggressively (or
peacefully), in the context of competition over resources, depending on
the quantitative morphological trait they perceive in their opponent
(badge). Over evolutionary time individuals evolve the size of their
badge, and whether it depends on their quality. Under this simple set
up, individuals can use the badge as a signal of quality. We use the
model to assess under what conditions of interaction structure we expect
different communication signals to evolve.

\hypertarget{the-model}{%
\subsection{The model}\label{the-model}}

We model the evolution of signals of quality in the context of agonistic
interactions. For simplicity we consider a population of haploid
individuals with non-overlapping generations. Individuals are born every
generation with a level of quality (\(Q_i\), where \(i\) is subscript of
the population vector of size \(N\)) given by a number between zero and
one, which is drawn from a truncated normal distribution
\(N(\mu = 0.5,\sigma)\). An individual with quality \(0\) has the lowest
RHP, while an individual with quality \(1\) has the highest. As part of
development processes individuals produce a morphological signal
(badge), the size of such depends on their level of quality according to
a reaction norm given by \(B_i=1/(1+e^{\alpha_i-\beta_i Q_i})\); where
are \(\alpha_i\) and \(\beta_i\) are individual specific traits that
determine the shape of the reaction norm (Fig. \ref{fig:model-struc} A).
Variation in \(\alpha_i\) and \(\beta_i\) means individuals can have
either uninformative (flat) or informative (logistic) norms. The size of
the signal is constrained between zero and one. We assume different
values of these traits are given by different alleles, and are inherited
from mother to offspring. Unless, with a small probability, mutation
changes the allelic value of the offspring. After birth individuals go
through a round of viability selection. Individual specific survival
probability is given by \(s_i=1/(1+e^{-k_1-k_2(Q_i-B_i)})\). Where
\(k_1\) and \(k_2\) are parameter values determining the shape of the
survival function. Importantly, if \(k_2=0\) survival probability is
independent of quality, while if \(k_2>0\) lower quality individuals pay
a higher price for similar-sized badge, fulfilling the core assumption
of the handicap principle (Botero et al. 2010; Grafen 1990; Johnstone
and Norris 1993).

Individuals who survive engage in a series of pairwise interactions
where they compete for resources. Individuals in real populations
typically do not interact at random nor with all individuals in the
population (Kurvers et al. 2014). In order to allow for non-random
interactions, we use the population vector to define an interaction
neighbourhood for the focal individual, the size of the neighbourhood
(\(g\)) is the number of positions around the focal's from which the
partner is drawn. So, if \(g=N\) interactions are global; as \(g\) gets
smaller interactions are more local. In an interaction each individual
must decide whether to escalate a fight or not. Following the classic
\emph{hawk-dove} game (Maynard-Smith 1982), if the focal individual
fights and its partner does not, the focal gets as pay-off the resource
of value \(V\) while its partner gets nothing; if both individuals
restrain from fighting they split up the resource in half; if both
decide to fight they split up the cost of fighting and the winner takes
over the resource. We further assume that the probability of wining a
fight for the focal individual depends on the difference in quality
between her and her interacting partner; specifically it is given by
\(p_{ij}=1/(1+e^{-k_3(Q_i-Q_j)})\), where \(k_3\) is a parameter
defining how strong the quality difference determines the winning
probability; and \(i\) and \(j\) denote the the position in the
population vector of the focal and its interacting partner respectively.

The decision of whether to escalate a fight against an interacting
partner can be dependent on the size of the partner's badge, the actual
dependency is determined by the focal´s experiences through a learning
process. We implement learning using the actor-critic approach from
reinforcement learning theory {[}RL; Sutton and Barto (2018);Quiñones et
al. (2019); Leimar and McNamara (2019){]}. Individuals estimate the
reward (pay-off) expected from interacting with partners of different
badge sizes (the critic in RL terminology). After every interaction they
update the estimate of reward proportionally to the difference between
their current estimate and the observed reward (prediction error
\(\delta\)) and to the speed of learning (\(\mathrm{A}\)). Furthermore,
they express different probabilities of retreating (attacking) depending
on the badge size of their opponent (the actor in RL). They update the
probability of retreating (attacking) up or down depending on whether
retreating (attacking) leads to an increase in the reward estimation.
So, if a focal individual decides to escalate a fight against an
individual with small badge, and that leads to an increase in the reward
estimation, the focal individual will increase the probability of
escalating fights with individuals of small badges in the future. Given
that badge size is a real number between 0 and 1, there are infinitely
many badge sizes. Thus, the reward estimation, as well as the
probability of retreating (escalating), must be generalized across
different values. To implement generalization we use the linear function
approximation method based on radial basis functions (Sutton and Barto
2018). Specifically, we pick \(c\) feature centres, which are evenly
spaced values along the badge size interval ({[}0,1{]}). Each one of
these features centres is associated with a weight for reward estimation
and tendency to play retreat in a given interaction. The reward
estimation and the tendency to retreat are calculated (in every
interaction) as the sum of the weights associated with each feature,
weighted by the response triggered by the feature (Fig.
\ref{fig:model-struc} B, dots represent the feature weights) The
response of each feature centre diminishes as a Gaussian function with
the distance between the feature centre and the badge size of the
partner (Fig. \ref{fig:model-struc} B, grey line). Formally the reward
estimation \(\hat{R}\) when the focal individual (\(i\)) faces
individual \(j\) is given by,\\
\begin{equation}
\hat{R}_{ij} =  \sum_{z=1}^{c} x_z e^{(-\frac{|B_i-c_z|}{2\theta^2})} 
\end{equation}

where \(x_z\) is the weight of feature centre \(z\) on the reward
estimation; and \(\theta\) is the width of the generalization function.
Similarly, the tendency to retreat is given by the sum of feature
weights associated with the actor, and the probability is obtained by
applying a logistic transformation (Fig. \ref{fig:model-struc} B, black
line). So formally, the log-odds to retreat when facing individual \(j\)
is

\begin{equation}
logit(q_{ij}) =  \sum_{z=1}^{c} y_z e^{(-\frac{|B_i-c_z|}{2\theta^2})} 
\end{equation}

where \(y_z\) is the weight of feature centre \(z\) on the tendency to
retreat.

After all their interactions, individuals in the population reproduce
proportionally to their fitness \(w_i\), which is a sum of the baseline
fitness (\(w_0\)) and all the pay-offs obtained throughout their life.
Thus, the combination of natural selection and genetic drift changes the
distribution of values in \(\alpha\) and \(\beta\) that segregates in
the population, effectively changing the badges expressed and the
communication system.

\begin{figure}

{\centering \includegraphics{manuscript_1.0_files/figure-latex/model-struc-1}

}

\caption{Model of communication in the context of aggressive interactions. In A the sender code: a reaction norm determines how the badge size is determined by the quality of the individual. Red shows an informative reaction norm, while the blue shows a uninformative reaction norm. B the receiver code: individuals have a behavioral reaction norm that determines their probability of retreating (black line). The reaction norm arises from generalizing the information from the feature weights (black dots). Generalization is represented by the grey line and its axis, which shows the response triggered by the fourth feature diminishes as the value evaluated is further from the feature center. The learning process moves the feature weights (black dots) up or down depending on whether that leads to an increase in the estimated reward. In C and D, effect of learning on the receiver strategy. Receivers in C face signallers with uninformative reaction norms (blue line in A), while in D they face signallers with informative reaction norms (red line in A). Colour scale in C (applies also for D) indicates the quality of the individual}\label{fig:model-struc}
\end{figure}

\hypertarget{results}{%
\subsection{Results}\label{results}}

We first present the outcome of simulations where we prevent the
evolution of the sender code (by setting the mutation rate to \(0\)),
and assume that all individuals in the population display either an
uninformative or informative badge size with respect to their quality.
These simulations show what type of receiver strategy develops through a
learning process. Figure \ref{fig:model-struc} (C and D) shows the
receiver strategy developed through learning; panel C is for receivers
that faced uninformative signals, and D informative ones. When learners
face uninformative signals, they change their probability of retreating
depending on their own quality. Individuals with higher quality (red
tones), after the learning process, have a high probability of
attacking; while individuals of lower quality (blue tones) mostly
retreat from confrontations. Thus, learning splits the population of
receivers into the two classic pure strategies of hawks and doves. Given
that we have a assumed a monomorphic population with unresponsive
reaction norms on the signalling side, the changes triggered by learning
only affect a small range of badge sizes. In contrast, when receivers
face informative reaction norms on the side of the signaller (panel B in
fig.~\ref{fig:model-struc}), receivers use the badge size of their
interacting partners to determine whether to retreat or attack. The
relationship is given by a threshold-like reaction norm, where the
change from retreating to attacking depends on the quality of the
receiver. As expected, the higher the quality of the receiver the larger
the badge size that triggers a retreat.

\begin{verbatim}
## Warning in knitr::include_graphics(here("Simulations", "betCostEvol1_", :
## It is highly recommended to use relative paths for images. You had
## absolute paths: "M:/Projects/SocialComp_morphCue/Simulations/betCostEvol1_/
## evolDyn0_betCost5.png"
\end{verbatim}

\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{M:/Projects/SocialComp_morphCue/Simulations/betCostEvol1_/evolDyn0_betCost5}

}

\caption{Evolution of the badge size as a handicap mediated by learning. Middle panels show the evolutionary dynamics of the sender code. On the left, changes in the distribution of values for the intercept of the reaction norm ($\alpha$); on the right changes in the distribution of the slope ($\beta$). Dashed lines in both panels show the mean of the distribution. Grey lines show the generation time corresponding to the panels above and below showing the sender and receiver code. In the panels above, the reaction norms correspond to individuals in the population after the interaction round. Color scale indicates quality just as in Fig. \ref{fig:model-struc}. In the panels below, colors represent a cluster clasification performed by the k-means algorithm}\label{fig:handicap}
\end{figure}

When we allow the evolution of the sender code (\(\alpha\) and \(\beta\)
change subject to natural selection and genetic drift), and the badge
size work as a handicap (i.e.~the cost of the badge is inversely
proportional to the quality of the individual), the sender code evolves
to produce an honest signal Fig \ref{fig:handicap}). The evolution of
the sender code does not happen immediately after the start of the
evolutionary process. Instead, the evolutionary dynamics of the reaction
norm parameters (Fig. \ref{fig:handicap}, X) show a set of steps. First,
the intercept of the reaction norm (\(\alpha\)) evolves to higher
values, reducing the average badge size in the populations. That makes
sense because reaction norms segregating in the population are flat at
this point, and consequently the receiver code does not respond to the
badge size, bigger badges are costlier and do not trigger lower attack
probabilities. During those first generations, the slope of the sender
reaction norm (\(\beta\)) remains around the initial value of zero. At
around generation 4000, the slope evolves toward positive values. Which
implies that larger badges correlate with higher quality. Receivers
learn to react to such correlation, reducing the probability of attack
towards individual of larger badges. Hence, natural selection favours
larger values of \(\beta\), eventually leading to an evolutionary
equilibrium where badge size is an honest signal of quality mediated by
the learned responses of receivers. The evolutionary trajectory
portrayed in figure \ref{fig:handicap} is not the only possible
trajectory. If the slope of the sender reaction norms (\(\beta\))
evolves toward negative values before badges become handicaps, receivers
never learn to react to the size of badges. Thus, badges are only costly
and do not provide information about the quality of an individual(Fig.
\ref{fig:noBadge} . Eventually, the badge disappears from the
population.

When production of the badge is not costly (i.e.~does not work as a
handicap), the parameters of the sender code go through evolutionary
branching event, triggering the evolution of different types of signals
in the population. Fig. \ref{fig:branching}A and B shows changes in the
distribution of the sender reaction norm (\(\alpha\) and \(\beta\))
along evolutionary time. Darker areas show trait values with high
frequency in the population. Populations start monomorphic with a value
of zero on both traits, and mutations quickly build up a normal
distribution around the starting value. Within the first 2000
generations the unimodal distribution splits into a bimodal one. In the
case of \(\alpha\), one of the peaks splits further, so at the end of
the evolutionary simulation the distribution of \(\alpha\) value in the
population shows three peaks. In the case of \(\beta\), the two initial
peaks coexist for around 4000 generations. Later on one of the types
goes extinct and the distribution goes back to being unimodal. These
changes in the distribution of the sender reaction norm parameters imply
that during the second half of the simulation individuals can be
classified into three distinct types. Two types express a flat reaction
norm, meaning that their badge size is not informative of their quality,
while the third type shows intermediate badge sizes which are determined
by the quality of the individual (Fig. \ref{fig:branching}).
Furthermore, the receiver reaction norms develop through learning,
particularly those of individuals of intermediate quality, respond to
the signal of their sender type by increasing the probability of
retreating from a fight with individuals with larger badges (Fig.
\ref{fig:branching}).

\begin{verbatim}
## Warning in knitr::include_graphics(here("Simulations", "nIntGroupNormQual_", :
## It is highly recommended to use relative paths for images. You had absolute
## paths: "M:/Projects/SocialComp_morphCue/Simulations/nIntGroupNormQual_/
## evolDyn7_nIntGroup2000.png"
\end{verbatim}

\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{M:/Projects/SocialComp_morphCue/Simulations/nIntGroupNormQual_/evolDyn7_nIntGroup2000}

}

\caption{The evolution of cheap signals. Portrait of the evolutionary dynamics of the sender code with snapsots of both sender and receiver codes just as in fig. \ref{fig:handicap}. The middle panels show changes in the distribution of values fir $\alpha$ and $\beta$ along evolutionary time. Panels above and bellow correspond to snapshots of the sender and receiver codes, respectively, generation time of the snapshots are indicated by the grey lines in the middle panels.}\label{fig:branching}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{../Simulations/initAct_/indVarScatter_}

}

\caption{WHy would you want to be recognized}\label{fig:startCond}
\end{figure}

\hypertarget{supplementary-material}{%
\section{Supplementary material}\label{supplementary-material}}

\beginsupplement

\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{M:/Projects/SocialComp_morphCue/Simulations/betCostEvol1_/evolDyn5_betCost5}

}

\caption{The evolution of cheap signals. Portrait of the evolutionary dynamics of the sender code with snapsots of both sender and receiver codes just as in fig. \ref{fig:handicap}. The middle panels show changes in the distribution of values fir $\alpha$ and $\beta$ along evolutionary time. Panels above and bellow correspond to snapshots of the sender and receiver codes, respectively, generation time of the snapshots are indicated by the grey lines in the middle panels.}\label{fig:noBadge}
\end{figure}

\hypertarget{references}{%
\subsection*{References}\label{references}}
\addcontentsline{toc}{subsection}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-behrens_Associative_2008}{}}%
Behrens, Timothy E. J., Laurence T. Hunt, Mark W. Woolrich, and Matthew
F. S. Rushworth. 2008. {``Associative Learning of Social Value.''}
\emph{Nature} 456 (7219): 245--49.
\url{https://doi.org/10.1038/nature07538}.

\leavevmode\vadjust pre{\hypertarget{ref-botero_Evolution_2010}{}}%
Botero, C. A, I. Pen, J. Komdeur, and F. J Weissing. 2010. {``The
Evolution of Individual Variation in Communication Strategies.''}
\emph{Evolution} 64 (11): 3123--33.

\leavevmode\vadjust pre{\hypertarget{ref-bradbury_Principles_2011}{}}%
Bradbury, Jack W., and Sandra L. Vehrencamp. 2011. \emph{Principles of
{Animal Communication}}. 2nd Edition. {Sunderland, Mass}: {Sinauer
Associates is an imprint of Oxford University Press}.

\leavevmode\vadjust pre{\hypertarget{ref-dridi_Environmental_2016}{}}%
Dridi, Slimane, and Laurent Lehmann. 2016. {``Environmental Complexity
Favors the Evolution of Learning.''} \emph{Behav Ecol} 27 (3): 842--50.
\url{https://doi.org/10.1093/beheco/arv184}.

\leavevmode\vadjust pre{\hypertarget{ref-enquist_Power_2016}{}}%
Enquist, Magnus, Johan Lind, and Stefano Ghirlanda. 2016. {``The Power
of Associative Learning and the Ontogeny of Optimal Behaviour.''}
\emph{Royal Society Open Science} 3 (11): 160734.
\url{https://doi.org/10.1098/rsos.160734}.

\leavevmode\vadjust pre{\hypertarget{ref-fawcett_Exposing_2013}{}}%
Fawcett, Tim W., Steven Hamblin, and Luc-Alain Giraldeau. 2013.
{``Exposing the Behavioral Gambit: The Evolution of Learning and
Decision Rules.''} \emph{Behavioral Ecology} 24 (1): 2--11.
\url{https://doi.org/10.1093/beheco/ars085}.

\leavevmode\vadjust pre{\hypertarget{ref-frankenhuis_Enriching_2018}{}}%
Frankenhuis, Willem E., Karthik Panchanathan, and Andrew G. Barto. 2018.
{``Enriching Behavioral Ecology with Reinforcement Learning Methods.''}
\emph{Behavioural Processes}, February.
\url{https://doi.org/10.1016/j.beproc.2018.01.008}.

\leavevmode\vadjust pre{\hypertarget{ref-grafen_Biological_1990}{}}%
Grafen, Alan. 1990. {``Biological Signals as Handicaps.''} \emph{Journal
of Theoretical Biology} 144 (4): 517--46.
\url{https://doi.org/10.1016/S0022-5193(05)80088-8}.

\leavevmode\vadjust pre{\hypertarget{ref-guilford_Receiver_1991}{}}%
Guilford, Tim, and Marian Stamp Dawkins. 1991. {``Receiver Psychology
and the Evolution of Animal Signals.''} \emph{Animal Behaviour} 42 (1):
1--14. \url{https://doi.org/10.1016/S0003-3472(05)80600-1}.

\leavevmode\vadjust pre{\hypertarget{ref-heyes_Simple_2012}{}}%
Heyes, Cecilia. 2012. {``Simple Minds: A Qualified Defence of
Associative Learning.''} \emph{Phil. Trans. R. Soc. B} 367 (1603):
2695--2703. \url{https://doi.org/10.1098/rstb.2012.0217}.

\leavevmode\vadjust pre{\hypertarget{ref-johnstone_Recognition_1997}{}}%
Johnstone, Rufus A. 1997. {``Recognition and the Evolution of
Distinctive Signatures: When Does It Pay to Reveal Identity?''}
\emph{Proceedings of the Royal Society of London. Series B: Biological
Sciences} 264 (1387): 1547--53.
\url{https://doi.org/10.1098/rspb.1997.0215}.

\leavevmode\vadjust pre{\hypertarget{ref-johnstone_Badges_1993}{}}%
Johnstone, Rufus A., and Ken Norris. 1993. {``Badges of Status and the
Cost of Aggression.''} \emph{Behav Ecol Sociobiol} 32 (2): 127--34.
\url{https://doi.org/10.1007/BF00164045}.

\leavevmode\vadjust pre{\hypertarget{ref-kamil_Optimal_1983}{}}%
Kamil, Alan C. 1983. {``Optimal {Foraging Theory} and the {Psychology}
of {Learning}.''} \emph{Integr Comp Biol} 23 (2): 291--302.
\url{https://doi.org/10.1093/icb/23.2.291}.

\leavevmode\vadjust pre{\hypertarget{ref-kurvers_Evolutionary_2014}{}}%
Kurvers, Ralf H. J. M., Jens Krause, Darren P. Croft, Alexander D. M.
Wilson, and Max Wolf. 2014. {``The Evolutionary and Ecological
Consequences of Animal Social Networks: Emerging Issues.''} \emph{Trends
in Ecology \& Evolution} 29 (6): 326--35.
\url{https://doi.org/10.1016/j.tree.2014.04.002}.

\leavevmode\vadjust pre{\hypertarget{ref-leimar_Learning_2019}{}}%
Leimar, Olof, and John M. McNamara. 2019. {``Learning Leads to Bounded
Rationality and the Evolution of Cognitive Bias in Public Goods
Games.''} \emph{Scientific Reports} 9 (1): 1--9.
\url{https://doi.org/10.1038/s41598-019-52781-7}.

\leavevmode\vadjust pre{\hypertarget{ref-macphail_Brain_1982}{}}%
Macphail, Euan M. 1982. \emph{Brain and {Intelligence} in
{Vertebrates}}. {Clarendon Press}.

\leavevmode\vadjust pre{\hypertarget{ref-maynard-smith_Evolution_1982}{}}%
Maynard-Smith, John. 1982. \emph{Evolution and the {Theory} of {Games}}.
1 edition. {Cambridge ; New York}: {Cambridge University Press}.

\leavevmode\vadjust pre{\hypertarget{ref-mcauliffe_Psychology_2015}{}}%
McAuliffe, K., and A. Thornton. 2015. {``The Psychology of Cooperation
in Animals: An Ecological Approach.''} \emph{J Zool} 295 (1): 23--35.
\url{https://doi.org/10.1111/jzo.12204}.

\leavevmode\vadjust pre{\hypertarget{ref-moller_Female_1988}{}}%
Møller, Anders Pape. 1988. {``Female Choice Selects for Male Sexual Tail
Ornaments in the Monogamous Swallow.''} \emph{Nature} 332 (6165):
640--42. \url{https://doi.org/10.1038/332640a0}.

\leavevmode\vadjust pre{\hypertarget{ref-parker_Assessment_1974}{}}%
Parker, G. A. 1974. {``Assessment Strategy and the Evolution of Fighting
Behaviour.''} \emph{Journal of Theoretical Biology} 47 (1): 223--43.
\url{https://doi.org/10.1016/0022-5193(74)90111-8}.

\leavevmode\vadjust pre{\hypertarget{ref-quinones_Negotiation_2016}{}}%
Quiñones, Andrés E., G. Sander van Doorn, Ido Pen, Franz J. Weissing,
and Michael Taborsky. 2016. {``Negotiation and Appeasement Can Be More
Effective Drivers of Sociality Than Kin Selection.''} \emph{Phil. Trans.
R. Soc. B} 371 (1687): 20150089.
\url{https://doi.org/10.1098/rstb.2015.0089}.

\leavevmode\vadjust pre{\hypertarget{ref-quinones_Reinforcement_2019}{}}%
Quiñones, Andrés E., Olof Leimar, Arnon Lotem, and Redouan Bshary. 2019.
{``Reinforcement {Learning Theory Reveals} the {Cognitive Requirements}
for {Solving} the {Cleaner Fish Market Task}.''} \emph{The American
Naturalist}, December, 000--000. \url{https://doi.org/10.1086/707519}.

\leavevmode\vadjust pre{\hypertarget{ref-rohwer_Social_1975a}{}}%
Rohwer, Sievert. 1975. {``The {Social Significance} of {Avian Winter
Plumage Variability}.''} \emph{Evolution} 29 (4): 593--610.
\url{https://doi.org/10.2307/2407071}.

\leavevmode\vadjust pre{\hypertarget{ref-staddon_Adaptive_2016}{}}%
Staddon, J. E. R. 2016. \emph{Adaptive {Behavior} and {Learning}}. 2
edition. {Cambridge}: {Cambridge University Press}.

\leavevmode\vadjust pre{\hypertarget{ref-sutton_Reinforcement_2018}{}}%
Sutton, Richard S., and Andrew G. Barto. 2018. \emph{Reinforcement
{Learning}: {An Introduction}}. Edited by Francis Bach. Second edition
edition. {Cambridge, MA}: {A Bradford Book}.

\leavevmode\vadjust pre{\hypertarget{ref-tibbetts_Socially_2004}{}}%
Tibbetts, Elizabeth A., and James Dale. 2004. {``A Socially Enforced
Signal of Quality in a Paper Wasp.''} \emph{Nature} 432 (7014): 218--22.
\url{https://doi.org/10.1038/nature02949}.

\leavevmode\vadjust pre{\hypertarget{ref-zahavi_Mate_1975}{}}%
Zahavi, Amotz. 1975. {``Mate Selection\textemdash -{A} Selection for a
Handicap.''} \emph{Journal of Theoretical Biology} 53 (1): 205--14.
\url{https://doi.org/10.1016/0022-5193(75)90111-3}.

\end{CSLReferences}

\end{document}

### 2022-07-20 AccFunc.R

@@ -241,52 +241,52 @@ get_clusters<-function(DT,vars,k.max=5,Bsamples=500,iterMax=100){
   listTimeSeed<-split(t(as.matrix(listTimeSeed)), 
         rep(1:ncol(t(as.matrix(listTimeSeed))), 
             each = nrow(t(as.matrix(listTimeSeed)))))
+  if(!is.list(listTimeSeed)){
+    print("STOPPP!!!")
+  }
   totClusters<-do.call(rbind,sapply(listTimeSeed, function(x){
-    if(x[1]==1&x[2]==20000){
-      print("STOPPP!!!")
+    print(x)
+    error<-FALSE
+    error<-tryCatch(expr = 
+      { clus_gap<-cluster::clusGap(as.matrix(
+        DT[seed==x[1]&time==x[2],.SD,.SDcols=vars]),kmeans,
+        K.max = k.max,B = Bsamples,iter.max=iterMax)
+        nclustersGap<-cluster::maxSE(f = clus_gap$Tab[, "gap"], 
+                                    SE.f = clus_gap$Tab[, "SE.sim"])
+        FALSE
+      },error = function(e) {TRUE}
+    )
+    if(error) {
+      clustersMatch<-cbind(which(DT[,seed] %in% x[1] & DT[,time] %in% x[2]),
+                           rep(1,dim(DT[seed==x[1]&time==x[2],.SD,.SDcols=vars])[1]))
     }
+    else if (nclustersGap==1){
+      clusters<-kmeans(as.matrix(
+      DT[seed==x[1]&time==x[2],.SD,.SDcols=vars]),
+      centers = nclustersGap,iter.max = 100)
+      clustersMatch<-cbind(which(DT[,seed] %in% x[1] & DT[,time] %in% x[2]),
+                         clusters$cluster)
+    }
+    else{
       print(x)
-      error<-FALSE
-      error<-tryCatch(expr = 
-        { clus_gap<-cluster::clusGap(as.matrix(
-          DT[seed==x[1]&time==x[2],.SD,.SDcols=vars]),kmeans,
-          K.max = k.max,B = Bsamples,iter.max=iterMax)
-          nclustersGap<-cluster::maxSE(f = clus_gap$Tab[, "gap"], 
-                                      SE.f = clus_gap$Tab[, "SE.sim"])
-          FALSE
-        },error = function(e) {TRUE}
-      )
-      if(error) {
-        clustersMatch<-cbind(which(DT[,seed] %in% x[1] & DT[,time] %in% x[2]),
[…]

### 2022-07-20 Figure_dyn_AGA.R

@@ -12,7 +12,7 @@ require("jsonlite")
 
 # Scenario to be plotted - corresponds to folders where simulations are stored
 
-scenario<-"betCostEvol2"
+scenario<-"betCostEvol4"
 
 extSimsDir<-#here("Simulations",paste0(scenario,"_"))
   paste0("M:/BadgeSims/",scenario,"_")
@@ -33,7 +33,7 @@ param<-fromJSON(here("Simulations",paste0(scenario,"_"),paramName[1]))
   # fromJSON(paramName[1])
 
 
-val<-2
+val<-1
 fileId<-val
 
 
@@ -54,8 +54,8 @@ pop<-do.call(rbind,lapply(indList_runs, fread))
   # pop<-fread(here("Simulations",paste0(scenario,"_"),indList[fileId]))
   # 
   # External sims folder
-  evol<-fread(evolList[fileId])
-  pop<-fread(indList[fileId])
+  # evol<-fread(evolList[fileId])
+  # pop<-fread(indList[fileId])
 
   # Valpar<-gsub("[[:alpha:]]",gsub(".txt","",tail(strsplit(indList[val],"_")[[1]],1)),
   #              replacement = "")
@@ -70,7 +70,7 @@ pop<-do.call(rbind,lapply(indList_runs, fread))
   
   ## Calculate clustering for all generations  -------------------
   
-  runChoi<-1
+  runChoi<-4
   
   # Choose which interaction to visualize
   lastInt<-tail(pop[,unique(nInteract)],2)[1]

### 2022-07-20 Presentations/ISBE2022.Rmd
new plots for the ppt

@@
 ```{r, echo=FALSE,fig.align='center', out.width='90%'}
-knitr::include_graphics("../Simulations/betCostEvol1_/evolDyn0_betCost5.png")
+# knitr::include_graphics("../Simulations/betCostEvol1_/evolDyn0_betCost5.png")
+knitr::include_graphics("../Simulations/betCostEvol2_/evolDyn18_betCost5.png")
 ```
@@
 ```{r, echo=FALSE,fig.align='center', out.height='90%'}
-knitr::include_graphics("../Simulations/betCostEvol1_/evolDyn5_betCost5.png")
+knitr::include_graphics("../Simulations/betCostEvol2_/evolDyn19_betCost5.png")
 ```

### 2022-07-20 compPar.R

@@ -10,14 +10,15 @@ source(here("AccFunc.R"))
 
 scenario<-"initAct"
 
-extSimsDir<-#here("Simulations",paste0(scenario,"_"))
-  paste0("e:/BadgeSims/",scenario,"_")
+# extSimsDir<-#here("Simulations",paste0(scenario,"_"))
+#   paste0("e:/BadgeSims/",scenario,"_")
 
 # Load files -------------------------------------------------------------------
 
-(listTest<-list.files(here("Simulations",paste0(scenario,"_"))))
+(listTest<-list.files(here("Simulations",paste0(scenario,"_")),full.names = TRUE))
+(listTest<-grep(".txt",listTest,value = TRUE))
 
-(listTest<-list.files(extSimsDir,full.names = TRUE))
+# (listTest<-list.files(extSimsDir,full.names = TRUE))
 
 (sdList<-grep("evol",listTest,value=TRUE))
 (indList<-grep("ind",listTest,value=TRUE))
@@ -211,9 +212,15 @@ lastInt<-tail(inds[,unique(nInteract)],2)[1]
 
 popOneInd<-inds[nInteract==lastInt]
 
-png(here("Simulations",paste0(scenario,"_"),
-         paste0("indVarScatter","_1",".png")),
-    width = 1400,height = 1200)
+# png(here("Simulations",paste0(scenario,"_"),
+#          paste0("indVarScatter","_1",".png")),
+#     width = 1400,height = 1200)
+
+popOneInd$idClust<-get_clusters(popOneInd,vars,k.max = 5,
+                                Bsamples =500,iterMax = 500)
+
+seedsInset<-c()
+
 
 yaxtAll<-c("s","n","n")
 xlabAll<-c("",expression(beta[s]),"")
@@ -228,35 +235,71 @@ for(PAr in popOneInd[,sort(unique(get(shortSce)),decreasing = TRUE)]){
       las=1,new=TRUE,
       yaxt=yaxtAll[match(PAr,popOneInd[,sort(unique(get(shortSce)),decreasing = TRUE)])],
       cex.axis=1.5,new=TRUE)
-  plot(data=popOneInd[(get(shortSce)==PAr&time>max(time)*0.8)&seed %in% 12:15],alpha~beta,ylab="",
-       xlab="", pch=20,cex.lab=3,cex.axis=2,las=1,cex=1,xaxt="s",
-       ylim=range(popOneInd[,alpha])*c(1.1,1.1),
[…]

### 2022-07-20 parTOjson.R

@@ -12,15 +12,17 @@ here()
 (param1<-#fromJSON(paste0("E:/BadgeSims","/betCostNoLearn_","/parameters.json"))
      fromJSON(here("Simulations","betCostEvol1_","parameters.json")))
 
-param<-list(totGen   = 20000,   nRep    = 1, seed = 10,
+traitsTrajs[,.SD,.SDcol=c(paste0(c("meanAlpha_"),runChoi),paste0(c("meanBeta_"),runChoi))]
+
+param<-list(totGen   = 10000,   nRep    = 1, seed = 10,
             printGen = 1000,   printLearn = 1000,
             printLearnInt = 500,
             popSize  = 3000,  baselineFit = 2,   
-            MutSd    = 0.3,  nInt        = 500,  init     = c(0,0,1,0),
+            MutSd    = 0.3,  nInt        = 1500,  init     = c(0,0,1,0),
             mutRate  = 0.001,  mutType  = 0,
             sampleSize = 100,   strQual  = 10,
-            errorQual = 0,    alphaBad	 = I(c(0)),    
-            betaBad	 = I(c(0)),
+            errorQual = 0,    alphaBad	 = I(c(1.8)),    
+            betaBad	 = I(c(5.3)),
             alphaRes =-10,betaRes=-20, gammaRes=0,
             alphaCrit  = 0.4,  alphaAct = 0.4,
             sigSq   	 = 0.01, nCenters = 6,
@@ -31,15 +33,12 @@ param<-list(totGen   = 20000,   nRep    = 1, seed = 10,
             payoff_matrix = c(1.5,1,0,0.5),
             shareCost = 0,
             namParam = "betCost",
-            rangParam = I(c(0)),
+            rangParam = I(c(5)),
             typeAgent = 2, #//0. hawk 1.dove 2. learner 3. evaluator
             folderL=paste(here("Simulations"),"/",sep="")) # comment for debug
 # folderL=paste(here("Simulations","test_"),"/",sep="")) # comment for release
 
-
-
-
-apendScenar<-"Evol5"
+apendScenar<-"Evol6"
 param$folderL<-paste0(param$folderL,
        param$namParam,apendScenar,"_/")
 # runTime<-"360:00:00"# "10:00:00"# 

### 2022-07-25 Presentations/ISBE2022.Rmd
Conference presentation done ISBE2022

@@
 title: | 
-    | Learning can mediate the evolution of 
-    | (cheap and costly) signals of quality
-author: "Andrés Quiñones"
+  | Learning can mediate the evolution of  (cheap and costly) signals of quality
+author: |
+  | Andrés Quiñones
+institute: |
+  | Universitè de Neuchâtel
 bibliography: ../SocialMorphCue.bib
@@
   | ISBE - Stockholm 2022
-    | August 2 - 2020
 output: 
@@
     keep_tex: yes
+classoption: "aspectratio=169"
 header-includes: 
@@
   \usepackage{xcolor}
+  \usepackage{makecell}
 ---
@@
 
-```{r,echo=FALSE,fig.align="center",out.width='40%'}
+```{r,echo=FALSE,fig.align="center",out.width='40%',fig.show='hold'}
 
 knitr::include_graphics("../Images/sexSelec.jpg")
-knitr::include_graphics("../Images/dominance.jpg")
+knitr::include_graphics("../Images/dominance2.jpg")
 ```
@@
 
-```{r,echo=FALSE,fig.align='center',out.width='45%',fig.show='hold'}
-knitr::include_graphics("../Images/HouseSparrow_strong.jpg")
-knitr::include_graphics("../Images/HouseSparrow_weak.jpg")
+```{r,echo=FALSE,fig.align='center',out.width='70%',fig.show='hold'}
+knitr::include_graphics("../Images/houseSparrows.png")
+# knitr::include_graphics("../Images/HouseSparrow_weak.jpg")
 ```
@@
 ```
+\vspace{-0.3cm}
 [@zahavi_Mate_1975;@grafen_Biological_1990]
@@
 
-\vspace{-0.3cm}
+### Conflict: Hawk-Dove game
 
-### Conflict
-
-\vspace{0.3cm}
 \begin{columns}
 \column{0.5\linewidth}
-hawk-dove game
 \tiny
 \begin{center}
-\begin{tabular}{ >{\centering\arraybackslash}p{0.2cm} | >{\centering\arraybackslash}p{2.5cm} | >{\centering\arraybackslash}p{0.8cm} }
-& H & D \\ \hline
-H & $p_w V\frac{-C}{2} + (1-p_w) \frac{-C}{2}$ & $V$ \\ \hline
-V & $0$ & $\frac{V}{2}$\\
+\begin{tabular}{ >{\centering\arraybackslash}p{0.6cm} | >{\centering\arraybackslash}p{1.5cm} | >{\centering\arraybackslash}p{1.5cm} }
+& Hawk & Dove \\ \hline
+Hawk & $ \makecell{p_w V\frac{-C}{2}  +  \\ (1-p_w) \frac{-C}{2}}$ & $V$ \\ \hline
+Dove & $0$ & $\frac{V}{2}$\\
 \end{tabular}
@@
 
-```{r,echo=FALSE,fig.align='center',out.width='60%'}
+```{r,echo=FALSE,fig.align='center',out.width='50%'}
 par(plt=posPlot(),las=1)
@@
 
-```{r, echo=FALSE,fig.align='center', out.width = '60%'}
+```{r, echo=FALSE,fig.align='center', out.width = '50%'}
 
@@
 
-```{r, echo=FALSE,fig.align='center', out.width = '100%'}
+```{r, echo=FALSE,fig.align='center', out.width = '90%'}
 knitr::include_graphics("../Simulations/alphaAct_/alphaAct0.4learnDyn.png")
@@
 
-```{r, echo=FALSE,fig.align='center', out.height='90%'}
+```{r, echo=FALSE,fig.align='center', out.width='90%'}
 knitr::include_graphics("../Simulations/betCostEvol2_/evolDyn19_betCost5.png")
@@
 ```{r,out.width='45%',fig.show='hold'}
-knitr::include_graphics("../Simulations/alphaAct_/alphaAct0.4learnDyn.png")
+knitr::include_graphics("../Simulations/alphaAct_/alphaAct0.4learnDyn.png",)
 # knitr::include_graphics("../Images/cartoonRBF_InitAct.png")
@@
 
-par(plt=posPlot()*0.95)
-matplot(x=rangx,y=as.matrix(actor),type='l',col=c("black","red","blue","springgreen4"),lty=1,
+par(plt=posPlot()*0.92)
+matplot(x=rangx,y=as.matrix(actor),type='l',col=c("black","#E31A1C","dodgerblue2","green4"),lty=1,
      xlab="",ylab="",ylim=c(0,1.1),lwd=5,xaxt="s",cex.lab=1.5,cex.axis=2)
@@
 points(y=logist(weights$neutral,alpha = 0,beta=1),x=centers,cex=3)
-text(y = actor[1,]+0.1,x=rep(0.5,4),col=c("black","red","blue","springgreen4"),
+text(y = actor[1,]+0.1,x=rep(0.5,4),col=c("black","#E31A1C","dodgerblue2","green4"),
      labels = names(weights),cex=2)
@@
 ```{r,echo=FALSE,out.width='90%',fig.show='hold'}
-knitr::include_graphics("../Simulations/initAct_/indVarScatter_1.png")
+knitr::include_graphics("../Simulations/initAct_/indVarScatter0_.png")
 
@@
 ```{r,out.width='90%',fig.show='hold'}
-knitr::include_graphics("../Simulations/initAct_/indVarScatter_.png")
+knitr::include_graphics("../Simulations/initAct_/indVarScatter1_.png")
+
+```
+
+## Why would you want to be recognized?
+
+### The peaceful, the aggressive and the clever
+
+```{r,out.width='90%',fig.show='hold'}
+knitr::include_graphics("../Simulations/initAct_/indVarScatter2_.png")
 
@@
 
-```{r, echo=FALSE,fig.align='center', out.height='80%',fig.show='hold'}
-  knitr::include_graphics("../Images/BizarroParrot.jpg")
-```
+<!-- ```{r, echo=FALSE,fig.align='center', out.height='50%',fig.show='hold'} -->
+<!--   knitr::include_graphics("../Images/BizarroParrot.jpg") -->
+<!--   knitr::include_graphics("../Images/Socialnet1.png") -->
+<!-- ``` -->
+
+\begin{columns}[T]
+ \column{0.7\linewidth}
+  \includegraphics[height=0.8\textwidth]{../Images/BizarroParrot.jpg}
+ \column{0.2\linewidth}
+  \includegraphics[height=1.5\textheight]{../Images//Socialnet.png}
+\end{columns}
 

### 2022-07-25 Presentations/ISBE2022.tex
Conference presentation done ISBE2022

@@
   ignorenonframetext,
-]{beamer}
+  aspectratio=169]{beamer}
 \usepackage{pgfpages}
@@
 \hypersetup{
-  pdfauthor={Andrés Quiñones},
   hidelinks,
@@
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
-\usepackage{float} \usepackage{multicol} \usepackage{amsmath} \usepackage{graphicx} \usepackage{array} \usepackage{xcolor}
+\usepackage{float} \usepackage{multicol} \usepackage{amsmath} \usepackage{graphicx} \usepackage{array} \usepackage{xcolor} \usepackage{makecell}
 \ifLuaTeX
@@
 
-\title{Learning can mediate the evolution of\\
-(cheap and costly) signals of quality}
+\title{Learning can mediate the evolution of (cheap and costly) signals
+of quality}
 \author{Andrés Quiñones}
@@
 \hfill\break
-ISBE - Stockholm 2022\\
-August 2 - 2020}
+ISBE - Stockholm 2022}
+\institute{Universitè de Neuchâtel}
 
@@
 \protect\hypertarget{the-need-to-match-partners-with-actions}{}
-\begin{center}\includegraphics[width=0.4\linewidth]{../Images/sexSelec} \end{center}
-
-\begin{center}\includegraphics[width=0.4\linewidth]{../Images/dominance} \end{center}
+\begin{center}\includegraphics[width=0.4\linewidth]{../Images/sexSelec} \includegraphics[width=0.4\linewidth]{../Images/dominance2} \end{center}
 \end{block}
@@
 \protect\hypertarget{badge-of-status}{}
-\begin{center}\includegraphics[width=0.45\linewidth]{../Images/HouseSparrow_strong} \includegraphics[width=0.45\linewidth]{../Images/HouseSparrow_weak} \end{center}
+\begin{center}\includegraphics[width=0.7\linewidth]{../Images/houseSparrows} \end{center}
 \end{block}
@@
 \begin{center}\includegraphics[height=0.35\textheight]{ISBE2022_files/figure-beamer/unnamed-chunk-4-1} \end{center}
+\vspace{-0.3cm}
 
@@
 \protect\hypertarget{a-model-of-communication-in-animal-conflict}{}
-\vspace{-0.3cm}
-
-\begin{block}{Conflict}
-\protect\hypertarget{conflict}{}
-\vspace{0.3cm}
+\begin{block}{Conflict: Hawk-Dove game}
+\protect\hypertarget{conflict-hawk-dove-game}{}
 \begin{columns}
 \column{0.5\linewidth}
-hawk-dove game
 \tiny
 \begin{center}
-\begin{tabular}{ >{\centering\arraybackslash}p{0.2cm} | >{\centering\arraybackslash}p{2.5cm} | >{\centering\arraybackslash}p{0.8cm} }
-& H & D \\ \hline
-H & $p_w V\frac{-C}{2} + (1-p_w) \frac{-C}{2}$ & $V$ \\ \hline
-V & $0$ & $\frac{V}{2}$\\
+\begin{tabular}{ >{\centering\arraybackslash}p{0.6cm} | >{\centering\arraybackslash}p{1.5cm} | >{\centering\arraybackslash}p{1.5cm} }
+& Hawk & Dove \\ \hline
+Hawk & $ \makecell{p_w V\frac{-C}{2}  +  \\ (1-p_w) \frac{-C}{2}}$ & $V$ \\ \hline
+Dove & $0$ & $\frac{V}{2}$\\
 \end{tabular}
@@
 
-\begin{center}\includegraphics[width=0.6\linewidth]{ISBE2022_files/figure-beamer/unnamed-chunk-17-1} \end{center}
+\begin{center}\includegraphics[width=0.5\linewidth]{ISBE2022_files/figure-beamer/unnamed-chunk-17-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=0.6\linewidth]{ISBE2022_files/figure-beamer/unnamed-chunk-19-1} \end{center}
+\begin{center}\includegraphics[width=0.5\linewidth]{ISBE2022_files/figure-beamer/unnamed-chunk-19-1} \end{center}
 
@@
 \protect\hypertarget{learning-under-honest-signalling}{}
-\begin{center}\includegraphics[width=1\linewidth]{../Simulations/alphaAct_/alphaAct0.4learnDyn} \end{center}
+\begin{center}\includegraphics[width=0.9\linewidth]{../Simulations/alphaAct_/alphaAct0.4learnDyn} \end{center}
 \end{frame}
@@
 \protect\hypertarget{learning-mediates-the-evolution-of-badges-of-status-handicaps}{}
-\begin{center}\includegraphics[width=0.9\linewidth]{../Simulations/betCostEvol1_/evolDyn0_betCost5} \end{center}
+\begin{center}\includegraphics[width=0.9\linewidth]{../Simulations/betCostEvol2_/evolDyn18_betCost5} \end{center}
 \end{frame}
@@
 \protect\hypertarget{individual-replicates-no-badges}{}
-\begin{center}\includegraphics[width=19.44in,height=0.9\textheight]{../Simulations/betCostEvol1_/evolDyn5_betCost5} \end{center}
+\begin{center}\includegraphics[width=0.9\linewidth]{../Simulations/betCostEvol2_/evolDyn19_betCost5} \end{center}
 \end{block}
@@
 \protect\hypertarget{the-peaceful-the-aggressive-and-the-clever-1}{}
-\includegraphics[width=0.9\linewidth]{../Simulations/initAct_/indVarScatter_1}
+\includegraphics[width=0.9\linewidth]{../Simulations/initAct_/indVarScatter0_}
 \end{block}
@@
 \protect\hypertarget{the-peaceful-the-aggressive-and-the-clever-2}{}
-\includegraphics[width=0.9\linewidth]{../Simulations/initAct_/indVarScatter_}
+\includegraphics[width=0.9\linewidth]{../Simulations/initAct_/indVarScatter1_}
+\end{block}
+\end{frame}
+
+\begin{frame}{Why would you want to be recognized?}
+\protect\hypertarget{why-would-you-want-to-be-recognized-3}{}
+\begin{block}{The peaceful, the aggressive and the clever}
+\protect\hypertarget{the-peaceful-the-aggressive-and-the-clever-3}{}
+\includegraphics[width=0.9\linewidth]{../Simulations/initAct_/indVarScatter2_}
 \end{block}
@@
 \protect\hypertarget{questions}{}
-\begin{center}\includegraphics[height=0.8\textheight]{../Images/BizarroParrot} \end{center}
+\begin{columns}[T]
+ \column{0.7\linewidth}
+  \includegraphics[height=0.8\textwidth]{../Images/BizarroParrot.jpg}
+ \column{0.2\linewidth}
+  \includegraphics[height=1.5\textheight]{../Images//Socialnet.png}
+\end{columns}
 \end{frame}

### 2022-07-25 Dyn_reac_learn.R

@@ -12,9 +12,9 @@ require("jsonlite")
 
 # Scenario to be plotted - corresponds to folders where simulations are stored
 
-scenario<-"betCostEvol4"
+scenario<-"betCostEvol5"
 
-extSimsDir<-here("Simulations",paste0(scenario,"_"))
+# extSimsDir<-here("Simulations",paste0(scenario,"_"))
   # paste0("e:/BadgeSims/",scenario,"_")
 
 

### 2022-07-25 Figure_dyn_AGA.R

@@ -12,7 +12,7 @@ require("jsonlite")
 
 # Scenario to be plotted - corresponds to folders where simulations are stored
 
-scenario<-"nIntGroupNormQual"#"betCostEvol2"
+scenario<-"alphaAct"#"nIntGroupNormQual"#
 
 # extSimsDir<-#here("Simulations",paste0(scenario,"_"))
 #   paste0("M:/BadgeSims/",scenario,"_")
@@ -29,7 +29,7 @@ scenario<-"nIntGroupNormQual"#"betCostEvol2"
 paramName<-list.files(here("Simulations",paste0(scenario,"_")))
 # paramName<-list.files(extSimsDir,full.names = TRUE)
 paramName<-grep(".json",paramName,value=TRUE)
-param<-fromJSON(here("Simulations",paste0(scenario,"_"),paramName[1]))
+param<-fromJSON(here("Simulations",paste0(scenario,"_"),paramName[2]))
   # fromJSON(paramName[1])
 
 
@@ -67,11 +67,14 @@ nampar<-param$namParam
 nCenters<-param$nCenters
 sigSquar<-param$sigSq
   
-  
+# Aesthetic paremeters ---------------------------------------------------------
+
+cexAxis<-2.2
+
+
   ## Calculate clustering for all generations  -------------------
   
-runChoi<-19
-  
+runChoi<-7
 # Choose which interaction to visualize
 lastInt<-tail(pop[,unique(nInteract)],4)[1]
 # pop[,max(nInteract),by=.(seed,time)][,min(V1)]
@@ -125,7 +128,7 @@ evolStats<-evol[, as.list(unlist(lapply(.SD, my.summary))),
 ## Plot mean and IQRs among individuals for each replicate --------------------
 
 
-cexAxis<-2.2
+# cexAxis<-1.3
 
 # get the trajectories for individual runs
 traitsTrajs<-dcast(evol,time~seed,
@@ -142,7 +145,7 @@ genstoPrint<-round(seq(1,length(unique(evolStats$time)),length.out = 5))[2:5]
 
 png(here("Simulations",paste0(scenario,"_"),
          paste0("evolDyn",runChoi,"_",nampar,Valpar,".png")),
-    width = 1400,height = 1200)
+    width = 1920,height = 1080)
 
 
 # Evolutionary dynamics for alpha - intercept of the sender --------------------------------------------
[…]

### 2022-08-08 Presentations/ISBE2022.Rmd
FInal version of the ISBE2022 slides

@@
 
-```{r,echo=FALSE,fig.align='center',out.height='35%',fig.show='hold'}
+```{r,echo=FALSE,fig.align='center',out.height='45%',fig.show='hold'}
 knitr::include_graphics("../Images/Amotz_Zahavi_profile.jpg")
@@
 
-```{r,echo=FALSE,fig.align='center',out.height='35%'}
+<!-- # ```{r,echo=FALSE,fig.align='center',out.height='35%'} -->
+<!-- #  -->
+<!-- # badQualRang<-seq(0,1,length=1000) -->
+<!-- # alpha<-0 -->
+<!-- # beta<-6 -->
+<!-- # costFunc<-function(beta,alpha,x) {d<-1/(1+exp(-alpha-beta*x))} -->
+<!-- # par(plt=posPlot()) -->
+<!-- # betRang<-c(5) -->
+<!-- # matplot(y=sapply(betRang,FUN = costFunc,alpha=alpha,x=badQualRang),x=badQualRang, -->
+<!-- #         type="l",cex.lab=1.5, cex.axis=2,lwd=3,lty=1,las=1, -->
+<!-- #      col=c(2,3),ylim=c(0.5,1.1),xaxt="s",yaxt="s",ylab="",xlab="") #,col=i -->
+<!-- # # lines(badgeFunc(qualRange,3,-60,1.05)~qualRange,cex.lab=1.5,cex.axis=2,lwd=3, -->
+<!-- # #       col="blue") -->
+<!-- # # legend("bottomright",legend = betRang,title = expression(beta),col=1:3, -->
+<!-- #        # lty=1) -->
+<!-- # mtext(side=1,text="Quality",line=3.5,cex=3) -->
+<!-- # mtext(side=2,text="Survival",line=3.5,cex=3) -->
+<!-- #  -->
+<!-- # ``` -->
 
-badQualRang<-seq(0,1,length=1000)
-alpha<-0
-beta<-6
-costFunc<-function(beta,alpha,x) {d<-1/(1+exp(-alpha-beta*x))}
-par(plt=posPlot())
-betRang<-c(5)
-matplot(y=sapply(betRang,FUN = costFunc,alpha=alpha,x=badQualRang),x=badQualRang,
-        type="l",cex.lab=1.5, cex.axis=2,lwd=3,lty=1,las=1,
-     col=c(2,3),ylim=c(0.5,1.1),xaxt="s",yaxt="s",ylab="",xlab="") #,col=i
-# lines(badgeFunc(qualRange,3,-60,1.05)~qualRange,cex.lab=1.5,cex.axis=2,lwd=3,
-#       col="blue")
-# legend("bottomright",legend = betRang,title = expression(beta),col=1:3,
-       # lty=1)
-mtext(side=1,text="Quality",line=3.5,cex=3)
-mtext(side=2,text="Survival",line=3.5,cex=3)
-
-```
-\vspace{-0.3cm}
+\vspace{0.6cm}
 [@zahavi_Mate_1975;@grafen_Biological_1990]
@@
 
-## What about learning?
+<!-- ## What about learning? -->
 
-```{r,echo=FALSE,fig.align='center',out.height='40%'}
-knitr::include_graphics("../Images//pavlov.jpg")
-```
+<!-- ```{r,echo=FALSE,fig.align='center',out.height='40%'} -->
+<!-- knitr::include_graphics("../Images//pavlov.jpg") -->
+<!-- ``` -->
 
-```{r,echo=FALSE,fig.align='center',out.height='35%'}
-knitr::include_graphics("../Images/learn_cartoon_2.png")
-```
+<!-- ```{r,echo=FALSE,fig.align='center',out.height='35%'} -->
+<!-- knitr::include_graphics("../Images/learn_cartoon_2.png") -->
+<!-- ``` -->
 
@@
 
-## What about learning?
+<!-- ## What about learning? -->
 
-```{r,echo=FALSE,fig.align='center',out.height='40%'}
-knitr::include_graphics("../Images//pavlov.jpg")
-```
+<!-- ```{r,echo=FALSE,fig.align='center',out.height='40%'} -->
+<!-- knitr::include_graphics("../Images//pavlov.jpg") -->
+<!-- ``` -->
 
-```{r,echo=FALSE,fig.align='center',out.height='35%'}
-knitr::include_graphics("../Images/learn_cartoon_5.png")
-```
+<!-- ```{r,echo=FALSE,fig.align='center',out.height='35%'} -->
+<!-- knitr::include_graphics("../Images/learn_cartoon_5.png") -->
+<!-- ``` -->
 
@@
 
-<!-- \small -->
-<!-- One of the replicates of the previous slide, the badge as a handicap -->
-```{r, echo=FALSE,fig.align='center', out.width='90%'}
-# knitr::include_graphics("../Simulations/betCostEvol1_/evolDyn0_betCost5.png")
-knitr::include_graphics("../Simulations/betCostEvol2_/evolDyn18_betCost5.png")
-```
+\begin{columns}[T]
+
+\column{0.05\linewidth}
+
+\tiny
+\vspace{1cm}
+
+Receiver
+
+\vspace{2cm}
+
+Evolutionary  \\
+Dynamics
+
+\vspace{2cm}
+
+Sender
+
+\column{0.95\linewidth}
+
+\includegraphics{../Simulations/betCostEvol2_/evolDyn18_betCost5_1.png}
+
+\end{columns}
+
+## Learning mediates the evolution of badges of status (handicaps)
+
+\begin{columns}[T]
+
+\column{0.05\linewidth}
+
+\tiny
+\vspace{1cm}
+
+Receiver
+
+\vspace{2cm}
+
+Evolutionary  \\
+Dynamics
+
+\vspace{2cm}
+
+Sender
+
+\column{0.95\linewidth}
+
+\includegraphics{../Simulations/betCostEvol2_/evolDyn18_betCost5.png}
+
+\end{columns}
 
@@
 
-## Learning can mediate the evolution of Badges of Status
-### Individual replicates: no badges
-<!-- \small -->
-<!-- One of the replicates where the badge did not evolve -->
+<!-- ## Learning can mediate the evolution of Badges of Status -->
+<!-- ### Individual replicates: no badges -->
+<!-- <!-- \small --> 
+<!-- <!-- One of the replicates where the badge did not evolve -->
 
-```{r, echo=FALSE,fig.align='center', out.width='90%'}
-knitr::include_graphics("../Simulations/betCostEvol2_/evolDyn19_betCost5.png")
-```
+<!-- \begin{columns}[T] -->
+
+<!-- \column{0.05\linewidth} -->
+
+<!-- \tiny -->
+<!-- \vspace{1cm} -->
+
+<!-- Receiver   -->
+
+<!-- \vspace{2cm} -->
+
+<!-- Evolutionary  \\ -->
+<!-- Dynamics -->
+
+<!-- \vspace{2cm} -->
+
+<!-- Sender -->
+
+<!-- \column{0.95\linewidth} -->
+
+<!-- \includegraphics{../Simulations/betCostEvol2_/evolDyn19_betCost5.png} -->
+
+<!-- \end{columns} -->
 
@@
 
-<!-- \small -->
-<!-- Here simulations where the signal has no cost. -->
+\begin{columns}[T]
 
-<!-- There seems to be some consistency in the three type of sender codes -->
+\column{0.05\linewidth}
 
-```{r, echo=FALSE,fig.align='center', out.width='90%'}
-knitr::include_graphics("../Simulations/nIntGroupNormQual_/evolDyn7_nIntGroup2000.png")
-```
+\tiny
+\vspace{1cm}
+
+Receiver
+
+\vspace{2cm}
+
+Evolutionary  \\
+Dynamics
+
+\vspace{2cm}
+
+Sender
+
+\column{0.95\linewidth}
+
+\includegraphics{../Simulations/nIntGroupNormQual_/evolDyn7_nIntGroup2000.png}
+
+\end{columns}
 
@@
 ```{r,echo=FALSE,out.width='90%',fig.show='hold'}
-knitr::include_graphics("../Simulations/initAct_/indVarScatter0_.png")
+knitr::include_graphics("../Simulations/initAct_/indVarScatter01_.png")
 
@@
 ```{r,out.width='90%',fig.show='hold'}
-knitr::include_graphics("../Simulations/initAct_/indVarScatter1_.png")
+knitr::include_graphics("../Simulations/initAct_/indVarScatter11_.png")
 
@@
 
-## Why would you want to be recognized?
+<!-- ## Why would you want to be recognized? -->
 
-### The peaceful, the aggressive and the clever
+<!-- ### The peaceful, the aggressive and the clever -->
 
-```{r,out.width='90%',fig.show='hold'}
-knitr::include_graphics("../Simulations/initAct_/indVarScatter2_.png")
+<!-- ```{r,out.width='90%',fig.show='hold'} -->
+<!-- knitr::include_graphics("../Simulations/initAct_/indVarScatter2_.png") -->
 
-```
+<!-- ``` -->
 

### 2022-08-08 Presentations/ISBE2022.tex
FInal version of the ISBE2022 slides

@@
 \usepackage{xcolor}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\hypersetup{
-  hidelinks,
-  pdfcreator={LaTeX via pandoc}}
-\urlstyle{same} % disable monospaced font for URLs
 \newif\ifbibliography
@@
 \fi
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same} % disable monospaced font for URLs
+\hypersetup{
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
 
@@
   >{\centering\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.5000}}@{}}
-\toprule
+\toprule()
 \begin{minipage}[b]{\linewidth}\centering
@@
 \end{minipage} \\
-\midrule
+\midrule()
 \endhead
@@
 \includegraphics[width=0.3\textwidth,height=\textheight]{../Images/logo_largo.png} \\
-\bottomrule
+\bottomrule()
 \end{longtable}
@@
 \protect\hypertarget{the-handicap-principle}{}
-\begin{center}\includegraphics[height=0.35\textheight]{../Images/Amotz_Zahavi_profile} \includegraphics[height=0.35\textheight]{../Images/sexualSelection} \includegraphics[height=0.35\textheight]{../Images/Grafen} \end{center}
+\begin{center}\includegraphics[height=0.45\textheight]{../Images/Amotz_Zahavi_profile} \includegraphics[height=0.45\textheight]{../Images/sexualSelection} \includegraphics[height=0.45\textheight]{../Images/Grafen} \end{center}
 
-\begin{center}\includegraphics[height=0.35\textheight]{ISBE2022_files/figure-beamer/unnamed-chunk-4-1} \end{center}
-\vspace{-0.3cm}
+\vspace{0.6cm}
 
@@
 
-\begin{center}\includegraphics[width=16.76in,height=0.35\textheight]{../Images/learn_cartoon_0} \end{center}
+\begin{center}\includegraphics[height=0.35\textheight]{../Images/learn_cartoon_0} \end{center}
 \end{frame}
@@
 
-\begin{center}\includegraphics[width=16.76in,height=0.35\textheight]{../Images/learn_cartoon_1} \end{center}
+\begin{center}\includegraphics[height=0.35\textheight]{../Images/learn_cartoon_1} \end{center}
 \end{frame}
@@
 
-\begin{center}\includegraphics[width=16.76in,height=0.35\textheight]{../Images/learn_cartoon_2} \end{center}
+\begin{center}\includegraphics[height=0.35\textheight]{../Images/learn_cartoon_3} \end{center}
 \end{frame}
@@
 
-\begin{center}\includegraphics[width=16.76in,height=0.35\textheight]{../Images/learn_cartoon_3} \end{center}
-\end{frame}
-
-\begin{frame}{What about learning?}
-\protect\hypertarget{what-about-learning-4}{}
-\begin{center}\includegraphics[height=0.4\textheight]{../Images//pavlov} \end{center}
-
-\begin{center}\includegraphics[width=16.76in,height=0.35\textheight]{../Images/learn_cartoon_5} \end{center}
-\end{frame}
-
-\begin{frame}{What about learning?}
-\protect\hypertarget{what-about-learning-5}{}
-\begin{center}\includegraphics[height=0.4\textheight]{../Images//pavlov} \end{center}
-
-\begin{center}\includegraphics[width=16.76in,height=0.35\textheight]{../Images/learn_cartoon_6} \end{center}
+\begin{center}\includegraphics[height=0.35\textheight]{../Images/learn_cartoon_6} \end{center}
 \end{frame}
@@
 
-\begin{center}\includegraphics[width=0.5\linewidth]{ISBE2022_files/figure-beamer/unnamed-chunk-17-1} \end{center}
+\begin{center}\includegraphics[width=0.5\linewidth]{ISBE2022_files/figure-beamer/unnamed-chunk-12-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=0.5\linewidth]{ISBE2022_files/figure-beamer/unnamed-chunk-18-1} \end{center}
+\begin{center}\includegraphics[width=0.5\linewidth]{ISBE2022_files/figure-beamer/unnamed-chunk-13-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=0.5\linewidth]{ISBE2022_files/figure-beamer/unnamed-chunk-19-1} \end{center}
+\begin{center}\includegraphics[width=0.5\linewidth]{ISBE2022_files/figure-beamer/unnamed-chunk-14-1} \end{center}
 
@@
 \protect\hypertarget{learning-mediates-the-evolution-of-badges-of-status-handicaps}{}
-\begin{center}\includegraphics[width=0.9\linewidth]{../Simulations/betCostEvol2_/evolDyn18_betCost5} \end{center}
+\begin{columns}[T]
+
+\column{0.05\linewidth}
+
+\tiny
+\vspace{1cm}
+
+Receiver
+
+\vspace{2cm}
+
+Evolutionary  \\
+Dynamics
+
+\vspace{2cm}
+
+Sender
+
+\column{0.95\linewidth}
+
+\includegraphics{../Simulations/betCostEvol2_/evolDyn18_betCost5_1.png}
+
+\end{columns}
 \end{frame}
 
-\begin{frame}{Learning can mediate the evolution of Badges of Status}
-\protect\hypertarget{learning-can-mediate-the-evolution-of-badges-of-status}{}
-\begin{block}{Individual replicates: no badges}
-\protect\hypertarget{individual-replicates-no-badges}{}
-\begin{center}\includegraphics[width=0.9\linewidth]{../Simulations/betCostEvol2_/evolDyn19_betCost5} \end{center}
-\end{block}
+\begin{frame}{Learning mediates the evolution of badges of status
+(handicaps)}
+\protect\hypertarget{learning-mediates-the-evolution-of-badges-of-status-handicaps-1}{}
+\begin{columns}[T]
+
+\column{0.05\linewidth}
+
+\tiny
+\vspace{1cm}
+
+Receiver
+
+\vspace{2cm}
+
+Evolutionary  \\
+Dynamics
+
+\vspace{2cm}
+
+Sender
+
+\column{0.95\linewidth}
+
+\includegraphics{../Simulations/betCostEvol2_/evolDyn18_betCost5.png}
+
+\end{columns}
 \end{frame}
@@
 \protect\hypertarget{the-evolution-of-cheap-badges-of-status}{}
-\begin{center}\includegraphics[width=0.9\linewidth]{../Simulations/nIntGroupNormQual_/evolDyn7_nIntGroup2000} \end{center}
+\begin{columns}[T]
+
+\column{0.05\linewidth}
+
+\tiny
+\vspace{1cm}
+
+Receiver
+
+\vspace{2cm}
+
+Evolutionary  \\
+Dynamics
+
+\vspace{2cm}
+
+Sender
+
+\column{0.95\linewidth}
+
+\includegraphics{../Simulations/nIntGroupNormQual_/evolDyn7_nIntGroup2000.png}
+
+\end{columns}
 \end{frame}
@@
 \includegraphics[width=0.45\linewidth]{../Simulations/alphaAct_/alphaAct0.4learnDyn}
-\includegraphics[width=0.45\linewidth]{ISBE2022_files/figure-beamer/unnamed-chunk-24-2}
+\includegraphics[width=0.45\linewidth]{ISBE2022_files/figure-beamer/unnamed-chunk-16-2}
 \end{block}
@@
 \protect\hypertarget{the-peaceful-the-aggressive-and-the-clever-1}{}
-\includegraphics[width=0.9\linewidth]{../Simulations/initAct_/indVarScatter0_}
+\includegraphics[width=0.9\linewidth]{../Simulations/initAct_/indVarScatter01_}
 \end{block}
@@
 \protect\hypertarget{the-peaceful-the-aggressive-and-the-clever-2}{}
-\includegraphics[width=0.9\linewidth]{../Simulations/initAct_/indVarScatter1_}
-\end{block}
-\end{frame}
-
-\begin{frame}{Why would you want to be recognized?}
-\protect\hypertarget{why-would-you-want-to-be-recognized-3}{}
-\begin{block}{The peaceful, the aggressive and the clever}
-\protect\hypertarget{the-peaceful-the-aggressive-and-the-clever-3}{}
-\includegraphics[width=0.9\linewidth]{../Simulations/initAct_/indVarScatter2_}
+\includegraphics[width=0.9\linewidth]{../Simulations/initAct_/indVarScatter11_}
 \end{block}

### 2022-08-08 Figure_dyn_AGA.R

@@ -12,7 +12,7 @@ require("jsonlite")
 
 # Scenario to be plotted - corresponds to folders where simulations are stored
 
-scenario<-"alphaAct"#"nIntGroupNormQual"#
+scenario<-"betCostEvol2"#"alphaAct"#"nIntGroupNormQual"#
 
 # extSimsDir<-#here("Simulations",paste0(scenario,"_"))
 #   paste0("M:/BadgeSims/",scenario,"_")
@@ -33,7 +33,7 @@ param<-fromJSON(here("Simulations",paste0(scenario,"_"),paramName[2]))
   # fromJSON(paramName[1])
 
 
-val<-1
+val<-2
 fileId<-val
 
 
@@ -74,7 +74,7 @@ cexAxis<-2.2
 
   ## Calculate clustering for all generations  -------------------
   
-runChoi<-7
+runChoi<-18
 # Choose which interaction to visualize
 lastInt<-tail(pop[,unique(nInteract)],4)[1]
 # pop[,max(nInteract),by=.(seed,time)][,min(V1)]
@@ -144,7 +144,7 @@ genstoPrint<-round(seq(1,length(unique(evolStats$time)),length.out = 5))[2:5]
 
 
 png(here("Simulations",paste0(scenario,"_"),
-         paste0("evolDyn",runChoi,"_",nampar,Valpar,".png")),
+         paste0("evolDyn",runChoi,"_",nampar,Valpar,"_1.png")),
     width = 1920,height = 1080)
 
 
@@ -169,7 +169,7 @@ legend("topleft",legend = c(expression(alpha[s])),col=colGenesLin[1],lwd=2,
 # grey lines to show the generations shown in the upper and lower panels
 
 matlines(x=matrix(rep(evolStats[genstoPrint,time],each=2),nrow=2),
-         y=matrix(rep(c(-2,3),4),nrow = 2),lty=1,col = "grey",lwd=2)
+         y=matrix(rep(c(-2,3),4),nrow = 2),lty=1,col = "grey",lwd=4)
 
 axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1))-1)],
      labels = axTicks(1)[2:(length(axTicks(1))-1)]/1000,cex.axis=cexAxis)
@@ -210,7 +210,7 @@ legend("topleft",legend = c(expression(beta[s])),cex=2,
[…]

### 2022-08-08 compPar.R

@@ -250,7 +250,7 @@ names(list.DataReact)<-
 
 
 png(here("Simulations",paste0(scenario,"_"),
-         paste0("indVarScatter","1_",".png")),
+         paste0("indVarScatter","11_",".png")),
     width = 1920,height = 1080)
 
 
@@ -262,63 +262,68 @@ colorAll<-c("dodgerblue2","green4","#E31A1C")
 plot.new()
 for(PAr in popOneInd[,sort(unique(get(shortSce)),decreasing = TRUE)]){
   count<-match(PAr,popOneInd[,sort(unique(get(shortSce)),decreasing = TRUE)])
-  par(plt=posPlot(numploty = 2,idploty = 2,
-                  numplotx = length(popOneInd[,unique(get(shortSce))]),
-                  idplotx = count)+c(0,0,-0.05,-0.05),
-      las=1,new=TRUE,
-      yaxt=yaxtAll[match(PAr,popOneInd[,sort(unique(get(shortSce)),decreasing = TRUE)])],
-      cex.axis=1.5,new=TRUE)
-  plot(data=popOneInd[(get(shortSce)==PAr&time>max(time)*0.8)&seed %in% 12:15],
-       alpha~beta,ylab="",
-       xlab="", pch=20,cex.lab=3,cex.axis=2,las=1,cex=2,xaxt="n",
-       ylim=range(popOneInd[,alpha])*c(1.05,1.05),
-       xlim=range(popOneInd[,beta]),
-       # col=alpha(multDiscrPallet[seed+1],0.3))
-       col=multDiscrPallet[seed+1])
-  # points(data=popOneInd[(get(shortSce)==PAr&time>max(time)*0.8)&
-  #                         seed == repsScen[count]],
-  #      alpha~beta,col=multDiscrPallet[seed+1],cex=3,pch=20)
-  lines(x=c(0,0),y=range(popOneInd[,alpha]),col="grey",
-        lwd=2)
-  lines(y=c(0,0),x=range(popOneInd[,beta]),col="grey",
-        lwd=2)
-  mtext(text = ylabAll[count],side = 2,line = 3.5,las=1,cex=3.5)
-  mtext(text = titleAll[count],side = 3,cex = 3,line = 1,
-        col = colorAll[count])
-  
-  par(plt=posPlot(numploty = 2,idploty = 1,
+  # par(plt=posPlot(numploty = 2,idploty = 2,
[…]

### 2023-01-16 manuscript/manuscript_1.0.Rmd
Figures directly included in the Rmardown document. Need to do this only ofr the last figure

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE,cache=TRUE,warnings=FALSE)
+knitr::opts_chunk$set(echo = FALSE,cache=TRUE,warnings=FALSE)
 require(png)
@@
 of aggressive interactions, is the cognitive aspects of the receiver module. 
-Theoretical models often assume that Communication systems relying on badges of 
+Theoretical models often assume that communication systems relying on badges of 
 status to have reaction norms as their 
@@
 which is drawn from a truncated normal distribution 
-$N(\mu = 0.5,\sigma)$. An individual with quality $0$ has the lowest RHP, 
+$N( 0.5,\sigma)$. An individual with quality $0$ has the lowest RHP, 
 while an individual with quality $1$ has the highest. As part of 
@@
 of quality according to a reaction norm given by 
-$B_i=1/(1+e^{\alpha_i-\beta_i Q_i})$; where are $\alpha_i$ and 
+$B_i=1/(1+e^{\alpha_i-\beta_i Q_i})$; where $\alpha_i$ and 
 $\beta_i$ are individual specific traits that determine the shape of 
@@
 are given by different alleles, and are inherited from mother to offspring. 
-Unless, with a small probability, mutation changes the allelic value 
-of the offspring. After birth individuals go through a round of 
+Unless, with a small probability ($\mu$), mutation changes the allelic value 
+of the offspring by an amount taken from a normal distribution $N(0,\sigma_\mu)$.
+After birth individuals go through a round of 
 viability selection. Individual specific survival probability is 
@@
 through a learning process. We implement learning using 
-the actor-critic approach from reinforcement learning theory [RL; @sutton_Reinforcement_2018;@quinones_Reinforcement_2019; @leimar_Learning_2019].
+the actor-critic approach from reinforcement learning (RL) theory [@sutton_Reinforcement_2018;@quinones_Reinforcement_2019; @leimar_Learning_2019].
 Individuals estimate the reward (pay-off) expected from  interacting 
@@
 
-```{r model-struc,echo=FALSE,fig.width=8,fig.align='center',fig.height=8,fig.show='hold',fig.cap= "Model of communication in the context of aggressive interactions.
+```{r model-struc,echo=FALSE,warning=FALSE,fig.width=8,fig.align='center',fig.height=7,fig.show='hold',fig.cap= "Model of communication in the context of aggressive interactions.
 In A the sender code: a reaction norm determines how the badge size is determined by the quality of the individual.
@@
 
+library(dplyr)
 honestBig<-readPNG(here("Images/cart_honestBig.png"))
@@
 par(plt=posPlot(numplotx = 2,idplotx = 1,numploty = 1,idploty = 1,upboundx = 91),
-    las=1,mfrow=c(2,1))
+    las=1,mfrow=c(2,1),mgp=c(2,0.7,0))
   rangPars<-data.frame(honest=c(4,2),dishonest=c(0,0))
@@
             findInterval(tempPop[nInteract==1700,Quality],colorbreaksQual)],
-          lwd=0.5,ylim=c(0,1.1))
+          lwd=3,ylim=c(0,1.1))
 text(x = 0.08,y=1.05,"C",cex = 1.5)
@@
             findInterval(tempPop[nInteract==1700,Quality],colorbreaksQual)],
-          lwd=0.5,ylim=c(0,1.1))
+          lwd=3,ylim=c(0,1.1))
 text(x = 0.08,y=1.05,"D",cex = 1.5)
-text(x = -0.1,y = -0.23,labels = "Badge size")
+text(x = -0.1,y = -0.25,labels = "Badge size")
 
@@
 
-```{r handicap,echo=FALSE,out.width='90%',fig.align='center',fig.show='hold',fig.cap="Evolution of the badge size as a handicap mediated by learning.
+```{r handicap,echo=FALSE,out.width='100%',fig.height=6,fig.align='center',fig.show='hold',fig.cap="Evolution of the badge size as a handicap mediated by learning.
 Middle panels show the evolutionary dynamics of the sender code.
@@
 \\ref{fig:model-struc}.
-In the panels below, colors represent a cluster clasification performed by the k-means algorithm" }
+In the panels below, colors represent a cluster clasification performed by the k-means algorithm" ,warning=FALSE}
 
-knitr::include_graphics(here("Simulations","betCostEvol1_","evolDyn0_betCost5.png"))
+scenario<-"betCostEvol4"#"alphaAct"#"nIntGroupNormQual"#
 
+# Load files -------------------------------------------------------------------
+# Project folder
+listTest<-list.files(here("Simulations",paste0(scenario,"_")),full.names = TRUE)
+evolList<-grep("evolLearn",listTest,value=TRUE)
+indList<-grep("indLearn",listTest,value=TRUE)
+paramName<-list.files(here("Simulations",paste0(scenario,"_")))
+# paramName<-list.files(extSimsDir,full.names = TRUE)
+paramName<-grep(".json",paramName,value=TRUE) %>% grep(pattern = "*5_*",value = TRUE)
+
+param<-fromJSON(here("Simulations",paste0(scenario,"_"),paramName[2]))
+  # fromJSON(paramName[1])
+val<-2
+fileId<-val
+evolList_runs<-grep(paste0(param$namParam,param$rangParam),
+                    evolList,value =TRUE)
+indList_runs<-grep(paste0(param$namParam,param$rangParam),
+                   indList,value =TRUE)
+evol<-do.call(rbind,lapply(evolList_runs,fread))
+pop<-do.call(rbind,lapply(indList_runs, fread))
+
+Valpar<-param$rangParam
+
+nampar<-param$namParam
+
+nCenters<-param$nCenters
+sigSquar<-param$sigSq
+
+# Aesthetic paremeters ---------------------------------------------------------
+
+## Calculate clustering for all generations  -------------------
+
+runChoi<-1
+# Choose which interaction to visualize
+# lastInt<-tail(pop[,unique(nInteract)],4)[2]
+# # pop[,max(nInteract),by=.(seed,time)][,min(V1)]
+# # to get last interaction: tail(pop[,unique(nInteract)],1)
+# 
+# popOneInd<-pop[nInteract==lastInt]
+# vars<-c("alpha","beta","Badge")
+# # ,"Badge",grep("WeightAct",names(popFinal),value = TRUE))
+# popOneInd$idClust<-get_clusters(popOneInd,vars,k.max = 5,
+#                                 Bsamples =500,iterMax = 500)
+# clusSummary<-popOneInd[,.(meanAlph=mean(alpha),meanBet=mean(beta)),
+#                        by=.(idClust,time,seed)]
+# clusSummary[,orderClus:=0]
+# invisible(capture.output(
+#   for(i in 1:dim(clusSummary)[1]){
+#   set(clusSummary,i,6L,match(clusSummary[i,meanAlph],
+#         sort(clusSummary[time==clusSummary[i,time]&
+#                            seed==clusSummary[i,seed]]$meanAlph)))
+# }
+# ))
+# 
+# 
+# popOneInd<-merge(popOneInd,clusSummary,all = TRUE)
+# 
+# fwrite(popOneInd,file = here("Simulations",paste0(scenario,"_"),
+#                              paste0("popOneInd",nampar,Valpar,".txt")))
+
+popOneInd<-fread(here("Simulations",paste0(scenario,"_"),
+                             paste0("popOneInd",nampar,Valpar,".txt")))
+
+# Get stats from the evolutionary simulations ----------------------------------
+# names(evol)
+
+cols<-c("freqGenHawks","freqGenDove",  "freqGenEval",  #"freqGenLearn",
+        "freqFenHawks", "freqFenDoves", "freqHH", "freqHD", "freqDD", "meanCue",
+        "meanAlpha", "meanBeta",
+        #"meanFit", #"meanInitCrit", "sdInitCrit", "meanInitAct", "sdInitAct",
+        "WeightAct_0","WeightCrit_0",
+        "WeightAct_1","WeightCrit_1","WeightAct_2",  "WeightCrit_2",
+        "WeightAct_3",  "WeightCrit_3", "WeightAct_4",  "WeightCrit_4",
+        "WeightAct_5","WeightCrit_5")
+
+my.summary<- function(x) list(mean = mean(x), lowIQR = fivenum(x)[2], 
+                              upIQR = fivenum(x)[4])
+
+evolStats<-evol[, as.list(unlist(lapply(.SD, my.summary))), 
+                .SDcols = cols,by=time]
+
+## Plot mean and IQRs among individuals for each replicate --------------------
+
+# get the trajectories for individual runs
+traitsTrajs<-dcast(evol,time~seed,
+                   value.var = c("meanAlpha","meanBeta",
+                                 #"meanInitCrit",#"meanFit",
+                                 # "meanInitAct","sdInitCrit","sdInitAct",
+                                 "sdAlpha","sdBeta","freqHH",
+                                 "freqHD","freqDD"))
+
+finReps<-evol[time==max(time),seed]
+
+genstoPrint<-round(seq(1,length(unique(evolStats$time)),length.out = 5))[2:5]
+
+# Evolutionary dynamics for alpha - intercept of the sender --------------------------------------------
+cexAxis<-1
+
+par(plt=posPlot(numploty = 3,idploty = 2,numplotx = 2,idplotx = 1,
+                lowboundx = 8, title = TRUE,lowboundy = 10,
+                upboundy = 90),
+    mfrow=c(1,1), xaxt="s",las=1,mgp=c(2,0.5,0))
+
+evolDist(indData = pop[seed==runChoi],variable = "alpha",nbins = 10,pal = pal_dist,
+         nlevels=10,cexAxis = cexAxis,xlab="",ylab = "",xaxt="n",yaxt = "s",
+         ylim = range(pop[seed==runChoi,alpha])*c(2,1.4),xlim=c(0,22000))
+
+mtext(text = "Trait value",cex = cexAxis,line = 1,side = 2,las=0)
+legend("top",legend = c(expression(alpha[s])),col=colGenesLin[1],lwd=2,
+       bty = "n",cex=cexAxis)
+
+axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
+     labels = axTicks(1)[2:(length(axTicks(1)))]/1000,cex.axis=cexAxis,
+     hadj = 1.3,line = -0.5,tick = FALSE)
+
+# grey lines to show the generations shown in the upper and lower panels
+
+matlines(x=matrix(rep(evolStats[genstoPrint,time],each=2),nrow=2),
+         y=matrix(rep(c(-2,3),4),nrow = 2),lty=1,col = "grey",lwd=2)
+
+# Variation among individuals
+
+matlines(x=traitsTrajs[,time],
+         y=traitsTrajs[,.SD,
+                       .SDcol=paste0(c("meanAlpha_"),runChoi)],
+         col=colGenesLin,lty = 2,type="l",lwd=5)
+
+par(plt=posPlot(numploty = 3,idploty = 2,numplotx = 2, idplotx = 2,
+                lowboundx = 8, title = TRUE,lowboundy = 10,
+                upboundy = 90),
+    xaxt="s",las=1,new=TRUE)
+
+evolDist(indData = pop[seed==runChoi],variable = "beta",nbins = 10,pal = pal_dist,
+         nlevels=10,cexAxis = 1.5,xlab="",ylab = "",xaxt="n",yaxt = "n",
+         ylim = range(pop[seed==runChoi,beta])*c(2,1.4),xlim=c(0,22000))
+legend("top",legend = c(expression(beta[s])),cex=cexAxis,
+       col=colGenesLin[2],lwd=2,bty = "n")
+# grey lines to show the generations shown in the upper and lower panels
+matlines(x=matrix(rep(evolStats[genstoPrint,time],each=2),nrow=2),
+         y=matrix(rep(c(-2,6),4),nrow = 2),lty=1,col = "grey",lwd=2)
+
+axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
+     labels = axTicks(1)[2:(length(axTicks(1)))]/1000,cex.axis=cexAxis,
+     hadj = 1.3,line = -0.5,tick = FALSE)
+
+axis(4,cex.axis=cexAxis)
+
+matlines(x=traitsTrajs[,time],
+         y=traitsTrajs[,.SD,
+                       .SDcol=paste0(c("meanBeta_"),runChoi)],
+         col=colGenesLin[2],lty = 2,type="l",lwd=4)
+
+# Choose time range
+gen2plot<-round(seq(1,length(unique(evolStats$time)),length.out = 5))[2:5]
+# Plor the actor
+seqYax<-c("s",rep("n",3))
+# seqYlabUp<-c("Badge",rep("",3))
+seqYlabUp<-c("P(dove)",rep("",3))
+seqYlabDown<-c("Badge",rep("",3))
+seqXlabDown<-c("","Quality","")#paste0("seed: ",runChoi))
+seqXlabUp<-c("","Badge","")
+rangQual<-seq(0,1,length.out = 50)
+interv<-1/(nCenters-1)
+centers<-interv*seq(0,nCenters-1)
+rangx<-seq(0,1,length=50)
+count<-0
+for(genC in genstoPrint){
+  count<-count+1
+  par(plt=posPlot(numplotx = 4,numploty = 3,idplotx = count,idploty = 3,
+                  lowboundx = 8, title = TRUE,lowboundy = 10,
+                  upboundy = 90),
+      xaxt="s",las=1,new=TRUE)
+  weightsAct<-as.double(evol[(time==unique(time)[genC])&seed==runChoi,.SD,
+                             .SDcols=grep("WeightAct",
+                                          names(evol),value = TRUE)])
+  tempPop<-popOneInd[(time==unique(time)[genC]&seed==runChoi),.SD[.N],
+               .SDcol=c(grep("WeightAct",
+                             names(evol),value = TRUE),"Quality","alpha","beta",
+                        "orderClus"),
+               by=indId]
+
+  randInds<-sample(tempPop$indId, length(tempPop$indId), replace = FALSE)
+
+  dataIndAct<-sapply(as.list(tempPop[indId %in%randInds,indId]),
+                     function(x){x=
+                       logist(totRBF(rangx,
+                                     centers,sigSquar,
+                                     as.double(
+                                       tempPop[indId==x,.SD,
+                                               .SDcol=grep("WeightAct",
+                                                           names(tempPop),
+                                                           value = TRUE)
+                                               ])),alpha=0,beta = 1)})
+  matplot(x=rangx,y=dataIndAct,
+          yaxt=seqYax[count],ylab="",xlab="",type="l",cex.lab=cexAxis,
+          lwd=2,xaxt="n",ylim=c(-0.06,1),col = paletteMeans(100)[
+            findInterval(tempPop[,Quality],colorbreaksQual)],lty = 1,
+          cex.axis=cexAxis)
+  axis(side = 3,cex.axis=cexAxis,padj = 0.1)
+  mtext(text = seqYlabUp[count],cex = cexAxis,line = 1.7,side = 2,las=0)
+  mtext(text = seqXlabUp[count],cex = cexAxis,line = 1.7,side = 3 ,las=0)
+  lines(logist(totRBF(rangx,centers,sigSquar,weightsAct),alpha=0,beta=1)~rangx,
+        col=1,lwd=3,lty=2)
+  # text(x=0.5,y=-0.015,labels = paste0("time=",unique(evolStats$time)[genC]/1000),
+  #      cex=2)
+  par(plt=posPlot(numplotx = 4,numploty = 3,idplotx = count,idploty = 1,
+                  lowboundx = 8,lowboundy = 10,
+                  upboundy = 90), las=1,new=TRUE)
+
+  # meansClustmp<-tempPop[,.(alphaMean=mean(alpha),betaMean=mean(beta)),by=orderClus]
+  # reacNormpClust<-sapply(meansClustmp[,orderClus], function(x){
+  #   logist(rangx,alpha = meansClustmp[orderClus==x,alphaMean],
+  #          beta = meansClustmp[orderClus==x,betaMean])})
+  # matplot(x = rangx,y = reacNormpClust,col=colRuns[meansClustmp$orderClus],
+  #          type='l',cex.axis=cexAxis,
+  #          xlab="",ylab="",ylim=c(0,1),lty=1,
+  #          lwd=2,yaxt=seqYax[count],xaxt="s",cex.lab=cexAxis)
+  reacNormInd<-sapply(tempPop[indId %in%randInds,indId], function(x){
+    logist(rangx,alpha = tempPop[indId==x,alpha],
+           beta = tempPop[indId==x,beta])})
+  matplot(x = rangx,y = reacNormInd,
+          type='l',cex.axis=cexAxis,
+          xlab="",ylab="",ylim=c(0,1),lty=1,col=1,
+          lwd=2,yaxt=seqYax[count],xaxt="s",cex.lab=cexAxis)
+  mtext(text = seqYlabDown[count],cex = cexAxis,line = 1.5,side = 2,las=0)
+  mtext(text = seqXlabDown[count],cex = cexAxis,line = 1.5,side = 1,las=0)
+
+  }
+
+rm(list=grep("temp",ls(),value = T))
 ```
@@
 
-knitr::include_graphics(here("Simulations","nIntGroupNormQual_","evolDyn7_nIntGroup2000.png"))
+# knitr::include_graphics(here("Simulations","nIntGroupNormQual_","evolDyn7_nIntGroup2000.png"))
+val<-2
+fileId<-val
+
+paramName<-list.files(here("Simulations",paste0(scenario,"_")))
+paramName<-grep(".json",paramName,value=TRUE) %>% grep(pattern = "*s0_*",value = TRUE)
+
+param<-fromJSON(here("Simulations",paste0(scenario,"_"),paramName[2]))
+
+evolList_runs<-grep(paste0(param$namParam,param$rangParam),
+                    evolList,value =TRUE)
+indList_runs<-grep(paste0(param$namParam,param$rangParam),
+                   indList,value =TRUE)
+
+evol<-do.call(rbind,lapply(evolList_runs,fread))
+pop<-do.call(rbind,lapply(indList_runs, fread))
+
+Valpar<-param$rangParam
+
+nampar<-param$namParam
+
+nCenters<-param$nCenters
+sigSquar<-param$sigSq
+
+# Aesthetic paremeters ---------------------------------------------------------
+
+## Calculate clustering for all generations  -------------------
+
+runChoi<-2
+### Choose which interaction to visualize
+# lastInt<-tail(pop[,unique(nInteract)],4)[2]
+# # pop[,max(nInteract),by=.(seed,time)][,min(V1)]
+# # to get last interaction: tail(pop[,unique(nInteract)],1)
+# 
+# popOneInd<-pop[nInteract==lastInt]
+# vars<-c("alpha","beta","Badge")
+# # ,"Badge",grep("WeightAct",names(popFinal),value = TRUE))
+# popOneInd$idClust<-get_clusters(popOneInd,vars,k.max = 10,
+#                                 Bsamples =500,iterMax = 500)
+# clusSummary<-popOneInd[,.(meanAlph=mean(alpha),meanBet=mean(beta)),
+#                        by=.(idClust,time,seed)]
+# clusSummary[,orderClus:=0]
+# invisible(capture.output(
+#   for(i in 1:dim(clusSummary)[1]){
+#   set(clusSummary,i,6L,match(clusSummary[i,meanAlph],
+#         sort(clusSummary[time==clusSummary[i,time]&
+#                            seed==clusSummary[i,seed]]$meanAlph)))
+# }
+# ))
+# 
+# 
+# popOneInd<-merge(popOneInd,clusSummary,all = TRUE)
+# 
+# 
+# fwrite(popOneInd,file = here("Simulations",paste0(scenario,"_"),
+#                              paste0("popOneInd",nampar,Valpar,".txt")))
+
+popOneInd<-fread(here("Simulations",paste0(scenario,"_"),
+                             paste0("popOneInd",nampar,Valpar,".txt")))
+
+# Get stats from the evolutionary simulations ----------------------------------
+# names(evol)
+
+cols<-c("freqGenHawks","freqGenDove",  "freqGenEval",  #"freqGenLearn",
+        "freqFenHawks", "freqFenDoves", "freqHH", "freqHD", "freqDD", "meanCue",
+        "meanAlpha", "meanBeta",
+        #"meanFit", #"meanInitCrit", "sdInitCrit", "meanInitAct", "sdInitAct",
+        "WeightAct_0","WeightCrit_0",
+        "WeightAct_1","WeightCrit_1","WeightAct_2",  "WeightCrit_2",
+        "WeightAct_3",  "WeightCrit_3", "WeightAct_4",  "WeightCrit_4",
+        "WeightAct_5","WeightCrit_5")
+
+my.summary<- function(x) list(mean = mean(x), lowIQR = fivenum(x)[2], 
+                              upIQR = fivenum(x)[4])
+
+evolStats<-evol[, as.list(unlist(lapply(.SD, my.summary))), 
+                .SDcols = cols,by=time]
+
+## Plot mean and IQRs among individuals for each replicate --------------------
+
+# get the trajectories for individual runs
+traitsTrajs<-dcast(evol,time~seed,
+                   value.var = c("meanAlpha","meanBeta",
+                                 #"meanInitCrit",#"meanFit",
+                                 # "meanInitAct","sdInitCrit","sdInitAct",
+                                 "sdAlpha","sdBeta","freqHH",
+                                 "freqHD","freqDD"))
+
+finReps<-evol[time==max(time),seed]
+
+genstoPrint<-round(seq(1,length(unique(evolStats$time)),length.out = 5))[2:5]
+
+# Evolutionary dynamics for alpha - intercept of the sender --------------------------------------------
+cexAxis<-1
+
+par(plt=posPlot(numploty = 3,idploty = 2,numplotx = 2,idplotx = 1,
+                lowboundx = 8, title = TRUE,lowboundy = 10,
+                upboundy = 90),
+    mfrow=c(1,1), xaxt="s",las=1,mgp=c(2,0.5,0))
+
+evolDist(indData = pop[seed==runChoi],variable = "alpha",nbins = 10,pal = pal_dist,
+         nlevels=10,cexAxis = cexAxis,xlab="",ylab = "",xaxt="n",yaxt = "s",
+         ylim = range(pop[seed==runChoi,alpha])*c(2,1.4),xlim=c(0,22000))
+
+mtext(text = "Trait value",cex = cexAxis,line = 1,side = 2,las=0)
+legend("top",legend = c(expression(alpha[s])),col=colGenesLin[1],lwd=2,
+       bty = "n",cex=cexAxis)
+
+axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
+     labels = axTicks(1)[2:(length(axTicks(1)))]/1000,cex.axis=cexAxis,
+     hadj = 1.3,line = -0.5,tick = FALSE)
+
+# grey lines to show the generations shown in the upper and lower panels
+
+matlines(x=matrix(rep(evolStats[genstoPrint,time],each=2),nrow=2),
+         y=matrix(rep(c(-2,3),4),nrow = 2),lty=1,col = "grey",lwd=2)
+
+# Variation among individuals
+
+matlines(x=traitsTrajs[,time],
+         y=traitsTrajs[,.SD,
+                       .SDcol=paste0(c("meanAlpha_"),runChoi)],
+         col=colGenesLin,lty = 2,type="l",lwd=5)
+
+par(plt=posPlot(numploty = 3,idploty = 2,numplotx = 2, idplotx = 2,
+                lowboundx = 8, title = TRUE,lowboundy = 10,
+                upboundy = 90),
+    xaxt="s",las=1,new=TRUE)
+
+evolDist(indData = pop[seed==runChoi],variable = "beta",nbins = 10,pal = pal_dist,
+         nlevels=10,cexAxis = 1.5,xlab="",ylab = "",xaxt="n",yaxt = "n",
+         ylim = range(pop[seed==runChoi,beta])*c(2,1.4),xlim=c(0,22000))
+legend("top",legend = c(expression(beta[s])),cex=cexAxis,
+       col=colGenesLin[2],lwd=2,bty = "n")
+# grey lines to show the generations shown in the upper and lower panels
+matlines(x=matrix(rep(evolStats[genstoPrint,time],each=2),nrow=2),
+         y=matrix(rep(c(-2,6),4),nrow = 2),lty=1,col = "grey",lwd=2)
+
+axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
+     labels = axTicks(1)[2:(length(axTicks(1)))]/1000,cex.axis=cexAxis,
+     hadj = 1.3,line = -0.5,tick = FALSE)
+
+axis(4,cex.axis=cexAxis)
+
+matlines(x=traitsTrajs[,time],
+         y=traitsTrajs[,.SD,
+                       .SDcol=paste0(c("meanBeta_"),runChoi)],
+         col=colGenesLin[2],lty = 2,type="l",lwd=4)
+
+# Choose time range
+gen2plot<-round(seq(1,length(unique(evolStats$time)),length.out = 5))[2:5]
+# Plor the actor
+seqYax<-c("s",rep("n",3))
+# seqYlabUp<-c("Badge",rep("",3))
+seqYlabUp<-c("P(dove)",rep("",3))
+seqYlabDown<-c("Badge",rep("",3))
+seqXlabDown<-c("","Quality","")#paste0("seed: ",runChoi))
+seqXlabUp<-c("","Badge","")
+rangQual<-seq(0,1,length.out = 50)
+interv<-1/(nCenters-1)
+sizeSamp<-30
+centers<-interv*seq(0,nCenters-1)
+rangx<-seq(0,1,length=50)
+count<-0
+for(genC in genstoPrint){
+
+  count<-count+1
+  par(plt=posPlot(numplotx = 4,numploty = 3,idplotx = count,idploty = 3,
+                  lowboundx = 8, title = TRUE,lowboundy = 10,
+                  upboundy = 90),
+      xaxt="s",las=1,new=TRUE)
+  weightsAct<-as.double(evol[(time==unique(time)[genC])&seed==runChoi,.SD,
+                             .SDcols=grep("WeightAct",
+                                          names(evol),value = TRUE)])
+  tempPop<-popOneInd[(time==unique(time)[genC]&seed==runChoi),.SD[.N],
+               .SDcol=c(grep("WeightAct",
+                             names(evol),value = TRUE),"Quality","alpha","beta",
+                        "orderClus"),
+               by=indId]
+
+  randInds<-sample(tempPop$indId, size = sizeSamp, replace = FALSE)
+
+  dataIndAct<-sapply(as.list(tempPop[indId %in%randInds,indId]),
+                     function(x){x=
+                       logist(totRBF(rangx,
+                                     centers,sigSquar,
+                                     as.double(
+                                       tempPop[indId==x,.SD,
+                                               .SDcol=grep("WeightAct",
+                                                           names(tempPop),
+                                                           value = TRUE)
+                                               ])),alpha=0,beta = 1)})
+  matplot(x=rangx,y=dataIndAct,
+          yaxt=seqYax[count],ylab="",xlab="",type="l",cex.lab=cexAxis,
+          lwd=2,xaxt="n",ylim=c(-0.06,1),col = paletteMeans(100)[
+            findInterval(tempPop[indId %in%randInds,Quality],colorbreaksQual)],
+          lty = 1,cex.axis=cexAxis)
+  axis(side = 3,cex.axis=cexAxis,padj = 0.1)
+  mtext(text = seqYlabUp[count],cex = cexAxis,line = 1.7,side = 2,las=0)
+  mtext(text = seqXlabUp[count],cex = cexAxis,line = 1.5,side = 3 ,las=0)
+  lines(logist(totRBF(rangx,centers,sigSquar,weightsAct),alpha=0,beta=1)~rangx,
+        col=1,lwd=3,lty=2)
+  # text(x=0.5,y=-0.015,labels = paste0("time=",unique(evolStats$time)[genC]/1000),
+  #      cex=2)
+  par(plt=posPlot(numplotx = 4,numploty = 3,idplotx = count,idploty = 1,
+                  lowboundx = 8,lowboundy = 10,
+                  upboundy = 90), las=1,new=TRUE)
+  # dataIndReact<-sapply(as.list(tempPop[,indId]),
+  #                      function(x){x=
+  #                        sapply(rangx, 
+  #                               function(y)
+  #                                 do.call(logist,
+  #                                         as.list(
+  #                                           c(y,as.double(
+  #                                             tempPop[indId==x,.SD,
+  #                                                     .SDcol=c("alpha","beta")])))))})
+  # meansClustmp<-tempPop[,.(alphaMean=mean(alpha),betaMean=mean(beta)),by=orderClus]
+  # reacNormpClust<-sapply(meansClustmp[,orderClus], function(x){
+  #   logist(rangx,alpha = meansClustmp[orderClus==x,alphaMean],
+  #          beta = meansClustmp[orderClus==x,betaMean])})
+  # matplot(x = rangx,y = reacNormpClust,col=colRuns[meansClustmp$orderClus],
+  #          type='l',cex.axis=cexAxis,
+  #          xlab="",ylab="",ylim=c(0,1),lty=1,
+  #          lwd=2,yaxt=seqYax[count],xaxt="s",cex.lab=cexAxis)
+   reacNormInd<-sapply(tempPop[indId %in%randInds,indId], function(x){
+    logist(rangx,alpha = tempPop[indId==x,alpha],
+           beta = tempPop[indId==x,beta])})
+  matplot(x = rangx,y = reacNormInd,
+          type='l',cex.axis=cexAxis,
+          xlab="",ylab="",ylim=c(0,1),lty=1,col=1,
+          lwd=2,yaxt=seqYax[count],xaxt="s",cex.lab=cexAxis)
+  mtext(text = seqYlabDown[count],cex = cexAxis,line = 1.5,side = 2,las=0)
+  mtext(text = seqXlabDown[count],cex = cexAxis,line = 1.5,side = 1,las=0)
+
+  }
+
+# rm(list=grep("temp",ls(),value = T))
 
@@
 
-knitr::include_graphics(here("Simulations","betCostEvol1_","evolDyn5_betCost5.png"))
+# knitr::include_graphics(here("Simulations","betCostEvol1_","evolDyn5_betCost5.png"))
+val<-2
+fileId<-val
+
+paramName<-list.files(here("Simulations",paste0(scenario,"_")))
+paramName<-grep(".json",paramName,value=TRUE) %>% grep(pattern = "*s0_*",value = TRUE)
+
+param<-fromJSON(here("Simulations",paste0(scenario,"_"),paramName[2]))
+
+evolList_runs<-grep(paste0(param$namParam,param$rangParam),
+                    evolList,value =TRUE)
+indList_runs<-grep(paste0(param$namParam,param$rangParam),
+                   indList,value =TRUE)
+
+evol<-do.call(rbind,lapply(evolList_runs,fread))
+pop<-do.call(rbind,lapply(indList_runs, fread))
+
+Valpar<-param$rangParam
+
+nampar<-param$namParam
+
+nCenters<-param$nCenters
+sigSquar<-param$sigSq
+
+# Aesthetic paremeters ---------------------------------------------------------
+
+## Calculate clustering for all generations  -------------------
+
+runChoi<-3
+### Choose which interaction to visualize
+# lastInt<-tail(pop[,unique(nInteract)],4)[2]
+# # pop[,max(nInteract),by=.(seed,time)][,min(V1)]
+# # to get last interaction: tail(pop[,unique(nInteract)],1)
+# 
+# popOneInd<-pop[nInteract==lastInt]
+# vars<-c("alpha","beta","Badge")
+# # ,"Badge",grep("WeightAct",names(popFinal),value = TRUE))
+# popOneInd$idClust<-get_clusters(popOneInd,vars,k.max = 10,
+#                                 Bsamples =500,iterMax = 500)
+# clusSummary<-popOneInd[,.(meanAlph=mean(alpha),meanBet=mean(beta)),
+#                        by=.(idClust,time,seed)]
+# clusSummary[,orderClus:=0]
+# invisible(capture.output(
+#   for(i in 1:dim(clusSummary)[1]){
+#   set(clusSummary,i,6L,match(clusSummary[i,meanAlph],
+#         sort(clusSummary[time==clusSummary[i,time]&
+#                            seed==clusSummary[i,seed]]$meanAlph)))
+# }
+# ))
+# 
+# 
+# popOneInd<-merge(popOneInd,clusSummary,all = TRUE)
+# 
+# 
+# fwrite(popOneInd,file = here("Simulations",paste0(scenario,"_"),
+#                              paste0("popOneInd",nampar,Valpar,".txt")))
+
+popOneInd<-fread(here("Simulations",paste0(scenario,"_"),
+                             paste0("popOneInd",nampar,Valpar,".txt")))
+
+# Get stats from the evolutionary simulations ----------------------------------
+# names(evol)
+
+cols<-c("freqGenHawks","freqGenDove",  "freqGenEval",  #"freqGenLearn",
+        "freqFenHawks", "freqFenDoves", "freqHH", "freqHD", "freqDD", "meanCue",
+        "meanAlpha", "meanBeta",
+        #"meanFit", #"meanInitCrit", "sdInitCrit", "meanInitAct", "sdInitAct",
+        "WeightAct_0","WeightCrit_0",
+        "WeightAct_1","WeightCrit_1","WeightAct_2",  "WeightCrit_2",
+        "WeightAct_3",  "WeightCrit_3", "WeightAct_4",  "WeightCrit_4",
+        "WeightAct_5","WeightCrit_5")
+
+my.summary<- function(x) list(mean = mean(x), lowIQR = fivenum(x)[2], 
+                              upIQR = fivenum(x)[4])
+
+evolStats<-evol[, as.list(unlist(lapply(.SD, my.summary))), 
+                .SDcols = cols,by=time]
+
+## Plot mean and IQRs among individuals for each replicate --------------------
+
+# get the trajectories for individual runs
+traitsTrajs<-dcast(evol,time~seed,
+                   value.var = c("meanAlpha","meanBeta",
+                                 #"meanInitCrit",#"meanFit",
+                                 # "meanInitAct","sdInitCrit","sdInitAct",
+                                 "sdAlpha","sdBeta","freqHH",
+                                 "freqHD","freqDD"))
+
+finReps<-evol[time==max(time),seed]
+
+genstoPrint<-round(seq(1,length(unique(evolStats$time)),length.out = 5))[2:5]
+
+# Evolutionary dynamics for alpha - intercept of the sender --------------------------------------------
+cexAxis<-1
+
+par(plt=posPlot(numploty = 3,idploty = 2,numplotx = 2,idplotx = 1,
+                lowboundx = 8, title = TRUE,lowboundy = 10,
+                upboundy = 90),
+    mfrow=c(1,1), xaxt="s",las=1,mgp=c(2,0.5,0))
+
+evolDist(indData = pop[seed==runChoi],variable = "alpha",nbins = 10,pal = pal_dist,
+         nlevels=10,cexAxis = cexAxis,xlab="",ylab = "",xaxt="n",yaxt = "s",
+         ylim = range(pop[seed==runChoi,alpha])*c(2,1.4),xlim=c(0,22000))
+
+mtext(text = "Trait value",cex = cexAxis,line = 1,side = 2,las=0)
+legend("top",legend = c(expression(alpha[s])),col=colGenesLin[1],lwd=2,
+       bty = "n",cex=cexAxis)
+
+axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
+     labels = axTicks(1)[2:(length(axTicks(1)))]/1000,cex.axis=cexAxis,
+     hadj = 1.3,line = -0.5,tick = FALSE)
+
+# grey lines to show the generations shown in the upper and lower panels
+
+matlines(x=matrix(rep(evolStats[genstoPrint,time],each=2),nrow=2),
+         y=matrix(rep(c(-2,3),4),nrow = 2),lty=1,col = "grey",lwd=2)
+
+# Variation among individuals
+
+matlines(x=traitsTrajs[,time],
+         y=traitsTrajs[,.SD,
+                       .SDcol=paste0(c("meanAlpha_"),runChoi)],
+         col=colGenesLin,lty = 2,type="l",lwd=5)
+
+par(plt=posPlot(numploty = 3,idploty = 2,numplotx = 2, idplotx = 2,
+                lowboundx = 8, title = TRUE,lowboundy = 10,
+                upboundy = 90),
+    xaxt="s",las=1,new=TRUE)
+
+evolDist(indData = pop[seed==runChoi],variable = "beta",nbins = 10,pal = pal_dist,
+         nlevels=10,cexAxis = 1.5,xlab="",ylab = "",xaxt="n",yaxt = "n",
+         ylim = range(pop[seed==runChoi,beta])*c(2,1.4),xlim=c(0,22000))
+legend("top",legend = c(expression(beta[s])),cex=cexAxis,
+       col=colGenesLin[2],lwd=2,bty = "n")
+# grey lines to show the generations shown in the upper and lower panels
+matlines(x=matrix(rep(evolStats[genstoPrint,time],each=2),nrow=2),
+         y=matrix(rep(c(-2,6),4),nrow = 2),lty=1,col = "grey",lwd=2)
+
+axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
+     labels = axTicks(1)[2:(length(axTicks(1)))]/1000,cex.axis=cexAxis,
+     hadj = 1.3,line = -0.5,tick = FALSE)
+
+axis(4,cex.axis=cexAxis)
+
+matlines(x=traitsTrajs[,time],
+         y=traitsTrajs[,.SD,
+                       .SDcol=paste0(c("meanBeta_"),runChoi)],
+         col=colGenesLin[2],lty = 2,type="l",lwd=4)
+
+# Choose time range
+gen2plot<-round(seq(1,length(unique(evolStats$time)),length.out = 5))[2:5]
+# Plor the actor
+seqYax<-c("s",rep("n",3))
+# seqYlabUp<-c("Badge",rep("",3))
+seqYlabUp<-c("P(dove)",rep("",3))
+seqYlabDown<-c("Badge",rep("",3))
+seqXlabDown<-c("","Quality","")#paste0("seed: ",runChoi))
+seqXlabUp<-c("","Badge","")
+rangQual<-seq(0,1,length.out = 50)
+interv<-1/(nCenters-1)
+sizeSamp<-30
+centers<-interv*seq(0,nCenters-1)
+rangx<-seq(0,1,length=50)
+count<-0
+for(genC in genstoPrint){
+
+  count<-count+1
+  par(plt=posPlot(numplotx = 4,numploty = 3,idplotx = count,idploty = 3,
+                  lowboundx = 8, title = TRUE,lowboundy = 10,
+                  upboundy = 90),
+      xaxt="s",las=1,new=TRUE)
+  weightsAct<-as.double(evol[(time==unique(time)[genC])&seed==runChoi,.SD,
+                             .SDcols=grep("WeightAct",
+                                          names(evol),value = TRUE)])
+  tempPop<-popOneInd[(time==unique(time)[genC]&seed==runChoi),.SD[.N],
+               .SDcol=c(grep("WeightAct",
+                             names(evol),value = TRUE),"Quality","alpha","beta",
+                        "orderClus"),
+               by=indId]
+
+  randInds<-sample(tempPop$indId, size = sizeSamp, replace = FALSE)
+
+  dataIndAct<-sapply(as.list(tempPop[indId %in%randInds,indId]),
+                     function(x){x=
+                       logist(totRBF(rangx,
+                                     centers,sigSquar,
+                                     as.double(
+                                       tempPop[indId==x,.SD,
+                                               .SDcol=grep("WeightAct",
+                                                           names(tempPop),
+                                                           value = TRUE)
+                                               ])),alpha=0,beta = 1)})
+  matplot(x=rangx,y=dataIndAct,
+          yaxt=seqYax[count],ylab="",xlab="",type="l",cex.lab=cexAxis,
+          lwd=2,xaxt="n",ylim=c(-0.06,1),col = paletteMeans(100)[
+            findInterval(tempPop[indId %in%randInds,Quality],colorbreaksQual)],
+          lty = 1,cex.axis=cexAxis)
+  axis(side = 3,cex.axis=cexAxis,padj = 0.1)
+  mtext(text = seqYlabUp[count],cex = cexAxis,line = 1.7,side = 2,las=0)
+  mtext(text = seqXlabUp[count],cex = cexAxis,line = 1.5,side = 3 ,las=0)
+  lines(logist(totRBF(rangx,centers,sigSquar,weightsAct),alpha=0,beta=1)~rangx,
+        col=1,lwd=3,lty=2)
+  # text(x=0.5,y=-0.015,labels = paste0("time=",unique(evolStats$time)[genC]/1000),
+  #      cex=2)
+  par(plt=posPlot(numplotx = 4,numploty = 3,idplotx = count,idploty = 1,
+                  lowboundx = 8,lowboundy = 10,
+                  upboundy = 90), las=1,new=TRUE)
+  # dataIndReact<-sapply(as.list(tempPop[,indId]),
+  #                      function(x){x=
+  #                        sapply(rangx, 
+  #                               function(y)
+  #                                 do.call(logist,
+  #                                         as.list(
+  #                                           c(y,as.double(
+  #                                             tempPop[indId==x,.SD,
+  #                                                     .SDcol=c("alpha","beta")])))))})
+  # meansClustmp<-tempPop[,.(alphaMean=mean(alpha),betaMean=mean(beta)),by=orderClus]
+  # reacNormpClust<-sapply(meansClustmp[,orderClus], function(x){
+  #   logist(rangx,alpha = meansClustmp[orderClus==x,alphaMean],
+  #          beta = meansClustmp[orderClus==x,betaMean])})
+  # matplot(x = rangx,y = reacNormpClust,col=colRuns[meansClustmp$orderClus],
+  #          type='l',cex.axis=cexAxis,
+  #          xlab="",ylab="",ylim=c(0,1),lty=1,
+  #          lwd=2,yaxt=seqYax[count],xaxt="s",cex.lab=cexAxis)
+   reacNormInd<-sapply(tempPop[indId %in%randInds,indId], function(x){
+    logist(rangx,alpha = tempPop[indId==x,alpha],
+           beta = tempPop[indId==x,beta])})
+  matplot(x = rangx,y = reacNormInd,
+          type='l',cex.axis=cexAxis,
+          xlab="",ylab="",ylim=c(0,1),lty=1,col=1,
+          lwd=2,yaxt=seqYax[count],xaxt="s",cex.lab=cexAxis)
+  mtext(text = seqYlabDown[count],cex = cexAxis,line = 1.5,side = 2,las=0)
+  mtext(text = seqXlabDown[count],cex = cexAxis,line = 1.5,side = 1,las=0)
+
+  }
 
@@
 
+| Symbol | Description |
+|-------:|------------:|
+| $Q_i$ | Quality of individual $i$ |
+| $N$   | Population size |
+| $\sigma$ | Standard deviation of the truncated normal distribution from which quality is drawn |
+| $B_i$ | Badge size of individual $i$ |
+| $\alpha_i$ | Intercept of the badge size reaction norm for individual $i$ |
+| $\beta_i$  | Slope of the badge size reaction norm for individual $i$ |
+| $ \mu$     | Mutation rate | 
+| $\sigma_\mu$ |Standard deviation of the normal distribution from which quality is mutations are drawn |
+| $s_i$ | Survival probability of individual $i$ |
+| $k_1$ | Intercept of the survival function |
+| $K_2$ | Slope of the survival function |
+| $V$   | Value of the contested resource |
+| $C$   | Cost of an escalated fight |
+| $p_{ij}$ | Probability that individual $i$ wins an escalated fight against individual $j$ |
+| $k_3$ | Parameter determining how important is quality defining the probability of wining |
+| $\delta$ | Prediction error |
+| $A$ | Speed of learning | 
+| $c$ | Number of feature centres | 
+| $\hat{R}$ | Reward estimate | 
+|
+
 ## References

### 2023-01-16 Dyn_reac_learn.R

@@ -9,11 +9,11 @@ source(here("AccFunc.R"))
 require("foreach")
 require("doParallel")
 require("jsonlite")
-require("d")
+library(dplyr)
 
 # Scenario to be plotted - corresponds to folders where simulations are stored
 
-scenario<-"betCostalphaL"
+scenario<-"betCostalphaL2"
 
 # extSimsDir<-here("Simulations",paste0(scenario,"_"))
   # paste0("e:/BadgeSims/",scenario,"_")
@@ -32,7 +32,7 @@ scenario<-"betCostalphaL"
 (indList<-grep("indLearn",listTest,value=TRUE))
 paramName<-list.files(here("Simulations",paste0(scenario,"_")))
 # paramName<-list.files(extSimsDir,full.names = TRUE)
-  paramName<-grep(".json",paramName,value=TRUE)
+paramName<-grep(".json",paramName,value=TRUE)
 paramRang<-sapply(paramName,function(json){
   tmpJson<-fromJSON(here("Simulations",paste0(scenario,"_"),json))
   return(tmpJson$rangParam)
@@ -174,7 +174,7 @@ with(evolStats,{
   # lines(time,m.meanInAct,col=colGenesLin[2],lwd=3)
 })
 legend("topleft",legend = c(expression(alpha),expression(beta)),
-       col=colGenesLin,lwd=2,bty = "n")
+       col=colGenesLin,lwd=2,bty = "n",ncol = 2)
 axis(side=1,padj = -3.5,cex=0.8,at=axTicks(1),labels = axTicks(1)/100)
 
 # Evolutionary Dynamics of learning parameters
@@ -183,7 +183,7 @@ par(plt=posPlot(numploty = 3,idploty = 2,numplotx = 3,idplotx = 2,
                 lowboundx = 8,upboundx = 93),xaxt="s",las=1,new=TRUE)
 plot(x=c(0,max(evolStats$time)),y=c(0,0),type="l",lwd=2,col="grey",
      # ylim=fivenum(as.matrix(evol[,meanInitCrit,meanInitAct]))[c(1,5)],yaxt="n",
-     ylim=fivenum(as.matrix(evol[,meanAlphaLearn]))[c(1,5)],yaxt="n",
[…]

### 2023-01-16 aesth.R

@@ -23,9 +23,8 @@ colGenesLin<-c(rgb(241/255,163/255,64/255),
                rgb(153/255,142/255,195/255),
                rgb(127/255,191/255,123/255))
 
-paletteMeans <- colorRampPalette(c('#d73027','#fc8d59','#fee090',
-                                   '#e0f3f8','#91bfdb','#4575b4')[6:1],
-                                 alpha=TRUE)
+paletteMeans <- colorRampPalette(c('#d73027','#fc8d59','#fdcb44',
+                                   '#a2dbea','#91bfdb','#4559b4')[6:1])
 colboxes<- c('#31a354','#d7191c','#fdae61','#2b83ba',"black")
 
 colRuns<-c('#e41a1c','#377eb8','#4daf4a', '#984ea3','#ff7f00')

### 2023-02-01 manuscript/manuscript_1.0.Rmd
Update the figures. Changing the order of the panels and adding labels

--- 2022-07-20 manuscript/manuscript_1.0.Rmd
@@
 output:
-  pdf_document: default
-  word_document: default
+  pdf_document: 
+    latex_engine: xelatex  # word_document: default
 fig_caption: yes
@@
     \usepackage{float}
+    \usepackage{unicode-math}
     \usepackage{setspace}\doublespacing
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE,cache=TRUE,warnings=FALSE)
+knitr::opts_chunk$set(echo = FALSE,cache=TRUE,warnings=FALSE)
 require(png)
 require(here)
+library(dplyr)
 source(here("../R_files/posPlots.r"))
 source(here("AccFunc.r"))
+library("scales")
 ```
@@
 Specifically, we pick $c$ feature centres, which are evenly spaced values 
-along the badge size interval ([0,1]). Each one of these features centres 
+along the badge size interval ([0,1]). This feature centres are the same for all individuals and stored in vector $\symbf{b}$ . Each one of these features centres 
 is associated with a weight for reward estimation and tendency to play 
@@
 \begin{equation}
-\hat{R}_{ij} =  \sum_{z=1}^{c} x_z e^{(-\frac{|B_i-c_z|}{2\theta^2})} 
+\hat{R}_{ij} =  \sum_{z=1}^{c} x_z e^{(-\frac{|B_i-b_z|}{2\theta^2})} 
 \end{equation}
@@
 \begin{equation}
-logit(q_{ij}) =  \sum_{z=1}^{c} y_z e^{(-\frac{|B_i-c_z|}{2\theta^2})} 
+logit(q_{ij}) =  \sum_{z=1}^{c} y_z e^{(-\frac{|B_i-b_z|}{2\theta^2})} 
 \end{equation}
@@
 After all their interactions, individuals in the population 
-reproduce proportionally to their fitness $w_i$, which is a sum of the 
-baseline fitness ($w_0$) and all the pay-offs obtained throughout their 
+reproduce proportionally to their total pay-off $w_i$, which is a sum of the 
+baseline pay-off ($w_0$) and all the pay-offs obtained throughout their 
 life. Thus, the combination of natural selection and genetic drift 
@@
 probability of retreating depending on their own quality. Individuals 
-with higher quality (red tones), after the learning process, have a 
+with higher quality (red tones) after the learning process have a 
 high probability of attacking; while individuals of lower quality 
@@
 \\ref{fig:model-struc}.
-In the panels below, colors represent a cluster clasification performed by the k-means algorithm" }
-
-knitr::include_graphics(here("Simulations","betCostEvol1_","evolDyn0_betCost5.png"))
+In the panels below, colors represent a cluster clasification performed by the k-means algorithm" ,warning=FALSE}
+
+scenario<-"betCostEvol3"#"alphaAct"#"nIntGroupNormQual"#
+runChoi<-9 #1
+
+# Load files -------------------------------------------------------------------
+# Project folder
+listTest<-list.files(here("Simulations",paste0(scenario,"_")),full.names = TRUE)
+evolList<-grep("evolLearn",listTest,value=TRUE)
+indList<-grep("indLearn",listTest,value=TRUE)
+paramName<-list.files(here("Simulations",paste0(scenario,"_")))
+# paramName<-list.files(extSimsDir,full.names = TRUE)
+paramName<-grep(".json",paramName,value=TRUE) %>% grep(pattern = "*5_*",value = TRUE)
+
+param<-fromJSON(here("Simulations",paste0(scenario,"_"),paramName[2]))
+  # fromJSON(paramName[1])
+val<-2
+fileId<-val
+evolList_runs<-grep(paste0(param$namParam,param$rangParam[2]),
+                    evolList,value =TRUE)
+indList_runs<-grep(paste0(param$namParam,param$rangParam[2]),
+                   indList,value =TRUE)
+evol<-do.call(rbind,lapply(evolList_runs,fread))
+pop<-do.call(rbind,lapply(indList_runs, fread))
+
+Valpar<-param$rangParam[2]
+
+nampar<-param$namParam
+
+nCenters<-param$nCenters
+sigSquar<-param$sigSq
+
+# Aesthetic paremeters ---------------------------------------------------------
+
+## Calculate clustering for all generations  -------------------
+
+# # Choose which interaction to visualize
+# lastInt<-tail(pop[,unique(nInteract)],4)[2]
+# # # pop[,max(nInteract),by=.(seed,time)][,min(V1)]
+# # # to get last interaction: tail(pop[,unique(nInteract)],1)
+# #
+# popOneInd<-pop[nInteract==lastInt]
+# vars<-c("alpha","beta","Badge")
+# # ,"Badge",grep("WeightAct",names(popFinal),value = TRUE))
+# popOneInd$idClust<-get_clusters(popOneInd,vars,k.max = 5,
+#                                 Bsamples =500,iterMax = 500)
+# clusSummary<-popOneInd[,.(meanAlph=mean(alpha),meanBet=mean(beta)),
+#                        by=.(idClust,time,seed)]
+# clusSummary[,orderClus:=0]
+# invisible(capture.output(
+#   for(i in 1:dim(clusSummary)[1]){
+#   set(clusSummary,i,6L,match(clusSummary[i,meanAlph],
+#         sort(clusSummary[time==clusSummary[i,time]&
+#                            seed==clusSummary[i,seed]]$meanAlph)))
+# }
+# ))
+# 
+# 
+# popOneInd<-merge(popOneInd,clusSummary,all = TRUE)
+# 
+# fwrite(popOneInd,file = here("Simulations",paste0(scenario,"_"),
+#                              paste0("popOneInd",nampar,Valpar,".txt")))
+
+popOneInd<-fread(here("Simulations",paste0(scenario,"_"),
+paste0("popOneInd",nampar,Valpar,".txt")))
+
+# Get stats from the evolutionary simulations ----------------------------------
+# names(evol)
+
+cols<-c("freqGenHawks","freqGenDove",  "freqGenEval",  #"freqGenLearn",
+        "freqFenHawks", "freqFenDoves", "freqHH", "freqHD", "freqDD", "meanCue",
+        "meanAlpha", "meanBeta",
+        #"meanFit", #"meanInitCrit", "sdInitCrit", "meanInitAct", "sdInitAct",
+        "WeightAct_0","WeightCrit_0",
+        "WeightAct_1","WeightCrit_1","WeightAct_2",  "WeightCrit_2",
+        "WeightAct_3",  "WeightCrit_3", "WeightAct_4",  "WeightCrit_4",
+        "WeightAct_5","WeightCrit_5")
+
+my.summary<- function(x) list(mean = mean(x), lowIQR = fivenum(x)[2], 
+                              upIQR = fivenum(x)[4])
+
+evolStats<-evol[, as.list(unlist(lapply(.SD, my.summary))), 
+                .SDcols = cols,by=time]
+
+## Plot mean and IQRs among individuals for each replicate --------------------
+
+# get the trajectories for individual runs
+traitsTrajs<-dcast(evol,time~seed,
+                   value.var = c("meanAlpha","meanBeta",
+                                 #"meanInitCrit",#"meanFit",
+                                 # "meanInitAct","sdInitCrit","sdInitAct",
+                                 "sdAlpha","sdBeta","freqHH",
+                                 "freqHD","freqDD"))
+
+finReps<-evol[time==max(time),seed]
+
+genstoPrint<-round(seq(1,length(unique(evolStats$time)),length.out = 5))[2:5]
+
+# Evolutionary dynamics for alpha - intercept of the sender --------------------------------------------
+cexAxis<-0.9
+
+nInd2print<-40
+
+par(plt=posPlot(numploty = 3,idploty = 3,numplotx = 2,idplotx = 1,
+                lowboundx = 8, title = TRUE,lowboundy = 15,
+                upboundy = 99),
+    mfrow=c(1,1), xaxt="s",las=1,mgp=c(2,0.5,0))
+
+adjsYlims<-abs(range(pop[seed==runChoi,alpha])[1]-
+                 range(pop[seed==runChoi,alpha])[2])*c(-0.1,0.2)
+
+evolDist(indData = pop[seed==runChoi],variable = "alpha",nbins = 10,
+         pal = pal_dist,
+         nlevels=10,cexAxis = cexAxis,xlab="",ylab = "",xaxt="n",yaxt = "s",
+         ylim = range(pop[seed==runChoi,alpha])+adjsYlims,
+         xlim=c(0,22000))
+
+mtext(text = "Trait value",cex = cexAxis,line = 1,side = 2,las=0)
+legend("topright",legend = c(expression(alpha[s])),col=colGenesLin[1],lwd=2,
+       bty = "n",cex=cexAxis)
+legend("topleft",legend = c("A"),cex=cexAxis,bty = "n",inset = -0.06)
+
+axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
+     labels = axTicks(1)[2:(length(axTicks(1)))]/1000,cex.axis=cexAxis,
+     hadj = 1.3,line = -0.5,tick = FALSE)
+
+# grey lines to show the generations shown in the upper and lower panels
+
+matlines(x=matrix(rep(evolStats[genstoPrint,time],each=2),nrow=2),
+         y=matrix(rep(range(pop[seed==runChoi,alpha]),4),nrow = 2),
+         lty=1,col = "grey",lwd=2)
+
+# Variation among individuals
+
+matlines(x=traitsTrajs[,time],
+         y=traitsTrajs[,.SD,
+                       .SDcol=paste0(c("meanAlpha_"),runChoi)],
+         col=colGenesLin,lty = 2,type="l",lwd=5)
+
+par(plt=posPlot(numploty = 3,idploty = 3,numplotx = 2, idplotx = 2,
+                lowboundx = 8, title = TRUE,lowboundy = 15,
+                upboundy = 99),
+    xaxt="s",las=1,new=TRUE)
+
+adjsYlims<-abs(range(pop[seed==runChoi,beta])[1]-
+                 range(pop[seed==runChoi,beta])[2])*c(-0.1,0.2)
+
+evolDist(indData = pop[seed==runChoi],variable = "beta",nbins = 10,pal = pal_dist,
+         nlevels=10,cexAxis = 1.5,xlab="",ylab = "",xaxt="n",yaxt = "n",
+         ylim = range(pop[seed==runChoi,beta])+adjsYlims,xlim=c(0,22000))
+legend("topright",legend = c(expression(beta[s])),cex=cexAxis,
+       col=colGenesLin[2],lwd=2,bty = "n")
+legend("topleft",legend = c("B"),cex=cexAxis,bty = "n",inset = -0.06)
+# grey lines to show the generations shown in the upper and lower panels
+matlines(x=matrix(rep(evolStats[genstoPrint,time],each=2),nrow=2),
+         y=matrix(rep(range(pop[seed==runChoi,beta]))),
+                      lty=1,col = "grey",lwd=2)
+
+axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
+     labels = axTicks(1)[2:(length(axTicks(1)))]/1000,cex.axis=cexAxis,
+     hadj = 1.3,line = -0.5,tick = FALSE)
+
+axis(4,cex.axis=cexAxis)
+
+matlines(x=traitsTrajs[,time],
+         y=traitsTrajs[,.SD,
+                       .SDcol=paste0(c("meanBeta_"),runChoi)],
+         col=colGenesLin[2],lty = 2,type="l",lwd=4)
+
+# Choose time range
+gen2plot<-round(seq(1,length(unique(evolStats$time)),length.out = 5))[2:5]
+# Plor the actor
+seqYax<-c("s",rep("n",3))
+# seqYlabUp<-c("Badge",rep("",3))
+seqYlabUp<-c("P(dove)",rep("",3))
+seqYlabDown<-c("Badge",rep("",3))
+seqXlabDown<-c("","Quality","")#paste0("seed: ",runChoi))
+seqXlabUp<-c("","Badge","")
+labPanels<-c("C","D","E","F","G")
+rangQual<-seq(0,1,length.out = 50)
+interv<-1/(nCenters-1)
+centers<-interv*seq(0,nCenters-1)
+rangx<-seq(0,1,length=50)
+count<-0
+for(genC in genstoPrint){
+  count<-count+1
+  par(plt=posPlot(numplotx = 4,numploty = 3,idplotx = count,idploty = 1,
+                  lowboundx = 8, title = TRUE,lowboundy = 10,
+                  upboundy = 87),
+      xaxt="s",las=1,new=TRUE)
+  weightsAct<-as.double(evol[(time==unique(time)[genC])&seed==runChoi,.SD,
+                             .SDcols=grep("WeightAct",
+                                          names(evol),value = TRUE)])
+  tempPop<-popOneInd[(time==unique(time)[genC]&seed==runChoi),.SD[.N],
+               .SDcol=c(grep("WeightAct",
+                             names(evol),value = TRUE),"Quality","alpha","beta",
+                        "orderClus"),
+               by=indId]
+
+  randInds<-sample(tempPop$indId, nInd2print, replace = FALSE)
+
+  dataIndAct<-sapply(as.list(tempPop[indId %in%randInds,indId]),
+                     function(x){x=
+                       logist(totRBF(rangx,
+                                     centers,sigSquar,
+                                     as.double(
+                                       tempPop[indId==x,.SD,
+                                               .SDcol=grep("WeightAct",
+                                                           names(tempPop),
+                                                           value = TRUE)
+                                               ])),alpha=0,beta = 1)})
+  matplot(x=rangx,y=dataIndAct,
+          yaxt=seqYax[count],ylab="",xlab="",type="l",cex.lab=cexAxis,
+          lwd=2,xaxt="n",ylim=c(-0.06,1.1),col = paletteMeans(100)[
+            findInterval(tempPop[indId %in%randInds,Quality],
+                         colorbreaksQual)],lty = 1,
+          cex.axis=cexAxis)
+  axis(side = 1,cex.axis=cexAxis,padj = 0)
+  mtext(text = seqYlabUp[count],cex = cexAxis,line = 1.7,side = 2,las=0)
+  mtext(text = seqXlabUp[count],cex = cexAxis,line = 1.2,side = 1 ,las=0)
+  lines(logist(totRBF(rangx,centers,sigSquar,weightsAct),alpha=0,beta=1)~rangx,
+        col=1,lwd=3,lty=2)
+  legend("topleft",legend = c(paste0("D.",count)),bty = "n",inset = -0.06)
+  # text(x=0.5,y=-0.015,labels = paste0("time=",unique(evolStats$time)[genC]/1000),
+  #      cex=2)
+  par(plt=posPlot(numplotx = 4,numploty = 3,idplotx = count,idploty = 2,
+                  lowboundx = 8,lowboundy = 15,
+                  upboundy = 99), las=1,new=TRUE)
+
+# meansClustmp<-tempPop[,.(alphaMean=mean(alpha),betaMean=mean(beta)),by=orderClus]
+# reacNormpClust<-sapply(meansClustmp[,orderClus], function(x){
+#   logist(rangx,alpha = meansClustmp[orderClus==x,alphaMean],
+#          beta = meansClustmp[orderClus==x,betaMean])})
+# matplot(x = rangx,y = reacNormpClust,col=colRuns[meansClustmp$orderClus],
+#          type='l',cex.axis=cexAxis,
+#          xlab="",ylab="",ylim=c(0,1.1),lty=1,
+#          lwd=2,yaxt=seqYax[count],xaxt="s",cex.lab=cexAxis)
+  reacNormInd<-sapply(tempPop[indId %in%randInds,indId], function(x){
+    logist(rangx,alpha = tempPop[indId==x,alpha],
+           beta = tempPop[indId==x,beta])})
+  matplot(x = rangx,y = reacNormInd,
+          type='l',cex.axis=cexAxis,
+          xlab="",ylab="",ylim=c(0,1.1),lty=1,col=1,
+          lwd=2,yaxt=seqYax[count],xaxt="s",cex.lab=cexAxis)
+  mtext(text = seqYlabDown[count],cex = cexAxis,line = 1.5,side = 2,las=0)
+  mtext(text = seqXlabDown[count],cex = cexAxis,line = 1.3,side = 1,las=0)
+  legend("topleft",legend =  c(paste0("C.",count)),bty = "n",inset = -0.06)
+  }
+
+rm(list=grep("temp",ls(),value = T))
 
@@
 size work as a handicap (i.e. the cost of the badge is inversely 
-proportional to the quality of the individual), the sender code evolves 
+proportional to the quality of the individual), the sender code can evolve 
 to produce an honest signal Fig \ref{fig:handicap}). The evolution of 
@@
 Thus, badges are only costly and do not provide information about the 
-quality of an individual(Fig. \ref{fig:noBadge} . Eventually, the badge disappears from the population.
-
-When production of the badge is not costly (i.e. does not work as a handicap), 
-the parameters of the sender code go through evolutionary branching event, 
-triggering the evolution of different types of signals in the population. 
-Fig. \ref{fig:branching}A and B shows changes in the distribution of the 
-sender reaction norm ($\alpha$ and $\beta$) along evolutionary time. 
+quality of an individual (Fig. \ref{fig:noBadge} . Eventually, the badge 
+disappears from the population. In contrast, when the badge does not work as a handicap but instead is cost-free, the evolutionary process never leads to the 
+establishment of an honest signal. Instead, subject to genetic drift the 
+population either evolves towards the disappearance of the badge, or to the 
+maximum badge size. In either of this cases reaction norms are flat, so the 
+badge does not provide any quality information.
+
+The amount of information that agents are able to collect through learning 
+throughout their life time can strongly change the outcome of 
+the evolutionary dynamics. In the simulations presented in figure 
+\ref{fig:handicap} and \ref{fig:noBadge}, 
+individuals learned with high learning rate ($\alpha = 0.4$) and interacted 
+repeatedly along their lifetime (on average 2000 interactions). When we reduce 
+the number of interactions that individuals have over their lifetime to 300 
+interactions on average, we see an drastic increase in the morphological and genetic
+variation present in the population. Fig. \ref{fig:branching} shows 
+changes in the distribution of the sender reaction norm
+along evolutionary time assuming the badge is cost-free. 
 Darker areas show trait values with high frequency in the population. 
@@
 and mutations quickly build up a normal distribution around the starting 
-value. Within the first 2000 generations the unimodal distribution splits 
-into a bimodal one. In the case of $\alpha$, one of the peaks splits further, 
+value. Within the first 2000 generations the unimodal distribution in 
+the intercept ($\alpha$) splits into a bimodal one. 
+Later in evolutionary time one of the peaks splits further, 
 so at the end of the evolutionary simulation the distribution of $\alpha$ 
-value in the population shows three peaks. In the case of $\beta$, the two 
-initial peaks coexist for around 4000 generations. Later on one of the 
-types goes extinct and the distribution goes back to being unimodal. 
-These changes in the distribution of the sender reaction norm parameters 
+value in the population shows three peaks. In the case of $\beta$, the 
+peaks in the distribution are not so clear-cut, but it's clear that there is an 
+increase in the variance of the distribution. These changes in the 
+distribution of the sender reaction norm parameters 
 imply that during the second half of the simulation individuals can be 
@@
 
-```{r branching,echo=FALSE,out.width='90%',fig.align='center',fig.show='hold',fig.cap= "The evolution of cheap signals.
+```{r branching,echo=FALSE,out.width='100%',fig.height=6,fig.align='center',fig.show='hold',fig.cap= "The evolution of cheap signals.
 Portrait of the evolutionary dynamics of the sender code with snapsots of both sender and receiver codes just as in fig.
@@
 The middle panels show changes in the distribution of values fir $\\alpha$ and $\\beta$ along evolutionary time.
-Panels above and bellow correspond to snapshots of the sender and receiver codes, respectively, generation time of the snapshots are indicated by the grey lines in the middle panels."}
-
-knitr::include_graphics(here("Simulations","nIntGroupNormQual_","evolDyn7_nIntGroup2000.png"))
+Panels above and below correspond to snapshots of the sender and receiver codes, respectively, generation time of the snapshots are indicated by the grey lines in the middle panels."}
+
+# knitr::include_graphics(here("Simulations","nIntGroupNormQual_","evolDyn7_nIntGroup2000.png"))
+
+scenario<-"betCostEvol4"#"alphaAct"#"nIntGroupNormQual"#
+runChoi<-12 #1
+
+# Load files -------------------------------------------------------------------
+# Project folder
+listTest<-list.files(here("Simulations",paste0(scenario,"_")),full.names = TRUE)
+evolList<-grep("evolLearn",listTest,value=TRUE)
+indList<-grep("indLearn",listTest,value=TRUE)
+
+val<-2
+fileId<-val
+
+paramName<-list.files(here("Simulations",paste0(scenario,"_")))
+paramName<-grep(".json",paramName,value=TRUE) %>% grep(pattern = "*s0_*",value = TRUE)
+
+param<-fromJSON(here("Simulations",paste0(scenario,"_"),paramName[2]))
+
+evolList_runs<-grep(paste0(param$namParam,param$rangParam),
+                    evolList,value =TRUE)
+indList_runs<-grep(paste0(param$namParam,param$rangParam),
+                   indList,value =TRUE)
+
+evol<-do.call(rbind,lapply(evolList_runs,fread))
+pop<-do.call(rbind,lapply(indList_runs, fread))
+
+Valpar<-param$rangParam
+
+nampar<-param$namParam
+
+nCenters<-param$nCenters
+sigSquar<-param$sigSq
+
+# Aesthetic paremeters ---------------------------------------------------------
+
+## Calculate clustering for all generations  -------------------
+
+### Choose which interaction to visualize
+# lastInt<-tail(pop[,unique(nInteract)],4)[2]
+# # pop[,max(nInteract),by=.(seed,time)][,min(V1)]
+# # to get last interaction: tail(pop[,unique(nInteract)],1)
+# 
+# popOneInd<-pop[nInteract==lastInt]
+# vars<-c("alpha","beta","Badge")
+# # ,"Badge",grep("WeightAct",names(popFinal),value = TRUE))
+# popOneInd$idClust<-get_clusters(popOneInd,vars,k.max = 10,
+#                                 Bsamples =500,iterMax = 500)
+# clusSummary<-popOneInd[,.(meanAlph=mean(alpha),meanBet=mean(beta)),
+#                        by=.(idClust,time,seed)]
+# clusSummary[,orderClus:=0]
+# invisible(capture.output(
+#   for(i in 1:dim(clusSummary)[1]){
+#   set(clusSummary,i,6L,match(clusSummary[i,meanAlph],
+#         sort(clusSummary[time==clusSummary[i,time]&
+#                            seed==clusSummary[i,seed]]$meanAlph)))
+# }
+# ))
+# 
+# 
+# popOneInd<-merge(popOneInd,clusSummary,all = TRUE)
+# 
+# 
+# fwrite(popOneInd,file = here("Simulations",paste0(scenario,"_"),
+#                              paste0("popOneInd",nampar,Valpar,".txt")))
+
+popOneInd<-fread(here("Simulations",paste0(scenario,"_"),
+                             paste0("popOneInd",nampar,Valpar,".txt")))
+
+# Get stats from the evolutionary simulations ----------------------------------
+# names(evol)
+
+cols<-c("freqGenHawks","freqGenDove",  "freqGenEval",  #"freqGenLearn",
+        "freqFenHawks", "freqFenDoves", "freqHH", "freqHD", "freqDD", "meanCue",
+        "meanAlpha", "meanBeta",
+        #"meanFit", #"meanInitCrit", "sdInitCrit", "meanInitAct", "sdInitAct",
+        "WeightAct_0","WeightCrit_0",
+        "WeightAct_1","WeightCrit_1","WeightAct_2",  "WeightCrit_2",
+        "WeightAct_3",  "WeightCrit_3", "WeightAct_4",  "WeightCrit_4",
+        "WeightAct_5","WeightCrit_5")
+
+my.summary<- function(x) list(mean = mean(x), lowIQR = fivenum(x)[2], 
+                              upIQR = fivenum(x)[4])
+
+evolStats<-evol[, as.list(unlist(lapply(.SD, my.summary))), 
+                .SDcols = cols,by=time]
+
+## Plot mean and IQRs among individuals for each replicate --------------------
+
+# get the trajectories for individual runs
+traitsTrajs<-dcast(evol,time~seed,
+                   value.var = c("meanAlpha","meanBeta",
+                                 #"meanInitCrit",#"meanFit",
+                                 # "meanInitAct","sdInitCrit","sdInitAct",
+                                 "sdAlpha","sdBeta","freqHH",
+                                 "freqHD","freqDD"))
+
+finReps<-evol[time==max(time),seed]
+
+genstoPrint<-round(seq(1,length(unique(evolStats$time)),length.out = 5))[2:5]
+
+# Evolutionary dynamics for alpha - intercept of the sender --------------------------------------------
+par(plt=posPlot(numploty = 3,idploty = 3,numplotx = 2,idplotx = 1,
+                lowboundx = 8, title = TRUE,lowboundy = 15,
+                upboundy = 99),
+    mfrow=c(1,1), xaxt="s",las=1,mgp=c(2,0.5,0))
+
+adjsYlims<-abs(range(pop[seed==runChoi,alpha])[1]-
+                 range(pop[seed==runChoi,alpha])[2])*c(-0.1,0.2)
+
+evolDist(indData = pop[seed==runChoi],variable = "alpha",nbins = 10,
+         pal = pal_dist,
+         nlevels=10,cexAxis = cexAxis,xlab="",ylab = "",xaxt="n",yaxt = "s",
+         ylim = range(pop[seed==runChoi,alpha])+adjsYlims,
+         xlim=c(0,22000))
+
+mtext(text = "Trait value",cex = cexAxis,line = 1,side = 2,las=0)
+legend("topright",legend = c(expression(alpha[s])),col=colGenesLin[1],lwd=2,
+       bty = "n",cex=cexAxis)
+legend("topleft",legend = c("A"),cex=cexAxis,bty = "n",inset = -0.06)
+
+axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
+     labels = axTicks(1)[2:(length(axTicks(1)))]/1000,cex.axis=cexAxis,
+     hadj = 1.3,line = -0.5,tick = FALSE)
+
+# grey lines to show the generations shown in the upper and lower panels
+
+matlines(x=matrix(rep(evolStats[genstoPrint,time],each=2),nrow=2),
+         y=matrix(rep(range(pop[seed==runChoi,alpha]),4),nrow = 2),
+         lty=1,col = "grey",lwd=2)
+
+# Variation among individuals
+
+matlines(x=traitsTrajs[,time],
+         y=traitsTrajs[,.SD,
+                       .SDcol=paste0(c("meanAlpha_"),runChoi)],
+         col=colGenesLin,lty = 2,type="l",lwd=5)
+
+par(plt=posPlot(numploty = 3,idploty = 3,numplotx = 2, idplotx = 2,
+                lowboundx = 8, title = TRUE,lowboundy = 15,
+                upboundy = 99),
+    xaxt="s",las=1,new=TRUE)
+
+adjsYlims<-abs(range(pop[seed==runChoi,beta])[1]-
+                 range(pop[seed==runChoi,beta])[2])*c(-0.1,0.2)
+
+evolDist(indData = pop[seed==runChoi],variable = "beta",nbins = 10,pal = pal_dist,
+         nlevels=10,cexAxis = 1.5,xlab="",ylab = "",xaxt="n",yaxt = "n",
+         ylim = range(pop[seed==runChoi,beta])+adjsYlims,xlim=c(0,22000))
+legend("topright",legend = c(expression(beta[s])),cex=cexAxis,
+       col=colGenesLin[2],lwd=2,bty = "n")
+legend("topleft",legend = c("B"),cex=cexAxis,bty = "n",inset = -0.06)
+# grey lines to show the generations shown in the upper and lower panels
+matlines(x=matrix(rep(evolStats[genstoPrint,time],each=2),nrow=2),
+         y=matrix(rep(range(pop[seed==runChoi,beta]))),
+                      lty=1,col = "grey",lwd=2)
+
+axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
+     labels = axTicks(1)[2:(length(axTicks(1)))]/1000,cex.axis=cexAxis,
+     hadj = 1.3,line = -0.5,tick = FALSE)
+
+axis(4,cex.axis=cexAxis)
+
+matlines(x=traitsTrajs[,time],
+         y=traitsTrajs[,.SD,
+                       .SDcol=paste0(c("meanBeta_"),runChoi)],
+         col=colGenesLin[2],lty = 2,type="l",lwd=4)
+
+# Choose time range
+gen2plot<-round(seq(1,length(unique(evolStats$time)),length.out = 5))[2:5]
+# Plor the actor
+seqYax<-c("s",rep("n",3))
+# seqYlabUp<-c("Badge",rep("",3))
+seqYlabUp<-c("P(dove)",rep("",3))
+seqYlabDown<-c("Badge",rep("",3))
+seqXlabDown<-c("","Quality","")#paste0("seed: ",runChoi))
+seqXlabUp<-c("","Badge","")
+labPanels<-c("C","D","E","F","G")
+rangQual<-seq(0,1,length.out = 50)
+interv<-1/(nCenters-1)
+centers<-interv*seq(0,nCenters-1)
+rangx<-seq(0,1,length=50)
+count<-0
+for(genC in genstoPrint){
+  count<-count+1
+  par(plt=posPlot(numplotx = 4,numploty = 3,idplotx = count,idploty = 1,
+                  lowboundx = 8, title = TRUE,lowboundy = 10,
+                  upboundy = 87),
+      xaxt="s",las=1,new=TRUE)
+  weightsAct<-as.double(evol[(time==unique(time)[genC])&seed==runChoi,.SD,
+                             .SDcols=grep("WeightAct",
+                                          names(evol),value = TRUE)])
+  tempPop<-popOneInd[(time==unique(time)[genC]&seed==runChoi),.SD[.N],
+               .SDcol=c(grep("WeightAct",
+                             names(evol),value = TRUE),"Quality","alpha","beta",
+                        "orderClus"),
+               by=indId]
+
+  randInds<-sample(tempPop$indId, nInd2print, replace = FALSE)
+
+  dataIndAct<-sapply(as.list(tempPop[indId %in%randInds,indId]),
+                     function(x){x=
+                       logist(totRBF(rangx,
+                                     centers,sigSquar,
+                                     as.double(
+                                       tempPop[indId==x,.SD,
+                                               .SDcol=grep("WeightAct",
+                                                           names(tempPop),
+                                                           value = TRUE)
+                                               ])),alpha=0,beta = 1)})
+  matplot(x=rangx,y=dataIndAct,
+          yaxt=seqYax[count],ylab="",xlab="",type="l",cex.lab=cexAxis,
+          lwd=2,xaxt="n",ylim=c(-0.06,1.1),col = paletteMeans(100)[
+            findInterval(tempPop[indId %in%randInds,Quality],
+                         colorbreaksQual)],lty = 1,
+          cex.axis=cexAxis)
+  axis(side = 1,cex.axis=cexAxis,padj = 0)
+  mtext(text = seqYlabUp[count],cex = cexAxis,line = 1.7,side = 2,las=0)
+  mtext(text = seqXlabUp[count],cex = cexAxis,line = 1.2,side = 1 ,las=0)
+  lines(logist(totRBF(rangx,centers,sigSquar,weightsAct),alpha=0,beta=1)~rangx,
+        col=1,lwd=3,lty=2)
+  legend("topleft",legend = c(paste0("D.",count)),bty = "n",inset = -0.06)
+  # text(x=0.5,y=-0.015,labels = paste0("time=",unique(evolStats$time)[genC]/1000),
+  #      cex=2)
+  par(plt=posPlot(numplotx = 4,numploty = 3,idplotx = count,idploty = 2,
+                  lowboundx = 8,lowboundy = 15,
+                  upboundy = 99), las=1,new=TRUE)
+
+# meansClustmp<-tempPop[,.(alphaMean=mean(alpha),betaMean=mean(beta)),by=orderClus]
+# reacNormpClust<-sapply(meansClustmp[,orderClus], function(x){
+#   logist(rangx,alpha = meansClustmp[orderClus==x,alphaMean],
+#          beta = meansClustmp[orderClus==x,betaMean])})
+# matplot(x = rangx,y = reacNormpClust,col=colRuns[meansClustmp$orderClus],
+#          type='l',cex.axis=cexAxis,
+#          xlab="",ylab="",ylim=c(0,1.1),lty=1,
+#          lwd=2,yaxt=seqYax[count],xaxt="s",cex.lab=cexAxis)
+  reacNormInd<-sapply(tempPop[indId %in%randInds,indId], function(x){
+    logist(rangx,alpha = tempPop[indId==x,alpha],
+           beta = tempPop[indId==x,beta])})
+  matplot(x = rangx,y = reacNormInd,
+          type='l',cex.axis=cexAxis,
+          xlab="",ylab="",ylim=c(0,1.1),lty=1,col=1,
+          lwd=2,yaxt=seqYax[count],xaxt="s",cex.lab=cexAxis)
+  mtext(text = seqYlabDown[count],cex = cexAxis,line = 1.5,side = 2,las=0)
+  mtext(text = seqXlabDown[count],cex = cexAxis,line = 1.3,side = 1,las=0)
+  legend("topleft",legend =  c(paste0("C.",count)),bty = "n",inset = -0.06)
+  }
+
+rm(list=grep("temp",ls(),value = T))
 
 ```
-
-```{r startCond, echo=FALSE, out.width='90%',fig.align='center',fig.show='hold',fig.cap="WHy would you want to be recognized"}
-
-knitr::include_graphics("../Simulations/initAct_/indVarScatter_.png")
+The effect of a limited number of interactions on the amount of variation in 
+the evolving parameters plays a role when the signal follows the handicap
+principle. In fig \ref{fig:honestDiv}, we show simulations where individuals 
+have on average 300 interaction in their life and the cost of the signal
+is proportional to the quality, following the handicap principle. 
+The evolutionary process leads to a combination of reaction norm parameters where 
+there is a positive correlation between the size of the badge and the quality 
+of the individual. This relationship however is muddle by the fact that there 
+are two clusters of values for the slope of the reaction norm ($\beta$) 
+in the population. Thus, there are two types of reaction norms. One with 
+a steeper slope, meaning that it expresses a higher badge size for a 
+given quality. This effect of an increased variance in the trait distribution 
+is not only triggered by lower number of interactions. Larger variances are also
+found When we assume a lower speed of learning (data not shown). This suggest
+that, limits to the amount of information that individuals acquire through
+learning, allows the coexistence of different communication strategies.
+
+```{r honestDiv,echo=FALSE,out.width='100%',fig.height=6,fig.align='center',fig.show='hold',fig.cap= "The evolution of diverse costly signals.
+Portrait of the evolutionary dynamics of the sender code with snapsots of both sender and receiver codes just as in fig.
+\\ref{fig:handicap}.
+The middle panels show changes in the distribution of values for $\\alpha$ and $\\beta$ along evolutionary time.
+Panels above and below correspond to snapshots of the sender and receiver codes, respectively, generation time of the snapshots are indicated by the grey lines in the middle panels."}
+
+# knitr::include_graphics(here("Simulations","nIntGroupNormQual_","evolDyn7_nIntGroup2000.png"))
+
+scenario<-"betCostEvol4"#"alphaAct"#"nIntGroupNormQual"#
+runChoi<-1
+
+# Load files -------------------------------------------------------------------
+# Project folder
+listTest<-list.files(here("Simulations",paste0(scenario,"_")),full.names = TRUE)
+evolList<-grep("evolLearn",listTest,value=TRUE)
+indList<-grep("indLearn",listTest,value=TRUE)
+paramName<-list.files(here("Simulations",paste0(scenario,"_")))
+# paramName<-list.files(extSimsDir,full.names = TRUE)
+paramName<-grep(".json",paramName,value=TRUE) %>% grep(pattern = "*5_*",value = TRUE)
+
+param<-fromJSON(here("Simulations",paste0(scenario,"_"),paramName[2]))
+  # fromJSON(paramName[1])
+
+val<-2
+fileId<-val
+
+evolList_runs<-grep(paste0(param$namParam,param$rangParam),
+                    evolList,value =TRUE)
+indList_runs<-grep(paste0(param$namParam,param$rangParam),
+                   indList,value =TRUE)
+
+evol<-do.call(rbind,lapply(evolList_runs,fread))
+pop<-do.call(rbind,lapply(indList_runs, fread))
+
+Valpar<-param$rangParam
+
+nampar<-param$namParam
+
+nCenters<-param$nCenters
+sigSquar<-param$sigSq
+
+# Aesthetic paremeters ---------------------------------------------------------
+
+## Calculate clustering for all generations  -------------------
+
+### Choose which interaction to visualize
+# lastInt<-tail(pop[,unique(nInteract)],4)[2]
+# # pop[,max(nInteract),by=.(seed,time)][,min(V1)]
+# # to get last interaction: tail(pop[,unique(nInteract)],1)
+# 
+# popOneInd<-pop[nInteract==lastInt]
+# vars<-c("alpha","beta","Badge")
+# # ,"Badge",grep("WeightAct",names(popFinal),value = TRUE))
+# popOneInd$idClust<-get_clusters(popOneInd,vars,k.max = 10,
+#                                 Bsamples =500,iterMax = 500)
+# clusSummary<-popOneInd[,.(meanAlph=mean(alpha),meanBet=mean(beta)),
+#                        by=.(idClust,time,seed)]
+# clusSummary[,orderClus:=0]
+# invisible(capture.output(
+#   for(i in 1:dim(clusSummary)[1]){
+#   set(clusSummary,i,6L,match(clusSummary[i,meanAlph],
+#         sort(clusSummary[time==clusSummary[i,time]&
+#                            seed==clusSummary[i,seed]]$meanAlph)))
+# }
+# ))
+# 
+# 
+# popOneInd<-merge(popOneInd,clusSummary,all = TRUE)
+# 
+# 
+# fwrite(popOneInd,file = here("Simulations",paste0(scenario,"_"),
+#                              paste0("popOneInd",nampar,Valpar,".txt")))
+
+popOneInd<-fread(here("Simulations",paste0(scenario,"_"),
+                             paste0("popOneInd",nampar,Valpar,".txt")))
+
+# Get stats from the evolutionary simulations ----------------------------------
+# names(evol)
+
+cols<-c("freqGenHawks","freqGenDove",  "freqGenEval",  #"freqGenLearn",
+        "freqFenHawks", "freqFenDoves", "freqHH", "freqHD", "freqDD", "meanCue",
+        "meanAlpha", "meanBeta",
+        #"meanFit", #"meanInitCrit", "sdInitCrit", "meanInitAct", "sdInitAct",
+        "WeightAct_0","WeightCrit_0",
+        "WeightAct_1","WeightCrit_1","WeightAct_2",  "WeightCrit_2",
+        "WeightAct_3",  "WeightCrit_3", "WeightAct_4",  "WeightCrit_4",
+        "WeightAct_5","WeightCrit_5")
+
+my.summary<- function(x) list(mean = mean(x), lowIQR = fivenum(x)[2], 
+                              upIQR = fivenum(x)[4])
+
+evolStats<-evol[, as.list(unlist(lapply(.SD, my.summary))), 
+                .SDcols = cols,by=time]
+
+## Plot mean and IQRs among individuals for each replicate --------------------
+
+# get the trajectories for individual runs
+traitsTrajs<-dcast(evol,time~seed,
+                   value.var = c("meanAlpha","meanBeta",
+                                 #"meanInitCrit",#"meanFit",
+                                 # "meanInitAct","sdInitCrit","sdInitAct",
+                                 "sdAlpha","sdBeta","freqHH",
+                                 "freqHD","freqDD"))
+
+finReps<-evol[time==max(time),seed]
+
+genstoPrint<-round(seq(1,length(unique(evolStats$time)),length.out = 5))[2:5]
+
+# Evolutionary dynamics for alpha - intercept of the sender --------------------------------------------
+par(plt=posPlot(numploty = 3,idploty = 3,numplotx = 2,idplotx = 1,
+                lowboundx = 8, title = TRUE,lowboundy = 15,
+                upboundy = 99),
+    mfrow=c(1,1), xaxt="s",las=1,mgp=c(2,0.5,0))
+
+adjsYlims<-abs(range(pop[seed==runChoi,alpha])[1]-
+                 range(pop[seed==runChoi,alpha])[2])*c(-0.1,0.2)
+
+evolDist(indData = pop[seed==runChoi],variable = "alpha",nbins = 10,
+         pal = pal_dist,
+         nlevels=10,cexAxis = cexAxis,xlab="",ylab = "",xaxt="n",yaxt = "s",
+         ylim = range(pop[seed==runChoi,alpha])+adjsYlims,
+         xlim=c(0,22000))
+
+mtext(text = "Trait value",cex = cexAxis,line = 1,side = 2,las=0)
+legend("topright",legend = c(expression(alpha[s])),col=colGenesLin[1],lwd=2,
+       bty = "n",cex=cexAxis)
+legend("topleft",legend = c("A"),cex=cexAxis,bty = "n",inset = -0.06)
+
+axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
+     labels = axTicks(1)[2:(length(axTicks(1)))]/1000,cex.axis=cexAxis,
+     hadj = 1.3,line = -0.5,tick = FALSE)
+
+# grey lines to show the generations shown in the upper and lower panels
+
+matlines(x=matrix(rep(evolStats[genstoPrint,time],each=2),nrow=2),
+         y=matrix(rep(range(pop[seed==runChoi,alpha]),4),nrow = 2),
+         lty=1,col = "grey",lwd=2)
+
+# Variation among individuals
+
+matlines(x=traitsTrajs[,time],
+         y=traitsTrajs[,.SD,
+                       .SDcol=paste0(c("meanAlpha_"),runChoi)],
+         col=colGenesLin,lty = 2,type="l",lwd=5)
+
+par(plt=posPlot(numploty = 3,idploty = 3,numplotx = 2, idplotx = 2,
+                lowboundx = 8, title = TRUE,lowboundy = 15,
+                upboundy = 99),
+    xaxt="s",las=1,new=TRUE)
+
+adjsYlims<-abs(range(pop[seed==runChoi,beta])[1]-
+                 range(pop[seed==runChoi,beta])[2])*c(-0.1,0.2)
+
+evolDist(indData = pop[seed==runChoi],variable = "beta",nbins = 10,pal = pal_dist,
+         nlevels=10,cexAxis = 1.5,xlab="",ylab = "",xaxt="n",yaxt = "n",
+         ylim = range(pop[seed==runChoi,beta])+adjsYlims,xlim=c(0,22000))
+legend("topright",legend = c(expression(beta[s])),cex=cexAxis,
+       col=colGenesLin[2],lwd=2,bty = "n")
+legend("topleft",legend = c("B"),cex=cexAxis,bty = "n",inset = -0.06)
+# grey lines to show the generations shown in the upper and lower panels
+matlines(x=matrix(rep(evolStats[genstoPrint,time],each=2),nrow=2),
+         y=matrix(rep(range(pop[seed==runChoi,beta]))),
+                      lty=1,col = "grey",lwd=2)
+
+axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
+     labels = axTicks(1)[2:(length(axTicks(1)))]/1000,cex.axis=cexAxis,
+     hadj = 1.3,line = -0.5,tick = FALSE)
+
+axis(4,cex.axis=cexAxis)
+
+matlines(x=traitsTrajs[,time],
+         y=traitsTrajs[,.SD,
+                       .SDcol=paste0(c("meanBeta_"),runChoi)],
+         col=colGenesLin[2],lty = 2,type="l",lwd=4)
+
+# Choose time range
+gen2plot<-round(seq(1,length(unique(evolStats$time)),length.out = 5))[2:5]
+# Plor the actor
+seqYax<-c("s",rep("n",3))
+# seqYlabUp<-c("Badge",rep("",3))
+seqYlabUp<-c("P(dove)",rep("",3))
+seqYlabDown<-c("Badge",rep("",3))
+seqXlabDown<-c("","Quality","")#paste0("seed: ",runChoi))
+seqXlabUp<-c("","Badge","")
+labPanels<-c("C","D","E","F","G")
+rangQual<-seq(0,1,length.out = 50)
+interv<-1/(nCenters-1)
+centers<-interv*seq(0,nCenters-1)
+rangx<-seq(0,1,length=50)
+count<-0
+for(genC in genstoPrint){
+  count<-count+1
+  par(plt=posPlot(numplotx = 4,numploty = 3,idplotx = count,idploty = 1,
+                  lowboundx = 8, title = TRUE,lowboundy = 10,
+                  upboundy = 87),
+      xaxt="s",las=1,new=TRUE)
+  weightsAct<-as.double(evol[(time==unique(time)[genC])&seed==runChoi,.SD,
+                             .SDcols=grep("WeightAct",
+                                          names(evol),value = TRUE)])
+  tempPop<-popOneInd[(time==unique(time)[genC]&seed==runChoi),.SD[.N],
+               .SDcol=c(grep("WeightAct",
+                             names(evol),value = TRUE),"Quality","alpha","beta",
+                        "orderClus"),
+               by=indId]
+
+  randInds<-sample(tempPop$indId,nInd2print, replace = FALSE)
+
+  dataIndAct<-sapply(as.list(tempPop[indId %in%randInds,indId]),
+                     function(x){x=
+                       logist(totRBF(rangx,
+                                     centers,sigSquar,
+                                     as.double(
+                                       tempPop[indId==x,.SD,
+                                               .SDcol=grep("WeightAct",
+                                                           names(tempPop),
+                                                           value = TRUE)
+                                               ])),alpha=0,beta = 1)})
+  matplot(x=rangx,y=dataIndAct,
+          yaxt=seqYax[count],ylab="",xlab="",type="l",cex.lab=cexAxis,
+          lwd=2,xaxt="n",ylim=c(-0.06,1.1),col = paletteMeans(100)[
+            findInterval(tempPop[indId %in%randInds,Quality],colorbreaksQual)],lty = 1,
+          cex.axis=cexAxis)
+  axis(side = 1,cex.axis=cexAxis,padj = 0)
+  mtext(text = seqYlabUp[count],cex = cexAxis,line = 1.7,side = 2,las=0)
+  mtext(text = seqXlabUp[count],cex = cexAxis,line = 1.2,side = 1 ,las=0)
+  lines(logist(totRBF(rangx,centers,sigSquar,weightsAct),alpha=0,beta=1)~rangx,
+        col=1,lwd=3,lty=2)
+  legend("topleft",legend = c(paste0("D.",count)),bty = "n",inset = -0.06)
+  # text(x=0.5,y=-0.015,labels = paste0("time=",unique(evolStats$time)[genC]/1000),
+  #      cex=2)
+  par(plt=posPlot(numplotx = 4,numploty = 3,idplotx = count,idploty = 2,
+                  lowboundx = 8,lowboundy = 15,
+                  upboundy = 99), las=1,new=TRUE)
+
+# meansClustmp<-tempPop[,.(alphaMean=mean(alpha),betaMean=mean(beta)),by=orderClus]
+# reacNormpClust<-sapply(meansClustmp[,orderClus], function(x){
+#   logist(rangx,alpha = meansClustmp[orderClus==x,alphaMean],
+#          beta = meansClustmp[orderClus==x,betaMean])})
+# matplot(x = rangx,y = reacNormpClust,col=colRuns[meansClustmp$orderClus],
+#          type='l',cex.axis=cexAxis,
+#          xlab="",ylab="",ylim=c(0,1.1),lty=1,
+#          lwd=2,yaxt=seqYax[count],xaxt="s",cex.lab=cexAxis)
+  reacNormInd<-sapply(tempPop[indId %in%randInds,indId], function(x){
+    logist(rangx,alpha = tempPop[indId==x,alpha],
+           beta = tempPop[indId==x,beta])})
+  matplot(x = rangx,y = reacNormInd,
+          type='l',cex.axis=cexAxis,
+          xlab="",ylab="",ylim=c(0,1.1),lty=1,col=1,
+          lwd=2,yaxt=seqYax[count],xaxt="s",cex.lab=cexAxis)
+  mtext(text = seqYlabDown[count],cex = cexAxis,line = 1.5,side = 2,las=0)
+  mtext(text = seqXlabDown[count],cex = cexAxis,line = 1.3,side = 1,las=0)
+  legend("topleft",legend =  c(paste0("C.",count)),bty = "n",inset = -0.06)
+  }
+
+rm(list=grep("temp",ls(),value = T))
+
+```
+
+The behaviour expressed by naive individuals (those who have not learned) 
+imposes negative-frequency dependent selection, allowing for the build up of
+genetic variation in reaction norms. In the simulations, presented so far we 
+assumed that individuals start with a flat behavioural reaction norm such that 
+they escalate aggressively a fight with a $0.5$ probability regardless of 
+the badge size of the interacting partner. To asses if this initial 
+conditions of the communication system had any role 
+in the build up of genetic variation, we ran a series of simulations varying 
+the initial conditions of the actor module in the learning model. Specifically,
+We let naive individuals have a flat reaction norm with a 1) low (peaceful),
+2) low (aggressive) probability of escalating a fight and 3) corresponding to the
+ESS of the original hawk-dove game (clever). Fig. \ref{fig:startCond} shows 
+the distribution of values of the intercept ($\alpha$) and slope ($\beta$) 
+evolved in different replicates of the simulations. The left-hand side panel, 
+that corresponds to peaceful naive individuals, is the only one where the  
+distribution of values is split in different clusters. That is, the majority of 
+the variation occurs within clusters. In contrast, when naive individuals behave
+either aggressively or cleverly, the badges evolve toward minimum and maximum 
+values, and so the variation is driven mainly by among replicate differences. 
+We can make sense of these results by realizing that individuals change their 
+naive behaviour quickly in ranges of the badge size that are common in 
+the population. Thus, one way to repeatedly interact with the naive behaviour
+is to have a rare badge size. When the naive behaviour is peaceful, individual
+with a rare badge size have a fitness advantage. This triggers 
+negative-frequency dependent selection and the evolution of different types of 
+badge sizes. In situations where individuals learn fast and interact repeatedly 
+the strength of frequency dependent selection diminishes. The diversification 
+process presented here is fully dependent on the learning process and how this
+process is implemented.
+
+```{r startCond, echo=FALSE, out.width='100%',fig.align='center',fig.show='hold',fig.cap="The peaceful, the clever and the aggressive.
+Distribution of values of the intercept $\\alpha$ and slope $\\beta$ for individuals at the end of the evolutionary simulations.
+The panels show the three different initial conditions for the behaviour of individuals: peaceful (blue), clever (green) and aggressive (red), see maintext for details.
+Colours indicate the replicate simulation.
+In the inset, the resulting reaction norms corresponding to the intecept and slope values for one of the replicates, which replicate is shown is indicated by the line colour."}
+
+# knitr::include_graphics("../Simulations/initAct_/indVarScatter_.png")
+
+scenario<-"initAct4"
+
+# extSimsDir<-#here("Simulations",paste0(scenario,"_"))
+#   paste0("e:/BadgeSims/",scenario,"_")
+
+# Load files -------------------------------------------------------------------
+
+listTest<-list.files(here("Simulations",paste0(scenario,"_")),full.names = TRUE)
+listTest<-grep(".txt",listTest,value = TRUE)
+
+# (listTest<-list.files(extSimsDir,full.names = TRUE))
+
+sdList<-grep("evol",listTest,value=TRUE)
+indList<-grep("ind",listTest,value=TRUE)
+
+inds<-rbindlist(lapply(indList,filesScenar,scenario,full.name=TRUE),fill = TRUE)
+
+inds[,diffActWeight:=abs(WeightAct_0-WeightAct_4)]
+
+shortSce<-gsub("[^[:alpha:]]",gsub(".txt","",tail(strsplit(indList[1],"_")[[1]],2)[1]),
+               replacement = "")
+
+# Individual variation for different parameters ------------------------------
+
+lastInt<-tail(inds[,unique(nInteract)],2)[1]
+
+popOneInd<-inds[nInteract==lastInt]
+
+# vars<-c("alpha","beta","Badge")
+# popOneInd$idClust<-get_clusters(popOneInd,vars,k.max = 5,
+#                                 Bsamples =500,iterMax = 500)
+
+# repsScen<-c(12,14,13)
+repsScen<-c(5,2,7)
+
+rangx<-seq(0,1,length=50)
+
+list.DataReact<-lapply(popOneInd[,sort(unique(get(shortSce)),decreasing = TRUE)],
+function(x){
+  count<-match(x,popOneInd[,sort(unique(get(shortSce)),decreasing = TRUE)])
+  tempPop<-popOneInd[get(shortSce)==x&(seed==repsScen[count]&time==max(time))]
+  tempPop<-tempPop[,.SD[.N],
+                   .SDcol=c(grep("WeightAct",
+                                 names(popOneInd),value = TRUE),"Quality","alpha","beta"
+                            # ,
+                            # "orderClus"),
+                   ),
+                   by=indId]
+  dataIndReact<-sapply(as.list(tempPop[,indId]),
+                       function(x){x=
+                         sapply(rangx, 
+                                function(y)
+                                  do.call(logist,
+                                          as.list(
+                                            c(y,as.double(
+                                              tempPop[indId==x,.SD,
+                                                      .SDcol=c("alpha","beta")])))))})
+  return(dataIndReact)
+})
+
+names(list.DataReact)<-
+  popOneInd[,sort(unique(get(shortSce)),decreasing = TRUE)]
+
+yaxtAll<-c("s","n","n")
+xlabAll<-c("",expression(beta[s]),"")
+ylabAll<-c(expression(alpha[s]),"","")
+titleAll<-c("Peaceful","Clever","Aggressive")
+colorAll<-c("dodgerblue2","green4","#E31A1C")
+multDiscrPallet[repsScen+1]<-colorAll
+plot.new()
+for(PAr in popOneInd[,sort(unique(get(shortSce)),decreasing = TRUE)]){
+  count<-match(PAr,popOneInd[,sort(unique(get(shortSce)),decreasing = TRUE)])
+  # par(plt=posPlot(numploty = 2,idploty = 2,
+  #                 numplotx = length(popOneInd[,unique(get(shortSce))]),
+  #                 idplotx = count)+c(0,0,-0.05,-0.05),
+  #     las=1,new=TRUE,
+  #     yaxt=yaxtAll[match(PAr,popOneInd[,sort(unique(get(shortSce)),decreasing = TRUE)])],
+  #     cex.axis=1.5,new=TRUE)
+  # plot(data=popOneInd[(get(shortSce)==PAr&time>max(time)*0.8)&seed %in% 12:15],
+  #      alpha~beta,ylab="",
+  #      xlab="", pch=20,cex.lab=3,cex.axis=2,las=1,cex=2,xaxt="n",
+  #      ylim=range(popOneInd[,alpha])*c(1.05,1.05),
+  #      xlim=range(popOneInd[,beta]),
+  #      # col=alpha(multDiscrPallet[seed+1],0.3))
+  #      col=multDiscrPallet[seed+1])
+  # # points(data=popOneInd[(get(shortSce)==PAr&time>max(time)*0.8)&
+  # #                         seed == repsScen[count]],
+  # #      alpha~beta,col=multDiscrPallet[seed+1],cex=3,pch=20)
+  # lines(x=c(0,0),y=range(popOneInd[,alpha]),col="grey",
+  #       lwd=2)
+  # lines(y=c(0,0),x=range(popOneInd[,beta]),col="grey",
+  #       lwd=2)
+
+  par(plt=posPlot(numploty = 1,idploty = 1,
+                  numplotx = length(popOneInd[,unique(get(shortSce))]),
+                  idplotx = count,upboundx = 98,lowboundx = 13)
+      +c(-0.03,-0.03,-0.025,-0.025),
+      las=1,new=TRUE,yaxt=yaxtAll[count],
+      cex.axis=1.5,new=TRUE)
+  plot(data=popOneInd[get(shortSce)==PAr&time>max(time)*0.8],alpha~beta,ylab="",
+       xlab="", pch=20,cex.lab=cexAxis,cex.axis=cexAxis,las=1,cex=1,
+       ylim=range(popOneInd[,alpha])*c(1.1,1.05),
+       xlim=range(popOneInd[,beta]),col=multDiscrPallet[seed+1])
+  lines(x=c(0,0),y=range(popOneInd[,alpha]),col="grey",
+        lwd=2)
+  lines(y=c(0,0),x=range(popOneInd[,beta]),col="grey",
+        lwd=2)
+  mtext(text = xlabAll[count],side = 1,line = 2,cex = cexAxis)
+  mtext(text = ylabAll[count],side = 2,line = 2,las=1,cex=cexAxis)
+  mtext(text = titleAll[count],side = 3,cex = cexAxis,line = 0.2,
+        col = colorAll[count])
+
+  points(data=popOneInd[(get(shortSce)==PAr&time>max(time)*0.8)&
+                          seed == repsScen[count]],
+       alpha~beta,col=multDiscrPallet[seed+1],cex=2,pch=20)
+
+}
+for(PAr in popOneInd[,sort(unique(get(shortSce)),decreasing = TRUE)]){
+  count<-match(PAr,popOneInd[,sort(unique(get(shortSce)),decreasing = TRUE)])
+
+  par(new=TRUE,plt=posPlot(numplotx = 6,numploty = 4,idplotx = count*2-1,
+                           idploty = 1,upboundx = 96,lowboundx =11,
+                           upboundy = 98)
+      +c(0.02,0.02,0.03,0.03),xpd=T,xaxt="s",yaxt="s")
+
+  plot(x=0,y=0,type='l',xlab="", ylab="",ylim=c(0,1),xlim=c(0,1),
+       xaxt="n",yaxt="n")
+  polygon(x=c(par("usr")[1],par("usr")[1],par("usr")[2],par("usr")[2]),
+          y=c(par("usr")[3],par("usr")[4],par("usr")[4],par("usr")[3]),
+          col = "white",border=T)
+  matlines(x=rangx,y=list.DataReact[[count]],
+    #        col = paletteMeans(100)[
+    # findInterval(tempPop[,Quality],colorbreaksQual)],
+    col=alpha(multDiscrPallet[repsScen[count]+1],1),
+    type='l',xlab="",
+    ylab="",ylim=c(0,1),lty=1,lwd=2,new=T)
+  axis(1,at = c(0,1),tick = TRUE,cex.axis=.7,padj=-1.7)
+  axis(2,at =c(0,1),tick = TRUE,cex.axis=.7,hadj=-0.2)
+}
 
@@
 
-```{r noBadge,echo=FALSE,warning=FALSE,out.width='90%',fig.align='center',fig.show='hold',fig.cap= "The evolution of cheap signals.
+```{r noBadge,echo=FALSE,warning=FALSE,out.width='100%',fig.height=6,fig.align='center',fig.show='hold',fig.cap= "The evolution of cheap signals.
 Portrait of the evolutionary dynamics of the sender code with snapsots of both sender and receiver codes just as in fig.
 \\ref{fig:handicap}.
-The middle panels show changes in the distribution of values fir $\\alpha$ and $\\beta$ along evolutionary time.
-Panels above and bellow correspond to snapshots of the sender and receiver codes, respectively, generation time of the snapshots are indicated by the grey lines in the middle panels."}
-
-knitr::include_graphics(here("Simulations","betCostEvol1_","evolDyn5_betCost5.png"))
+The middle panels show changes in the distribution of values for $\\alpha$ and $\\beta$ along evolutionary time.
+Panels above and below correspond to snapshots of the sender and receiver codes, respectively, generation time of the snapshots are indicated by the grey lines in the middle panels."}
+
+# knitr::include_graphics(here("Simulations","betCostEvol1_","evolDyn5_betCost5.png"))
+
+scenario<-"betCostEvol3"#"alphaAct"#"nIntGroupNormQual"#
+runChoi<-12 #1
+
+# Load files -------------------------------------------------------------------
+# Project folder
+listTest<-list.files(here("Simulations",paste0(scenario,"_")),full.names = TRUE)
+evolList<-grep("evolLearn",listTest,value=TRUE)
+indList<-grep("indLearn",listTest,value=TRUE)
+
+val<-2
+fileId<-val
+
+paramName<-list.files(here("Simulations",paste0(scenario,"_")))
+paramName<-grep(".json",paramName,value=TRUE) %>% grep(pattern = "*s0_*",value = TRUE)
+
+param<-fromJSON(here("Simulations",paste0(scenario,"_"),paramName[1]))
+
+evolList_runs<-grep(paste0(param$namParam,param$rangParam),
+                    evolList,value =TRUE)
+indList_runs<-grep(paste0(param$namParam,param$rangParam),
+                   indList,value =TRUE)
+
+evol<-do.call(rbind,lapply(evolList_runs,fread))
+pop<-do.call(rbind,lapply(indList_runs, fread))
+
+Valpar<-param$rangParam[1]
+
+nampar<-param$namParam
+
+nCenters<-param$nCenters
+sigSquar<-param$sigSq
+
+# Aesthetic paremeters ---------------------------------------------------------
+
+## Calculate clustering for all generations  -------------------
+
+runChoi<-9#3
+## Choose which interaction to visualize
+# lastInt<-tail(pop[,unique(nInteract)],4)[2]
+# # pop[,max(nInteract),by=.(seed,time)][,min(V1)]
+# # to get last interaction: tail(pop[,unique(nInteract)],1)
+# 
+# popOneInd<-pop[nInteract==lastInt]
+# vars<-c("alpha","beta","Badge")
+# # ,"Badge",grep("WeightAct",names(popFinal),value = TRUE))
+# popOneInd$idClust<-get_clusters(popOneInd,vars,k.max = 10,
+#                                 Bsamples =500,iterMax = 500)
+# clusSummary<-popOneInd[,.(meanAlph=mean(alpha),meanBet=mean(beta)),
+#                        by=.(idClust,time,seed)]
+# clusSummary[,orderClus:=0]
+# invisible(capture.output(
+#   for(i in 1:dim(clusSummary)[1]){
+#   set(clusSummary,i,6L,match(clusSummary[i,meanAlph],
+#         sort(clusSummary[time==clusSummary[i,time]&
+#                            seed==clusSummary[i,seed]]$meanAlph)))
+# }
+# ))
+# 
+# 
+# popOneInd<-merge(popOneInd,clusSummary,all = TRUE)
+# 
+# 
+# fwrite(popOneInd,file = here("Simulations",paste0(scenario,"_"),
+#                              paste0("popOneInd",nampar,Valpar,".txt")))
+
+popOneInd<-fread(here("Simulations",paste0(scenario,"_"),
+                             paste0("popOneInd",nampar,Valpar,".txt")))
+
+# Get stats from the evolutionary simulations ----------------------------------
+# names(evol)
+
+cols<-c("freqGenHawks","freqGenDove",  "freqGenEval",  #"freqGenLearn",
+        "freqFenHawks", "freqFenDoves", "freqHH", "freqHD", "freqDD", "meanCue",
+        "meanAlpha", "meanBeta",
+        #"meanFit", #"meanInitCrit", "sdInitCrit", "meanInitAct", "sdInitAct",
+        "WeightAct_0","WeightCrit_0",
+        "WeightAct_1","WeightCrit_1","WeightAct_2",  "WeightCrit_2",
+        "WeightAct_3",  "WeightCrit_3", "WeightAct_4",  "WeightCrit_4",
+        "WeightAct_5","WeightCrit_5")
+
+my.summary<- function(x) list(mean = mean(x), lowIQR = fivenum(x)[2], 
+                              upIQR = fivenum(x)[4])
+
+evolStats<-evol[, as.list(unlist(lapply(.SD, my.summary))), 
+                .SDcols = cols,by=time]
+
+## Plot mean and IQRs among individuals for each replicate --------------------
+
+# get the trajectories for individual runs
+traitsTrajs<-dcast(evol,time~seed,
+                   value.var = c("meanAlpha","meanBeta",
+                                 #"meanInitCrit",#"meanFit",
+                                 # "meanInitAct","sdInitCrit","sdInitAct",
+                                 "sdAlpha","sdBeta","freqHH",
+                                 "freqHD","freqDD"))
+
+finReps<-evol[time==max(time),seed]
+
+genstoPrint<-round(seq(1,length(unique(evolStats$time)),length.out = 5))[2:5]
+
+# Evolutionary dynamics for alpha - intercept of the sender --------------------------------------------
+par(plt=posPlot(numploty = 3,idploty = 3,numplotx = 2,idplotx = 1,
+                lowboundx = 8, title = TRUE,lowboundy = 15,
+                upboundy = 99),
+    mfrow=c(1,1), xaxt="s",las=1,mgp=c(2,0.5,0))
+
+adjsYlims<-abs(range(pop[seed==runChoi,alpha])[1]-
+                 range(pop[seed==runChoi,alpha])[2])*c(-0.1,0.2)
+
+evolDist(indData = pop[seed==runChoi],variable = "alpha",nbins = 10,
+         pal = pal_dist,
+         nlevels=10,cexAxis = cexAxis,xlab="",ylab = "",xaxt="n",yaxt = "s",
+         ylim = range(pop[seed==runChoi,alpha])+adjsYlims,
+         xlim=c(0,22000))
+
+mtext(text = "Trait value",cex = cexAxis,line = 1,side = 2,las=0)
+legend("topright",legend = c(expression(alpha[s])),col=colGenesLin[1],lwd=2,
+       bty = "n",cex=cexAxis)
+legend("topleft",legend = c("A"),cex=cexAxis,bty = "n",inset = -0.06)
+
+axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
+     labels = axTicks(1)[2:(length(axTicks(1)))]/1000,cex.axis=cexAxis,
+     hadj = 1.3,line = -0.5,tick = FALSE)
+
+# grey lines to show the generations shown in the upper and lower panels
+
+matlines(x=matrix(rep(evolStats[genstoPrint,time],each=2),nrow=2),
+         y=matrix(rep(range(pop[seed==runChoi,alpha]),4),nrow = 2),
+         lty=1,col = "grey",lwd=2)
+
+# Variation among individuals
+
+matlines(x=traitsTrajs[,time],
+         y=traitsTrajs[,.SD,
+                       .SDcol=paste0(c("meanAlpha_"),runChoi)],
+         col=colGenesLin,lty = 2,type="l",lwd=5)
+
+par(plt=posPlot(numploty = 3,idploty = 3,numplotx = 2, idplotx = 2,
+                lowboundx = 8, title = TRUE,lowboundy = 15,
+                upboundy = 99),
+    xaxt="s",las=1,new=TRUE)
+
+adjsYlims<-abs(range(pop[seed==runChoi,beta])[1]-
+                 range(pop[seed==runChoi,beta])[2])*c(-0.1,0.2)
+
+evolDist(indData = pop[seed==runChoi],variable = "beta",nbins = 10,pal = pal_dist,
+         nlevels=10,cexAxis = 1.5,xlab="",ylab = "",xaxt="n",yaxt = "n",
+         ylim = range(pop[seed==runChoi,beta])+adjsYlims,xlim=c(0,22000))
+legend("topright",legend = c(expression(beta[s])),cex=cexAxis,
+       col=colGenesLin[2],lwd=2,bty = "n")
+legend("topleft",legend = c("B"),cex=cexAxis,bty = "n",inset = -0.06)
+# grey lines to show the generations shown in the upper and lower panels
+matlines(x=matrix(rep(evolStats[genstoPrint,time],each=2),nrow=2),
+         y=matrix(rep(range(pop[seed==runChoi,beta]))),
+                      lty=1,col = "grey",lwd=2)
+
+axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
+     labels = axTicks(1)[2:(length(axTicks(1)))]/1000,cex.axis=cexAxis,
+     hadj = 1.3,line = -0.5,tick = FALSE)
+
+axis(4,cex.axis=cexAxis)
+
+matlines(x=traitsTrajs[,time],
+         y=traitsTrajs[,.SD,
+                       .SDcol=paste0(c("meanBeta_"),runChoi)],
+         col=colGenesLin[2],lty = 2,type="l",lwd=4)
+
+# Choose time range
+gen2plot<-round(seq(1,length(unique(evolStats$time)),length.out = 5))[2:5]
+# Plor the actor
+seqYax<-c("s",rep("n",3))
+# seqYlabUp<-c("Badge",rep("",3))
+seqYlabUp<-c("P(dove)",rep("",3))
+seqYlabDown<-c("Badge",rep("",3))
+seqXlabDown<-c("","Quality","")#paste0("seed: ",runChoi))
+seqXlabUp<-c("","Badge","")
+labPanels<-c("C","D","E","F","G")
+rangQual<-seq(0,1,length.out = 50)
+interv<-1/(nCenters-1)
+centers<-interv*seq(0,nCenters-1)
+rangx<-seq(0,1,length=50)
+count<-0
+for(genC in genstoPrint){
+  count<-count+1
+  par(plt=posPlot(numplotx = 4,numploty = 3,idplotx = count,idploty = 1,
+                  lowboundx = 8, title = TRUE,lowboundy = 10,
+                  upboundy = 87),
+      xaxt="s",las=1,new=TRUE)
+  weightsAct<-as.double(evol[(time==unique(time)[genC])&seed==runChoi,.SD,
+                             .SDcols=grep("WeightAct",
+                                          names(evol),value = TRUE)])
+  tempPop<-popOneInd[(time==unique(time)[genC]&seed==runChoi),.SD[.N],
+               .SDcol=c(grep("WeightAct",
+                             names(evol),value = TRUE),"Quality","alpha","beta",
+                        "orderClus"),
+               by=indId]
+
+  randInds<-sample(tempPop$indId, nInd2print, replace = FALSE)
+
+  dataIndAct<-sapply(as.list(tempPop[indId %in%randInds,indId]),
+                     function(x){x=
+                       logist(totRBF(rangx,
+                                     centers,sigSquar,
+                                     as.double(
+                                       tempPop[indId==x,.SD,
+                                               .SDcol=grep("WeightAct",
+                                                           names(tempPop),
+                                                           value = TRUE)
+                                               ])),alpha=0,beta = 1)})
+  matplot(x=rangx,y=dataIndAct,
+          yaxt=seqYax[count],ylab="",xlab="",type="l",cex.lab=cexAxis,
+          lwd=2,xaxt="n",ylim=c(-0.06,1.1),col = paletteMeans(100)[
+            findInterval(tempPop[indId %in%randInds,Quality],
+                         colorbreaksQual)],lty = 1,
+          cex.axis=cexAxis)
+  axis(side = 1,cex.axis=cexAxis,padj = 0)
+  mtext(text = seqYlabUp[count],cex = cexAxis,line = 1.7,side = 2,las=0)
+  mtext(text = seqXlabUp[count],cex = cexAxis,line = 1.2,side = 1 ,las=0)
+  lines(logist(totRBF(rangx,centers,sigSquar,weightsAct),alpha=0,beta=1)~rangx,
+        col=1,lwd=3,lty=2)
+  legend("topleft",legend = c(paste0("D.",count)),bty = "n",inset = -0.06)
+  # text(x=0.5,y=-0.015,labels = paste0("time=",unique(evolStats$time)[genC]/1000),
+  #      cex=2)
+  par(plt=posPlot(numplotx = 4,numploty = 3,idplotx = count,idploty = 2,
+                  lowboundx = 8,lowboundy = 15,
+                  upboundy = 99), las=1,new=TRUE)
+
+# meansClustmp<-tempPop[,.(alphaMean=mean(alpha),betaMean=mean(beta)),by=orderClus]
+# reacNormpClust<-sapply(meansClustmp[,orderClus], function(x){
+#   logist(rangx,alpha = meansClustmp[orderClus==x,alphaMean],
+#          beta = meansClustmp[orderClus==x,betaMean])})
+# matplot(x = rangx,y = reacNormpClust,col=colRuns[meansClustmp$orderClus],
+#          type='l',cex.axis=cexAxis,
+#          xlab="",ylab="",ylim=c(0,1.1),lty=1,
+#          lwd=2,yaxt=seqYax[count],xaxt="s",cex.lab=cexAxis)
+  reacNormInd<-sapply(tempPop[indId %in%randInds,indId], function(x){
+    logist(rangx,alpha = tempPop[indId==x,alpha],
+           beta = tempPop[indId==x,beta])})
+  matplot(x = rangx,y = reacNormInd,
+          type='l',cex.axis=cexAxis,
+          xlab="",ylab="",ylim=c(0,1.1),lty=1,col=1,
+          lwd=2,yaxt=seqYax[count],xaxt="s",cex.lab=cexAxis)
+  mtext(text = seqYlabDown[count],cex = cexAxis,line = 1.5,side = 2,las=0)
+  mtext(text = seqXlabDown[count],cex = cexAxis,line = 1.3,side = 1,las=0)
+  legend("topleft",legend =  c(paste0("C.",count)),bty = "n",inset = -0.06)
+  }
+
+rm(list=grep("temp",ls(),value = T))
 
@@
 
+\newpage
+\small
+
+| Symbol |      Description |
+|:-------:|:--------------------------------|
+| $Q_i$ | Quality of individual $i$ |
+| $N$   | Population size |
+| $\sigma$ | Standard deviation of the truncated normal distribution from which quality is drawn |
+| $B_i$ | Badge size of individual $i$ |
+| $\alpha_i$ | Intercept of the badge size reaction norm for individual $i$ |
+| $\beta_i$  | Slope of the badge size reaction norm for individual $i$ |
+| $\mu$     | Mutation rate | 
+| $\sigma_\mu$ |Standard deviation of the normal distribution from which quality is mutations are drawn |
+| $s_i$ | Survival probability of individual $i$ |
+| $k_1$ | Intercept of the survival function |
+| $k_2$ | Slope of the survival function |
+| $V$   | Value of the contested resource |
+| $C$   | Cost of an escalated fight |
+| $p_{ij}$ | Probability that individual $i$ wins an escalated fight against individual $j$ |
+| $k_3$ | Parameter determining how important is quality defining the probability of wining |
+| $\delta$ | Prediction error |
+| $A$ | Speed of learning | 
+| $c$ | Number of feature centres |
+| $b_z$ | badge size for feature centre $z$  | 
+| $\hat{R}$ | Reward estimate | 
+| $x_z$ | weight of feature centre $z$ on the reward estimation |
+| $\theta$ | width of the generalization function | 
+| $q_{ij}$ | probability that individual $i$ retreats from a fight in an encounter with individual $j$ |
+| $y_z$ | weight of feature centre z on the tendency to retreat |
+| $w_i$ | total pay-off of individual $i$ |
+| $w_0$ | base line pay-off |
+
+Table: Notation of the model parameters and variables
+
 ## References

### 2023-02-01 Dyn_reac_learn.R

@@ -13,7 +13,7 @@ library(dplyr)
 
 # Scenario to be plotted - corresponds to folders where simulations are stored
 
-scenario<-"betCostalphaL2"
+scenario<-"initAct5"
 
 # extSimsDir<-here("Simulations",paste0(scenario,"_"))
   # paste0("e:/BadgeSims/",scenario,"_")
@@ -49,7 +49,7 @@ val<-1
 
 
 # loop to produce pdfs for parameter values
-numCores <- length(param$rangParam)
+numCores <- length(paramRang)
 registerDoParallel(numCores)
 
 foreach(val = 1:length(paramRang),

### 2023-02-01 Figure_dyn_AGA.R

@@ -373,7 +373,7 @@ dev.off()
   
 ## Effect of the alternative equilibria in behavioural interactions ----------
 
-indVarScatter
+#indVarScatter
 evolFinal<-evol[time>max(time)*0.8]
 
 names(evolFinal)

### 2023-03-10 manuscript/manuscript_1.0.Rmd
New version of the discussion finished. Started addressing Daniel´s comments

@@
   pdf_document:
-    latex_engine: xelatex  # word_document: default
+    latex_engine: xelatex
+  word_document: default
 fig_caption: yes
 fontsize: 12pt
-header-includes: |
-  ```{=latex}
-    \usepackage{float}
-    \usepackage{unicode-math}
-    \usepackage{setspace}\doublespacing
-    \usepackage[backref=page]{hyperref}
-    \linespread{2}
-    \usepackage{lineno}
-    \linenumbers
-    \floatplacement{figure}{H}
-    \newcommand{\beginsupplement}{ \setcounter{table}{0}         \renewcommand{\thetable}{S\arabic{table}}\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
-  ```
+header-includes: "```{=latex}\n  \\usepackage{float}\n  \\usepackage{unicode-math}\n
+  \ \\usepackage{setspace}\\doublespacing\n  \\usepackage[backref=page]{hyperref}\n
+  \ \\linespread{2}\n  \\usepackage{lineno}\n  \\linenumbers\n  \\floatplacement{figure}{H}\n
+  \ \\newcommand{\\beginsupplement}{ \\setcounter{table}{0}         \\renewcommand{\\thetable}{S\\arabic{table}}\\setcounter{figure}{0}
+  \\renewcommand{\\thefigure}{S\\arabic{figure}}}\n```\n"
 keep_tex: yes
@@
 style: C:\Users\andre\Zotero\styles\elife.csl
-abstract: | 
-  Adaptive behavioural responses often depend on qualities of the interacting partner of an individual.
-For example, when competing for resources, an individual might be better off escalating fights with individuals of lower quality, while restraining from fighting individuals of higher quality.
-Communication systems involving signals of quality allow individuals to reduce uncertainty regarding the fighting ability of their partners and make more adaptive behavioral decisions.
-However, dishonest individuals can destabilize such communications systems.
-One open question is whether cognitive mechanisms, such as learning, can maintain the honesty of signals, thus, favoring their evolutionary stability.
-Here we present evolutionary simulations, where individuals can produce a signal proportional to their quality; and learn along their lifetime the best response to the signal emitted by their peers.
-In these simulations, learning on the receiver side can mediate the evolution of signals of quality on the sender side.
-When the cost of the signal is proportional to the quality of the sender populations are only composed of honest signalers.
-When the cost is not proportional to the quality of the signaler, the population is composed of both honest and dishonest signalers.
-We argue that learning can be a general cognitive mechanism playing a role in a wide range of communication systems.
+abstract: "Adaptive behavioural responses often depend on qualities of the interacting
+  partner of an individual. For example, when competing for resources, an individual
+  might be better off escalating fights with individuals of lower quality, while restraining
+  from fighting individuals of higher quality. Communication systems involving signals
+  of quality allow individuals to reduce uncertainty regarding the fighting ability
+  of their partners and make more adaptive behavioral decisions. However, dishonest
+  individuals can destabilize such communications systems. One open question is whether
+  cognitive mechanisms, such as learning, can maintain the honesty of signals, thus,
+  favoring their evolutionary stability. Here we present evolutionary simulations,
+  where individuals can produce a signal proportional to their quality; and learn
+  along their lifetime the best response to the signal emitted by their peers. In
+  these simulations, learning on the receiver side can mediate the evolution of signals
+  of quality on the sender side. When the cost of the signal is proportional to the
+  quality of the sender populations are only composed of honest signalers. When the
+  cost is not proportional to the quality of the signaler, the population is composed
+  of both honest and dishonest signalers. We argue that learning can be a general
+  cognitive mechanism playing a role in a wide range of communication systems.  \n"
 ---
@@
 
-The outcome of interactions with con-specifics is an important 
+The outcome of interactions with conspecifics individuals is an important 
 determinant of fitness in social animals. Irrespective of whether 
@@
 their reproductive success. However, the best action for an individual 
-might vary depending on its condition, and that of individuals with 
+engaging in a social interaction might vary depending on its 
+own condition, and that of individuals with 
 whom it interacts. Thus, acting based on information about interacting 
 partners is typically adaptive [@quinones_Negotiation_2016].
-Acquiring this information, however, is far from trivial. 
+Acquiring such information, however, is far from trivial. 
 In some cases, interacting partners (signallers) might be 'willing' to 
@@
 own interest to conceal information [@johnstone_Recognition_1997], 
-or to provide erroneous information [@johnstone_Badges_1993]. 
-Typically, in a given context, some proportion of individuals will benefit 
-from broadcasting accurate information, whereas others will benefit 
-from hiding it. Take, for instance, an interaction between two individuals 
+or to provide deceptive information [@johnstone_Badges_1993]. 
+Typically, in a given context, some individuals benefit 
+from broadcasting accurate information, whereas others benefit 
+from concealing it. Take, for instance, an interaction between two individuals 
 where one can help the other. Given some costs and benefits, 
-a potential donor will be interested in helping 
+a potential donor would be interested in helping 
 related individuals. Therefore, for a relative of the donor, 
-broadcasting kinship is advantageous, whereas for an unrelated individual, 
-concealing the lack of kinship is better. Similar scenarios may apply to 
-many aspects of social life such as finding mates, 
-feeding offspring, dominance relationships and aggressive interactions 
+broadcasting kinship would be advantageous, whereas concealing the lack 
+of kinship would be better for an unrelated individual. 
+Similar scenarios may apply to a variety of aspects of social 
+life such as finding mates, feeding offspring, or engaging in
+dominance relationships or aggressive interactions 
 [@bradbury_Principles_2011; @moller_Female_1988; @tibbetts_Socially_2004].
-In dominance and aggressive interactions, an important piece of
-information to guide individual actions is interacting 
-partners fighting ability, which is sometimes referred to as 
-Resource Holding Potential (RHP) 
-[@parker_Assessment_1974], or simply as quality. 
-Responsiveness to the quality of the partner is central 
-to communication systems such as badges of status, 
-where an arbitrary signal conveys quality 
+In dominance and aggressive interactions, a crucial piece of
+information to guide individual actions is the fighting ability of interacting 
+partners, which is sometimes referred to as Resource Holding Potential 
+[RHP, Parker -@parker_Assessment_1974], or simply as quality. 
+Responsiveness to the quality of an opponent is central 
+to communication systems such as those in which individuals 
+exhibit badges of status, where a signal conveys quality 
 [@johnstone_Badges_1993; @rohwer_Social_1975a]. 
-Here, arbitrary refers to a signal which is not 
+Such signals must be arbitrary, meaning they are not 
 ontogenetically correlated with quality. 
-Badges of status can be evolutionarily stable whenever the 
-signal imposes a fitness cost which decreases with the 
-quality of an individual  [@botero_Evolution_2010; @johnstone_Badges_1993], 
-such that it's no longer in the interest of low quality individuals 
-to fake quality by producing a large badge. The cost of honest signals 
-is a more general principle of communications systems, 
-usually referred to as the handicap principle 
-[@grafen_Biological_1990; @zahavi_Mate_1975]. 
+Generally signals can be evolutionarily stable whenever low quality individuals 
+pay a higher fitness cost for carrying the signal. 
+So, it's no longer in the interest of low quality individuals 
+to fake quality by producing a large badge. Signal costs inversely proportional
+to quality con come about by the production costs of the signal, 
+in which case the signal works like a handicap [@botero_Evolution_2010; @johnstone_Badges_1993;@grafen_Biological_1990; @zahavi_Mate_1975]. 
+Alternatively, the costs can come from 
+the aggressive reaction of receivers when the convention of the communication 
+is broken [@enquist_Signaling_2010;@tibbetts_Socially_2004]. 
 <!-- should this last part of the paragraph stay ?-->
@@
 
+Continue here!!!!
+
 A somewhat ignored component of communication systems, in the context 
@@
 individuals associate cue of their peers to their quality. These association 
-must be learn thoughout individuals lifes, thus it has the usual 
+must be learned throughout and individual's life, thus it has the usual 
 cognitive requirements of an associative learning process. An alternative view 
@@
 mechanism whose underpinnings show interspecific variation 
-[@enquist_Power_2016; @quinones_Reinforcement_2019], so presumably, it has been modified 
-by natural selection. Despite all this, associative learning is 
+[@enquist_Power_2016; @quinones_Reinforcement_2019]. Associative learning 
+is not a single mechanisms, but rather a set of cognitive structures 
+and processes that can vary in their scope and complexity. Presumably, 
+these structures and processes have been modified 
+by natural selection and could provide new explanations for behavioural 
+variation. Despite all this, associative learning is 
 not often included in the narrative of evolutionary explanations of 
@@
 generate theoretical predictions of the effect that learning can 
-have in behavioural evolution [@leimar_Learning_2019].
+have in behavioural evolution [@leimar_Learning_2019;@leimar_Effects_2022].
 
@@
 Unless, with a small probability ($\mu$), mutation changes the allelic value 
-of the offspring by an amount taken from a normal distribution $N(0,\sigma_\mu)$.
+of the offspring by an amount taken from a normal distribution $TN(0,\sigma_\mu)$.
 After birth individuals go through a round of 
@@
 Individuals who survive engage in a series of pairwise interactions 
-where they compete for resources. Individuals in real populations typically 
-do not interact at random nor with all individuals in the population 
-[@kurvers_Evolutionary_2014]. In order to allow for non-random interactions, 
-we use the population vector to define an interaction neighbourhood 
-for the focal individual, the size of the neighbourhood ($g$)
-is the number of positions around the focal's from which the partner is drawn. 
-So, if $g=N$ interactions are global; as $g$ gets smaller 
-interactions are more local. In an interaction each individual must 
+where they compete for resources. 
+<!-- Individuals in real populations typically  -->
+<!-- do not interact at random nor with all individuals in the population  -->
+<!-- [@kurvers_Evolutionary_2014]. In order to allow for non-random interactions,  -->
+<!-- we use the population vector to define an interaction neighbourhood  -->
+<!-- for the focal individual, the size of the neighbourhood ($g$) -->
+<!-- is the number of positions around the focal's from which the partner is drawn.  -->
+<!-- So, if $g=N$ interactions are global; as $g$ gets smaller  -->
+<!-- interactions are more local.  -->
+In an interaction each individual must 
 decide whether to escalate a fight or not. 
@@
 to retreat are calculated (in every interaction) as the sum of the weights 
-associated with each feature, weighted by the response 
+associated with each feature centre, weighted by the response 
 triggered by the feature (Fig. \ref{fig:model-struc} B, dots 
@@
           lwd=3,ylim=c(0,1.1))
-text(x = 0.08,y=1.05,"C",cex = 1.5)
+text(x = 0.08,y=1.05,"B",cex = 1.5)
 par(new=FALSE)
@@
 ```{r handicap,echo=FALSE,out.width='100%',fig.height=6,fig.align='center',fig.show='hold',fig.cap="Evolution of the badge size as a handicap mediated by learning.
-Middle panels show the evolutionary dynamics of the sender code.
-On the left, changes in the distribution of values for the intercept of the reaction norm ($\\alpha$); on the right changes in the distribution of the slope ($\\beta$).
-Dashed lines in both panels show the mean of the distribution.
-Grey lines show the generation time corresponding to the panels above and below showing the sender and receiver code.
-In the panels above, the reaction norms correspond to individuals in the population after the interaction round.
+Top panels (A.1 and A.2) show the evolutionary dynamics of the sender code.
+On the left (A.1), changes in the distribution of values for the intercept of the reaction norm ($\\alpha$); on the right (A.2) changes in the distribution of the slope ($\\beta$).
+Darker areas of the background correspond to values with high frequency, while dashed lines in both panels show the mean of the distribution.
+X axis in the evolutionary dynamics are given in thousands of generations.
+Grey lines show the generation time corresponding to the panels below portraing the sender (B) and receiver code(C).
+In the bottom panels (C1-4), the learned reaction norms correspond to individuals in the population after the interaction round.
 Color scale indicates quality just as in Fig.
-\\ref{fig:model-struc}.
-In the panels below, colors represent a cluster clasification performed by the k-means algorithm" ,warning=FALSE}
+\\ref{fig:model-struc}." ,warning=FALSE}
 
@@
        bty = "n",cex=cexAxis)
-legend("topleft",legend = c("A"),cex=cexAxis,bty = "n",inset = -0.06)
+legend("topleft",legend = c("A.1"),cex=cexAxis,bty = "n",inset = -0.06)
 
@@
        col=colGenesLin[2],lwd=2,bty = "n")
-legend("topleft",legend = c("B"),cex=cexAxis,bty = "n",inset = -0.06)
+legend("topleft",legend = c("A.2"),cex=cexAxis,bty = "n",inset = -0.06)
 # grey lines to show the generations shown in the upper and lower panels
@@
 seqXlabUp<-c("","Badge","")
-labPanels<-c("C","D","E","F","G")
 rangQual<-seq(0,1,length.out = 50)
@@
         col=1,lwd=3,lty=2)
-  legend("topleft",legend = c(paste0("D.",count)),bty = "n",inset = -0.06)
+  legend("topleft",legend = c(paste0("C.",count)),bty = "n",inset = -0.06)
   # text(x=0.5,y=-0.015,labels = paste0("time=",unique(evolStats$time)[genC]/1000),
@@
   mtext(text = seqXlabDown[count],cex = cexAxis,line = 1.3,side = 1,las=0)
-  legend("topleft",legend =  c(paste0("C.",count)),bty = "n",inset = -0.06)
+  legend("topleft",legend =  c(paste0("B.",count)),bty = "n",inset = -0.06)
   }
@@
 proportional to the quality of the individual), the sender code can evolve 
-to produce an honest signal Fig \ref{fig:handicap}). The evolution of 
+to produce an honest signal (Fig. \ref{fig:handicap}). The evolution of 
 the sender code does not happen immediately after the start of the 
 evolutionary process. Instead, the evolutionary dynamics of the reaction 
-norm parameters (Fig. \ref{fig:handicap}, X) show a set of 
+norm parameters (Fig. \ref{fig:handicap}, A) show a set of 
 steps. First, the intercept of the reaction norm ($\alpha$) evolves to 
@@
 At around generation 4000, the slope evolves toward positive values. 
-Which implies that larger badges correlate with higher quality. Receivers 
-learn to react to such correlation, reducing the probability of attack towards 
-individual of larger badges. Hence, natural selection favours larger 
-values of $\beta$, eventually leading to an evolutionary equilibrium where 
-badge size is an honest signal of quality mediated by the learned responses 
-of receivers. The evolutionary trajectory portrayed in figure 
-\ref{fig:handicap} is not the only possible trajectory. If the slope of the 
+Positive values in the slope imply that larger badges correlate 
+with higher quality (Fig. \ref{fig:handicap} B). Receivers learn to react to such 
+correlation, increasing the probability of retreat when facing 
+individual of larger badges (Fig. \ref{fig:handicap} C). Hence, natural selection 
+favours larger values of $\beta$, eventually leading to an evolutionary 
+equilibrium where badge size is an honest signal of quality mediated
+by the learned responses of receivers (Fig. \ref{fig:handicap} B.4 and C.4). 
+The evolutionary trajectory portrayed in figure 
+\ref{fig:handicap} A is not the only possible trajectory. If the slope of the 
 sender reaction norms ($\beta$) evolves toward negative values before badges 
@@
 Thus, badges are only costly and do not provide information about the 
-quality of an individual (Fig. \ref{fig:noBadge} . Eventually, the badge 
-disappears from the population. In contrast, when the badge does not work as a handicap but instead is cost-free, the evolutionary process never leads to the 
-establishment of an honest signal. Instead, subject to genetic drift the 
+quality of an individual (Fig. \ref{fig:noBadge}) . Eventually, the badge 
+disappears from the population. In contrast, when the badge does not 
+work as a handicap but instead is cost-free, the evolutionary process 
+never leads to the establishment of an honest signal. 
+Instead, subject to genetic drift the 
 population either evolves towards the disappearance of the badge, or to the 
@@
 variation present in the population. Fig. \ref{fig:branching} shows 
-changes in the distribution of the sender reaction norm
-along evolutionary time assuming the badge is cost-free. 
-Darker areas show trait values with high frequency in the population. 
+changes in the distribution of the sender reaction norm parameters
+along evolutionary time, assuming the badge is cost-free. 
+In panels A1 and 2, darker areas show trait values with high frequency in the population. 
 Populations start monomorphic with a value of zero on both traits, 
@@
 Later in evolutionary time one of the peaks splits further, 
-so at the end of the evolutionary simulation the distribution of $\alpha$ 
-value in the population shows three peaks. In the case of $\beta$, the 
+so at the end of the evolutionary simulation the distribution of the intercept 
+($\alpha$) in the population shows three peaks. In the case of $\beta$, the 
 peaks in the distribution are not so clear-cut, but it's clear that there is an 
@@
 distribution of the sender reaction norm parameters 
-imply that during the second half of the simulation individuals can be 
-classified into three distinct types. Two types express a flat reaction 
-norm, meaning that their badge size is not informative of their quality, 
-while the third type shows intermediate badge sizes which are determined 
-by the quality of the individual (Fig. \ref{fig:branching}). Furthermore, 
-the receiver reaction norms develop through learning, particularly those 
+imply that individuals can generally be classified into three 
+distinct types (\ref{fig:branching}). Two types express a flat reaction 
+norm with extreme values for the badge size, meaning that their badge size 
+is not informative of their quality, while the third type shows 
+intermediate badge sizes which are determined 
+by the quality of the individual (Fig. \ref{fig:branching} B.3-4). 
+In this particular simulation, the values of the slope in the informative group
+are negative, which implies that badge size correlates negatively with quality. 
+Furthermore, the receiver reaction norms developed through learning, particularly those 
 of individuals of intermediate quality, respond to the signal of their 
 sender type by increasing the probability of retreating from a fight with 
-individuals with larger badges (Fig. \ref{fig:branching}).
+individuals with smaller badges (Fig. \ref{fig:branching}).
 
@@
 \\ref{fig:handicap}.
-The middle panels show changes in the distribution of values fir $\\alpha$ and $\\beta$ along evolutionary time.
-Panels above and below correspond to snapshots of the sender and receiver codes, respectively, generation time of the snapshots are indicated by the grey lines in the middle panels."}
+The top panels (A) show changes in the distribution of values for $\\alpha$ (A.1) and $\\beta$ (A.2) along evolutionary time.
+Panels below correspond to snapshots of the sender (B) and receiver codes (C).
+Generation time of the snapshots are indicated by the grey lines in the top panels."}
 
@@
 the evolving parameters plays a role when the signal follows the handicap
-principle. In fig \ref{fig:honestDiv}, we show simulations where individuals 
-have on average 300 interaction in their life and the cost of the signal
+principle as well. In fig \ref{fig:honestDiv}, we show simulations where individuals 
+have on average 300 interactions in their life and the cost of the signal
 is proportional to the quality, following the handicap principle. 
@@
 there is a positive correlation between the size of the badge and the quality 
-of the individual. This relationship however is muddle by the fact that there 
-are two clusters of values for the slope of the reaction norm ($\beta$) 
-in the population. Thus, there are two types of reaction norms. One with 
-a steeper slope, meaning that it expresses a higher badge size for a 
-given quality. This effect of an increased variance in the trait distribution 
-is not only triggered by lower number of interactions. Larger variances are also
-found When we assume a lower speed of learning (data not shown). This suggest
+of the individual. This relationship however is muddled by the fact that there 
+are two clusters of values for the intercept ($\alpha$) and slope ($\beta$) 
+of the reaction norm in the population. Thus, there are two types 
+of reaction norms. One with a steeper slope, meaning that it expresses 
+a higher badge size for a given quality. This effect of an increased 
+variance in the trait distribution is not only triggered by lower 
+number of interactions. Larger variances are also found When we assume 
+a lower speed of learning (data not shown). This suggest
 that, limits to the amount of information that individuals acquire through
-learning, allows the coexistence of different communication strategies.
+learning, allows the coexistence of different communication strategies 
+[@botero_Evolution_2010].
 
@@
 the initial conditions of the actor module in the learning model. Specifically,
-We let naive individuals have a flat reaction norm with a 1) low (peaceful),
-2) low (aggressive) probability of escalating a fight and 3) corresponding to the
-ESS of the original hawk-dove game (clever). Fig. \ref{fig:startCond} shows 
-the distribution of values of the intercept ($\alpha$) and slope ($\beta$) 
-evolved in different replicates of the simulations. The left-hand side panel, 
+we let naive individuals have a flat reaction norm with a 1) low (peaceful),
+2) high (aggressive) probability of escalating a fight and 3) a probability 
+corresponding to the ESS of the original hawk-dove game (clever). 
+Fig. \ref{fig:startCond} shows the distribution of values of the 
+intercept ($\alpha$) and slope ($\beta$) evolved in different 
+replicates of the simulations. The left-hand side panel, 
 that corresponds to peaceful naive individuals, is the only one where the  
@@
 
+## Discussion
+
+We have presented here an evolutionary model where the response of a signal in a 
+communication system is mediated by learning processes. Individuals 
+throughout their life learn the best way to respond to a quantitative 
+trait of their interactive partners. The learning process can mediate the evolution 
+of an honest signal under the handicap 
+principle [@grafen_Biological_1990;@zahavi_Mate_1975]. This conclusion is not 
+different from the classical genetic models, where responses to the signal are
+inherent to the individual. However, unlike classical genetic models,
+our model shows that learning can also mediate a signal polymorphism in the 
+absence of the handicap principle. Under this polymorphic equilibrium different 
+types of individuals, with honest and dishonest signals, compose the population. 
+The amount of information gathered through learning, as a well as the 
+initial conditions of the learning process, are crucial for the emergence
+and maintenance of this variation. The emergence of the polymorphism requires
+a limit in the amount of information collected by individuals, as well as
+peaceful behaviour in naive individuals.
+
+Associative learning is a powerful mechanism to learn about the world, about 
+social partners and about one's abilities relative to the social context. 
+By associating cues and signals with fitness relevant outcomes individuals
+collect information that allows them to improve their reproductive potential. 
+This is very obvious in classical examples such as when animals learn to avoid 
+food that makes them sick, or use environmental cues to find food. 
+More recent ideas have highlighted the potential role that associations 
+can have in a social contexts such as: hierarchy formation 
+[@leimar_Reproductive_2022;@leimar_Effects_2022],
+and cooperation [@leimar_Learning_2019;@dridi_Learning_2018]. In this 
+theoretical examples individuals use various sources of reward to make
+adaptive choices in a given social context. Here we extend this logic to the 
+evolution of a communication system mediated by a quality signal. 
+Previous evolutionary models of badges of status assumed that individuals 
+responded using a behavioural reaction norm, where the opponent's badge 
+and individual's own quality determined the behaviour [@botero_Evolution_2010]. 
+However, there isn't a clear mechanism to justify the assumption that 
+individuals inherently know their own quality, particularly relative to their peers.
+In our model, individuals not only learn about the quality signal, but also 
+learn about their own quality relative to others. This can be seen throughout
+our simulations in the responses developed by individuals of different quality. 
+Because individuals learn about their own quality, 
+even in the absence of an honest signal, they are able to make adaptive choices. 
+Thus the learning process we have modelled and the response it mediates 
+has fitness relevant consequences even in the absence of an honest signal.
+
+Learning processes collecting information on a population wide level promote 
+frequency dependent selection, in the communication system we model frequency 
+dependency translates to the coexistence of different communication strategies. 
+The frequency-dependence triggered by learning process was highlighted by the  
+classic study of [@bond_Visual_2002] where live predators drive the evolution
+of *in silico* polymorphic prey. The key to the evolution of this polymorphisms is 
+that predators learn to discriminate better preys that are frequently encountered.
+Thus, preys found in low frequency have a selective advantage. 
+An analogous process occurs in our simulations, the learning algorithm of the 
+receiver individual learns the best response for herself towards individuals 
+that have a badge size common in the population. When an individual has a
+rare badge size it can be favoured or unfavoured by the naive behaviour. When 
+the naive behaviour is peaceful, polymorphism is promoted, when the naive 
+behaviour is aggressive polymorphism is prevented. These two examples show how
+behaviours dependent on learning processes seem to generally trigger frequency 
+dependent selection. That is because the learning algorithm collects more 
+information on the frequent values of the the trait distribution. For the sake  
+of simplicity we modelled a single dimension (badge size), however this effect 
+could potentially be more important when individuals learn from 
+multidimensional signals.
+
+Are bibs real signals of quality? if so are they handicaps or badge of status? 
+Originally @rohwer_Social_1975a proposed the idea that certain morphological 
+traits could be used by animals as status signals or signals of quality. 
+However, from an evolutionary perspective it is unclear how the stability of 
+such communication systems can be maintained. One option is that honesty is 
+maintained by the cost of signal production; that is, when the signal works as a 
+handicap [@botero_Evolution_2010;@grafen_Biological_1990;@johnstone_Badges_1993]. 
+Alternatively, if the signal does not carry production costs inversely 
+proportional to quality it may work as a convention.
+In this later case, honesty is supposedly 
+maintained by the aggressive reaction of receivers when the convention is 
+broken [@enquist_Signaling_2010;@tibbetts_Socially_2004]. Our model, portraits 
+both kinds of signals, although with some nuances. When we assume the handicap 
+principle, we get the evolution of honest signals mediated by the learned 
+response. When we assume cost-free signal, under certain conditions the learned 
+response mediated the emergence of phenotypic variation. This variation, 
+particularly in the mid-range of the distribution facilitates the establishment 
+of a convention, in this range individuals respond appropriately to the trait of their 
+peer. The house sparrow (*Passer domesticus*) has been a text-book example 
+of badges of status. According to the prevailing narrative the bib size 
+in males is a signal of the dominance rank and quality.
+However, a recent meta-analysis called into question 
+this narrative by showing that the effect size of the association between bib 
+size and dominance rank is small and uncertain [@sanchez-tojar_Metaanalysis_2018a]. 
+The simple straight correlation between bib size and dominance rank
+( and quality) is expected under the handicap principle, as it is shown here and 
+elsewhere [@botero_Evolution_2010;@grafen_Biological_1990;@johnstone_Badges_1993].
+But, the diversification process presented here shows that conventions can 
+emerge in a less straightforward way when mediated by learning processes. This 
+nuances could potentially make sense of seemingly contradictory evidence on the 
+correlation between plumage traits and quality. For example, conventions in 
+our simulations arose only in an intermediate range of the signalling trait. 
+While correlations are often performed across the whole range of variation.
+
+The focus of theoretical and empirical work on communication systems, 
+and particularly badges of status, is often explaining the presence and 
+absence of certain morphological traits within populations. However, the 
+morphological traits hypothesized to play a role in communication systems 
+regularly present interesting patters at the phylogeographic
+level. One example of this is the leapfrog pattern [@remsen_High_1984],
+whereas a certain morphological trait alternates its presence and 
+absence in a set of geographically adjunct populations. Patterns of 
+molecular variation along geographic clines suggest that natural selection 
+is involved in the emergence of these patterns 
+[@cadena_Testing_2011;@cadena_Evolutionary_2007]. However, it is unclear what 
+type of ecological process is behind these selective regimes. 
+Given the wide variation in outcomes found in our model, that depend on both 
+stochasticity and cognitive parameters, we think variation in cognition and 
+learning in particular could provide some explanatory power in this respect. 
+One first step, for example is to evaluate the way individuals in different 
+populations respond to novel traits [@avendano_Territorial_2021].
+
+We have presented here an evolutionary model where the evolution of a 
+communication system is mediated through the learned responses of
+receivers. This is a novel way to understand the evolution of communication
+that integrates classical cognitive process in learning with 
+evolutionary explanations of communication. This approach contributes to the 
+further integration of proximate and ultimate explanations in behavioral and 
+evolutionary biology.
+
 # Supplementary material

### 2023-03-17 manuscript/manuscript_1.0.Rmd
Done with reviewing main text. Need to check figure captions

@@
 
-A somewhat ignored component of communication systems, in the context 
-of aggressive interactions, is the cognitive aspects of the receiver module. 
+A somewhat ignored component of communication systems in the context 
+of aggressive interactions are cognitive aspects of the receiver module. 
 Theoretical models often assume that communication systems relying on badges of 
-status to have reaction norms as their 
-mechanistic underpinning [@botero_Evolution_2010], with 
+status have reaction norms as their mechanistic underpinning 
+[@botero_Evolution_2010], with 
 individuals using their own quality and the opponent's badge to determine 
-whether to aggressively engage in contest. Reaction norms 
-allow individuals to respond to the available information without a big
-cognitive burden. That is at least in contrast to other information processing 
-mechanisms like individual recognition. With individual recognition 
-individuals associate cue of their peers to their quality. These association 
-must be learned throughout and individual's life, thus it has the usual 
-cognitive requirements of an associative learning process. An alternative view 
-of badges of status is that individuals learn 
+whether to aggressively engage in contests. Reaction norms 
+allow individuals to respond to the available information without a large
+cognitive burden. The computations involved in the reaction norm response do not
+required a lot of memory or advanced algorithmic processes. 
+This is in contrast to other information processing 
+mechanisms like individual recognition. With individual recognition, 
+individuals associate cues of their peers to their quality. Because these 
+associations must be learnt throughout the life of individual, 
+they have the usual cognitive requirements of associative learning processes. 
+An alternative view of badges of status is that individuals learn 
 to react to them based on their experiences [@guilford_Receiver_1991], 
@@
 (e.g. in size or intensity), fighting ability may increase monotonically 
-with attributes of the badge and would be reinforced by every interaction, 
-so in principle learning would be faster than in systems involving individual 
-recognition in which the association between signals and their meaning would 
-vary depending on the interacting partners. In any case, learning could be 
+with attributes of the badge and would be reinforced by every interaction.
+Therefore, in principle, learning in such scenarios would be faster 
+than in systems involving individual 
+recognition in which the association between signals and their meaning 
+varies depending on the interacting partners. In any case, learning may be 
 a central cognitive mechanism in both types of communication systems, 
@@
 
-Associative learning is a key cognitive mechanism that allows 
-individuals to associate rewards with environmental stimuli 
-and thus behave adaptively [@staddon_Adaptive_2016]. Associative 
-learning exists in all major vertebrate taxa, and in many invertebrates
-as well [@heyes_Simple_2012;@macphail_Brain_1982; @staddon_Adaptive_2016; 
+Associative learning is a taxonomically widespread cognitive 
+mechanism that allows individuals to associate rewards with environmental stimuli 
+and thus behave adaptively 
+[@heyes_Simple_2012;@macphail_Brain_1982; @staddon_Adaptive_2016; 
 @behrens_Associative_2008].
-Theory has shown that natural selection favours these associations 
-in complex environments where conditions are hard to predict 
-[@dridi_Environmental_2016]. Besides its wide taxonomic and
-ecological relevance, associative learning is a flexible cognitive 
+Theory has shown that natural selection favours such associations 
+in complex environments where conditions are difficult to predict 
+[@dridi_Environmental_2016]. Associative learning is a flexible cognitive 
 mechanism whose underpinnings show interspecific variation 
@@
 on the psychological and neurological mechanisms of associative 
-learning [@sutton_Reinforcement_2018]. These set of algorithms allow the 
+learning [@sutton_Reinforcement_2018]. This set of algorithms allows the 
 implementation of biologically realistic problems, 
@@
 
-In here we present an evolutionary model where individuals use 
+Here we present an evolutionary model where individuals use 
 associative learning to develop a tendency to behave aggressively 
-(or  peacefully), in the context of competition over resources, 
-depending on the quantitative morphological trait they 
-perceive in their opponent (badge). Over evolutionary time 
-individuals evolve the size of their badge, and whether it depends 
-on their quality. Under this simple
-set up, individuals can use the badge as a signal of quality. 
+(or  peacefully) in the context of competition over resource, 
+depending on a quantitative morphological trait 
+(i.e.: a badge indicating quality) they 
+perceive in their opponents. Over evolutionary time the size of 
+the badge evolves, and its dependency on the individual's quality. 
+Under this simple set up, individuals can use the badge as a signal of quality. 
 We use the model to assess under what conditions 
-of interaction structure we expect different communication signals to evolve.
+we expect communication signals to evolve as handicaps or conventions.
 
@@
 
-We model the evolution of signals of quality in the context of agonistic 
-interactions. For simplicity we consider a population of haploid 
+We model the evolution of signals indicating individual quality 
+in the context of agonistic interactions. 
+For simplicity we consider a population of haploid 
 individuals with non-overlapping generations. Individuals are born 
@@
 $N( 0.5,\sigma)$. An individual with quality $0$ has the lowest RHP, 
-while an individual with quality $1$ has the highest. As part of 
-development processes individuals produce a morphological signal (badge), 
-the size of such depends on their level 
-of quality according to a reaction norm given by 
+while an individual with quality $1$ has the highest. As they 
+develop, individuals produce a phenotypic signal (badge), 
+the size of which depends on their quality according to a reaction norm given by 
 $B_i=1/(1+e^{\alpha_i-\beta_i Q_i})$; where $\alpha_i$ and 
@@
 and $\beta_i$ means individuals can have either uninformative (flat) or 
-informative (logistic) norms. The size of the signal is constrained 
-between zero and one. We assume different values of these traits 
-are given by different alleles, and are inherited from mother to offspring. 
-Unless, with a small probability ($\mu$), mutation changes the allelic value 
+informative (logistic) norms. Informative reaction norms represent a 
+developmental program where the phenotype of the individual is determined by
+the environmental conditions under which it grows, thus causing a 
+correlation between quality and signal. The size of the signal is constrained 
+to take values between zero and one. We assume different values of these traits 
+are given by different alleles and are inherited from mother to offspring 
+unless, with a small probability ($\mu$), mutation changes the allelic value 
 of the offspring by an amount taken from a normal distribution $TN(0,\sigma_\mu)$.
-After birth individuals go through a round of 
-viability selection. Individual specific survival probability is 
-given by $s_i=1/(1+e^{-k_1-k_2(Q_i-B_i)})$. Where $k_1$ and $k_2$ are parameter 
+
+After birth, individuals go through a round of 
+viability selection. The survival probability of an individual is 
+given by $s_i=1/(1+e^{-k_1-k_2(Q_i-B_i)})$, where $k_1$ and $k_2$ are parameter 
 values determining the shape of the survival function. 
@@
 <!-- interactions are more local.  -->
-In an interaction each individual must 
+In each interaction individuals must 
 decide whether to escalate a fight or not. 
@@
 as pay-off the resource of value $V$ while its partner gets nothing; 
-if both individuals restrain  from fighting they split up the resource 
+if both individuals restrain  from fighting, then they split up the resource 
 in half; if both decide to fight they split up the cost of fighting and 
@@
 probability of wining a fight for the focal individual depends on the 
-difference in quality between her and her interacting partner; specifically 
+difference in quality between it and its interacting partner; specifically 
 it is given by $p_{ij}=1/(1+e^{-k_3(Q_i-Q_j)})$, where $k_3$ is a 
 parameter defining how strong the quality difference determines the 
-winning probability; and $i$ and $j$ denote the the position 
+winning probability, and $i$ and $j$ denote the the position 
 in the population vector of the focal and its interacting 
-partner respectively.
+partner, respectively.
 
 The decision of whether to escalate a fight against an interacting 
-partner can be dependent on the size of the partner's badge, 
-the actual dependency is determined by the focal´s experiences 
+partner can depend on the size of the partner's badge, 
+the dependency is determined by the focal's experiences 
 through a learning process. We implement learning using 
@@
 with partners of different badge sizes (the critic in RL terminology). 
-After every interaction they update the estimate of reward proportionally 
+After each interaction they update the estimate of reward proportionally 
 to the difference between their current estimate and the observed 
 reward (prediction error $\delta$) and to the speed of learning 
-($\mathrm{A}$). Furthermore, they express different probabilities of 
-retreating (attacking) depending on the badge size of their opponent 
+($\mathrm{A}$). Furthermore, individuals express different probabilities of 
+retreating/attacking depending on the badge size of their opponent 
 (the actor in RL). They update the probability of 
-retreating (attacking)  up or down depending on whether retreating 
-(attacking) leads to an increase in the reward estimation. 
+retreating/attacking  up or down depending on whether retreating/attacking
+leads to an increase in the reward estimation. 
 So, if a focal individual decides to escalate a fight against 
-an individual with small badge, and that leads to an increase in 
-the reward estimation, the focal individual will increase the probability 
+an individual with small badge and this leads to an increase in 
+the reward estimation, then the focal individual will increase the probability 
 of escalating fights with individuals of small badges in the future. 
@@
 many badge sizes. Thus, the reward estimation, as well as the probability 
-of retreating (escalating), must be generalized across different values. 
+of retreating/attacking, must be generalized across different values. 
 To implement generalization we use the linear function approximation 
@@
 Specifically, we pick $c$ feature centres, which are evenly spaced values 
-along the badge size interval ([0,1]). This feature centres are the same for all individuals and stored in vector $\symbf{b}$ . Each one of these features centres 
+along the badge size interval ([0,1]) where the updates are focused. 
+For the sake of simplicity we keep the location of these feature centres 
+constant and  the same for all individuals, the are stored in the 
+vector $\symbf{b}$ . Each of these feature centres 
 is associated with a weight for reward estimation and tendency to play 
@@
 triggered by the feature (Fig. \ref{fig:model-struc} B, dots 
-represent the feature weights) The response of each feature centre 
+represent the feature weights). The response of each feature centre 
 diminishes as a Gaussian function with the distance between the 
 feature centre and the badge size of 
-the partner (Fig. \ref{fig:model-struc} B, grey line). Formally the 
+the partner (Fig. \ref{fig:model-struc} B, grey line). Formally, the 
 reward estimation $\hat{R}$ when the focal individual ($i$) faces 
@@
 
-After all their interactions, individuals in the population 
-reproduce proportionally to their total pay-off $w_i$, which is a sum of the 
+Individuals interact as the focal individual $n$ times, and the interaction 
+partner is chosen at random. Thus, the expected number of interactions for 
+each individual is $n+\frac{n(N+1)}{N}$. 
+After the interaction round , individuals in the population 
+reproduce with a probability proportional to their total pay-off $w_i$, which is a sum of the 
 baseline pay-off ($w_0$) and all the pay-offs obtained throughout their 
@@
 ```{r model-struc,echo=FALSE,warning=FALSE,fig.width=8,fig.align='center',fig.height=7,fig.show='hold',fig.cap= "Model of communication in the context of aggressive interactions.
-In A the sender code: a reaction norm determines how the badge size is determined by the quality of the individual.
+In A, the reaction norm determines the badge size.
 Red shows an informative reaction norm, while the blue shows a uninformative reaction norm.
-B the receiver code: individuals have a behavioral reaction norm that determines their probability of retreating (black line).
+B individuals perceiving the signal have a behavioral reaction norm that determines their probability of retreating (black line).
 The reaction norm arises from generalizing the information from the feature weights (black dots).
-Generalization is represented by the grey line and its axis, which shows the response triggered by the fourth feature diminishes as the value evaluated is further from the feature center.
+Feature weaights are values associated with a particular badge size and are updated as individuals interact with each other.
+The generalization of the information associated with each feature weight is represented by the grey line and its axis, which shows the response triggered by the fourth feature diminishes as the value evaluated is further from the feature center.
 The learning process moves the feature weights (black dots) up or down depending on whether that leads to an increase in the estimated reward.
@@
 Receivers in C face signallers with uninformative reaction norms (blue line in A), while in D they face signallers with informative reaction norms (red line in A).
-Colour scale in C (applies also for D) indicates the quality of the individual"}
+Colour scale in C and D indicates the quality of the individual"}
 
@@
           lwd=3,ylim=c(0,1.1))
-text(x = 0.08,y=1.05,"B",cex = 1.5)
+text(x = 0.08,y=1.05,"C",cex = 1.5)
 par(new=FALSE)
@@
 We first present the outcome of simulations where we prevent the 
-evolution of the sender code (by setting the mutation rate to $0$), 
+evolution of the badge size (by setting the mutation rate to $0$), 
 and assume that all individuals in the population display either 
@@
 receivers that faced uninformative signals, and D informative ones. 
-When learners face uninformative signals, they change their 
-probability of retreating depending on their own quality. Individuals 
-with higher quality (red tones) after the learning process have a 
+The simulations reveal that when learners face uninformative signals, 
+they modify their probability of retreating depending on their own quality. 
+Individuals with higher quality (red tones) after the learning process have a 
 high probability of attacking; while individuals of lower quality 
-(blue tones) mostly retreat from confrontations. Thus, learning splits
-the population of receivers into the two classic pure strategies of 
-hawks and doves. Given that we have a assumed a monomorphic population 
+(blue tones) mostly retreat from confrontations; in both cases irrespective of 
+their opponent. Thus, learning splits the population of receivers into 
+the two classic pure strategies of hawks and doves. 
+Given that we have a assumed a monomorphic population 
 with unresponsive reaction norms on the signalling side, the changes 
-triggered by learning only affect a small range of badge sizes. 
+triggered by learning only affect a small range of badge sizes. This can be seen 
+in panel C, all learning has occurred around a badge size of $0.5$, because all
+badges in the population are this size. 
 In contrast, when receivers face informative reaction norms 
@@
 whether to retreat or attack. The relationship is given by a 
-threshold-like reaction norm, where the change from retreating to 
-attacking depends on the quality of the receiver. As expected, 
+threshold-like reaction norm, where the decision to retreat or  
+attack depends on the quality of the receiver. As expected, 
 the higher the quality of the receiver the larger the badge size 
@@
 
-When we allow the evolution of the sender code ($\alpha$ and $\beta$ 
-change subject to natural selection and genetic drift), and the badge 
-size work as a handicap (i.e. the cost of the badge is inversely 
+When we allow the badge size to evolve ($\alpha$ and $\beta$ 
+change subject to natural selection and genetic drift), and the signal
+works as a handicap (i.e. the cost of the badge is inversely 
 proportional to the quality of the individual), the sender code can evolve 
@@
 evolutionary process. Instead, the evolutionary dynamics of the reaction 
-norm parameters (Fig. \ref{fig:handicap}, A) show a set of 
-steps. First, the intercept of the reaction norm ($\alpha$) evolves to 
+norm parameters (Fig. \ref{fig:handicap}, A) appear to involve a set of 
+steps. First, the intercept of the reaction norm ($\alpha$) 
+describing the relationship between individual quality and badge size evolves to 
 higher values, reducing the average badge size in the populations. 
-That makes sense because reaction norms segregating in the population are 
-flat at this point, and consequently the receiver code does not respond to 
-the badge size, bigger badges are costlier and do not trigger lower attack 
-probabilities. During those first generations, the slope of the sender 
-reaction norm ($\beta$)  remains around the initial value of zero. 
+This is because reaction norms segregating in the population are 
+flat at the beginning of the simulation, consequently, 
+the receiver code does not respond to the badge size, 
+and larger badges are costlier and do not trigger lower attack 
+probabilities. During the first generations, the slope of the sender 
+reaction norm ($\beta$)  remains close to the initial value of zero. 
 At around generation 4000, the slope evolves toward positive values. 
 Positive values in the slope imply that larger badges correlate 
-with higher quality (Fig. \ref{fig:handicap} B). Receivers learn to react to such 
+with higher quality (Fig. \ref{fig:handicap} B). Receivers then learn to react to such 
 correlation, increasing the probability of retreat when facing 
-individual of larger badges (Fig. \ref{fig:handicap} C). Hence, natural selection 
+individuals with larger badges (Fig. \ref{fig:handicap} C). Hence, natural selection 
 favours larger values of $\beta$, eventually leading to an evolutionary 
-equilibrium where badge size is an honest signal of quality mediated
+equilibrium in which badge size is an honest signal of quality mediated
 by the learned responses of receivers (Fig. \ref{fig:handicap} B.4 and C.4).
+
 The evolutionary trajectory portrayed in figure 
-\ref{fig:handicap} A is not the only possible trajectory. If the slope of the 
-sender reaction norms ($\beta$) evolves toward negative values before badges 
-become handicaps, receivers never learn to react to the size of badges. 
+\ref{fig:handicap} A is not the only possible outcome. If the slope of the 
+sender reaction norms ($\beta$) evolves, subject to genetic drift,
+toward negative values before badges become handicaps, 
+receivers never learn to react to the size of badges. 
 Thus, badges are only costly and do not provide information about the 
-quality of an individual (Fig. \ref{fig:noBadge}) . Eventually, the badge 
+quality of individuals (Fig. \ref{fig:noBadge}) . Eventually, the badge 
 disappears from the population. In contrast, when the badge does not 
@@
 never leads to the establishment of an honest signal. 
-Instead, subject to genetic drift the 
+Instead, subject to genetic drift, the 
 population either evolves towards the disappearance of the badge, or to the 
 maximum badge size. In either of this cases reaction norms are flat, so the 
-badge does not provide any quality information.
+badge does not provide any information about quality.
 
@@
 throughout their life time can strongly change the outcome of 
-the evolutionary dynamics. In the simulations presented in figure 
+evolutionary dynamics. In the simulations presented in figure 
 \ref{fig:handicap} and \ref{fig:noBadge}, 
-individuals learned with high learning rate ($\alpha = 0.4$) and interacted 
-repeatedly along their lifetime (on average 2000 interactions). When we reduce 
+individuals learned with high speed of learning ($\mathrm{A} = 0.4$) and interacted 
+repeatedly along their lifetime (on average 2000 interactions). When we reduced 
 the number of interactions that individuals have over their lifetime to 300 
-interactions on average, we see an drastic increase in the morphological and genetic
+on average, we saw a drastic increase in the phenotypic and genetic
 variation present in the population. Fig. \ref{fig:branching} shows 
@@
 In panels A1 and 2, darker areas show trait values with high frequency in the population. 
-Populations start monomorphic with a value of zero on both traits, 
-and mutations quickly build up a normal distribution around the starting 
-value. Within the first 2000 generations the unimodal distribution in 
+Populations start monomorphic with a value of zero on both the intercept and
+the slope, and mutations quickly build up a normal distribution around the starting 
+value. Within the first 2000 generations, the unimodal distribution in 
 the intercept ($\alpha$) splits into a bimodal one. 
@@
 ($\alpha$) in the population shows three peaks. In the case of $\beta$, the 
-peaks in the distribution are not so clear-cut, but it's clear that there is an 
+peaks in the distribution are not so clear-cut, but it is clear that there is an 
 increase in the variance of the distribution. These changes in the 
-distribution of the sender reaction norm parameters 
+distribution of the parameters of the sender reaction norm 
 imply that individuals can generally be classified into three 
@@
 norm with extreme values for the badge size, meaning that their badge size 
-is not informative of their quality, while the third type shows 
-intermediate badge sizes which are determined 
-by the quality of the individual (Fig. \ref{fig:branching} B.3-4). 
+is not informative of their quality, whereas the third type shows 
+intermediate badge sizes determined by the quality of the individual 
+(Fig. \ref{fig:branching} B.3-4). 
 In this particular simulation, the values of the slope in the informative group
@@
 ```
-The effect of a limited number of interactions on the amount of variation in 
-the evolving parameters plays a role when the signal follows the handicap
+A limited number of interactions on the amount of variation has an effect on 
+the evolving parameters when the signal follows the handicap
 principle as well. In fig \ref{fig:honestDiv}, we show simulations where individuals 
@@
 there is a positive correlation between the size of the badge and the quality 
-of the individual. This relationship however is muddled by the fact that there 
+of the individual. This relationship, however, is muddled by the fact that there 
 are two clusters of values for the intercept ($\alpha$) and slope ($\beta$) 
 of the reaction norm in the population. Thus, there are two types 
-of reaction norms. One with a steeper slope, meaning that it expresses 
+of reaction norms. One of the types has a steeper slope, meaning that it expresses 
 a higher badge size for a given quality. This effect of an increased 
 variance in the trait distribution is not only triggered by lower 
-number of interactions. Larger variances are also found When we assume 
+number of interactions. Larger variances are also found when we assume 
 a lower speed of learning (data not shown). This suggest
-that, limits to the amount of information that individuals acquire through
-learning, allows the coexistence of different communication strategies 
-[@botero_Evolution_2010].
+that limits to the amount of information that individuals acquire through
+learning allows the coexistence of different communication strategies within 
+a population [@botero_Evolution_2010].
 
@@
 imposes negative-frequency dependent selection, allowing for the build up of
-genetic variation in reaction norms. In the simulations, presented so far we 
+genetic variation in reaction norms. In the simulations presented so far, we 
 assumed that individuals start with a flat behavioural reaction norm such that 
 they escalate aggressively a fight with a $0.5$ probability regardless of 
-the badge size of the interacting partner. To asses if this initial 
+the badge size of the interacting partner. To asses whether such initial 
 conditions of the communication system had any role 
@@
 the initial conditions of the actor module in the learning model. Specifically,
-we let naive individuals have a flat reaction norm with a 1) low (peaceful),
-2) high (aggressive) probability of escalating a fight and 3) a probability 
+we let naive individuals have a flat reaction norm with either a 1) low (peaceful),
+2) high (aggressive) probability of escalating a fight or 3) a probability 
 corresponding to the ESS of the original hawk-dove game (clever). 
@@
 replicates of the simulations. The left-hand side panel, 
-that corresponds to peaceful naive individuals, is the only one where the  
-distribution of values is split in different clusters. That is, the majority of 
+corresponding to peaceful naive individuals, is the only one where the  
+distribution of values is split in different clusters. That is, most of 
 the variation occurs within clusters. In contrast, when naive individuals behave
 either aggressively or cleverly, the badges evolve toward minimum and maximum 
-values, and so the variation is driven mainly by among replicate differences. 
+values, and so the variation is driven mainly by differences among replicates. 
 We can make sense of these results by realizing that individuals change their 
@@
 the population. Thus, one way to repeatedly interact with the naive behaviour
-is to have a rare badge size. When the naive behaviour is peaceful, individual
+is to have a rare badge size. When the naive behaviour is peaceful, individuals
 with a rare badge size have a fitness advantage. This triggers 
@@
 badge sizes. In situations where individuals learn fast and interact repeatedly 
-the strength of frequency dependent selection diminishes. The diversification 
-process presented here is fully dependent on the learning process and how this
-process is implemented.
+the strength of frequency dependent selection diminishes.
 
@@
 | $N$   | Population size |
+| $n$   | Number of interactions as the focal individual |
 | $\sigma$ | Standard deviation of the truncated normal distribution from which quality is drawn |

### 2023-03-31 manuscript/manuscript_1.0.Rmd
addressing Daniel´s comments on the manuscript

@@
 title: |
-  | Learning can mediate both badges of status and individual recognition
+  | The role of learning in status signalling: a modeling approach
 author:
@@
 output:
-  pdf_document:
-    latex_engine: xelatex
   word_document: default
+  # pdf_document:
+  #   latex_engine: xelatex
 fig_caption: yes
@@
 whom it interacts. Thus, acting based on information about interacting 
-partners is typically adaptive [@quinones_Negotiation_2016].
-Acquiring such information, however, is far from trivial. 
+partners is typically adaptive [@quinones_Negotiation_2016], yet
+acquiring such information is far from trivial. 
 In some cases, interacting partners (signallers) might be 'willing' to 
@@
 where one can help the other. Given some costs and benefits, 
-a potential donor would be interested in helping 
+a potential donor is likely to be interested in helping 
 related individuals. Therefore, for a relative of the donor, 
@@
 [@bradbury_Principles_2011; @moller_Female_1988; @tibbetts_Socially_2004].
+
 In dominance and aggressive interactions, a crucial piece of
@@
 [@johnstone_Badges_1993; @rohwer_Social_1975a]. 
-Such signals must be arbitrary, meaning they are not 
-ontogenetically correlated with quality. 
+Such signals must be arbitrary, meaning the signals could potentially be 
+produced by high and low quality individuals alike. 
 Generally signals can be evolutionarily stable whenever low quality individuals 
-pay a higher fitness cost for carrying the signal. 
-So, it's no longer in the interest of low quality individuals 
+pay a higher fitness cost for carrying the signal such that it is not 
+in the interest of low quality individuals 
 to fake quality by producing a large badge. Signal costs inversely proportional
-to quality con come about by the production costs of the signal, 
+to quality can come about as a result of the production costs of the signal, 
 in which case the signal works like a handicap [@botero_Evolution_2010; @johnstone_Badges_1993;@grafen_Biological_1990; @zahavi_Mate_1975]. 
@@
 
-Continue here!!!!
-
 A somewhat ignored component of communication systems in the context 
@@
 [@botero_Evolution_2010], with 
-individuals using their own quality and the opponent's badge to determine 
+individuals using their own quality and the badges of opponents to determine 
 whether to aggressively engage in contests. Reaction norms 
 allow individuals to respond to the available information without a large
-cognitive burden. The computations involved in the reaction norm response do not
-required a lot of memory or advanced algorithmic processes. 
-This is in contrast to other information processing 
+cognitive burden because the computations involved in the reaction 
+norm response do not require much memory nor advanced algorithmic processes. 
+This contrasts with other information-processing 
 mechanisms like individual recognition. With individual recognition, 
-individuals associate cues of their peers to their quality. Because these 
-associations must be learnt throughout the life of individual, 
+where individuals associate cues of their peers to their quality. Because these 
+associations must be learnt throughout the life of individuals, 
 they have the usual cognitive requirements of associative learning processes. 
@@
 explored in the empirical nor theoretical literature.
+
+<!-- Continuar aquí en la revisión -->
 
@@
 In C and D, effect of learning on the receiver strategy.
-Receivers in C face signallers with uninformative reaction norms (blue line in A), while in D they face signallers with informative reaction norms (red line in A).
-Colour scale in C and D indicates the quality of the individual"}
+Receivers in C face signallers with uninformative reaction norms (blue line in A).
+The uninformative reaction norm yields badge size of 0.5, thus the behavioural reaction norm in C is only updated around that value.D Receivers face signallers with informative reaction norms (red line in A).
+Accordingly, receivers develop a threashold-like reaction norm where the decision to retreat increases with the badge size of the interacting partner.
+Colour scale in C and D indicates the quality of the individual.
+"}
 
@@
 Color scale indicates quality just as in Fig.
-\\ref{fig:model-struc}." ,warning=FALSE}
+\\ref{fig:model-struc}.
+On the first half of the simulation (B.1 and B.2) the badge size evolves to its minimum value.
+In the second half, the value of the slope takes positive values and reaction norms evolve to produce an honest signal of quality.
+Correspondingly, individuals learn to react to badge size by increasing the probability of retreat with increasing badge sizes"  ,warning=FALSE}
 
@@
 Panels below correspond to snapshots of the sender (B) and receiver codes (C).
-Generation time of the snapshots are indicated by the grey lines in the top panels."}
+Generation time of the snapshots are indicated by the grey lines in the top panels.
+The unimodal distribution with which $\\alpha$ (A.1) starts the simulation, quickly turns into a bimondal distribution, and at about 10000 generations it splits further into three modes.
+These three peaks correspond to three type of reaction norms in panels B.1-3."}
 
@@
 
-```{r honestDiv,echo=FALSE,out.width='100%',fig.height=6,fig.align='center',fig.show='hold',fig.cap= "The evolution of diverse costly signals.
+```{r honestDiv,echo=FALSE,out.width='100%',fig.height=6,fig.align='center',fig.show='hold',fig.cap= "The evolution of costly signals.
 Portrait of the evolutionary dynamics of the sender code with snapsots of both sender and receiver codes just as in fig.
@@
 The middle panels show changes in the distribution of values for $\\alpha$ and $\\beta$ along evolutionary time.
-Panels above and below correspond to snapshots of the sender and receiver codes, respectively, generation time of the snapshots are indicated by the grey lines in the middle panels."}
+Panels above and below correspond to snapshots of the sender and receiver codes, respectively, generation time of the snapshots are indicated by the grey lines in the middle panels.
+Similarly to the evolutionary dynamics in Fig.
+\\ref{fig:handicap}, first the badge size evolves towards its minimum value.
+Then, the slope evolves positive values determining an increasing reaction norm.
+In contrast to \\ref{fig:handicap}, the normal distribution splits up into two modes.
+This translates into two types of reaction norms assorting in the population where both code for a positive relation between quality and badge size."}
 
@@
 Distribution of values of the intercept $\\alpha$ and slope $\\beta$ for individuals at the end of the evolutionary simulations.
-The panels show the three different initial conditions for the behaviour of individuals: peaceful (blue), clever (green) and aggressive (red), see maintext for details.
+The panels show the three different initial conditions for the behaviour of individuals: peaceful, \"clever\" and aggressive, see maintext for details.
 Colours indicate the replicate simulation.
-In the inset, the resulting reaction norms corresponding to the intecept and slope values for one of the replicates, which replicate is shown is indicated by the line colour."}
+In the inset, the resulting reaction norms corresponding to the intecept and slope values for one of the replicates, which replicate is shown is indicated by the line colour.
+The simulations where individual disply a \"clever\" and agressive behaviour, before the learning process, show all the points of a single replicate cluster together, while in simulation where individuals start being peaceful al replicates show distinct groups of individuals spread throughout the x and y axis.
+Only in this condition there is a frequency-dependent process that favour diversity in the reaction norm."}
 

### 2023-04-27 manuscript/manuscript_1.0.Rmd
New figure with simulations where learning pars evolve

@@
 title: |
-  | The role of learning in status signalling: a modeling approach
+  The role of learning in status signalling: a modeling approach
 author:
@@
 output:
-  word_document: default
-  # pdf_document:
-  #   latex_engine: xelatex
+  pdf_document:
+    latex_engine: xelatex # default  # word_document: default
 fig_caption: yes
 fontsize: 12pt
-header-includes: "```{=latex}\n  \\usepackage{float}\n  \\usepackage{unicode-math}\n
-  \ \\usepackage{setspace}\\doublespacing\n  \\usepackage[backref=page]{hyperref}\n
-  \ \\linespread{2}\n  \\usepackage{lineno}\n  \\linenumbers\n  \\floatplacement{figure}{H}\n
-  \ \\newcommand{\\beginsupplement}{ \\setcounter{table}{0}         \\renewcommand{\\thetable}{S\\arabic{table}}\\setcounter{figure}{0}
-  \\renewcommand{\\thefigure}{S\\arabic{figure}}}\n```\n"
+header-includes: |
+  ```{=latex}
+    \usepackage{float}
+    \usepackage{unicode-math}
+    \usepackage{setspace}\doublespacing
+    \usepackage[backref=page]{hyperref}
+    \linespread{2}
+    \usepackage{lineno}
+    \linenumbers
+    \floatplacement{figure}{H}
+    \newcommand{\beginsupplement}{ \setcounter{table}{0}         \renewcommand{\thetable}{S\arabic{table}}\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
+  ```
 keep_tex: yes
@@
 style: C:\Users\andre\Zotero\styles\elife.csl
-abstract: "Adaptive behavioural responses often depend on qualities of the interacting
+abstract: |
+  Adaptive behavioural responses often depend on qualities of the interacting
   partner of an individual. For example, when competing for resources, an individual
@@
   of both honest and dishonest signalers. We argue that learning can be a general
-  cognitive mechanism playing a role in a wide range of communication systems.  \n"
+  cognitive mechanism playing a role in a wide range of communication systems."
 ---
@@
 if both individuals restrain  from fighting, then they split up the resource 
-in half; if both decide to fight they split up the cost of fighting and 
-the winner takes over the resource. We further assume that the 
-probability of wining a fight for the focal individual depends on the 
+in half.
+If both individuals decide to fight, the winner takes over the resource and the cost of fighting is split among the two players.
+We further assume that the probability of wining a fight for the focal individual depends on the
 difference in quality between it and its interacting partner; specifically 
@@
 
+### Learning under a monomorphic population
+
 We first present the outcome of simulations where we prevent the 
@@
 
+### Badges as Handicaps
+
 When we allow the badge size to evolve ($\alpha$ and $\beta$ 
@@
 badge does not provide any information about quality.
+
+### The evolution of polymorphism mediated by learning
 
@@
 rm(list=grep("temp",ls(),value = T))
-
 ```
+
+### The peaceful, the clever and the aggressive
 
@@
 
-We have presented here an evolutionary model where the response of a signal in a 
+Here we presented a model that integrates the basic mechanisms of 
+learning with the fundamental rules of evolutionary processes. 
+In the model the response of a signal in a 
 communication system is mediated by learning processes. Individuals 
 throughout their life learn the best way to respond to a quantitative 
-trait of their interactive partners. The learning process can mediate the evolution 
+trait of their interactive partners. 
+The learning process can mediate the evolution 
 of an honest signal under the handicap 
@@
 
+```{r learnEvol,echo=FALSE,out.width='100%',fig.height=6,fig.align='center',fig.show='hold',fig.cap= "The evolution of costly signals.
+Portrait of the evolutionary dynamics of the sender code with snapsots of both sender and receiver codes just as in fig.
+\\ref{fig:handicap}.
+The middle panels show changes in the distribution of values for $\\alpha$ and $\\beta$ along evolutionary time.
+Panels above and below correspond to snapshots of the sender and receiver codes, respectively, generation time of the snapshots are indicated by the grey lines in the middle panels.
+Similarly to the evolutionary dynamics in Fig.
+\\ref{fig:handicap}, first the badge size evolves towards its minimum value.
+Then, the slope evolves positive values determining an increasing reaction norm.
+In contrast to \\ref{fig:handicap}, the normal distribution splits up into two modes.
+This translates into two types of reaction norms assorting in the population where both code for a positive relation between quality and badge size."}
+
+# knitr::include_graphics(here("Simulations","nIntGroupNormQual_","evolDyn7_nIntGroup2000.png"))
+
+scenario<-"betCostalphaL"#"alphaAct"#"nIntGroupNormQual"#
+runChoi<-13
+
+# Load files -------------------------------------------------------------------
+# Project folder
+listTest<-list.files(here("Simulations",paste0(scenario,"_")),full.names = TRUE)
+evolList<-grep("evolLearn",listTest,value=TRUE)
+indList<-grep("indLearn",listTest,value=TRUE)
+paramName<-list.files(here("Simulations",paste0(scenario,"_")))
+# paramName<-list.files(extSimsDir,full.names = TRUE)
+paramName<-grep(".json",paramName,value=TRUE) %>% grep(pattern = "*0_*",value = TRUE)
+
+param<-fromJSON(here("Simulations",paste0(scenario,"_"),paramName[1]))
+  # fromJSON(paramName[1])
+
+val<-1
+fileId<-val
+
+evolList_runs<-grep(paste0(param$namParam,param$rangParam),
+                    evolList,value =TRUE)
+indList_runs<-grep(paste0(param$namParam,param$rangParam),
+                   indList,value =TRUE)
+
+evol<-do.call(rbind,lapply(evolList_runs,fread))
+pop<-do.call(rbind,lapply(indList_runs, fread))
+
+Valpar<-param$rangParam
+
+nampar<-param$namParam
+
+nCenters<-param$nCenters
+sigSquar<-param$sigSq
+
+# Aesthetic paremeters ---------------------------------------------------------
+
+## Calculate clustering for all generations  -------------------
+
+### Choose which interaction to visualize
+# lastInt<-tail(pop[,unique(nInteract)],4)[2]
+# # pop[,max(nInteract),by=.(seed,time)][,min(V1)]
+# # to get last interaction: tail(pop[,unique(nInteract)],1)
+# 
+# popOneInd<-pop[nInteract==lastInt]
+# vars<-c("alpha","beta","Badge")
+# # ,"Badge",grep("WeightAct",names(popFinal),value = TRUE))
+# popOneInd$idClust<-get_clusters(popOneInd,vars,k.max = 10,
+#                                 Bsamples =500,iterMax = 500)
+# clusSummary<-popOneInd[,.(meanAlph=mean(alpha),meanBet=mean(beta)),
+#                        by=.(idClust,time,seed)]
+# clusSummary[,orderClus:=0]
+# invisible(capture.output(
+#   for(i in 1:dim(clusSummary)[1]){
+#   set(clusSummary,i,6L,match(clusSummary[i,meanAlph],
+#         sort(clusSummary[time==clusSummary[i,time]&
+#                            seed==clusSummary[i,seed]]$meanAlph)))
+# }
+# ))
+# 
+# 
+# popOneInd<-merge(popOneInd,clusSummary,all = TRUE)
+# 
+# 
+# fwrite(popOneInd,file = here("Simulations",paste0(scenario,"_"),
+#                              paste0("popOneInd",nampar,Valpar,".txt")))
+# 
+popOneInd<-fread(here("Simulations",paste0(scenario,"_"),
+                             paste0("popOneInd",nampar,Valpar,".txt")))
+
+# Get stats from the evolutionary simulations ----------------------------------
+# names(evol)
+
+cols<-c("freqGenHawks","freqGenDove",  "freqGenEval",  #"freqGenLearn",
+        "freqFenHawks", "freqFenDoves", "freqHH", "freqHD", "freqDD", "meanCue",
+        "meanAlpha", "meanBeta",
+        #"meanFit", #"meanInitCrit", "sdInitCrit", 
+        "meanInitAct", "sdInitAct", "meanAlphaLearn","sdAlphaLearn",
+        "WeightAct_0","WeightCrit_0",
+        "WeightAct_1","WeightCrit_1","WeightAct_2",  "WeightCrit_2",
+        "WeightAct_3",  "WeightCrit_3", "WeightAct_4",  "WeightCrit_4",
+        "WeightAct_5","WeightCrit_5")
+
+my.summary<- function(x) list(mean = mean(x), lowIQR = fivenum(x)[2], 
+                              upIQR = fivenum(x)[4])
+
+evolStats<-evol[, as.list(unlist(lapply(.SD, my.summary))), 
+                .SDcols = cols,by=time]
+
+## Plot mean and IQRs among individuals for each replicate --------------------
+
+# get the trajectories for individual runs
+traitsTrajs<-dcast(evol,time~seed,
+                   value.var = c("meanAlpha","meanBeta",
+                                 #"meanInitCrit",#"meanFit",
+                                 "meanInitAct", "sdInitAct",
+                                 "meanAlphaLearn","sdAlphaLearn",
+                                 "sdAlpha","sdBeta","freqHH",
+                                 "freqHD","freqDD"))
+
+finReps<-evol[time==max(time),seed]
+
+genstoPrint<-round(seq(1,length(unique(evolStats$time)),length.out = 5))[2:5]
+
+# Evolutionary dynamics for alpha - intercept of the sender --------------------------------------------
+par(plt=posPlot(numploty = 3,idploty = 3,numplotx = 2,idplotx = 1,
+                lowboundx = 8, title = TRUE,lowboundy = 10,
+                upboundy = 92,upboundx = 92),
+    mfrow=c(1,1), xaxt="s",las=1,mgp=c(2,0.8,0))
+
+adjsYlims<-abs(range(pop[seed==runChoi,alpha])[1]-
+                 range(pop[seed==runChoi,alpha])[2])*c(-0.1,0.2)
+
+evolDist(indData = pop[seed==runChoi],variable = "alpha",nbins = 10,
+         pal = pal_dist,
+         nlevels=10,cexAxis = cexAxis,xlab="",ylab = "",xaxt="n",yaxt = "s",
+         ylim = range(pop[seed==runChoi,alpha])+adjsYlims,
+         xlim=c(0,22000))
+
+mtext(text = "Trait value",cex = cexAxis,line = 1.8,side = 2,las=0)
+legend("topright",legend =
+      expression(paste("intercept (",alpha[s],")")),col=colGenesLin[1],lwd=2,
+       bty = "n",cex=cexAxis)
+legend("topleft",legend = c("A.1"),cex=cexAxis,bty = "n",inset = -0.06)
+
+axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
+     labels = axTicks(1)[2:(length(axTicks(1)))]/1000,cex.axis=cexAxis,
+     hadj = 1.3,line = -0.5,tick = FALSE)
+
+# grey lines to show the generations shown in the upper and lower panels
+
+matlines(x=matrix(rep(evolStats[genstoPrint,time],each=2),nrow=2),
+         y=matrix(rep(range(pop[seed==runChoi,alpha]),4),nrow = 2),
+         lty=1,col = "grey",lwd=2)
+
+# Variation among individuals
+
+matlines(x=traitsTrajs[,time],
+         y=traitsTrajs[,.SD,
+                       .SDcol=paste0(c("meanAlpha_"),runChoi)],
+         col=colGenesLin,lty = 1,type="l",lwd=5)
+
+par(plt=posPlot(numploty = 3,idploty = 3,numplotx = 2, idplotx = 2,
+                lowboundx = 8, title = TRUE,lowboundy = 10,
+                upboundy = 92,upboundx = 92),
+    xaxt="s",las=1,new=TRUE)
+
+adjsYlims<-abs(range(pop[seed==runChoi,beta])[1]-
+                 range(pop[seed==runChoi,beta])[2])*c(-0.1,0.2)
+
+evolDist(indData = pop[seed==runChoi],variable = "beta",nbins = 10,pal = pal_dist,
+         nlevels=10,cexAxis = 1.5,xlab="",ylab = "",xaxt="n",yaxt = "n",
+         ylim = range(pop[seed==runChoi,beta])+adjsYlims,xlim=c(0,22000))
+legend("topright",legend = expression(paste("Slope (",beta[s],")")),cex=cexAxis,
+       col=colGenesLin[2],lwd=2,bty = "n")
+legend("topleft",legend = c("A.2"),cex=cexAxis,bty = "n",inset = -0.06)
+# grey lines to show the generations shown in the upper and lower panels
+matlines(x=matrix(rep(evolStats[genstoPrint,time],each=2),nrow=2),
+         y=matrix(rep(range(pop[seed==runChoi,beta]))),
+                      lty=1,col = "grey",lwd=2)
+
+axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
+     labels = axTicks(1)[2:(length(axTicks(1)))]/1000,cex.axis=cexAxis,
+     hadj = 1.3,line = -0.5,tick = FALSE)
+
+axis(4,cex.axis=cexAxis)
+
+matlines(x=traitsTrajs[,time],
+         y=traitsTrajs[,.SD,
+                       .SDcol=paste0(c("meanBeta_"),runChoi)],
+         col=colGenesLin[2],lty = 1,type="l",lwd=4)
+
+# Evolutionary dynamics for the speed of learning-------------------------------
+
+par(plt=posPlot(numploty = 3,idploty = 2,numplotx = 2,idplotx = 1,
+                lowboundx = 8, title = TRUE,lowboundy = 10,
+                upboundy = 92,upboundx = 92),
+    mfrow=c(1,1), xaxt="s",las=1,mgp=c(2,0.8,0),new=TRUE)
+
+adjsYlims<-abs(range(pop[seed==runChoi,alphaLear])[1]-
+                 range(pop[seed==runChoi,alphaLear])[2])*c(-0.1,0.2)
+
+evolDist(indData = pop[seed==runChoi],variable = "alphaLear",nbins = 10,
+         pal = pal_dist,
+         nlevels=10,cexAxis = cexAxis,xlab="",ylab = "",xaxt="n",yaxt = "s",
+         ylim = range(pop[seed==runChoi,alphaLear])+adjsYlims,
+         xlim=c(0,22000))
+
+mtext(text = "Trait value",cex = cexAxis,line = 1.8,side = 2,las=0)
+legend("topright",legend = "Speed of learning",col=colGenesLin[1],lwd=2,
+       bty = "n",cex=cexAxis)
+legend("topleft",legend = c("B.1"),cex=cexAxis,bty = "n",inset = -0.06)
+
+axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
+     labels = axTicks(1)[2:(length(axTicks(1)))]/1000,cex.axis=cexAxis,
+     hadj = 1.3,line = -0.5,tick = FALSE)
+
+# black line showing the value 0 as a benchmark
+
+lines(x = pop[,range(time)],y=c(0,0))
+
+# grey lines to show the generations shown in the upper and lower panels
+
+matlines(x=matrix(rep(evolStats[genstoPrint,time],each=2),nrow=2),
+         y=matrix(rep(range(pop[seed==runChoi,alphaLear]),4),nrow = 2),
+         lty=1,col = "grey",lwd=2)
+
+# Variation among individuals
+
+matlines(x=traitsTrajs[,time],
+         y=traitsTrajs[,.SD,
+                       .SDcol=paste0(c("meanAlphaLearn_"),runChoi)],         col=colGenesLin,lty = 1,type="l",lwd=5)
+
+## Plot evolutionary dynamics of the naive behaviour of inds -------------------
+
+par(plt=posPlot(numploty = 3,idploty = 2,numplotx = 2, idplotx = 2,
+                lowboundx = 8, title = TRUE,lowboundy = 10,
+                upboundy = 92,upboundx = 92),
+    xaxt="s",las=1,new=TRUE)
+
+adjsYlims<-abs(range(pop[seed==runChoi,initAct])[1]-
+                 range(pop[seed==runChoi,initAct])[2])*c(-0.1,0.2)
+
+evolDist(indData = pop[seed==runChoi],variable = "initAct",nbins = 10,
+         pal = pal_dist, nlevels=10,cexAxis = 1.5,xlab="",ylab = "",
+         xaxt="n",yaxt = "n",
+         ylim = range(pop[seed==runChoi,initAct])+adjsYlims,xlim=c(0,22000))
+legend("topright",legend = c("Behavioural tendency before learning"),cex=cexAxis,
+       col=colGenesLin[2],lwd=2,bty = "n")
+legend("topleft",legend = c("B.2"),cex=cexAxis,bty = "n",inset = -0.06)
+# grey lines to show the generations shown in the upper and lower panels
+matlines(x=matrix(rep(evolStats[genstoPrint,time],each=2),nrow=2),
+         y=matrix(rep(range(pop[seed==runChoi,initAct]))),
+                      lty=1,col = "grey",lwd=2)
+
+axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
+     labels = axTicks(1)[2:(length(axTicks(1)))]/1000,cex.axis=cexAxis,
+     hadj = 1.3,line = -0.5,tick = FALSE)
+
+axis(4,cex.axis=cexAxis)
+
+matlines(x=traitsTrajs[,time],
+         y=traitsTrajs[,.SD,
+                       .SDcol=paste0(c("meanInitAct_"),runChoi)],
+         col=colGenesLin[2],lty = 1,type="l",lwd=4)
+
+# black line showing the value 0 as a benchmark
+
+lines(x = pop[,range(time)],y=c(0,0))
+
+# Choose time range
+gen2plot<-round(seq(1,length(unique(evolStats$time)),length.out = 5))[2:5]
+# Plor the actor
+seqYax<-c("s",rep("n",3))
+# seqYlabUp<-c("Badge",rep("",3))
+seqYlabUp<-c("P(dove)",rep("",3))
+seqYlabDown<-c("Badge",rep("",3))
+seqXlabDown<-c("","Quality","")#paste0("seed: ",runChoi))
+seqXlabUp<-c("","Badge","")
+rangQual<-seq(0,1,length.out = 50)
+interv<-1/(nCenters-1)
+centers<-interv*seq(0,nCenters-1)
+rangx<-seq(0,1,length=50)
+count<-0
+for(genC in genstoPrint){
+  count<-count+1
+  par(plt=posPlot(numplotx = 4,numploty = 3,idplotx = count,idploty = 1,
+                  lowboundx = 8, title = TRUE,lowboundy = 10,
+                  upboundy = 92,upboundx = 92),
+      xaxt="s",las=1,new=TRUE)
+  tempPop<-popOneInd[(time==unique(time)[genC]&seed==runChoi),.SD[.N],
+               .SDcol=c(grep("WeightAct",
+                             names(evol),value = TRUE),"Quality","alpha","beta",
+                        "orderClus"),
+               by=indId]
+
+  randInds<-sample(tempPop$indId,nInd2print, replace = FALSE)
+
+# meansClustmp<-tempPop[,.(alphaMean=mean(alpha),betaMean=mean(beta)),by=orderClus]
+# reacNormpClust<-sapply(meansClustmp[,orderClus], function(x){
+#   logist(rangx,alpha = meansClustmp[orderClus==x,alphaMean],
+#          beta = meansClustmp[orderClus==x,betaMean])})
+# matplot(x = rangx,y = reacNormpClust,col=colRuns[meansClustmp$orderClus],
+#          type='l',cex.axis=cexAxis,
+#          xlab="",ylab="",ylim=c(0,1.1),lty=1,
+#          lwd=2,yaxt=seqYax[count],xaxt="s",cex.lab=cexAxis)
+  reacNormInd<-sapply(tempPop[indId %in%randInds,indId], function(x){
+    logist(rangx,alpha = tempPop[indId==x,alpha],
+           beta = tempPop[indId==x,beta])})
+  matplot(x = rangx,y = reacNormInd,
+          type='l',cex.axis=cexAxis,
+          xlab="",ylab="",ylim=c(0,1.1),lty=1,col=1,
+          lwd=2,yaxt=seqYax[count],xaxt="s",cex.lab=cexAxis)
+  mtext(text = seqYlabDown[count],cex = cexAxis,line = 1.9,side = 2,las=0)
+  mtext(text = seqXlabDown[count],cex = cexAxis,line = 1.7,side = 1,las=0)
+  legend("topleft",legend = c(paste0("C.",count)),bty = "n",inset = -0.06)
+  }
+
+rm(list=grep("temp",ls(),value = T))
+```
+
 \newpage

### 2023-04-27 dynamics.R

@@ -6,18 +6,20 @@
 
 library(here)
 here()
+library(dplyr)
 source(here("AccFunc.R"))
 
 
 # Scenario to be plotted - corresponds to folders where simulations are stored
 
-scenario<-"QualStDvEvol"
+scenario<-"betCostalphaL2"
 
 
 # Load files -------------------------------------------------------------------
 
 (listTest<-list.files(here("Simulations",paste0(scenario,"_"))))
-(sdList<-grep("evol",listTest,value=TRUE))
+(sdList<-grep("evol",listTest,value=TRUE)%>%
+    grep(pattern = ".txt",value = TRUE))
 
 evol<-fread(here("Simulations",paste0(scenario,"_"),sdList[1]))
 

### 2023-04-27 quadraticSelection.R

@@ -1,0 +1,66 @@
+# Playing around to visualize frequancy- dependent selection
+# 
+
+# Required libraries -----------------------------------------------------------
+
+require(here)
+here()
+source(here("AccFunc.R"))
+require("jsonlite")
+
+# Scenario to be plotted - corresponds to folders where simulations are stored
+
+scenario<-"betCostEvol4"#"alphaAct"#"nIntGroupNormQual"#
+
+# extSimsDir<-#here("Simulations",paste0(scenario,"_"))
+#   paste0("M:/BadgeSims/",scenario,"_")
+
+
+# Load files -------------------------------------------------------------------
+# Project folder
+(listTest<-list.files(here("Simulations",paste0(scenario,"_")),full.names = TRUE))
+# External sims folder
+# (listTest<-list.files(extSimsDir,full.names = TRUE))
+
+(evolList<-grep("evolLearn",listTest,value=TRUE))
+(indList<-grep("indLearn",listTest,value=TRUE))
+paramName<-list.files(here("Simulations",paste0(scenario,"_")))
+# paramName<-list.files(extSimsDir,full.names = TRUE)
+paramName<-grep(".json",paramName,value=TRUE)
+param<-fromJSON(here("Simulations",paste0(scenario,"_"),paramName[2]))
+# fromJSON(paramName[1])
+
+
+val<-1
+fileId<-val
+
+
+evolList_runs<-grep(paste0(param$namParam,param$rangParam[val]),
+                    evolList,value =TRUE)
+indList_runs<-grep(paste0(param$namParam,param$rangParam[val]),
+                   indList,value =TRUE)
+
+
+evol<-do.call(rbind,lapply(evolList_runs,fread))
+pop<-do.call(rbind,lapply(indList_runs, fread))
+
+
+lastInt<-tail(pop[,unique(nInteract)],1)
+
+popOneInd<-pop[nInteract==lastInt]
+
+pop[,unique(cumPayoff)]
+
+popOneInd.gen<-popOneInd[time==3000]
+
+str(popOneInd)
+
+
+
+multReg.gen <- lm(data = popOneInd.1000,cumPayoff/ntotInteract~Quality+Badge+I(Badge^2))
+
+anova(multReg.gen)
+
+
+
+

### 2023-04-27 manuscript/manuscript_1.0.Rmd
Changes to the reference list

--- 2023-03-31 manuscript/manuscript_1.0.Rmd
@@
 mechanism whose underpinnings show interspecific variation 
-[@enquist_Power_2016; @quinones_Reinforcement_2019]. Associative learning 
+[@enquist_Power_2016; @quinones_Reinforcement_2020]. Associative learning 
 is not a single mechanisms, but rather a set of cognitive structures 
@@
 through a learning process. We implement learning using 
-the actor-critic approach from reinforcement learning (RL) theory [@sutton_Reinforcement_2018;@quinones_Reinforcement_2019; @leimar_Learning_2019].
+the actor-critic approach from reinforcement learning (RL) theory [@sutton_Reinforcement_2018;@quinones_Reinforcement_2020; @leimar_Learning_2019;
+@leimar_Evolution_2021].
 Individuals estimate the reward (pay-off) expected from  interacting 
@@
 ```{r model-struc,echo=FALSE,warning=FALSE,fig.width=8,fig.align='center',fig.height=7,fig.show='hold',fig.cap= "Model of communication in the context of aggressive interactions.
-In A, the reaction norm determines the badge size.
+In A, the reaction norm determines the bib (badge) size, black band on birds chest.
 Red shows an informative reaction norm, while the blue shows a uninformative reaction norm.
@@
 can have in a social contexts such as: hierarchy formation 
-[@leimar_Reproductive_2022;@leimar_Effects_2022],
+[@leimar_Evolution_2021;@leimar_Reproductive_2022;@leimar_Effects_2022],
 and cooperation [@leimar_Learning_2019;@dridi_Learning_2018]. In this 

### 2023-05-03 manuscript/manuscript_1.0.Rmd
New ways to asses the effect of frequency dependent selection on the signal

--- 2023-04-27 manuscript/manuscript_1.0.Rmd
@@
   partner of an individual. For example, when competing for resources, an individual
-  might be better off escalating fights with individuals of lower quality, while restraining
-  from fighting individuals of higher quality. Communication systems involving signals
-  of quality allow individuals to reduce uncertainty regarding the fighting ability
-  of their partners and make more adaptive behavioral decisions. However, dishonest
-  individuals can destabilize such communications systems. One open question is whether
-  cognitive mechanisms, such as learning, can maintain the honesty of signals, thus,
-  favoring their evolutionary stability. Here we present evolutionary simulations,
-  where individuals can produce a signal proportional to their quality; and learn
-  along their lifetime the best response to the signal emitted by their peers. In
+  might be better off escalating fights with individuals of lower quality, while
+  restraining  from fighting individuals of higher quality. Communication systems
+  involving signals of quality allow individuals to reduce uncertainty regarding the  
+  fighting ability of their partners and make more adaptive behavioral decisions. 
+  However, dishonest individuals can destabilize such communications systems. 
+  One open question is whether cognitive mechanisms, such as learning, 
+  can maintain the honesty of signals, thus, favoring their evolutionary 
+  stability. Here we present evolutionary simulations, where individuals 
+  can produce a signal proportional to their quality; and learn along 
+  their lifetime the best response to the signal emitted by their peers. In
   these simulations, learning on the receiver side can mediate the evolution of signals
@@
 mtext(text = "Trait value",cex = cexAxis,line = 1,side = 2,las=0)
-legend("topright",legend = c(expression(alpha[s])),col=colGenesLin[1],lwd=2,
-       bty = "n",cex=cexAxis)
+legend("topright",legend = c(expression(alpha[s])), bty = "n",cex=cexAxis)
 legend("topleft",legend = c("A.1"),cex=cexAxis,bty = "n",inset = -0.06)
@@
                        .SDcol=paste0(c("meanAlpha_"),runChoi)],
-         col=colGenesLin,lty = 2,type="l",lwd=5)
+         col=colGenesLin,lty = 1,type="l",lwd=5)
 
@@
          ylim = range(pop[seed==runChoi,beta])+adjsYlims,xlim=c(0,22000))
-legend("topright",legend = c(expression(beta[s])),cex=cexAxis,
-       col=colGenesLin[2],lwd=2,bty = "n")
+legend("topright",legend = c(expression(beta[s])),cex=cexAxis,bty = "n")
 legend("topleft",legend = c("A.2"),cex=cexAxis,bty = "n",inset = -0.06)
@@
                        .SDcol=paste0(c("meanBeta_"),runChoi)],
-         col=colGenesLin[2],lty = 2,type="l",lwd=4)
+         col=colGenesLin,lty = 1,type="l",lwd=4)
 
@@
           type='l',cex.axis=cexAxis,
-          xlab="",ylab="",ylim=c(0,1.1),lty=1,col=1,
+          xlab="",ylab="",ylim=c(0,1.1),lty=1,col="grey",
           lwd=2,yaxt=seqYax[count],xaxt="s",cex.lab=cexAxis)
@@
 imply that individuals can generally be classified into three 
-distinct types (\ref{fig:branching}). Two types express a flat reaction 
+distinct types (Fig. \ref{fig:branching}). Two types express a flat reaction 
 norm with extreme values for the badge size, meaning that their badge size 
@@
 mtext(text = "Trait value",cex = cexAxis,line = 1,side = 2,las=0)
-legend("topright",legend = c(expression(alpha[s])),col=colGenesLin[1],lwd=2,
+legend("topright",legend = c(expression(alpha[s])),
        bty = "n",cex=cexAxis)
@@
                           seed == repsScen[count]],
-       alpha~beta,col=multDiscrPallet[seed+1],cex=2,pch=20)
+       alpha~beta,col=multDiscrPallet[seed+1],cex=1,pch=20)
 
@@
 ```
+
+The action of natural selection and genetic drift do not override the conditions 
+for the evolutionary branching described earlier. Given how crucial naive 
+behaviour is for the diversification process described earlier, we tested whether 
+evolutionary processes could lead key parameters of the learning process toward
+values that prevented the genetic variation to build up. 
+Figure \ref{fig:learnEvol} shows the dynamics from a set of simulations where we 
+allow for the coevolution of the reaction norm as well as key parameters of
+the learning module, namely the speed of learning ($A$) and the behavioural 
+tendency expressed by individuals before learning. Top panels in 
+figure \ref{fig:learnEvol} show that there is diversification process in this 
+coevolutionary scenario as well. That is because, as is shown in the middle 
+panels (B) of the same figure, the learning parameters stay within the range 
+necessary for the diversification process. The speed of learning 
+(Fig. \ref{fig:learnEvol} B.1) stays well above zero throughout the simulation. 
+In contrast, the initial behavioral tendency maintain values around zero. 
+Thus, evolutionary processes acting on the learning parameteres 
+did not override the conditions necessary for frequency dependent to boost 
+genetic variation.
 
@@
 mtext(text = "Trait value",cex = cexAxis,line = 1,side = 2,las=0)
-legend("topright",legend = c(expression(alpha[s])),col=colGenesLin[1],lwd=2,
-       bty = "n",cex=cexAxis)
+legend("topright",legend = c(expression(alpha[s])),bty = "n",cex=cexAxis)
 legend("topleft",legend = c("A.1"),cex=cexAxis,bty = "n",inset = -0.06)
@@
          ylim = range(pop[seed==runChoi,beta])+adjsYlims,xlim=c(0,22000))
-legend("topright",legend = c(expression(beta[s])),cex=cexAxis,
-       col=colGenesLin[2],lwd=2,bty = "n")
-legend("topleft",legend = c("B"),cex=cexAxis,bty = "n",inset = -0.06)
+legend("topright",legend = c(expression(beta[s])),cex=cexAxis,bty = "n")
+legend("topleft",legend = c("A.2"),cex=cexAxis,bty = "n",inset = -0.06)
 # grey lines to show the generations shown in the upper and lower panels
@@
 
-```{r learnEvol,echo=FALSE,out.width='100%',fig.height=6,fig.align='center',fig.show='hold',fig.cap= "The evolution of costly signals.
-Portrait of the evolutionary dynamics of the sender code with snapsots of both sender and receiver codes just as in fig.
-\\ref{fig:handicap}.
-The middle panels show changes in the distribution of values for $\\alpha$ and $\\beta$ along evolutionary time.
-Panels above and below correspond to snapshots of the sender and receiver codes, respectively, generation time of the snapshots are indicated by the grey lines in the middle panels.
-Similarly to the evolutionary dynamics in Fig.
-\\ref{fig:handicap}, first the badge size evolves towards its minimum value.
-Then, the slope evolves positive values determining an increasing reaction norm.
-In contrast to \\ref{fig:handicap}, the normal distribution splits up into two modes.
-This translates into two types of reaction norms assorting in the population where both code for a positive relation between quality and badge size."}
+```{r learnEvol,echo=FALSE,out.width='100%',fig.height=6,fig.align='center',fig.show='hold',fig.cap= "Coevolutionary dynamics of both the signal reaction norm and key parameters of the learning module.
+Dynamics are portrait as changes in the distribution of values assorting in the population.
+In the top panels (A)  evolutionary dynamics of the sender code; A.1 shows the intercept and A.2 the slope of the logistic reaction norm.
+Both of these parameters experience evolutionary branching, whereby in the end three distincs types are assorting in the population.
+The middle panels (B) show changes in the distribution of values for the speed of learning ($A$;B1) and the behavioural tendency before learning ($y_z^0$; B2) along evolutionary time.
+The speed of learning does not change drastically, it maintains values well above 0 (black line).
+In contrast, the initial behavioural tendency does not evolve away from 0 (black line).
+Both of these conditions favour evolutionary branching in the signal reaction norm.
+Bottom Panels correspond to snapshots of the sender code.
+Generation time of the snapshots are indicated by the grey lines in the top and middle panels."}
 
@@
 legend("topright",legend =
-      expression(paste("intercept (",alpha[s],")")),col=colGenesLin[1],lwd=2,
-       bty = "n",cex=cexAxis)
+      expression(paste("intercept (",alpha[s],")")),bty = "n",cex=cexAxis)
 legend("topleft",legend = c("A.1"),cex=cexAxis,bty = "n",inset = -0.06)
@@
 legend("topright",legend = expression(paste("Slope (",beta[s],")")),cex=cexAxis,
-       col=colGenesLin[2],lwd=2,bty = "n")
+       bty = "n")
 legend("topleft",legend = c("A.2"),cex=cexAxis,bty = "n",inset = -0.06)
@@
                        .SDcol=paste0(c("meanBeta_"),runChoi)],
-         col=colGenesLin[2],lty = 1,type="l",lwd=4)
+         col=colGenesLin,lty = 1,type="l",lwd=4)
 
@@
 mtext(text = "Trait value",cex = cexAxis,line = 1.8,side = 2,las=0)
-legend("topright",legend = "Speed of learning",col=colGenesLin[1],lwd=2,
-       bty = "n",cex=cexAxis)
+legend("topright",legend = "Speed of learning", bty = "n",cex=cexAxis)
 legend("topleft",legend = c("B.1"),cex=cexAxis,bty = "n",inset = -0.06)
@@
          ylim = range(pop[seed==runChoi,initAct])+adjsYlims,xlim=c(0,22000))
-legend("topright",legend = c("Behavioural tendency before learning"),cex=cexAxis,
-       col=colGenesLin[2],lwd=2,bty = "n")
+legend("topright",legend = c("Naive behavioural tendency"),cex=cexAxis,
+       bty = "n")
 legend("topleft",legend = c("B.2"),cex=cexAxis,bty = "n",inset = -0.06)
@@
                        .SDcol=paste0(c("meanInitAct_"),runChoi)],
-         col=colGenesLin[2],lty = 1,type="l",lwd=4)
+         col=colGenesLin,lty = 1,type="l",lwd=4)
 

### 2023-05-03 quadraticSelection.R

@@ -4,13 +4,13 @@
 # Required libraries -----------------------------------------------------------
 
 require(here)
-here()
+library(ggplot2)
 source(here("AccFunc.R"))
 require("jsonlite")
 
 # Scenario to be plotted - corresponds to folders where simulations are stored
 
-scenario<-"betCostEvol4"#"alphaAct"#"nIntGroupNormQual"#
+scenario<-"betCostEvol5"#"alphaAct"#"nIntGroupNormQual"#
 
 # extSimsDir<-#here("Simulations",paste0(scenario,"_"))
 #   paste0("M:/BadgeSims/",scenario,"_")
@@ -51,15 +51,92 @@ popOneInd<-pop[nInteract==lastInt]
 
 pop[,unique(cumPayoff)]
 
-popOneInd.gen<-popOneInd[time==3000]
+popOneInd[,unique(time)]
+
+popOneInd.gen<-popOneInd
+
+popOneInd.gen[,dist:=sqrt((Badge-mean(Badge))^2)]
 
 str(popOneInd)
 
+ggplot(popOneInd.gen,aes(y=cumPayoff,x=Quality))+
+  geom_point(aes(color=as.factor(seed)))+
+  geom_smooth()+
+  theme_classic()
+
+ggplot(popOneInd.gen,aes(y=cumPayoff,x=Badge))+
+  geom_point(aes(color=as.factor(seed)))+
+  geom_vline(xintercept = median(popOneInd.gen$Badge))+
+  geom_smooth()+
+  theme_classic()
+
+ggplot(popOneInd.gen,aes(y=cumPayoff,x=dist))+
+  geom_point(aes(color=as.factor(seed)))+
+  geom_smooth()+
+  theme_classic()
+
+nbins<-30
+
+ggplot(popOneInd.gen,aes(x=Badge))+
+  geom_histogram(bins = nbins)+
+  theme_classic()
+
+par(mfrow=c(1,1),plt=posPlot())
+
+hist.badge<-hist(popOneInd.gen$Badge,breaks = nbins,freq = TRUE)
+
+popOneInd.gen[,binFreq:=sapply(Badge,function(x){
+  hist.badge$density[which.min(abs(hist.badge$mids-x))]
+})]
+
+str(popOneInd.gen)
+
+ggplot(popOneInd.gen,aes(y=cumPayoff,x=binFreq))+
+  geom_point(aes(color=as.factor(seed)))+
+  geom_smooth()+
+  theme_classic()
[…]

### 2023-05-29 manuscript/manuscript_1.0.Rmd
New version of the manuacrip. New files to manage the authors affiliations

@@
 author:
-- Andrés E. Quiñones^[Laboratorio de Biología Evolutiva de Vertebrados, Departamento
-  de Ciencias Biológicas, Universidad de los Andes, Bogotá, Colombia]
-- C. Daniel Cadena*
-- Redouan Bshary^[Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland]
+- Andrés E. Quiñones:
+    email: "andresqp@gmail.com"
+    institute: [andes, unine]
+    correspondence: true
+- Redouan Bshary:
+    institute: unine
+- C. Daniel Cadena:
+    institute: andes
+institute:
+- andes: Laboratorio de Biología Evolutiva de Vertebrados, Departamento  de Ciencias Biológicas, Universidad de los Andes, Bogotá, Colombia
+- unine: Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland
 date: "`r format(Sys.time(), '%B %d, %Y')`"
 output:
+  # word_document: default
   pdf_document:
-    latex_engine: xelatex # default  # word_document: default
+    latex_engine: xelatex # default  #
+    pandoc_args:
+      - '--lua-filter=scholarly-metadata.lua'
+      - '--lua-filter=author-info-blocks.lua'
 fig_caption: yes
@@
   stability. Here we present evolutionary simulations, where individuals 
-  can produce a signal proportional to their quality; and learn along 
-  their lifetime the best response to the signal emitted by their peers. In
-  these simulations, learning on the receiver side can mediate the evolution of signals
+  can produce a signal proportional to their quality; and learn along their
+  lifetime the best response to the signal emitted by their peers. Our simulations
+  replicate previous results where the handicap principle mediates the evolution 
+  of signals. As in traditional models without learning, in our simulations 
+  learning on the receiver side can mediate the evolution of signals
   of quality on the sender side. When the cost of the signal is proportional to the
   quality of the sender populations are only composed of honest signalers. When the
-  cost is not proportional to the quality of the signaler, the population is composed
-  of both honest and dishonest signalers. We argue that learning can be a general
-  cognitive mechanism playing a role in a wide range of communication systems."
+  cost is not proportional to the quality of the signaler tranditional models 
+  predict the absense of a signal, in contrast learning facilitates the 
+  evolution of a polymorphim where population is composed of both honest and 
+  dishonest signalers. We argue that learning can be a general cognitive mechanism 
+  role wide range of communication systems and more generally
+  behavioral responses."
 ---
@@
 exhibit badges of status, where a signal conveys quality 
-[@johnstone_Badges_1993; @rohwer_Social_1975a]. 
+[@johnstone_Badges_1993; @rohwer_Social_1975]. 
 Such signals must be arbitrary, meaning the signals could potentially be 
@@
 capturing the essence of learning processes 
-[@frankenhuis_Enriching_2018; @quinones_Reinforcement_2019]. 
+[@frankenhuis_Enriching_2018; @quinones_Reinforcement_2020]. 
 Furthermore, these algorithms can be embedded in evolutionary simulations to 
@@
 # Top panels -  A the sender code and  B learning for the receiver code
-
+# load(here("manuscript","manuscript_1.0_cache","latex","model-struc_6905f4797a1faa21041511a8dcdf1eb9.RData"))
 library(dplyr)
@@
   legend("topleft",legend = c(paste0("C.",count)),bty = "n",inset = -0.06)
+  if(count==1){
+   color.bar.aeqp(paletteMeans(100),min =min(colorbreaksQual),
+               max = max(colorbreaksQual),nticks = 3,
+               cex.tit = 0.9,title = "",
+               numplotx = 20,numploty = 14,idplotx =3,idploty = 1) 
+  }
   # text(x=0.5,y=-0.015,labels = paste0("time=",unique(evolStats$time)[genC]/1000),
@@
   #      cex=2)
+  if(count==1){
+   color.bar.aeqp(paletteMeans(100),min =min(colorbreaksQual),
+               max = max(colorbreaksQual),nticks = 3,
+               cex.tit = 0.9,title = "",
+               numplotx = 20,numploty = 14,idplotx =3,idploty = 1) 
+  }
   par(plt=posPlot(numplotx = 4,numploty = 3,idplotx = count,idploty = 2,
@@
 
-### The peaceful, the clever and the aggressive
+### The peaceful, the aggressive and the clever
 
@@
 ylabAll<-c(expression(alpha[s]),"","")
-titleAll<-c("Peaceful","Clever","Aggressive")
-colorAll<-c("dodgerblue2","green4","#E31A1C")
+titleAll<-c("Peaceful","Aggressive","Clever")
+colorAll<-c("dodgerblue2","#E31A1C","green4")
 multDiscrPallet[repsScen+1]<-colorAll
 plot.new()
-for(PAr in popOneInd[,sort(unique(get(shortSce)),decreasing = TRUE)]){
+for(PAr in popOneInd[,unique(get(shortSce))][c(3,2,1)]){
   count<-match(PAr,popOneInd[,sort(unique(get(shortSce)),decreasing = TRUE)])
@@
 a limit in the amount of information collected by individuals, as well as
-peaceful behaviour in naive individuals.
+peaceful behaviour in naive individuals. Furthermore, we have shown that those 
+conditions can be reached when evolutionary processes drive changes in the learning
+system. Thus, the diversification process we present might play a role in natural
+populations.
 
@@
 Because individuals learn about their own quality, 
-even in the absence of an honest signal, they are able to make adaptive choices. 
-Thus the learning process we have modelled and the response it mediates 
-has fitness relevant consequences even in the absence of an honest signal.
+even in the absence of an honest signal, they are able to make more 
+adaptive choices. Thus the learning process we have modelled and the 
+response it mediates has fitness relevant consequences even in the 
+absence of an honest signal. This is confirmed by simulation where 
+the speed of learning is allowed to change subject to evolutionary processes. 
+In those simulations, natural selection always
+maintained learning rates above zero. Our simulations however, 
+only captured the effect of selection mediated by the social 
+game on the learning parameters. Inter-specific variation in cognitive 
+abilities mediated by environmental differences has been reported elsewhere
+[@sonnenberg_Natural_2019]. If such variation is partly mediated by changes 
+that affect general cognitive processes, they could impact the outcome 
+of communication systems like the one we modelled.
 
@@
 
-Are bibs real signals of quality? if so are they handicaps or badge of status? 
-Originally @rohwer_Social_1975a proposed the idea that certain morphological 
+Originally @rohwer_Social_1975 proposed the idea that certain morphological 
 traits could be used by animals as status signals or signals of quality. 
-However, from an evolutionary perspective it is unclear how the stability of 
-such communication systems can be maintained. One option is that honesty is 
+This idea has been tested repeatedly on the dark patches of some bird species 
+(bibs). So, are bibs real signals of quality? if so are they handicaps or 
+badge of status? From an evolutionary perspective it is unclear how the stability of 
+status signals can be maintained. One option is that honesty is 
 maintained by the cost of signal production; that is, when the signal works as a 
@@
 particularly in the mid-range of the distribution facilitates the establishment 
-of a convention, in this range individuals respond appropriately to the trait of their 
-peer. The house sparrow (*Passer domesticus*) has been a text-book example 
-of badges of status. According to the prevailing narrative the bib size 
-in males is a signal of the dominance rank and quality.
+of a convention, in this range individuals respond appropriately to the 
+trait of their peer. The house sparrow (*Passer domesticus*) 
+has been a text-book example of badges of status. According to the prevailing narrative the bib size in males is a signal of the dominance rank and quality.
 However, a recent meta-analysis called into question 
 this narrative by showing that the effect size of the association between bib 
-size and dominance rank is small and uncertain [@sanchez-tojar_Metaanalysis_2018a]. 
+size and dominance rank is small and uncertain [@sanchez-tojar_Metaanalysis_2018]. 
 The simple straight correlation between bib size and dominance rank
-( and quality) is expected under the handicap principle, as it is shown here and 
+(and quality) is expected under the handicap principle, as it is shown here and 
 elsewhere [@botero_Evolution_2010;@grafen_Biological_1990;@johnstone_Badges_1993].

### 2023-06-27 manuscript/manuscript_1.0.Rmd
New word file for Daniel to comment

@@
 output:
+  word_document: default
   # word_document: default

### 2023-07-19 manuscript/manuscript_1.0.Rmd
Addressed Daniel's comments

--- 2023-05-29 manuscript/manuscript_1.0.Rmd
@@
 output:
+  word_document: default
   # word_document: default
-  pdf_document:
+  # pdf_document:
     latex_engine: xelatex # default  #
@@
   However, dishonest individuals can destabilize such communications systems. 
-  One open question is whether cognitive mechanisms, such as learning, 
-  can maintain the honesty of signals, thus, favoring their evolutionary 
-  stability. Here we present evolutionary simulations, where individuals 
-  can produce a signal proportional to their quality; and learn along their
-  lifetime the best response to the signal emitted by their peers. Our simulations
+  An open question is whether cognitive mechanisms, such as learning, 
+  can maintain the honesty of signals, thus favoring their evolutionary 
+  stability. Here we present evolutionary simulations where individuals 
+  can produce a signal proportional to their quality and learn along their
+  lifetime the best response to the signals emitted by their peers. Our simulations
   replicate previous results where the handicap principle mediates the evolution 
@@
   of quality on the sender side. When the cost of the signal is proportional to the
-  quality of the sender populations are only composed of honest signalers. When the
-  cost is not proportional to the quality of the signaler tranditional models 
-  predict the absense of a signal, in contrast learning facilitates the 
-  evolution of a polymorphim where population is composed of both honest and 
-  dishonest signalers. We argue that learning can be a general cognitive mechanism 
-  role wide range of communication systems and more generally
-  behavioral responses."
+  quality of the sender, populations are only composed of honest signalers. In contrast 
+  to traditional models which predict the absence of signals when the
+  cost is not proportional to the quality of the signaler, our model revealed that
+  learning facilitates the evolution of a polymorphism in which populations are
+  composed of both honest and dishonest signalers. We argue that learning can 
+  be a general cognitive mechanism having a role in a wide range of communication systems 
+  and more generally in behavioral responses.
 ---
@@
 
-The outcome of interactions with conspecifics individuals is an important 
+The outcome of interactions with conspecific individuals is an important 
 determinant of fitness in social animals. Irrespective of whether 
@@
 acquiring such information is far from trivial. 
-In some cases, interacting partners (signallers) might be 'willing' to 
+In some cases, interacting partners (*i.e.* signallers) might be 'willing' to 
 provide accurate information, but in others it might be in their 
-own interest to conceal information [@johnstone_Recognition_1997], 
+own interest to conceal information [@johnstone_Recognition_1997] 
 or to provide deceptive information [@johnstone_Badges_1993]. 
@@
 exhibit badges of status, where a signal conveys quality 
-[@johnstone_Badges_1993; @rohwer_Social_1975]. 
+[@johnstone_Badges_1993; @rohwer_Social_1975a]. 
 Such signals must be arbitrary, meaning the signals could potentially be 
@@
 pay a higher fitness cost for carrying the signal such that it is not 
-in the interest of low quality individuals 
-to fake quality by producing a large badge. Signal costs inversely proportional
-to quality can come about as a result of the production costs of the signal, 
-in which case the signal works like a handicap [@botero_Evolution_2010; @johnstone_Badges_1993;@grafen_Biological_1990; @zahavi_Mate_1975]. 
+in the interest of low quality individuals to fake quality by producing a large badge.
+Signal costs inversely proportional to quality may come about as a result of the
+production costs of the signal, in which case the signal works like a handicap
+[@botero_Evolution_2010; @johnstone_Badges_1993;@grafen_Biological_1990;@zahavi_Mate_1975]. 
 Alternatively, the costs can come from 
 the aggressive reaction of receivers when the convention of the communication 
-is broken [@enquist_Signaling_2010;@tibbetts_Socially_2004]. 
+is broken [@enquist_Signaling_2010;@tibbetts_Socially_2004]. For example in paper 
+wasps, subordinate individuals with experimentally manipulated dominant-like facial
+patterns received more aggression from the dominant. In this case, the costs are 
+socially mediated rather than developmentally and are triggered by the violation 
+of the convention. It is still incluer under this narrative, how the convenvention 
+is stablished in the first place.
+
 <!-- should this last part of the paragraph stay ?-->
@@
 This contrasts with other information-processing 
-mechanisms like individual recognition. With individual recognition, 
-where individuals associate cues of their peers to their quality. Because these 
+mechanisms like individual recognition. Where individuals associate cues of 
+their peers to their quality. Because these 
 associations must be learnt throughout the life of individuals, 
@@
 recognition. In cases where badges of status vary quantitatively 
-(e.g. in size or intensity), fighting ability may increase monotonically 
+(*e.g.* in size or intensity), fighting ability may increase monotonically 
 with attributes of the badge and would be reinforced by every interaction.
@@
 Associative learning is a taxonomically widespread cognitive 
-mechanism that allows individuals to associate rewards with environmental stimuli 
-and thus behave adaptively 
+mechanism that allows individuals to associate rewards with 
+environmental stimuli and thus behave adaptively 
 [@heyes_Simple_2012;@macphail_Brain_1982; @staddon_Adaptive_2016; 
-@behrens_Associative_2008].
-Theory has shown that natural selection favours such associations 
-in complex environments where conditions are difficult to predict 
-[@dridi_Environmental_2016]. Associative learning is a flexible cognitive 
+@behrens_Associative_2008]. Theory has shown that natural selection 
+favours such associations in complex environments where conditions 
+are difficult to predict [@dridi_Environmental_2016]. 
+Associative learning is a flexible cognitive 
 mechanism whose underpinnings show interspecific variation 
-[@enquist_Power_2016; @quinones_Reinforcement_2020]. Associative learning 
-is not a single mechanisms, but rather a set of cognitive structures 
-and processes that can vary in their scope and complexity. Presumably, 
-these structures and processes have been modified 
+[@enquist_Power_2016; @quinones_Reinforcement_2020;@pretot_Factors_2016;
+pretot_Comparative_2021]. For example, species can have different learning rates
+[@hoedjes_Natural_2010],
+some may vary these rates during learning trials which allows them to have more 
+flexible learning [@leimar_Flexibility_2023a], and some may include future rewards in their associations [raby_Planning_2007;osvath_Spontaneous_2009].
+Associative learning is not a single mechanism, but rather a set of 
+cognitive structures and processes that can vary in their scope and complexity. Presumably, these structures and processes have been modified 
 by natural selection and could provide new explanations for behavioural 
-variation. Despite all this, associative learning is 
+variation. Nonetheless, associative learning is 
 not often included in the narrative of evolutionary explanations of 
@@
 Here we present an evolutionary model where individuals use 
-associative learning to develop a tendency to behave aggressively 
-(or  peacefully) in the context of competition over resource, 
+associative learning to develop a tendency to behave either aggressively 
+or  peacefully in the context of competition over resources, 
 depending on a quantitative morphological trait 
-(i.e.: a badge indicating quality) they 
-perceive in their opponents. Over evolutionary time the size of 
-the badge evolves, and its dependency on the individual's quality. 
+(*i.e.* a badge indicating quality) they perceive in their opponents. 
+Over evolutionary time, the size of 
+the badge evolves as does its dependency on the individual's quality. 
 Under this simple set up, individuals can use the badge as a signal of quality. 
@@
 unless, with a small probability ($\mu$), mutation changes the allelic value 
-of the offspring by an amount taken from a normal distribution $TN(0,\sigma_\mu)$.
+of the offspring by an amount drawn from a normal distribution $TN(0,\sigma_\mu)$.
 
@@
 Importantly, if $k_2=0$ survival probability is independent of quality, 
-while if $k_2>0$ lower quality individuals pay a higher price for similar-sized 
-badge, fulfilling the core assumption of the handicap principle 
+while if $k_2>0$ lower-quality individuals pay a higher price for similar-sized 
+badges, fulfilling the core assumption of the handicap principle 
 [@botero_Evolution_2010; @grafen_Biological_1990; @johnstone_Badges_1993].
@@
 Following the classic *hawk-dove* game [@maynard-smith_Evolution_1982], 
-if the focal individual fights and its partner does not, the focal gets 
+if the focal individual fights and its partner does not, then the focal gets 
 as pay-off the resource of value $V$ while its partner gets nothing; 
 if both individuals restrain  from fighting, then they split up the resource 
-in half.
-If both individuals decide to fight, the winner takes over the resource and the cost of fighting is split among the two players.
-We further assume that the probability of wining a fight for the focal individual depends on the
-difference in quality between it and its interacting partner; specifically 
+in half. If both individuals decide to fight, then the winner takes over 
+the resource and the cost of fighting is split between the two players. 
+We further assume that the probability of wining a fight for the focal 
+individual depends on the difference in quality between it and its 
+interacting partner; specifically 
 it is given by $p_{ij}=1/(1+e^{-k_3(Q_i-Q_j)})$, where $k_3$ is a 
@@
 partner can depend on the size of the partner's badge, 
-the dependency is determined by the focal's experiences 
+the dependency is determined by the focal's experiences acquired
 through a learning process. We implement learning using 
@@
 (the actor in RL). They update the probability of 
-retreating/attacking  up or down depending on whether retreating/attacking
+retreating/attacking depending on whether retreating/attacking
 leads to an increase in the reward estimation. 
 So, if a focal individual decides to escalate a fight against 
-an individual with small badge and this leads to an increase in 
+an individual with a small badge and this leads to an increase in 
 the reward estimation, then the focal individual will increase the probability 
@@
 transformation (Fig. \ref{fig:model-struc} B, black line). 
-So formally, the log-odds to retreat when facing individual $j$ is
+Thus formally, the log-odds to retreat when facing individual $j$ is
 
@@
 
-Individuals interact as the focal individual $n$ times, and the interaction 
+Individuals interact as the focal individual $n$ times and the interaction 
 partner is chosen at random. Thus, the expected number of interactions for 
@@
 In A, the reaction norm determines the bib (badge) size, black band on birds chest.
-Red shows an informative reaction norm, while the blue shows a uninformative reaction norm.
+Red shows an informative reaction norm, whereas blue shows a uninformative reaction norm.
 B individuals perceiving the signal have a behavioral reaction norm that determines their probability of retreating (black line).
@@
 Feature weaights are values associated with a particular badge size and are updated as individuals interact with each other.
-The generalization of the information associated with each feature weight is represented by the grey line and its axis, which shows the response triggered by the fourth feature diminishes as the value evaluated is further from the feature center.
-The learning process moves the feature weights (black dots) up or down depending on whether that leads to an increase in the estimated reward.
+The generalization of the information associated with each feature weight is represented by the grey line and its axis, which shows that the response triggered by the fourth feature diminishes as the value evaluated is further from the feature center.
+The learning process moves the feature weights (black dots) up or down depending on whether this leads to an increase in the estimated reward.
 In C and D, effect of learning on the receiver strategy.
 Receivers in C face signallers with uninformative reaction norms (blue line in A).
-The uninformative reaction norm yields badge size of 0.5, thus the behavioural reaction norm in C is only updated around that value.D Receivers face signallers with informative reaction norms (red line in A).
+The uninformative reaction norm yields a badge size of 0.5, thus the behavioural reaction norm in C is only updated around that value.
+D Receivers face signallers with informative reaction norms (red line in A).
 Accordingly, receivers develop a threashold-like reaction norm where the decision to retreat increases with the badge size of the interacting partner.
@@
 
-### Learning under a monomorphic population
-
-We first present the outcome of simulations where we prevent the 
+### Learning in a monomorphic population
+
+We first present the outcome of simulations where we prevented the 
 evolution of the badge size (by setting the mutation rate to $0$), 
-and assume that all individuals in the population display either 
+and assumed that all individuals in the population display either 
 an uninformative or informative badge size with respect to their quality. 
 These simulations show what type of receiver strategy develops through 
-a learning process. Figure \ref{fig:model-struc} (C and D) shows 
+a learning process in such monomorphic population. 
+Figure \ref{fig:model-struc} (C and D) shows 
 the receiver strategy developed through learning; panel C is for 
 receivers that faced uninformative signals, and D informative ones. 
-The simulations reveal that when learners face uninformative signals, 
-they modify their probability of retreating depending on their own quality. 
-Individuals with higher quality (red tones) after the learning process have a 
-high probability of attacking; while individuals of lower quality 
-(blue tones) mostly retreat from confrontations; in both cases irrespective of 
+The simulations reveal that when learners face uninformative signals (Fig. \ref{fig:model-struc} C), they modify their probability of retreating 
+depending on their own quality. 
+Individuals with higher quality (red tones) have a 
+high probability of attacking after the learning process; 
+while individuals of lower quality 
+(blue tones) mostly retreat from confrontations; in both cases 
+individuals behave equally irrespective of the quality of
 their opponent. Thus, learning splits the population of receivers into 
@@
 with unresponsive reaction norms on the signalling side, the changes 
-triggered by learning only affect a small range of badge sizes. This can be seen 
-in panel C, all learning has occurred around a badge size of $0.5$, because all
-badges in the population are this size. 
+triggered by learning only affect a small range of badge sizes. Specifically, 
+all learning has occurred around a badge size of $0.5$, because all
+badges in the population are of this size (panel C). 
 In contrast, when receivers face informative reaction norms 
-on the side of the signaller (panel B in fig. \ref{fig:model-struc}), 
-receivers use the badge size of their interacting partners to determine 
-whether to retreat or attack. The relationship is given by a 
-threshold-like reaction norm, where the decision to retreat or 
-attack depends on the quality of the receiver. As expected, 
-the higher the quality of the receiver the larger the badge size 
-that triggers a retreat.
+on the side of the signaller (Fig. \ref{fig:model-struc} B), 
+receivers use information on the badge size of their interacting 
+partners to determine whether to retreat or attack. 
+The relationship is given by a threshold-like reaction norm, 
+where the decision to retreat or to attack depends on the quality of the receiver. 
+As expected, the higher the quality of the receiver, the larger the badge size 
+of the signaller that triggers a retreat.
 
@@
 On the left (A.1), changes in the distribution of values for the intercept of the reaction norm ($\\alpha$); on the right (A.2) changes in the distribution of the slope ($\\beta$).
-Darker areas of the background correspond to values with high frequency, while dashed lines in both panels show the mean of the distribution.
+Darker areas of the background correspond to values with high frequency, while orange lines in both panels show the mean of the distribution.
 X axis in the evolutionary dynamics are given in thousands of generations.
@@
 On the first half of the simulation (B.1 and B.2) the badge size evolves to its minimum value.
-In the second half, the value of the slope takes positive values and reaction norms evolve to produce an honest signal of quality.
-Correspondingly, individuals learn to react to badge size by increasing the probability of retreat with increasing badge sizes"  ,warning=FALSE}
+In the second half, the slope takes positive values and reaction norms evolve to produce an honest signal of quality.
+Correspondingly, individuals learn to react to badge size by increasing the probability of retreating with increasing badge sizes"  ,warning=FALSE}
 
@@
 # seqYlabUp<-c("Badge",rep("",3))
-seqYlabUp<-c("P(dove)",rep("",3))
-seqYlabDown<-c("Badge",rep("",3))
+seqYlabUp<-c("P(retreat)",rep("",3))
+seqYlabDown<-c("Badge size",rep("",3))
 seqXlabDown<-c("","Quality","")#paste0("seed: ",runChoi))
-seqXlabUp<-c("","Badge","")
+seqXlabUp<-c("","Badge size","")
 rangQual<-seq(0,1,length.out = 50)
@@
 
-When we allow the badge size to evolve ($\alpha$ and $\beta$ 
+When we allowed the badge size to evolve ($\alpha$ and $\beta$ 
 change subject to natural selection and genetic drift), and the signal
-works as a handicap (i.e. the cost of the badge is inversely 
-proportional to the quality of the individual), the sender code can evolve 
+worked as a handicap (*i.e.* the cost of the badge is inversely 
+proportional to the quality of the individual), the sender code evolved often
 to produce an honest signal (Fig. \ref{fig:handicap}). The evolution of 
-the sender code does not happen immediately after the start of the 
+the sender code did not happen immediately after the start of the 
 evolutionary process. Instead, the evolutionary dynamics of the reaction 
-norm parameters (Fig. \ref{fig:handicap}, A) appear to involve a set of 
+norm parameters (Fig. \ref{fig:handicap}, A) appeared to involve a set of 
 steps. First, the intercept of the reaction norm ($\alpha$) 
-describing the relationship between individual quality and badge size evolves to 
+describing the relationship between individual quality and badge size evolved to 
 higher values, reducing the average badge size in the populations. 
 This is because reaction norms segregating in the population are 
-flat at the beginning of the simulation, consequently, 
-the receiver code does not respond to the badge size, 
+flat at the beginning of the simulation; consequently, 
+the receiver code does not respond to the badge size 
 and larger badges are costlier and do not trigger lower attack 
@@
 with higher quality (Fig. \ref{fig:handicap} B). Receivers then learn to react to such 
-correlation, increasing the probability of retreat when facing 
+correlation, increasing the probability of retreating when facing 
 individuals with larger badges (Fig. \ref{fig:handicap} C). Hence, natural selection 
@@
 Thus, badges are only costly and do not provide information about the 
-quality of individuals (Fig. \ref{fig:noBadge}) . Eventually, the badge 
-disappears from the population. In contrast, when the badge does not 
+quality of individuals (Fig. \ref{fig:noBadge}) . Eventually, therefore, badges 
+disappear from the population. In contrast, when the badge does not 
 work as a handicap but instead is cost-free, the evolutionary process 
@@
 Instead, subject to genetic drift, the 
-population either evolves towards the disappearance of the badge, or to the 
-maximum badge size. In either of this cases reaction norms are flat, so the 
+population either evolves towards the disappearance of the badge or to the 
+maximum badge size. In either of these cases reaction norms are flat, so the 
 badge does not provide any information about quality.
@@
 individuals learned with high speed of learning ($\mathrm{A} = 0.4$) and interacted 
-repeatedly along their lifetime (on average 2000 interactions). When we reduced 
-the number of interactions that individuals have over their lifetime to 300 
+repeatedly along their lifetime (2000 interactions on average). When we reduced 
+the number of interactions that individuals experienced over their lifetime to 300 
 on average, we saw a drastic increase in the phenotypic and genetic
-variation present in the population. Fig. \ref{fig:branching} shows 
-changes in the distribution of the sender reaction norm parameters
-along evolutionary time, assuming the badge is cost-free. 
-In panels A1 and 2, darker areas show trait values with high frequency in the population. 
-Populations start monomorphic with a value of zero on both the intercept and
+variation present in the population. Populations start monomorphic 
+with a value of zero on both the intercept and
 the slope, and mutations quickly build up a normal distribution around the starting 
@@
 ($\alpha$) in the population shows three peaks. In the case of $\beta$, the 
-peaks in the distribution are not so clear-cut, but it is clear that there is an 
-increase in the variance of the distribution. These changes in the 
+peaks in the distribution are not so clear-cut, but there is an 
+increase in the variance of the distribution over time. These changes in the 
 distribution of the parameters of the sender reaction norm 
@@
 (Fig. \ref{fig:branching} B.3-4). 
-In this particular simulation, the values of the slope in the informative group
+In this simulation, the values of the slope in the informative group
 are negative, which implies that badge size correlates negatively with quality. 
@@
 ```
-A limited number of interactions on the amount of variation has an effect on 
+
+A limited number of interactions has an effect on the amount of variation in
 the evolving parameters when the signal follows the handicap
@@
 of the reaction norm in the population. Thus, there are two types 
-of reaction norms. One of the types has a steeper slope, meaning that it expresses 
-a higher badge size for a given quality. This effect of an increased 
-variance in the trait distribution is not only triggered by lower 
-number of interactions. Larger variances are also found when we assume 
+of reaction norms. One of the types has a steeper slope compared to the other, 
+meaning that it expresses a higher badge size for a given quality. 
+This effect of an increased variance in the trait distribution
+is not only triggered by lower number of interactions. 
+Larger variances are also found when we assume 
 a lower speed of learning (data not shown). This suggest
@@
 learning allows the coexistence of different communication strategies within 
-a population [@botero_Evolution_2010].
+a population, something previously shown by @botero_Evolution_2010.
 
@@
 
-The behaviour expressed by naive individuals (those who have not learned) 
-imposes negative-frequency dependent selection, allowing for the build up of
-genetic variation in reaction norms. In the simulations presented so far, we 
-assumed that individuals start with a flat behavioural reaction norm such that 
-they escalate aggressively a fight with a $0.5$ probability regardless of 
-the badge size of the interacting partner. To asses whether such initial 
+Our model revealed that the behaviour expressed by naive individuals 
+(*i.e.* those who have not learned) imposes negative-frequency dependent selection, 
+allowing for the build up of genetic variation in reaction norms. 
+In the simulations presented so far, we assumed that individuals start with a flat behavioural reaction norm such that 
+they escalate fights aggressively with a $0.5$ probability regardless of 
+the badge size of the interacting partner. To assess whether such initial 
 conditions of the communication system had any role 
@@
 2) high (aggressive) probability of escalating a fight or 3) a probability 
-corresponding to the ESS of the original hawk-dove game (clever). 
+corresponding to the ESS of the original hawk-dove game 
+(which we call clever, due to information and computation necessary to know the ESS). 
 Fig. \ref{fig:startCond} shows the distribution of values of the 
@@
 the variation occurs within clusters. In contrast, when naive individuals behave
-either aggressively or cleverly, the badges evolve toward minimum and maximum 
+either aggressively or cleverly, badges evolve toward minimum and maximum 
 values, and so the variation is driven mainly by differences among replicates. 
@@
 naive behaviour quickly in ranges of the badge size that are common in 
-the population. Thus, one way to repeatedly interact with the naive behaviour
-is to have a rare badge size. When the naive behaviour is peaceful, individuals
+the population. In contrast, an individual expressing a rare badge size 
+will most likely experience the naive behaviour of their partner. 
+When the naive behaviour is peaceful, individuals
 with a rare badge size have a fitness advantage. This triggers 
 negative-frequency dependent selection and the evolution of different types of 
-badge sizes. In situations where individuals learn fast and interact repeatedly 
-the strength of frequency dependent selection diminishes.
-
-```{r startCond, echo=FALSE, out.width='100%',fig.align='center',fig.show='hold',fig.cap="The peaceful, the clever and the aggressive.
+badge sizes. According to this narrative, situations where individuals 
+learn fast and interact repeatedly will diminish the strength 
+of frequency dependent selection.
+
+```{r startCond, echo=FALSE, out.width='100%',fig.align='center',fig.show='hold',fig.cap="The peaceful, the aggressive and the clever.
 Distribution of values of the intercept $\\alpha$ and slope $\\beta$ for individuals at the end of the evolutionary simulations.
@@
 
-The action of natural selection and genetic drift do not override the conditions 
-for the evolutionary branching described earlier. Given how crucial naive 
-behaviour is for the diversification process described earlier, we tested whether 
-evolutionary processes could lead key parameters of the learning process toward
+The action of natural selection and genetic drift on the learning parameters 
+do not override the conditions for the evolutionary branching 
+[@dieckmann_Dynamical_1996] described earlier. 
+Given how crucial naive behaviour is for the diversification process 
+described earlier, we tested whether evolutionary processes could lead 
+key parameters of the learning process toward
 values that prevented the genetic variation to build up. 
@@
 In contrast, the initial behavioral tendency maintain values around zero. 
-Thus, evolutionary processes acting on the learning parameteres 
+Thus, evolutionary processes acting on the learning parameters 
 did not override the conditions necessary for frequency dependent to boost 
@@
 
-Here we presented a model that integrates the basic mechanisms of 
-learning with the fundamental rules of evolutionary processes. 
-In the model the response of a signal in a 
+We presented a model that integrates learning with 
+evolutionary processes. In the model the response of a signal in a 
 communication system is mediated by learning processes. Individuals 
@@
 trait of their interactive partners. 
-The learning process can mediate the evolution 
-of an honest signal under the handicap 
-principle [@grafen_Biological_1990;@zahavi_Mate_1975]. This conclusion is not 
-different from the classical genetic models, where responses to the signal are
-inherent to the individual. However, unlike classical genetic models,
-our model shows that learning can also mediate a signal polymorphism in the 
-absence of the handicap principle. Under this polymorphic equilibrium different 
-types of individuals, with honest and dishonest signals, compose the population. 
+We show that the learning process can mediate the evolution 
+of an honest signal under the handicap principle
+[@grafen_Biological_1990;@zahavi_Mate_1975]. This conclusion is not 
+different from classical genetic models, where responses to signals are
+innate to the individual. However, unlike classical genetic models,
+our model shows that learning can also mediate the origin of a signal 
+polymorphism in the absence of the handicap principle. 
+Under this polymorphic equilibrium the population consist of distinct 
+types of individuals, with honest and dishonest signals. 
 The amount of information gathered through learning, as a well as the 
 initial conditions of the learning process, are crucial for the emergence
-and maintenance of this variation. The emergence of the polymorphism requires
+and maintenance of variation in the signalling system. 
+We also show that the emergence of the polymorphism requires
 a limit in the amount of information collected by individuals, as well as
-peaceful behaviour in naive individuals. Furthermore, we have shown that those 
-conditions can be reached when evolutionary processes drive changes in the learning
-system. Thus, the diversification process we present might play a role in natural
-populations.
+peaceful behaviour in naive individuals. Furthermore, we have shown that the 
+conditions leading to the origin of the polymorphism can be reached when evolutionary
+processes are able to change the learning system.
 
 Associative learning is a powerful mechanism to learn about the world, about 
-social partners and about one's abilities relative to the social context. 
-By associating cues and signals with fitness relevant outcomes individuals
+social partners and about one's abilities in a given social context. 
+By associating cues and signals with fitness-relevant outcomes, individuals
 collect information that allows them to improve their reproductive potential. 
-This is very obvious in classical examples such as when animals learn to avoid 
+This is quite obvious in classical examples such as when animals learn to avoid 
 food that makes them sick, or use environmental cues to find food. 
-More recent ideas have highlighted the potential role that associations 
-can have in a social contexts such as: hierarchy formation 
+More recent theoretical work has highlighted the potential role that associations 
+can have in social contexts such as hierarchy formation 
 [@leimar_Evolution_2021;@leimar_Reproductive_2022;@leimar_Effects_2022],
-and cooperation [@leimar_Learning_2019;@dridi_Learning_2018]. In this 
-theoretical examples individuals use various sources of reward to make
-adaptive choices in a given social context. Here we extend this logic to the 
+and cooperation [@leimar_Learning_2019;@dridi_Learning_2018]. In these 
+examples, models involve individuals who use various sources of reward to make
+adaptive choices. Here we extended this logic to the 
 evolution of a communication system mediated by a quality signal. 
@@
 and individual's own quality determined the behaviour [@botero_Evolution_2010]. 
-However, there isn't a clear mechanism to justify the assumption that 
+However, no clear mechanism justify the assumption that 
 individuals inherently know their own quality, particularly relative to their peers.
@@
 even in the absence of an honest signal, they are able to make more 
-adaptive choices. Thus the learning process we have modelled and the 
+adaptive choices. Thus, the learning process we have modelled and the 
 response it mediates has fitness relevant consequences even in the 
-absence of an honest signal. This is confirmed by simulation where 
+absence of an honest signal. This is confirmed by simulations where 
 the speed of learning is allowed to change subject to evolutionary processes. 
@@
 Learning processes collecting information on a population wide level promote 
-frequency dependent selection, in the communication system we model frequency 
-dependency translates to the coexistence of different communication strategies. 
-The frequency-dependence triggered by learning process was highlighted by the 
-classic study of [@bond_Visual_2002] where live predators drive the evolution
-of *in silico* polymorphic prey. The key to the evolution of this polymorphisms is 
-that predators learn to discriminate better preys that are frequently encountered.
-Thus, preys found in low frequency have a selective advantage. 
-An analogous process occurs in our simulations, the learning algorithm of the 
-receiver individual learns the best response for herself towards individuals 
-that have a badge size common in the population. When an individual has a
+frequency dependent selection. Frequency-dependence triggered by 
+learning process was highlighted by a 
+classic study where live predators drive the evolution
+of *in silico* polymorphic prey [@bond_Visual_2002]. 
+The key to the evolution of prey polymorphism is 
+that predators are better able to discriminate preys from the background, 
+when preys are frequently encountered.
+Thus, preys found in low frequency in a population have a selective advantage. 
+An analogous process emerged in our simulations, in which the 
+receiver individual learned the appropriate response towards individuals 
+with a common badge size in the population. When an individual has a
 rare badge size it can be favoured or unfavoured by the naive behaviour. When 
@@
 
-Originally @rohwer_Social_1975 proposed the idea that certain morphological 
+Originally @rohwer_Social_1975a proposed the idea that certain morphological 
 traits could be used by animals as status signals or signals of quality. 
@@
 this narrative by showing that the effect size of the association between bib 
-size and dominance rank is small and uncertain [@sanchez-tojar_Metaanalysis_2018]. 
+size and dominance rank is small and uncertain [@sanchez-tojar_Metaanalysis_2018a]. 
 The simple straight correlation between bib size and dominance rank

### 2023-09-06 manuscript/manuscript_1.0.Rmd
Addressing DAniel´s comments

@@
 output:
-  word_document: default
-  # word_document: default
-  # pdf_document:
-    latex_engine: xelatex # default  #
+  pdf_document:
+    latex_engine: xelatex  # default  ##word_document: default  # word_document:default
     pandoc_args:
@@
       - '--lua-filter=author-info-blocks.lua'
+  word_document: default
 fig_caption: yes
@@
   can maintain the honesty of signals, thus favoring their evolutionary 
-  stability. Here we present evolutionary simulations where individuals 
+  stability. We present evolutionary simulations where individuals 
   can produce a signal proportional to their quality and learn along their
@@
   replicate previous results where the handicap principle mediates the evolution 
-  of signals. As in traditional models without learning, in our simulations 
-  learning on the receiver side can mediate the evolution of signals
-  of quality on the sender side. When the cost of the signal is proportional to the
-  quality of the sender, populations are only composed of honest signalers. In contrast 
-  to traditional models which predict the absence of signals when the
+  of signals. In our simulations learning on the receiver side can mediate the evolution
+  of signals  of quality on the sender side. When the cost of the signal is proportional
+  to the  quality of the sender, all individuals in populations are honest signalers.   
+  In contrast to traditional models which predict the absence of signals when the
   cost is not proportional to the quality of the signaler, our model revealed that
-  learning facilitates the evolution of a polymorphism in which populations are
-  composed of both honest and dishonest signalers. We argue that learning can 
-  be a general cognitive mechanism having a role in a wide range of communication systems 
-  and more generally in behavioral responses.
+  learning facilitates the evolution of a polymorphism in which populations comprise both
+  honest and dishonest signalers. We argue that learning may have a role in the evolution
+  and dynamics of a wide range of communication systems and more generally in 
+  behavioral responses.
 ---
@@
 
-The outcome of interactions with conspecific individuals is an important 
+The outcome of interactions with conspecific individuals is a critical
 determinant of fitness in social animals. Irrespective of whether 
 such interactions are cooperative or competitive, the actions of 
-interacting individuals partly determine 
+interacting individuals influence 
 their reproductive success. However, the best action for an individual 
@@
 partners, which is sometimes referred to as Resource Holding Potential 
-[RHP, Parker -@parker_Assessment_1974], or simply as quality. 
+[RHP, Parker -@parker_Assessment_1974] or simply as quality. 
 Responsiveness to the quality of an opponent is central 
-to communication systems such as those in which individuals 
+to communication systems in which individuals 
 exhibit badges of status, where a signal conveys quality 
@@
 produced by high and low quality individuals alike. 
-Generally signals can be evolutionarily stable whenever low quality individuals 
+Generally signals can be evolutionarily stable whenever low-quality individuals 
 pay a higher fitness cost for carrying the signal such that it is not 
-in the interest of low quality individuals to fake quality by producing a large badge.
+in the interest of low-quality individuals to fake quality by producing a large badge.
 Signal costs inversely proportional to quality may come about as a result of the
@@
 wasps, subordinate individuals with experimentally manipulated dominant-like facial
-patterns received more aggression from the dominant. In this case, the costs are 
-socially mediated rather than developmentally and are triggered by the violation 
-of the convention. It is still incluer under this narrative, how the convenvention 
-is stablished in the first place.
+patterns received more aggression from the dominant[@tibbetts_Socially_2004]. 
+In this case, costs are socially rather than developmentally mediated and are 
+triggered by the violation of the convention, although how the convention
+is established in the first place is unclear.
 
@@
 
-<!-- Continuar aquí en la revisión -->
-
 Associative learning is a taxonomically widespread cognitive 
@@
 [@enquist_Power_2016; @quinones_Reinforcement_2020;@pretot_Factors_2016;
-pretot_Comparative_2021]. For example, species can have different learning rates
-[@hoedjes_Natural_2010],
-some may vary these rates during learning trials which allows them to have more 
-flexible learning [@leimar_Flexibility_2023a], and some may include future rewards in their associations [raby_Planning_2007;osvath_Spontaneous_2009].
+@pretot_Comparative_2021]. For example, species can have different learning rates
+[@hoedjes_Natural_2010], some may vary these rates during learning 
+trials which allows them to have more 
+flexible learning [@leimar_Flexibility_2023], and some may include future rewards in their associations [@raby_Planning_2007;@osvath_Spontaneous_2009].
 Associative learning is not a single mechanism, but rather a set of 
 cognitive structures and processes that can vary in their scope and complexity. Presumably, these structures and processes have been modified 
-by natural selection and could provide new explanations for behavioural 
+by natural selection and could provide novel explanations for behavioural 
 variation. Nonetheless, associative learning is 
@@
 behavioural patterns [@fawcett_Exposing_2013; @kamil_Optimal_1983; 
-@mcauliffe_Psychology_2015]. Computational models of evolution 
-can play an important role to overcome the lack of integration 
-between learning and evolution. Reinforcement learning theory 
-encompasses a series of computational methods inspired 
+@mcauliffe_Psychology_2015].
+
+Computational models of evolution can overcome the lack of integration 
+between learning and evolution. Reinforcement 
+learning theory encompasses a series of computational methods inspired 
 on the psychological and neurological mechanisms of associative 
@@
 Furthermore, these algorithms can be embedded in evolutionary simulations to 
-generate theoretical predictions of the effect that learning can 
-have in behavioural evolution [@leimar_Learning_2019;@leimar_Effects_2022].
+generate theoretical predictions of the effect of learning on behavioural evolution [@leimar_Learning_2019;@leimar_Effects_2022].
 
@@
 partner can depend on the size of the partner's badge, 
-the dependency is determined by the focal's experiences acquired
+with the dependency being determined by the focal's experiences acquired
 through a learning process. We implement learning using 
@@
 leads to an increase in the reward estimation. 
-So, if a focal individual decides to escalate a fight against 
+Thus, if a focal individual decides to escalate a fight against 
 an individual with a small badge and this leads to an increase in 
@@
 For the sake of simplicity we keep the location of these feature centres 
-constant and  the same for all individuals, the are stored in the 
+constant and  the same for all individuals; they are stored in 
 vector $\symbf{b}$ . Each of these feature centres 
@@
 ```{r model-struc,echo=FALSE,warning=FALSE,fig.width=8,fig.align='center',fig.height=7,fig.show='hold',fig.cap= "Model of communication in the context of aggressive interactions.
-In A, the reaction norm determines the bib (badge) size, black band on birds chest.
+In A, the reaction norm determines the badge size, *e.g* the size of a black band on a bird's chest.
 Red shows an informative reaction norm, whereas blue shows a uninformative reaction norm.
@@
 The reaction norm arises from generalizing the information from the feature weights (black dots).
-Feature weaights are values associated with a particular badge size and are updated as individuals interact with each other.
+Feature weights are values associated with a particular badge size and are updated as individuals interact with each other.
 The generalization of the information associated with each feature weight is represented by the grey line and its axis, which shows that the response triggered by the fourth feature diminishes as the value evaluated is further from the feature center.
 The learning process moves the feature weights (black dots) up or down depending on whether this leads to an increase in the estimated reward.
-In C and D, effect of learning on the receiver strategy.
+Panels C and D show the effect of learning on the receiver strategy.
 Receivers in C face signallers with uninformative reaction norms (blue line in A).
@@
 These simulations show what type of receiver strategy develops through 
-a learning process in such monomorphic population. 
+a learning process in such a monomorphic population. 
 Figure \ref{fig:model-struc} (C and D) shows 
 the receiver strategy developed through learning; panel C is for 
-receivers that faced uninformative signals, and D informative ones. 
+receivers that faced uninformative signals, and D for those facing informative ones. 
 The simulations reveal that when learners face uninformative signals (Fig. \ref{fig:model-struc} C), they modify their probability of retreating 
@@
 the two classic pure strategies of hawks and doves. 
-Given that we have a assumed a monomorphic population 
+Given that we assumed a monomorphic population 
 with unresponsive reaction norms on the signalling side, the changes 
 triggered by learning only affect a small range of badge sizes. Specifically, 
-all learning has occurred around a badge size of $0.5$, because all
+all learning occurred around a badge size of $0.5$, because all
 badges in the population are of this size (panel C). 
@@
 When we allowed the badge size to evolve ($\alpha$ and $\beta$ 
-change subject to natural selection and genetic drift), and the signal
-worked as a handicap (*i.e.* the cost of the badge is inversely 
+changed subject to natural selection and genetic drift) and the signal
+worked as a handicap (*i.e.* the cost of the badge was inversely 
 proportional to the quality of the individual), the sender code evolved often
@@
 The amount of information that agents are able to collect through learning 
-throughout their life time can strongly change the outcome of 
+over their life time can strongly change the outcome of 
 evolutionary dynamics. In the simulations presented in figure 
 \ref{fig:handicap} and \ref{fig:noBadge}, 
-individuals learned with high speed of learning ($\mathrm{A} = 0.4$) and interacted 
+individuals learned with high speed ($\mathrm{A} = 0.4$) and interacted 
 repeatedly along their lifetime (2000 interactions on average). When we reduced 
@@
 variation present in the population. Populations start monomorphic 
-with a value of zero on both the intercept and
-the slope, and mutations quickly build up a normal distribution around the starting 
+with a value of zero for both the intercept and
+the slope of the reaction norm, and mutations quickly build up a 
+normal distribution around the starting 
 value. Within the first 2000 generations, the unimodal distribution in 
@@
 so at the end of the evolutionary simulation the distribution of the intercept 
-($\alpha$) in the population shows three peaks. In the case of $\beta$, the 
-peaks in the distribution are not so clear-cut, but there is an 
-increase in the variance of the distribution over time. These changes in the 
+($\alpha$) in the population shows three distinct peaks. In the case of $\beta$, the 
+peaks in the distribution are not so clear-cut, but the variance of the 
+distribution increases over time. These changes in the 
 distribution of the parameters of the sender reaction norm 
@@
 is not informative of their quality, whereas the third type shows 
-intermediate badge sizes determined by the quality of the individual 
+intermediate badge sizes determined by individual quality 
 (Fig. \ref{fig:branching} B.3-4). 
@@
 have on average 300 interactions in their life and the cost of the signal
-is proportional to the quality, following the handicap principle. 
+is proportional to quality, following the handicap principle. 
 The evolutionary process leads to a combination of reaction norm parameters where 
@@
 of the reaction norm in the population. Thus, there are two types 
-of reaction norms. One of the types has a steeper slope compared to the other, 
-meaning that it expresses a higher badge size for a given quality. 
+of reaction norms. One of the types has a steeper slope compared to the other. 
 This effect of an increased variance in the trait distribution
@@
 Our model revealed that the behaviour expressed by naive individuals 
-(*i.e.* those who have not learned) imposes negative-frequency dependent selection, 
+(*i.e.* those who have not yet learned) imposes negative-frequency dependent selection, 
 allowing for the build up of genetic variation in reaction norms. 
@@
 Distribution of values of the intercept $\\alpha$ and slope $\\beta$ for individuals at the end of the evolutionary simulations.
-The panels show the three different initial conditions for the behaviour of individuals: peaceful, \"clever\" and aggressive, see maintext for details.
+The panels show the three different initial conditions for the behaviour of individuals: peaceful, agressive and  \"clever\", see maintext for details.
 Colours indicate the replicate simulation.
@@
 [@dieckmann_Dynamical_1996] described earlier. 
-Given how crucial naive behaviour is for the diversification process 
+Given how crucial naive behaviour is for the branching event 
 described earlier, we tested whether evolutionary processes could lead 
@@
 the learning module, namely the speed of learning ($A$) and the behavioural 
-tendency expressed by individuals before learning. Top panels in 
+tendency expressed by individuals prior to learning. Top panels in 
 figure \ref{fig:learnEvol} show that there is diversification process in this 
@@
 panels (B) of the same figure, the learning parameters stay within the range 
-necessary for the diversification process. The speed of learning 
+necessary for the branching event to take place. The speed of learning 
 (Fig. \ref{fig:learnEvol} B.1) stays well above zero throughout the simulation. 
@@
 Thus, evolutionary processes acting on the learning parameters 
-did not override the conditions necessary for frequency dependent to boost 
+did not override the conditions necessary for frequency-dependent selection to boost 
 genetic variation.
@@
 The amount of information gathered through learning, as a well as the 
-initial conditions of the learning process, are crucial for the emergence
+initial conditions of the learning process, were crucial for the emergence
 and maintenance of variation in the signalling system. 
-We also show that the emergence of the polymorphism requires
+Namely the polymorphism required
 a limit in the amount of information collected by individuals, as well as
@@
 conditions leading to the origin of the polymorphism can be reached when evolutionary
-processes are able to change the learning system.
+processes are able to change the parameters of the learning system (the speed of
+learning and the naive behaviour).
 
 Associative learning is a powerful mechanism to learn about the world, about 
-social partners and about one's abilities in a given social context. 
+social partners, and about one's abilities in a particular social context. 
 By associating cues and signals with fitness-relevant outcomes, individuals
-collect information that allows them to improve their reproductive potential. 
-This is quite obvious in classical examples such as when animals learn to avoid 
-food that makes them sick, or use environmental cues to find food. 
+collect information allowing them to improve their reproductive potential. 
+This is evident in classical situations such as when animals learn to avoid 
+food that makes them sick or use environmental cues to find food. 
 More recent theoretical work has highlighted the potential role that associations 
 can have in social contexts such as hierarchy formation 
-[@leimar_Evolution_2021;@leimar_Reproductive_2022;@leimar_Effects_2022],
+[@leimar_Evolution_2021;@leimar_Reproductive_2022;@leimar_Effects_2022]
 and cooperation [@leimar_Learning_2019;@dridi_Learning_2018]. In these 
@@
 adaptive choices. Here we extended this logic to the 
-evolution of a communication system mediated by a quality signal. 
+evolution of a communication system mediated by a signal of quality. 
 Previous evolutionary models of badges of status assumed that individuals 
 responded using a behavioural reaction norm, where the opponent's badge 
-and individual's own quality determined the behaviour [@botero_Evolution_2010]. 
-However, no clear mechanism justify the assumption that 
+and the individual's own quality determined behaviour [@botero_Evolution_2010]. 
+However, there is no clear mechanism justifying the assumption that 
 individuals inherently know their own quality, particularly relative to their peers.
 In our model, individuals not only learn about the quality signal, but also 
-learn about their own quality relative to others. This can be seen throughout
+learn about their own quality relative to others. This can be seen in
 our simulations in the responses developed by individuals of different quality. 
@@
 even in the absence of an honest signal, they are able to make more 
-adaptive choices. Thus, the learning process we have modelled and the 
+adaptive choices. Specifically, lower quality individuals refrained more from
+escalating fights, while higher quality took over resources they 
+could win in a fight. Thus, the learning process we modelled and the 
 response it mediates has fitness relevant consequences even in the 
-absence of an honest signal. This is confirmed by simulations where 
-the speed of learning is allowed to change subject to evolutionary processes. 
+absence of an honest signal. This was confirmed by simulations where 
+the speed of learning was allowed to change subject to evolutionary processes. 
 In those simulations, natural selection always
@@
 only captured the effect of selection mediated by the social 
-game on the learning parameters. Inter-specific variation in cognitive 
-abilities mediated by environmental differences has been reported elsewhere
+game on the learning parameters. For example, We did not account for 
+potential costs of faster updating, or improved performance in foraginf tasks. 
+Inter-specific variation in cognitive abilities mediated by environmental 
+differences has been reported elsewhere
 [@sonnenberg_Natural_2019]. If such variation is partly mediated by changes 
-that affect general cognitive processes, they could impact the outcome 
-of communication systems like the one we modelled.
+that affect general cognitive processes, like the speed of learning,
+they could impact the outcome of communication systems like the one we modelled.
 
@@
 The key to the evolution of prey polymorphism is 
-that predators are better able to discriminate preys from the background, 
+that predators are better able to discriminate prey from the background, 
 when preys are frequently encountered.
-Thus, preys found in low frequency in a population have a selective advantage. 
+Thus, prey morphs found in low frequency in a population have a selective advantage. 
 An analogous process emerged in our simulations, in which the 
@@
 rare badge size it can be favoured or unfavoured by the naive behaviour. When 
-the naive behaviour is peaceful, polymorphism is promoted, when the naive 
-behaviour is aggressive polymorphism is prevented. These two examples show how
-behaviours dependent on learning processes seem to generally trigger frequency 
-dependent selection. That is because the learning algorithm collects more 
-information on the frequent values of the the trait distribution. For the sake 
-of simplicity we modelled a single dimension (badge size), however this effect 
-could potentially be more important when individuals learn from 
-multidimensional signals.
-
-Originally @rohwer_Social_1975a proposed the idea that certain morphological 
+the naive behaviour is peaceful, polymorphism is promoted, whereas when the naive 
+behaviour is aggressive polymorphism is prevented. These two situations show how
+behaviours dependent on learning processes seem to generally trigger 
+frequency-dependent selection. That is because the learning algorithm 
+collects more information on the more frequent values of the the trait distribution. 
+For the sake of simplicity we modelled a single dimension (badge size), 
+however the frequency-dependen effect of learning could potentially be 
+more important when individuals learn from multidimensional signals.
+
+Originally, @rohwer_Social_1975a proposed the idea that certain phenotypic 
 traits could be used by animals as status signals or signals of quality. 
-This idea has been tested repeatedly on the dark patches of some bird species 
-(bibs). So, are bibs real signals of quality? if so are they handicaps or 
+This idea has been tested repeatedly on the dark patches of some bird species, 
+such as the bibs exhibited by species in various families, particularly sparrows.
+So, are bibs real signals of quality? if so are they handicaps or 
 badge of status? From an evolutionary perspective it is unclear how the stability of 
@@
 proportional to quality it may work as a convention.
-In this later case, honesty is supposedly 
-maintained by the aggressive reaction of receivers when the convention is 
+In this later case, honesty is presumably maintained by the aggressive 
+reaction of receivers when the convention is 
 broken [@enquist_Signaling_2010;@tibbetts_Socially_2004]. Our model, portraits 
 both kinds of signals, although with some nuances. When we assume the handicap 
-principle, we get the evolution of honest signals mediated by the learned 
-response. When we assume cost-free signal, under certain conditions the learned 
-response mediated the emergence of phenotypic variation. This variation, 
-particularly in the mid-range of the distribution facilitates the establishment 
-of a convention, in this range individuals respond appropriately to the 
-trait of their peer. The house sparrow (*Passer domesticus*) 
-has been a text-book example of badges of status. According to the prevailing narrative the bib size in males is a signal of the dominance rank and quality.
+principle, simulations result in the evolution of honest signals mediated by the
+learned response. When we assume cost-free signal, under certain conditions the
+learned response mediated the emergence of phenotypic variation in the signal. 
+This variation, particularly in the mid-range of the distribution facilitates 
+the establishment of a convention, in this range individuals respond 
+appropriately to the trait of their peer. The house sparrow (*Passer domesticus*) 
+has been a text-book example of badges of status. According to the prevailing narrative the bib size in males is a signal of dominance rank and quality.
 However, a recent meta-analysis called into question 
@@
 size and dominance rank is small and uncertain [@sanchez-tojar_Metaanalysis_2018a]. 
-The simple straight correlation between bib size and dominance rank
+A simple correlation between bib size and dominance rank
 (and quality) is expected under the handicap principle, as it is shown here and 
 elsewhere [@botero_Evolution_2010;@grafen_Biological_1990;@johnstone_Badges_1993].
-But, the diversification process presented here shows that conventions can 
-emerge in a less straightforward way when mediated by learning processes. This 
+However, the branching event presented here shows that conventions can 
+emerge in a less straightforward way when mediated by learning processes. These 
 nuances could potentially make sense of seemingly contradictory evidence on the 
-correlation between plumage traits and quality. For example, conventions in 
-our simulations arose only in an intermediate range of the signalling trait. 
-While correlations are often performed across the whole range of variation.
+correlation between plumage traits and quality. For example, conventions in our
+simulations arose only in an intermediate range of the signalling trait, yet 
+empirical studies seeing for correlations between phenotypic traits and quality 
+are often performed across the whole range of variation.
 

### 2023-09-07 manuscript/manuscript_1.0.Rmd
New version of the manuscript

@@
     \floatplacement{figure}{H}
-    \newcommand{\beginsupplement}{ \setcounter{table}{0}         \renewcommand{\thetable}{S\arabic{table}}\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
+    \newcommand{\beginsupplement}{ \setcounter{table}{0}            
+    \renewcommand{\thetable}{S\arabic{table}}\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
   ```
@@
 exhibit badges of status, where a signal conveys quality 
-[@johnstone_Badges_1993; @rohwer_Social_1975a]. 
-Such signals must be arbitrary, meaning the signals could potentially be 
-produced by high and low quality individuals alike. 
-Generally signals can be evolutionarily stable whenever low-quality individuals 
+[@johnstone_Badges_1993; @rohwer_Social_1975]. 
+Such signals tend to be arbitrary, meaning the signals could potentially be 
+produced by high- and low-quality individuals alike. This symmetry opens up the 
+possibility of dishonest signalling and the question of evolutionary stability. 
+Signals can be evolutionarily stable whenever low-quality individuals 
 pay a higher fitness cost for carrying the signal such that it is not 
@@
 (Fig. \ref{fig:learnEvol} B.1) stays well above zero throughout the simulation. 
-In contrast, the initial behavioral tendency maintain values around zero. 
+In contrast, the initial behavioural tendency maintain values around zero. 
 Thus, evolutionary processes acting on the learning parameters 
 did not override the conditions necessary for frequency-dependent selection to boost 
-genetic variation.
+genetic and phenotypic variation.
 
@@
 
-Originally, @rohwer_Social_1975a proposed the idea that certain phenotypic 
+Originally, @rohwer_Social_1975 proposed the idea that certain phenotypic 
 traits could be used by animals as status signals or signals of quality. 
@@
 reaction of receivers when the convention is 
-broken [@enquist_Signaling_2010;@tibbetts_Socially_2004]. Our model, portraits 
+broken [@enquist_Signaling_2010;@tibbetts_Socially_2004]. Our model, captures 
 both kinds of signals, although with some nuances. When we assume the handicap 
@@
 this narrative by showing that the effect size of the association between bib 
-size and dominance rank is small and uncertain [@sanchez-tojar_Metaanalysis_2018a]. 
+size and dominance rank is small and uncertain [@sanchez-tojar_Metaanalysis_2018]. 
 A simple correlation between bib size and dominance rank
@@
 and particularly badges of status, is often explaining the presence and 
-absence of certain morphological traits within populations. However, the 
-morphological traits hypothesized to play a role in communication systems 
-regularly present interesting patters at the phylogeographic
-level. One example of this is the leapfrog pattern [@remsen_High_1984],
-whereas a certain morphological trait alternates its presence and 
-absence in a set of geographically adjunct populations. Patterns of 
-molecular variation along geographic clines suggest that natural selection 
-is involved in the emergence of these patterns 
-[@cadena_Testing_2011;@cadena_Evolutionary_2007]. However, it is unclear what 
+absence of certain morphological traits within populations. Phenotypic traits
+hypothesized to play a role in communication systems regularly exhibit 
+patterns of geographic variation within species that have long called 
+for explanations. One example of this is the leapfrog pattern [@remsen_High_1984],
+whereby a certain morphological trait alternates its presence and 
+absence in a set of geographically adjunct populations, likely as 
+a consequence of selection [@cadena_Testing_2011;@marquez_Divergence_2020]. 
+However, it is unclear what 
 type of ecological process is behind these selective regimes. 
-Given the wide variation in outcomes found in our model, that depend on both 
-stochasticity and cognitive parameters, we think variation in cognition and 
-learning in particular could provide some explanatory power in this respect. 
-One first step, for example is to evaluate the way individuals in different 
-populations respond to novel traits [@avendano_Territorial_2021].
+Given the wide variation in outcomes found in our model, which depend on both
+stochasticity and cognitive parameters which may vary in space owing to various
+processes, we speculate that variation in cognition and learning could 
+provide some explanatory power in this respect. A first step to assess this 
+hypothesis would be to evaluate the way individuals in different populations 
+respond to novel traits [*e.g.* @avendano_Territorial_2021].
 

### 2023-09-14 manuscript/Title_page.Rmd
Title_page rmd file and readme md file, getting readu for realease and submission

---
title: |
   \large \textbf{Title:} The role of learning in the evolution of status signalling: a modeling approach
subtitle: |
  \normalsize \textbf{Abbreviated title:} Learning and evolution in status signalling
author: 
  - name: Andrés E. Quiñones 
    email: andresqp@gmail.com
    affiliation: 
      - a
      - b
    corresponding: andresqp@gmail.com 
    fulladress: Rue Emile-Argand 11, CH-2000, Neuchatel, Switzerland, tel. +41325586486
  - name: Redouan Bshary
    affiliation: a
  - name: C. Daniel Cadena
    affiliation: 
      - b
address:
  - code: a
    address: Institute of Biology, University of Neuchatel, Neuchatel, Switzerland
  - code: b
    address: Laboratorio de Biología Evolutiva de Vertebrados, Departamento  de Ciencias Biológicas, Universidad de los Andes, Bogotá, Colombia
output:    
  pdf_document:
    keep_tex: true
    latex_engine: xelatex
    template: template.latex
---

## Funding
This work was supported by the posdoctoral assistant funding from the Depatamento de Ciencias Biológicas from the Universidad de los Andes granted to A.Q. and by the by the Swiss National Science Foundation grants (grant number: 310030_192673) to R.B.

## Acknowledgements
AQ and CDC kindly thank the members of Evolvert, in particular
Jorge Avendaño for pictures of the *Arremon* and discussion on
signal evolution. AQ thanks C. Botero for fruitful discussions, and thecluster of the Institute of Computer Science of the University of Neuchâtel (IIUN) for the computing resources provided.

### 2023-09-14 manuscript/Title_page.tex
Title_page rmd file and readme md file, getting readu for realease and submission

--- 2022-07-20 manuscript/manuscript_1.0.tex
@@
 \documentclass[
-  12pt,
 ]{article}
@@
 \usepackage{xcolor}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\hypersetup{
-  pdfauthor={Andrés E. Quiñones; C. Daniel Cadena*; Redouan Bshary},
-  pdfkeywords={My keywords},
-  hidelinks,
-  pdfcreator={LaTeX via pandoc}}
-\urlstyle{same} % disable monospaced font for URLs
-\usepackage[margin=1in]{geometry}
-\usepackage{graphicx}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-% Set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
 \setlength{\emergencystretch}{3em} % prevent overfull lines
@@
 \setcounter{secnumdepth}{-\maxdimen} % remove section numbering
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newlength{\cslentryspacingunit} % times entry-spacing
-\setlength{\cslentryspacingunit}{\parskip}
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1
-  \let\oldpar\par
-  \def\par{\hangindent=\cslhangindent\oldpar}
-  \fi
-  % set entry spacing
-  \setlength{\parskip}{#2\cslentryspacingunit}
- }%
- {}
-\usepackage{calc}
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
-  \usepackage{float}
-  \usepackage{setspace}\doublespacing
-  \usepackage[backref=page]{hyperref}
-  \linespread{2}
-  \usepackage{lineno}
-  \linenumbers
-  \floatplacement{figure}{H}
- \newcommand{\beginsupplement}{ \setcounter{table}{0}     \renewcommand{\thetable}{S\arabic{table}}\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
-
 \ifLuaTeX
@@
 \fi
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same} % disable monospaced font for URLs
+\hypersetup{
+  pdftitle={ The role of learning in the evolution of status signalling: a modeling approach},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
 
-\title{Learning can mediate both badges of status and individual
-recognition}
-\author{Andrés E. Quiñones\footnote{Laboratorio de Biología Evolutiva de
-  Vertebrados, Departamento de Ciencias Biológicas, Universidad de los
-  Andes, Bogotá, Colombia} \and C. Daniel Cadena* \and Redouan
-Bshary\footnote{Institute of Biology, University of Neuchâtel,
-  Neuchâtel, Switzerland}}
-\date{July 18, 2022}
+\title{\large \textbf{Title:} The role of learning in the evolution of
+status signalling: a modeling approach}
+\usepackage{etoolbox}
+\makeatletter
+\providecommand{\subtitle}[1]{% add subtitle to \maketitle
+  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
+}
+\makeatother
+\subtitle{\normalsize \textbf{Abbreviated title:} Learning and evolution
+in status signalling}
+\date{}
+
+\usepackage{forarray}
+\usepackage{xstring}
+\newcommand{\getIndex}[2]{
+  \ForEach{,}{\IfEq{#1}{\thislevelitem}{\number\thislevelcount\ExitForEach}{}}{#2}
+}
+
+\newcommand{\getAff}[1]{
+  \getIndex{#1}{a,b}
+}
 
@@
 \maketitle
-\begin{abstract}
-Adaptive behavioural responses often depend on qualities of the
-interacting partner of an individual. For example, when competing for
-resources, an individual might be better off escalating fights with
-individuals of lower quality, while restraining from fighting
-individuals of higher quality. Communication systems involving signals
-of quality allow individuals to reduce uncertainty regarding the
-fighting ability of their partners and make more adaptive behavioral
-decisions. However, dishonest individuals can destabilize such
-communications systems. One open question is whether cognitive
-mechanisms, such as learning, can maintain the honesty of signals, thus,
-favoring their evolutionary stability. Here we present evolutionary
-simulations, where individuals can produce a signal proportional to
-their quality; and learn along their lifetime the best response to the
-signal emitted by their peers. In these simulations, learning on the
-receiver side can mediate the evolution of signals of quality on the
-sender side. When the cost of the signal is proportional to the quality
-of the sender populations are only composed of honest signalers. When
-the cost is not proportional to the quality of the signaler, the
-population is composed of both honest and dishonest signalers. We argue
-that learning can be a general cognitive mechanism playing a role in a
-wide range of communication systems.
-\end{abstract}
+\begin{flushleft}
 
-\hypertarget{introduction}{%
-\subsection{Introduction}\label{introduction}}
+Andrés E.
+Quiñones\textsuperscript{\getAff{a}, \getAff{b}}\textsuperscript{*},
+Redouan Bshary\textsuperscript{\getAff{a}},
+C. Daniel Cadena\textsuperscript{\getAff{b}}\\
+\bigskip
+\textbf{\getAff{a}}Institute of Biology, University of Neuchatel,
+Neuchatel, Switzerland\\
+\textbf{\getAff{b}}Laboratorio de Biología Evolutiva de Vertebrados,
+Departamento de Ciencias Biológicas, Universidad de los Andes, Bogotá,
+Colombia\\
+\bigskip
+* Corresponding author: \href{mailto:andresqp@gmail.com}{\nolinkurl{andresqp@gmail.com}}, Rue
+Emile-Argand 11, CH-2000, Neuchatel, Switzerland, tel. +41325586486\\
+\end{flushleft}
 
-The outcome of interactions with con-specifics is an important
-determinant of fitness in social animals. Irrespective of whether such
-interactions are cooperative or competitive, the actions of interacting
-individuals partly determine their reproductive success. However, the
-best action for an individual might vary depending on its condition, and
-that of individuals with whom it interacts. Thus, acting based on
-information about interacting partners is typically adaptive (Quiñones
-et al. 2016). Acquiring this information, however, is far from trivial.
-In some cases, interacting partners (signallers) might be `willing' to
-provide accurate information, but in others it might be in their own
-interest to conceal information (Johnstone 1997), or to provide
-erroneous information (Johnstone and Norris 1993). Typically, in a given
-context, some proportion of individuals will benefit from broadcasting
-accurate information, whereas others will benefit from hiding it. Take,
-for instance, an interaction between two individuals where one can help
-the other. Given some costs and benefits, a potential donor will be
-interested in helping related individuals. Therefore, for a relative of
-the donor, broadcasting kinship is advantageous, whereas for an
-unrelated individual, concealing the lack of kinship is better. Similar
-scenarios may apply to many aspects of social life such as finding
-mates, feeding offspring, dominance relationships and aggressive
-interactions (Bradbury and Vehrencamp 2011; Møller 1988; Tibbetts and
-Dale 2004). In dominance and aggressive interactions, an important piece
-of information to guide individual actions is interacting partners
-fighting ability, which is sometimes referred to as Resource Holding
-Potential (RHP) (Parker 1974), or simply as quality. Responsiveness to
-the quality of the partner is central to communication systems such as
-badges of status, where an arbitrary signal conveys quality (Johnstone
-and Norris 1993; Rohwer 1975). Here, arbitrary refers to a signal which
-is not ontogenetically correlated with quality. Badges of status can be
-evolutionarily stable whenever the signal imposes a fitness cost which
-decreases with the quality of an individual (Botero et al. 2010;
-Johnstone and Norris 1993), such that it's no longer in the interest of
-low quality individuals to fake quality by producing a large badge. The
-cost of honest signals is a more general principle of communications
-systems, usually referred to as the handicap principle (Grafen 1990;
-Zahavi 1975).
+\hypertarget{funding}{%
+\subsection{Funding}\label{funding}}
 
-A somewhat ignored component of communication systems, in the context of
-aggressive interactions, is the cognitive aspects of the receiver
-module. Theoretical models often assume that Communication systems
-relying on badges of status to have reaction norms as their mechanistic
-underpinning (Botero et al. 2010), with individuals using their own
-quality and the opponent's badge to determine whether to aggressively
-engage in contest. Reaction norms allow individuals to respond to the
-available information without a big cognitive burden. That is at least
-in contrast to other information processing mechanisms like individual
-recognition. With individual recognition individuals associate cue of
-their peers to their quality. These association must be learn thoughout
-individuals lifes, thus it has the usual cognitive requirements of an
-associative learning process. An alternative view of badges of status is
-that individuals learn to react to them based on their experiences
-(Guilford and Dawkins 1991), which would imply that receivers must learn
-to associate signals with the fighting ability of bearers just as in
-systems based on individual recognition. In cases where badges of status
-vary quantitatively (e.g.~in size or intensity), fighting ability may
-increase monotonically with attributes of the badge and would be
-reinforced by every interaction, so in principle learning would be
-faster than in systems involving individual recognition in which the
-association between signals and their meaning would vary depending on
-the interacting partners. In any case, learning could be a central
-cognitive mechanism in both types of communication systems, but the role
-of learning in these contexts has not been thoroughly explored in the
-empirical nor theoretical literature.
+This work was supported by the posdoctoral assistant funding from the
+Depatamento de Ciencias Biológicas from the Universidad de los Andes
+granted to A.Q. and by the by the Swiss National Science Foundation
+grants (grant number: 310030\_192673) to R.B.
 
-Associative learning is a key cognitive mechanism that allows
-individuals to associate rewards with environmental stimuli and thus
-behave adaptively (Staddon 2016). Associative learning exists in all
-major vertebrate taxa, and in many invertebrates as well (Heyes 2012;
-Macphail 1982; Staddon 2016; Behrens et al. 2008). Theory has shown that
-natural selection favours these associations in complex environments
-where conditions are hard to predict (Dridi and Lehmann 2016). Besides
-its wide taxonomic and ecological relevance, associative learning is a
-flexible cognitive mechanism whose underpinnings show interspecific
-variation (Enquist, Lind, and Ghirlanda 2016; Quiñones et al. 2019), so
-presumably, it has been modified by natural selection. Despite all this,
-associative learning is not often included in the narrative of
-evolutionary explanations of behavioural patterns (Fawcett, Hamblin, and
-Giraldeau 2013; Kamil 1983; McAuliffe and Thornton 2015). Computational
-models of evolution can play an important role to overcome the lack of
-integration between learning and evolution. Reinforcement learning
-theory encompasses a series of computational methods inspired on the
-psychological and neurological mechanisms of associative learning
-(Sutton and Barto 2018). These set of algorithms allow the
-implementation of biologically realistic problems, capturing the essence
-of learning processes (Frankenhuis, Panchanathan, and Barto 2018;
-Quiñones et al. 2019). Furthermore, these algorithms can be embedded in
-evolutionary simulations to generate theoretical predictions of the
-effect that learning can have in behavioural evolution (Leimar and
-McNamara 2019).
+\hypertarget{acknowledgements}{%
+\subsection{Acknowledgements}\label{acknowledgements}}
 
-In here we present an evolutionary model where individuals use
-associative learning to develop a tendency to behave aggressively (or
-peacefully), in the context of competition over resources, depending on
-the quantitative morphological trait they perceive in their opponent
-(badge). Over evolutionary time individuals evolve the size of their
-badge, and whether it depends on their quality. Under this simple set
-up, individuals can use the badge as a signal of quality. We use the
-model to assess under what conditions of interaction structure we expect
-different communication signals to evolve.
-
-\hypertarget{the-model}{%
-\subsection{The model}\label{the-model}}
-
-We model the evolution of signals of quality in the context of agonistic
-interactions. For simplicity we consider a population of haploid
-individuals with non-overlapping generations. Individuals are born every
-generation with a level of quality (\(Q_i\), where \(i\) is subscript of
-the population vector of size \(N\)) given by a number between zero and
-one, which is drawn from a truncated normal distribution
-\(N(\mu = 0.5,\sigma)\). An individual with quality \(0\) has the lowest
-RHP, while an individual with quality \(1\) has the highest. As part of
-development processes individuals produce a morphological signal
-(badge), the size of such depends on their level of quality according to
-a reaction norm given by \(B_i=1/(1+e^{\alpha_i-\beta_i Q_i})\); where
-are \(\alpha_i\) and \(\beta_i\) are individual specific traits that
-determine the shape of the reaction norm (Fig. \ref{fig:model-struc} A).
-Variation in \(\alpha_i\) and \(\beta_i\) means individuals can have
-either uninformative (flat) or informative (logistic) norms. The size of
-the signal is constrained between zero and one. We assume different
-values of these traits are given by different alleles, and are inherited
-from mother to offspring. Unless, with a small probability, mutation
-changes the allelic value of the offspring. After birth individuals go
-through a round of viability selection. Individual specific survival
-probability is given by \(s_i=1/(1+e^{-k_1-k_2(Q_i-B_i)})\). Where
-\(k_1\) and \(k_2\) are parameter values determining the shape of the
-survival function. Importantly, if \(k_2=0\) survival probability is
-independent of quality, while if \(k_2>0\) lower quality individuals pay
-a higher price for similar-sized badge, fulfilling the core assumption
-of the handicap principle (Botero et al. 2010; Grafen 1990; Johnstone
-and Norris 1993).
-
-Individuals who survive engage in a series of pairwise interactions
-where they compete for resources. Individuals in real populations
-typically do not interact at random nor with all individuals in the
-population (Kurvers et al. 2014). In order to allow for non-random
-interactions, we use the population vector to define an interaction
-neighbourhood for the focal individual, the size of the neighbourhood
-(\(g\)) is the number of positions around the focal's from which the
-partner is drawn. So, if \(g=N\) interactions are global; as \(g\) gets
-smaller interactions are more local. In an interaction each individual
-must decide whether to escalate a fight or not. Following the classic
-\emph{hawk-dove} game (Maynard-Smith 1982), if the focal individual
-fights and its partner does not, the focal gets as pay-off the resource
-of value \(V\) while its partner gets nothing; if both individuals
-restrain from fighting they split up the resource in half; if both
-decide to fight they split up the cost of fighting and the winner takes
-over the resource. We further assume that the probability of wining a
-fight for the focal individual depends on the difference in quality
-between her and her interacting partner; specifically it is given by
-\(p_{ij}=1/(1+e^{-k_3(Q_i-Q_j)})\), where \(k_3\) is a parameter
-defining how strong the quality difference determines the winning
-probability; and \(i\) and \(j\) denote the the position in the
-population vector of the focal and its interacting partner respectively.
-
-The decision of whether to escalate a fight against an interacting
-partner can be dependent on the size of the partner's badge, the actual
-dependency is determined by the focal´s experiences through a learning
-process. We implement learning using the actor-critic approach from
-reinforcement learning theory {[}RL; Sutton and Barto (2018);Quiñones et
-al. (2019); Leimar and McNamara (2019){]}. Individuals estimate the
-reward (pay-off) expected from interacting with partners of different
-badge sizes (the critic in RL terminology). After every interaction they
-update the estimate of reward proportionally to the difference between
-their current estimate and the observed reward (prediction error
-\(\delta\)) and to the speed of learning (\(\mathrm{A}\)). Furthermore,
-they express different probabilities of retreating (attacking) depending
-on the badge size of their opponent (the actor in RL). They update the
-probability of retreating (attacking) up or down depending on whether
-retreating (attacking) leads to an increase in the reward estimation.
-So, if a focal individual decides to escalate a fight against an
-individual with small badge, and that leads to an increase in the reward
-estimation, the focal individual will increase the probability of
-escalating fights with individuals of small badges in the future. Given
-that badge size is a real number between 0 and 1, there are infinitely
-many badge sizes. Thus, the reward estimation, as well as the
-probability of retreating (escalating), must be generalized across
-different values. To implement generalization we use the linear function
-approximation method based on radial basis functions (Sutton and Barto
-2018). Specifically, we pick \(c\) feature centres, which are evenly
-spaced values along the badge size interval ({[}0,1{]}). Each one of
-these features centres is associated with a weight for reward estimation
-and tendency to play retreat in a given interaction. The reward
-estimation and the tendency to retreat are calculated (in every
-interaction) as the sum of the weights associated with each feature,
-weighted by the response triggered by the feature (Fig.
-\ref{fig:model-struc} B, dots represent the feature weights) The
-response of each feature centre diminishes as a Gaussian function with
-the distance between the feature centre and the badge size of the
-partner (Fig. \ref{fig:model-struc} B, grey line). Formally the reward
-estimation \(\hat{R}\) when the focal individual (\(i\)) faces
-individual \(j\) is given by,\\
-\begin{equation}
-\hat{R}_{ij} =  \sum_{z=1}^{c} x_z e^{(-\frac{|B_i-c_z|}{2\theta^2})} 
-\end{equation}
-
-where \(x_z\) is the weight of feature centre \(z\) on the reward
-estimation; and \(\theta\) is the width of the generalization function.
-Similarly, the tendency to retreat is given by the sum of feature
-weights associated with the actor, and the probability is obtained by
-applying a logistic transformation (Fig. \ref{fig:model-struc} B, black
-line). So formally, the log-odds to retreat when facing individual \(j\)
-is
-
-\begin{equation}
-logit(q_{ij}) =  \sum_{z=1}^{c} y_z e^{(-\frac{|B_i-c_z|}{2\theta^2})} 
-\end{equation}
-
-where \(y_z\) is the weight of feature centre \(z\) on the tendency to
-retreat.
-
-After all their interactions, individuals in the population reproduce
-proportionally to their fitness \(w_i\), which is a sum of the baseline
-fitness (\(w_0\)) and all the pay-offs obtained throughout their life.
-Thus, the combination of natural selection and genetic drift changes the
-distribution of values in \(\alpha\) and \(\beta\) that segregates in
-the population, effectively changing the badges expressed and the
-communication system.
-
-\begin{figure}
-
-{\centering \includegraphics{manuscript_1.0_files/figure-latex/model-struc-1}
-
-}
-
-\caption{Model of communication in the context of aggressive interactions.
-In A the sender code: a reaction norm determines how the badge size is determined by the quality of the individual.
-Red shows an informative reaction norm, while the blue shows a uninformative reaction norm.
-B the receiver code: individuals have a behavioral reaction norm that determines their probability of retreating (black line).
-The reaction norm arises from generalizing the information from the feature weights (black dots).
-Generalization is represented by the grey line and its axis, which shows the response triggered by the fourth feature diminishes as the value evaluated is further from the feature center.
-The learning process moves the feature weights (black dots) up or down depending on whether that leads to an increase in the estimated reward.
-In C and D, effect of learning on the receiver strategy.
-Receivers in C face signallers with uninformative reaction norms (blue line in A), while in D they face signallers with informative reaction norms (red line in A).
-Colour scale in C (applies also for D) indicates the quality of the individual}\label{fig:model-struc}
-\end{figure}
-
-\hypertarget{results}{%
-\subsection{Results}\label{results}}
-
-We first present the outcome of simulations where we prevent the
-evolution of the sender code (by setting the mutation rate to \(0\)),
-and assume that all individuals in the population display either an
-uninformative or informative badge size with respect to their quality.
-These simulations show what type of receiver strategy develops through a
-learning process. Figure \ref{fig:model-struc} (C and D) shows the
-receiver strategy developed through learning; panel C is for receivers
-that faced uninformative signals, and D informative ones. When learners
-face uninformative signals, they change their probability of retreating
-depending on their own quality. Individuals with higher quality (red
-tones), after the learning process, have a high probability of
-attacking; while individuals of lower quality (blue tones) mostly
-retreat from confrontations. Thus, learning splits the population of
-receivers into the two classic pure strategies of hawks and doves. Given
-that we have a assumed a monomorphic population with unresponsive
-reaction norms on the signalling side, the changes triggered by learning
-only affect a small range of badge sizes. In contrast, when receivers
-face informative reaction norms on the side of the signaller (panel B in
-fig.~\ref{fig:model-struc}), receivers use the badge size of their
-interacting partners to determine whether to retreat or attack. The
-relationship is given by a threshold-like reaction norm, where the
-change from retreating to attacking depends on the quality of the
-receiver. As expected, the higher the quality of the receiver the larger
-the badge size that triggers a retreat.
-
-\begin{verbatim}
-## Warning in knitr::include_graphics(here("Simulations", "betCostEvol1_", :
-## It is highly recommended to use relative paths for images. You had
-## absolute paths: "M:/Projects/SocialComp_morphCue/Simulations/betCostEvol1_/
-## evolDyn0_betCost5.png"
-\end{verbatim}
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.9\linewidth]{M:/Projects/SocialComp_morphCue/Simulations/betCostEvol1_/evolDyn0_betCost5}
-
-}
-
-\caption{Evolution of the badge size as a handicap mediated by learning.
-Middle panels show the evolutionary dynamics of the sender code.
-On the left, changes in the distribution of values for the intercept of the reaction norm ($\alpha$); on the right changes in the distribution of the slope ($\beta$).
-Dashed lines in both panels show the mean of the distribution.
-Grey lines show the generation time corresponding to the panels above and below showing the sender and receiver code.
-In the panels above, the reaction norms correspond to individuals in the population after the interaction round.
-Color scale indicates quality just as in Fig.
-\ref{fig:model-struc}.
-In the panels below, colors represent a cluster clasification performed by the k-means algorithm}\label{fig:handicap}
-\end{figure}
-
-When we allow the evolution of the sender code (\(\alpha\) and \(\beta\)
-change subject to natural selection and genetic drift), and the badge
-size work as a handicap (i.e.~the cost of the badge is inversely
-proportional to the quality of the individual), the sender code evolves
-to produce an honest signal Fig \ref{fig:handicap}). The evolution of
-the sender code does not happen immediately after the start of the
-evolutionary process. Instead, the evolutionary dynamics of the reaction
-norm parameters (Fig. \ref{fig:handicap}, X) show a set of steps. First,
-the intercept of the reaction norm (\(\alpha\)) evolves to higher
-values, reducing the average badge size in the populations. That makes
-sense because reaction norms segregating in the population are flat at
-this point, and consequently the receiver code does not respond to the
-badge size, bigger badges are costlier and do not trigger lower attack
-probabilities. During those first generations, the slope of the sender
-reaction norm (\(\beta\)) remains around the initial value of zero. At
-around generation 4000, the slope evolves toward positive values. Which
-implies that larger badges correlate with higher quality. Receivers
-learn to react to such correlation, reducing the probability of attack
-towards individual of larger badges. Hence, natural selection favours
-larger values of \(\beta\), eventually leading to an evolutionary
-equilibrium where badge size is an honest signal of quality mediated by
-the learned responses of receivers. The evolutionary trajectory
-portrayed in figure \ref{fig:handicap} is not the only possible
-trajectory. If the slope of the sender reaction norms (\(\beta\))
-evolves toward negative values before badges become handicaps, receivers
-never learn to react to the size of badges. Thus, badges are only costly
-and do not provide information about the quality of an individual(Fig.
-\ref{fig:noBadge} . Eventually, the badge disappears from the
-population.
-
-When production of the badge is not costly (i.e.~does not work as a
-handicap), the parameters of the sender code go through evolutionary
-branching event, triggering the evolution of different types of signals
-in the population. Fig. \ref{fig:branching}A and B shows changes in the
-distribution of the sender reaction norm (\(\alpha\) and \(\beta\))
-along evolutionary time. Darker areas show trait values with high
-frequency in the population. Populations start monomorphic with a value
-of zero on both traits, and mutations quickly build up a normal
-distribution around the starting value. Within the first 2000
-generations the unimodal distribution splits into a bimodal one. In the
-case of \(\alpha\), one of the peaks splits further, so at the end of
-the evolutionary simulation the distribution of \(\alpha\) value in the
-population shows three peaks. In the case of \(\beta\), the two initial
-peaks coexist for around 4000 generations. Later on one of the types
-goes extinct and the distribution goes back to being unimodal. These
-changes in the distribution of the sender reaction norm parameters imply
-that during the second half of the simulation individuals can be
-classified into three distinct types. Two types express a flat reaction
-norm, meaning that their badge size is not informative of their quality,
-while the third type shows intermediate badge sizes which are determined
-by the quality of the individual (Fig. \ref{fig:branching}).
-Furthermore, the receiver reaction norms develop through learning,
-particularly those of individuals of intermediate quality, respond to
-the signal of their sender type by increasing the probability of
-retreating from a fight with individuals with larger badges (Fig.
-\ref{fig:branching}).
-
-\begin{verbatim}
-## Warning in knitr::include_graphics(here("Simulations", "nIntGroupNormQual_", :
-## It is highly recommended to use relative paths for images. You had absolute
-## paths: "M:/Projects/SocialComp_morphCue/Simulations/nIntGroupNormQual_/
-## evolDyn7_nIntGroup2000.png"
-\end{verbatim}
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.9\linewidth]{M:/Projects/SocialComp_morphCue/Simulations/nIntGroupNormQual_/evolDyn7_nIntGroup2000}
-
-}
-
-\caption{The evolution of cheap signals.
-Portrait of the evolutionary dynamics of the sender code with snapsots of both sender and receiver codes just as in fig.
-\ref{fig:handicap}.
-The middle panels show changes in the distribution of values fir $\alpha$ and $\beta$ along evolutionary time.
-Panels above and bellow correspond to snapshots of the sender and receiver codes, respectively, generation time of the snapshots are indicated by the grey lines in the middle panels.}\label{fig:branching}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.9\linewidth]{../Simulations/initAct_/indVarScatter_}
-
-}
-
-\caption{WHy would you want to be recognized}\label{fig:startCond}
-\end{figure}
-
-\hypertarget{supplementary-material}{%
-\section{Supplementary material}\label{supplementary-material}}
-
-\beginsupplement
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.9\linewidth]{M:/Projects/SocialComp_morphCue/Simulations/betCostEvol1_/evolDyn5_betCost5}
-
-}
-
-\caption{The evolution of cheap signals.
-Portrait of the evolutionary dynamics of the sender code with snapsots of both sender and receiver codes just as in fig.
-\ref{fig:handicap}.
-The middle panels show changes in the distribution of values fir $\alpha$ and $\beta$ along evolutionary time.
-Panels above and bellow correspond to snapshots of the sender and receiver codes, respectively, generation time of the snapshots are indicated by the grey lines in the middle panels.}\label{fig:noBadge}
-\end{figure}
-
-\hypertarget{references}{%
-\subsection*{References}\label{references}}
-\addcontentsline{toc}{subsection}{References}
-
-\hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-behrens_Associative_2008}{}}%
-Behrens, Timothy E. J., Laurence T. Hunt, Mark W. Woolrich, and Matthew
-F. S. Rushworth. 2008. {``Associative Learning of Social Value.''}
-\emph{Nature} 456 (7219): 245--49.
-\url{https://doi.org/10.1038/nature07538}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-botero_Evolution_2010}{}}%
-Botero, C. A, I. Pen, J. Komdeur, and F. J Weissing. 2010. {``The
-Evolution of Individual Variation in Communication Strategies.''}
-\emph{Evolution} 64 (11): 3123--33.
-
-\leavevmode\vadjust pre{\hypertarget{ref-bradbury_Principles_2011}{}}%
-Bradbury, Jack W., and Sandra L. Vehrencamp. 2011. \emph{Principles of
-{Animal Communication}}. 2nd Edition. {Sunderland, Mass}: {Sinauer
-Associates is an imprint of Oxford University Press}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-dridi_Environmental_2016}{}}%
-Dridi, Slimane, and Laurent Lehmann. 2016. {``Environmental Complexity
-Favors the Evolution of Learning.''} \emph{Behav Ecol} 27 (3): 842--50.
-\url{https://doi.org/10.1093/beheco/arv184}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-enquist_Power_2016}{}}%
-Enquist, Magnus, Johan Lind, and Stefano Ghirlanda. 2016. {``The Power
-of Associative Learning and the Ontogeny of Optimal Behaviour.''}
-\emph{Royal Society Open Science} 3 (11): 160734.
-\url{https://doi.org/10.1098/rsos.160734}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-fawcett_Exposing_2013}{}}%
-Fawcett, Tim W., Steven Hamblin, and Luc-Alain Giraldeau. 2013.
-{``Exposing the Behavioral Gambit: The Evolution of Learning and
-Decision Rules.''} \emph{Behavioral Ecology} 24 (1): 2--11.
-\url{https://doi.org/10.1093/beheco/ars085}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-frankenhuis_Enriching_2018}{}}%
-Frankenhuis, Willem E., Karthik Panchanathan, and Andrew G. Barto. 2018.
-{``Enriching Behavioral Ecology with Reinforcement Learning Methods.''}
-\emph{Behavioural Processes}, February.
-\url{https://doi.org/10.1016/j.beproc.2018.01.008}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-grafen_Biological_1990}{}}%
-Grafen, Alan. 1990. {``Biological Signals as Handicaps.''} \emph{Journal
-of Theoretical Biology} 144 (4): 517--46.
-\url{https://doi.org/10.1016/S0022-5193(05)80088-8}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-guilford_Receiver_1991}{}}%
-Guilford, Tim, and Marian Stamp Dawkins. 1991. {``Receiver Psychology
-and the Evolution of Animal Signals.''} \emph{Animal Behaviour} 42 (1):
-1--14. \url{https://doi.org/10.1016/S0003-3472(05)80600-1}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-heyes_Simple_2012}{}}%
-Heyes, Cecilia. 2012. {``Simple Minds: A Qualified Defence of
-Associative Learning.''} \emph{Phil. Trans. R. Soc. B} 367 (1603):
-2695--2703. \url{https://doi.org/10.1098/rstb.2012.0217}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-johnstone_Recognition_1997}{}}%
-Johnstone, Rufus A. 1997. {``Recognition and the Evolution of
-Distinctive Signatures: When Does It Pay to Reveal Identity?''}
-\emph{Proceedings of the Royal Society of London. Series B: Biological
-Sciences} 264 (1387): 1547--53.
-\url{https://doi.org/10.1098/rspb.1997.0215}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-johnstone_Badges_1993}{}}%
-Johnstone, Rufus A., and Ken Norris. 1993. {``Badges of Status and the
-Cost of Aggression.''} \emph{Behav Ecol Sociobiol} 32 (2): 127--34.
-\url{https://doi.org/10.1007/BF00164045}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-kamil_Optimal_1983}{}}%
-Kamil, Alan C. 1983. {``Optimal {Foraging Theory} and the {Psychology}
-of {Learning}.''} \emph{Integr Comp Biol} 23 (2): 291--302.
-\url{https://doi.org/10.1093/icb/23.2.291}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-kurvers_Evolutionary_2014}{}}%
-Kurvers, Ralf H. J. M., Jens Krause, Darren P. Croft, Alexander D. M.
-Wilson, and Max Wolf. 2014. {``The Evolutionary and Ecological
-Consequences of Animal Social Networks: Emerging Issues.''} \emph{Trends
-in Ecology \& Evolution} 29 (6): 326--35.
-\url{https://doi.org/10.1016/j.tree.2014.04.002}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-leimar_Learning_2019}{}}%
-Leimar, Olof, and John M. McNamara. 2019. {``Learning Leads to Bounded
-Rationality and the Evolution of Cognitive Bias in Public Goods
-Games.''} \emph{Scientific Reports} 9 (1): 1--9.
-\url{https://doi.org/10.1038/s41598-019-52781-7}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-macphail_Brain_1982}{}}%
-Macphail, Euan M. 1982. \emph{Brain and {Intelligence} in
-{Vertebrates}}. {Clarendon Press}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-maynard-smith_Evolution_1982}{}}%
-Maynard-Smith, John. 1982. \emph{Evolution and the {Theory} of {Games}}.
-1 edition. {Cambridge ; New York}: {Cambridge University Press}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-mcauliffe_Psychology_2015}{}}%
-McAuliffe, K., and A. Thornton. 2015. {``The Psychology of Cooperation
-in Animals: An Ecological Approach.''} \emph{J Zool} 295 (1): 23--35.
-\url{https://doi.org/10.1111/jzo.12204}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-moller_Female_1988}{}}%
-Møller, Anders Pape. 1988. {``Female Choice Selects for Male Sexual Tail
-Ornaments in the Monogamous Swallow.''} \emph{Nature} 332 (6165):
-640--42. \url{https://doi.org/10.1038/332640a0}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-parker_Assessment_1974}{}}%
-Parker, G. A. 1974. {``Assessment Strategy and the Evolution of Fighting
-Behaviour.''} \emph{Journal of Theoretical Biology} 47 (1): 223--43.
-\url{https://doi.org/10.1016/0022-5193(74)90111-8}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-quinones_Negotiation_2016}{}}%
-Quiñones, Andrés E., G. Sander van Doorn, Ido Pen, Franz J. Weissing,
-and Michael Taborsky. 2016. {``Negotiation and Appeasement Can Be More
-Effective Drivers of Sociality Than Kin Selection.''} \emph{Phil. Trans.
-R. Soc. B} 371 (1687): 20150089.
-\url{https://doi.org/10.1098/rstb.2015.0089}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-quinones_Reinforcement_2019}{}}%
-Quiñones, Andrés E., Olof Leimar, Arnon Lotem, and Redouan Bshary. 2019.
-{``Reinforcement {Learning Theory Reveals} the {Cognitive Requirements}
-for {Solving} the {Cleaner Fish Market Task}.''} \emph{The American
-Naturalist}, December, 000--000. \url{https://doi.org/10.1086/707519}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-rohwer_Social_1975a}{}}%
-Rohwer, Sievert. 1975. {``The {Social Significance} of {Avian Winter
-Plumage Variability}.''} \emph{Evolution} 29 (4): 593--610.
-\url{https://doi.org/10.2307/2407071}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-staddon_Adaptive_2016}{}}%
-Staddon, J. E. R. 2016. \emph{Adaptive {Behavior} and {Learning}}. 2
-edition. {Cambridge}: {Cambridge University Press}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-sutton_Reinforcement_2018}{}}%
-Sutton, Richard S., and Andrew G. Barto. 2018. \emph{Reinforcement
-{Learning}: {An Introduction}}. Edited by Francis Bach. Second edition
-edition. {Cambridge, MA}: {A Bradford Book}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-tibbetts_Socially_2004}{}}%
-Tibbetts, Elizabeth A., and James Dale. 2004. {``A Socially Enforced
-Signal of Quality in a Paper Wasp.''} \emph{Nature} 432 (7014): 218--22.
-\url{https://doi.org/10.1038/nature02949}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-zahavi_Mate_1975}{}}%
-Zahavi, Amotz. 1975. {``Mate Selection\textemdash -{A} Selection for a
-Handicap.''} \emph{Journal of Theoretical Biology} 53 (1): 205--14.
-\url{https://doi.org/10.1016/0022-5193(75)90111-3}.
-
-\end{CSLReferences}
+AQ and CDC kindly thank the members of Evolvert, in particular Jorge
+Avendaño for pictures of the \emph{Arremon} and discussion on signal
+evolution. AQ thanks C. Botero for fruitful discussions, and thecluster
+of the Institute of Computer Science of the University of Neuchâtel
+(IIUN) for the computing resources provided.
 

### 2023-09-14 manuscript/manuscript_1.0.Rmd
Title_page rmd file and readme md file, getting readu for realease and submission

@@
 title: |
-  The role of learning in status signalling: a modeling approach
+   \large \textbf{Title:} The role of learning in the evolution of status signalling: a modeling approach
+subtitle: |
+  \normalsize \textbf{Abbreviated title:} Learning and evolution in status signalling
 author: 
-- Andrés E. Quiñones:
-    email: "andresqp@gmail.com"
-    institute: [andes, unine]
-    correspondence: true
-- Redouan Bshary:
-    institute: unine
-- C. Daniel Cadena:
-    institute: andes
-institute:
-- andes: Laboratorio de Biología Evolutiva de Vertebrados, Departamento  de Ciencias Biológicas, Universidad de los Andes, Bogotá, Colombia
-- unine: Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland
-date: "`r format(Sys.time(), '%B %d, %Y')`"
+# author:
+# - Andrés E. Quiñones:
+#     email: "andresqp@gmail.com"
+#     institute: [andes, unine]
+#     correspondence: true
+# - Redouan Bshary:
+#     institute: unine
+# - C. Daniel Cadena:
+#     institute: andes
+# institute:
+# - andes: Laboratorio de Biología Evolutiva de Vertebrados, Departamento  de Ciencias Biológicas, Universidad de los Andes, Bogotá, Colombia
+# - unine: Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland
+# date: "`r format(Sys.time(), '%B %d, %Y')`"
+# output:
+#   pdf_document:
+#     latex_engine: xelatex  # default  ##word_document: default  # word_document:default
+#     pandoc_args:
+#       - '--lua-filter=scholarly-metadata.lua'
+#       - '--lua-filter=author-info-blocks.lua'
+#   word_document: default
 output:
   pdf_document:
-    latex_engine: xelatex  # default  ##word_document: default  # word_document:default
-    pandoc_args:
-      - '--lua-filter=scholarly-metadata.lua'
-      - '--lua-filter=author-info-blocks.lua'
+    keep_tex: yes
+    latex_engine: xelatex
+    template: template.latex
   word_document: default
+keep_tex: yes
 fig_caption: yes
@@
   ```{=latex}
+    \AddToHook{env/abstract/before}{\vspace*{-4.5cm}}
     \usepackage{float}
+    \usepackage{graphicx}
     \usepackage{unicode-math}
@@
   ```
-keep_tex: yes
 keywords: My keywords
@@
 spacing: double
-style: C:\Users\andre\Zotero\styles\elife.csl
+style: "M:\\Zotero\\styles\\behavioral-ecology.csl"
 abstract: |
@@
   behavioral responses.
+lay_summary:  |
+  Biological signals allow individuals to respond appropietly. For example, a 
+  signal of fighting hability would allow peers to decide whether to fight 
+  another individual. But, what prevents weaklings from faking to be 
+  strong? Using a theoretical approach we assess whether learning has 
+  a role in maintaining a signal honest. We find that when individuals 
+  learn how best to respond to a signal, populations are composed of a 
+  mixed of honest and dishonest signals. 
+editor_options:
+  markdown:
+    wrap: 80
 ---
@@
 
+\newpage
+
 ## Introduction
@@
 engaging in a social interaction might vary depending on its 
-own condition, and that of individuals with 
+own condition and that of individuals with 
 whom it interacts. Thus, acting based on information about interacting 
@@
 or to provide deceptive information [@johnstone_Badges_1993]. 
-Typically, in a given context, some individuals benefit 
-from broadcasting accurate information, whereas others benefit 
+Typically, in a given context, some individuals benefit may 
+from broadcasting accurate information, whereas others may benefit 
 from concealing it. Take, for instance, an interaction between two individuals 
@@
 to communication systems in which individuals 
-exhibit badges of status, where a signal conveys quality 
-[@johnstone_Badges_1993; @rohwer_Social_1975]. 
-Such signals tend to be arbitrary, meaning the signals could potentially be 
-produced by high- and low-quality individuals alike. This symmetry opens up the 
-possibility of dishonest signalling and the question of evolutionary stability. 
+exhibit badges of status, where a signal typically conveys quality 
+[@johnstone_Badges_1993; @rohwer_Social_1975a]. 
+Such signals tend to be arbitrary, however, meaning the signals could potentially be 
+produced by high- and low-quality individuals alike, opening up the 
+possibility of dishonest signalling and raising the question of evolutionary stability. 
 Signals can be evolutionarily stable whenever low-quality individuals 
@@
 wasps, subordinate individuals with experimentally manipulated dominant-like facial
-patterns received more aggression from the dominant[@tibbetts_Socially_2004]. 
+patterns received more aggression from the dominant [@tibbetts_Socially_2004]. 
 In this case, costs are socially rather than developmentally mediated and are 
@@
 This contrasts with other information-processing 
-mechanisms like individual recognition. Where individuals associate cues of 
+mechanisms like individual recognition where individuals associate cues of 
 their peers to their quality. Because these 
@@
 
-Here we present an evolutionary model where individuals use 
+Here, we present an evolutionary model where individuals use 
 associative learning to develop a tendency to behave either aggressively 
@@
 We use the model to assess under what conditions 
-we expect communication signals to evolve as handicaps or conventions.
+one might expect communication signals to evolve as handicaps or conventions.
 
@@
 in the context of agonistic interactions. 
-For simplicity we consider a population of haploid 
+For simplicity, we consider a population of haploid 
 individuals with non-overlapping generations. Individuals are born 
@@
 $\beta_i$ are individual specific traits that determine the shape of 
-the reaction norm (Fig. \ref{fig:model-struc} A). Variation in $\alpha_i$ 
+the reaction norm (Fig. \ref{fig:modelStruc} A). Variation in $\alpha_i$ 
 and $\beta_i$ means individuals can have either uninformative (flat) or 
@@
 along the badge size interval ([0,1]) where the updates are focused. 
-For the sake of simplicity we keep the location of these feature centres 
+For simplicity, we keep the location of these feature centres 
 constant and  the same for all individuals; they are stored in 
@@
 associated with each feature centre, weighted by the response 
-triggered by the feature (Fig. \ref{fig:model-struc} B, dots 
+triggered by the feature (Fig. \ref{fig:modelStruc} B, dots 
 represent the feature weights). The response of each feature centre 
@@
 feature centre and the badge size of 
-the partner (Fig. \ref{fig:model-struc} B, grey line). Formally, the 
+the partner (Fig. \ref{fig:modelStruc} B, grey line). Formally, the 
 reward estimation $\hat{R}$ when the focal individual ($i$) faces 
@@
 with the actor, and the probability is obtained by applying a logistic 
-transformation (Fig. \ref{fig:model-struc} B, black line). 
+transformation (Fig. \ref{fig:modelStruc} B, black line). 
 Thus formally, the log-odds to retreat when facing individual $j$ is
@@
 
-```{r model-struc,echo=FALSE,warning=FALSE,fig.width=8,fig.align='center',fig.height=7,fig.show='hold',fig.cap= "Model of communication in the context of aggressive interactions.
+```{r modelStruc,echo=FALSE,out.width='100%',fig.height=5,fig.align='center',fig.show='hold',fig.cap= "Model of communication in the context of aggressive interactions.
 In A, the reaction norm determines the badge size, *e.g* the size of a black band on a bird's chest.
@@
 Accordingly, receivers develop a threashold-like reaction norm where the decision to retreat increases with the badge size of the interacting partner.
-Colour scale in C and D indicates the quality of the individual.
-"}
+Colour scale in C and D indicates the quality of the individual." ,warning=FALSE}
 
 # Top panels -  A the sender code and  B learning for the receiver code
-# load(here("manuscript","manuscript_1.0_cache","latex","model-struc_6905f4797a1faa21041511a8dcdf1eb9.RData"))
+# load(here("manuscript","manuscript_1.0_cache","latex","modelStruc_6905f4797a1faa21041511a8dcdf1eb9.RData"))
 library(dplyr)
@@
 dishonest<-readPNG(here("Images/cart_dishonest.png"))
-par(plt=posPlot(numplotx = 2,idplotx = 1,numploty = 1,idploty = 1,upboundx = 91),
+par(plt=posPlot(numplotx = 2,idplotx = 1,numploty = 1,idploty = 1,upboundx = 91,
+                lowboundy = 18),
     las=1,mfrow=c(2,1),mgp=c(2,0.7,0))
@@
  text(x = 0.08,y=1.05,"A",cex = 1.5)
- mtext("Quality",1,line = 2,cex = 1,las = 0)
+ mtext("Quality",1,line = 1.5,cex = 1,las = 0)
  matlines(x=rangQual,y=linesSender,lwd=4,col=colRuns,lty=1)
- mtext("Badge size",2,line = 2.3,cex = 1,las = 0)
+ mtext("Badge size",2,line = 2,cex = 1,las = 0)
  # include bird cartoons
@@
 par(plt=posPlot(numplotx = 2,idplotx = 2,numploty = 1,idploty = 1,
-                 upboundx = 91),las=1,new=TRUE)
+                 upboundx = 91,lowboundy = 18),las=1,new=TRUE)
   nCenters<-6
@@
 mtext("prob. retreat",4,line = 2.3,cex = 1,las = 0)
-mtext("Badge size",1,line = 2,cex = 1,las = 0)
+mtext("Badge size",1,line = 1.5,cex = 1,las = 0)
 
@@
   # par(plt=posPlot(numploty = 4,numplotx = 4,idploty = 2,idplotx = 1))
-  par(plt=posPlot(numploty = 1,numplotx = 2,idploty = 1,idplotx = 1,upboundx = 91))
+par(plt=posPlot(numploty = 1,numplotx = 2,idploty = 1,idplotx = 1,upboundx = 91,
+                  lowboundy = 18))
 matplot(x=rangx,y=dataIndsAct,type='l',xlab="",ylab="prob. retreat",
@@
   # par(plt=posPlot(numploty = 4,numplotx = 4,idploty = 2,idplotx = 1))
-  par(plt=posPlot(numploty = 1,numplotx = 2,idploty = 1,idplotx = 2,upboundx = 91)
+  par(plt=posPlot(numploty = 1,numplotx = 2,idploty = 1,idplotx = 2,upboundx = 91,
+                  lowboundy = 18)
       ,new=TRUE,xpd=TRUE)
@@
 text(x = 0.08,y=1.05,"D",cex = 1.5)
-text(x = -0.1,y = -0.25,labels = "Badge size")
+text(x = -0.1,y = -0.3,labels = "Badge size")
 
@@
 a learning process in such a monomorphic population. 
-Figure \ref{fig:model-struc} (C and D) shows 
+Figure \ref{fig:modelStruc} (C and D) shows 
 the receiver strategy developed through learning; panel C is for 
 receivers that faced uninformative signals, and D for those facing informative ones. 
-The simulations reveal that when learners face uninformative signals (Fig. \ref{fig:model-struc} C), they modify their probability of retreating 
+The simulations reveal that when learners face uninformative signals (Fig. \ref{fig:modelStruc} C), they modify their probability of retreating 
 depending on their own quality. 
 Individuals with higher quality (red tones) have a 
-high probability of attacking after the learning process; 
+high probability of attacking after the learning process  
 while individuals of lower quality 
-(blue tones) mostly retreat from confrontations; in both cases 
+(blue tones) mostly retreat from confrontations; in both cases, 
 individuals behave equally irrespective of the quality of
@@
 with unresponsive reaction norms on the signalling side, the changes 
-triggered by learning only affect a small range of badge sizes. Specifically, 
+triggered by learning only affected a small range of badge sizes. Specifically, 
 all learning occurred around a badge size of $0.5$, because all
@@
 In contrast, when receivers face informative reaction norms 
-on the side of the signaller (Fig. \ref{fig:model-struc} B), 
+on the side of the signaller (Fig. \ref{fig:modelStruc} B), 
 receivers use information on the badge size of their interacting 
@@
 Color scale indicates quality just as in Fig.
-\\ref{fig:model-struc}.
+\\ref{fig:modelStruc}.
 On the first half of the simulation (B.1 and B.2) the badge size evolves to its minimum value.
@@
 worked as a handicap (*i.e.* the cost of the badge was inversely 
-proportional to the quality of the individual), the sender code evolved often
+proportional to the quality of the individual), the sender code often evolved 
 to produce an honest signal (Fig. \ref{fig:handicap}). The evolution of 
@@
 
-The amount of information that agents are able to collect through learning 
+The amount of information that agents are able to collect via learning 
 over their life time can strongly change the outcome of 
@@
 the intercept ($\alpha$) splits into a bimodal one. 
-Later in evolutionary time one of the peaks splits further, 
+Later in evolutionary time, one of the peaks splits further, 
 so at the end of the evolutionary simulation the distribution of the intercept 
@@
 the evolving parameters when the signal follows the handicap
-principle as well. In fig \ref{fig:honestDiv}, we show simulations where individuals 
+principle as well. In fig. \ref{fig:honestDiv}, we show simulations where individuals 
 have on average 300 interactions in their life and the cost of the signal
@@
 learn fast and interact repeatedly will diminish the strength 
-of frequency dependent selection.
+of frequency-dependent selection.
 
@@
 do not override the conditions for the evolutionary branching 
-[@dieckmann_Dynamical_1996] described earlier. 
+[@dieckmann_Dynamical_1996] described above. 
 Given how crucial naive behaviour is for the branching event 
@@
 We presented a model that integrates learning with 
-evolutionary processes. In the model the response of a signal in a 
-communication system is mediated by learning processes. Individuals 
-throughout their life learn the best way to respond to a quantitative 
-trait of their interactive partners. 
-We show that the learning process can mediate the evolution 
+evolutionary processes. In the model, the response to a signal in a 
+communication system is mediated by learning processes: individuals 
+learn how to respond to a quantitative trait of their interactive partners. 
+We showed that the learning process can mediate the evolution 
 of an honest signal under the handicap principle
@@
 a limit in the amount of information collected by individuals, as well as
-peaceful behaviour in naive individuals. Furthermore, we have shown that the 
+peaceful behaviour in naive individuals. Furthermore, we showed that the 
 conditions leading to the origin of the polymorphism can be reached when evolutionary
@@
 adaptive choices. Specifically, lower quality individuals refrained more from
-escalating fights, while higher quality took over resources they 
+escalating fights, while higher-quality individuals took over resources they 
 could win in a fight. Thus, the learning process we modelled and the 
-response it mediates has fitness relevant consequences even in the 
+response it mediates has fitness-relevant consequences even in the 
 absence of an honest signal. This was confirmed by simulations where 
@@
 only captured the effect of selection mediated by the social 
-game on the learning parameters. For example, We did not account for 
-potential costs of faster updating, or improved performance in foraginf tasks. 
+game on the learning parameters. For example, we did not account for 
+potential costs of faster updating, or improved performance in foraging tasks. 
 Inter-specific variation in cognitive abilities mediated by environmental 
@@
 
-Learning processes collecting information on a population wide level promote 
-frequency dependent selection. Frequency-dependence triggered by 
+Learning processes which involve collecting information on a population-wide level promote 
+frequency-dependent selection. For example, frequency-dependence triggered by 
 learning process was highlighted by a 
@@
 of *in silico* polymorphic prey [@bond_Visual_2002]. 
-The key to the evolution of prey polymorphism is 
-that predators are better able to discriminate prey from the background, 
-when preys are frequently encountered.
+The key to the evolution of prey polymorphism in such system is 
+that predators are better able to discriminate prey from the background 
+when prey are frequently encountered.
 Thus, prey morphs found in low frequency in a population have a selective advantage. 
@@
 behaviours dependent on learning processes seem to generally trigger 
-frequency-dependent selection. That is because the learning algorithm 
+frequency-dependent selection because the learning algorithm 
 collects more information on the more frequent values of the the trait distribution. 
-For the sake of simplicity we modelled a single dimension (badge size), 
-however the frequency-dependen effect of learning could potentially be 
+For simplicity, we modelled a single dimension (badge size), 
+yet the frequency-dependent effect of learning could potentially be 
 more important when individuals learn from multidimensional signals.
 
-Originally, @rohwer_Social_1975 proposed the idea that certain phenotypic 
+Originally, @rohwer_Social_1975a proposed the idea that certain phenotypic 
 traits could be used by animals as status signals or signals of quality. 
@@
 such as the bibs exhibited by species in various families, particularly sparrows.
-So, are bibs real signals of quality? if so are they handicaps or 
+So, are bibs real signals of quality? If so, are they handicaps or 
 badge of status? From an evolutionary perspective it is unclear how the stability of 
@@
 Alternatively, if the signal does not carry production costs inversely 
-proportional to quality it may work as a convention.
-In this later case, honesty is presumably maintained by the aggressive 
+proportional to quality, then it may work as a convention.
+In this latter case, honesty is presumably maintained by the aggressive 
 reaction of receivers when the convention is 
-broken [@enquist_Signaling_2010;@tibbetts_Socially_2004]. Our model, captures 
+broken [@enquist_Signaling_2010;@tibbetts_Socially_2004]. Our model captures 
 both kinds of signals, although with some nuances. When we assume the handicap 
 principle, simulations result in the evolution of honest signals mediated by the
-learned response. When we assume cost-free signal, under certain conditions the
+learned response. When we assume a cost-free signal, under certain conditions the
 learned response mediated the emergence of phenotypic variation in the signal. 
-This variation, particularly in the mid-range of the distribution facilitates 
-the establishment of a convention, in this range individuals respond 
-appropriately to the trait of their peer. The house sparrow (*Passer domesticus*) 
-has been a text-book example of badges of status. According to the prevailing narrative the bib size in males is a signal of dominance rank and quality.
-However, a recent meta-analysis called into question 
-this narrative by showing that the effect size of the association between bib 
-size and dominance rank is small and uncertain [@sanchez-tojar_Metaanalysis_2018]. 
-A simple correlation between bib size and dominance rank
+This variation, particularly in the mid-range of the distribution, facilitates 
+the establishment of a convention: in this range, individuals respond 
+appropriately to the trait of their peer. 
+A simple correlation between the size os a signal and dominance rank
 (and quality) is expected under the handicap principle, as it is shown here and 
 elsewhere [@botero_Evolution_2010;@grafen_Biological_1990;@johnstone_Badges_1993].
-However, the branching event presented here shows that conventions can 
+However, the branching event we documented shows that conventions can 
 emerge in a less straightforward way when mediated by learning processes. These 
 nuances could potentially make sense of seemingly contradictory evidence on the 
-correlation between plumage traits and quality. For example, conventions in our
-simulations arose only in an intermediate range of the signalling trait, yet 
+correlation between plumage traits and quality. A case in point is that of the 
+House Sparrow (*Passer domesticus*), a textbook example of badges of status. 
+According to the prevailing narrative [@3062aab1-6f44-361c-a548-6bc040ba8ca2;@searcy_Evolution_2005], 
+the bib size of males of this species is a signal of dominance rank and quality. 
+However, a recent meta-analysis called into question this narrative by showing 
+that the effect size of the association between bib size and dominance rank 
+is small and uncertain [@sanchez-tojar_Metaanalysis_2018a].
+Our simulations revealed that, conventions 
+arose only in an intermediate range of the signalling trait, yet 
 empirical studies seeing for correlations between phenotypic traits and quality 
 are often performed across the whole range of variation. 
+Whether such a situation may apply empirically to bib sizes in House Sparrows 
+remains to be studied.
 
@@
 
-We have presented here an evolutionary model where the evolution of a 
+In sum, we have presented here an evolutionary model where the evolution of a 
 communication system is mediated through the learned responses of

### 2023-09-14 manuscript/manuscript_1.0.tex
Title_page rmd file and readme md file, getting readu for realease and submission

@@
 \usepackage{xcolor}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\hypersetup{
-  pdfauthor={Andrés E. Quiñones; C. Daniel Cadena*; Redouan Bshary},
-  pdfkeywords={My keywords},
-  hidelinks,
-  pdfcreator={LaTeX via pandoc}}
-\urlstyle{same} % disable monospaced font for URLs
-\usepackage[margin=1in]{geometry}
-\usepackage{graphicx}
+\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
 \makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
 \makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-% Set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
 \setlength{\emergencystretch}{3em} % prevent overfull lines
@@
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+  \AddToHook{env/abstract/before}{\vspace*{-4.5cm}}
   \usepackage{float}
+  \usepackage{graphicx}
+  \usepackage{unicode-math}
   \usepackage{setspace}\doublespacing
@@
   \floatplacement{figure}{H}
- \newcommand{\beginsupplement}{ \setcounter{table}{0}     \renewcommand{\thetable}{S\arabic{table}}\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
+  \newcommand{\beginsupplement}{ \setcounter{table}{0}            
+  \renewcommand{\thetable}{S\arabic{table}}\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
 
@@
 \fi
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same} % disable monospaced font for URLs
+\hypersetup{
+  pdftitle={ The role of learning in the evolution of status signalling: a modeling approach},
+  pdfkeywords={My keywords},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
 
-\title{Learning can mediate both badges of status and individual
-recognition}
-\author{Andrés E. Quiñones\footnote{Laboratorio de Biología Evolutiva de
-  Vertebrados, Departamento de Ciencias Biológicas, Universidad de los
-  Andes, Bogotá, Colombia} \and C. Daniel Cadena* \and Redouan
-Bshary\footnote{Institute of Biology, University of Neuchâtel,
-  Neuchâtel, Switzerland}}
-\date{July 18, 2022}
+\title{\large \textbf{Title:} The role of learning in the evolution of
+status signalling: a modeling approach}
+\usepackage{etoolbox}
+\makeatletter
+\providecommand{\subtitle}[1]{% add subtitle to \maketitle
+  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
+}
+\makeatother
+\subtitle{\normalsize \textbf{Abbreviated title:} Learning and evolution
+in status signalling}
+\date{}
+
+\usepackage{forarray}
+\usepackage{xstring}
+\newcommand{\getIndex}[2]{
+  \ForEach{,}{\IfEq{#1}{\thislevelitem}{\number\thislevelcount\ExitForEach}{}}{#2}
+}
+
+\newcommand{\getAff}[1]{
+  \getIndex{#1}{}
+}
 
@@
 \maketitle
+\begin{flushleft}
+\end{flushleft}
 \begin{abstract}
@@
 decisions. However, dishonest individuals can destabilize such
-communications systems. One open question is whether cognitive
-mechanisms, such as learning, can maintain the honesty of signals, thus,
-favoring their evolutionary stability. Here we present evolutionary
-simulations, where individuals can produce a signal proportional to
-their quality; and learn along their lifetime the best response to the
-signal emitted by their peers. In these simulations, learning on the
-receiver side can mediate the evolution of signals of quality on the
-sender side. When the cost of the signal is proportional to the quality
-of the sender populations are only composed of honest signalers. When
-the cost is not proportional to the quality of the signaler, the
-population is composed of both honest and dishonest signalers. We argue
-that learning can be a general cognitive mechanism playing a role in a
-wide range of communication systems.
+communications systems. An open question is whether cognitive
+mechanisms, such as learning, can maintain the honesty of signals, thus
+favoring their evolutionary stability. We present evolutionary
+simulations where individuals can produce a signal proportional to their
+quality and learn along their lifetime the best response to the signals
+emitted by their peers. Our simulations replicate previous results where
+the handicap principle mediates the evolution of signals. In our
+simulations learning on the receiver side can mediate the evolution of
+signals of quality on the sender side. When the cost of the signal is
+proportional to the quality of the sender, all individuals in
+populations are honest signalers. In contrast to traditional models
+which predict the absence of signals when the cost is not proportional
+to the quality of the signaler, our model revealed that learning
+facilitates the evolution of a polymorphism in which populations
+comprise both honest and dishonest signalers. We argue that learning may
+have a role in the evolution and dynamics of a wide range of
+communication systems and more generally in behavioral responses.
 \end{abstract}
+\subsection*{Lay summary}
+Biological signals allow individuals to respond appropietly. For
+example, a singal of fighting hability would allow peers to decide
+whether to fight another individual. But, what prevents a weaklings from
+faking to be strong? Using a theoretical approach we assess whether
+learning has a role in maintaining signals honest. We find that when
+individuals learn how best to respond to a signal, populations are
+composed of a mixed of honest and dishonest signals.
+
+\newpage
 
@@
 
-The outcome of interactions with con-specifics is an important
+The outcome of interactions with conspecific individuals is a critical
 determinant of fitness in social animals. Irrespective of whether such
 interactions are cooperative or competitive, the actions of interacting
-individuals partly determine their reproductive success. However, the
-best action for an individual might vary depending on its condition, and
-that of individuals with whom it interacts. Thus, acting based on
-information about interacting partners is typically adaptive (Quiñones
-et al. 2016). Acquiring this information, however, is far from trivial.
-In some cases, interacting partners (signallers) might be `willing' to
-provide accurate information, but in others it might be in their own
-interest to conceal information (Johnstone 1997), or to provide
-erroneous information (Johnstone and Norris 1993). Typically, in a given
-context, some proportion of individuals will benefit from broadcasting
-accurate information, whereas others will benefit from hiding it. Take,
-for instance, an interaction between two individuals where one can help
-the other. Given some costs and benefits, a potential donor will be
-interested in helping related individuals. Therefore, for a relative of
-the donor, broadcasting kinship is advantageous, whereas for an
-unrelated individual, concealing the lack of kinship is better. Similar
-scenarios may apply to many aspects of social life such as finding
-mates, feeding offspring, dominance relationships and aggressive
-interactions (Bradbury and Vehrencamp 2011; Møller 1988; Tibbetts and
-Dale 2004). In dominance and aggressive interactions, an important piece
-of information to guide individual actions is interacting partners
-fighting ability, which is sometimes referred to as Resource Holding
-Potential (RHP) (Parker 1974), or simply as quality. Responsiveness to
-the quality of the partner is central to communication systems such as
-badges of status, where an arbitrary signal conveys quality (Johnstone
-and Norris 1993; Rohwer 1975). Here, arbitrary refers to a signal which
-is not ontogenetically correlated with quality. Badges of status can be
-evolutionarily stable whenever the signal imposes a fitness cost which
-decreases with the quality of an individual (Botero et al. 2010;
-Johnstone and Norris 1993), such that it's no longer in the interest of
-low quality individuals to fake quality by producing a large badge. The
-cost of honest signals is a more general principle of communications
-systems, usually referred to as the handicap principle (Grafen 1990;
-Zahavi 1975).
+individuals influence their reproductive success. However, the best
+action for an individual engaging in a social interaction might vary
+depending on its own condition and that of individuals with whom it
+interacts. Thus, acting based on information about interacting partners
+is typically adaptive (Quiñones et al. 2016), yet acquiring such
+information is far from trivial. In some cases, interacting partners
+(\emph{i.e.} signallers) might be `willing' to provide accurate
+information, but in others it might be in their own interest to conceal
+information (Johnstone 1997) or to provide deceptive information
+(Johnstone and Norris 1993). Typically, in a given context, some
+individuals benefit may from broadcasting accurate information, whereas
+others may benefit from concealing it. Take, for instance, an
+interaction between two individuals where one can help the other. Given
+some costs and benefits, a potential donor is likely to be interested in
+helping related individuals. Therefore, for a relative of the donor,
+broadcasting kinship would be advantageous, whereas concealing the lack
+of kinship would be better for an unrelated individual. Similar
+scenarios may apply to a variety of aspects of social life such as
+finding mates, feeding offspring, or engaging in dominance relationships
+or aggressive interactions (Bradbury and Vehrencamp 2011; Møller 1988;
+Tibbetts and Dale 2004).
 
-A somewhat ignored component of communication systems, in the context of
-aggressive interactions, is the cognitive aspects of the receiver
-module. Theoretical models often assume that Communication systems
-relying on badges of status to have reaction norms as their mechanistic
-underpinning (Botero et al. 2010), with individuals using their own
-quality and the opponent's badge to determine whether to aggressively
-engage in contest. Reaction norms allow individuals to respond to the
-available information without a big cognitive burden. That is at least
-in contrast to other information processing mechanisms like individual
-recognition. With individual recognition individuals associate cue of
-their peers to their quality. These association must be learn thoughout
-individuals lifes, thus it has the usual cognitive requirements of an
-associative learning process. An alternative view of badges of status is
-that individuals learn to react to them based on their experiences
-(Guilford and Dawkins 1991), which would imply that receivers must learn
-to associate signals with the fighting ability of bearers just as in
-systems based on individual recognition. In cases where badges of status
-vary quantitatively (e.g.~in size or intensity), fighting ability may
+In dominance and aggressive interactions, a crucial piece of information
+to guide individual actions is the fighting ability of interacting
+partners, which is sometimes referred to as Resource Holding Potential
+(RHP, Parker 1974) or simply as quality. Responsiveness to the quality
+of an opponent is central to communication systems in which individuals
+exhibit badges of status, where a signal typically conveys quality
+(Johnstone and Norris 1993; Rohwer 1975). Such signals tend to be
+arbitrary, however, meaning the signals could potentially be produced by
+high- and low-quality individuals alike, opening up the possibility of
+dishonest signalling and raising the question of evolutionary stability.
+Signals can be evolutionarily stable whenever low-quality individuals
+pay a higher fitness cost for carrying the signal such that it is not in
+the interest of low-quality individuals to fake quality by producing a
+large badge. Signal costs inversely proportional to quality may come
+about as a result of the production costs of the signal, in which case
+the signal works like a handicap (Botero et al. 2010; Johnstone and
+Norris 1993; Grafen 1990; Zahavi 1975). Alternatively, the costs can
+come from the aggressive reaction of receivers when the convention of
+the communication is broken (Enquist, Ghirlanda, and Hurd 2010; Tibbetts
+and Dale 2004). For example in paper wasps, subordinate individuals with
+experimentally manipulated dominant-like facial patterns received more
+aggression from the dominant (Tibbetts and Dale 2004). In this case,
+costs are socially rather than developmentally mediated and are
+triggered by the violation of the convention, although how the
+convention is established in the first place is unclear.
+
+A somewhat ignored component of communication systems in the context of
+aggressive interactions are cognitive aspects of the receiver module.
+Theoretical models often assume that communication systems relying on
+badges of status have reaction norms as their mechanistic underpinning
+(Botero et al. 2010), with individuals using their own quality and the
+badges of opponents to determine whether to aggressively engage in
+contests. Reaction norms allow individuals to respond to the available
+information without a large cognitive burden because the computations
+involved in the reaction norm response do not require much memory nor
+advanced algorithmic processes. This contrasts with other
+information-processing mechanisms like individual recognition where
+individuals associate cues of their peers to their quality. Because
+these associations must be learnt throughout the life of individuals,
+they have the usual cognitive requirements of associative learning
+processes. An alternative view of badges of status is that individuals
+learn to react to them based on their experiences (Guilford and Dawkins
+1991), which would imply that receivers must learn to associate signals
+with the fighting ability of bearers just as in systems based on
+individual recognition. In cases where badges of status vary
+quantitatively (\emph{e.g.} in size or intensity), fighting ability may
 increase monotonically with attributes of the badge and would be
-reinforced by every interaction, so in principle learning would be
-faster than in systems involving individual recognition in which the
-association between signals and their meaning would vary depending on
-the interacting partners. In any case, learning could be a central
-cognitive mechanism in both types of communication systems, but the role
-of learning in these contexts has not been thoroughly explored in the
-empirical nor theoretical literature.
+reinforced by every interaction. Therefore, in principle, learning in
+such scenarios would be faster than in systems involving individual
+recognition in which the association between signals and their meaning
+varies depending on the interacting partners. In any case, learning may
+be a central cognitive mechanism in both types of communication systems,
+but the role of learning in these contexts has not been thoroughly
+explored in the empirical nor theoretical literature.
 
-Associative learning is a key cognitive mechanism that allows
-individuals to associate rewards with environmental stimuli and thus
-behave adaptively (Staddon 2016). Associative learning exists in all
-major vertebrate taxa, and in many invertebrates as well (Heyes 2012;
-Macphail 1982; Staddon 2016; Behrens et al. 2008). Theory has shown that
-natural selection favours these associations in complex environments
-where conditions are hard to predict (Dridi and Lehmann 2016). Besides
-its wide taxonomic and ecological relevance, associative learning is a
-flexible cognitive mechanism whose underpinnings show interspecific
-variation (Enquist, Lind, and Ghirlanda 2016; Quiñones et al. 2019), so
-presumably, it has been modified by natural selection. Despite all this,
-associative learning is not often included in the narrative of
-evolutionary explanations of behavioural patterns (Fawcett, Hamblin, and
-Giraldeau 2013; Kamil 1983; McAuliffe and Thornton 2015). Computational
-models of evolution can play an important role to overcome the lack of
-integration between learning and evolution. Reinforcement learning
-theory encompasses a series of computational methods inspired on the
+Associative learning is a taxonomically widespread cognitive mechanism
+that allows individuals to associate rewards with environmental stimuli
+and thus behave adaptively (Heyes 2012; Macphail 1982; Staddon 2016;
+Behrens et al. 2008). Theory has shown that natural selection favours
+such associations in complex environments where conditions are difficult
+to predict (Dridi and Lehmann 2016). Associative learning is a flexible
+cognitive mechanism whose underpinnings show interspecific variation
+(Enquist, Lind, and Ghirlanda 2016; Quiñones et al. 2020; Prétôt,
+Bshary, and Brosnan 2016; Prétôt et al. 2021). For example, species can
+have different learning rates (Hoedjes et al. 2010), some may vary these
+rates during learning trials which allows them to have more flexible
+learning (Leimar, Quiñones, and Bshary 2023), and some may include
+future rewards in their associations (Raby et al. 2007; Osvath 2009).
+Associative learning is not a single mechanism, but rather a set of
+cognitive structures and processes that can vary in their scope and
+complexity. Presumably, these structures and processes have been
+modified by natural selection and could provide novel explanations for
+behavioural variation. Nonetheless, associative learning is not often
+included in the narrative of evolutionary explanations of behavioural
+patterns (Fawcett, Hamblin, and Giraldeau 2013; Kamil 1983; McAuliffe
+and Thornton 2015).
+
+Computational models of evolution can overcome the lack of integration
+between learning and evolution. Reinforcement learning theory
+encompasses a series of computational methods inspired on the
 psychological and neurological mechanisms of associative learning
-(Sutton and Barto 2018). These set of algorithms allow the
+(Sutton and Barto 2018). This set of algorithms allows the
 implementation of biologically realistic problems, capturing the essence
 of learning processes (Frankenhuis, Panchanathan, and Barto 2018;
-Quiñones et al. 2019). Furthermore, these algorithms can be embedded in
+Quiñones et al. 2020). Furthermore, these algorithms can be embedded in
 evolutionary simulations to generate theoretical predictions of the
-effect that learning can have in behavioural evolution (Leimar and
-McNamara 2019).
+effect of learning on behavioural evolution (Leimar and McNamara 2019;
+Leimar and Bshary 2022b).
 
-In here we present an evolutionary model where individuals use
-associative learning to develop a tendency to behave aggressively (or
-peacefully), in the context of competition over resources, depending on
-the quantitative morphological trait they perceive in their opponent
-(badge). Over evolutionary time individuals evolve the size of their
-badge, and whether it depends on their quality. Under this simple set
-up, individuals can use the badge as a signal of quality. We use the
-model to assess under what conditions of interaction structure we expect
-different communication signals to evolve.
+Here, we present an evolutionary model where individuals use associative
+learning to develop a tendency to behave either aggressively or
+peacefully in the context of competition over resources, depending on a
+quantitative morphological trait (\emph{i.e.} a badge indicating
+quality) they perceive in their opponents. Over evolutionary time, the
+size of the badge evolves as does its dependency on the individual's
+quality. Under this simple set up, individuals can use the badge as a
+signal of quality. We use the model to assess under what conditions one
+might expect communication signals to evolve as handicaps or
+conventions.
 
@@
 
-We model the evolution of signals of quality in the context of agonistic
-interactions. For simplicity we consider a population of haploid
-individuals with non-overlapping generations. Individuals are born every
-generation with a level of quality (\(Q_i\), where \(i\) is subscript of
-the population vector of size \(N\)) given by a number between zero and
-one, which is drawn from a truncated normal distribution
-\(N(\mu = 0.5,\sigma)\). An individual with quality \(0\) has the lowest
-RHP, while an individual with quality \(1\) has the highest. As part of
-development processes individuals produce a morphological signal
-(badge), the size of such depends on their level of quality according to
-a reaction norm given by \(B_i=1/(1+e^{\alpha_i-\beta_i Q_i})\); where
-are \(\alpha_i\) and \(\beta_i\) are individual specific traits that
-determine the shape of the reaction norm (Fig. \ref{fig:model-struc} A).
-Variation in \(\alpha_i\) and \(\beta_i\) means individuals can have
-either uninformative (flat) or informative (logistic) norms. The size of
-the signal is constrained between zero and one. We assume different
-values of these traits are given by different alleles, and are inherited
-from mother to offspring. Unless, with a small probability, mutation
-changes the allelic value of the offspring. After birth individuals go
-through a round of viability selection. Individual specific survival
-probability is given by \(s_i=1/(1+e^{-k_1-k_2(Q_i-B_i)})\). Where
-\(k_1\) and \(k_2\) are parameter values determining the shape of the
-survival function. Importantly, if \(k_2=0\) survival probability is
-independent of quality, while if \(k_2>0\) lower quality individuals pay
-a higher price for similar-sized badge, fulfilling the core assumption
-of the handicap principle (Botero et al. 2010; Grafen 1990; Johnstone
-and Norris 1993).
+We model the evolution of signals indicating individual quality in the
+context of agonistic interactions. For simplicity, we consider a
+population of haploid individuals with non-overlapping generations.
+Individuals are born every generation with a level of quality (\(Q_i\),
+where \(i\) is subscript of the population vector of size \(N\)) given
+by a number between zero and one, which is drawn from a truncated normal
+distribution \(N( 0.5,\sigma)\). An individual with quality \(0\) has
+the lowest RHP, while an individual with quality \(1\) has the highest.
+As they develop, individuals produce a phenotypic signal (badge), the
+size of which depends on their quality according to a reaction norm
+given by \(B_i=1/(1+e^{\alpha_i-\beta_i Q_i})\); where \(\alpha_i\) and
+\(\beta_i\) are individual specific traits that determine the shape of
+the reaction norm (Fig. \ref{fig:modelStruc} A). Variation in
+\(\alpha_i\) and \(\beta_i\) means individuals can have either
+uninformative (flat) or informative (logistic) norms. Informative
+reaction norms represent a developmental program where the phenotype of
+the individual is determined by the environmental conditions under which
+it grows, thus causing a correlation between quality and signal. The
+size of the signal is constrained to take values between zero and one.
+We assume different values of these traits are given by different
+alleles and are inherited from mother to offspring unless, with a small
+probability (\(\mu\)), mutation changes the allelic value of the
+offspring by an amount drawn from a normal distribution
+\(TN(0,\sigma_\mu)\).
+
+After birth, individuals go through a round of viability selection. The
+survival probability of an individual is given by
+\(s_i=1/(1+e^{-k_1-k_2(Q_i-B_i)})\), where \(k_1\) and \(k_2\) are
+parameter values determining the shape of the survival function.
+Importantly, if \(k_2=0\) survival probability is independent of
+quality, while if \(k_2>0\) lower-quality individuals pay a higher price
+for similar-sized badges, fulfilling the core assumption of the handicap
+principle (Botero et al. 2010; Grafen 1990; Johnstone and Norris 1993).
 
 Individuals who survive engage in a series of pairwise interactions
-where they compete for resources. Individuals in real populations
-typically do not interact at random nor with all individuals in the
-population (Kurvers et al. 2014). In order to allow for non-random
-interactions, we use the population vector to define an interaction
-neighbourhood for the focal individual, the size of the neighbourhood
-(\(g\)) is the number of positions around the focal's from which the
-partner is drawn. So, if \(g=N\) interactions are global; as \(g\) gets
-smaller interactions are more local. In an interaction each individual
-must decide whether to escalate a fight or not. Following the classic
+where they compete for resources. In each interaction individuals must
+decide whether to escalate a fight or not. Following the classic
 \emph{hawk-dove} game (Maynard-Smith 1982), if the focal individual
-fights and its partner does not, the focal gets as pay-off the resource
-of value \(V\) while its partner gets nothing; if both individuals
-restrain from fighting they split up the resource in half; if both
-decide to fight they split up the cost of fighting and the winner takes
-over the resource. We further assume that the probability of wining a
-fight for the focal individual depends on the difference in quality
-between her and her interacting partner; specifically it is given by
+fights and its partner does not, then the focal gets as pay-off the
+resource of value \(V\) while its partner gets nothing; if both
+individuals restrain from fighting, then they split up the resource in
+half. If both individuals decide to fight, then the winner takes over
+the resource and the cost of fighting is split between the two players.
+We further assume that the probability of wining a fight for the focal
+individual depends on the difference in quality between it and its
+interacting partner; specifically it is given by
 \(p_{ij}=1/(1+e^{-k_3(Q_i-Q_j)})\), where \(k_3\) is a parameter
 defining how strong the quality difference determines the winning
-probability; and \(i\) and \(j\) denote the the position in the
-population vector of the focal and its interacting partner respectively.
+probability, and \(i\) and \(j\) denote the the position in the
+population vector of the focal and its interacting partner,
+respectively.
 
 The decision of whether to escalate a fight against an interacting
-partner can be dependent on the size of the partner's badge, the actual
-dependency is determined by the focal´s experiences through a learning
-process. We implement learning using the actor-critic approach from
-reinforcement learning theory {[}RL; Sutton and Barto (2018);Quiñones et
-al. (2019); Leimar and McNamara (2019){]}. Individuals estimate the
-reward (pay-off) expected from interacting with partners of different
-badge sizes (the critic in RL terminology). After every interaction they
-update the estimate of reward proportionally to the difference between
-their current estimate and the observed reward (prediction error
-\(\delta\)) and to the speed of learning (\(\mathrm{A}\)). Furthermore,
-they express different probabilities of retreating (attacking) depending
-on the badge size of their opponent (the actor in RL). They update the
-probability of retreating (attacking) up or down depending on whether
-retreating (attacking) leads to an increase in the reward estimation.
-So, if a focal individual decides to escalate a fight against an
-individual with small badge, and that leads to an increase in the reward
-estimation, the focal individual will increase the probability of
-escalating fights with individuals of small badges in the future. Given
-that badge size is a real number between 0 and 1, there are infinitely
-many badge sizes. Thus, the reward estimation, as well as the
-probability of retreating (escalating), must be generalized across
-different values. To implement generalization we use the linear function
-approximation method based on radial basis functions (Sutton and Barto
-2018). Specifically, we pick \(c\) feature centres, which are evenly
-spaced values along the badge size interval ({[}0,1{]}). Each one of
-these features centres is associated with a weight for reward estimation
-and tendency to play retreat in a given interaction. The reward
-estimation and the tendency to retreat are calculated (in every
-interaction) as the sum of the weights associated with each feature,
-weighted by the response triggered by the feature (Fig.
-\ref{fig:model-struc} B, dots represent the feature weights) The
+partner can depend on the size of the partner's badge, with the
+dependency being determined by the focal's experiences acquired through
+a learning process. We implement learning using the actor-critic
+approach from reinforcement learning (RL) theory (Sutton and Barto 2018;
+Quiñones et al. 2020; Leimar and McNamara 2019; Leimar 2021).
+Individuals estimate the reward (pay-off) expected from interacting with
+partners of different badge sizes (the critic in RL terminology). After
+each interaction they update the estimate of reward proportionally to
+the difference between their current estimate and the observed reward
+(prediction error \(\delta\)) and to the speed of learning
+(\(\mathrm{A}\)). Furthermore, individuals express different
+probabilities of retreating/attacking depending on the badge size of
+their opponent (the actor in RL). They update the probability of
+retreating/attacking depending on whether retreating/attacking leads to
+an increase in the reward estimation. Thus, if a focal individual
+decides to escalate a fight against an individual with a small badge and
+this leads to an increase in the reward estimation, then the focal
+individual will increase the probability of escalating fights with
+individuals of small badges in the future. Given that badge size is a
+real number between 0 and 1, there are infinitely many badge sizes.
+Thus, the reward estimation, as well as the probability of
+retreating/attacking, must be generalized across different values. To
+implement generalization we use the linear function approximation method
+based on radial basis functions (Sutton and Barto 2018). Specifically,
+we pick \(c\) feature centres, which are evenly spaced values along the
+badge size interval ({[}0,1{]}) where the updates are focused. For
+simplicity, we keep the location of these feature centres constant and
+the same for all individuals; they are stored in vector \(\symbf{b}\) .
+Each of these feature centres is associated with a weight for reward
+estimation and tendency to play retreat in a given interaction. The
+reward estimation and the tendency to retreat are calculated (in every
+interaction) as the sum of the weights associated with each feature
+centre, weighted by the response triggered by the feature (Fig.
+\ref{fig:modelStruc} B, dots represent the feature weights). The
 response of each feature centre diminishes as a Gaussian function with
 the distance between the feature centre and the badge size of the
-partner (Fig. \ref{fig:model-struc} B, grey line). Formally the reward
+partner (Fig. \ref{fig:modelStruc} B, grey line). Formally, the reward
 estimation \(\hat{R}\) when the focal individual (\(i\)) faces
@@
 weights associated with the actor, and the probability is obtained by
-applying a logistic transformation (Fig. \ref{fig:model-struc} B, black
-line). So formally, the log-odds to retreat when facing individual \(j\)
-is
+applying a logistic transformation (Fig. \ref{fig:modelStruc} B, black
+line). Thus formally, the log-odds to retreat when facing individual
+\(j\) is
 
 \begin{equation}
-logit(q_{ij}) =  \sum_{z=1}^{c} y_z e^{(-\frac{|B_i-c_z|}{2\theta^2})} 
+logit(q_{ij}) =  \sum_{z=1}^{c} y_z e^{(-\frac{|B_i-b_z|}{2\theta^2})} 
 \end{equation}
@@
 
-After all their interactions, individuals in the population reproduce
-proportionally to their fitness \(w_i\), which is a sum of the baseline
-fitness (\(w_0\)) and all the pay-offs obtained throughout their life.
-Thus, the combination of natural selection and genetic drift changes the
-distribution of values in \(\alpha\) and \(\beta\) that segregates in
-the population, effectively changing the badges expressed and the
-communication system.
+Individuals interact as the focal individual \(n\) times and the
+interaction partner is chosen at random. Thus, the expected number of
+interactions for each individual is \(n+\frac{n(N+1)}{N}\). After the
+interaction round , individuals in the population reproduce with a
+probability proportional to their total pay-off \(w_i\), which is a sum
+of the baseline pay-off (\(w_0\)) and all the pay-offs obtained
+throughout their life. Thus, the combination of natural selection and
+genetic drift changes the distribution of values in \(\alpha\) and
+\(\beta\) that segregates in the population, effectively changing the
+badges expressed and the communication system.
 
@@
 
-{\centering \includegraphics{manuscript_1.0_files/figure-latex/model-struc-1}
+{\centering \includegraphics[width=1\linewidth]{manuscript_1.0_files/figure-latex/modelStruc-1}
 
@@
 \caption{Model of communication in the context of aggressive interactions.
-In A the sender code: a reaction norm determines how the badge size is determined by the quality of the individual.
-Red shows an informative reaction norm, while the blue shows a uninformative reaction norm.
-B the receiver code: individuals have a behavioral reaction norm that determines their probability of retreating (black line).
+In A, the reaction norm determines the badge size, *e.g* the size of a black band on a bird's chest.
+Red shows an informative reaction norm, whereas blue shows a uninformative reaction norm.
+B individuals perceiving the signal have a behavioral reaction norm that determines their probability of retreating (black line).
 The reaction norm arises from generalizing the information from the feature weights (black dots).
-Generalization is represented by the grey line and its axis, which shows the response triggered by the fourth feature diminishes as the value evaluated is further from the feature center.
-The learning process moves the feature weights (black dots) up or down depending on whether that leads to an increase in the estimated reward.
-In C and D, effect of learning on the receiver strategy.
-Receivers in C face signallers with uninformative reaction norms (blue line in A), while in D they face signallers with informative reaction norms (red line in A).
-Colour scale in C (applies also for D) indicates the quality of the individual}\label{fig:model-struc}
+Feature weights are values associated with a particular badge size and are updated as individuals interact with each other.
+The generalization of the information associated with each feature weight is represented by the grey line and its axis, which shows that the response triggered by the fourth feature diminishes as the value evaluated is further from the feature center.
+The learning process moves the feature weights (black dots) up or down depending on whether this leads to an increase in the estimated reward.
+Panels C and D show the effect of learning on the receiver strategy.
+Receivers in C face signallers with uninformative reaction norms (blue line in A).
+The uninformative reaction norm yields a badge size of 0.5, thus the behavioural reaction norm in C is only updated around that value.
+D Receivers face signallers with informative reaction norms (red line in A).
+Accordingly, receivers develop a threashold-like reaction norm where the decision to retreat increases with the badge size of the interacting partner.
+Colour scale in C and D indicates the quality of the individual.}\label{fig:modelStruc}
 \end{figure}
@@
 
-We first present the outcome of simulations where we prevent the
-evolution of the sender code (by setting the mutation rate to \(0\)),
-and assume that all individuals in the population display either an
+\hypertarget{learning-in-a-monomorphic-population}{%
+\subsubsection{Learning in a monomorphic
+population}\label{learning-in-a-monomorphic-population}}
+
+We first present the outcome of simulations where we prevented the
+evolution of the badge size (by setting the mutation rate to \(0\)), and
+assumed that all individuals in the population display either an
 uninformative or informative badge size with respect to their quality.
 These simulations show what type of receiver strategy develops through a
-learning process. Figure \ref{fig:model-struc} (C and D) shows the
-receiver strategy developed through learning; panel C is for receivers
-that faced uninformative signals, and D informative ones. When learners
-face uninformative signals, they change their probability of retreating
-depending on their own quality. Individuals with higher quality (red
-tones), after the learning process, have a high probability of
-attacking; while individuals of lower quality (blue tones) mostly
-retreat from confrontations. Thus, learning splits the population of
-receivers into the two classic pure strategies of hawks and doves. Given
-that we have a assumed a monomorphic population with unresponsive
+learning process in such a monomorphic population. Figure
+\ref{fig:modelStruc} (C and D) shows the receiver strategy developed
+through learning; panel C is for receivers that faced uninformative
+signals, and D for those facing informative ones. The simulations reveal
+that when learners face uninformative signals (Fig. \ref{fig:modelStruc}
+C), they modify their probability of retreating depending on their own
+quality. Individuals with higher quality (red tones) have a high
+probability of attacking after the learning process\\
+while individuals of lower quality (blue tones) mostly retreat from
+confrontations; in both cases, individuals behave equally irrespective
+of the quality of their opponent. Thus, learning splits the population
+of receivers into the two classic pure strategies of hawks and doves.
+Given that we assumed a monomorphic population with unresponsive
 reaction norms on the signalling side, the changes triggered by learning
-only affect a small range of badge sizes. In contrast, when receivers
-face informative reaction norms on the side of the signaller (panel B in
-fig.~\ref{fig:model-struc}), receivers use the badge size of their
-interacting partners to determine whether to retreat or attack. The
-relationship is given by a threshold-like reaction norm, where the
-change from retreating to attacking depends on the quality of the
-receiver. As expected, the higher the quality of the receiver the larger
-the badge size that triggers a retreat.
-
-\begin{verbatim}
-## Warning in knitr::include_graphics(here("Simulations", "betCostEvol1_", :
-## It is highly recommended to use relative paths for images. You had
-## absolute paths: "M:/Projects/SocialComp_morphCue/Simulations/betCostEvol1_/
-## evolDyn0_betCost5.png"
-\end{verbatim}
+only affected a small range of badge sizes. Specifically, all learning
+occurred around a badge size of \(0.5\), because all badges in the
+population are of this size (panel C). In contrast, when receivers face
+informative reaction norms on the side of the signaller (Fig.
+\ref{fig:modelStruc} B), receivers use information on the badge size of
+their interacting partners to determine whether to retreat or attack.
+The relationship is given by a threshold-like reaction norm, where the
+decision to retreat or to attack depends on the quality of the receiver.
+As expected, the higher the quality of the receiver, the larger the
+badge size of the signaller that triggers a retreat.
 
@@
 
-{\centering \includegraphics[width=0.9\linewidth]{M:/Projects/SocialComp_morphCue/Simulations/betCostEvol1_/evolDyn0_betCost5}
+{\centering \includegraphics[width=1\linewidth]{manuscript_1.0_files/figure-latex/handicap-1}
 
@@
 \caption{Evolution of the badge size as a handicap mediated by learning.
-Middle panels show the evolutionary dynamics of the sender code.
-On the left, changes in the distribution of values for the intercept of the reaction norm ($\alpha$); on the right changes in the distribution of the slope ($\beta$).
-Dashed lines in both panels show the mean of the distribution.
-Grey lines show the generation time corresponding to the panels above and below showing the sender and receiver code.
-In the panels above, the reaction norms correspond to individuals in the population after the interaction round.
+Top panels (A.1 and A.2) show the evolutionary dynamics of the sender code.
+On the left (A.1), changes in the distribution of values for the intercept of the reaction norm ($\alpha$); on the right (A.2) changes in the distribution of the slope ($\beta$).
+Darker areas of the background correspond to values with high frequency, while orange lines in both panels show the mean of the distribution.
+X axis in the evolutionary dynamics are given in thousands of generations.
+Grey lines show the generation time corresponding to the panels below portraing the sender (B) and receiver code(C).
+In the bottom panels (C1-4), the learned reaction norms correspond to individuals in the population after the interaction round.
 Color scale indicates quality just as in Fig.
-\ref{fig:model-struc}.
-In the panels below, colors represent a cluster clasification performed by the k-means algorithm}\label{fig:handicap}
+\ref{fig:modelStruc}.
+On the first half of the simulation (B.1 and B.2) the badge size evolves to its minimum value.
+In the second half, the slope takes positive values and reaction norms evolve to produce an honest signal of quality.
+Correspondingly, individuals learn to react to badge size by increasing the probability of retreating with increasing badge sizes}\label{fig:handicap}
 \end{figure}
 
-When we allow the evolution of the sender code (\(\alpha\) and \(\beta\)
-change subject to natural selection and genetic drift), and the badge
-size work as a handicap (i.e.~the cost of the badge is inversely
-proportional to the quality of the individual), the sender code evolves
-to produce an honest signal Fig \ref{fig:handicap}). The evolution of
-the sender code does not happen immediately after the start of the
-evolutionary process. Instead, the evolutionary dynamics of the reaction
-norm parameters (Fig. \ref{fig:handicap}, X) show a set of steps. First,
-the intercept of the reaction norm (\(\alpha\)) evolves to higher
-values, reducing the average badge size in the populations. That makes
-sense because reaction norms segregating in the population are flat at
-this point, and consequently the receiver code does not respond to the
-badge size, bigger badges are costlier and do not trigger lower attack
-probabilities. During those first generations, the slope of the sender
-reaction norm (\(\beta\)) remains around the initial value of zero. At
-around generation 4000, the slope evolves toward positive values. Which
-implies that larger badges correlate with higher quality. Receivers
-learn to react to such correlation, reducing the probability of attack
-towards individual of larger badges. Hence, natural selection favours
-larger values of \(\beta\), eventually leading to an evolutionary
-equilibrium where badge size is an honest signal of quality mediated by
-the learned responses of receivers. The evolutionary trajectory
-portrayed in figure \ref{fig:handicap} is not the only possible
-trajectory. If the slope of the sender reaction norms (\(\beta\))
-evolves toward negative values before badges become handicaps, receivers
-never learn to react to the size of badges. Thus, badges are only costly
-and do not provide information about the quality of an individual(Fig.
-\ref{fig:noBadge} . Eventually, the badge disappears from the
-population.
+\hypertarget{badges-as-handicaps}{%
+\subsubsection{Badges as Handicaps}\label{badges-as-handicaps}}
 
-When production of the badge is not costly (i.e.~does not work as a
-handicap), the parameters of the sender code go through evolutionary
-branching event, triggering the evolution of different types of signals
-in the population. Fig. \ref{fig:branching}A and B shows changes in the
-distribution of the sender reaction norm (\(\alpha\) and \(\beta\))
-along evolutionary time. Darker areas show trait values with high
-frequency in the population. Populations start monomorphic with a value
-of zero on both traits, and mutations quickly build up a normal
-distribution around the starting value. Within the first 2000
-generations the unimodal distribution splits into a bimodal one. In the
-case of \(\alpha\), one of the peaks splits further, so at the end of
-the evolutionary simulation the distribution of \(\alpha\) value in the
-population shows three peaks. In the case of \(\beta\), the two initial
-peaks coexist for around 4000 generations. Later on one of the types
-goes extinct and the distribution goes back to being unimodal. These
-changes in the distribution of the sender reaction norm parameters imply
-that during the second half of the simulation individuals can be
-classified into three distinct types. Two types express a flat reaction
-norm, meaning that their badge size is not informative of their quality,
-while the third type shows intermediate badge sizes which are determined
-by the quality of the individual (Fig. \ref{fig:branching}).
-Furthermore, the receiver reaction norms develop through learning,
-particularly those of individuals of intermediate quality, respond to
-the signal of their sender type by increasing the probability of
-retreating from a fight with individuals with larger badges (Fig.
+When we allowed the badge size to evolve (\(\alpha\) and \(\beta\)
+changed subject to natural selection and genetic drift) and the signal
+worked as a handicap (\emph{i.e.} the cost of the badge was inversely
+proportional to the quality of the individual), the sender code often
+evolved to produce an honest signal (Fig. \ref{fig:handicap}). The
+evolution of the sender code did not happen immediately after the start
+of the evolutionary process. Instead, the evolutionary dynamics of the
+reaction norm parameters (Fig. \ref{fig:handicap}, A) appeared to
+involve a set of steps. First, the intercept of the reaction norm
+(\(\alpha\)) describing the relationship between individual quality and
+badge size evolved to higher values, reducing the average badge size in
+the populations. This is because reaction norms segregating in the
+population are flat at the beginning of the simulation; consequently,
+the receiver code does not respond to the badge size and larger badges
+are costlier and do not trigger lower attack probabilities. During the
+first generations, the slope of the sender reaction norm (\(\beta\))
+remains close to the initial value of zero. At around generation 4000,
+the slope evolves toward positive values. Positive values in the slope
+imply that larger badges correlate with higher quality (Fig.
+\ref{fig:handicap} B). Receivers then learn to react to such
+correlation, increasing the probability of retreating when facing
+individuals with larger badges (Fig. \ref{fig:handicap} C). Hence,
+natural selection favours larger values of \(\beta\), eventually leading
+to an evolutionary equilibrium in which badge size is an honest signal
+of quality mediated by the learned responses of receivers (Fig.
+\ref{fig:handicap} B.4 and C.4).
+
+The evolutionary trajectory portrayed in figure \ref{fig:handicap} A is
+not the only possible outcome. If the slope of the sender reaction norms
+(\(\beta\)) evolves, subject to genetic drift, toward negative values
+before badges become handicaps, receivers never learn to react to the
+size of badges. Thus, badges are only costly and do not provide
+information about the quality of individuals (Fig. \ref{fig:noBadge}) .
+Eventually, therefore, badges disappear from the population. In
+contrast, when the badge does not work as a handicap but instead is
+cost-free, the evolutionary process never leads to the establishment of
+an honest signal. Instead, subject to genetic drift, the population
+either evolves towards the disappearance of the badge or to the maximum
+badge size. In either of these cases reaction norms are flat, so the
+badge does not provide any information about quality.
+
+\hypertarget{the-evolution-of-polymorphism-mediated-by-learning}{%
+\subsubsection{The evolution of polymorphism mediated by
+learning}\label{the-evolution-of-polymorphism-mediated-by-learning}}
+
+The amount of information that agents are able to collect via learning
+over their life time can strongly change the outcome of evolutionary
+dynamics. In the simulations presented in figure \ref{fig:handicap} and
+\ref{fig:noBadge}, individuals learned with high speed
+(\(\mathrm{A} = 0.4\)) and interacted repeatedly along their lifetime
+(2000 interactions on average). When we reduced the number of
+interactions that individuals experienced over their lifetime to 300 on
+average, we saw a drastic increase in the phenotypic and genetic
+variation present in the population. Populations start monomorphic with
+a value of zero for both the intercept and the slope of the reaction
+norm, and mutations quickly build up a normal distribution around the
+starting value. Within the first 2000 generations, the unimodal
+distribution in the intercept (\(\alpha\)) splits into a bimodal one.
+Later in evolutionary time, one of the peaks splits further, so at the
+end of the evolutionary simulation the distribution of the intercept
+(\(\alpha\)) in the population shows three distinct peaks. In the case
+of \(\beta\), the peaks in the distribution are not so clear-cut, but
+the variance of the distribution increases over time. These changes in
+the distribution of the parameters of the sender reaction norm imply
+that individuals can generally be classified into three distinct types
+(Fig. \ref{fig:branching}). Two types express a flat reaction norm with
+extreme values for the badge size, meaning that their badge size is not
+informative of their quality, whereas the third type shows intermediate
+badge sizes determined by individual quality (Fig. \ref{fig:branching}
+B.3-4). In this simulation, the values of the slope in the informative
+group are negative, which implies that badge size correlates negatively
+with quality. Furthermore, the receiver reaction norms developed through
+learning, particularly those of individuals of intermediate quality,
+respond to the signal of their sender type by increasing the probability
+of retreating from a fight with individuals with smaller badges (Fig.
 \ref{fig:branching}).
-
-\begin{verbatim}
-## Warning in knitr::include_graphics(here("Simulations", "nIntGroupNormQual_", :
-## It is highly recommended to use relative paths for images. You had absolute
-## paths: "M:/Projects/SocialComp_morphCue/Simulations/nIntGroupNormQual_/
-## evolDyn7_nIntGroup2000.png"
-\end{verbatim}
 
@@
 
-{\centering \includegraphics[width=0.9\linewidth]{M:/Projects/SocialComp_morphCue/Simulations/nIntGroupNormQual_/evolDyn7_nIntGroup2000}
+{\centering \includegraphics[width=1\linewidth]{manuscript_1.0_files/figure-latex/branching-1}
 
@@
 \ref{fig:handicap}.
-The middle panels show changes in the distribution of values fir $\alpha$ and $\beta$ along evolutionary time.
-Panels above and bellow correspond to snapshots of the sender and receiver codes, respectively, generation time of the snapshots are indicated by the grey lines in the middle panels.}\label{fig:branching}
+The top panels (A) show changes in the distribution of values for $\alpha$ (A.1) and $\beta$ (A.2) along evolutionary time.
+Panels below correspond to snapshots of the sender (B) and receiver codes (C).
+Generation time of the snapshots are indicated by the grey lines in the top panels.
+The unimodal distribution with which $\alpha$ (A.1) starts the simulation, quickly turns into a bimondal distribution, and at about 10000 generations it splits further into three modes.
+These three peaks correspond to three type of reaction norms in panels B.1-3.}\label{fig:branching}
 \end{figure}
+
+A limited number of interactions has an effect on the amount of
+variation in the evolving parameters when the signal follows the
+handicap principle as well. In fig.~\ref{fig:honestDiv}, we show
+simulations where individuals have on average 300 interactions in their
+life and the cost of the signal is proportional to quality, following
+the handicap principle. The evolutionary process leads to a combination
+of reaction norm parameters where there is a positive correlation
+between the size of the badge and the quality of the individual. This
+relationship, however, is muddled by the fact that there are two
+clusters of values for the intercept (\(\alpha\)) and slope (\(\beta\))
+of the reaction norm in the population. Thus, there are two types of
+reaction norms. One of the types has a steeper slope compared to the
+other. This effect of an increased variance in the trait distribution is
+not only triggered by lower number of interactions. Larger variances are
+also found when we assume a lower speed of learning (data not shown).
+This suggest that limits to the amount of information that individuals
+acquire through learning allows the coexistence of different
+communication strategies within a population, something previously shown
+by Botero et al. (2010).
 
@@
 
-{\centering \includegraphics[width=0.9\linewidth]{../Simulations/initAct_/indVarScatter_}
+{\centering \includegraphics[width=1\linewidth]{manuscript_1.0_files/figure-latex/honestDiv-1}
 
@@
 
-\caption{WHy would you want to be recognized}\label{fig:startCond}
+\caption{The evolution of costly signals.
+Portrait of the evolutionary dynamics of the sender code with snapsots of both sender and receiver codes just as in fig.
+\ref{fig:handicap}.
+The middle panels show changes in the distribution of values for $\alpha$ and $\beta$ along evolutionary time.
+Panels above and below correspond to snapshots of the sender and receiver codes, respectively, generation time of the snapshots are indicated by the grey lines in the middle panels.
+Similarly to the evolutionary dynamics in Fig.
+\ref{fig:handicap}, first the badge size evolves towards its minimum value.
+Then, the slope evolves positive values determining an increasing reaction norm.
+In contrast to \ref{fig:handicap}, the normal distribution splits up into two modes.
+This translates into two types of reaction norms assorting in the population where both code for a positive relation between quality and badge size.}\label{fig:honestDiv}
 \end{figure}
+
+\hypertarget{the-peaceful-the-aggressive-and-the-clever}{%
+\subsubsection{The peaceful, the aggressive and the
+clever}\label{the-peaceful-the-aggressive-and-the-clever}}
+
+Our model revealed that the behaviour expressed by naive individuals
+(\emph{i.e.} those who have not yet learned) imposes negative-frequency
+dependent selection, allowing for the build up of genetic variation in
+reaction norms. In the simulations presented so far, we assumed that
+individuals start with a flat behavioural reaction norm such that they
+escalate fights aggressively with a \(0.5\) probability regardless of
+the badge size of the interacting partner. To assess whether such
+initial conditions of the communication system had any role in the build
+up of genetic variation, we ran a series of simulations varying the
+initial conditions of the actor module in the learning model.
+Specifically, we let naive individuals have a flat reaction norm with
+either a 1) low (peaceful), 2) high (aggressive) probability of
+escalating a fight or 3) a probability corresponding to the ESS of the
+original hawk-dove game (which we call clever, due to information and
+computation necessary to know the ESS). Fig. \ref{fig:startCond} shows
+the distribution of values of the intercept (\(\alpha\)) and slope
+(\(\beta\)) evolved in different replicates of the simulations. The
+left-hand side panel, corresponding to peaceful naive individuals, is
+the only one where the distribution of values is split in different
+clusters. That is, most of the variation occurs within clusters. In
+contrast, when naive individuals behave either aggressively or cleverly,
+badges evolve toward minimum and maximum values, and so the variation is
+driven mainly by differences among replicates. We can make sense of
+these results by realizing that individuals change their naive behaviour
+quickly in ranges of the badge size that are common in the population.
+In contrast, an individual expressing a rare badge size will most likely
+experience the naive behaviour of their partner. When the naive
+behaviour is peaceful, individuals with a rare badge size have a fitness
+advantage. This triggers negative-frequency dependent selection and the
+evolution of different types of badge sizes. According to this
+narrative, situations where individuals learn fast and interact
+repeatedly will diminish the strength of frequency-dependent selection.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{manuscript_1.0_files/figure-latex/startCond-1}
+
+}
+
+\caption{The peaceful, the aggressive and the clever.
+Distribution of values of the intercept $\alpha$ and slope $\beta$ for individuals at the end of the evolutionary simulations.
+The panels show the three different initial conditions for the behaviour of individuals: peaceful, agressive and  "clever", see maintext for details.
+Colours indicate the replicate simulation.
+In the inset, the resulting reaction norms corresponding to the intecept and slope values for one of the replicates, which replicate is shown is indicated by the line colour.
+The simulations where individual disply a "clever" and agressive behaviour, before the learning process, show all the points of a single replicate cluster together, while in simulation where individuals start being peaceful al replicates show distinct groups of individuals spread throughout the x and y axis.
+Only in this condition there is a frequency-dependent process that favour diversity in the reaction norm.}\label{fig:startCond}
+\end{figure}
+
+The action of natural selection and genetic drift on the learning
+parameters do not override the conditions for the evolutionary branching
+(Dieckmann and Law 1996) described above. Given how crucial naive
+behaviour is for the branching event described earlier, we tested
+whether evolutionary processes could lead key parameters of the learning
+process toward values that prevented the genetic variation to build up.
+Figure \ref{fig:learnEvol} shows the dynamics from a set of simulations
+where we allow for the coevolution of the reaction norm as well as key
+parameters of the learning module, namely the speed of learning (\(A\))
+and the behavioural tendency expressed by individuals prior to learning.
+Top panels in figure \ref{fig:learnEvol} show that there is
+diversification process in this coevolutionary scenario as well. That is
+because, as is shown in the middle panels (B) of the same figure, the
+learning parameters stay within the range necessary for the branching
+event to take place. The speed of learning (Fig. \ref{fig:learnEvol}
+B.1) stays well above zero throughout the simulation. In contrast, the
+initial behavioural tendency maintain values around zero. Thus,
+evolutionary processes acting on the learning parameters did not
+override the conditions necessary for frequency-dependent selection to
+boost genetic and phenotypic variation.
+
+\hypertarget{discussion}{%
+\subsection{Discussion}\label{discussion}}
+
+We presented a model that integrates learning with evolutionary
+processes. In the model, the response to a signal in a communication
+system is mediated by learning processes: individuals learn how to
+respond to a quantitative trait of their interactive partners. We showed
+that the learning process can mediate the evolution of an honest signal
+under the handicap principle (Grafen 1990; Zahavi 1975). This conclusion
+is not different from classical genetic models, where responses to
+signals are innate to the individual. However, unlike classical genetic
+models, our model shows that learning can also mediate the origin of a
+signal polymorphism in the absence of the handicap principle. Under this
+polymorphic equilibrium the population consist of distinct types of
+individuals, with honest and dishonest signals. The amount of
+information gathered through learning, as a well as the initial
+conditions of the learning process, were crucial for the emergence and
+maintenance of variation in the signalling system. Namely the
+polymorphism required a limit in the amount of information collected by
+individuals, as well as peaceful behaviour in naive individuals.
+Furthermore, we showed that the conditions leading to the origin of the
+polymorphism can be reached when evolutionary processes are able to
+change the parameters of the learning system (the speed of learning and
+the naive behaviour).
+
+Associative learning is a powerful mechanism to learn about the world,
+about social partners, and about one's abilities in a particular social
+context. By associating cues and signals with fitness-relevant outcomes,
+individuals collect information allowing them to improve their
+reproductive potential. This is evident in classical situations such as
+when animals learn to avoid food that makes them sick or use
+environmental cues to find food. More recent theoretical work has
+highlighted the potential role that associations can have in social
+contexts such as hierarchy formation (Leimar 2021; Leimar and Bshary
+2022a, 2022b) and cooperation (Leimar and McNamara 2019; Dridi and Akçay
+2018). In these examples, models involve individuals who use various
+sources of reward to make adaptive choices. Here we extended this logic
+to the evolution of a communication system mediated by a signal of
+quality. Previous evolutionary models of badges of status assumed that
+individuals responded using a behavioural reaction norm, where the
+opponent's badge and the individual's own quality determined behaviour
+(Botero et al. 2010). However, there is no clear mechanism justifying
+the assumption that individuals inherently know their own quality,
+particularly relative to their peers. In our model, individuals not only
+learn about the quality signal, but also learn about their own quality
+relative to others. This can be seen in our simulations in the responses
+developed by individuals of different quality. Because individuals learn
+about their own quality, even in the absence of an honest signal, they
+are able to make more adaptive choices. Specifically, lower quality
+individuals refrained more from escalating fights, while higher-quality
+individuals took over resources they could win in a fight. Thus, the
+learning process we modelled and the response it mediates has
+fitness-relevant consequences even in the absence of an honest signal.
+This was confirmed by simulations where the speed of learning was
+allowed to change subject to evolutionary processes. In those
+simulations, natural selection always maintained learning rates above
+zero. Our simulations however, only captured the effect of selection
+mediated by the social game on the learning parameters. For example, we
+did not account for potential costs of faster updating, or improved
+performance in foraging tasks. Inter-specific variation in cognitive
+abilities mediated by environmental differences has been reported
+elsewhere (Sonnenberg et al. 2019). If such variation is partly mediated
+by changes that affect general cognitive processes, like the speed of
+learning, they could impact the outcome of communication systems like
+the one we modelled.
+
+Learning processes which involve collecting information on a
+population-wide level promote frequency-dependent selection. For
+example, frequency-dependence triggered by learning process was
+highlighted by a classic study where live predators drive the evolution
+of \emph{in silico} polymorphic prey (Bond and Kamil 2002). The key to
+the evolution of prey polymorphism in such system is that predators are
+better able to discriminate prey from the background when prey are
+frequently encountered. Thus, prey morphs found in low frequency in a
+population have a selective advantage. An analogous process emerged in
+our simulations, in which the receiver individual learned the
+appropriate response towards individuals with a common badge size in the
+population. When an individual has a rare badge size it can be favoured
+or unfavoured by the naive behaviour. When the naive behaviour is
+peaceful, polymorphism is promoted, whereas when the naive behaviour is
+aggressive polymorphism is prevented. These two situations show how
+behaviours dependent on learning processes seem to generally trigger
+frequency-dependent selection because the learning algorithm collects
+more information on the more frequent values of the the trait
+distribution. For simplicity, we modelled a single dimension (badge
+size), yet the frequency-dependent effect of learning could potentially
+be more important when individuals learn from multidimensional signals.
+
+Originally, Rohwer (1975) proposed the idea that certain phenotypic
+traits could be used by animals as status signals or signals of quality.
+This idea has been tested repeatedly on the dark patches of some bird
+species, such as the bibs exhibited by species in various families,
+particularly sparrows. So, are bibs real signals of quality? If so, are
+they handicaps or badge of status? From an evolutionary perspective it
+is unclear how the stability of status signals can be maintained. One
+option is that honesty is maintained by the cost of signal production;
+that is, when the signal works as a handicap (Botero et al. 2010; Grafen
+1990; Johnstone and Norris 1993). Alternatively, if the signal does not
+carry production costs inversely proportional to quality, then it may
+work as a convention. In this latter case, honesty is presumably
+maintained by the aggressive reaction of receivers when the convention
+is broken (Enquist, Ghirlanda, and Hurd 2010; Tibbetts and Dale 2004).
+Our model captures both kinds of signals, although with some nuances.
+When we assume the handicap principle, simulations result in the
+evolution of honest signals mediated by the learned response. When we
+assume a cost-free signal, under certain conditions the learned response
+mediated the emergence of phenotypic variation in the signal. This
+variation, particularly in the mid-range of the distribution,
+facilitates the establishment of a convention: in this range,
+individuals respond appropriately to the trait of their peer. A simple
+correlation between the size os a signal and dominance rank (and
+quality) is expected under the handicap principle, as it is shown here
+and elsewhere (Botero et al. 2010; Grafen 1990; Johnstone and Norris
+1993). However, the branching event we documented shows that conventions
+can emerge in a less straightforward way when mediated by learning
+processes. These nuances could potentially make sense of seemingly
+contradictory evidence on the correlation between plumage traits and
+quality. A case in point is that of the House Sparrow (\emph{Passer
+domesticus}), a textbook example of badges of status. According to the
+prevailing narrative (Senar 2006; Searcy and Nowicki 2005), the bib size
+of males of this species is a signal of dominance rank and quality.
+However, a recent meta-analysis called into question this narrative by
+showing that the effect size of the association between bib size and
+dominance rank is small and uncertain (Sánchez-Tójar et al. 2018). Our
+simulations revealed that, conventions arose only in an intermediate
+range of the signalling trait, yet empirical studies seeing for
+correlations between phenotypic traits and quality are often performed
+across the whole range of variation. Whether such a situation may apply
+empirically to bib sizes in House Sparrows remains to be studied.
+
+The focus of theoretical and empirical work on communication systems,
+and particularly badges of status, is often explaining the presence and
+absence of certain morphological traits within populations. Phenotypic
+traits hypothesized to play a role in communication systems regularly
+exhibit patterns of geographic variation within species that have long
+called for explanations. One example of this is the leapfrog pattern
+(Remsen 1984), whereby a certain morphological trait alternates its
+presence and absence in a set of geographically adjunct populations,
+likely as a consequence of selection (Cadena, Cheviron, and Funk 2011;
+Márquez et al. 2020). However, it is unclear what type of ecological
+process is behind these selective regimes. Given the wide variation in
+outcomes found in our model, which depend on both stochasticity and
+cognitive parameters which may vary in space owing to various processes,
+we speculate that variation in cognition and learning could provide some
+explanatory power in this respect. A first step to assess this
+hypothesis would be to evaluate the way individuals in different
+populations respond to novel traits (\emph{e.g.} Avendaño and Cadena
+2021).
+
+In sum, we have presented here an evolutionary model where the evolution
+of a communication system is mediated through the learned responses of
+receivers. This is a novel way to understand the evolution of
+communication that integrates classical cognitive process in learning
+with evolutionary explanations of communication. This approach
+contributes to the further integration of proximate and ultimate
+explanations in behavioral and evolutionary biology.
 
@@
 
-{\centering \includegraphics[width=0.9\linewidth]{M:/Projects/SocialComp_morphCue/Simulations/betCostEvol1_/evolDyn5_betCost5}
+{\centering \includegraphics[width=1\linewidth]{manuscript_1.0_files/figure-latex/noBadge-1}
 
@@
 \ref{fig:handicap}.
-The middle panels show changes in the distribution of values fir $\alpha$ and $\beta$ along evolutionary time.
-Panels above and bellow correspond to snapshots of the sender and receiver codes, respectively, generation time of the snapshots are indicated by the grey lines in the middle panels.}\label{fig:noBadge}
+The middle panels show changes in the distribution of values for $\alpha$ and $\beta$ along evolutionary time.
+Panels above and below correspond to snapshots of the sender and receiver codes, respectively, generation time of the snapshots are indicated by the grey lines in the middle panels.}\label{fig:noBadge}
 \end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{manuscript_1.0_files/figure-latex/learnEvol-1}
+
+}
+
+\caption{Coevolutionary dynamics of both the signal reaction norm and key parameters of the learning module.
+Dynamics are portrait as changes in the distribution of values assorting in the population.
+In the top panels (A)  evolutionary dynamics of the sender code; A.1 shows the intercept and A.2 the slope of the logistic reaction norm.
+Both of these parameters experience evolutionary branching, whereby in the end three distincs types are assorting in the population.
+The middle panels (B) show changes in the distribution of values for the speed of learning ($A$;B1) and the behavioural tendency before learning ($y_z^0$; B2) along evolutionary time.
+The speed of learning does not change drastically, it maintains values well above 0 (black line).
+In contrast, the initial behavioural tendency does not evolve away from 0 (black line).
+Both of these conditions favour evolutionary branching in the signal reaction norm.
+Bottom Panels correspond to snapshots of the sender code.
+Generation time of the snapshots are indicated by the grey lines in the top and middle panels.}\label{fig:learnEvol}
+\end{figure}
+
+\newpage
+\small
+
+\begin{longtable}[]{@{}
+  >{\centering\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.2143}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.7857}}@{}}
+\caption{Notation of the model parameters and variables}\tabularnewline
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\centering
+Symbol
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Description
+\end{minipage} \\
+\midrule\noalign{}
+\endfirsthead
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\centering
+Symbol
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Description
+\end{minipage} \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+\(Q_i\) & Quality of individual \(i\) \\
+\(N\) & Population size \\
+\(n\) & Number of interactions as the focal individual \\
+\(\sigma\) & Standard deviation of the truncated normal distribution
+from which quality is drawn \\
+\(B_i\) & Badge size of individual \(i\) \\
+\(\alpha_i\) & Intercept of the badge size reaction norm for individual
+\(i\) \\
+\(\beta_i\) & Slope of the badge size reaction norm for individual
+\(i\) \\
+\(\mu\) & Mutation rate \\
+\(\sigma_\mu\) & Standard deviation of the normal distribution from
+which quality is mutations are drawn \\
+\(s_i\) & Survival probability of individual \(i\) \\
+\(k_1\) & Intercept of the survival function \\
+\(k_2\) & Slope of the survival function \\
+\(V\) & Value of the contested resource \\
+\(C\) & Cost of an escalated fight \\
+\(p_{ij}\) & Probability that individual \(i\) wins an escalated fight
+against individual \(j\) \\
+\(k_3\) & Parameter determining how important is quality defining the
+probability of wining \\
+\(\delta\) & Prediction error \\
+\(A\) & Speed of learning \\
+\(c\) & Number of feature centres \\
+\(b_z\) & badge size for feature centre \(z\) \\
+\(\hat{R}\) & Reward estimate \\
+\(x_z\) & weight of feature centre \(z\) on the reward estimation \\
+\(\theta\) & width of the generalization function \\
+\(q_{ij}\) & probability that individual \(i\) retreats from a fight in
+an encounter with individual \(j\) \\
+\(y_z\) & weight of feature centre z on the tendency to retreat \\
+\(w_i\) & total pay-off of individual \(i\) \\
+\(w_0\) & base line pay-off \\
+\end{longtable}
 
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-avendano_Territorial_2021}{}}%
+Avendaño, Jorge Enrique, and Carlos Daniel Cadena. 2021. {``Territorial
+Males Do Not Discriminate Between Local and Novel Plumage Phenotypes in
+a Tropical Songbird Species Complex: Implications for the Role of Social
+Selection in Trait Evolution.''} \emph{Behavioral Ecology and
+Sociobiology} 75 (2): 37.
+\url{https://doi.org/10.1007/s00265-021-02976-8}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-behrens_Associative_2008}{}}%
@@
 \url{https://doi.org/10.1038/nature07538}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-bond_Visual_2002}{}}%
+Bond, Alan B., and Alan C. Kamil. 2002. {``Visual Predators Select for
+Crypticity and Polymorphism in Virtual Prey.''} \emph{Nature} 415
+(6872): 609--13. \url{https://doi.org/10.1038/415609a}.
 
@@
 Bradbury, Jack W., and Sandra L. Vehrencamp. 2011. \emph{Principles of
-{Animal Communication}}. 2nd Edition. {Sunderland, Mass}: {Sinauer
-Associates is an imprint of Oxford University Press}.
+{Animal} {Communication}}. 2nd Edition. Sunderland, Mass: Sinauer
+Associates is an imprint of Oxford University Press.
+
+\leavevmode\vadjust pre{\hypertarget{ref-cadena_Testing_2011}{}}%
+Cadena, C. D., Z. A. Cheviron, and W. C. Funk. 2011. {``Testing the
+Molecular and Evolutionary Causes of a {`Leapfrog'} Pattern of
+Geographical Variation in Coloration.''} \emph{Journal of Evolutionary
+Biology} 24 (2): 402--14.
+\url{https://doi.org/10.1111/j.1420-9101.2010.02175.x}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-dieckmann_Dynamical_1996}{}}%
+Dieckmann, Ulf, and Richard Law. 1996. {``The Dynamical Theory of
+Coevolution: A Derivation from Stochastic Ecological Processes.''}
+\emph{Journal of Mathematical Biology} 34 (5): 579--612.
+\url{https://doi.org/10.1007/BF02409751}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-dridi_Learning_2018}{}}%
+Dridi, Slimane, and Erol Akçay. 2018. {``Learning to {Cooperate}: {The}
+{Evolution} of {Social} {Rewards} in {Repeated} {Interactions}.''}
+\emph{The American Naturalist} 191 (1): 58--73.
+\url{https://doi.org/10.1086/694822}.
 
@@
 Dridi, Slimane, and Laurent Lehmann. 2016. {``Environmental Complexity
-Favors the Evolution of Learning.''} \emph{Behav Ecol} 27 (3): 842--50.
-\url{https://doi.org/10.1093/beheco/arv184}.
+Favors the Evolution of Learning.''} \emph{Behavioral Ecology} 27 (3):
+842--50. \url{https://doi.org/10.1093/beheco/arv184}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-enquist_Signaling_2010}{}}%
+Enquist, Magnus, Stefano Ghirlanda, and Peter L. Hurd. 2010.
+{``Signaling.''} In \emph{Evolutionary {Behavioral} {Ecology}}. Oxford
+University Press.
 
@@
 Heyes, Cecilia. 2012. {``Simple Minds: A Qualified Defence of
-Associative Learning.''} \emph{Phil. Trans. R. Soc. B} 367 (1603):
-2695--2703. \url{https://doi.org/10.1098/rstb.2012.0217}.
+Associative Learning.''} \emph{Philosophical Transactions of the Royal
+Society B: Biological Sciences} 367 (1603): 2695--2703.
+\url{https://doi.org/10.1098/rstb.2012.0217}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-hoedjes_Natural_2010}{}}%
+Hoedjes, Katja M., H. Marjolein Kruidhof, Martinus E. Huigens, Marcel
+Dicke, Louise E. M. Vet, and Hans M. Smid. 2010. {``Natural Variation in
+Learning Rate and Memory Dynamics in Parasitoid Wasps: Opportunities for
+Converging Ecology and Neuroscience.''} \emph{Proceedings of the Royal
+Society B: Biological Sciences} 278 (1707): 889--97.
+\url{https://doi.org/10.1098/rspb.2010.2199}.
 
@@
 Johnstone, Rufus A., and Ken Norris. 1993. {``Badges of Status and the
-Cost of Aggression.''} \emph{Behav Ecol Sociobiol} 32 (2): 127--34.
-\url{https://doi.org/10.1007/BF00164045}.
+Cost of Aggression.''} \emph{Behavioral Ecology and Sociobiology} 32
+(2): 127--34. \url{https://doi.org/10.1007/BF00164045}.
 
 \leavevmode\vadjust pre{\hypertarget{ref-kamil_Optimal_1983}{}}%
-Kamil, Alan C. 1983. {``Optimal {Foraging Theory} and the {Psychology}
-of {Learning}.''} \emph{Integr Comp Biol} 23 (2): 291--302.
-\url{https://doi.org/10.1093/icb/23.2.291}.
+Kamil, Alan C. 1983. {``Optimal {Foraging} {Theory} and the {Psychology}
+of {Learning}.''} \emph{Integrative and Comparative Biology} 23 (2):
+291--302. \url{https://doi.org/10.1093/icb/23.2.291}.
 
-\leavevmode\vadjust pre{\hypertarget{ref-kurvers_Evolutionary_2014}{}}%
-Kurvers, Ralf H. J. M., Jens Krause, Darren P. Croft, Alexander D. M.
-Wilson, and Max Wolf. 2014. {``The Evolutionary and Ecological
-Consequences of Animal Social Networks: Emerging Issues.''} \emph{Trends
-in Ecology \& Evolution} 29 (6): 326--35.
-\url{https://doi.org/10.1016/j.tree.2014.04.002}.
+\leavevmode\vadjust pre{\hypertarget{ref-leimar_Evolution_2021}{}}%
+Leimar, Olof. 2021. {``The {Evolution} of {Social} {Dominance} Through
+{Reinforcement} {Learning}.''} \emph{The American Naturalist} 197 (5):
+560--75. \url{https://doi.org/10.1086/713758}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-leimar_Reproductive_2022}{}}%
+Leimar, Olof, and Redouan Bshary. 2022a. {``Reproductive Skew, Fighting
+Costs and Winner--Loser Effects in Social Dominance Evolution.''}
+\emph{Journal of Animal Ecology} n/a (n/a).
+\url{https://doi.org/10.1111/1365-2656.13691}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-leimar_Effects_2022}{}}%
+---------. 2022b. {``Effects of Local Versus Global Competition on
+Reproductive Skew and Sex Differences in Social Dominance Behaviour.''}
+\emph{Proceedings of the Royal Society B: Biological Sciences} 289
+(1987): 20222081. \url{https://doi.org/10.1098/rspb.2022.2081}.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-leimar_Flexibility_2023}{}}%
+Leimar, Olof, Andrés E. Quiñones, and Redouan Bshary. 2023.
+{``Flexibility of Learning in Complex Worlds.''} bioRxiv.
+\url{https://doi.org/10.1101/2023.06.12.544544}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-macphail_Brain_1982}{}}%
 Macphail, Euan M. 1982. \emph{Brain and {Intelligence} in
-{Vertebrates}}. {Clarendon Press}.
+{Vertebrates}}. Clarendon Press.
+
+\leavevmode\vadjust pre{\hypertarget{ref-marquez_Divergence_2020}{}}%
+Márquez, Roberto, Tyler P. Linderoth, Daniel Mejía-Vargas, Rasmus
+Nielsen, Adolfo Amézquita, and Marcus R. Kronforst. 2020. {``Divergence,
+Gene Flow, and the Origin of Leapfrog Geographic Distributions: {The}
+History of Colour Pattern Variation in {Phyllobates} Poison-Dart
+Frogs.''} \emph{Molecular Ecology} 29 (19): 3702--19.
+\url{https://doi.org/10.1111/mec.15598}.
 
@@
 Maynard-Smith, John. 1982. \emph{Evolution and the {Theory} of {Games}}.
-1 edition. {Cambridge ; New York}: {Cambridge University Press}.
+1 edition. Cambridge ; New York: Cambridge University Press.
 
@@
 McAuliffe, K., and A. Thornton. 2015. {``The Psychology of Cooperation
-in Animals: An Ecological Approach.''} \emph{J Zool} 295 (1): 23--35.
-\url{https://doi.org/10.1111/jzo.12204}.
+in Animals: An Ecological Approach.''} \emph{Journal of Zoology} 295
+(1): 23--35. \url{https://doi.org/10.1111/jzo.12204}.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-osvath_Spontaneous_2009}{}}%
+Osvath, Mathias. 2009. {``Spontaneous Planning for Future Stone Throwing
+by a Male Chimpanzee.''} \emph{Current Biology} 19 (5): R190--91.
+\url{https://doi.org/10.1016/j.cub.2009.01.010}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-parker_Assessment_1974}{}}%
@@
 \url{https://doi.org/10.1016/0022-5193(74)90111-8}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-pretot_Factors_2016}{}}%
+Prétôt, Laurent, Redouan Bshary, and Sarah F. Brosnan. 2016. {``Factors
+Influencing the Different Performance of Fish and Primates on a
+Dichotomous Choice Task.''} \emph{Animal Behaviour} 119 (September):
+189--99. \url{https://doi.org/10.1016/j.anbehav.2016.06.023}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-pretot_Comparative_2021}{}}%
+Prétôt, Laurent, Jennifer Mickelberg, Jodi Carrigan, Tara Stoinski,
+Redouan Bshary, and Sarah F. Brosnan. 2021. {``Comparative Performance
+of Orangutans ({Pongo} Spp.), Gorillas ({Gorilla} Gorilla Gorilla), and
+Drills ({Mandrillus} Leucophaeus), in an Ephemeral Foraging Task.''}
+\emph{American Journal of Primatology} 83 (1): e23212.
+\url{https://doi.org/10.1002/ajp.23212}.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-quinones_Reinforcement_2019}{}}%
-Quiñones, Andrés E., Olof Leimar, Arnon Lotem, and Redouan Bshary. 2019.
-{``Reinforcement {Learning Theory Reveals} the {Cognitive Requirements}
-for {Solving} the {Cleaner Fish Market Task}.''} \emph{The American
-Naturalist}, December, 000--000. \url{https://doi.org/10.1086/707519}.
+\leavevmode\vadjust pre{\hypertarget{ref-quinones_Reinforcement_2020}{}}%
+Quiñones, Andrés E., Olof Leimar, Arnon Lotem, and Redouan Bshary. 2020.
+{``Reinforcement {Learning} {Theory} {Reveals} the {Cognitive}
+{Requirements} for {Solving} the {Cleaner} {Fish} {Market} {Task}.''}
+\emph{The American Naturalist} 195 (4): 664--77.
+\url{https://doi.org/10.1086/707519}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-raby_Planning_2007}{}}%
+Raby, C. R., D. M. Alexis, A. Dickinson, and N. S. Clayton. 2007.
+{``Planning for the Future by Western Scrub-Jays.''} \emph{Nature} 445
+(7130): 919. \url{https://doi.org/10.1038/nature05575}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-remsen_High_1984}{}}%
+Remsen, J. V. 1984. {``High {Incidence} of "{Leapfrog}" {Pattern} of
+{Geographic} {Variation} in {Andean} {Birds}: {Implications} for the
+{Speciation} {Process}.''} \emph{Science} 224 (4645): 171--73.
+\url{https://doi.org/10.1126/science.224.4645.171}.
 
 \leavevmode\vadjust pre{\hypertarget{ref-rohwer_Social_1975a}{}}%
-Rohwer, Sievert. 1975. {``The {Social Significance} of {Avian Winter
-Plumage Variability}.''} \emph{Evolution} 29 (4): 593--610.
+Rohwer, Sievert. 1975. {``The {Social} {Significance} of {Avian}
+{Winter} {Plumage} {Variability}.''} \emph{Evolution} 29 (4): 593--610.
 \url{https://doi.org/10.2307/2407071}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-sanchez-tojar_Metaanalysis_2018a}{}}%
+Sánchez-Tójar, Alfredo, Shinichi Nakagawa, Moisès Sánchez-Fortún,
+Dominic A Martin, Sukanya Ramani, Antje Girndt, Veronika Bókony, et al.
+2018. {``Meta-Analysis Challenges a Textbook Example of Status
+Signalling and Demonstrates Publication Bias.''} Edited by Diethard
+Tautz and Tim Parker. \emph{eLife} 7 (November): e37385.
+\url{https://doi.org/10.7554/eLife.37385}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-searcy_Evolution_2005}{}}%
+Searcy, William A., and Stephen Nowicki. 2005. \emph{The {Evolution} of
+{Animal} {Communication}: {Reliability} and {Deception} in {Signaling}
+{Systems}}. Princeton: Princeton University Press.
+
+\leavevmode\vadjust pre{\hypertarget{ref-3062aab1-6f44-361c-a548-6bc040ba8ca2}{}}%
+Senar, Juan Carlos. 2006. {``Color Displays as Intrasexual Signals of
+Aggression and Dominance.''} In \emph{Bird Coloration, Volume 2:
+{Function} and Evolution}, 87--136. Harvard University Press.
+\url{http://www.jstor.org/stable/j.ctv22jnr8k.6}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-sonnenberg_Natural_2019}{}}%
+Sonnenberg, Benjamin R., Carrie L. Branch, Angela M. Pitera, Eli Bridge,
+and Vladimir V. Pravosudov. 2019. {``Natural {Selection} and {Spatial}
+{Cognition} in {Wild} {Food}-{Caching} {Mountain} {Chickadees}.''}
+\emph{Current Biology} 29 (4): 670--676.e3.
+\url{https://doi.org/10.1016/j.cub.2019.01.006}.
 
@@
 Staddon, J. E. R. 2016. \emph{Adaptive {Behavior} and {Learning}}. 2
-edition. {Cambridge}: {Cambridge University Press}.
+edition. Cambridge: Cambridge University Press.
 
@@
 Sutton, Richard S., and Andrew G. Barto. 2018. \emph{Reinforcement
-{Learning}: {An Introduction}}. Edited by Francis Bach. Second edition
-edition. {Cambridge, MA}: {A Bradford Book}.
+{Learning}: {An} {Introduction}}. Edited by Francis Bach. Second edition
+edition. Cambridge, MA: A Bradford Book.
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-zahavi_Mate_1975}{}}%
-Zahavi, Amotz. 1975. {``Mate Selection\textemdash -{A} Selection for a
+Zahavi, Amotz. 1975. {``Mate Selection----{A} Selection for a
 Handicap.''} \emph{Journal of Theoretical Biology} 53 (1): 205--14.

### 2023-09-14 AccFunc.R

@@ -218,7 +218,8 @@ evolDist<-function(indData,variable,nbins,range=NULL,pal, nlevels =10,
   #                 new=NewP)
   
  
-  filled.contour3(x=as.numeric(timeseq),y = as.numeric(bins),z = log(0.01+t(contour1)),
+  filled.contour3(x=as.numeric(timeseq),y = as.numeric(bins),
+                  z = log(0.01+t(contour1)),
                  color.palette = pal, nlevels = nlevels,ylim=ylim,
                  plot.axes={
                    axis(1,cex.axis=cexAxis,xaxt=xaxt)
@@ -312,35 +313,6 @@ ESS<-function(pay_offsM){
 }
 
 
-
-# clus_silou<-c(0,clus_silou)
-# nclustersSil<-which.max(clus_silou)
-
-
-# DT[,idClust:=as.numeric(idClust)]
-
-# for(i in which(DT[,seed] %in% runChoi)){
-#   set(popFinal,i,match( "idClust",names(DT)),
-#       clustersMatch[clustersMatch[,2]==i,1])
-# }
-
-
-# #SBATCH --job-name=TestJOB		#Nombre del job
-# #SBATCH -p short			#Cola a usar, Default=short (Ver colas y límites en /hpcfs/shared/README/partitions.txt)
-# #SBATCH -N 1				#Nodos requeridos, Default=1
-# #SBATCH -n 1				#Tasks paralelos, recomendado para MPI, Default=1
-# #SBATCH --cpus-per-task=1		#Cores requeridos por task, recomendado para multi-thread, Default=1
-# #SBATCH --mem=2000		#Memoria en Mb por CPU, Default=2048
-# #SBATCH --time=00:10:00			#Tiempo máximo de corrida, Default=2 horas
-# #SBATCH --mail-user=USER@uniandes.edu.co
-# #SBATCH --mail-type=ALL			
-# #SBATCH -o TEST_job.o%j			#Nombre de archivo de salida
-# 
-# host=`/bin/hostname`
-# date=`/bin/date`
-# echo "Soy un JOB de prueba"
-# echo "Corri en la maquina: "$host
-# echo "Corri el: "$date
 
 
 

### 2023-09-14 tmp.R

@@ -1,21 +1,0 @@
-tmp<-function(x){
-  count<-match(x,popOneInd[,sort(unique(get(shortSce)),decreasing = TRUE)])
-  tempPop<-popOneInd[get(shortSce)==x&(seed==repsScen[count]&time==max(time))]
-  tempPop<-tempPop[,.SD[.N],
-                   .SDcol=c(grep("WeightAct",
-                                 names(popOneInd),value = TRUE),"Quality","alpha","beta"
-                            # ,
-                            # "orderClus"),
-                   ),
-                   by=indId]
-  dataIndReact<-sapply(as.list(tempPop[,indId]),
-                       function(x){x=
-                         sapply(rangx, 
-                                function(y)
-                                  do.call(logist,
-                                          as.list(
-                                            c(y,as.double(
-                                              tempPop[indId==x,.SD,
-                                                      .SDcol=c("alpha","beta")])))))})
-  return(dataIndReact)
-}

### 2023-09-14 manuscript/manuscript_1.0.Rmd
Forgot to save some changes

@@
 lay_summary:  |
-  Biological signals allow individuals to respond appropietly. For example, a 
-  signal of fighting hability would allow peers to decide whether to fight 
-  another individual. But, what prevents weaklings from faking to be 
-  strong? Using a theoretical approach we assess whether learning has 
-  a role in maintaining a signal honest. We find that when individuals 
-  learn how best to respond to a signal, populations are composed of a 
-  mixed of honest and dishonest signals. 
+  Biological signals allow animals to respond appropriately when they compete 
+  with each other. For example, a signal of fighting ability would allow 
+  individuals to decide whether to fight another individual. But, what prevents 
+  weaklings from faking to be strong? Using a theoretical approach, we assess 
+  whether learning has a role in maintaining honest signals. We find that when 
+  individuals learn how best to respond to a signal, populations evolve to be 
+  composed of a mix of honest and dishonest signals.
+
 editor_options:

### 2023-09-14 quadraticSelection.R

@@ -128,7 +128,7 @@ ggplot(popOneInd.gen,aes(y=resid.loess,x=binFreq))+
   geom_smooth()+
   theme_classic()
 
-ggplot(popOneInd.gen,aes(y=binFreq,x=Quality))+
+%>% ggplot(popOneInd.gen,aes(y=binFreq,x=Quality))+
   geom_point(aes(color=as.factor(seed)))+
   geom_smooth()+
   theme_classic()

### 2023-09-19 manuscript/manuscript_1.0.tex
Elimate files no longer needed for the publication

@@
 \subsection*{Lay summary}
-Biological signals allow individuals to respond appropietly. For
-example, a singal of fighting hability would allow peers to decide
-whether to fight another individual. But, what prevents a weaklings from
-faking to be strong? Using a theoretical approach we assess whether
-learning has a role in maintaining signals honest. We find that when
-individuals learn how best to respond to a signal, populations are
-composed of a mixed of honest and dishonest signals.
+Biological signals allow animals to respond appropriately when they
+compete with each other. For example, a signal of fighting ability would
+allow individuals to decide whether to fight another individual. But,
+what prevents weaklings from faking to be strong? Using a theoretical
+approach, we assess whether learning has a role in maintaining honest
+signals. We find that when individuals learn how best to respond to a
+signal, populations evolve to be composed of a mix of honest and
+dishonest signals.
 

### 2023-09-19 Dyn_reac_learn.R

@@ -1,1516 +1,0 @@
-# Evolutionary dynamics for averages and individual runs, reaction norms of sampled 
-# individuals and their learning profiles
-
-# Required libraries -----------------------------------------------------------
-
-require(here)
-here()
-source(here("AccFunc.R"))
-require("foreach")
-require("doParallel")
-require("jsonlite")
-library(dplyr)
-
-# Scenario to be plotted - corresponds to folders where simulations are stored
-
-scenario<-"betCostalphaL"
-
-# extSimsDir<-here("Simulations",paste0(scenario,"_"))
-  # paste0("e:/BadgeSims/",scenario,"_")
-
-
-
-# Load files -------------------------------------------------------------------
-# when program was run with internal paralellization
-
-# Project folder
-(listTest<-list.files(here("Simulations",paste0(scenario,"_")),full.names = TRUE))
-# External sims folder
-# (listTest<-list.files(extSimsDir,full.names = TRUE))
-
-(evolList<-grep("evolLearn",listTest,value=TRUE))
-(indList<-grep("indLearn",listTest,value=TRUE))
-paramName<-list.files(here("Simulations",paste0(scenario,"_")))
-# paramName<-list.files(extSimsDir,full.names = TRUE)
-paramName<-grep(".json",paramName,value=TRUE)
-paramRang<-sapply(paramName,function(json){
-  tmpJson<-fromJSON(here("Simulations",paste0(scenario,"_"),json))
-  return(tmpJson$rangParam)
-})  %>%  unique
-param<-fromJSON(here("Simulations",paste0(scenario,"_"),paramName[1]))
-# Choose which parameter to plot
-
-val<-1
-
-
-# Load files -------------------------------------------------------------------
-# when program was run with external paralellization
-
-
-
-# loop to produce pdfs for parameter values
-numCores <- length(paramRang)
-registerDoParallel(numCores)
-
-foreach(val = 1:length(paramRang),
[…]

### 2023-09-19 Dyn_reac_learn_noLearn.R

@@ -1,440 +1,0 @@
-# Evolutionary dynamics for averages and individual runs
-# Model with reaction norms on both sides of the communication systems
-
-# Required libraries -----------------------------------------------------------
-
-require(here)
-here()
-source(here("AccFunc.R"))
-require("foreach")
-require("doParallel")
-require("jsonlite")
-
-# Scenario to be plotted - corresponds to folders where simulations are stored
-
-scenario<-"betCostNoLearnShare0"
-SimDir<-"Simulations"
-
-extSimsDir<-#here("Simulations",paste0(scenario,"_"))
-  paste0("e:/BadgeSims/",scenario,"_")
-
-
-# Load files -------------------------------------------------------------------
-# Project folder
-# (listTest<-list.files(here(SimDir,paste0(scenario,"_"))))
-# External sims folder
-(listTest<-list.files(extSimsDir,full.names = TRUE))
-
-(evolList<-grep("evolLearn",listTest,value=TRUE))
-
-paramName<-list.files(here(SimDir,paste0(scenario,"_")))
-paramName<-grep(".json",paramName,value=TRUE)
-param<-fromJSON(here(SimDir,paste0(scenario,"_"),paramName[9]))
-
-
-numCores <- length(evolList)
-registerDoParallel(numCores)
-
-val<-2
-
-
-# loop to produce pdfs for parameter values
-foreach(val = 1:2,.packages = c("data.table","here")) %dopar% {
-  source(here("AccFunc.R"))
-  
-  
-
-fileId<-val
-
-
-# Project folder
-# evol<-fread(here(SimDir,paste0(scenario,"_"),evolList[fileId]))
-# pop<-fread(here(SimDir,paste0(scenario,"_"),indList[fileId]))
-# External sims folder
-
-# evol<-fread(evolList[fileId])
-# pop<-fread(indList[fileId])
-
-
-(Valpar<-param$rangParam[val])
-
-(nampar<-param$namParam)
-
-evolList_runs<-grep(paste0(param$namParam,param$rangParam[val]),
-                    evolList,value =TRUE)
-
-evol<-do.call(rbind,lapply(evolList_runs,fread))
-
[…]

### 2023-09-19 Filled.contour3.R

@@ -1,0 +1,63 @@
+filled.contour3 <-
+  function (x = seq(0, 1, length.out = nrow(z)),
+            y = seq(0, 1, length.out = ncol(z)), z, xlim = range(x, finite = TRUE), 
+            ylim = range(y, finite = TRUE), zlim = range(z, finite = TRUE), 
+            levels = pretty(zlim, nlevels), nlevels = 20, color.palette = cm.colors, 
+            col = color.palette(length(levels) - 1), plot.title, plot.axes, 
+            key.title, key.axes, asp = NA, xaxs = "i", yaxs = "i", las = 1, 
+            axes = TRUE, frame.plot = axes,mar, ...) 
+{
+  # modification by Ian Taylor of the filled.contour function
+  # to remove the key and facilitate overplotting with contour()
+  # further modified by Carey McGilliard and Bridget Ferris
+  # to allow multiple plots on one page
+
+  if (missing(z)) {
+    if (!missing(x)) {
+      if (is.list(x)) {
+        z <- x$z
+        y <- x$y
+        x <- x$x
+      }
+      else {
+        z <- x
+        x <- seq.int(0, 1, length.out = nrow(z))
+      }
+    }
+    else stop("no 'z' matrix specified")
+  }
+  else if (is.list(x)) {
+    y <- x$y
+    x <- x$x
+  }
+  if (any(diff(x) <= 0) || any(diff(y) <= 0)) 
+    stop("increasing 'x' and 'y' values expected")
+ # mar.orig <- (par.orig <- par(c("mar", "las", "mfrow")))$mar
+ # on.exit(par(par.orig))
+ # w <- (3 + mar.orig[2]) * par("csi") * 2.54
+ # par(las = las)
+ # mar <- mar.orig
+ plot.new()
+ # par(mar=mar)
+  plot.window(xlim, ylim, "", xaxs = xaxs, yaxs = yaxs, asp = asp)
+  if (!is.matrix(z) || nrow(z) <= 1 || ncol(z) <= 1) 
+    stop("no proper 'z' matrix specified")
+  if (!is.double(z)) 
+    storage.mode(z) <- "double"
+  .filled.contour(as.double(x), as.double(y), z, as.double(levels),
+                        col = col)
+	  if (missing(plot.axes)) {
+    if (axes) {
+      title(main = "", xlab = "", ylab = "")
+      Axis(x, side = 1)
+      Axis(y, side = 2)
+    }
+  }
[…]

### 2023-09-19 cartoons_reactNorms.R

@@ -1,66 +1,0 @@
-########## Cartoon plot of reaction norms #####################################
-
-# Load libraries and external functions -----------------------------------
-library(here)
-# directory where source files are saved
-
-source(here("..","R_files","posPlots.R"))
-
-
-## Reaction norms ------------------------------------------------------------
-
-# output file
-
-# png(here("reacNorm.png"),width = 800,height = 800)
-
-
-qualRange<-seq(0,1,length=100)
-alpha<-3
-beta<-6
-badgeFunc<-function(x,alpha,beta,maxBadge) {d<-maxBadge/(1+exp(alpha-beta*x))}
-par(plt=posPlot())
-plot(badgeFunc(qualRange,alpha,beta,1)~qualRange,type="l",cex.lab=1.5,
-     cex.axis=2,lwd=3,
-     col="red",ylim=c(0,1.1),xaxt="s",yaxt="s",ylab="",xlab="") #,col=i
-# lines(badgeFunc(qualRange,3,-60,1.05)~qualRange,cex.lab=1.5,cex.axis=2,lwd=3,
-#       col="blue")
-mtext(side=1,text="Quality",line=2,cex=5)
-mtext(side=2,text="Badge",line=2,cex=5)
-# text(0.01,1.17,"A", cex=2)
-dev.off()
-
-# Receiver 3d reaction norm ----------------------------------------------------
-
-rivalBadge<-qualRange
-
-
-
-perspTmp<-outer(X=qualRange,Y = rivalBadge,FUN = logist3d, 
-                alpha=-10,beta=-20,gamma=0)
-
-persp(x = qualRange,y=rivalBadge,z=perspTmp,phi = 35, theta=40 ,
-      xlab="Quality",ylab="Rival's badge",zlim=c(0,1),
-      zlab="p(Dove)",col="lightblue")
-
-## Survival function -----------------------------------------------------------
-
-# png(here("reacNorm.png"),width = 800,height = 800)
-
-badQualRang<-seq(0,1,length=1000)
-alpha<-2
-beta<-0
-costFunc<-function(beta,alpha,x) {d<-1/(1+exp(-alpha-beta*x))}
-par(plt=posPlot())
-betRang<-c(0)
-matplot(y=sapply(betRang,FUN = costFunc,alpha=alpha,x=badQualRang),x=badQualRang,
[…]

### 2023-09-19 clusterAnalysis.R

@@ -1,277 +1,0 @@
-require(here)
-here()
-source(here("AccFunc.R"))
-source(here("..","R_files","NbClust.R"))
-require("foreach")
-require("doParallel")
-require("jsonlite")
-require("factoextra")
-require("cluster")
-
-# Scenario to be plotted - corresponds to folders where simulations are stored
-
-scenario<-"nIntGroupNormQual2"
-
-extSimsDir<-#here("Simulations",paste0(scenario,"_"))
-  paste0("e:/BadgeSims/",scenario,"_")
-
-
-# Load files -------------------------------------------------------------------
-# Project folder
-# (listTest<-list.files(here("Simulations",paste0(scenario,"_"))))
-# External sims folder
-(listTest<-list.files(extSimsDir,full.names = TRUE))
-
-(evolList<-grep("evolLearn",listTest,value=TRUE))
-(indList<-grep("indLearn",listTest,value=TRUE))
-paramName<-list.files(here("Simulations",paste0(scenario,"_")))
-paramName<-list.files(extSimsDir,full.names = TRUE)
-paramName<-grep(".json",paramName,value=TRUE)
-param<-#fromJSON(here("Simulations",paste0(scenario,"_"),paramName))
-  fromJSON(paramName[1])
-
-
-val<-1
-
-# loop to produce pdfs for parameter values
-# numCores <- length(indList)
-# registerDoParallel(numCores)
-# foreach(val = 1:length(indList),.packages = c("data.table","here")) %dopar% {
-#   source(here("AccFunc.R"))
-  
-  fileId<-val
-  
-  # Project folder
-  # evol<-fread(here("Simulations",paste0(scenario,"_"),evolList[fileId]))
-  # pop<-fread(here("Simulations",paste0(scenario,"_"),indList[fileId]))
-  
-  # External sims folder
-  # evol<-fread(evolList[fileId])
-  # pop<-fread(indList[fileId])
-
-  evolList_runs<-grep(paste0(param$namParam,param$rangParam[val]),
-                      evolList,value =TRUE)
-  indList_runs<-grep(paste0(param$namParam,param$rangParam[val]),
-                     indList,value =TRUE)
-  
-  
[…]

### 2023-09-19 compPar.R

@@ -1,356 +1,0 @@
-# Varying parameter of a single scenario ---------------------------------------
-
-# Required libraries -----------------------------------------------------------
-
-library(here)
-here()
-source(here("AccFunc.R"))
-library("scales")
-
-# Scenario to be plotted - corresponds to folders where simulations are stored
-
-scenario<-"initAct"
-
-# extSimsDir<-#here("Simulations",paste0(scenario,"_"))
-#   paste0("e:/BadgeSims/",scenario,"_")
-
-# Load files -------------------------------------------------------------------
-
-(listTest<-list.files(here("Simulations",paste0(scenario,"_")),full.names = TRUE))
-(listTest<-grep(".txt",listTest,value = TRUE))
-
-# (listTest<-list.files(extSimsDir,full.names = TRUE))
-
-(sdList<-grep("evol",listTest,value=TRUE))
-(indList<-grep("ind",listTest,value=TRUE))
-
-fread(indList[2])
-
-inds<-rbindlist(lapply(indList,filesScenar,scenario,full.name=TRUE),fill = TRUE)
-
-inds[,diffActWeight:=abs(WeightAct_0-WeightAct_4)]
-
-shortSce<-gsub("[^[:alpha:]]",gsub(".txt","",tail(strsplit(indList[1],"_")[[1]],1)),
-               replacement = "")
-  
-
-# Graphs -----------------------------------------------------------------------
-
-inds[,unique(get(shortSce))]
-
-inds[get(shortSce)==unique(get(shortSce))[3]]
-
-gener<-tail(inds[,unique(time)],1)
-lastInt<-tail(inds[,unique(nInteract)],1)
-runChoi<-0
-nCenters<-6
-interv<-1/(nCenters-1)
-# centers when extremes are included
-centers<-interv*seq(0,nCenters-1) 
-# centers when extremes are not included
-# centers<-interv*0.5+interv*seq(0,nCenters-1) 
-yaxtRang<-c("s",rep("n",10))
-yaxlabs<-c("p(D)",rep("",10))
-tempPop<-inds[time==gener&nInteract==lastInt[1],.SD[.N],
-             .SDcol=c(grep("Weight",
-                           names(inds),value = TRUE),"Quality",
-                      as.character(shortSce),"diffActWeight",
-                      "alpha","beta"),
-             by=indId]
-tempPop[,unique(get(shortSce))]
[…]

### 2023-09-19 compScen.R

@@ -1,38 +1,0 @@
-# Comparing results from different scenarios -----------------------------------
-
-# Required libraries -----------------------------------------------------------
-
-library(here)
-here()
-source(here("AccFunc.R"))
-
-
-# Choove scenarios to be plotted - corresponds to folders where simulations are stored
-
-scenarios<-c("nIntGroupEvol1","betCostEvol1")
-
-variable<-
-  lapply(scenarios,FUN=gsub,pattern="[^[:alpha:]]",replacement = '')
-variable1<-gsub("Evol",variable1,replacement = '')
-extSimsDir<-paste0("e:/BadgeSims/",scenario1,"_")
-
-variable2<-gsub("[^[:alpha:]]",scenario2,replacement = '')
-variable2<-gsub("Evol",variable2,replacement = '')
-extSimsDir<-paste0("e:/BadgeSims/",scenario2,"_")
-
-
-# Load files -------------------------------------------------------------------
-
-(listTest<-list.files(extSimsDir,full.names = TRUE))
-(evolList<-grep("evolLearn",listTest,value=TRUE))
-(indList<-grep("indLearn",listTest,value=TRUE))
-
-nampar<-gsub("[^[:alpha:]]",gsub(".txt","",tail(strsplit(indList[1],"_")[[1]],1)),
-             replacement = "")
-
-# (listTest<-list.files(here("Simulations",paste0(scenario,"_"))))
-# (sdList<-grep("evol",listTest,value=TRUE))
-
-evol<-do.call(rbind,lapply(evolList,filesScenar,scenario,full.name=TRUE))
-
-pop<-do.call(rbind,lapply(indList, filesScenar,scenario,full.name=TRUE))

### 2023-09-19 dynamics.R

@@ -1,487 +1,0 @@
-# Evolutionary dynamics of the hawk dove game with an extra strategy that 
-# learns a preference for the phenotypic strategy depending on a morphological
-# trait
-
-# Required libraries -----------------------------------------------------------
-
-library(here)
-here()
-library(dplyr)
-source(here("AccFunc.R"))
-
-
-# Scenario to be plotted - corresponds to folders where simulations are stored
-
-scenario<-"betCostalphaL2"
-
-
-# Load files -------------------------------------------------------------------
-
-(listTest<-list.files(here("Simulations",paste0(scenario,"_"))))
-(sdList<-grep("evol",listTest,value=TRUE)%>%
-    grep(pattern = ".txt",value = TRUE))
-
-evol<-fread(here("Simulations",paste0(scenario,"_"),sdList[1]))
-
-# Extract means and IQR for the dynamic variables ------------------------------
-
-evolStats<-evol[,.(m.freqGenHawk=mean(freqGenHawks),
-                     upIQR.freqGenHawk=fivenum(freqGenHawks)[4],
-                     lowIQR.freqGenHawk=fivenum(freqGenHawks)[2],
-                     m.freqGenDove=mean(freqGenDove),
-                     upIQR.freqGenDove=fivenum(freqGenDove)[4],
-                     lowIQR.freqGenDove=fivenum(freqGenDove)[2],
-                     m.freqEval=mean(freqGenEval),
-                     upIQR.freqGenEval=fivenum(freqGenEval)[4],
-                     lowIQR.freqGenEval=fivenum(freqGenEval)[2], 
-                     m.freqFenHawk=mean(freqFenHawks),
-                     upIQR.freqFenHawk=fivenum(freqFenHawks)[4],
-                     lowIQR.freqFenHawk=fivenum(freqFenHawks)[2],
-                     m.freqFenDove=mean(freqFenDoves),
-                     upIQR.freqFenDove=fivenum(freqFenDoves)[4],
-                     lowIQR.freqFenDove=fivenum(freqFenDoves)[2],
[…]

### 2023-09-19 posPlots.R

@@ -1,0 +1,23 @@
+posPlot<-function(numplotx=1,numploty=1,idplotx=1,idploty=1,title=F,lowboundx=15,
+upboundx=95,lowboundy=15,upboundy=95)
+{
+  if(idplotx>numplotx){print("error x id does not match number of plots")}
+  else{
+  if(idploty>numploty){print("error y id does not match number of plots")}
+  else{
+  #lowboundx<-15
+  #if(title){upboundy<-95}
+  #else {upboundy<-98}
+  #lowboundy<-15
+  upxs<--1#rep(-1,2)numplotx)
+  lowxs<-1#rep(1,2)numplotx)
+  upys<--1#rep(-1,2)numploty)
+  lowys<-1#rep(1,2)numploty)
+  stepx<-(upboundx-lowboundx)/numplotx
+  stepy<-(upboundy-lowboundy)/numploty
+  upxs<-lowboundx+idplotx*stepx+upxs*0.5
+  lowxs<-lowboundx+(idplotx-1)*stepx+lowxs*0.5
+  upys<-lowboundy+idploty*stepy+upys*0.5
+  lowys<-lowboundy+(idploty-1)*stepy+lowys*0.5
+  return(c(lowxs,upxs,lowys,upys)/100)}}
+}

### 2023-09-19 quadraticSelection.R

@@ -1,147 +1,0 @@
-# Playing around to visualize frequancy- dependent selection
-# 
-
-# Required libraries -----------------------------------------------------------
-
-require(here)
-library(ggplot2)
-source(here("AccFunc.R"))
-require("jsonlite")
-library(dplyr)
-# Scenario to be plotted - corresponds to folders where simulations are stored
-
-scenario<-"betCostEvol5"#"alphaAct"#"nIntGroupNormQual"#
-
-# extSimsDir<-#here("Simulations",paste0(scenario,"_"))
-#   paste0("M:/BadgeSims/",scenario,"_")
-
-
-# Load files -------------------------------------------------------------------
-# Project folder
-(listTest<-list.files(here("Simulations",paste0(scenario,"_")),full.names = TRUE))
-# External sims folder
-# (listTest<-list.files(extSimsDir,full.names = TRUE))
-
-(evolList<-grep("evolLearn",listTest,value=TRUE))
-(indList<-grep("indLearn",listTest,value=TRUE))
-paramName<-list.files(here("Simulations",paste0(scenario,"_")))
-# paramName<-list.files(extSimsDir,full.names = TRUE)
-paramName<-grep(".json",paramName,value=TRUE)
-paramName<-grep(".json",paramName,value=TRUE) %>% grep(pattern = "*s0_*",value = TRUE)
-param<-fromJSON(here("Simulations",paste0(scenario,"_"),paramName[2]))
-# fromJSON(paramName[1])
-
-
-val<-1
-fileId<-val
-
-
-evolList_runs<-grep(paste0(param$namParam,param$rangParam[val]),
-                    evolList,value =TRUE)
-indList_runs<-grep(paste0(param$namParam,param$rangParam[val]),
-                   indList,value =TRUE)
-
-
-evol<-do.call(rbind,lapply(evolList_runs,fread))
-pop<-do.call(rbind,lapply(indList_runs, fread))
-
-
-lastInt<-tail(pop[,unique(nInteract)],1)
-
-popOneInd<-pop[nInteract==lastInt]
-
-pop[,unique(cumPayoff)]
-
-popOneInd[,unique(time)]
-
[…]

### 2023-11-15 manuscript/manuscript_1.0.Rmd
submitted version 1.0

@@
 title: |
-   \large \textbf{Title:} The role of learning in the evolution of status signalling: a modeling approach
-subtitle: |
-  \normalsize \textbf{Abbreviated title:} Learning and evolution in status signalling
-author: 
+   The role of learning in the evolution of status signaling: a modeling approach
+# subtitle: |
+#   \normalsize \textbf{Abbreviated title:} Learning and evolution in status signalling
 # author:
@@
     \renewcommand{\thetable}{S\arabic{table}}\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
+    \providecommand{\keywords}[1]{\textbf{\textit{Keywords---}} #1}
   ```
-keywords: My keywords
+keywords: |
+  Biological signals, evolutionary branching, agent-based modeling, animal communication, badges of status,
+  handicap principle
 linespread: 2 
@@
   composed of a mix of honest and dishonest signals.
-
 editor_options:
@@
 ---
+
+```{=latex}
+\keywords{Signals, evolutionary branching, agent-based modeling, animal communication, badges of status, handicap principle}
+```
 
@@
 
+The code necessary to run simulations as well as to produce visualizations can be found 
+in: \url{https://doi.org/10.5281/zenodo.8359809}
+
 ```{r modelStruc,echo=FALSE,out.width='100%',fig.height=5,fig.align='center',fig.show='hold',fig.cap= "Model of communication in the context of aggressive interactions.
@@
  # include bird cartoons
- rasterImage(honestBig,0.75,0.87,0.95,1.07)
- rasterImage(honestSmall,0.05,0.01,0.25,0.21)
- rasterImage(dishonest,0.75,0.5,0.95,0.7)
- rasterImage(dishonest,0.05,0.5,0.25,0.7)
+ rasterImage(honestBig,0.77,0.87,0.95,1.07)
+ rasterImage(honestSmall,0.05,0.01,0.22,0.21)
+ rasterImage(dishonest,0.77,0.5,0.95,0.7)
+ rasterImage(dishonest,0.05,0.5,0.22,0.7)
  # legend("left",legend=rangeBeta,col=colRuns,lwd=3,title = expression(beta),cex=2,bty="n")

### 2023-11-24 manuscript/manuscript_2.0.Rmd
starting revision changes

--- 2023-11-15 manuscript/manuscript_1.0.Rmd
@@
 \begin{equation}
-\hat{R}_{ij} =  \sum_{z=1}^{c} x_z e^{(-\frac{|B_i-b_z|}{2\theta^2})} 
+\hat{R}_{ij} =  \sum_{z=1}^{c} x_z e^{(-\frac{|B_j-b_z|}{2\theta^2})} 
 \end{equation}
@@
 partner is chosen at random. Thus, the expected number of interactions for 
-each individual is $n+\frac{n(N+1)}{N}$. 
+each individual is $2n$. 
 After the interaction round , individuals in the population 
@@
 
-```{r handicap,echo=FALSE,out.width='100%',fig.height=6,fig.align='center',fig.show='hold',fig.cap="Evolution of the badge size as a handicap mediated by learning.
+```{r handicap,echo=FALSE,out.width='100%',fig.height=7,fig.align='center',fig.show='hold',fig.cap="Evolution of the badge size as a handicap mediated by learning.
 Top panels (A.1 and A.2) show the evolutionary dynamics of the sender code.
+In the y axis the trait value, in the x axis the generation time in thousands of generations.
 On the left (A.1), changes in the distribution of values for the intercept of the reaction norm ($\\alpha$); on the right (A.2) changes in the distribution of the slope ($\\beta$).
 Darker areas of the background correspond to values with high frequency, while orange lines in both panels show the mean of the distribution.
-X axis in the evolutionary dynamics are given in thousands of generations.
-Grey lines show the generation time corresponding to the panels below portraing the sender (B) and receiver code(C).
+Grey lines show the generation time corresponding to the panels below portraing the sender (B) and receiver code (C).
 In the bottom panels (C1-4), the learned reaction norms correspond to individuals in the population after the interaction round.
@@
 \\ref{fig:modelStruc}.
+The black dashed line shows the average reaction norm of the population.
 On the first half of the simulation (B.1 and B.2) the badge size evolves to its minimum value.
@@
 
+text(x = 2000,y=0,"time:")
 axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
@@
 
+text(x = 2000,y=-3.1,"time:")
 axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
@@
 The top panels (A) show changes in the distribution of values for $\\alpha$ (A.1) and $\\beta$ (A.2) along evolutionary time.
+In the y axis the trait value, in the x axis the generation time in thousands of generations.
 Panels below correspond to snapshots of the sender (B) and receiver codes (C).
 Generation time of the snapshots are indicated by the grey lines in the top panels.
+Color of the lines indicates the quality of the individual, as indicated in the color code in C.1.
+The dashed line represents the average reaction norm of the popúlations.
 The unimodal distribution with which $\\alpha$ (A.1) starts the simulation, quickly turns into a bimondal distribution, and at about 10000 generations it splits further into three modes.
@@
 
+text(x = 2000,y=-4,"time:")
 axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
@@
 
+text(x = 2000,y=-3.2,"time:")
 axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
@@
 
+text(x = 2000,y=-1,"time:")
 axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
@@
 
+text(x = 2000,y=-2,"time:")
 axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
@@
 
+text(x = 2000,y=-0.4,"time:")
 axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
@@
 
+text(x = 2000,y=-1.5,"time:")
 axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],

### 2023-11-24 manuscript/manuscript_2.0.tex
starting revision changes

--- 2023-09-19 manuscript/manuscript_1.0.tex
@@
   \renewcommand{\thetable}{S\arabic{table}}\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
+  \providecommand{\keywords}[1]{\textbf{\textit{Keywords---}} #1}
 
@@
 \hypersetup{
-  pdftitle={ The role of learning in the evolution of status signalling: a modeling approach},
-  pdfkeywords={My keywords},
+  pdftitle={The role of learning in the evolution of status signaling: a modeling approach},
+  pdfkeywords={Biological signals, evolutionary branching, agent-based
+modeling, animal communication, badges of status, handicap principle},
   hidelinks,
@@
 
-\title{\large \textbf{Title:} The role of learning in the evolution of
-status signalling: a modeling approach}
-\usepackage{etoolbox}
-\makeatletter
-\providecommand{\subtitle}[1]{% add subtitle to \maketitle
-  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
-}
-\makeatother
-\subtitle{\normalsize \textbf{Abbreviated title:} Learning and evolution
-in status signalling}
+\title{The role of learning in the evolution of status signaling: a
+modeling approach}
 \date{}
@@
 dishonest signals.
+
+\keywords{Signals, evolutionary branching, agent-based modeling, animal communication, badges of status, handicap principle}
 
@@
 interaction partner is chosen at random. Thus, the expected number of
-interactions for each individual is \(n+\frac{n(N+1)}{N}\). After the
-interaction round , individuals in the population reproduce with a
-probability proportional to their total pay-off \(w_i\), which is a sum
-of the baseline pay-off (\(w_0\)) and all the pay-offs obtained
-throughout their life. Thus, the combination of natural selection and
-genetic drift changes the distribution of values in \(\alpha\) and
-\(\beta\) that segregates in the population, effectively changing the
-badges expressed and the communication system.
+interactions for each individual is \(2n\). After the interaction round
+, individuals in the population reproduce with a probability
+proportional to their total pay-off \(w_i\), which is a sum of the
+baseline pay-off (\(w_0\)) and all the pay-offs obtained throughout
+their life. Thus, the combination of natural selection and genetic drift
+changes the distribution of values in \(\alpha\) and \(\beta\) that
+segregates in the population, effectively changing the badges expressed
+and the communication system.
+
+The code necessary to run simulations as well as to produce
+visualizations can be found in:
+\url{https://doi.org/10.5281/zenodo.8359809}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{manuscript_1.0_files/figure-latex/modelStruc-1}
+{\centering \includegraphics[width=1\linewidth]{manuscript_2.0_files/figure-latex/modelStruc-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{manuscript_1.0_files/figure-latex/handicap-1}
+{\centering \includegraphics[width=1\linewidth]{manuscript_2.0_files/figure-latex/handicap-1}
 
@@
 Top panels (A.1 and A.2) show the evolutionary dynamics of the sender code.
+In the y axis the trait value, in the x axis the generation time in thousands of generations.
 On the left (A.1), changes in the distribution of values for the intercept of the reaction norm ($\alpha$); on the right (A.2) changes in the distribution of the slope ($\beta$).
 Darker areas of the background correspond to values with high frequency, while orange lines in both panels show the mean of the distribution.
-X axis in the evolutionary dynamics are given in thousands of generations.
-Grey lines show the generation time corresponding to the panels below portraing the sender (B) and receiver code(C).
+Grey lines show the generation time corresponding to the panels below portraing the sender (B) and receiver code (C).
 In the bottom panels (C1-4), the learned reaction norms correspond to individuals in the population after the interaction round.
@@
 \ref{fig:modelStruc}.
+The black dashed line shows the average reaction norm of the population.
 On the first half of the simulation (B.1 and B.2) the badge size evolves to its minimum value.
@@
 
-{\centering \includegraphics[width=1\linewidth]{manuscript_1.0_files/figure-latex/branching-1}
+{\centering \includegraphics[width=1\linewidth]{manuscript_2.0_files/figure-latex/branching-1}
 
@@
 The top panels (A) show changes in the distribution of values for $\alpha$ (A.1) and $\beta$ (A.2) along evolutionary time.
+In the y axis the trait value, in the x axis the generation time in thousands of generations.
 Panels below correspond to snapshots of the sender (B) and receiver codes (C).
 Generation time of the snapshots are indicated by the grey lines in the top panels.
+Color of the lines indicates the quality of the individual, as indicated in the color code in C.1.
+The dashed line represents the average reaction norm of the popúlations.
 The unimodal distribution with which $\alpha$ (A.1) starts the simulation, quickly turns into a bimondal distribution, and at about 10000 generations it splits further into three modes.
@@
 
-{\centering \includegraphics[width=1\linewidth]{manuscript_1.0_files/figure-latex/honestDiv-1}
+{\centering \includegraphics[width=1\linewidth]{manuscript_2.0_files/figure-latex/honestDiv-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{manuscript_1.0_files/figure-latex/startCond-1}
+{\centering \includegraphics[width=1\linewidth]{manuscript_2.0_files/figure-latex/startCond-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{manuscript_1.0_files/figure-latex/noBadge-1}
+{\centering \includegraphics[width=1\linewidth]{manuscript_2.0_files/figure-latex/noBadge-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{manuscript_1.0_files/figure-latex/learnEvol-1}
+{\centering \includegraphics[width=1\linewidth]{manuscript_2.0_files/figure-latex/learnEvol-1}
 

### 2023-12-06 manuscript/manuscript_2.0.Rmd
further advances in hte revision

@@
 Signal costs inversely proportional to quality may come about as a result of the
-production costs of the signal, in which case the signal works like a handicap
+production costs of the signal, in which case the signal works like a 'handicap'
 [@botero_Evolution_2010; @johnstone_Badges_1993;@grafen_Biological_1990;@zahavi_Mate_1975].
+Throughout this manuscript, we use the word 'handicap' to refer to the differential 
+costs model by Grafen [-@grafen_Biological_1990]
 Alternatively, the costs can come from 

### 2024-02-06 manuscript/manuscript_2.0.Rmd
Wrapping up new manuscript version

@@
     \usepackage{unicode-math}
+    \usepackage{geometry}
     \usepackage{setspace}\doublespacing
@@
     \providecommand{\keywords}[1]{\textbf{\textit{Keywords---}} #1}
+    \geometry{top=1in,left=1in,bottom=1in,right=1in}
   ```
@@
 Throughout this manuscript, we use the word 'handicap' to refer to the differential 
-costs model by Grafen [-@grafen_Biological_1990]
-Alternatively, the costs can come from 
-the aggressive reaction of receivers when the convention of the communication 
-is broken [@enquist_Signaling_2010;@tibbetts_Socially_2004]. For example in paper 
+costs model by Grafen [-@grafen_Biological_1990]; however, note that the 
+accuracy of this terminology has been questioned [@penn_Handicap_2020]. 
+Alternatively, the costs can come from the aggressive reaction of receivers 
+when the convention of the communication 
+is broken [@enquist_Signaling_2010;@tibbetts_Socially_2004]. For example, in paper 
 wasps, subordinate individuals with experimentally manipulated dominant-like facial
@@
 trials which allows them to have more 
-flexible learning [@leimar_Flexibility_2023], and some may include future rewards in their associations [@raby_Planning_2007;@osvath_Spontaneous_2009].
+flexible learning [@leimar_Flexibility_2023], and some may include 
+future rewards in their associations [@raby_Planning_2007;@osvath_Spontaneous_2009].
 Associative learning is not a single mechanism, but rather a set of 
-cognitive structures and processes that can vary in their scope and complexity. Presumably, these structures and processes have been modified 
+cognitive structures and processes that can vary in their scope and complexity. 
+Presumably, these structures and processes have been modified 
 by natural selection and could provide novel explanations for behavioural 
@@
 For simplicity, we consider a population of haploid 
-individuals with non-overlapping generations. Individuals are born 
-every generation with a level of quality ($Q_i$, where $i$ is subscript 
-of the population vector of size $N$) given by a number between zero and one, 
+individuals with non-overlapping generations of size $N=3000$. Individuals are born 
+every generation with a level of quality $Q_i$ (where $i$ is subscript 
+of the population vector) given by a number between zero and one, 
 which is drawn from a truncated normal distribution 
-$N( 0.5,\sigma)$. An individual with quality $0$ has the lowest RHP, 
+$TN( 0.5,\sigma)$. An individual with quality $0$ has the lowest RHP, 
 while an individual with quality $1$ has the highest. As they 
 develop, individuals produce a phenotypic signal (badge), 
-the size of which depends on their quality according to a reaction norm given by 
-$B_i=1/(1+e^{\alpha_i-\beta_i Q_i})$; where $\alpha_i$ and 
-$\beta_i$ are individual specific traits that determine the shape of 
-the reaction norm (Fig. \ref{fig:modelStruc} A). Variation in $\alpha_i$ 
+the size of which depends on their quality according to a logistic reaction norm given by
+\begin{equation}
+B_i=1/(1+e^{-\alpha_i-\beta_i Q_i}) \textrm{;}\label{eq:reacNorm}
+\end{equation}
+
+where $\alpha_i$ and $\beta_i$ are individual specific traits 
+that determine the shape of the reaction norm (Fig. \ref{fig:modelStruc} 
+A). Variation in $\alpha_i$ 
 and $\beta_i$ means individuals can have either uninformative (flat) or 
@@
 unless, with a small probability ($\mu$), mutation changes the allelic value 
-of the offspring by an amount drawn from a normal distribution $TN(0,\sigma_\mu)$.
+of the offspring by an amount drawn from a normal distribution $N(0,\sigma_\mu)$.
 
@@
 viability selection. The survival probability of an individual is 
-given by $s_i=1/(1+e^{-k_1-k_2(Q_i-B_i)})$, where $k_1$ and $k_2$ are parameter 
+given by
+
+\begin{equation}
+s_i=1/(1+e^{-k_1-k_2(Q_i-B_i)})\textrm{,} \label{eq:viability}
+\end{equation}
+
+where $k_1$ and $k_2$ are parameter 
 values determining the shape of the survival function. 
@@
 The decision of whether to escalate a fight against an interacting 
-partner can depend on the size of the partner's badge, 
-with the dependency being determined by the focal's experiences acquired
-through a learning process. We implement learning using 
-the actor-critic approach from reinforcement learning (RL) theory [@sutton_Reinforcement_2018;@quinones_Reinforcement_2020; @leimar_Learning_2019;
-@leimar_Evolution_2021].
-Individuals estimate the reward (pay-off) expected from  interacting 
-with partners of different badge sizes (the critic in RL terminology). 
-After each interaction they update the estimate of reward proportionally 
-to the difference between their current estimate and the observed 
-reward (prediction error $\delta$) and to the speed of learning 
-($\mathrm{A}$). Furthermore, individuals express different probabilities of 
-retreating/attacking depending on the badge size of their opponent 
-(the actor in RL). They update the probability of 
-retreating/attacking depending on whether retreating/attacking
-leads to an increase in the reward estimation. 
-Thus, if a focal individual decides to escalate a fight against 
-an individual with a small badge and this leads to an increase in 
-the reward estimation, then the focal individual will increase the probability 
-of escalating fights with individuals of small badges in the future. 
-Given that badge size is a real number between 0 and 1, there are infinitely 
-many badge sizes. Thus, the reward estimation, as well as the probability 
-of retreating/attacking, must be generalized across different values. 
+partner depends on the size of the partner's badge, with the dependency 
+being determined by the focal's experiences acquired
+through a learning process. Individuals use previous experiences to estimate
+the reward expected from encountering a partner with a particular badge size. 
+However, given that badge size is a real number between 0 and 1, previous 
+experiences must be generalized across the badge size range. 
 To implement generalization we use the linear function approximation 
@@
 constant and  the same for all individuals; they are stored in 
-vector $\symbf{b}$ . Each of these feature centres 
-is associated with a weight for reward estimation and tendency to play 
-retreat in a given interaction. The reward estimation and the tendency 
-to retreat are calculated (in every interaction) as the sum of the weights 
-associated with each feature centre, weighted by the response 
-triggered by the feature (Fig. \ref{fig:modelStruc} B, dots 
-represent the feature weights). The response of each feature centre 
+vector $\symbf{b}$. Each of these feature centres 
+is associated with a weight for reward estimation 
+($x_z$, $z$ is an index that goes from 1 to $c$) and tendency to play 
+retreat in a given interaction ($y_z$). For each badge size ecountered the reward 
+estimate and the associated tendency to retreat are calculated 
+as the sum of the weights associated with each feature centre, 
+weighted by the response triggered by the feature (Fig. \ref{fig:modelStruc} B, 
+dots  represent the feature weights). The response of each feature centre 
 diminishes as a Gaussian function with the distance between the 
@@
 individual $j$ is given by,
-\
+
 \begin{equation}
@@
 
+The weights vectors for reward estimation ($\symbf{x}$) and tendency 
+to retreat ($\symbf{y}$) are the focus of the learning update. In the baseline
+implementation of the model, these two vectors are initialized with values
+of zero. For the reward estimation ($\symbf{x}$), this means that
+before any interaction individuals assume they will get a payoff of 
+zero. As for the behavioral policy, these initial values
+translate into a $0.5$ probability of playing dove. Beyond the baseline scenario,
+we explore initializing the policy vector with a) more positive values triggering 
+more peaceful behaviour, b) more negative values triggering more aggressive 
+behaviour, c) values that trigger the ESS of the original *hawk-dove* game and 
+d) letting these values be defined by evolutionary processes 
+(see Results for more details). The later was 
+implemented by letting an extra locus pass on the initial value from parent to 
+offspring and changed by mutation. Regardless of the initialization, all these 
+values change with every interaction as a function of the reward experienced. 
+Formally the change is given by,
+
+\begin{equation}
+  \Delta x_z = \Alpha \delta e^{(-\frac{|B_j-b_z|}{2\theta^2})} \textrm{.}
+  \label{eq:critic}
+\end{equation}
+
+Where $\delta$ is the *prediction error*, which is the difference between the 
+predicted and experienced reward, formally
+
+\begin{equation}
+  \delta = R - \hat{R} \textrm{.} \label{eq:predError}
+\end{equation}
+
+The experienced reward ($R$) is the outcome from the *hawk-dove* 
+game described earlier.
+
+Thus, the learning update depends on the speed of learning ($A$) 
+and the prediction error. The exponential term assures, just as in 
+the reward estimation, that the size of update is weighted by 
+the distance of the partner's badge size to each feature centre.
+
+The weights for the probability of retreat ($\symbf{z}$) are updated in a similar
+fashion according to
+
+\begin{equation}
+  \Delta y_z = \Alpha \delta (1-\pi) e^{(-\frac{|B_j-b_z|}{2\theta^2})} \textrm{.}
+  \label{eq:actor}
+\end{equation}
+
+where $\pi$ is the probability of choosing the action chosen by the individual. 
+If an individual chooses to behave as a dove and avoid an overt conflict,
+
+\begin{equation}
+  \pi = q_{ij} = \frac{1}{1+e^{\sum_{z=1}^{c} y_z e^{-\frac{|B_i-b_z|}{2\theta^2}}}} \textrm{;}\label{eq:pi}
+\end{equation}
+
+if alternatively, the individual chooses to escalate the fight, $\pi$ corresponds 
+to $1-q_{ij}$. The combination of $\pi$ and the *prediction error* ($\delta$)
+assures that behavioral policy updates are carried increasing the probability of 
+behaviours that predict higher reward.
+
+This implementation corresponds to the *actor-critic* approach from reinforcement learning (RL) theory [@sutton_Reinforcement_2018;@quinones_Reinforcement_2020; @leimar_Learning_2019;@leimar_Evolution_2021].
+In the *actor-critic* approach
+the agent updates its behavioral policy (*actor*) and in parallel its 
+reward estimation (*critic*). In our model the actor update is implemented by equations \ref{eq:actor} and \ref{eq:pi}; while the *critic* update 
+is implemented by equation \ref{eq:critic}. 
+Both of these updates include the *prediction error* 
+(equation \ref{eq:predError}). The *prediction error* corresponds to the Rescorla-Wagner 
+rule [@rescorla_Theory_1972], which is a key component of learning 
+theory and widely applied in neurobiology to understand learning processes
+[@holland_Minireview_2016].
+
 Individuals interact as the focal individual $n$ times and the interaction 
@@
 each individual is $2n$. 
-After the interaction round , individuals in the population 
-reproduce with a probability proportional to their total pay-off $w_i$, which is a sum of the 
-baseline pay-off ($w_0$) and all the pay-offs obtained throughout their 
-life. Thus, the combination of natural selection and genetic drift 
-changes the distribution of values in $\alpha$ and $\beta$ that segregates 
+After the interaction round, individuals in the population 
+reproduce with a probability proportional to their total pay-off $w_i$, 
+which is a sum of the baseline pay-off ($w_0$) and all the pay-offs 
+obtained from interactions throughout their life. Thus, the total fitness
+is composed of the viability component (equation \ref{eq:viability}) and the 
+total payoff collected in the interaction round. Therefore, the combination of natural 
+selection and genetic drift changes the distribution of 
+values in $\alpha$ and $\beta$ that segregates 
 in the population, effectively changing  the badges expressed and the 
@@
 
-The code necessary to run simulations as well as to produce visualizations can be found 
-in: \url{https://doi.org/10.5281/zenodo.8359809}
+The code necessary to run simulations as well as to produce visualizations 
+can be found in: \url{https://doi.org/10.5281/zenodo.8359809}
 
 ```{r modelStruc,echo=FALSE,out.width='100%',fig.height=5,fig.align='center',fig.show='hold',fig.cap= "Model of communication in the context of aggressive interactions.
-In A, the reaction norm determines the badge size, *e.g* the size of a black band on a bird's chest.
+In A, the reaction norm  (equation \\ref{eq:reacNorm}) determines the badge size, \\textit{e.g} the size of a black band on a bird's chest.
 Red shows an informative reaction norm, whereas blue shows a uninformative reaction norm.
 B individuals perceiving the signal have a behavioral reaction norm that determines their probability of retreating (black line).
-The reaction norm arises from generalizing the information from the feature weights (black dots).
+The reaction norm arises from generalizing the information from the policy feature weights (black dots).
 Feature weights are values associated with a particular badge size and are updated as individuals interact with each other.
@@
 
-```{r handicap,echo=FALSE,out.width='100%',fig.height=7,fig.align='center',fig.show='hold',fig.cap="Evolution of the badge size as a handicap mediated by learning.
+```{r handicap,echo=FALSE,out.width='100%',fig.height=6,fig.align='center',fig.show='hold',fig.cap="Evolution of the badge size as a handicap mediated by learning.
 Top panels (A.1 and A.2) show the evolutionary dynamics of the sender code.
-In the y axis the trait value, in the x axis the generation time in thousands of generations.
+In y axis shows the trait value, in the x axis shows the generation time in thousands of generations.
 On the left (A.1), changes in the distribution of values for the intercept of the reaction norm ($\\alpha$); on the right (A.2) changes in the distribution of the slope ($\\beta$).
@@
 In the second half, the slope takes positive values and reaction norms evolve to produce an honest signal of quality.
-Correspondingly, individuals learn to react to badge size by increasing the probability of retreating with increasing badge sizes"  ,warning=FALSE}
+Correspondingly, individuals learn to react to badge size by increasing the probability of retreating with increasing badge sizes."  ,warning=FALSE}
 
@@
 pop<-do.call(rbind,lapply(indList_runs, fread))
+
+evol[,meanAlpha:=meanAlpha*-1]
+pop[,alpha:=alpha*-1]
 
@@
 
-mtext(text = "Trait value",cex = cexAxis,line = 1,side = 2,las=0)
+mtext(text = "Trait value",cex = cexAxis,line = 1.2,side = 2,las=0)
 legend("topright",legend = c(expression(alpha[s])), bty = "n",cex=cexAxis)
@@
 
-text(x = 2000,y=0,"time:")
+text(x = 2000,y=-4.2,"time:")
 axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
@@
 
-### Badges as Handicaps
+### Badges as Handicaps (costly signals)
 
@@
 norm parameters (Fig. \ref{fig:handicap}, A) appeared to involve a set of 
-steps. First, the intercept of the reaction norm ($\alpha$) 
+steps. First, the intercept of the logistic reaction norm ($\alpha$) 
 describing the relationship between individual quality and badge size evolved to 
-higher values, reducing the average badge size in the populations. 
+lower values, reducing the average badge size in the populations. 
 This is because reaction norms segregating in the population are 
@@
 maximum badge size. In either of these cases reaction norms are flat, so the 
-badge does not provide any information about quality.
-
-### The evolution of polymorphism mediated by learning
+badge does not provide any information about quality. This means that learning
+is not sufficient to yield a communication system, but rather is requires 
+the badge to have a cost proportional to the quality of the individual.
+
+We have described here  two replicate simulations representative of the 
+two outcomes found in a series of 20 replicates run with the same parameter
+combination. The exact evolutionary trajectory differs in each of these replicates 
+due to the stochastic effects of the dynamical system. Particularly the timing 
+of the changes was different in each replicate. Nevertheless, all of the 
+replicates could be classified as either evolving a functioning communication 
+system, or not having the badge present in the population.
+
+### The evolution of polymorphism mediated by learning when the signal is cost-free
 
@@
 value. Within the first 2000 generations, the unimodal distribution in 
-the intercept ($\alpha$) splits into a bimodal one. 
+the logistic intercept ($\alpha$) splits into a bimodal one. 
 Later in evolutionary time, one of the peaks splits further, 
-so at the end of the evolutionary simulation the distribution of the intercept 
-($\alpha$) in the population shows three distinct peaks. In the case of $\beta$, the 
-peaks in the distribution are not so clear-cut, but the variance of the 
-distribution increases over time. These changes in the 
-distribution of the parameters of the sender reaction norm 
+so at the end of the evolutionary simulation the distribution of the 
+logistic intercept ($\alpha$) in the population shows three distinct peaks. 
+In the case of $\beta$, the peaks in the distribution are not so clear-cut, 
+but the variance of the distribution increases over time. 
+These changes in the distribution of the parameters of the sender reaction norm 
 imply that individuals can generally be classified into three 
@@
 (Fig. \ref{fig:branching} B.3-4).
+
 In this simulation, the values of the slope in the informative group
 are negative, which implies that badge size correlates negatively with quality. 
-Furthermore, the receiver reaction norms developed through learning, particularly those 
+The receiver reaction norms developed through learning, particularly those 
 of individuals of intermediate quality, respond to the signal of their 
 sender type by increasing the probability of retreating from a fight with 
-individuals with smaller badges (Fig. \ref{fig:branching}).
+individuals with smaller badges (Fig. \ref{fig:branching}). However, 
+given the absence of costs, the sign of the slope is arbitrary. In our replicate 
+simulations we found examples where the third type had positive 
+and negative slopes.
 
@@
 The top panels (A) show changes in the distribution of values for $\\alpha$ (A.1) and $\\beta$ (A.2) along evolutionary time.
-In the y axis the trait value, in the x axis the generation time in thousands of generations.
+The y axis shows the trait value, and  the x axis shows the generation time in thousands of generations.
 Panels below correspond to snapshots of the sender (B) and receiver codes (C).
@@
 Color of the lines indicates the quality of the individual, as indicated in the color code in C.1.
-The dashed line represents the average reaction norm of the popúlations.
+The dashed line represents the average reaction norm of the populations.
 The unimodal distribution with which $\\alpha$ (A.1) starts the simulation, quickly turns into a bimondal distribution, and at about 10000 generations it splits further into three modes.
@@
 
-text(x = 2000,y=-3.2,"time:")
+text(x = 2000,y=-3.3,"time:")
 axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
@@
 of the individual. This relationship, however, is muddled by the fact that there 
-are two clusters of values for the intercept ($\alpha$) and slope ($\beta$) 
-of the reaction norm in the population. Thus, there are two types 
-of reaction norms. One of the types has a steeper slope compared to the other. 
+are two clusters of values for the logistic intercept 
+($\alpha$) and slope ($\beta$) of the reaction norm in the population. 
+Thus, there are two types of reaction norms. 
+One of the types has a steeper slope compared to the other. 
 This effect of an increased variance in the trait distribution
@@
 
-text(x = 2000,y=0,"time:")
+text(x = 2000,y=-4,"time:")
 axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
@@
 
-text(x = 2000,y=-1,"time:")
+text(x = 2000,y=-1.2,"time:")
 axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
@@
 
-### The peaceful, the aggressive and the clever
+### The peaceful, the aggressive and the clever, when signals are cost-free
 
@@
 corresponding to the ESS of the original hawk-dove game 
-(which we call clever, due to information and computation necessary to know the ESS). 
+(which we call clever, due to information and computation 
+necessary to know the ESS). 
 Fig. \ref{fig:startCond} shows the distribution of values of the 
@@
 distribution of values is split in different clusters. That is, most of 
-the variation occurs within clusters. In contrast, when naive individuals behave
+the variation occurs between clusters of each replicate. 
+In contrast, when naive individuals behave
 either aggressively or cleverly, badges evolve toward minimum and maximum 
@@
 with a rare badge size have a fitness advantage. This triggers 
-negative-frequency dependent selection and the evolution of different types of 
+negative frequency-dependent selection and the evolution of different types of 
 badge sizes. According to this narrative, situations where individuals 
@@
 ```{r startCond, echo=FALSE, out.width='100%',fig.align='center',fig.show='hold',fig.cap="The peaceful, the aggressive and the clever.
-Distribution of values of the intercept $\\alpha$ and slope $\\beta$ for individuals at the end of the evolutionary simulations.
+Distribution of values of the logistic intercept $\\alpha$ and slope $\\beta$ for individuals at the end of the evolutionary simulations.
 The panels show the three different initial conditions for the behaviour of individuals: peaceful, agressive and  \"clever\", see maintext for details.
@@
 
+inds[,negAlpha:=alpha*-1]
 inds[,diffActWeight:=abs(WeightAct_0-WeightAct_4)]
 
-shortSce<-gsub("[^[:alpha:]]",gsub(".txt","",tail(strsplit(indList[1],"_")[[1]],2)[1]),
+shortSce<-gsub("[^[:alpha:]]",
+               gsub(".txt","",tail(strsplit(indList[1],"_")[[1]],2)[1]),
                replacement = "")
@@
       cex.axis=1.5,new=TRUE)
-  plot(data=popOneInd[get(shortSce)==PAr&time>max(time)*0.8],alpha~beta,ylab="",
+  plot(data=popOneInd[get(shortSce)==PAr&time>max(time)*0.8],negAlpha~beta,ylab="",
        xlab="", pch=20,cex.lab=cexAxis,cex.axis=cexAxis,las=1,cex=1,
-       ylim=range(popOneInd[,alpha])*c(1.1,1.05),
+       ylim=range(popOneInd[,negAlpha])*c(1.1,1.2),
        xlim=range(popOneInd[,beta]),col=multDiscrPallet[seed+1])
-  lines(x=c(0,0),y=range(popOneInd[,alpha]),col="grey",
+  lines(x=c(0,0),y=range(popOneInd[,negAlpha]),col="grey",
         lwd=2)
@@
                           seed == repsScen[count]],
-       alpha~beta,col=multDiscrPallet[seed+1],cex=1,pch=20)
+       negAlpha~beta,col=multDiscrPallet[seed+1],cex=1,pch=20)
 
@@
   par(new=TRUE,plt=posPlot(numplotx = 6,numploty = 4,idplotx = count*2-1,
-                           idploty = 1,upboundx = 96,lowboundx =11,
-                           upboundy = 98)
-      +c(0.02,0.02,0.03,0.03),xpd=T,xaxt="s",yaxt="s")
+                           idploty = 3,upboundx = 96,lowboundx =11,
+                           upboundy = 98)+c(0.0,0.0,0.13,0.13),
+      xpd=T,xaxt="s",yaxt="s")
 
@@
   axis(1,at = c(0,1),tick = TRUE,cex.axis=.7,padj=-1.7)
-  axis(2,at =c(0,1),tick = TRUE,cex.axis=.7,hadj=-0.2)
+  axis(4,at =c(0,1),tick = TRUE,cex.axis=.7,hadj=-0.2)
 }
@@
 
-The action of natural selection and genetic drift on the learning parameters 
-do not override the conditions for the evolutionary branching 
-[@dieckmann_Dynamical_1996] described above. 
 Given how crucial naive behaviour is for the branching event 
-described earlier, we tested whether evolutionary processes could lead 
-key parameters of the learning process toward
-values that prevented the genetic variation to build up. 
-Figure \ref{fig:learnEvol} shows the dynamics from a set of simulations where we 
-allow for the coevolution of the reaction norm as well as key parameters of
-the learning module, namely the speed of learning ($A$) and the behavioural 
-tendency expressed by individuals prior to learning. Top panels in 
-figure \ref{fig:learnEvol} show that there is diversification process in this 
-coevolutionary scenario as well. That is because, as is shown in the middle 
-panels (B) of the same figure, the learning parameters stay within the range 
+described earlier, we tested whether the occurrence of branching was robust to 
+the actions of natural selection. If natural selection favoured high 
+learning rates or extreme values in the initial policy vector, then
+the evolutionary branching we find would not be as relevant in 
+biological populations. Thus, we developed an extension of the model 
+where the speed of learning ($\Alpha$) and the initial value of the 
+policy vector $\symbf{y}$ are allowed to change
+subject to natural selection and genetic drift. We do this by assuming 
+that the values of the learning parameters are inherited from parent to 
+offspring and occasionally undergo mutation, just like the parameters 
+of the reaction norm. The action of natural selection and genetic drift 
+on the learning parameters do not override the conditions 
+for the evolutionary branching [@dieckmann_Dynamical_1996] described above. 
+Top panels in figure \ref{fig:learnEvol} show that there is diversification 
+process in this coevolutionary scenario as well. That is because, 
+as is shown in the middle panels (B) of the same figure, 
+the learning parameters stay within the range 
 necessary for the branching event to take place. The speed of learning 
@@
 We showed that the learning process can mediate the evolution 
-of an honest signal under the handicap principle
-[@grafen_Biological_1990;@zahavi_Mate_1975]. This conclusion is not 
+of an honest signal under the handicap principle, but it cannot without it
+[@grafen_Biological_1990;@zahavi_Mate_1975]. Learning is not sufficient to 
+trigger the evolution of the communication system.  This conclusion is not 
 different from classical genetic models, where responses to signals are
@@
 and maintenance of variation in the signalling system. 
-Namely the polymorphism required
+Namely, the polymorphism required
 a limit in the amount of information collected by individuals, as well as
@@
 
-The focus of theoretical and empirical work on communication systems, 
-and particularly badges of status, is often explaining the presence and 
-absence of certain morphological traits within populations. Phenotypic traits
+The focus of classical theoretical and empirical work on communication systems
+is often explaining the presence and absence of certain 
+morphological traits (ornaments) within populations. Phenotypic traits
 hypothesized to play a role in communication systems regularly exhibit 
@@
 
-mtext(text = "Trait value",cex = cexAxis,line = 1,side = 2,las=0)
+mtext(text = "Trait value",cex = cexAxis,line = 1.2,side = 2,las=0)
 legend("topright",legend = c(expression(alpha[s])),bty = "n",cex=cexAxis)
@@
 
+evol[,meanAlpha:=meanAlpha*-1]
+pop[,alpha:=alpha*-1]
+
 Valpar<-param$rangParam
@@
 
-| Symbol |      Description |
-|:-------:|:--------------------------------|
-| $Q_i$ | Quality of individual $i$ |
-| $N$   | Population size |
+| Symbol |      Description |      Value |
+|:-------:|:--------------------------------|:----------------|
+| $Q_i$ | Quality of individual $i$ |                         |
+| $N$   | Population size | $3000$                            |
 | $n$   | Number of interactions as the focal individual |
-| $\sigma$ | Standard deviation of the truncated normal distribution from which quality is drawn |
-| $B_i$ | Badge size of individual $i$ |
-| $\alpha_i$ | Intercept of the badge size reaction norm for individual $i$ |
+| $\sigma$ | Standard deviation of the truncated normal distribution from which quality is drawn |    $0.15$                                             |
+| $B_i$ | Badge size of individual $i$ |                      |
+| $\alpha_i$ | Intercept of the badge size logistic reaction norm for individual $i$ |
+                                                              |
 | $\beta_i$  | Slope of the badge size reaction norm for individual $i$ |
-| $\mu$     | Mutation rate | 
-| $\sigma_\mu$ |Standard deviation of the normal distribution from which quality is mutations are drawn |
-| $s_i$ | Survival probability of individual $i$ |
-| $k_1$ | Intercept of the survival function |
-| $k_2$ | Slope of the survival function |
-| $V$   | Value of the contested resource |
-| $C$   | Cost of an escalated fight |
-| $p_{ij}$ | Probability that individual $i$ wins an escalated fight against individual $j$ |
-| $k_3$ | Parameter determining how important is quality defining the probability of wining |
-| $\delta$ | Prediction error |
-| $A$ | Speed of learning | 
-| $c$ | Number of feature centres |
-| $b_z$ | badge size for feature centre $z$  | 
-| $\hat{R}$ | Reward estimate | 
+                                                              |
+| $\mu$     | Mutation rate | $0.001$                         |
+| $\sigma_\mu$ |Standard deviation of the normal distribution from which quality is mutations are drawn | $0.3$                                   |
+| $s_i$ | Survival probability of individual $i$ |            |
+| $k_1$ | Intercept of the survival function |  $3$           |     
+| $k_2$ | Slope of the survival function |                    |
+| $V$   | Value of the contested resource | $1$               |
+| $C$   | Cost of an escalated fight |    $1.5$               |
+| $p_{ij}$ | Probability that individual $i$ wins an escalated fight against individual $j$ |                                                             |
+| $k_3$ | Parameter determining how important quality is defining the probability of wining | $10$                                                        |
+| $\delta$ | Prediction error |                               |
+| $A$ | Speed of learning |                                   |
+| $c$ | Number of feature centres | $6$                       |   
+| $b_z$ | badge size for feature centre $z$  |                |          
+| $\hat{R}$ | Reward estimate |                               |
 | $x_z$ | weight of feature centre $z$ on the reward estimation |
-| $\theta$ | width of the generalization function | 
-| $q_{ij}$ | probability that individual $i$ retreats from a fight in an encounter with individual $j$ |
-| $y_z$ | weight of feature centre z on the tendency to retreat |
-| $w_i$ | total pay-off of individual $i$ |
-| $w_0$ | base line pay-off |
-
-Table: Notation of the model parameters and variables
+                                                              |
+| $\theta$ | width of the generalization function | $0.01$   |          
+| $q_{ij}$ | probability that individual $i$ retreats from a fight in an encounter with individual $j$ |                                              |
+| $y_z$ | weight of feature centre $z$ on the tendency to retreat |
+                                                              |
+| $w_i$ | total pay-off of individual $i$ |                   |
+| $w_0$ | base line pay-off | $2$                             |
+
+Table: Notation of the model parameters and variables. Values for the parameters
+that are kept constant throughout all simulations are given in the third column.
 

### 2024-02-06 manuscript/manuscript_2.0.tex
Wrapping up new manuscript version

@@
   \usepackage{unicode-math}
+  \usepackage{geometry}
   \usepackage{setspace}\doublespacing
@@
   \providecommand{\keywords}[1]{\textbf{\textit{Keywords---}} #1}
+  \geometry{top=1in,left=1in,bottom=1in,right=1in}
 
@@
 about as a result of the production costs of the signal, in which case
-the signal works like a handicap (Botero et al. 2010; Johnstone and
-Norris 1993; Grafen 1990; Zahavi 1975). Alternatively, the costs can
+the signal works like a `handicap' (Botero et al. 2010; Johnstone and
+Norris 1993; Grafen 1990; Zahavi 1975). Throughout this manuscript, we
+use the word `handicap' to refer to the differential costs model by
+Grafen (1990); however, note that the accuracy of this terminology has
+been questioned (Penn and Számadó 2020). Alternatively, the costs can
 come from the aggressive reaction of receivers when the convention of
 the communication is broken (Enquist, Ghirlanda, and Hurd 2010; Tibbetts
-and Dale 2004). For example in paper wasps, subordinate individuals with
-experimentally manipulated dominant-like facial patterns received more
-aggression from the dominant (Tibbetts and Dale 2004). In this case,
-costs are socially rather than developmentally mediated and are
+and Dale 2004). For example, in paper wasps, subordinate individuals
+with experimentally manipulated dominant-like facial patterns received
+more aggression from the dominant (Tibbetts and Dale 2004). In this
+case, costs are socially rather than developmentally mediated and are
 triggered by the violation of the convention, although how the
@@
 context of agonistic interactions. For simplicity, we consider a
-population of haploid individuals with non-overlapping generations.
-Individuals are born every generation with a level of quality (\(Q_i\),
-where \(i\) is subscript of the population vector of size \(N\)) given
-by a number between zero and one, which is drawn from a truncated normal
-distribution \(N( 0.5,\sigma)\). An individual with quality \(0\) has
-the lowest RHP, while an individual with quality \(1\) has the highest.
-As they develop, individuals produce a phenotypic signal (badge), the
-size of which depends on their quality according to a reaction norm
-given by \(B_i=1/(1+e^{\alpha_i-\beta_i Q_i})\); where \(\alpha_i\) and
-\(\beta_i\) are individual specific traits that determine the shape of
-the reaction norm (Fig. \ref{fig:modelStruc} A). Variation in
-\(\alpha_i\) and \(\beta_i\) means individuals can have either
-uninformative (flat) or informative (logistic) norms. Informative
+population of haploid individuals with non-overlapping generations of
+size \(N=3000\). Individuals are born every generation with a level of
+quality \(Q_i\) (where \(i\) is subscript of the population vector)
+given by a number between zero and one, which is drawn from a truncated
+normal distribution \(TN( 0.5,\sigma)\). An individual with quality
+\(0\) has the lowest RHP, while an individual with quality \(1\) has the
+highest. As they develop, individuals produce a phenotypic signal
+(badge), the size of which depends on their quality according to a
+logistic reaction norm given by \begin{equation}
+B_i=1/(1+e^{-\alpha_i-\beta_i Q_i}) \textrm{;}\label{eq:reacNorm}
+\end{equation}
+
+where \(\alpha_i\) and \(\beta_i\) are individual specific traits that
+determine the shape of the reaction norm (Fig. \ref{fig:modelStruc} A).
+Variation in \(\alpha_i\) and \(\beta_i\) means individuals can have
+either uninformative (flat) or informative (logistic) norms. Informative
 reaction norms represent a developmental program where the phenotype of
@@
 offspring by an amount drawn from a normal distribution
-\(TN(0,\sigma_\mu)\).
+\(N(0,\sigma_\mu)\).
 
@@
 survival probability of an individual is given by
-\(s_i=1/(1+e^{-k_1-k_2(Q_i-B_i)})\), where \(k_1\) and \(k_2\) are
-parameter values determining the shape of the survival function.
-Importantly, if \(k_2=0\) survival probability is independent of
-quality, while if \(k_2>0\) lower-quality individuals pay a higher price
-for similar-sized badges, fulfilling the core assumption of the handicap
-principle (Botero et al. 2010; Grafen 1990; Johnstone and Norris 1993).
+
+\begin{equation}
+s_i=1/(1+e^{-k_1-k_2(Q_i-B_i)})\textrm{,} \label{eq:viability}
+\end{equation}
+
+where \(k_1\) and \(k_2\) are parameter values determining the shape of
+the survival function. Importantly, if \(k_2=0\) survival probability is
+independent of quality, while if \(k_2>0\) lower-quality individuals pay
+a higher price for similar-sized badges, fulfilling the core assumption
+of the handicap principle (Botero et al. 2010; Grafen 1990; Johnstone
+and Norris 1993).
 
@@
 The decision of whether to escalate a fight against an interacting
-partner can depend on the size of the partner's badge, with the
-dependency being determined by the focal's experiences acquired through
-a learning process. We implement learning using the actor-critic
-approach from reinforcement learning (RL) theory (Sutton and Barto 2018;
-Quiñones et al. 2020; Leimar and McNamara 2019; Leimar 2021).
-Individuals estimate the reward (pay-off) expected from interacting with
-partners of different badge sizes (the critic in RL terminology). After
-each interaction they update the estimate of reward proportionally to
-the difference between their current estimate and the observed reward
-(prediction error \(\delta\)) and to the speed of learning
-(\(\mathrm{A}\)). Furthermore, individuals express different
-probabilities of retreating/attacking depending on the badge size of
-their opponent (the actor in RL). They update the probability of
-retreating/attacking depending on whether retreating/attacking leads to
-an increase in the reward estimation. Thus, if a focal individual
-decides to escalate a fight against an individual with a small badge and
-this leads to an increase in the reward estimation, then the focal
-individual will increase the probability of escalating fights with
-individuals of small badges in the future. Given that badge size is a
-real number between 0 and 1, there are infinitely many badge sizes.
-Thus, the reward estimation, as well as the probability of
-retreating/attacking, must be generalized across different values. To
+partner depends on the size of the partner's badge, with the dependency
+being determined by the focal's experiences acquired through a learning
+process. Individuals use previous experiences to estimate the reward
+expected from encountering a partner with a particular badge size.
+However, given that badge size is a real number between 0 and 1,
+previous experiences must be generalized across the badge size range. To
 implement generalization we use the linear function approximation method
@@
 simplicity, we keep the location of these feature centres constant and
-the same for all individuals; they are stored in vector \(\symbf{b}\) .
+the same for all individuals; they are stored in vector \(\symbf{b}\).
 Each of these feature centres is associated with a weight for reward
-estimation and tendency to play retreat in a given interaction. The
-reward estimation and the tendency to retreat are calculated (in every
-interaction) as the sum of the weights associated with each feature
-centre, weighted by the response triggered by the feature (Fig.
+estimation (\(x_z\), \(z\) is an index that goes from 1 to \(c\)) and
+tendency to play retreat in a given interaction (\(y_z\)). For each
+badge size ecountered the reward estimate and the associated tendency to
+retreat are calculated as the sum of the weights associated with each
+feature centre, weighted by the response triggered by the feature (Fig.
 \ref{fig:modelStruc} B, dots represent the feature weights). The
@@
 estimation \(\hat{R}\) when the focal individual (\(i\)) faces
-individual \(j\) is given by,\\
+individual \(j\) is given by,
+
 \begin{equation}
@@
 
+The weights vectors for reward estimation (\(\symbf{x}\)) and tendency
+to retreat (\(\symbf{y}\)) are the focus of the learning update. In the
+baseline implementation of the model, these two vectors are initialized
+with values of zero. For the reward estimation (\(\symbf{x}\)), this
+means that before any interaction individuals assume they will get a
+payoff of zero. As for the behavioral policy, these initial values
+translate into a \(0.5\) probability of playing dove. Beyond the
+baseline scenario, we explore initializing the policy vector with a)
+more positive values triggering more peaceful behaviour, b) more
+negative values triggering more aggressive behaviour, c) values that
+trigger the ESS of the original \emph{hawk-dove} game and d) letting
+these values be defined by evolutionary processes (see Results for more
+details). The later was implemented by letting an extra locus pass on
+the initial value from parent to offspring and changed by mutation.
+Regardless of the initialization, all these values change with every
+interaction as a function of the reward experienced. Formally the change
+is given by,
+
+\begin{equation}
+  \Delta x_z = \Alpha \delta e^{(-\frac{|B_j-b_z|}{2\theta^2})} \textrm{.}
+  \label{eq:critic}
+\end{equation}
+
+Where \(\delta\) is the \emph{prediction error}, which is the difference
+between the predicted and experienced reward, formally
+
+\begin{equation}
+  \delta = R - \hat{R} \textrm{.} \label{eq:predError}
+\end{equation}
+
+The experienced reward (\(R\)) is the outcome from the \emph{hawk-dove}
+game described earlier.
+
+Thus, the learning update depends on the speed of learning (\(A\)) and
+the prediction error. The exponential term assures, just as in the
+reward estimation, that the size of update is weighted by the distance
+of the partner's badge size to each feature centre.
+
+The weights for the probability of retreat (\(\symbf{z}\)) are updated
+in a similar fashion according to
+
+\begin{equation}
+  \Delta y_z = \Alpha \delta (1-\pi) e^{(-\frac{|B_j-b_z|}{2\theta^2})} \textrm{.}
+  \label{eq:actor}
+\end{equation}
+
+where \(\pi\) is the probability of choosing the action chosen by the
+individual. If an individual chooses to behave as a dove and avoid an
+overt conflict,
+
+\begin{equation}
+  \pi = q_{ij} = \frac{1}{1+e^{\sum_{z=1}^{c} y_z e^{-\frac{|B_i-b_z|}{2\theta^2}}}} \textrm{;}\label{eq:pi}
+\end{equation}
+
+if alternatively, the individual chooses to escalate the fight, \(\pi\)
+corresponds to \(1-q_{ij}\). The combination of \(\pi\) and the
+\emph{prediction error} (\(\delta\)) assures that behavioral policy
+updates are carried increasing the probability of behaviours that
+predict higher reward.
+
+This implementation corresponds to the \emph{actor-critic} approach from
+reinforcement learning (RL) theory (Sutton and Barto 2018; Quiñones et
+al. 2020; Leimar and McNamara 2019; Leimar 2021). In the
+\emph{actor-critic} approach the agent updates its behavioral policy
+(\emph{actor}) and in parallel its reward estimation (\emph{critic}). In
+our model the actor update is implemented by equations \ref{eq:actor}
+and \ref{eq:pi}; while the \emph{critic} update is implemented by
+equation \ref{eq:critic}. Both of these updates include the
+\emph{prediction error} (equation \ref{eq:predError}). The
+\emph{prediction error} corresponds to the Rescorla-Wagner rule
+(Rescorla and Wagner 1972), which is a key component of learning theory
+and widely applied in neurobiology to understand learning processes
+(Holland and Schiffino 2016).
+
 Individuals interact as the focal individual \(n\) times and the
 interaction partner is chosen at random. Thus, the expected number of
-interactions for each individual is \(2n\). After the interaction round
-, individuals in the population reproduce with a probability
-proportional to their total pay-off \(w_i\), which is a sum of the
-baseline pay-off (\(w_0\)) and all the pay-offs obtained throughout
-their life. Thus, the combination of natural selection and genetic drift
-changes the distribution of values in \(\alpha\) and \(\beta\) that
-segregates in the population, effectively changing the badges expressed
-and the communication system.
+interactions for each individual is \(2n\). After the interaction round,
+individuals in the population reproduce with a probability proportional
+to their total pay-off \(w_i\), which is a sum of the baseline pay-off
+(\(w_0\)) and all the pay-offs obtained from interactions throughout
+their life. Thus, the total fitness is composed of the viability
+component (equation \ref{eq:viability}) and the total payoff collected
+in the interaction round. Therefore, the combination of natural
+selection and genetic drift changes the distribution of values in
+\(\alpha\) and \(\beta\) that segregates in the population, effectively
+changing the badges expressed and the communication system.
 
@@
 \caption{Model of communication in the context of aggressive interactions.
-In A, the reaction norm determines the badge size, *e.g* the size of a black band on a bird's chest.
+In A, the reaction norm  (equation \ref{eq:reacNorm}) determines the badge size, \textit{e.g} the size of a black band on a bird's chest.
 Red shows an informative reaction norm, whereas blue shows a uninformative reaction norm.
 B individuals perceiving the signal have a behavioral reaction norm that determines their probability of retreating (black line).
-The reaction norm arises from generalizing the information from the feature weights (black dots).
+The reaction norm arises from generalizing the information from the policy feature weights (black dots).
 Feature weights are values associated with a particular badge size and are updated as individuals interact with each other.
@@
 Top panels (A.1 and A.2) show the evolutionary dynamics of the sender code.
-In the y axis the trait value, in the x axis the generation time in thousands of generations.
+In y axis shows the trait value, in the x axis shows the generation time in thousands of generations.
 On the left (A.1), changes in the distribution of values for the intercept of the reaction norm ($\alpha$); on the right (A.2) changes in the distribution of the slope ($\beta$).
@@
 In the second half, the slope takes positive values and reaction norms evolve to produce an honest signal of quality.
-Correspondingly, individuals learn to react to badge size by increasing the probability of retreating with increasing badge sizes}\label{fig:handicap}
+Correspondingly, individuals learn to react to badge size by increasing the probability of retreating with increasing badge sizes.}\label{fig:handicap}
 \end{figure}
 
-\hypertarget{badges-as-handicaps}{%
-\subsubsection{Badges as Handicaps}\label{badges-as-handicaps}}
+\hypertarget{badges-as-handicaps-costly-signals}{%
+\subsubsection{Badges as Handicaps (costly
+signals)}\label{badges-as-handicaps-costly-signals}}
 
@@
 reaction norm parameters (Fig. \ref{fig:handicap}, A) appeared to
-involve a set of steps. First, the intercept of the reaction norm
-(\(\alpha\)) describing the relationship between individual quality and
-badge size evolved to higher values, reducing the average badge size in
-the populations. This is because reaction norms segregating in the
+involve a set of steps. First, the intercept of the logistic reaction
+norm (\(\alpha\)) describing the relationship between individual quality
+and badge size evolved to lower values, reducing the average badge size
+in the populations. This is because reaction norms segregating in the
 population are flat at the beginning of the simulation; consequently,
@@
 badge size. In either of these cases reaction norms are flat, so the
-badge does not provide any information about quality.
-
-\hypertarget{the-evolution-of-polymorphism-mediated-by-learning}{%
-\subsubsection{The evolution of polymorphism mediated by
-learning}\label{the-evolution-of-polymorphism-mediated-by-learning}}
+badge does not provide any information about quality. This means that
+learning is not sufficient to yield a communication system, but rather
+is requires the badge to have a cost proportional to the quality of the
+individual.
+
+We have described here two replicate simulations representative of the
+two outcomes found in a series of 20 replicates run with the same
+parameter combination. The exact evolutionary trajectory differs in each
+of these replicates due to the stochastic effects of the dynamical
+system. Particularly the timing of the changes was different in each
+replicate. Nevertheless, all of the replicates could be classified as
+either evolving a functioning communication system, or not having the
+badge present in the population.
+
+\hypertarget{the-evolution-of-polymorphism-mediated-by-learning-when-the-signal-is-cost-free}{%
+\subsubsection{The evolution of polymorphism mediated by learning when
+the signal is
+cost-free}\label{the-evolution-of-polymorphism-mediated-by-learning-when-the-signal-is-cost-free}}
 
@@
 starting value. Within the first 2000 generations, the unimodal
-distribution in the intercept (\(\alpha\)) splits into a bimodal one.
-Later in evolutionary time, one of the peaks splits further, so at the
-end of the evolutionary simulation the distribution of the intercept
-(\(\alpha\)) in the population shows three distinct peaks. In the case
-of \(\beta\), the peaks in the distribution are not so clear-cut, but
-the variance of the distribution increases over time. These changes in
-the distribution of the parameters of the sender reaction norm imply
-that individuals can generally be classified into three distinct types
-(Fig. \ref{fig:branching}). Two types express a flat reaction norm with
-extreme values for the badge size, meaning that their badge size is not
-informative of their quality, whereas the third type shows intermediate
-badge sizes determined by individual quality (Fig. \ref{fig:branching}
-B.3-4). In this simulation, the values of the slope in the informative
-group are negative, which implies that badge size correlates negatively
-with quality. Furthermore, the receiver reaction norms developed through
-learning, particularly those of individuals of intermediate quality,
-respond to the signal of their sender type by increasing the probability
-of retreating from a fight with individuals with smaller badges (Fig.
-\ref{fig:branching}).
+distribution in the logistic intercept (\(\alpha\)) splits into a
+bimodal one. Later in evolutionary time, one of the peaks splits
+further, so at the end of the evolutionary simulation the distribution
+of the logistic intercept (\(\alpha\)) in the population shows three
+distinct peaks. In the case of \(\beta\), the peaks in the distribution
+are not so clear-cut, but the variance of the distribution increases
+over time. These changes in the distribution of the parameters of the
+sender reaction norm imply that individuals can generally be classified
+into three distinct types (Fig. \ref{fig:branching}). Two types express
+a flat reaction norm with extreme values for the badge size, meaning
+that their badge size is not informative of their quality, whereas the
+third type shows intermediate badge sizes determined by individual
+quality (Fig. \ref{fig:branching} B.3-4).
+
+In this simulation, the values of the slope in the informative group are
+negative, which implies that badge size correlates negatively with
+quality. The receiver reaction norms developed through learning,
+particularly those of individuals of intermediate quality, respond to
+the signal of their sender type by increasing the probability of
+retreating from a fight with individuals with smaller badges (Fig.
+\ref{fig:branching}). However, given the absence of costs, the sign of
+the slope is arbitrary. In our replicate simulations we found examples
+where the third type had positive and negative slopes.
 
@@
 The top panels (A) show changes in the distribution of values for $\alpha$ (A.1) and $\beta$ (A.2) along evolutionary time.
-In the y axis the trait value, in the x axis the generation time in thousands of generations.
+The y axis shows the trait value, and  the x axis shows the generation time in thousands of generations.
 Panels below correspond to snapshots of the sender (B) and receiver codes (C).
@@
 Color of the lines indicates the quality of the individual, as indicated in the color code in C.1.
-The dashed line represents the average reaction norm of the popúlations.
+The dashed line represents the average reaction norm of the populations.
 The unimodal distribution with which $\alpha$ (A.1) starts the simulation, quickly turns into a bimondal distribution, and at about 10000 generations it splits further into three modes.
@@
 relationship, however, is muddled by the fact that there are two
-clusters of values for the intercept (\(\alpha\)) and slope (\(\beta\))
-of the reaction norm in the population. Thus, there are two types of
-reaction norms. One of the types has a steeper slope compared to the
-other. This effect of an increased variance in the trait distribution is
-not only triggered by lower number of interactions. Larger variances are
-also found when we assume a lower speed of learning (data not shown).
-This suggest that limits to the amount of information that individuals
-acquire through learning allows the coexistence of different
-communication strategies within a population, something previously shown
-by Botero et al. (2010).
+clusters of values for the logistic intercept (\(\alpha\)) and slope
+(\(\beta\)) of the reaction norm in the population. Thus, there are two
+types of reaction norms. One of the types has a steeper slope compared
+to the other. This effect of an increased variance in the trait
+distribution is not only triggered by lower number of interactions.
+Larger variances are also found when we assume a lower speed of learning
+(data not shown). This suggest that limits to the amount of information
+that individuals acquire through learning allows the coexistence of
+different communication strategies within a population, something
+previously shown by Botero et al. (2010).
 
@@
 
-\hypertarget{the-peaceful-the-aggressive-and-the-clever}{%
-\subsubsection{The peaceful, the aggressive and the
-clever}\label{the-peaceful-the-aggressive-and-the-clever}}
+\hypertarget{the-peaceful-the-aggressive-and-the-clever-when-signals-are-cost-free}{%
+\subsubsection{The peaceful, the aggressive and the clever, when signals
+are
+cost-free}\label{the-peaceful-the-aggressive-and-the-clever-when-signals-are-cost-free}}
 
@@
 the only one where the distribution of values is split in different
-clusters. That is, most of the variation occurs within clusters. In
-contrast, when naive individuals behave either aggressively or cleverly,
-badges evolve toward minimum and maximum values, and so the variation is
-driven mainly by differences among replicates. We can make sense of
-these results by realizing that individuals change their naive behaviour
-quickly in ranges of the badge size that are common in the population.
-In contrast, an individual expressing a rare badge size will most likely
-experience the naive behaviour of their partner. When the naive
-behaviour is peaceful, individuals with a rare badge size have a fitness
-advantage. This triggers negative-frequency dependent selection and the
-evolution of different types of badge sizes. According to this
-narrative, situations where individuals learn fast and interact
-repeatedly will diminish the strength of frequency-dependent selection.
+clusters. That is, most of the variation occurs between clusters of each
+replicate. In contrast, when naive individuals behave either
+aggressively or cleverly, badges evolve toward minimum and maximum
+values, and so the variation is driven mainly by differences among
+replicates. We can make sense of these results by realizing that
+individuals change their naive behaviour quickly in ranges of the badge
+size that are common in the population. In contrast, an individual
+expressing a rare badge size will most likely experience the naive
+behaviour of their partner. When the naive behaviour is peaceful,
+individuals with a rare badge size have a fitness advantage. This
+triggers negative frequency-dependent selection and the evolution of
+different types of badge sizes. According to this narrative, situations
+where individuals learn fast and interact repeatedly will diminish the
+strength of frequency-dependent selection.
 
@@
 \caption{The peaceful, the aggressive and the clever.
-Distribution of values of the intercept $\alpha$ and slope $\beta$ for individuals at the end of the evolutionary simulations.
+Distribution of values of the logistic intercept $\alpha$ and slope $\beta$ for individuals at the end of the evolutionary simulations.
 The panels show the three different initial conditions for the behaviour of individuals: peaceful, agressive and  "clever", see maintext for details.
@@
 
-The action of natural selection and genetic drift on the learning
+Given how crucial naive behaviour is for the branching event described
+earlier, we tested whether the occurrence of branching was robust to the
+actions of natural selection. If natural selection favoured high
+learning rates or extreme values in the initial policy vector, then the
+evolutionary branching we find would not be as relevant in biological
+populations. Thus, we developed an extension of the model where the
+speed of learning (\(\Alpha\)) and the initial value of the policy
+vector \(\symbf{y}\) are allowed to change subject to natural selection
+and genetic drift. We do this by assuming that the values of the
+learning parameters are inherited from parent to offspring and
+occasionally undergo mutation, just like the parameters of the reaction
+norm. The action of natural selection and genetic drift on the learning
 parameters do not override the conditions for the evolutionary branching
-(Dieckmann and Law 1996) described above. Given how crucial naive
-behaviour is for the branching event described earlier, we tested
-whether evolutionary processes could lead key parameters of the learning
-process toward values that prevented the genetic variation to build up.
-Figure \ref{fig:learnEvol} shows the dynamics from a set of simulations
-where we allow for the coevolution of the reaction norm as well as key
-parameters of the learning module, namely the speed of learning (\(A\))
-and the behavioural tendency expressed by individuals prior to learning.
-Top panels in figure \ref{fig:learnEvol} show that there is
-diversification process in this coevolutionary scenario as well. That is
-because, as is shown in the middle panels (B) of the same figure, the
-learning parameters stay within the range necessary for the branching
-event to take place. The speed of learning (Fig. \ref{fig:learnEvol}
-B.1) stays well above zero throughout the simulation. In contrast, the
-initial behavioural tendency maintain values around zero. Thus,
-evolutionary processes acting on the learning parameters did not
-override the conditions necessary for frequency-dependent selection to
-boost genetic and phenotypic variation.
+(Dieckmann and Law 1996) described above. Top panels in figure
+\ref{fig:learnEvol} show that there is diversification process in this
+coevolutionary scenario as well. That is because, as is shown in the
+middle panels (B) of the same figure, the learning parameters stay
+within the range necessary for the branching event to take place. The
+speed of learning (Fig. \ref{fig:learnEvol} B.1) stays well above zero
+throughout the simulation. In contrast, the initial behavioural tendency
+maintain values around zero. Thus, evolutionary processes acting on the
+learning parameters did not override the conditions necessary for
+frequency-dependent selection to boost genetic and phenotypic variation.
 
@@
 that the learning process can mediate the evolution of an honest signal
-under the handicap principle (Grafen 1990; Zahavi 1975). This conclusion
-is not different from classical genetic models, where responses to
-signals are innate to the individual. However, unlike classical genetic
-models, our model shows that learning can also mediate the origin of a
-signal polymorphism in the absence of the handicap principle. Under this
-polymorphic equilibrium the population consist of distinct types of
-individuals, with honest and dishonest signals. The amount of
-information gathered through learning, as a well as the initial
-conditions of the learning process, were crucial for the emergence and
-maintenance of variation in the signalling system. Namely the
-polymorphism required a limit in the amount of information collected by
-individuals, as well as peaceful behaviour in naive individuals.
-Furthermore, we showed that the conditions leading to the origin of the
-polymorphism can be reached when evolutionary processes are able to
-change the parameters of the learning system (the speed of learning and
-the naive behaviour).
+under the handicap principle, but it cannot without it (Grafen 1990;
+Zahavi 1975). Learning is not sufficient to trigger the evolution of the
+communication system. This conclusion is not different from classical
+genetic models, where responses to signals are innate to the individual.
+However, unlike classical genetic models, our model shows that learning
+can also mediate the origin of a signal polymorphism in the absence of
+the handicap principle. Under this polymorphic equilibrium the
+population consist of distinct types of individuals, with honest and
+dishonest signals. The amount of information gathered through learning,
+as a well as the initial conditions of the learning process, were
+crucial for the emergence and maintenance of variation in the signalling
+system. Namely, the polymorphism required a limit in the amount of
+information collected by individuals, as well as peaceful behaviour in
+naive individuals. Furthermore, we showed that the conditions leading to
+the origin of the polymorphism can be reached when evolutionary
+processes are able to change the parameters of the learning system (the
+speed of learning and the naive behaviour).
 
@@
 
-The focus of theoretical and empirical work on communication systems,
-and particularly badges of status, is often explaining the presence and
-absence of certain morphological traits within populations. Phenotypic
-traits hypothesized to play a role in communication systems regularly
-exhibit patterns of geographic variation within species that have long
-called for explanations. One example of this is the leapfrog pattern
-(Remsen 1984), whereby a certain morphological trait alternates its
-presence and absence in a set of geographically adjunct populations,
-likely as a consequence of selection (Cadena, Cheviron, and Funk 2011;
-Márquez et al. 2020). However, it is unclear what type of ecological
-process is behind these selective regimes. Given the wide variation in
-outcomes found in our model, which depend on both stochasticity and
-cognitive parameters which may vary in space owing to various processes,
-we speculate that variation in cognition and learning could provide some
+The focus of classical theoretical and empirical work on communication
+systems is often explaining the presence and absence of certain
+morphological traits (ornaments) within populations. Phenotypic traits
+hypothesized to play a role in communication systems regularly exhibit
+patterns of geographic variation within species that have long called
+for explanations. One example of this is the leapfrog pattern (Remsen
+1984), whereby a certain morphological trait alternates its presence and
+absence in a set of geographically adjunct populations, likely as a
+consequence of selection (Cadena, Cheviron, and Funk 2011; Márquez et
+al. 2020). However, it is unclear what type of ecological process is
+behind these selective regimes. Given the wide variation in outcomes
+found in our model, which depend on both stochasticity and cognitive
+parameters which may vary in space owing to various processes, we
+speculate that variation in cognition and learning could provide some
 explanatory power in this respect. A first step to assess this
@@
 \begin{longtable}[]{@{}
-  >{\centering\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.2143}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.7857}}@{}}
-\caption{Notation of the model parameters and variables}\tabularnewline
+  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1525}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.5593}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2881}}@{}}
+\caption{Notation of the model parameters and variables. Values for the
+parameters that are kept constant throughout all simulations are given
+in the third column.}\tabularnewline
 \toprule\noalign{}
@@
 Description
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Value
 \end{minipage} \\
@@
 \endlastfoot
-\(Q_i\) & Quality of individual \(i\) \\
-\(N\) & Population size \\
-\(n\) & Number of interactions as the focal individual \\
+\(Q_i\) & Quality of individual \(i\) & \\
+\(N\) & Population size & \(3000\) \\
+\(n\) & Number of interactions as the focal individual & \\
 \(\sigma\) & Standard deviation of the truncated normal distribution
-from which quality is drawn \\
-\(B_i\) & Badge size of individual \(i\) \\
-\(\alpha_i\) & Intercept of the badge size reaction norm for individual
-\(i\) \\
-\(\beta_i\) & Slope of the badge size reaction norm for individual
-\(i\) \\
-\(\mu\) & Mutation rate \\
+from which quality is drawn & \(0.15\) \\
+\(B_i\) & Badge size of individual \(i\) & \\
+\(\alpha_i\) & Intercept of the badge size logistic reaction norm for
+individual \(i\) & \\
+& & \\
+\(\beta_i\) & Slope of the badge size reaction norm for individual \(i\)
+& \\
+& & \\
+\(\mu\) & Mutation rate & \(0.001\) \\
 \(\sigma_\mu\) & Standard deviation of the normal distribution from
-which quality is mutations are drawn \\
-\(s_i\) & Survival probability of individual \(i\) \\
-\(k_1\) & Intercept of the survival function \\
-\(k_2\) & Slope of the survival function \\
-\(V\) & Value of the contested resource \\
-\(C\) & Cost of an escalated fight \\
+which quality is mutations are drawn & \(0.3\) \\
+\(s_i\) & Survival probability of individual \(i\) & \\
+\(k_1\) & Intercept of the survival function & \(3\) \\
+\(k_2\) & Slope of the survival function & \\
+\(V\) & Value of the contested resource & \(1\) \\
+\(C\) & Cost of an escalated fight & \(1.5\) \\
 \(p_{ij}\) & Probability that individual \(i\) wins an escalated fight
-against individual \(j\) \\
-\(k_3\) & Parameter determining how important is quality defining the
-probability of wining \\
-\(\delta\) & Prediction error \\
-\(A\) & Speed of learning \\
-\(c\) & Number of feature centres \\
-\(b_z\) & badge size for feature centre \(z\) \\
-\(\hat{R}\) & Reward estimate \\
-\(x_z\) & weight of feature centre \(z\) on the reward estimation \\
-\(\theta\) & width of the generalization function \\
+against individual \(j\) & \\
+\(k_3\) & Parameter determining how important quality is defining the
+probability of wining & \(10\) \\
+\(\delta\) & Prediction error & \\
+\(A\) & Speed of learning & \\
+\(c\) & Number of feature centres & \(6\) \\
+\(b_z\) & badge size for feature centre \(z\) & \\
+\(\hat{R}\) & Reward estimate & \\
+\(x_z\) & weight of feature centre \(z\) on the reward estimation & \\
+& & \\
+\(\theta\) & width of the generalization function & \(0.01\) \\
 \(q_{ij}\) & probability that individual \(i\) retreats from a fight in
-an encounter with individual \(j\) \\
-\(y_z\) & weight of feature centre z on the tendency to retreat \\
-\(w_i\) & total pay-off of individual \(i\) \\
-\(w_0\) & base line pay-off \\
+an encounter with individual \(j\) & \\
+\(y_z\) & weight of feature centre \(z\) on the tendency to retreat & \\
+& & \\
+\(w_i\) & total pay-off of individual \(i\) & \\
+\(w_0\) & base line pay-off & \(2\) \\
 \end{longtable}
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-holland_Minireview_2016}{}}%
+Holland, Peter C., and Felipe L. Schiffino. 2016. {``Mini-Review:
+{Prediction} Errors, Attention and Associative Learning.''}
+\emph{Neurobiology of Learning and Memory} 131 (May): 207--15.
+\url{https://doi.org/10.1016/j.nlm.2016.02.014}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-johnstone_Recognition_1997}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-penn_Handicap_2020}{}}%
+Penn, Dustin J., and Szabolcs Számadó. 2020. {``The {Handicap}
+{Principle}: How an Erroneous Hypothesis Became a Scientific
+Principle.''} \emph{Biological Reviews} 95 (1): 267--90.
+\url{https://doi.org/10.1111/brv.12563}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-pretot_Factors_2016}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-rescorla_Theory_1972}{}}%
+Rescorla, R. A., and A. R. A. Wagner. 1972. {``A Theory of {Pavlovian}
+Conditioning: Variations in the Effectiveness of Reinforcement and
+Non-Reinforcement.''} In \emph{Classical Conditioning {II}: Current
+Research and Theory}, edited by Abraham H. Black and William Frederick
+Prokasy. New York: Appleton-Century-Crofts.
+
 \leavevmode\vadjust pre{\hypertarget{ref-rohwer_Social_1975a}{}}%
